---
title: "ATAC_seq_integration"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(readxl)
```

#### Data input  


```{r loading data}
 
DOX_24_ATAC <- readRDS("data/Other_paper_data/Matthews_2025_DOX_24h_ATAC.RDS")

DOX_24_DAR_gr <- DOX_24_ATAC %>% 
  dplyr::filter(adj.P.Val<0.05) %>% 
  dplyr::select(peak) %>% 
  separate(peak, into = c("seqnames","start","stop"),remove = FALSE) %>% 
  mutate(start=as.numeric(start),
         stop=as.numeric(stop)) %>% 
  as_granges(seqnames = seqnames,
             start=start,
             end=stop,
             strand="*")
DOX_24_CAR_gr <- DOX_24_ATAC %>% 
  dplyr::filter(adj.P.Val>0.05) %>% 
  dplyr::select(peak) %>% 
  separate(peak, into = c("seqnames","start","stop"),remove = FALSE) %>% 
  mutate(start=as.numeric(start),
         stop=as.numeric(stop)) %>% 
  as_granges(seqnames = seqnames,
             start=start,
             end=stop,
             strand="*")
mcols(DOX_24_CAR_gr)$name <- mcols(DOX_24_CAR_gr)$peak
mcols(DOX_24_DAR_gr)$name <- mcols(DOX_24_DAR_gr)$peak

H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")
 H3K27ac_summit_gr_ext <- resize(H3K27ac_summit_gr, width(H3K27ac_summit_gr)+400, fix = "center")

peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)

H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})

##assigning Peakid as name of summit region
 mcols(H3K27ac_summit_gr)$name <- mcols(H3K27ac_summit_gr)$Peakid
 
 # H3K27ac_set_case_lfc <- readRDS("data/RDS_files/H3K27ac_set_case_lfc.RDS")


```

```{r export DARs and CARs bedlist, eval=FALSE}

library(GenomeInfoDb)
library(BSgenome.Hsapiens.UCSC.hg38)

export(DOX_24_DAR_gr, "data/Bed_exports/ATAC_DOX_24h_DAR.bed", format = "BED")
export(DOX_24_CAR_gr, "data/Bed_exports/ATAC_DOX_24h_CAR.bed", format = "BED")
genome_sl <- seqlengths(BSgenome.Hsapiens.UCSC.hg38)

genome_sl <- genome_sl[seqlevels(H3K27ac_summit_gr)]

seqlengths(H3K27ac_summit_gr) <-genome_sl
H3K27ac_summit_gr <- trim(H3K27ac_summit_gr)
H3K27ac_summit_gr_ext <- promoters(
  H3K27ac_summit_gr,
  upstream = 200,
  downstream = 200
)
export(H3K27ac_summit_gr_ext,
  "data/Bed_exports/H3K27ac_summit_gr_ext200.bed",
  format = "bed"
)

```
####  Overlapping CAR and DAR of ATAC data with ext200 ROIs of H3K27ac  



```{r CAR hits}

CAR_H3K27ac_hits <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, DOX_24_CAR_gr, ignore.strand = TRUE)
H3K27ac_CAR_overlap <- tibble(
  hit_id = queryHits(CAR_H3K27ac_hits),
  ROI_ID=mcols(H3K27ac_sets_gr$all_H3K27ac)$Peakid[queryHits((CAR_H3K27ac_hits))],
  cluster   = mcols(H3K27ac_sets_gr$all_H3K27ac)$cluster[queryHits(CAR_H3K27ac_hits)],
  CAR_ROI   = mcols(DOX_24_CAR_gr)$peak[subjectHits(CAR_H3K27ac_hits)])

H3K27ac_CAR_peakid <- H3K27ac_CAR_overlap %>% 
  group_by(ROI_ID) %>% 
  distinct(ROI_ID)

CAR_H3K27ac_hits_ext <- findOverlaps(H3K27ac_summit_gr_ext, DOX_24_CAR_gr, ignore.strand = TRUE)
H3K27ac_ext_CAR_overlap <- tibble(
  hit_id = queryHits(CAR_H3K27ac_hits_ext),
  ROI_ID=mcols(H3K27ac_summit_gr_ext)$Peakid[queryHits((CAR_H3K27ac_hits_ext))],
  cluster   = mcols(H3K27ac_summit_gr_ext)$cluster[queryHits(CAR_H3K27ac_hits_ext)],
  CAR_ROI   = mcols(DOX_24_CAR_gr)$peak[subjectHits(CAR_H3K27ac_hits_ext)])

H3K27ac_ext_CAR_peakid <- H3K27ac_ext_CAR_overlap %>% 
  group_by(ROI_ID) %>% 
  distinct(ROI_ID)

```



```{r DAR hits}

DAR_H3K27ac_hits <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, DOX_24_DAR_gr, ignore.strand = TRUE)
H3K27ac_DAR_overlap <- tibble(
  hit_id = queryHits(DAR_H3K27ac_hits),
  ROI_ID=mcols(H3K27ac_sets_gr$all_H3K27ac)$Peakid[queryHits((DAR_H3K27ac_hits))],
  cluster   = mcols(H3K27ac_sets_gr$all_H3K27ac)$cluster[queryHits(DAR_H3K27ac_hits)],
  DAR_ROI   = mcols(DOX_24_DAR_gr)$peak[subjectHits(DAR_H3K27ac_hits)])

H3K27ac_DAR_peakid <- H3K27ac_DAR_overlap %>% 
  group_by(ROI_ID) %>% 
  distinct(ROI_ID) 


DAR_H3K27ac_hits_ext <- findOverlaps(H3K27ac_summit_gr_ext, DOX_24_DAR_gr, ignore.strand = TRUE)
H3K27ac_ext_DAR_overlap <- tibble(
  hit_id = queryHits(DAR_H3K27ac_hits_ext),
  ROI_ID=mcols(H3K27ac_summit_gr_ext)$Peakid[queryHits((DAR_H3K27ac_hits_ext))],
  cluster   = mcols(H3K27ac_summit_gr_ext)$cluster[queryHits(DAR_H3K27ac_hits_ext)],
  CAR_ROI   = mcols(DOX_24_DAR_gr)$peak[subjectHits(DAR_H3K27ac_hits_ext)])

H3K27ac_ext_DAR_peakid <- H3K27ac_ext_DAR_overlap %>% 
  group_by(ROI_ID) %>% 
  distinct(ROI_ID)


```
  
  
### Plots of enrichment Full H3K27ac overlap  


```{r}
joint_H3K27ac <- intersect(H3K27ac_DAR_peakid$ROI_ID, H3K27ac_CAR_peakid$ROI_ID)

prop_ATAC_df <- peakAnnoList_H3K27ac$all_H3K27ac %>%
  as.data.frame() %>% 
  dplyr::select(Peakid) %>% 
  left_join(H3K27ac_lookup) %>% 
  mutate(cluster = if_else(is.na(cluster), "NA", cluster)) %>% 
  mutate(
    ATAC_type = case_when(Peakid %in% joint_H3K27ac ~ "joint",
      Peakid %in% H3K27ac_CAR_peakid$ROI_ID ~ "CAR",
      Peakid %in% H3K27ac_DAR_peakid$ROI_ID ~ "DAR",
      TRUE ~ "not_atac"
    )
  ) %>%
  count(cluster, ATAC_type, name = "n") %>%
  group_by(cluster) %>%
  mutate(
    percent = n / sum(n) * 100
  ) %>%
  ungroup()

ggplot(prop_ATAC_df, aes(x = cluster, y = percent, fill = ATAC_type)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  geom_text(
    aes(label = sprintf("%.1f%%", percent)),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  labs(y = "Proportion")+
  theme_bw()+
  ggtitle("Proportion of CARs and DARs in each H3K27ac set")
  

prop_ATAC_df %>% 
  pivot_wider(id_cols = c(cluster), names_from = ATAC_type, values_from = percent) %>% 
   datatable(.,rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE)) %>% 
  DT::formatRound(columns = c("CAR", "DAR", "joint", "not_atac"), # explicitly name the numeric columns
    digits = 2)
  

prop_ATAC_df %>% 
  pivot_wider(id_cols = c(cluster), names_from = ATAC_type, values_from = n) %>%
  rowwise() %>% 
  mutate(total=sum(c_across(any_of(c("CAR", "DAR","joint","not_atac"))))) %>% 
  datatable(.,rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE), caption = "counts of H3K27ac ROIs")
```

### Plots of enrichment 200ext H3K27ac overlap  


```{r}
joint_H3K27ac_ext <- intersect(H3K27ac_ext_DAR_peakid$ROI_ID, H3K27ac_ext_CAR_peakid$ROI_ID)

prop_ATAC_df_ext <- peakAnnoList_H3K27ac$all_H3K27ac %>%
  as.data.frame() %>% 
  dplyr::select(Peakid) %>% 
  left_join(H3K27ac_lookup) %>%
  mutate(cluster = if_else(is.na(cluster), "NA", cluster)) %>% 
  mutate(
    ATAC_type = case_when(Peakid %in% joint_H3K27ac_ext ~ "joint",
      Peakid %in% H3K27ac_ext_DAR_peakid$ROI_ID ~ "DAR",
      Peakid %in% H3K27ac_ext_CAR_peakid$ROI_ID ~ "CAR",
      TRUE ~ "not_atac"
    )
  ) %>%
  count(cluster, ATAC_type, name = "n") %>%
  group_by(cluster) %>%
  mutate(
    percent = n / sum(n) * 100
  ) %>%
  ungroup()

ggplot(prop_ATAC_df_ext, aes(x = cluster, y = percent, fill = ATAC_type)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  geom_text(
    aes(label = sprintf("%.1f%%", percent)),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  labs(y = "Proportion")+
  theme_bw()+
  ggtitle("Proportion of CARs and DARs in each H3K27ac ext_200 set")
  

prop_ATAC_df_ext %>% 
  pivot_wider(id_cols = c(cluster), names_from = ATAC_type, values_from = percent) %>% 
   datatable(.,rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE)) %>% 
  DT::formatRound(columns = c("CAR", "DAR", "joint", "not_atac"), # explicitly name the numeric columns
    digits = 2)
  

prop_ATAC_df_ext %>% 
  pivot_wider(id_cols = c(cluster), names_from = ATAC_type, values_from = n) %>%
  rowwise() %>% 
  mutate(total=sum(c_across(any_of(c("CAR", "DAR","joint","not_atac"))))) %>% 
  datatable(.,rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE), caption = "counts of H3K27ac ext ROIs")
```
```{r}
Peakid_df <- data.frame(
  Peakid = H3K27ac_summit_gr_ext$Peakid,
  cluster = H3K27ac_summit_gr_ext$cluster
)

Peakid_df$is_CAR <-Peakid_df$Peakid %in% H3K27ac_ext_CAR_peakid$ROI_ID
Peakid_df$is_DAR <-Peakid_df$Peakid %in% H3K27ac_ext_DAR_peakid$ROI_ID
```

#### Enrichment statistics  


```{r testing set2vset1 car}
df_sub_set2v1 <- Peakid_df %>%
  dplyr::filter(cluster %in% c("Set_1", "Set_2"))
table(df_sub_set2v1$cluster, df_sub_set2v1$is_CAR)

fisher.test(table(df_sub_set2v1$cluster, df_sub_set2v1$is_CAR))


```
so SEt2 is depeleted in CARs compared to SEt1

```{r}
df_sub_set3v1 <- Peakid_df %>%
  dplyr::filter(cluster %in% c("Set_1", "Set_3"))
table(df_sub_set3v1$cluster, df_sub_set3v1$is_CAR)

fisher.test(table(df_sub_set3v1$cluster, df_sub_set3v1$is_CAR))
```

Set 3 is also significantly depleted in CARs compared to SET 1

```{r}
DAR_sub_set2v1 <- Peakid_df %>%
  dplyr::filter(cluster %in% c("Set_1", "Set_2"))
table(DAR_sub_set2v1$cluster, DAR_sub_set2v1$is_DAR)

fisher.test(table(DAR_sub_set2v1$cluster, DAR_sub_set2v1$is_DAR))
```

SEt 2 significantly enriched in DARs compared to set 1

```{r}
DAR_sub_set3v1 <- Peakid_df %>%
  dplyr::filter(cluster %in% c("Set_1", "Set_3"))
table(DAR_sub_set3v1$cluster, DAR_sub_set3v1$is_DAR)

fisher.test(table(DAR_sub_set3v1$cluster, DAR_sub_set3v1$is_DAR))
```

Set 3 is significantly enriched in DARs vs Set 1

###  Evaluation of repName in CAR v DAR

This section is trying to test if the MER61E or any integrants in any repName (the ones I am concerned about in my Set 3 data (Acute response) have are DARs and whether they increase or decrease in ATAC LFC)

To type out what I am thinking a little more, my idea is that if MER61 elements are acting as promotors, I will find increased ATAC accessibility and increased H3K27ac marks of the overlapping integrants at 24T.  I know I already see a decrease of H3K27ac overtime in my other analysis for these integrants, but have not drilled down to specific sites.

```{r}
sat_lowc_simrpt_list <- readRDS( "data/RDS_files/sat_lowc_simrpt_repNames.RDS")

annotated_H3K27ac_ext200 <- readRDS("data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")


split_af_H3K27ac <- annotated_H3K27ac_ext200 %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName, repID), delim=";") %>% 
  dplyr::select(Peakid,repClass:repID,cluster) %>% 
  dplyr::filter(!repName %in% sat_lowc_simrpt_list)


anno_check_split_repName_H3K27ac <- split(split_af_H3K27ac,split_af_H3K27ac$repName)


rep_counts_H3K27ac <- split_af_H3K27ac %>%
  distinct(Peakid, repName) %>%   # avoid duplicate counting per peak
  count(repName, name = "n_peaks") %>%
  arrange(desc(n_peaks))

### finding higher count repNames
keep_repNames_H3K27ac <- rep_counts_H3K27ac %>%
  filter(n_peaks >= 10) %>%
  pull(repName)

### filter out lower count repNames


```




```{r}
repName_peak_list_H3K27ac <- lapply(anno_check_split_repName, function(df) {
  unique(df$Peakid)
})


repName_peak_list_filt_H3K27ac <- repName_peak_list_H3K27ac[names(repName_peak_list_H3K27ac) %in% keep_repNames_H3K27ac]


all_peaks_H3K27ac <- unique(annotated_H3K27ac_ext200$Peakid)

CAR_H3K27ac <- unique(H3K27ac_ext_CAR_peakid$ROI_ID)
DAR_H3K27ac <- unique(H3K27ac_ext_DAR_peakid$ROI_ID)

CAR_H3K27ac_only <- setdiff(CAR_H3K27ac, DAR_H3K27ac)
DAR_H3K27ac_only <- setdiff(DAR_H3K27ac, CAR_H3K27ac)

test_enrichment_H3K27ac <- function(TE_peaks, target_peaks, background) {
  a <- sum(background %in% TE_peaks & background %in% target_peaks)
  b <- sum(background %in% TE_peaks & !(background %in% target_peaks))
  c <- sum(!(background %in% TE_peaks) & background %in% target_peaks)
  d <- sum(!(background %in% TE_peaks) & !(background %in% target_peaks))
  
  mat <- matrix(c(a,b,c,d), nrow = 2)
  res <- fisher.test(mat)
  
  data.frame(
    overlap = a,
    odds_ratio = as.numeric(res$estimate),
    p_value = res$p.value
  )
}
```

```{r CAR H3K27ac repName test}
CAR_H3K27ac_results <- lapply(names(repName_peak_list_filt_H3K27ac), function(rep) {
  
  TE_peaks <- repName_peak_list_filt_H3K27ac[[rep]]
  
  res <- test_enrichment_H3K27ac(
    TE_peaks = TE_peaks,
    target_peaks = CAR_H3K27ac,
    background = all_peaks_H3K27ac
  )
  
  res$repName <- rep
  res$TE_total <- length(TE_peaks)   # add this
  
  res
}) %>% bind_rows()
```

```{r DAR H3K27ac repName test}

DAR_H3K27ac_results <- lapply(names(repName_peak_list_filt_H3K27ac), function(rep) {
  
  TE_peaks <- repName_peak_list_filt_H3K27ac[[rep]]
  
  res <- test_enrichment_H3K27ac(
    TE_peaks = TE_peaks,
    target_peaks = DAR_H3K27ac,
    background = all_peaks_H3K27ac
  )
  
  res$repName <- rep
  res$TE_total <- length(TE_peaks)   # add this
  
  res
}) %>% bind_rows()

```

```{r volcano plots}
CAR_H3K27ac_results <- CAR_H3K27ac_results %>%
  filter(!is.na(repName) & repName != "NA") %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

DAR_H3K27ac_results <- DAR_H3K27ac_results %>%
  filter(!is.na(repName) & repName != "NA") %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

# saveRDS(CAR_H3K27ac_results, "data/RDS_files/CAR_H3K27ac_repName_enrichment_results.RDS")
# saveRDS(DAR_H3K27ac_results, "data/RDS_files/DAR_H3K27ac_repName_enrichment_results.RDS")

```


```{r}
CAR_H3K27ac_volcano_df <- CAR_H3K27ac_results %>%
  mutate(
    log2_OR = log2(odds_ratio),
    neg_log10_FDR = -log10(FDR),
    sig = FDR < 0.05
  ) %>% 
 mutate(
    log2_OR = ifelse(is.infinite(log2_OR), NA, log2_OR)
  )


ggplot(CAR_H3K27ac_volcano_df, aes(x = log2_OR, y = neg_log10_FDR)) +
  geom_point(aes(color = sig), alpha = 0.7) +
  geom_text_repel(data = subset(CAR_H3K27ac_volcano_df, FDR < 0.05),  # only significant points
                  aes(label = repName),
    size = 3,
    max.overlaps = 100
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  scale_color_manual(values = c("grey70", "red")) +
  theme_classic() +
  labs(
    x = "log2(Odds Ratio)",
    y = "-log10(FDR)",
    title = "TE Enrichment in CAR H3K27ac Peaks"
  )
```

```{r}
DAR_H3K27ac_volcano_df <- DAR_H3K27ac_results %>%
  mutate(
    log2_OR = log2(odds_ratio),
    neg_log10_FDR = -log10(FDR),
    sig = FDR < 0.05
  ) %>% 
 mutate(
    log2_OR = ifelse(is.infinite(log2_OR), NA, log2_OR)
  )

ggplot(DAR_H3K27ac_volcano_df, aes(x = log2_OR, y = neg_log10_FDR)) +
  geom_point(aes(color = sig), alpha = 0.7) +
  geom_text_repel(data = subset(DAR_H3K27ac_volcano_df, FDR < 0.05),  # only significant points
                  aes(label = repName),
    size = 3,
    max.overlaps = 100
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  scale_color_manual(values = c("grey70", "red")) +
  theme_classic() +
  labs(
    x = "log2(Odds Ratio)",
    y = "-log10(FDR)",
    title = "TE Enrichment in DAR H3K27ac Peaks"
  )
```
