---
title: "ChromHMM_2"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 400px;
}
```


```{r}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(readxl)
library(ChIPseeker)
# library(regioneR)
library(cowplot)
library(GenomeInfoDb)
library(ggforce)
library(ggalluvial)
```




### Adding annotation to states

```{r annotation}

States_anno <- data.frame("State"=c(paste0("E",1:16)),
"Name"=c("Heterochromatin_1",
"Quiescent_Low_Coverage",
"Repressed_Heterochromatin",
"Repressed_ZFP_genes",
"Strong_Polycomb_Repressed",
 "Bivalent_Enhancer",
"Heterochromatin_2",
"Polycomb_Repressed",
 "Weak_Genic_Enhancer",
 "Bivalent_Poised_TSS",
 "Active_Enhancer",
  "Strong_Genic_Enhancer",
 "Active_TSS",
 "Very_Weak_Transcription",
 "Genic_Strong_Transcription",
 "Genic_Weak_Transcription"), "recode"=c("N","P","M","L","J","I","O","K","G","H","E","F","A","D","B","C")) %>% 
  mutate(recode=factor(recode,levels=c("A","B","C","D","E","F","G","H","I","J","K","L","M","N","O","P")))

state_colors <- c(
  # Promoters (strong → weak)
  "Active_TSS"  = "#d73027",  # strong red
    # Changed
  "Strong_Genic_Enhancer"  = "#ffff60",  
  
  # Enhancers
  "Active_Enhancer"  = "#ffff01",  # bright yellow
  "Weak_Genic_Enhancer"  = "#ffeda0",  # soft yellow
  
  # Transcribed regions
  "Genic_Strong_Transcription"  = "#1a9850",  # dark green
  "Genic_Weak_Transcription"  = "#66bd63",
  "Very_Weak_Transcription"  = "#a6d96a",
  
  # Bivalent / regulatory
  "Bivalent_Poised_TSS"  = "#c51b7d",
  "Bivalent_Enhancer"= "#de77ae",
  
  # Polycomb / repressed
  "Strong_Polycomb_Repressed" = "#762a83",  # deep purple
  "Polycomb_Repressed" = "#af8dc3",
  
  # Heterochromatin / ZNF/repeats
  "Repressed_ZFP_genes" = "#2c7bb6",  # blue
  "Repressed_Heterochromatin" = "#74abe1",
  "Heterochromatin_1"="#abd9e9",
  "Heterochromatin_2"="#abf9f9",
  
  # Quiescent / low signal
  "Quiescent_Low_Coverage" = "#bdbdbd"
 )

name_order <- States_anno %>%
  arrange(recode) %>%
  pull(Name)

```
### Loading interested regions for overlap

```{r}
MER61E_regions <- import("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed")

annotated_H3K27ac_ext200 <- readRDS("data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")
H3K27ac_ROI <- import("data/Bed_exportsH3K27ac_summit_gr_ext200.bed")

cluster_lookup <- annotated_H3K27ac_ext200 %>% 
  dplyr::select(Peakid, cluster)

# H3K27ac_ROI %>% as.data.frame()

# annotated_H3K27ac_ext200 %>% 
#   dplyr::filter(str_detect(repName,"^HERVIP"))
KZFP_tss_gr <- import("data/Bed_exports/KZFP_tss_ensembl.bed")
acute_repid <- annotated_H3K27ac_ext200 %>% 
  dplyr::filter(cluster=="Set_3") %>% 
   tidyr::separate_longer_delim(
    cols = c(repClass, repFamily, repName, repID),
    delim = ";"
  ) %>% 
  dplyr::filter(repName=="MER61E") %>% 
  pull(repID)

acute_LTR_repid <- annotated_H3K27ac_ext200 %>% 
  dplyr::filter(cluster=="Set_3") %>% 
   tidyr::separate_longer_delim(
    cols = c(repClass, repFamily, repName, repID),
    delim = ";"
  ) %>% 
  dplyr::filter(repName=="LTR10E") %>% 
  pull(repID)
```


### Loading State segment files
```{r}
chrom_24T_VEH <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24T_VEH_16_segments.bed")
chrom_24R_VEH <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24R_VEH_16_segments.bed")
chrom_144R_VEH <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/144R_VEH_16_segments.bed")
chrom_24T_DOX <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24T_DOX_16_segments.bed")
chrom_24R_DOX <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24R_DOX_16_segments.bed")
chrom_144R_DOX <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/144R_DOX_16_segments.bed")
### important for future overlapping
seqlevelsStyle(KZFP_tss_gr) <- seqlevelsStyle(chrom_144R_DOX)

```
###  Global MER61E

```{r}
global_hits <- findOverlaps(H3K27ac_ROI, chrom_24T_VEH, ignore.strand = TRUE)

peak_states <- data.frame(
  peak_id = queryHits(global_hits),
  Peak_name = mcols(H3K27ac_ROI)$name[queryHits(global_hits)],
  state = mcols(chrom_24T_VEH)$name[subjectHits(global_hits)],
  width   = width(pintersect(
    H3K27ac_ROI[queryHits(global_hits)],
    chrom_24T_VEH[subjectHits(global_hits)]
  ))
)
peak_states_summed <- peak_states %>%
  group_by(Peak_name, state) %>%
  summarize(total_width = sum(width), .groups = "drop")


# Instead of collapsing to one state per peak, keep the proportions:
peak_states_props <- peak_states_summed %>%
  mutate(frac = total_width / 401)

```

```{r}
peak_states_complete <- peak_states_props %>%
  complete(Peak_name, state, fill = list(frac = 0)) %>%
  left_join(cluster_lookup, by=c("Peak_name"="Peakid")) %>% 
  left_join(States_anno, by=c("state"="State")) %>% 
   mutate(Name = factor(Name, levels = name_order))

plot_df <- peak_states_complete %>% 
  group_by(cluster, Name) %>%
  summarize(mean_frac = mean(frac), .groups = "drop")
  
ggplot(plot_df, aes(x = cluster, y = mean_frac, fill = Name)) +
  geom_bar(stat = "identity") +
  ylab("Mean fraction of peak") +
  theme_classic()+
  scale_fill_manual(values = state_colors) 

ggplot(plot_df ,aes(x = Name, y = cluster, fill = mean_frac)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Mean fraction") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

```{r chromhmm function}
process_chromhmm_raw <- function(chrom_gr, roi_gr) {

  hits <- findOverlaps(roi_gr, chrom_gr, ignore.strand = TRUE)

  peak_states <- data.frame(
    peak_id   = queryHits(hits),
    Peak_name = mcols(roi_gr)$name[queryHits(hits)],
    state     = mcols(chrom_gr)$name[subjectHits(hits)],
    width     = width(pintersect(
      roi_gr[queryHits(hits)],
      chrom_gr[subjectHits(hits)]
    ))
  )

  peak_states_summed <- peak_states %>%
    group_by(Peak_name, state) %>%
    summarize(total_width = sum(width), .groups = "drop")

  peak_states_props <- peak_states_summed %>%
    mutate(frac = total_width / 401)
  
  return(peak_states_props)
}

annotate_states <- function(df) {

  df %>%
    complete(Peak_name, state = unique(States_anno$State),
             fill = list(frac = 0)) %>%
    left_join(cluster_lookup, by = c("Peak_name" = "Peakid")) %>%
    left_join(States_anno, by = c("state" = "State")) %>%
    mutate(Name = factor(Name, levels = name_order))
}
```

```{r}
chrom_list <- list(
  VEH_24T = chrom_24T_VEH,
  VEH_24R  = chrom_24R_VEH,
  VEH_144R  = chrom_144R_VEH,
  DOX_24T  = chrom_24T_DOX,
  DOX_24R  = chrom_24R_DOX,
  DOX_144R  = chrom_144R_DOX
)
results_list <- lapply(chrom_list, function(x)
  annotate_states(process_chromhmm_raw(x, H3K27ac_ROI))
)

all_results <- bind_rows(
  lapply(names(results_list), function(nm) {
    results_list[[nm]] %>%
      mutate(Set = nm)
  })
)
all_results_df <- all_results %>% 
  group_by(Set,cluster, Name) %>%
  summarize(mean_frac = mean(frac), .groups = "drop")


ggplot(all_results_df, aes(x = cluster, y = mean_frac, fill = Name)) +
  geom_bar(stat = "identity") +
  ylab("Mean fraction of peak") +
  theme_classic()+
  facet_wrap(~Set)+
  scale_fill_manual(values = state_colors) 


ggplot(all_results_df ,aes(x = Name, y = cluster, fill = mean_frac)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Mean fraction") +
  theme_minimal() +
  facet_wrap(~Set)+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


### Isolating MER61E

```{r}

### getting PeakIDs of regions that overlap MER61E
MER61E_Peaks <- annotated_H3K27ac_ext200 %>% 
  dplyr::filter(str_detect(repName, "MER61E")) %>% 
  distinct(Peakid)
###Filtering out peakids that do not overlap MER61E elements
MER61E_only_df <- all_results %>% 
  dplyr::filter(Peak_name %in% MER61E_Peaks$Peakid) %>% 
  group_by(Set,cluster, Name) %>%
  summarize(mean_frac = mean(frac), .groups = "drop") %>% 
  mutate(Condition = ifelse(str_detect(Set, "DOX"), "DOX", "VEH"),
         Timepoint = case_when(
           str_detect(Set, "24T")  ~ "24T",
           str_detect(Set, "24R")  ~ "24R",
           str_detect(Set, "144R") ~ "144R")) %>% 
  mutate(Timepoint = factor(Timepoint, levels = c("24T","24R","144R")))

### states I want to keep which contain data
states_to_keep <- MER61E_only_df %>%
  group_by(Name) %>%
  summarize(max_frac = max(mean_frac)) %>%
  dplyr::filter(max_frac > 0.05) %>%
  pull(Name)
### also filter out not_assigned sets here
MER61E_only_df_filt <- MER61E_only_df %>%
  dplyr::filter(Name %in% states_to_keep) %>% 
  dplyr::filter(cluster!="not_assigned")

ggplot(MER61E_only_df, aes(x = cluster, y = mean_frac, fill = Name)) +
  geom_bar(stat = "identity") +
  ylab("Mean fraction of peak") +
  theme_classic()+
  facet_wrap(~Set)+
  scale_fill_manual(values = state_colors) 

ggplot(MER61E_only_df ,aes(x = Name, y = cluster, fill = mean_frac)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Mean fraction") +
  theme_minimal() +
  facet_wrap(~Set)+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )



ggplot(MER61E_only_df_filt,
       aes(x = Timepoint,
           y = mean_frac,
           group = Condition,
           color = Condition,
           linetype = Condition)) +
  geom_line(size = 1) +
  geom_point() +
  facet_grid(cluster ~ Name) +
  scale_linetype_manual(values = c("DOX" = "solid", "VEH" = "dotted")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


all_results %>% 
  count(Peak_name,cluster) %>% 
          dplyr::filter((n<96))
```


```{r}
delta_df <- MER61E_only_df_filt %>%
  dplyr::select(Condition, Timepoint, cluster, Name, mean_frac) %>%
   tidyr::complete(Timepoint, cluster, Name, Condition,
                  fill = list(mean_frac = 0)) %>% 
  pivot_wider(names_from = Condition, values_from = mean_frac) %>% 
  mutate(delta= DOX-VEH)

summary(delta_df$DOX - delta_df$VEH)

ggplot(delta_df,
       aes(x = Timepoint, y = delta, group = Name, color = Name)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line() +
  geom_point() +
  facet_wrap(~cluster) +
  theme_minimal()+
  scale_color_manual(values = state_colors) 

```

```{r, fig.width=10}

### To look at MER61E only, filter out and reduce down to dominant state per region This is done without respect to cluster relationship
peak_state_calls <- all_results %>%
  dplyr::filter(Peak_name %in% MER61E_Peaks$Peakid) %>%
  group_by(Peak_name, Set, Name) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  group_by(Peak_name, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  # left_join(
  #   distinct(all_results[, c("Peak_name", "cluster")]),
  #   by = "Peak_name"
  # ) %>%
  mutate(
    Condition = ifelse(grepl("DOX", Set), "DOX", "VEH"),
    Timepoint = case_when(
      grepl("24T", Set) ~ "24T",
      grepl("24R", Set) ~ "24R",
      grepl("144R", Set) ~ "144R"))

state_transitions <- peak_state_calls %>%
 dplyr::select(Peak_name, Condition, Timepoint, Name, cluster) %>%
  pivot_wider(names_from = Timepoint, values_from = Name) 


ggplot(peak_state_calls,
       aes(x = Timepoint, y = Name, fill = Condition)) +
  geom_tile() +
  facet_wrap(~cluster) +
  theme_minimal()


ggplot(state_transitions,
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.5) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  facet_wrap(~Condition) +
  theme_minimal()


```

```{r}
state_map <- c(
  "Active_TSS" = "Transcription",
  "Very_Weak_Transcription"="Transcription",
  "Strong_Genic_Enhancer" = "Enhancer",
  "Repressed_ZFP_genes" = "Repressed",
  "Active_Enhancer" = "Enhancer",
  "Weak_Genic_Enhancer" = "Enhancer",
  "Bivalent_Poised_TSS" = "Poised",
  "Polycomb_Repressed" = "Repressed",
  "Strong_Polycomb_Repressed" = "Repressed",
  "Repressed_Heterochromatin" = "Repressed",
  "Heterochromatin_1" = "Repressed",
  "Heterochromatin_2" = "Repressed",
  "Quiescent_Low_Coverage" = "Quiescent",
  "Bivalent_Enhancer"="Poised",
"Genic_Weak_Transcription"="Transcription",
"Genic_Strong_Transcription"="Transcription"
)

peak_states <- all_results %>%
  dplyr::filter(Peak_name %in% MER61E_Peaks$Peakid) %>%
  group_by(Peak_name, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
    Condition = ifelse(grepl("DOX", Set), "DOX", "VEH"),
    Timepoint = case_when(
      grepl("24T", Set) ~ "24T",
      grepl("24R", Set) ~ "24R",
      grepl("144R", Set) ~ "144R"))
trajectory <- peak_states %>%
  dplyr::select(Peak_name, Condition, Timepoint, Name, cluster) %>%
  tidyr::pivot_wider(names_from = Timepoint, values_from = Name) %>% 
  mutate(across(c(`24T`,`24R`,`144R`),
                ~ unname(state_map[.x])))
ggplot(dplyr::filter(trajectory, Condition == "DOX"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.6) +
  geom_stratum(width = 0.2) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
  ggtitle("MER61E state transitions (DOX)") +
  theme_minimal()

ggplot(dplyr::filter(trajectory, Condition == "VEH"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.6) +
  geom_stratum(width = 0.2) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
  ggtitle("MER61E state transitions (VEH)") +
  theme_minimal()

peak_states %>% group_by(Peak_name) %>% count(
  
)
```


Cases NOT incomplete anymore.  Time to filter out. note there are 68 MER61E overlapping regions in my data.  For quantification purposes, 16 are not assigned a cluster, 40 are assigned Acute cluster (set 3) and 12 are assigned non response cluster (set 1)

Why are they incomplete:  maybe there is reduced chromatin state confidence?  I have tested to make sure that there is
```{r, clean}

trajectory_clean <- trajectory %>%
  dplyr::filter(complete.cases(.))

ggplot(dplyr::filter(trajectory_clean, Condition == "DOX"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.6) +
  geom_stratum(width = 0.2) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
  ggtitle("MER61E state transitions (DOX)") +
  theme_minimal()

ggplot(dplyr::filter(trajectory_clean, Condition == "VEH"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.6) +
  geom_stratum(width = 0.2) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
  ggtitle("MER61E state transitions (VEH)") +
  theme_minimal()
```

```{r}
qc_df <- trajectory %>%
  mutate(status = case_when(
    if_all(c(`24T`, `24R`, `144R`), ~ !is.na(.)) ~ "complete",
    if_all(c(`24T`, `24R`, `144R`), is.na) ~ "missing",
    TRUE ~ "partial"
  ))

qc_df %>% group_by(Condition,status) %>% tally

qc_peak <- trajectory %>%
  mutate(status = case_when(
    if_all(c(`24T`, `24R`, `144R`), ~ !is.na(.)) ~ "complete",
    if_all(c(`24T`, `24R`, `144R`), is.na) ~ "missing",
    TRUE ~ "partial"
  )) %>%
  distinct(Peak_name, status)

qc_counts <- qc_peak %>%
  count(status)
trajectory %>%
  summarise(
    n_peaks = dplyr::n(),
    complete = sum(complete.cases(.)),
    partially_missing = sum(!complete.cases(.))
  )

ggplot(qc_counts, aes(x = status, y = n, fill = status)) +
  geom_col(width = 0.6) +
  theme_minimal() +
  labs(
    x = NULL,
    y = "Number of MER61E peaks",
    title = "ChromHMM state coverage completeness"
  ) +
  theme(
    legend.position = "none"
  )
```

```{r}
traj <- trajectory %>%
  mutate(status = case_when(
    if_all(c(`24T`,`24R`,`144R`), ~ !is.na(.)) ~ "complete",
    TRUE ~ "partial"
  ))
traj_complete <- trajectory %>%
  dplyr::filter(if_all(c(`24T`,`24R`,`144R`), ~ !is.na(.)))

ggplot(traj_complete,
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`)) +
  geom_stratum() +
  facet_wrap(~Condition) +
  theme_minimal()


```
```{r}
plot_trajectory_TE_spec <- function(anno_df,results_df, TE) {
  # anno_df=annotated_H3K27ac_ext200
  # results_df=all_results 
  # TE="MER61A"
  state_colors <- c(
  "Transcription" = "#D73027",
  "Enhancer" = "#FC8D59",
  "Poised" = "#FEE08B",
  "Repressed" = "#4575B4",
  "Quiescent" = "#D9D9D9")
  
  ## retrieve the peakIDs with associated TE repName
  
  peak_ids <- anno_df %>% 
  dplyr::filter(str_detect(repName, TE)) %>% 
  pull(Peakid)
  
  ## filter results with peak_ids and break into Timepoint and Condition
  peak_states <- all_results %>%
  dplyr::filter(Peak_name %in% peak_ids) %>%
  group_by(Peak_name, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>% 
    mutate(
    Condition = ifelse(grepl("DOX", Set), "DOX", "VEH"),
    Timepoint = case_when(
      grepl("24T", Set) ~ "24T",
      grepl("24R", Set) ~ "24R",
      grepl("144R", Set) ~ "144R"))



## make trajectory dataframe
trajectory <- peak_states %>%
  dplyr::select(Peak_name, Condition, Timepoint, Name, cluster) %>%
  pivot_wider(
    id_cols = c(Peak_name, Condition, cluster),
    names_from = Timepoint,
    values_from = Name
  ) %>%
  mutate(across(c(`24T`,`24R`,`144R`),
                ~ unname(state_map[.x])))


DOX_plot <- ggplot(dplyr::filter(trajectory, Condition == "DOX"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
 geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 3) +
  ggtitle(paste0(TE," state transitions (DOX)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)

VEH_plot <- ggplot(dplyr::filter(trajectory, Condition == "VEH"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 3) +
  ggtitle(paste0(TE," state transitions (VEH)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)
#### clean plots:
trajectory_clean <- trajectory %>%
  dplyr::filter(complete.cases(.))

DOX_clean <- ggplot(dplyr::filter(trajectory_clean, Condition == "DOX"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
   ggtitle(paste0(TE," state transitions (DOX)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)

VEH_clean <- ggplot(dplyr::filter(trajectory_clean, Condition == "VEH"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
   ggtitle(paste0(TE," state transitions (VEH)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)



#### calculating incompletes and QC
traj <- trajectory %>%
  mutate(status = case_when(
    if_all(c(`24T`,`24R`,`144R`), ~ !is.na(.)) ~ "complete",
    TRUE ~ "partial"
  ))

traj_complete <- traj %>%
  dplyr::filter(status == "complete")

completion_plot <- ggplot(traj_complete,
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 3) +
  facet_wrap(~Condition) +
  theme_minimal()+
  ggtitle(paste0("Complete cases of ",TE))+
  scale_fill_manual(values = state_colors)
#### MORE QC

qc_df <- trajectory %>%
  mutate(status = case_when(
    complete.cases(across(c(`24T`,`24R`,`144R`))) ~ "complete",
    rowSums(is.na(across(c(`24T`,`24R`,`144R`)))) == 3 ~ "missing",
    TRUE ~ "partial"
  ))

qc_counts <- qc_df %>%
  distinct(Peak_name, status) %>%
  count(status)

traj_incomp <- trajectory %>%
  summarise(
    n_peaks = dplyr::n(),
    complete = sum(complete.cases(.)),
    partially_missing = sum(!complete.cases(.))
  )

completeness_plot <- ggplot(qc_counts, aes(x = status, y = n, fill = status)) +
  geom_col(width = 0.6) +
  theme_minimal() +
  labs(
    x = NULL,
    y = paste0("Number of ",TE," peaks"),
    title =paste0("ChromHMM state coverage completeness for ",TE)
  ) +
  theme(
    legend.position = "none"
  )


return(list(
  DOX_plot = DOX_plot,
  VEH_plot = VEH_plot,
  DOX_clean = DOX_clean,
  VEH_clean = VEH_clean,
  completion_plot = completion_plot,
  QC_data= qc_df,
  QC_data_2=qc_counts,
  QC_missing=traj_incomp,
  qc_plot=completeness_plot))
}
```
### MER61 family  


```{r}
MER61A_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61A")

MER61B_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61B")

MER61C_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61C")
MER61D_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61D")

MER61E_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61E")

MER61F_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61F")
MER61_int_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61-int")



```
```{r MER61B}
MER61A_results
MER61B_results
MER61C_results
MER61D_results
MER61E_results
MER61F_results
MER61_int_results
```


### LTR10 Families  


```{r}

LTR10E_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "LTR10E")

LTR10E_results
```

```{r}
LTR10B1_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "LTR10B1")

LTR10B1_results

LTR10B2_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "LTR10B2")

LTR10B2_results

HERVIP10B3_int_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "HERVIP10B3-int")

HERVIP10B3_int_results

HERV1_LTRa_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "HERV1_LTRa")

HERV1_LTRa_results


```

### Some Other sub-families
```{r}
SVAF_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "SVA_F")
SVAF_results

L1PA3_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "L1PA3")
L1PA3_results

ALUSX1_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "AluSx1")
ALUSX1_results

THE1C_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "THE1C")
THE1C_results
```
```{r}
Tigger7_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "Tigger7")
Tigger7_results

Chap1mam_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "Chap1_Mam")
Chap1mam_results
```




### KZFP TSS

```{r}


process_chromhmm_KZFP_raw <- function(chrom_gr, roi_gr) {
seqlevelsStyle(roi_gr) <- seqlevelsStyle(chrom_gr)
  hits <- findOverlaps(roi_gr, chrom_gr, ignore.strand = TRUE)

  tss_states <- data.frame(
    tx_id_num   = queryHits(hits),
    Tx_ID = mcols(roi_gr)$name[queryHits(hits)],
    state     = mcols(chrom_gr)$name[subjectHits(hits)],
    width     = width(pintersect(
      roi_gr[queryHits(hits)],
      chrom_gr[subjectHits(hits)]
    ))
  )

  tss_states_summed <- tss_states %>%
    group_by(Tx_ID, state) %>%
    summarize(total_width = sum(width), .groups = "drop")

  # compute actual ROI widths per Tx_ID (handles variable windows correctly)
  roi_widths <- data.frame(
    Tx_ID = mcols(roi_gr)$name,
    roi_width = width(roi_gr)
  ) %>%
    dplyr::distinct()

  peak_states_props <- tss_states_summed %>%
    dplyr::left_join(roi_widths, by = "Tx_ID") %>%
    dplyr::mutate(frac = total_width / roi_width)

   return(peak_states_props)
}

annotate_KZFP_states <- function(df) {

  df %>%
    complete(Tx_ID, state = unique(States_anno$State),
             fill = list(frac = 0)) %>%
    left_join(KZFP_lookup, by = c("Tx_ID" = "tx_name")) %>%
    left_join(States_anno, by = c("state" = "State")) %>%
    mutate(Name = factor(Name, levels = name_order))
}

```


```{r}
mcols(KZFP_tss_gr)
KZFP_transcript_ensembl_id <- readRDS("data/RDS_files/KZFP_transcript_ensembl_id.RDS")


KZFP_2026_list <- read_excel("data/Other_paper_data/human_protein_coding_KZFPs_Davis_natgen_2026.xlsx")

human_KZFP <- KZFP_2026_list %>%
  dplyr::select(Label, `Gene ID`, chrom, start, end, strand) %>%
  dplyr::rename("Gene_ID"=`Gene ID`) %>% 
  dplyr::select(Label, Gene_ID)
KZFP_lookup <-KZFP_transcript_ensembl_id %>%
  as.data.frame() %>% 
  left_join(human_KZFP,by=c("gene_id"="Gene_ID")) %>% 
  dplyr::select(gene_id:Label)


KZFP_results_list <- lapply(chrom_list, function(x)
  annotate_KZFP_states(process_chromhmm_KZFP_raw(x, KZFP_tss_gr))
)


KZFP_results <- bind_rows(
  lapply(names(KZFP_results_list), function(nm) {
    KZFP_results_list[[nm]] %>%
      mutate(Set = nm)
  })
)
KZFP_results_df <-
  KZFP_results %>% 
  group_by(Set,Label, Name) %>%
  summarize(mean_frac = mean(frac), .groups = "drop")

```

```{r}
plot_trajectory_KZFP <- function(results_df,title) {
  state_colors <- c(
  "Transcription" = "#D73027",
  "Enhancer" = "#FC8D59",
  "Poised" = "#FEE08B",
  "Repressed" = "#4575B4",
  "Quiescent" = "#D9D9D9")
  
    ##  break into Timepoint and Condition
  KZFP_states <-  results_df %>%
  group_by(Label, Set, Name) %>%
  summarize(frac = sum(mean_frac), .groups = "drop") %>%
  group_by(Label, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(
    Condition = ifelse(grepl("DOX", Set), "DOX", "VEH"),
    Timepoint = case_when(
      grepl("24T", Set) ~ "24T",
      grepl("24R", Set) ~ "24R",
      grepl("144R", Set) ~ "144R"))
  
  
  ## make trajectory dataframe
trajectory <- KZFP_states %>%
  dplyr::select(Label, Condition, Timepoint, Name) %>%
  pivot_wider(
    id_cols = c(Label, Condition),
    names_from = Timepoint,
    values_from = Name
  ) %>%
  mutate(across(c(`24T`,`24R`,`144R`),
                ~ unname(state_map[.x])))



DOX_plot <- ggplot(dplyr::filter(trajectory, Condition == "DOX"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
 geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 3) +
  ggtitle(paste0( title, " state transitions (DOX)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)

VEH_plot <- ggplot(dplyr::filter(trajectory, Condition == "VEH"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 3) +
  ggtitle(paste0(title, " state transitions (VEH)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)
#### clean plots:
trajectory_clean <- trajectory %>%
  dplyr::filter(complete.cases(.))


#### calculating incompletes and QC
traj <- trajectory %>%
  mutate(status = case_when(
    if_all(c(`24T`,`24R`,`144R`), ~ !is.na(.)) ~ "complete",
    TRUE ~ "partial"
  ))

traj_complete <- traj %>%
  dplyr::filter(status == "complete")

completion_plot <- ggplot(traj_complete,
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 3) +
  facet_wrap(~Condition) +
  theme_minimal()+
  ggtitle(paste0("Complete cases of ",title))+
  scale_fill_manual(values = state_colors)
#### MORE QC

qc_df <- trajectory %>%
  mutate(status = case_when(
    complete.cases(across(c(`24T`,`24R`,`144R`))) ~ "complete",
    rowSums(is.na(across(c(`24T`,`24R`,`144R`)))) == 3 ~ "missing",
    TRUE ~ "partial"
  ))

qc_counts <- qc_df %>%
  distinct(Label, status) %>%
  count(status)

traj_incomp <- trajectory %>%
  summarise(
    n_peaks = dplyr::n(),
    complete = sum(complete.cases(.)),
    partially_missing = sum(!complete.cases(.))
  )

completeness_plot <- ggplot(qc_counts, aes(x = status, y = n, fill = status)) +
  geom_col(width = 0.6) +
  theme_minimal() +
  labs(
    x = NULL,
    y = paste0("Number of  peaks"),
    title =paste0("ChromHMM state coverage completeness for ",title)
  ) +
  theme(
    legend.position = "none"
  )


return(list(
  DOX_plot = DOX_plot,
  VEH_plot = VEH_plot,
   # DOX_clean = DOX_clean,
  # VEH_clean = VEH_clean,
  completion_plot = completion_plot,
  QC_data= qc_df,
  QC_data_2=qc_counts,
  QC_missing=traj_incomp,
  qc_plot=completeness_plot))
}
```

```{r}
plot_trajectory_KZFP(KZFP_results_df, "KZFP TSS")
```
```{r}

KZFP_tss_50bp <- promoters(KZFP_tss_gr, upstream = 50, downstream = 50)
KZFP_results_50bp_list <- lapply(chrom_list, function(x)
  annotate_KZFP_states(process_chromhmm_KZFP_raw(x, KZFP_tss_50bp))
)


KZFP_50bp_results <- bind_rows(
  lapply(names(KZFP_results_50bp_list), function(nm) {
    KZFP_results_50bp_list[[nm]] %>%
      mutate(Set = nm)
  })
)
KZFP_50bp_results_df <-
  KZFP_50bp_results %>% 
  group_by(Set,Label, Name) %>%
  summarize(mean_frac = mean(frac), .groups = "drop")


KZFP_tss_1kb <- promoters(KZFP_tss_gr, upstream = 1000, downstream = 1000)
KZFP_results_1kb_list <- lapply(chrom_list, function(x)
  annotate_KZFP_states(process_chromhmm_KZFP_raw(x, KZFP_tss_1kb))
)


KZFP_1kb_results <- bind_rows(
  lapply(names(KZFP_results_1kb_list), function(nm) {
    KZFP_results_1kb_list[[nm]] %>%
      mutate(Set = nm)
  })
)
KZFP_1kb_results_df <-
  KZFP_1kb_results %>% 
  group_by(Set,Label, Name) %>%
  summarize(mean_frac = mean(frac), .groups = "drop")
```


```{r}
plot_trajectory_KZFP(KZFP_50bp_results_df, "KZFP TSS +/- 50bp")
```

```{r}
plot_trajectory_KZFP(KZFP_1kb_results_df, "KZFP TSS +/- 1kb")
```

```{r}

KZFP_results_50bp_list$VEH_24T %>% 
  dplyr::filter(!is.na(roi_width)) %>% 
  count(Tx_ID) %>% dplyr::filter(n>1)
 


  
```

### Around the genome

```{r}
##making a segmentation set that all files have
all_segments <- GenomicRanges::disjoin(c(chrom_24T_DOX, chrom_24R_DOX, chrom_144R_DOX, chrom_24T_VEH, chrom_24R_VEH,chrom_144R_VEH))

assign_state <- function(common_gr, sample_gr) {

  hits <- findOverlaps(common_gr, sample_gr)

  out <- rep(NA_character_, length(common_gr))

  out[queryHits(hits)] <- sample_gr$name[subjectHits(hits)]

  out
}

state_transitions <- data.frame(

  regionID = seq_along(all_segments),

  DOX_24T  = assign_state(all_segments, chrom_24T_DOX),
  DOX_24R  = assign_state(all_segments, chrom_24R_DOX),
  DOX_144R = assign_state(all_segments, chrom_144R_DOX),

  VEH_24T  = assign_state(all_segments, chrom_24T_VEH),
  VEH_24R  = assign_state(all_segments, chrom_24R_VEH),
  VEH_144R = assign_state(all_segments, chrom_144R_VEH))

### summarizing state transitions because of number
# transition_counts <- state_transitions %>%
#   count(DOX_24T, DOX_24R, DOX_144R, VEH_24T,VEH_24R,VEH_144R)

States_anno$simple_state <- state_map[States_anno$Name] 

mapping_simple_state <- setNames(
  States_anno$simple_state,
  States_anno$State)

state_transitions <- state_transitions %>%
  mutate(
    DOX_24T_anno  = mapping_simple_state[DOX_24T],
    DOX_24R_anno  = mapping_simple_state[DOX_24R],
    DOX_144R_anno = mapping_simple_state[DOX_144R],

    VEH_24T_anno  = mapping_simple_state[VEH_24T],
    VEH_24R_anno  = mapping_simple_state[VEH_24R],
    VEH_144R_anno = mapping_simple_state[VEH_144R])
  
```

```{r}
state_colors <- c(
  "Transcription" = "#D73027",
  "Enhancer" = "#FC8D59",
  "Poised" = "#FEE08B",
  "Repressed" = "#4575B4",
  "Quiescent" = "#D9D9D9")
  
state_props <- state_transitions %>% 
   pivot_longer(
    cols = ends_with("_anno"),
    names_to = "TX_timepoint",
    values_to = "State" ) %>% 
  count(TX_timepoint,State) %>% 
  mutate(prop= n/sum(n))

state_props %>% 
  dplyr::mutate(TX_timepoint=factor(TX_timepoint,levels=c("DOX_24T_anno","DOX_24R_anno","DOX_144R_anno","VEH_24T_anno","VEH_24R_anno","VEH_144R_anno"),labels=c("DOX_24T","DOX_24R","DOX_144R","VEH_24T","VEH_24R","VEH_144R"))) %>% 
ggplot( aes(x = TX_timepoint, y = prop, fill = State)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)

# names(state_transitions)
# grep("_anno", names(state_transitions), value = TRUE)
# str(state_transitions)
# setdiff(unique(state_transitions$VEH_24T), names(mapping_simple_state))
# unique(state_transitions$VEH_24T)
# nchar(unique(state_transitions$VEH_24T))
```


### All MER61E integrants
Working on application to all TEs
```{r}

process_chromhmm_TE_raw <- function(chrom_gr, te_gr, te_name) {

  hits <- findOverlaps(te_gr, chrom_gr, ignore.strand = TRUE)

  te_states <- data.frame(
    te_id    = mcols(te_gr)$name[queryHits(hits)],
    te_name  = te_name,
    state    = mcols(chrom_gr)$name[subjectHits(hits)],
    width    = width(pintersect(
      te_gr[queryHits(hits)],
      chrom_gr[subjectHits(hits)]
    )),
    TE_width = width(te_gr[queryHits(hits)])
  )

  te_states_summed <- te_states %>%
    group_by(te_id, te_name, state, TE_width) %>%
    summarize(total_width = sum(width), .groups = "drop")

  te_states_props <- te_states_summed %>%
    mutate(frac = total_width / TE_width)
  
  return(te_states_props)
}
```


```{r MER61E processing}
MER61E_regions <- import("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed")

MER61E_results_df <- purrr::imap_dfr(chrom_list, function(chrom_gr, set_name) {
  
  process_chromhmm_TE_raw(
    chrom_gr = chrom_gr,
    te_gr = MER61E_regions,
    te_name = "MER61E"
  ) %>%
    mutate(Set = set_name)
})


MER61E_dominant <- MER61E_results_df %>%
  group_by(te_id, te_name, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  left_join(States_anno, by = c("state" = "State")) %>%
  mutate(
    Broad_state = unname(state_map[Name]),
    Condition = ifelse(grepl("DOX", Set), "DOX", "VEH"),
    Timepoint = case_when(
      grepl("24T", Set) ~ "24T",
      grepl("24R", Set) ~ "24R",
      grepl("144R", Set) ~ "144R",
      TRUE ~ NA_character_
    )
  )

MER61E_trajectory_detailed <- MER61E_dominant %>%
  dplyr::select(te_id, te_name, Condition, Timepoint, Name) %>%
  pivot_wider(
    names_from = c(Condition, Timepoint),
    values_from = Name
  )

MER61E_trajectory_broad <- MER61E_dominant %>%
  dplyr::select(te_id, te_name, Condition, Timepoint, Broad_state) %>%
  pivot_wider(
    names_from = c(Condition, Timepoint),
    values_from = Broad_state)
  all_results
```

The above code created my dominant state dataframe, then a detailed version and a broad version of each dominant state.

We wanted to investigate how many states there are within MER61E groups and what the breakdown is. 

```{r MER61E processing states}
MER61E_state_wide <- MER61E_results_df %>%
  group_by(te_id, Set, state) %>%
  summarise(
    TE_width = dplyr::first(TE_width),
    total_width = sum(total_width, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    frac = total_width / TE_width
  ) %>%
  pivot_wider(
    id_cols = c(te_id, Set, TE_width),
    names_from = state,
    values_from = c(total_width, frac),
    names_glue = "{state}_{.value}",
    values_fill = list(
      total_width = 0,
      frac = 0
    )
  ) %>%
  arrange(te_id, Set)
 
 MER61E_state_mixed <- MER61E_state_wide %>%
  filter(
    !if_any(ends_with("_frac"), ~ .x == 1)
  )
 
 
 MER61E_fifty_fifty <- MER61E_state_wide %>%
  rowwise() %>%
  filter(
    sum(
      dplyr::near(
        c_across(ends_with("_frac")),
        0.5,
        tol = 1e-6
      )
    ) == 2,
    dplyr::near(
      sum(c_across(ends_with("_frac")), na.rm = TRUE),
      1,
      tol = 1e-6
    )
  ) %>%
  ungroup()
 
MER61E_state_counts <- MER61E_state_wide %>%
  rowwise() %>%
  mutate(
    n_states = sum(c_across(ends_with("_frac")) > 0, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  select(te_id, Set, TE_width, n_states)
  
  
  
  
MER61E_state_counts %>% 
     group_by(n_states, Set) %>% 
     tally() %>% 
     pivot_wider(id_cols = Set, values_from = n, names_from = n_states) %>% 
  datatable(.,caption = "MER61E state  breakdowns",
            options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE)
   


```
I also created a data frame of just the single states and the directly 50/50 split states.

```{r}
state_frac_long <- MER61E_state_wide %>%
   rowwise() %>%
  mutate(
    n_states = sum(c_across(ends_with("_frac")) > 0, na.rm = TRUE)
  ) %>%
  ungroup() %>% 
  pivot_longer(
    cols = ends_with("_frac"),
    names_to = "state",
    values_to = "frac"
  ) %>%
  mutate(
    state = sub("_frac$", "", state)
  ) %>%
  filter(frac > 0) %>% 
  left_join(States_anno[1:2], by=c("state"="State"))

ggplot(
  state_frac_long %>% filter(frac > 0),
  aes(x = Name, y = frac)
) +
  geom_violin(trim = TRUE, scale = "width") +
  geom_boxplot(
    width = 0.15,
    outlier.shape = NA
  ) +
  labs(
    x = "ChromHMM state",
    y = "Fraction of MER61E element covered",
    title = "Distribution of nonzero ChromHMM state fractions across all MER61E elements"
  ) +
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90))

two_state_long <- state_frac_long %>%
  filter(n_states == 2)
ggplot(two_state_long,
       aes(x = frac)) +
  geom_histogram(binwidth = 0.05) +
  labs(
    x = "Fraction of TE covered by two states",
    y = "Count"
  ) +
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90))

major_minor <- two_state_long %>%
  group_by(te_id, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  mutate(rank = c("Major", "Minor")) %>%
  ungroup()


major_minor_MER <- two_state_long %>%
  group_by(te_id, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  summarise(
    major_state = dplyr::first(state),
    major_frac  = dplyr::first(frac),
    minor_state = dplyr::last(state),
    minor_frac  = dplyr::last(frac),
    .groups = "drop"
  )

ggplot(major_minor,
       aes(rank, frac)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.4) +
  theme_classic()+
  ggtitle("Distribution of major and minor fractions")
   
   major_minor %>%
  group_by(te_id, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  summarise(
    major_state = dplyr::first(state),
    major_frac  = dplyr::first(frac),
    minor_state = dplyr::last(state),
    minor_frac  = dplyr::last(frac),
    .groups = "drop"
  ) %>%
  count(major_state, minor_state, sort = TRUE) %>% 
     left_join(States_anno[,1:2],by=c("major_state"="State")) %>% 
     left_join(States_anno[1:2], by=c("minor_state"="State")) %>% 
     dplyr::rename("Minor_name"=Name.y, "Major_name"=Name.x) %>% 
     dplyr::select(Major_name, Minor_name,n,major_state,minor_state) %>% 
   
   datatable(.,caption="MER61E state pairs (for 2 state fractions)",
               options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE)
   
```
The above plot is showing the overall distribution of fractions of MER61E elements by each state in a violin plot.  Note only 15 of 16 states are showing.  The absent state is E6 aka Bivalent_Enhancer.

```{r}
ggplot(two_state_long,
       aes(Name, frac)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x =
          element_text(angle = 45, hjust = 1))+
  ggtitle("Fraction of MER61E elements assigned to each ChromHMM state\n(all exactly two-state overlaps)")+
  theme(axis.text.x=element_text(angle = 90))




```
The Fraction of MER61E elements-twostate overlaps plot above is a boxplot showing the distribution of fractions of the 526 2-state instances.  Note this plot is also missing E6 Bivalent_Enhancer marks; E2 aka Quiescent_Low_Coverage,and E16 Genic_Weak_Transcription.


Now I am interested in how these two state (pairs, major/minor) are occuring...


```{r}
major_minor_MER %>%
   
  ggplot(., aes(x=major_frac))+
  geom_histogram(binwidth = 0.05)
```
This is the plot of the Major fraction amounts.  A fraction that is labelled major occurs when grouping by te_id and Set (treatment-timepoint) then ordering by the fraction.  The first value is labelled Major, the second value is labelled Minor.  In the case of ties (3 instances) the state that came first from left to right on the TE is considered the Major state. (note, in R the default order of states is (E1, E10, E11,...,  E16, E2, E3,......,E9)) (how r orders character strings)

```{r}
te_state_frequency_MER <- MER61E_state_counts %>%
  count(te_id, n_states, name = "n_sets") %>%
  arrange(te_id, n_states)

te_state_frequency_wide_MER <- MER61E_state_counts %>%
  count(te_id, n_states, name = "n_sets") %>%
  mutate(
    n_states = paste0("n_states_", n_states)
  ) %>%
  pivot_wider(
    names_from = n_states,
    values_from = n_sets,
    values_fill = 0
  )
te_state_frequency_wide_MER %>% 
datatable(.,caption="MER61E state frequency",
               options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE)


MER61E_state_counts %>%
  pivot_wider(
    id_cols = te_id,
    names_from = Set,
    values_from = n_states
  ) %>% 
  datatable(.,caption="MER61E state frequency by treatment-timepoint",
               options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE)

```


```{r func plot trajectory}
plot_trajectory <- function(dominant_df, TE, id_col="te_id",state_col="Broad_state") {
  
  state_colors <- c(
  "Transcription" = "#D73027",
  "Enhancer" = "#FC8D59",
  "Poised" = "#FEE08B",
  "Repressed" = "#4575B4",
  "Quiescent" = "#D9D9D9")
  
  name_color_key <- dominant_df %>%
  dplyr::distinct(Name, Broad_state) %>%
  dplyr::filter(!is.na(Name), !is.na(Broad_state)) %>%
  dplyr::mutate(color = unname(state_colors[Broad_state]))

name_colors <- stats::setNames(
  name_color_key$color,
  name_color_key$Name)

  trajectory <- dominant_df %>%
    dplyr::mutate(Condition=factor(Condition, levels=c("VEH","DOX"))) %>% 
  dplyr::select(all_of(id_col),te_name, Condition, Timepoint, all_of(state_col)) %>%
  tidyr::pivot_wider(
    names_from = Timepoint,
    values_from = all_of(state_col))
  
  
  ## checking if TE has all markings
  trajectory <- trajectory %>% 
    mutate(status=case_when(
        rowSums(is.na(across(c(`24T`, `24R`, `144R`)))) == 0 ~ "complete",
        rowSums(is.na(across(c(`24T`, `24R`, `144R`)))) == 3 ~ "missing",
        TRUE ~ "partial"))
  
  n_elements <- trajectory %>%
  distinct(.data[[id_col]]) %>%
  nrow()
  
  trajectory_complete <- trajectory %>%
    filter(status == "complete")
  
   alluvial_plot <- ggplot(trajectory_complete, aes(axis1 = `24T`,
                                                    axis2 = `24R`,
                                                    axis3 = `144R`,
                                                    y = 1)) +
     geom_alluvium(aes(fill = after_stat(stratum)),
                   alpha = 0.4) +
     geom_stratum(aes(fill = after_stat(stratum)),
                  color = "black") +
     geom_text(stat = "stratum",aes(label = after_stat(stratum)),
               size = 3) +
     facet_wrap(~ Condition) +
     scale_x_discrete(limits = c("24T", "24R", "144R"),
                      expand = c(0.1, 0.05)) +
     scale_fill_manual(values = name_colors,
                       na.value = "grey80") +
     labs(title = paste0("Complete ChromHMM state trajectories for ", TE," (n = ",n_elements," integrants)"),
          x = NULL,
          y = paste0("Number of ", TE, " elements"),
          fill = "State") +
     theme_minimal()
   #### calculating incompletes and QC
    qc_counts <- trajectory %>%
    count(Condition, status)
    
    qc_plot <- ggplot(qc_counts, aes(x = status, y = n, fill = status)) +
    geom_col(width = 0.6) +
    facet_wrap(~ Condition) +
    theme_minimal() +
    labs( x = NULL,
          y = paste0("Number of ", TE, " elements"),
          title = paste0("ChromHMM trajectory completeness for ", TE)) +
      theme(legend.position = "none")
    
    qc_summary <- trajectory %>%
    group_by(Condition) %>%
    summarise(n_elements = dplyr::n(),
              complete = sum(status == "complete"),
              partial = sum(status == "partial"),
              missing = sum(status == "missing"),
              frac_complete = complete / n_elements,
              .groups = "drop")
    
    return(list(
    alluvial_plot = alluvial_plot,
    qc_plot = qc_plot,
    trajectory = trajectory,
    trajectory_complete = trajectory_complete,
    QC_counts = qc_counts,
    QC_summary = qc_summary
  ))
}

```

```{r MER61E_plot_data}
MER61E_plot_data <- plot_trajectory(MER61E_dominant,TE="MER61E",id_col="te_id",state_col="Broad_state")

acute_MER61E_dominant <- MER61E_dominant %>% 
  dplyr::filter(te_id %in% acute_repid)

acute_MER61E_plot_data <- plot_trajectory(acute_MER61E_dominant,TE="MER61E",id_col="te_id",state_col="Name")

MER61E_plot_data

acute_MER61E_plot_data
```

### All LTR10E integrants  


```{r LTR10E processing}
LTR10E_regions <- import("data/Bed_exports/repeatmasker_sets/ERV1/LTR10E_ERV1_LTR.bed")

LTR10E_results_df <- purrr::imap_dfr(chrom_list, function(chrom_gr, set_name) {
  
  process_chromhmm_TE_raw(
    chrom_gr = chrom_gr,
    te_gr = LTR10E_regions,
    te_name = "LTR10E"
  ) %>%
    mutate(Set = set_name)
})


LTR10E_dominant <- LTR10E_results_df %>%
  group_by(te_id, te_name, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  left_join(States_anno, by = c("state" = "State")) %>%
  mutate(
    Broad_state = unname(state_map[Name]),
    Condition = ifelse(grepl("DOX", Set), "DOX", "VEH"),
    Timepoint = case_when(
      grepl("24T", Set) ~ "24T",
      grepl("24R", Set) ~ "24R",
      grepl("144R", Set) ~ "144R",
      TRUE ~ NA_character_
    )
  )

LTR10E_trajectory_detailed <- LTR10E_dominant %>%
  dplyr::select(te_id, te_name, Condition, Timepoint, Name) %>%
  pivot_wider(
    names_from = c(Condition, Timepoint),
    values_from = Name
  )

LTR10E_trajectory_broad <- LTR10E_dominant %>%
  dplyr::select(te_id, te_name, Condition, Timepoint, Broad_state) %>%
  pivot_wider(
    names_from = c(Condition, Timepoint),
    values_from = Broad_state)
```

The above code created my dominant state dataframe, then a detailed version and a broad version of each dominant state.

We wanted to investigate how many states there are within MER61E groups and what the breakdown is. 

```{r LTR10E processing states}
LTR10E_state_wide <- LTR10E_results_df %>%
  group_by(te_id, Set, state) %>%
  summarise(
    TE_width = dplyr::first(TE_width),
    total_width = sum(total_width, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    frac = total_width / TE_width
  ) %>%
  pivot_wider(
    id_cols = c(te_id, Set, TE_width),
    names_from = state,
    values_from = c(total_width, frac),
    names_glue = "{state}_{.value}",
    values_fill = list(
      total_width = 0,
      frac = 0
    )
  ) %>%
  arrange(te_id, Set)
 
 LTR10E_state_mixed <- LTR10E_state_wide %>%
  filter(
    !if_any(ends_with("_frac"), ~ .x == 1)
  )
 
 
 LTR10E_fifty_fifty <- LTR10E_state_wide %>%
  rowwise() %>%
  filter(
    sum(
      dplyr::near(
        c_across(ends_with("_frac")),
        0.5,
        tol = 1e-6
      )
    ) == 2,
    dplyr::near(
      sum(c_across(ends_with("_frac")), na.rm = TRUE),
      1,
      tol = 1e-6
    )
  ) %>%
  ungroup()
 
LTR10E_state_counts <- LTR10E_state_wide %>%
  rowwise() %>%
  mutate(
    n_states = sum(c_across(ends_with("_frac")) > 0, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  select(te_id, Set, TE_width, n_states)
  
  
  
  
LTR10E_state_counts %>% 
     group_by(n_states, Set) %>% 
     tally() %>% 
     pivot_wider(id_cols = Set, values_from = n, names_from = n_states) %>% 
  datatable(.,caption = "LTR10E state  breakdowns",
            options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE)
   

```
I also created a data frame of just the single states and the directly 50/50 split states.

I did not find any 50/50 splits in LTR10E elements

```{r}
state_frac_long_LTR <- LTR10E_state_wide %>%
   rowwise() %>%
  mutate(
    n_states = sum(c_across(ends_with("_frac")) > 0, na.rm = TRUE)
  ) %>%
  ungroup() %>% 
  pivot_longer(
    cols = ends_with("_frac"),
    names_to = "state",
    values_to = "frac"
  ) %>%
  mutate(
    state = sub("_frac$", "", state)
  ) %>%
  filter(frac > 0) %>% 
  left_join(States_anno[1:2], by=c("state"="State"))

ggplot((state_frac_long_LTR %>% filter(frac > 0)),
  aes(x = Name, y = frac)
) +
  geom_violin(trim = TRUE, scale = "width") +
  geom_boxplot(
    width = 0.15,
    outlier.shape = NA
  ) +
  labs(
    x = "ChromHMM state",
    y = "Fraction of LTR10E element covered",
    title = "Distribution of nonzero ChromHMM state fractions across all LTR10E elements"
  ) +
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90))

two_state_long_LTR <- state_frac_long_LTR %>%
  filter(n_states == 2)

ggplot(two_state_long_LTR,
       aes(x = frac)) +
  geom_histogram(binwidth = 0.05) +
  labs(
    x = "Fraction of TE covered by two states",
    y = "Count"
  ) +
  theme_classic()+
  theme(axis.text.x=element_text(angle = 90))

major_minor_LTR <- two_state_long_LTR %>%
  group_by(te_id, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  mutate(rank = c("Major", "Minor")) %>%
  ungroup()


major_minor_2_LTR <- two_state_long_LTR %>%
  group_by(te_id, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  summarise(
    major_state = dplyr::first(state),
    major_frac  = dplyr::first(frac),
    minor_state = dplyr::last(state),
    minor_frac  = dplyr::last(frac),
    .groups = "drop"
  )

ggplot(major_minor_LTR,
       aes(rank, frac)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.4) +
  theme_classic()+
  ggtitle("Distribution of major and minor fractions in LTR10E")
   
   major_minor_LTR %>%
  group_by(te_id, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  summarise(
    major_state = dplyr::first(state),
    major_frac  = dplyr::first(frac),
    minor_state = dplyr::last(state),
    minor_frac  = dplyr::last(frac),
    .groups = "drop"
  ) %>%
  count(major_state, minor_state, sort = TRUE) %>% 
     left_join(States_anno[,1:2],by=c("major_state"="State")) %>% 
     left_join(States_anno[1:2], by=c("minor_state"="State")) %>% 
     dplyr::rename("Minor_name"=Name.y, "Major_name"=Name.x) %>% 
     dplyr::select(Major_name, Minor_name,n,major_state,minor_state) %>% 
   
   datatable(.,caption="LTR10E state pairs (for 2 state fractions)",
               options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE)
   

```
The above plot is showing the overall distribution of fractions of LTR10E elements by each state in a violin plot.  All 16 states are showing. (unlike MER61E) 

```{r}
ggplot(two_state_long_LTR,
       aes(Name, frac)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x =
          element_text(angle = 45, hjust = 1))+
  ggtitle("Fraction of LTR10E elements assigned to each ChromHMM state\n(all exactly two-state overlaps)")+
  theme(axis.text.x=element_text(angle = 90))


```
The Fraction of LTR10E elements-twostate overlaps plot above is a boxplot showing the distribution of fractions of the 499 2-state instances.  Note this plot is missing E2 aka Quiescent_Low_Coverage.


Now I am interested in how these two state (pairs, major/minor) are occurring...


```{r}
major_minor_2_LTR %>%
   
  ggplot(., aes(x=major_frac))+
  geom_histogram(binwidth = 0.05)
```
This is the plot of the Major fraction amounts.  A fraction that is labelled major occurs when grouping by te_id and Set (treatment-timepoint) then ordering by the fraction.  The first value is labelled Major, the second value is labelled Minor.  There were no ties (ie 50/50 splits across LTR10E elements)

Lets look at frequency of te_id and states:
```{r}
te_state_frequency_LTR <- LTR10E_state_counts %>%
  count(te_id, n_states, name = "n_sets") %>%
  arrange(te_id, n_states)

te_state_frequency_wide_LTR <- LTR10E_state_counts %>%
  count(te_id, n_states, name = "n_sets") %>%
  mutate(
    n_states = paste0("n_states_", n_states)
  ) %>%
  pivot_wider(
    names_from = n_states,
    values_from = n_sets,
    values_fill = 0
  )
te_state_frequency_wide_MER %>% 
datatable(.,caption="LTR10E state frequency",
               options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE)


LTR10E_state_counts %>%
  pivot_wider(
    id_cols = te_id,
    names_from = Set,
    values_from = n_states
  ) %>% 
  datatable(.,caption="LTR10E state frequency by treatment-timepoint",
               options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE)

```

```{r LTR10E_plot_data}
LTR10E_plot_data <- plot_trajectory(LTR10E_dominant,TE="LTR10E",id_col="te_id",state_col="Broad_state")

acute_LTR10E_dominant <- LTR10E_dominant %>% 
  dplyr::filter(te_id %in% acute_LTR_repid)

acute_LTR10E_plot_data <- plot_trajectory(acute_LTR10E_dominant,TE="LTR10E",id_col="te_id",state_col="Name")


LTR10E_plot_data
acute_LTR10E_plot_data
```
```{r file saving, eval=FALSE}
saveRDS(LTR10E_trajectory_broad,"data/RDS_files/LTR10E_trajectory_broad.RDS")
saveRDS(LTR10E_trajectory_detailed,"data/RDS_files/LTR10E_trajectory_detailed.RDS")
saveRDS(MER61E_trajectory_broad,"data/RDS_files/MER61E_trajectory_broad.RDS")
saveRDS(MER61E_trajectory_detailed,"data/RDS_files/MER61E_trajectory_detailed.RDS")
```

