---
title: "ChromHMM_2"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 400px;
}
```


```{r}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(readxl)
library(ChIPseeker)
# library(regioneR)
library(cowplot)
library(GenomeInfoDb)
library(ggforce)
library(ggalluvial)
```




### Adding annotation to states

```{r annotation}

States_anno <- data.frame("State"=c(paste0("E",1:16)),
"Name"=c("Heterochromatin_1",
"Quiescent_Low_Coverage",
"Repressed_Heterochromatin",
"Repressed_ZFP_genes",
"Strong_Polycomb_Repressed",
 "Bivalent_Enhancer",
"Heterochromatin_2",
"Polycomb_Repressed",
 "Weak_Genic_Enhancer",
 "Bivalent_Poised_TSS",
 "Active_Enhancer",
  "Strong_Genic_Enhancer",
 "Active_TSS",
 "Very_Weak_Transcription",
 "Genic_Strong_Transcription",
 "Genic_Weak_Transcription"), "recode"=c("N","P","M","L","J","I","O","K","G","H","E","F","A","D","B","C")) %>% 
  mutate(recode=factor(recode,levels=c("A","B","C","D","E","F","G","H","I","J","K","L","M","N","O","P")))

state_colors <- c(
  # Promoters (strong → weak)
  "Active_TSS"  = "#d73027",  # strong red
    # Changed
  "Strong_Genic_Enhancer"  = "#ffff60",  
  
  # Enhancers
  "Active_Enhancer"  = "#ffff01",  # bright yellow
  "Weak_Genic_Enhancer"  = "#ffeda0",  # soft yellow
  
  # Transcribed regions
  "Genic_Strong_Transcription"  = "#1a9850",  # dark green
  "Genic_Weak_Transcription"  = "#66bd63",
  "Very_Weak_Transcription"  = "#a6d96a",
  
  # Bivalent / regulatory
  "Bivalent_Poised_TSS"  = "#c51b7d",
  "Bivalent_Enhancer"= "#de77ae",
  
  # Polycomb / repressed
  "Strong_Polycomb_Repressed" = "#762a83",  # deep purple
  "Polycomb_Repressed" = "#af8dc3",
  
  # Heterochromatin / ZNF/repeats
  "Repressed_ZFP_genes" = "#2c7bb6",  # blue
  "Repressed_Heterochromatin" = "#74abe1",
  "Heterochromatin_1"="#abd9e9",
  "Heterochromatin_2"="#abf9f9",
  
  # Quiescent / low signal
  "Quiescent_Low_Coverage" = "#bdbdbd"
 )

name_order <- States_anno %>%
  arrange(recode) %>%
  pull(Name)
```
### Loading interested regions for overlap

```{r}
MER61E_regions <- import("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed")

annotated_H3K27ac_ext200 <- readRDS("data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")
H3K27ac_ROI <- import("data/Bed_exportsH3K27ac_summit_gr_ext200.bed")

cluster_lookup <- annotated_H3K27ac_ext200 %>% 
  dplyr::select(Peakid, cluster)

# H3K27ac_ROI %>% as.data.frame()

# annotated_H3K27ac_ext200 %>% 
#   dplyr::filter(str_detect(repName,"^HERVIP"))
```


### Loading State segment files
```{r}
chrom_24T_VEH <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24T_VEH_16_segments.bed")
chrom_24R_VEH <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24R_VEH_16_segments.bed")
chrom_144R_VEH <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/144R_VEH_16_segments.bed")
chrom_24T_DOX <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24T_DOX_16_segments.bed")
chrom_24R_DOX <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24R_DOX_16_segments.bed")
chrom_144R_DOX <- import("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/144R_DOX_16_segments.bed")


colnames(mcols(chrom_24T_VEH))
```
###  Global MER61E

```{r}
global_hits <- findOverlaps(H3K27ac_ROI, chrom_24T_VEH, ignore.strand = TRUE)

peak_states <- data.frame(
  peak_id = queryHits(global_hits),
  Peak_name = mcols(H3K27ac_ROI)$name[queryHits(global_hits)],
  state = mcols(chrom_24T_VEH)$name[subjectHits(global_hits)],
  width   = width(pintersect(
    H3K27ac_ROI[queryHits(global_hits)],
    chrom_24T_VEH[subjectHits(global_hits)]
  ))
)
peak_states_summed <- peak_states %>%
  group_by(Peak_name, state) %>%
  summarize(total_width = sum(width), .groups = "drop")


# Instead of collapsing to one state per peak, keep the proportions:
peak_states_props <- peak_states_summed %>%
  mutate(frac = total_width / 401)

```

```{r}
peak_states_complete <- peak_states_props %>%
  complete(Peak_name, state, fill = list(frac = 0)) %>%
  left_join(cluster_lookup, by=c("Peak_name"="Peakid")) %>% 
  left_join(States_anno, by=c("state"="State")) %>% 
   mutate(Name = factor(Name, levels = name_order))

plot_df <- peak_states_complete %>% 
  group_by(cluster, Name) %>%
  summarize(mean_frac = mean(frac), .groups = "drop")
  
ggplot(plot_df, aes(x = cluster, y = mean_frac, fill = Name)) +
  geom_bar(stat = "identity") +
  ylab("Mean fraction of peak") +
  theme_classic()+
  scale_fill_manual(values = state_colors) 

ggplot(plot_df ,aes(x = Name, y = cluster, fill = mean_frac)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Mean fraction") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

```{r chromhmm function}
process_chromhmm_raw <- function(chrom_gr, roi_gr) {

  hits <- findOverlaps(roi_gr, chrom_gr, ignore.strand = TRUE)

  peak_states <- data.frame(
    peak_id   = queryHits(hits),
    Peak_name = mcols(roi_gr)$name[queryHits(hits)],
    state     = mcols(chrom_gr)$name[subjectHits(hits)],
    width     = width(pintersect(
      roi_gr[queryHits(hits)],
      chrom_gr[subjectHits(hits)]
    ))
  )

  peak_states_summed <- peak_states %>%
    group_by(Peak_name, state) %>%
    summarize(total_width = sum(width), .groups = "drop")

  peak_states_props <- peak_states_summed %>%
    mutate(frac = total_width / 401)
  
  return(peak_states_props)
}

annotate_states <- function(df) {

  df %>%
    complete(Peak_name, state = unique(States_anno$State),
             fill = list(frac = 0)) %>%
    left_join(cluster_lookup, by = c("Peak_name" = "Peakid")) %>%
    left_join(States_anno, by = c("state" = "State")) %>%
    mutate(Name = factor(Name, levels = name_order))
}
```

```{r}
chrom_list <- list(
  VEH_24T = chrom_24T_VEH,
  VEH_24R  = chrom_24R_VEH,
  VEH_144R  = chrom_144R_VEH,
  DOX_24T  = chrom_24T_DOX,
  DOX_24R  = chrom_24R_DOX,
  DOX_144R  = chrom_144R_DOX
)
results_list <- lapply(chrom_list, function(x)
  annotate_states(process_chromhmm_raw(x, H3K27ac_ROI))
)

all_results <- bind_rows(
  lapply(names(results_list), function(nm) {
    results_list[[nm]] %>%
      mutate(Set = nm)
  })
)
all_results_df <- all_results %>% 
  group_by(Set,cluster, Name) %>%
  summarize(mean_frac = mean(frac), .groups = "drop")


ggplot(all_results_df, aes(x = cluster, y = mean_frac, fill = Name)) +
  geom_bar(stat = "identity") +
  ylab("Mean fraction of peak") +
  theme_classic()+
  facet_wrap(~Set)+
  scale_fill_manual(values = state_colors) 


ggplot(all_results_df ,aes(x = Name, y = cluster, fill = mean_frac)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Mean fraction") +
  theme_minimal() +
  facet_wrap(~Set)+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


### Isolating MER61E

```{r}

### getting PeakIDs of regions that overlap MER61E
MER61E_Peaks <- annotated_H3K27ac_ext200 %>% 
  dplyr::filter(str_detect(repName, "MER61E")) %>% 
  distinct(Peakid)
###Filtering out peakids that do not overlap MER61E elements
MER61E_only_df <- all_results %>% 
  dplyr::filter(Peak_name %in% MER61E_Peaks$Peakid) %>% 
  group_by(Set,cluster, Name) %>%
  summarize(mean_frac = mean(frac), .groups = "drop") %>% 
  mutate(Condition = ifelse(str_detect(Set, "DOX"), "DOX", "VEH"),
         Timepoint = case_when(
           str_detect(Set, "24T")  ~ "24T",
           str_detect(Set, "24R")  ~ "24R",
           str_detect(Set, "144R") ~ "144R")) %>% 
  mutate(Timepoint = factor(Timepoint, levels = c("24T","24R","144R")))

### states I want to keep which contain data
states_to_keep <- MER61E_only_df %>%
  group_by(Name) %>%
  summarize(max_frac = max(mean_frac)) %>%
  filter(max_frac > 0.05) %>%
  pull(Name)
### also filter out not_assigned sets here
MER61E_only_df_filt <- MER61E_only_df %>%
  filter(Name %in% states_to_keep) %>% 
  dplyr::filter(cluster!="not_assigned")

ggplot(MER61E_only_df, aes(x = cluster, y = mean_frac, fill = Name)) +
  geom_bar(stat = "identity") +
  ylab("Mean fraction of peak") +
  theme_classic()+
  facet_wrap(~Set)+
  scale_fill_manual(values = state_colors) 

ggplot(MER61E_only_df ,aes(x = Name, y = cluster, fill = mean_frac)) +
  geom_tile() +
  scale_fill_viridis_c(name = "Mean fraction") +
  theme_minimal() +
  facet_wrap(~Set)+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )



ggplot(MER61E_only_df_filt,
       aes(x = Timepoint,
           y = mean_frac,
           group = Condition,
           color = Condition,
           linetype = Condition)) +
  geom_line(size = 1) +
  geom_point() +
  facet_grid(cluster ~ Name) +
  scale_linetype_manual(values = c("DOX" = "solid", "VEH" = "dotted")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```


```{r}
delta_df <- MER61E_only_df_filt %>%
  select(Condition, Timepoint, cluster, Name, mean_frac) %>%
   tidyr::complete(Timepoint, cluster, Name, Condition,
                  fill = list(mean_frac = 0)) %>% 
  pivot_wider(names_from = Condition, values_from = mean_frac) %>% 
  mutate(delta= DOX-VEH)

summary(delta_df$DOX - delta_df$VEH)

ggplot(delta_df,
       aes(x = Timepoint, y = delta, group = Name, color = Name)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line() +
  geom_point() +
  facet_wrap(~cluster) +
  theme_minimal()+
  scale_color_manual(values = state_colors) 

```

```{r, fig.width=10}

### To look at MER61E only, filter out and reduce down to dominant state per region This is done without respect to cluster relationship
peak_state_calls <- all_results %>%
  filter(Peak_name %in% MER61E_Peaks$Peakid) %>%
  group_by(Peak_name, Set, Name) %>%
  summarize(frac = sum(frac), .groups = "drop") %>%
  group_by(Peak_name, Set) %>%
  slice_max(frac, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  left_join(
    distinct(all_results[, c("Peak_name", "cluster")]),
    by = "Peak_name"
  ) %>%
  mutate(
    Condition = ifelse(grepl("DOX", Set), "DOX", "VEH"),
    Timepoint = case_when(
      grepl("24T", Set) ~ "24T",
      grepl("24R", Set) ~ "24R",
      grepl("144R", Set) ~ "144R"))

state_transitions <- peak_state_calls %>%
  select(Peak_name, Condition, Timepoint, Name, cluster) %>%
  pivot_wider(names_from = Timepoint, values_from = Name) 


ggplot(peak_state_calls,
       aes(x = Timepoint, y = Name, fill = Condition)) +
  geom_tile() +
  facet_wrap(~cluster) +
  theme_minimal()


ggplot(state_transitions,
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.5) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  facet_wrap(~Condition) +
  theme_minimal()
```

```{r}
state_map <- c(
  "Active_TSS" = "Active",
  "Very_Weak_Transcription"="Active",
  "Strong_Genic_Enhancer" = "Enhancer",
  "Repressed_ZFP_genes" = "Repressed",
  "Active_Enhancer" = "Enhancer",
  "Weak_Genic_Enhancer" = "Enhancer",
  "Bivalent_Poised_TSS" = "Poised",
  "Polycomb_Repressed" = "Repressed",
  "Strong_Polycomb_Repressed" = "Repressed",
  "Repressed_Heterochromatin" = "Repressed",
  "Heterochromatin_1" = "Repressed",
  "Heterochromatin_2" = "Repressed",
  "Quiescent_Low_Coverage" = "Quiescent"
)

peak_states <- all_results %>%
  filter(Peak_name %in% MER61E_Peaks$Peakid) %>%
  group_by(Peak_name, Set) %>%
  slice_max(frac, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    Condition = ifelse(grepl("DOX", Set), "DOX", "VEH"),
    Timepoint = case_when(
      grepl("24T", Set) ~ "24T",
      grepl("24R", Set) ~ "24R",
      grepl("144R", Set) ~ "144R"))
trajectory <- peak_states %>%
  select(Peak_name, Condition, Timepoint, Name, cluster) %>%
  tidyr::pivot_wider(names_from = Timepoint, values_from = Name) %>% 
  mutate(across(c(`24T`,`24R`,`144R`),
                ~ unname(state_map[.x])))
ggplot(filter(trajectory, Condition == "DOX"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.6) +
  geom_stratum(width = 0.2) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
  ggtitle("MER61E state transitions (DOX)") +
  theme_minimal()

ggplot(filter(trajectory, Condition == "VEH"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.6) +
  geom_stratum(width = 0.2) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
  ggtitle("MER61E state transitions (VEH)") +
  theme_minimal()
```


Cases are incomplete.  Time to filter out. note there are 68 MER61E overlapping regions in my data.  For quantification purposes, 16 are not assigned a cluster, 40 are assigned Acute cluster (set 3) and 12 are assigned non response cluster (set 1)

Why are they incomplete:  maybe there is reduced chromatin state confidence?  I have tested to make sure that there is
```{r, clean}

trajectory_clean <- trajectory %>%
  filter(complete.cases(.))

ggplot(filter(trajectory_clean, Condition == "DOX"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.6) +
  geom_stratum(width = 0.2) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
  ggtitle("MER61E state transitions (DOX)") +
  theme_minimal()

ggplot(filter(trajectory_clean, Condition == "VEH"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`), alpha = 0.6) +
  geom_stratum(width = 0.2) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
  ggtitle("MER61E state transitions (VEH)") +
  theme_minimal()
```

```{r}
qc_df <- trajectory %>%
  mutate(status = case_when(
    if_all(c(`24T`, `24R`, `144R`), ~ !is.na(.)) ~ "complete",
    if_all(c(`24T`, `24R`, `144R`), is.na) ~ "missing",
    TRUE ~ "partial"
  ))

qc_df %>% group_by(Condition,status) %>% tally

qc_peak <- trajectory %>%
  mutate(status = case_when(
    if_all(c(`24T`, `24R`, `144R`), ~ !is.na(.)) ~ "complete",
    if_all(c(`24T`, `24R`, `144R`), is.na) ~ "missing",
    TRUE ~ "partial"
  )) %>%
  distinct(Peak_name, status)

qc_counts <- qc_peak %>%
  count(status)
trajectory %>%
  summarise(
    n_peaks = dplyr::n(),
    complete = sum(complete.cases(.)),
    partially_missing = sum(!complete.cases(.))
  )

ggplot(qc_counts, aes(x = status, y = n, fill = status)) +
  geom_col(width = 0.6) +
  theme_minimal() +
  labs(
    x = NULL,
    y = "Number of MER61E peaks",
    title = "ChromHMM state coverage completeness"
  ) +
  theme(
    legend.position = "none"
  )
```

```{r}
traj <- trajectory %>%
  mutate(status = case_when(
    if_all(c(`24T`,`24R`,`144R`), ~ !is.na(.)) ~ "complete",
    TRUE ~ "partial"
  ))
traj_complete <- trajectory %>%
  filter(if_all(c(`24T`,`24R`,`144R`), ~ !is.na(.)))

ggplot(traj_complete,
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = `24T`)) +
  geom_stratum() +
  facet_wrap(~Condition) +
  theme_minimal()
```
```{r}
plot_trajectory_TE_spec <- function(anno_df,results_df, TE) {
  state_colors <- c(
  "Active" = "#D73027",
  "Enhancer" = "#FC8D59",
  "Poised" = "#FEE08B",
  "Repressed" = "#4575B4",
  "Quiescent" = "#D9D9D9")
  
  ## retrieve the peakIDs with associated TE repName
  
  peak_ids <- anno_df %>% 
  dplyr::filter(str_detect(repName, TE)) %>% 
  pull(Peakid)
  
  ## filter results with peak_ids and break into Timepoint and Condition
  peak_states <- results_df %>%
  filter(Peak_name %in% peak_ids) %>%
  group_by(Peak_name, Set, Name) %>%
  summarize(frac = sum(frac), .groups = "drop") %>%
  group_by(Peak_name, Set) %>%
  arrange(desc(frac), .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  left_join(
    distinct(results_df %>% select(Peak_name, cluster)),
    by = "Peak_name"
  ) %>%
  mutate(
    Condition = ifelse(grepl("DOX", Set), "DOX", "VEH"),
    Timepoint = case_when(
      grepl("24T", Set) ~ "24T",
      grepl("24R", Set) ~ "24R",
      grepl("144R", Set) ~ "144R"))
  
  ## make trajectory dataframe
trajectory <- peak_states %>%
  select(Peak_name, Condition, Timepoint, Name, cluster) %>%
  pivot_wider(
    id_cols = c(Peak_name, Condition, cluster),
    names_from = Timepoint,
    values_from = Name
  ) %>%
  mutate(across(c(`24T`,`24R`,`144R`),
                ~ unname(state_map[.x])))


DOX_plot <- ggplot(filter(trajectory, Condition == "DOX"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
 geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 3) +
  ggtitle(paste0(TE," state transitions (DOX)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)

VEH_plot <- ggplot(filter(trajectory, Condition == "VEH"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 3) +
  ggtitle(paste0(TE," state transitions (VEH)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)
#### clean plots:
trajectory_clean <- trajectory %>%
  filter(complete.cases(.))

DOX_clean <- ggplot(filter(trajectory_clean, Condition == "DOX"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
   ggtitle(paste0(TE," state transitions (DOX)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)

VEH_clean <- ggplot(filter(trajectory_clean, Condition == "VEH"),
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), size = 3) +
   ggtitle(paste0(TE," state transitions (VEH)")) +
  theme_minimal()+
  scale_fill_manual(values = state_colors)



#### calculating incompletes and QC
traj <- trajectory %>%
  mutate(status = case_when(
    if_all(c(`24T`,`24R`,`144R`), ~ !is.na(.)) ~ "complete",
    TRUE ~ "partial"
  ))

traj_complete <- traj %>%
  filter(status == "complete")

completion_plot <- ggplot(traj_complete,
       aes(axis1 = `24T`,
           axis2 = `24R`,
           axis3 = `144R`,
           y = 1)) +
  geom_alluvium(aes(fill = after_stat(stratum)), alpha = 0.4)+
  geom_stratum(aes(fill = after_stat(stratum))) + # color changes per axis
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 3) +
  facet_wrap(~Condition) +
  theme_minimal()+
  ggtitle(paste0("Complete cases of ",TE))+
  scale_fill_manual(values = state_colors)
#### MORE QC

qc_df <- trajectory %>%
  mutate(status = case_when(
    complete.cases(across(c(`24T`,`24R`,`144R`))) ~ "complete",
    rowSums(is.na(across(c(`24T`,`24R`,`144R`)))) == 3 ~ "missing",
    TRUE ~ "partial"
  ))

qc_counts <- qc_df %>%
  distinct(Peak_name, status) %>%
  count(status)

traj_incomp <- trajectory %>%
  summarise(
    n_peaks = dplyr::n(),
    complete = sum(complete.cases(.)),
    partially_missing = sum(!complete.cases(.))
  )

completeness_plot <- ggplot(qc_counts, aes(x = status, y = n, fill = status)) +
  geom_col(width = 0.6) +
  theme_minimal() +
  labs(
    x = NULL,
    y = paste0("Number of ",TE," peaks"),
    title =paste0("ChromHMM state coverage completeness for ",TE)
  ) +
  theme(
    legend.position = "none"
  )


return(list(
  DOX_plot = DOX_plot,
  VEH_plot = VEH_plot,
  DOX_clean = DOX_clean,
  VEH_clean = VEH_clean,
  completion_plot = completion_plot,
  QC_data= qc_df,
  QC_data_2=qc_counts,
  QC_missing=traj_incomp,
  qc_plot=completeness_plot))
}
```

```{r}
MER61A_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61A")

MER61B_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61B")

MER61C_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61C")
MER61D_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61D")

MER61E_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61E")

MER61F_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61F")
MER61_int_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "MER61-int")



```
```{r MER61B}
MER61A_results
MER61B_results
MER61C_results
MER61D_results
MER61E_results
MER61F_results
MER61_int_results
```

```{r}

LTR10E_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "LTR10E")

LTR10E_results
```

```{r}
LTR10B1_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "LTR10B1")

LTR10B1_results

LTR10B2_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "LTR10B2")

LTR10B2_results

HERVIP10B3_int_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "HERVIP10B3-int")

HERVIP10B3_int_results

HERV1_LTRa_results <- plot_trajectory_TE_spec(annotated_H3K27ac_ext200, all_results, "HERV1_LTRa")

HERV1_LTRa_results


```

