---
title: "ChromHMM"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 400px;
}
```


```{r}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(readxl)
library(ChIPseeker)
```

This is the AIC/BIC plots genrated to figure out how many states we wanted to use for our modeling of our data:
```{r TACC generated}
aic_bic_table_TACC <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TACC_models/model_selection_summary.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

aic_bic_table_TACC %>% 
   pivot_longer(cols=c(AIC,BIC), names_to = "test_type",values_to = "calculation") %>% 
  ggplot(.,aes(x=States,y=calculation))+
  geom_point()+
  geom_line()+
  facet_wrap(~test_type)
  
```

The results point to 16 states yielding the best results.

I ran the models for 16 states and here is the outcome of those results:

Here is the emission table of the results:
```{r emission of 16 states}

emission_16state <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TACC_models/Chrom_model_16states_final/emissions_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

emission_16state %>% 
  mutate(`State (Emission order)`=factor(`State (Emission order)`, levels=c(1:16))) %>% 
  pivot_longer(., !`State (Emission order)`, names_to = "Histone",values_to = "emission") %>% 
  mutate(Histone=factor(Histone,levels=c("H3K27ac","H3K36me3","H3K27me3","H3K9me3"))) %>% 
  ggplot(., aes(x=Histone, y=`State (Emission order)`, fill=emission))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(expand = c(0, 0)) +
scale_y_discrete(expand = c(0, 0), limits = rev)+
  coord_fixed()+
    theme_classic()+
  ggtitle("Emission Parameters")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Here is the transition plot of the results:

```{r transition of 16 states}

transition_16state <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TACC_models/Chrom_model_16states_final/transitions_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

transition_16state %>% 
  mutate(`State (from\\to) (Emission order)`=factor(`State (from\\to) (Emission order)`, levels=c(1:16)))%>%
  dplyr::rename(state_from_to=`State (from\\to) (Emission order)`) %>% 
  pivot_longer(., !state_from_to, names_to = "States",values_to = "transition") %>% 
  mutate(States=factor(States, levels=c(1:16))) %>% 
  
  ggplot(., aes(x=States, y=state_from_to, fill=transition))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(expand = c(0, 0)) +
scale_y_discrete(expand = c(0, 0), limits = rev)+
  coord_fixed()+
    theme_classic()+
  ggtitle("Transition Parameters")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
### Enrichment across the states:
looking at enrichment of Features across the states by each group.  This helps with the classification and interpretation of what each state represents.

```{r loading other features enrichment}
DOX_24T_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 
```
```{r loading gene enrichment}
DOX_24T_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 all_genes_states <- 
  VEH_144R_gene %>% 
  bind_rows(DOX_144R_gene) %>% 
   bind_rows(DOX_24R_gene) %>% 
   bind_rows(VEH_24R_gene) %>% 
   bind_rows(DOX_24T_gene) %>% 
   bind_rows(VEH_24T_gene) 
 
```
```{r loading ZFP_run data}
DOX_24T_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 all_ZFPs_states <- 
  VEH_144R_ZFP %>% 
  bind_rows(DOX_144R_ZFP) %>% 
   bind_rows(DOX_24R_ZFP) %>% 
   bind_rows(VEH_24R_ZFP) %>% 
   bind_rows(DOX_24T_ZFP) %>% 
   bind_rows(VEH_24T_ZFP) 
 
 just_ZFP <- all_ZFPs_states[,c("State (Emission order)","ZFP_proteins.bed","group")]
 just_ZFP %>% 
   dplyr::filter(`State (Emission order)` !="Base")

```

```{r making the large plotting dataframe}
all_files_states <- 
  VEH_144R_full %>% 
  bind_rows(DOX_144R_full) %>% 
   bind_rows(DOX_24R_full) %>% 
   bind_rows(VEH_24R_full) %>% 
   bind_rows(DOX_24T_full) %>% 
   bind_rows(VEH_24T_full) 

long_file <- all_files_states %>% 
  dplyr::select(`State (Emission order)`,
                `Genome %`,
                `GRCh38-cCREs.bed`,
                `SCREEN_hg38_CA-CTCF.bed`:SCREEN_hg38_pELS.bed,
                Set_1_H3K27ac_ROI.bed:Set_3_H3K27ac_ROI.bed,
                all_H3K27ac_H3K27ac_ROI.bed,
                LINE_rptmasker.bed,
                SINE_rptmasker.bed,
                LTR_rptmasker.bed,
                DNA_rptmasker.bed,
                Retroposon_rptmasker.bed,
                RC_rptmasker.bed,
                Low_complexity_rptmasker.bed,
                RNA_rptmasker.bed,
                Satellite_rptmasker.bed,
                Simple_repeat_rptmasker.bed,
                Unknown_rptmasker.bed, 
                rRNA_rptmasker.bed:group) %>% 
  mutate(`State (Emission order)`=factor(`State (Emission order)`, levels = c(paste0("E", 1:16),"Base"))) %>%
  dplyr::filter(`State (Emission order)` != "Base") %>% 
  pivot_longer(., cols=!c(`State (Emission order)`,group), names_to = "Identity",values_to = "Fold_enrichment") %>% 
  mutate(Identity=str_remove(Identity,".bed")) %>% 
  mutate(Identity=factor(Identity, 
                         levels=c("Genome %",
                                  "GRCh38-cCREs",
                                  "SCREEN_hg38_CA-CTCF",
                                  "SCREEN_hg38_CA-H3K4me3",
                                  "SCREEN_hg38_CA-TF",
                                  "SCREEN_hg38_CA",
                                  "SCREEN_hg38_PLS",
                                  "SCREEN_hg38_TF",
                                  "SCREEN_hg38_pELS",
                                  "SCREEN_hg38_dELS",
                                  "all_H3K27ac_H3K27ac_ROI",
                                  "Set_1_H3K27ac_ROI",
                                  "Set_2_H3K27ac_ROI",
                                  "Set_3_H3K27ac_ROI",
                                  "LINE_rptmasker",
                                  "SINE_rptmasker",
                                  "LTR_rptmasker",
                                  "DNA_rptmasker",
                                  "Retroposon_rptmasker",
                                  "RC_rptmasker",
                                  "Low_complexity_rptmasker",
                                  "RNA_rptmasker",
                                  "Satellite_rptmasker",
                                  "Simple_repeat_rptmasker",
                                  "Unknown_rptmasker", 
                                  "rRNA_rptmasker",
                                  "scRNA_rptmasker",
                                  "snRNA_rptmasker",
                                  "srpRNA_rptmasker",
                                  "tRNA_rptmasker"))) %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R")))

```

```{r plot dataframe for genes}

# columns to pivot (exclude ID columns)
id_cols <- c("State (Emission order)", "group")

all_genes_states_long <-
all_genes_states %>%
  mutate(`State (Emission order)`=paste0("E",`State (Emission order)`)) %>% 
  left_join(just_ZFP) %>% 
  
  mutate(`State (Emission order)` = factor(`State (Emission order)`,
                                           levels = c(paste0("E",1:16),"EBase"))) %>%
  filter(`State (Emission order)` != "EBase") %>%
  
  pivot_longer(
    cols = -all_of(id_cols),
    names_to = "Identity",
    values_to = "Fold_enrichment"
  ) %>%
  # preserve original order
  mutate(Identity = factor(Identity, levels =c( names(all_genes_states)[!names(all_genes_states) %in% id_cols],"ZFP_proteins.bed"))) %>% 
         # create cleaned label for plotting
         mutate(Identity_label = case_when(
      str_detect(Identity, "\\.hg38\\.bed\\.gz$") ~ str_remove(Identity, "\\.hg38\\.bed\\.gz$"),
      str_detect(Identity, "\\.bed$") ~ str_remove(Identity, "\\.bed$"),
      TRUE ~ Identity   # keep everything else unchanged
    ),
    group = factor(group,
                        levels = c("VEH_24T","VEH_24R","VEH_144R",
                                   "DOX_24T","DOX_24R","DOX_144R"))) 

  
final_order <- unique(all_genes_states_long$Identity_label)

all_genes_states_long <- all_genes_states_long %>% 
  mutate(Identity_label=factor(Identity_label,levels=final_order))

```

### Looking a genomic features:

```{r plot all genes states, fig.width=12, fig.height=6}

all_genes_states_long %>% 
  group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>% 
    ggplot(.,aes(x=Identity_label,y=`State (Emission order)`, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=1, ncol=6)+
    scale_y_discrete(limits=rev)+
  ggtitle("Genic region enrichment across states")+
  coord_fixed()


```


```{r enchacers plot, fig.height=8}
long_file %>% 
  dplyr::filter(stringr::str_detect(Identity, "SCREEN")) %>% 
   group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>% 
    ggplot(.,aes(x=Identity,y=`State (Emission order)`, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=1, ncol=6)+
    scale_y_discrete(limits=rev)+
  ggtitle("Enhancer enrichment acros states")+
  coord_fixed()
```
### Cormotif set enrichment across states  

Each Histone will have slices of their own states.  Here are all sets next to all states


```{r fig.width=12, fig.height=8}
order <- c("Genome %",
"all_H3K27ac_H3K27ac_ROI.bed",
"Set_1_H3K27ac_ROI.bed" ,
"Set_2_H3K27ac_ROI.bed",
"Set_3_H3K27ac_ROI.bed",
"all_H3K36me3_regions_H3K36me3_ROI.bed",
"Set_1_H3K36me3_ROI.bed",
"Set_2_H3K36me3_ROI.bed",
"all_H3K27me3_regions_H3K27me3_ROI.bed",
"Set_1_H3K27me3_ROI.bed",               
"all_H3K9me3_regions_H3K9me3_ROI.bed",           
"Set_2_H3K27me3_ROI.bed" ,              
"Set_1_H3K9me3_ROI.bed",               
"Set_2_H3K9me3_ROI.bed", 
"Set_3_H3K9me3_ROI.bed" ,               
"ZFP_proteins.bed")                  

all_ZFPs_states %>%
  mutate(
    `State (Emission order)` = as.character(`State (Emission order)`),
    `State (Emission order)` = case_when(
      `State (Emission order)` %in% as.character(1:16) ~ paste0("E", `State (Emission order)`),
      TRUE ~ `State (Emission order)`)) %>% 
  mutate(`State (Emission order)` = factor(`State (Emission order)`,
                                           levels = c(paste0("E",1:16), "Base"))) %>%
  filter(`State (Emission order)` != "Base") %>% 
  pivot_longer(., -c(`State (Emission order)`,group), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) %>% 
  mutate(Identity=factor(Identity, levels=order)) %>% 
  dplyr::filter(Identity!="ZFP_proteins.bed") %>% 
  group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>% 
    ggplot(.,aes(x=Identity,y=`State (Emission order)`, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=2, ncol=3)+
    scale_y_discrete(limits=rev)+
  ggtitle("Cormotif across states")+
  coord_fixed()

  
  
```


