---
title: "ChromHMM"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 400px;
}
```


```{r}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(readxl)
library(ChIPseeker)
library(regioneR)
library(patchwork)
library(GenomeInfoDb)
```

This is the AIC/BIC plots genrated to figure out how many states we wanted to use for our modeling of our data:
```{r TACC generated}
aic_bic_table_TACC <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TACC_models/model_selection_summary.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

aic_bic_table_TACC %>% 
   pivot_longer(cols=c(AIC,BIC), names_to = "test_type",values_to = "calculation") %>% 
  ggplot(.,aes(x=States,y=calculation))+
  geom_point()+
  geom_line()+
  facet_wrap(~test_type)
  
```

The results point to 16 states yielding the best results.

I ran the models for 16 states and here is the outcome of those results:

Here is the emission table of the results:
```{r emission of 16 states}

emission_16state <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TACC_models/Chrom_model_16states_final/emissions_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

emission_16state %>% 
  mutate(`State (Emission order)`=factor(`State (Emission order)`, levels=c(1:16))) %>% 
  pivot_longer(., !`State (Emission order)`, names_to = "Histone",values_to = "emission") %>% 
  mutate(Histone=factor(Histone,levels=c("H3K27ac","H3K36me3","H3K27me3","H3K9me3"))) %>% 
  ggplot(., aes(x=Histone, y=`State (Emission order)`, fill=emission))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(expand = c(0, 0)) +
scale_y_discrete(expand = c(0, 0), limits = rev)+
  coord_fixed()+
    theme_classic()+
  ggtitle("Emission Parameters")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Here is the transition plot of the results:

```{r transition of 16 states}

transition_16state <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TACC_models/Chrom_model_16states_final/transitions_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

transition_16state %>% 
  mutate(`State (from\\to) (Emission order)`=factor(`State (from\\to) (Emission order)`, levels=c(1:16)))%>%
  dplyr::rename(state_from_to=`State (from\\to) (Emission order)`) %>% 
  pivot_longer(., !state_from_to, names_to = "States",values_to = "transition") %>% 
  mutate(States=factor(States, levels=c(1:16))) %>% 
  
  ggplot(., aes(x=States, y=state_from_to, fill=transition))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(expand = c(0, 0)) +
scale_y_discrete(expand = c(0, 0), limits = rev)+
  coord_fixed()+
    theme_classic()+
  ggtitle("Transition Parameters")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### Adding annotation to states

```{r annotation}

States_anno <- data.frame("State"=c(paste0("E",1:16)),
"Name"=c("Heterochromatin_1",
"Quiescent_Low_Coverage",
"Repressed_Heterochromatin",
"Repressed_ZFP_regions",
"Strong_Polycomb_Repressed",
 "Bivalent_Enhancer",
"Heterochromatin_2",
"Bivalent_Poised_TSS2",
 "Weak_Genic_Enhancer",
 "Strong_Genic_Enhancer1",
 "Active_Proximal_Enhancer",
  "Strong_Genic_Enhancer2",
 "Active_TSS",
 "Very_Weak_Transcription",
 "Genic_Strong_Transcription",
 "Genic_Weak_Transcription"), "recode"=c("N","P","M","L","K","J","O","I","H","F","E","G","A","D","B","C")) %>% 
  mutate(recode=factor(recode,levels=c("A","B","C","D","E","F","G","H","I","J","K","L","M","N","O","P")))

# SOIs <- c("E4","E10","E11","E12","E13")

# States_anno

```

### Enrichment across the states:
looking at enrichment of Features across the states by each group.  This helps with the classification and interpretation of what each state represents.

```{r loading other features enrichment}
DOX_24T_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 
```
```{r loading gene enrichment}
DOX_24T_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 

 all_genes_states <- 
  VEH_144R_gene %>% 
  bind_rows(DOX_144R_gene) %>% 
   bind_rows(DOX_24R_gene) %>% 
   bind_rows(VEH_24R_gene) %>% 
   bind_rows(DOX_24T_gene) %>% 
   bind_rows(VEH_24T_gene) 
 
```

```{r loading ZFP_run data}
DOX_24T_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 all_ZFPs_states <- 
  VEH_144R_ZFP %>% 
  bind_rows(DOX_144R_ZFP) %>% 
   bind_rows(DOX_24R_ZFP) %>% 
   bind_rows(VEH_24R_ZFP) %>% 
   bind_rows(DOX_24T_ZFP) %>% 
   bind_rows(VEH_24T_ZFP) 
 
 just_ZFP <- all_ZFPs_states[,c("State (Emission order)","ZFP_proteins.bed","group")]
 just_ZFP %>% 
   dplyr::filter(`State (Emission order)` !="Base")

```

```{r loading ATAC_run}

DOX_24T_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 all_ATAC_states <- 
  VEH_144R_ATAC %>% 
  bind_rows(DOX_144R_ATAC) %>% 
   bind_rows(DOX_24R_ATAC) %>% 
   bind_rows(VEH_24R_ATAC) %>% 
   bind_rows(DOX_24T_ATAC) %>% 
   bind_rows(VEH_24T_ATAC) %>% 
   left_join(.,States_anno, by = c(`State (Emission order)`="State"))
 
```


```{r making the large plotting dataframe}
all_files_states <- 
  VEH_144R_full %>% 
  bind_rows(DOX_144R_full) %>% 
   bind_rows(DOX_24R_full) %>% 
   bind_rows(VEH_24R_full) %>% 
   bind_rows(DOX_24T_full) %>% 
   bind_rows(VEH_24T_full) %>% 
  left_join(.,States_anno, by = c(`State (Emission order)`="State"))

long_file <- all_files_states %>% 
  dplyr::select(`State (Emission order)`,
                `Genome %`,
                `GRCh38-cCREs.bed`,
                `SCREEN_hg38_CA-CTCF.bed`:SCREEN_hg38_pELS.bed,
                Set_1_H3K27ac_ROI.bed:Set_3_H3K27ac_ROI.bed,
                all_H3K27ac_H3K27ac_ROI.bed,
                LINE_rptmasker.bed,
                SINE_rptmasker.bed,
                LTR_rptmasker.bed,
                DNA_rptmasker.bed,
                Retroposon_rptmasker.bed,
                RC_rptmasker.bed,
                Low_complexity_rptmasker.bed,
                RNA_rptmasker.bed,
                Satellite_rptmasker.bed,
                Simple_repeat_rptmasker.bed,
                Unknown_rptmasker.bed, 
                rRNA_rptmasker.bed:group,Name,recode) %>% 
  mutate(`State (Emission order)`=factor(`State (Emission order)`, levels = c(paste0("E", 1:16),"Base"))) %>%
  dplyr::filter(`State (Emission order)` != "Base") %>% 
  pivot_longer(., cols=!c(`State (Emission order)`,group,Name,recode), names_to = "Identity",values_to = "Fold_enrichment") %>% 
  mutate(Identity=str_remove(Identity,".bed")) %>% 
  mutate(Identity=factor(Identity, 
                         levels=c("Genome %",
                                  "GRCh38-cCREs",
                                  "SCREEN_hg38_CA-CTCF",
                                  "SCREEN_hg38_CA-H3K4me3",
                                  "SCREEN_hg38_CA-TF",
                                  "SCREEN_hg38_CA",
                                  "SCREEN_hg38_PLS",
                                  "SCREEN_hg38_TF",
                                  "SCREEN_hg38_pELS",
                                  "SCREEN_hg38_dELS",
                                  "all_H3K27ac_H3K27ac_ROI",
                                  "Set_1_H3K27ac_ROI",
                                  "Set_2_H3K27ac_ROI",
                                  "Set_3_H3K27ac_ROI",
                                  "LINE_rptmasker",
                                  "SINE_rptmasker",
                                  "LTR_rptmasker",
                                  "DNA_rptmasker",
                                  "Retroposon_rptmasker",
                                  "RC_rptmasker",
                                  "Low_complexity_rptmasker",
                                  "RNA_rptmasker",
                                  "Satellite_rptmasker",
                                  "Simple_repeat_rptmasker",
                                  "Unknown_rptmasker", 
                                  "rRNA_rptmasker",
                                  "scRNA_rptmasker",
                                  "snRNA_rptmasker",
                                  "srpRNA_rptmasker",
                                  "tRNA_rptmasker"))) %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) 

```

```{r plot dataframe for genes}

# columns to pivot (exclude ID columns)
id_cols <- c("State (Emission order)", "group","Name","recode")

all_genes_states_long <-
all_genes_states %>%
  mutate(`State (Emission order)`=paste0("E",`State (Emission order)`)) %>% 
  left_join(just_ZFP) %>% 
  left_join(.,States_anno, by = c(`State (Emission order)`="State")) %>% 
  
  mutate(`State (Emission order)` = factor(`State (Emission order)`,
                                           levels = c(paste0("E",1:16),"EBase"))) %>%
  filter(`State (Emission order)` != "EBase") %>%
  
  pivot_longer(
    cols = -all_of(id_cols),
    names_to = "Identity",
    values_to = "Fold_enrichment"
  ) %>%
  # preserve original order
  mutate(Identity = factor(Identity, levels =c( names(all_genes_states)[!names(all_genes_states) %in% id_cols],"ZFP_proteins.bed"))) %>% 
         # create cleaned label for plotting
         mutate(Identity_label = case_when(
      str_detect(Identity, "\\.hg38\\.bed\\.gz$") ~ str_remove(Identity, "\\.hg38\\.bed\\.gz$"),
      str_detect(Identity, "\\.bed$") ~ str_remove(Identity, "\\.bed$"),
      TRUE ~ Identity   # keep everything else unchanged
    ),
    group = factor(group,
                        levels = c("VEH_24T","VEH_24R","VEH_144R",
                                   "DOX_24T","DOX_24R","DOX_144R"))) 

  
final_order <- unique(all_genes_states_long$Identity_label)

all_genes_states_long <- all_genes_states_long %>% 
  mutate(Identity_label=factor(Identity_label,levels=final_order))

```

### Looking a genomic features:

```{r plot all genes states, fig.width=12, fig.height=6}

all_genes_states_long %>% 
  group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>%
  arrange(recode) %>%
  mutate(
    Name = factor(Name,
                  levels = unique(Name[order(recode)])) ) %>% 
 
    # ggplot(.,aes(x=Identity_label,y=`State (Emission order)`, fill=chromhmm_scaled))+
  ggplot(.,aes(x=Identity_label,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=1, ncol=6)+
    scale_y_discrete(limits=rev)+
  ggtitle("Genic region enrichment across states")+
  coord_fixed()


```
The genic enrichment across states plots shows the exact enrichment across each condition, however, for summary purposes, I wish to know the general enrichment from all conditions.  I will take the geometric mean  across  DOX and VEH conditions (and all conditions) at each timepoint for each state and identity ( in this case for the CpGIslands, Exons, genes, TESs and TSS etc..).  This treats the fold-change enrichment and depletion symmetrically about 1 because enrichment behaves multiplicatively since it is already a ratio.  I am only using this to determine which states are constitutively TSS-enriched, with are CPG-biased, which are ZFP enriched etc...
```{r geometric mean, all conditions}
geom_mean_df <- all_genes_states_long %>% 
  group_by(Identity, Identity_label, `State (Emission order)`) %>% 
  summarize(geom_FE = exp(mean(log(Fold_enrichment + 1e-6), na.rm = TRUE)),
    .groups = "drop") %>% 
  group_by(Identity) %>% 
             mutate(
    FE_min = min(geom_FE, na.rm = TRUE),
    FE_max = max(geom_FE, na.rm = TRUE),
    chromhmm_scaled = (geom_FE - FE_min) / (FE_max - FE_min)) 
  
  ggplot(geom_mean_df,aes(x=Identity_label,y=`State (Emission order)`, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  
    scale_y_discrete(limits=rev)+
  ggtitle("Genic region relative enrichment pattern across conditions\n (geometric mean)")+
  coord_fixed()

```
```{r geom_mean unscaled log2}
geom_mean_df <- geom_mean_df %>%  mutate(log2_FE = log2(geom_FE))

ggplot(geom_mean_df,
       aes(x = Identity_label,
           y = `State (Emission order)`,
           fill = log2_FE)) +

  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "darkgreen",
    mid = "white",
    high = "#08306B",
    midpoint = 0,
    name = "log2(Fold Enrichment)"
  ) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +

  ggtitle("Geometric Mean ChromHMM Enrichment Across Conditions")+
  coord_fixed()
```


#### Poster future plots

```{r poster plot data frame}
full_states_data_long <- bind_rows(all_genes_states_long,long_file)

Identity_label_order <- c("Genome %", "CpGIsland","RefSeqGene","RefSeqTSS2kb","RefSeqTSS","RefSeqExon","RefSeqTES",   
      "ZFP_proteins","PLS","pELS","dELS")

geom_df_reorderd <-
full_states_data_long %>% 
  mutate(Identity_label = case_when(
      str_detect(Identity, "\\.hg38\\.bed\\.gz$") ~ str_remove(Identity, "\\.hg38\\.bed\\.gz$"),
      str_detect(Identity, "\\.bed$") ~ str_remove(Identity, "\\.bed$"),
      str_detect(Identity,"^SCREEN_hg38_")~ str_remove(Identity,"^SCREEN_hg38_"),
       str_detect(Identity,"_rptmasker")~ str_remove(Identity,"_rptmasker"),
      Identity =="all_H3K27ac_H3K27ac_ROI"~ "all_H3K27ac_regions",
      TRUE ~ Identity   # keep everything else unchanged
    ),
    group = factor(group,
                        levels = c("VEH_24T","VEH_24R","VEH_144R",
                                   "DOX_24T","DOX_24R","DOX_144R"))) %>% 
  dplyr::filter(Identity_label %in% Identity_label_order) %>% 
  arrange(recode) %>%
  mutate(
    Name = factor(Name,
                  levels = unique(Name[order(recode)])),
    Identity_label=factor(Identity_label,levels=Identity_label_order)) %>% 
   group_by(Identity, Identity_label, Name) %>% 
  summarize(geom_FE = exp(mean(log(Fold_enrichment + 1e-6), na.rm = TRUE)),
    .groups = "drop") %>% 
  group_by(Identity) %>% 
             mutate(
    FE_min = min(geom_FE, na.rm = TRUE),
    FE_max = max(geom_FE, na.rm = TRUE),
    chromhmm_scaled = (geom_FE - FE_min) / (FE_max - FE_min)) 


ggplot(geom_df_reorderd,aes(x=Identity_label,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  
    scale_y_discrete(limits=rev)+
  ggtitle("Genic region relative enrichment pattern across conditions\n (geometric mean)")+
  coord_fixed()

  
```

```{r suppdata for poster}
geom_df_reorderd <- geom_df_reorderd %>%  mutate(log2_FE = log2(geom_FE))

ggplot(geom_df_reorderd,
       aes(x = Identity_label,
           y = Name,
           fill = log2_FE)) +

  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "darkgreen",
    mid = "white",
    high = "#08306B",
    midpoint = 0,
    name = "log2(Fold Enrichment)"
  ) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +

  ggtitle("Geometric Mean ChromHMM Enrichment Across Conditions")+
  coord_fixed()
```


Now I want to look at the change of log2 Foldenrichment of DOX -VEH at each timepoint


```{r delta lfe}

df_1 <- all_genes_states_long %>%
  mutate(log2_FE = log2(Fold_enrichment + 1e-6),
  Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
  separate(group, into = c("trt","time"), remove = FALSE)

delta_df_1 <- df_1 %>%
  group_by(time, Identity, Identity_label,Name) %>%
  summarise(
    delta_log2FE =
      log2_FE[trt == "DOX"] -
      log2_FE[trt == "VEH"],
    .groups = "drop") %>% 
  mutate(time=factor(time, levels=c("24T","24R","144R")))

ggplot(delta_df_1,
       aes(x = Identity_label,
           y = Name,
           fill = delta_log2FE)) +

  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "#053061",
    mid = "white",
    high = "#67001F",
    midpoint = 0,
    name = "log2(DOX / VEH)"
  ) +

  facet_wrap(~time, nrow = 1) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Delta of log2(DOX/VEH) at each timepoint")

```


#### Enhancer plot  

```{r enchacers plot, fig.height=8}
long_file %>% 
  dplyr::filter(stringr::str_detect(Identity, "SCREEN")) %>% 
   group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min),
    Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
     ggplot(.,aes(x=Identity,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=1, ncol=6)+
    scale_y_discrete(limits=rev)+
  ggtitle("All Enhancer types enrichment across states")+
  coord_fixed()
```
### Cormotif set enrichment across states  

Each Histone will have slices of their own states.  Here are all sets next to all states


```{r fig.width=12, fig.height=8}
order <- c("Genome %",
"all_H3K27ac_H3K27ac_ROI.bed",
"Set_1_H3K27ac_ROI.bed" ,
"Set_2_H3K27ac_ROI.bed",
"Set_3_H3K27ac_ROI.bed",
"all_H3K36me3_regions_H3K36me3_ROI.bed",
"Set_1_H3K36me3_ROI.bed",
"Set_2_H3K36me3_ROI.bed",
"all_H3K27me3_regions_H3K27me3_ROI.bed",
"Set_1_H3K27me3_ROI.bed",               
"Set_2_H3K27me3_ROI.bed" ,              
"all_H3K9me3_regions_H3K9me3_ROI.bed",  
"Set_1_H3K9me3_ROI.bed",               
"Set_2_H3K9me3_ROI.bed", 
"Set_3_H3K9me3_ROI.bed" ,               
"ZFP_proteins.bed")                  

all_ZFPs_states %>%
  mutate(
    `State (Emission order)` = as.character(`State (Emission order)`),
    `State (Emission order)` = case_when(
      `State (Emission order)` %in% as.character(1:16) ~ paste0("E", `State (Emission order)`),
      TRUE ~ `State (Emission order)`)) %>% 
  mutate(`State (Emission order)` = factor(`State (Emission order)`,
                                           levels = c(paste0("E",1:16), "Base"))) %>%
  filter(`State (Emission order)` != "Base") %>% 
   left_join(.,States_anno, by = c(`State (Emission order)`="State")) %>% 
  mutate(Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
 
  pivot_longer(., -c(`State (Emission order)`,group, recode, Name), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) %>% 
  mutate(Identity=factor(Identity, levels=order),
         ) %>% 
  dplyr::filter(Identity!="ZFP_proteins.bed") %>% 
  group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>% 
    ggplot(.,aes(x=Identity,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=2, ncol=3)+
    scale_y_discrete(limits=rev)+
  ggtitle("Cormotif across states")+
  coord_fixed()

  
  
```

```{r repeatplot, fig.width=10}
long_file %>% 
  dplyr::filter(stringr::str_detect(Identity, "_rptmasker")) %>% 
  mutate(rpt=case_when(Identity=="LINE_rptmasker"~"LINE",
                       Identity=="SINE_rptmasker"~"SINE",
                       Identity=="LTR_rptmasker"~"LTR",
                       Identity=="DNA_rptmasker"~"DNA",
                       Identity=="Retroposon_rptmasker"~"SVA",
                       TRUE~"Other")) %>% 
  dplyr::filter(rpt != "Other") %>% 
  group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min),
    Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>%  
    ggplot(.,aes(x=rpt,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=2, ncol=3)+
    scale_y_discrete(limits=rev)+
  ggtitle("Transposable elements across states")+
  coord_fixed()
```
#### Looking at Delta TEs  

Now using the delta of DOX-VEH (the log2 FE of each) and plotting the states

```{r fig.height=10}
delta_TE_df <- long_file %>% 
  dplyr::filter(stringr::str_detect(Identity, "_rptmasker")) %>% 
  mutate(rpt=case_when(Identity=="LINE_rptmasker"~"LINE",
                       Identity=="SINE_rptmasker"~"SINE",
                       Identity=="LTR_rptmasker"~"LTR",
                       Identity=="DNA_rptmasker"~"DNA",
                       Identity=="Retroposon_rptmasker"~"SVA",
                       TRUE~"Other"),
         Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>%  
  mutate(log2_FE = log2(Fold_enrichment + 1e-6)) %>% 
  separate(group, into = c("trt","time"), remove = FALSE) %>% 
  dplyr::filter(rpt != "Other") %>% 
  group_by(time,  rpt,Name) %>%
  summarise(
    delta_log2FE =
      log2_FE[trt == "DOX"] -
      log2_FE[trt == "VEH"],
    .groups = "drop") %>% 
  mutate(time=factor(time, levels=c("24T","24R","144R")))



ggplot(delta_TE_df ,
       aes(x = time,
           y = Name,
           fill = delta_log2FE)) +
  geom_tile(color = "grey20", linewidth = 0.2) +
  
  scale_fill_gradient2(
    low = "#B2182B",   # Decrease
    mid = "white",     # No change
    high = "#2166AC",  # Increase
    midpoint = 0,
    limits = c(-4.5, 4.5),
    name = "Δ log2(FE)\n(DOX - VEH)"
  ) +
  
  labs(
    x = "Timepoint",
    y = "ChromHMM State",
    title = "TE Enrichment Changes After DNA Damage"
  ) +
  
  facet_wrap(~rpt, ncol = 5) +  # one facet per TE type
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5)
  ) +
  coord_fixed(ratio = 1.2)



```


```{r scaled TE}
TE_delta_df <- delta_TE_df %>%
  # filter(rpt %in% c("LTR","LINE","SINE","DNA","SVA")) %>%
  filter(rpt %in% c("LTR")) %>% 
  mutate(
    # `State (Emission order)` = forcats::fct_rev(`State (Emission order)`),
    time = factor(time, levels = c("24T","24R","144R"))
  ) %>%
  group_by(rpt) %>%
  mutate(
    # scale relative to max abs value per TE
    delta_log2FE_scaled = delta_log2FE / max(abs(delta_log2FE), na.rm = TRUE),
    # ensure no value goes outside -1..1
    delta_log2FE_scaled = pmax(pmin(delta_log2FE_scaled, .5), -.5)
  ) %>%
  ungroup()
# Scale across the entire dataset
# max_abs_delta <- max(abs(TE_delta_df$delta_log2FE_scaled), na.rm = TRUE)

# TE_delta_df <- TE_delta_df %>%
#   mutate(delta_log2FE_scaled = delta_log2FE / max_abs_delta)

ggplot(TE_delta_df,
       aes(x = time,
           y = Name,
           fill = delta_log2FE)) +
  geom_tile(color = "grey20", linewidth = 0.2) +
  
  scale_fill_gradient2(
  low = "#2166AC",
  mid = "white",
  high = "#B2182B",
  midpoint = 0,
  limits = c(-.5, .5),
  oob = scales::squish,
  name = "Δ log2(FE)\n"
)+
  
  facet_wrap(~rpt, ncol = 5) +  # one TE type per facet
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5)
  ) +
  labs(
    x = "Timepoint",
    y = "ChromHMM State",
    title = "Scaled Δ log2(FE) per TE type after DNA damage"
  ) +
   scale_y_discrete(limits=rev)+
  coord_fixed(ratio = 1.2)
```





### ATAC and cardiac enhancer enrichment

```{r cardiac enchacers plot, fig.height=8}

ATAC_long_file <- all_ATAC_states %>% 
  filter(`State (Emission order)` != "Base") %>% 
  pivot_longer(., -c(`State (Emission order)`,group, recode, Name), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R")),
          `State (Emission order)`=factor(`State (Emission order)`, levels = paste0("E",1:16)))


ATAC_long_file %>% 
  # dplyr::filter(stringr::str_detect(Identity, "SCREEN")) %>% 
   group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>% 
    ggplot(.,aes(x=Identity,y=`State (Emission order)`, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=1, ncol=6)+
    scale_y_discrete(limits=rev)+
  ggtitle("CAR/DAR and cardiac Enhancer types enrichment across states")+
  coord_fixed()
```

Looking at SINE elements that overlap H3K27ac ROI elements

### SOI data

Now to pull in the segmentation data for just the SOI (states of interest).

```{r  SOI 1, eval=FALSE}
##getting all segmentation file locations
seg_files <- list.files(
  "C:/Users/renee/Other_projects_data/DXR_data/TACC_models/Chrom_model_16states_final/",
  pattern = "segments.bed$",
  full.names = TRUE)
### coversion to grages list
chromhmm_gr <- map_df(seg_files, function(f) {
  cond <- basename(f) |> str_remove("_segments.bed")
  import(f) |> 
    as.data.frame() |>
    mutate(condition = cond)
})
# saveRDS(chromhmm_gr,"data/RDS_files/chromhmm_granges_segmentation_files.RDS")
autosomes <- paste0("chr", 1:22)
```

```{r SOI 2, eval=FALSE}
## filtering for the states I want and removing X and y chromosome
chromhmm_sub <- chromhmm_gr %>%
  filter(name %in% SOIs) %>% 
  filter(seqnames %in% autosomes)

repeatmasker <- read_delim("data/Other_paper_data/repeatmasker_20250911.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)


repeatmasker_clean <- repeatmasker %>% mutate(
    strand = ifelse(strand == "C", "-", "+")
  ) %>% 
   mutate(
    start = genoStart + 1,
    end   = genoEnd)%>% 
  mutate(repFamily= str_remove(repFamily, "\\?$")) %>%
  # mutate(repClass= str_remove(repClass, "\\?$")) %>%
  dplyr::filter(genoName %in% autosomes)
  

rpt_split <- split(repeatmasker_clean, repeatmasker_clean$repClass)

rpt_split_gr_list <- lapply(rpt_split, function(df) {
  GRanges(
    seqnames = df$genoName,
    ranges   = IRanges(start = df$start, end = df$end),
    strand   = df$strand,
    repName   = df$repName,
    repClass  = df$repClass,
    repFamily = df$repFamily,
    swScore   = df$swScore,
    milliDiv  = df$milliDiv,
    id        = df$id
  )
})

SINE_gr <- rpt_split_gr_list$SINE

LINE_gr <- rpt_split_gr_list$LINE

LTR_gr <- rpt_split_gr_list$LTR

SVA_gr <- rpt_split_gr_list$Retroposon

DNA_gr <- rpt_split_gr_list$DNA

subsetchromhmm_gr <- GRanges(
  seqnames = chromhmm_sub$seqnames,
  ranges   = IRanges(chromhmm_sub$start, chromhmm_sub$end),
  state    = chromhmm_sub$name,
  condition = chromhmm_sub$condition)


```


### Questions and overlapping section  

```{r loading ROIs, eval=FALSE}

H3K27ac_anno_ROIs <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

H3K27ac_anno_ROIs_gr <- do.call(c,
  lapply(names(H3K27ac_anno_ROIs), function(group_name) {

    cs <- H3K27ac_anno_ROIs[[group_name]]
    gr <- cs@anno

    # add metadata
    mcols(gr)$set  <- group_name
    mcols(gr)$name <- mcols(gr)$Peakid

    gr}))
seqlevelsStyle(H3K27ac_anno_ROIs_gr)
seqlevelsStyle(subsetchromhmm_gr)

##confirm both use UCSC
```
```{r overlap sets 1, eval=FALSE}
H3K27ac_hits <- findOverlaps(H3K27ac_anno_ROIs_gr, subsetchromhmm_gr)

H3K27ac_roi_state_annotated <- H3K27ac_anno_ROIs_gr[queryHits(H3K27ac_hits)]

mcols(H3K27ac_roi_state_annotated) <- cbind(
  mcols(H3K27ac_roi_state_annotated),
  mcols(subsetchromhmm_gr[subjectHits(H3K27ac_hits)]))
table(mcols(H3K27ac_roi_state_annotated)$state)
table(mcols(H3K27ac_roi_state_annotated)$set, mcols(H3K27ac_roi_state_annotated)$state)


H3K27ac_state_summary <-H3K27ac_roi_state_annotated %>%
  as.data.frame() %>%
  group_by(set, condition, state) %>%
  summarise(
    n_peaks = n(),
    total_bp = sum(width),
    .groups = "drop") %>%
  group_by(set, condition) %>%
  mutate(
    frac_peaks = n_peaks / sum(n_peaks),
    frac_bp    = total_bp / sum(total_bp)
  ) %>%
  ungroup()
```

```{r Looking at overlap data for my own purpose, eval=FALSE}

### Need to know that an ROI can overlap more than 1 state, say state 11 twice in a set of  ROIs.

H3K27ac_roi_state_annotated %>% 
  as.data.frame() %>% 
  # distinct(Peakid) ##146222
  dplyr::filter(set %in% c("Set_1","Set_2","Set_3")) %>% 
  # distinct(Peakid) ##114832
 dplyr::filter(state=="E10") %>% 
  # distinct(Peakid)##28,079
  group_by(set, condition, Peakid) %>% tally() %>% group_by(n) %>% tally()
### the above show grouping can lead to bp overcounting

H3K27ac_E10_roi_gr <- H3K27ac_roi_state_annotated[ mcols(H3K27ac_roi_state_annotated)$state=="E10" &
                                      mcols(H3K27ac_roi_state_annotated)$set %in% c("Set_1","Set_2","Set_3") ]

# reduce ROIs to merge any duplicates
H3K27ac_E10_roi_gr_unique <- GenomicRanges::reduce(H3K27ac_E10_roi_gr)  # merges overlapping ranges within each ROI

# compute bp per set
as.data.frame(H3K27ac_E10_roi_gr_unique)# %>%
  group_by(set, condition) %>%
  summarise(total_bp = sum(width))
  
  ### this just give the exact bp coverage of E10 for bp of ROI in each set and condition.

```
Now I am just going to only pull out unique Peakids by Set and then overlap with repeatmasker data

```{r Looking just at E10, eval=FALSE}
H3K27ac_E10_roi_df <- H3K27ac_E10_roi_gr %>% 
  as.data.frame() %>% 
 group_by(Peakid, set, condition, seqnames) %>%
  summarise(
    start = min(start),
    end = max(end),
    .groups = "drop")
 
H3K27ac_E10_roi_reduced_gr <- GRanges(
  seqnames = H3K27ac_E10_roi_df$seqnames,
  ranges = IRanges(start = H3K27ac_E10_roi_df$start, end = H3K27ac_E10_roi_df$end),
  Peakid = H3K27ac_E10_roi_df$Peakid,
  set = H3K27ac_E10_roi_df$set,
  condition = H3K27ac_E10_roi_df$condition)

te_list <- c("LINE","SINE","LTR","DNA","Retroposon")

H3K27ac_E10_roi_te_hits <- lapply(te_list, function(te_class) {
  te_gr <- rpt_split_gr_list[[te_class]]
  
  hits <- findOverlaps(H3K27ac_E10_roi_reduced_gr, te_gr)
  if (length(hits) == 0) return(NULL)

  roi_hits <- H3K27ac_E10_roi_reduced_gr[queryHits(hits)]

  mcols(roi_hits)$TE_class <- te_class
  mcols(roi_hits)$repClass <- mcols(te_gr)$repClass[subjectHits(hits)]
  mcols(roi_hits)$repFamily <- mcols(te_gr)$repFamily[subjectHits(hits)]
  mcols(roi_hits)$repName <- mcols(te_gr)$repName[subjectHits(hits)]

  roi_hits
})

names(H3K27ac_E10_roi_te_hits) <- te_list
```
```{r pulling permtest values function, eval=FALSE}
extract_enrichment <- function(perm_obj, condition_name = NULL) {
  # Extract observed overlaps
  obs <- perm_obj$numOverlaps$observed

  # Extract expected overlaps (mean of permuted)
  exp <- mean(perm_obj$numOverlaps$permuted)
  
  # Enrichment ratio
  enrich <- obs / exp
  
  # Z-score
  zscore <- perm_obj$numOverlaps$zscore
  
  # Return as a data frame
  df <- data.frame(
    Condition = condition_name,
    Observed = obs,
    Expected = exp,
    Enrichment = enrich,
    Zscore = zscore
  )
  
  return(df)
}
```


```{r, eval=FALSE}


genome_gr <- getGenomeAndMask("hg38")$genome
genome_autosomes <- genome_gr[seqnames(genome_gr) %in% autosomes]

test_set_E10_24T <- H3K27ac_E10_roi_te_hits$LINE[ mcols(H3K27ac_E10_roi_te_hits$LINE)$condition=="24T_DOX_16" &
                                      mcols(H3K27ac_E10_roi_te_hits$LINE)$set == "Set_2" ]

test_set_E10_24R <- H3K27ac_E10_roi_te_hits$LINE[ mcols(H3K27ac_E10_roi_te_hits$LINE)$condition=="24R_DOX_16" &
                                      mcols(H3K27ac_E10_roi_te_hits$LINE)$set == "Set_2" ]

test_set_E10_144R <- H3K27ac_E10_roi_te_hits$LINE[ mcols(H3K27ac_E10_roi_te_hits$LINE)$condition=="144R_DOX_16" &
                                      mcols(H3K27ac_E10_roi_te_hits$LINE)$set == "Set_2" ]




# 
perm_test_24T_E10_s2_H3K27ac <- permTest(A= test_set_E10_24T,
                          B= rpt_split_gr_list$LINE,
                          ntimes=1000,
                          randomize.function=randomizeRegions,
                          evaluate.function = numOverlaps,
                          genome=genome_autosomes,
                          count.once= TRUE,
                          verbose = TRUE)

perm_test_24R_E10_s2_H3K27ac <- permTest(A= test_set_E10_24R,
                          B= rpt_split_gr_list$LINE,
                          ntimes=1000,
                          randomize.function=randomizeRegions,
                          evaluate.function = numOverlaps,
                          genome=genome_autosomes,
                          count.once= TRUE,
                          verbose = TRUE)

perm_test_144R_E10_s2_H3K27ac <- permTest(A= test_set_E10_144R,
                          B= rpt_split_gr_list$LINE,
                          ntimes=1000,
                          randomize.function=randomizeRegions,
                          evaluate.function = numOverlaps,
                          genome=genome_autosomes,
                          count.once= TRUE,
                          verbose = TRUE)
# 
# A <- extract_enrichment(perm_test_24T_E10_s2_H3K27ac,"S2_E10_24T" )
# B <- extract_enrichment(perm_test_24R_E10_s2_H3K27ac,"S2_E10_24R" )
# C <- extract_enrichment(perm_test_144R_E10_s2_H3K27ac,"S2_E10_144R" )
# test <- bind_rows(A,B,C)
plot(perm_test_24R_E10_s2_H3K27ac)
plot(perm_test_24T_E10_s2_H3K27ac)
plot(perm_test_144R_E10_s2_H3K27ac)
# saveRDS(test, "data/RDS_files/permtest_H3K27ac_S2_LINE.RDS")
```


```{r overlap sets 3, eval=FALSE}
H3K27ac_144R_set1_summary <- H3K27ac_state_summary %>%
  filter(set == "Set_1", condition %in% c("144R_VEH_16", "144R_DOX_16"))


ggplot(H3K27ac_144R_set1_summary, aes(x = condition, y = frac_peaks, fill = state)) +
  geom_col(position = "stack") +
  ylab("Fraction of peaks per state") +
  xlab("Condition") +
  scale_fill_brewer(palette = "Set2") +
  theme_classic() +
  ggtitle("State composition in Set1: VEH144R vs DOX144R")

H3K27ac_144R_set1_wide <- H3K27ac_144R_set1_summary %>%
  select(condition, state, frac_peaks) %>%
  tidyr::pivot_wider(names_from = condition, values_from = frac_peaks)

H3K27ac_144R_set1_wide <- H3K27ac_144R_set1_wide %>%
  mutate(frac_ratio = `144R_DOX_16` / `144R_VEH_16`)



######-24R------------------------------------------------------
H3K27ac_24R_set1_summary <- H3K27ac_state_summary %>%
  filter(set == "Set_1", condition %in% c("24R_VEH_16", "24R_DOX_16"))


ggplot(H3K27ac_24R_set1_summary, aes(x = condition, y = frac_peaks, fill = state)) +
  geom_col(position = "stack") +
  ylab("Fraction of peaks per state") +
  xlab("Condition") +
  scale_fill_brewer(palette = "Set2") +
  theme_classic() +
  ggtitle("State composition in Set1: VEH24R vs DOX24R")

H3K27ac_24R_set1_wide <- H3K27ac_24R_set1_summary %>%
  select(condition, state, frac_peaks) %>%
  tidyr::pivot_wider(names_from = condition, values_from = frac_peaks)

H3K27ac_24R_set1_wide <- H3K27ac_24R_set1_wide %>%
  mutate(frac_ratio = `24R_DOX_16` / `24R_VEH_16`)


######-24TR------------------------------------------------------
H3K27ac_24T_set1_summary <- H3K27ac_state_summary %>%
  filter(set == "Set_1", condition %in% c("24T_VEH_16", "24T_DOX_16"))


ggplot(H3K27ac_24T_set1_summary, aes(x = condition, y = frac_peaks, fill = state)) +
  geom_col(position = "stack") +
  ylab("Fraction of peaks per state") +
  xlab("Condition") +
  scale_fill_brewer(palette = "Set2") +
  theme_classic() +
  ggtitle("State composition in Set1: VEH24T vs DOX24T")

H3K27ac_24T_set1_wide <- H3K27ac_24T_set1_summary %>%
  select(condition, state, frac_peaks) %>%
  tidyr::pivot_wider(names_from = condition, values_from = frac_peaks)

H3K27ac_24T_set1_wide <- H3K27ac_24T_set1_wide %>%
  mutate(frac_ratio = `24T_DOX_16` / `24T_VEH_16`)
```

Graphing what I have so far.  I grouped the overlapping data by sets, so I have all ROIs, all set_1 ROIs, all set_2 rois all set_3 rois.

Now I want to see the breakdown of these states across condtitions in my ROIs-  note I have already done a Fold enrichment, so this should mimic that.

```{r, eval=FALSE}
ggplot(H3K27ac_state_summary, aes(x = set, y = frac_peaks, fill = state)) +
  geom_col(position = "stack") +
  facet_wrap(~condition) +
  ylab("Fraction of peaks per state") +
  xlab("ROI Set") +
  scale_fill_brewer(palette = "Set2") +
  theme_classic()

table1 <- table(
  H3K27ac_roi_state_annotated$set[H3K27ac_roi_state_annotated$condition == "144R_VEH_16"],
  H3K27ac_roi_state_annotated$state[H3K27ac_roi_state_annotated$condition == "144R_VEH_16"]
)

chisq.test(table1)
```

