---
title: "ChromHMM_1"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 400px;
}
```


```{r}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(readxl)
library(ChIPseeker)
library(regioneR)
library(cowplot)
library(GenomeInfoDb)
library(ggforce)
```

This is the AIC/BIC plots generated to figure out how many states we wanted to use for our modeling of our data:
```{r TACC generated}
aic_bic_table_TACC <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TACC_models/model_selection_summary.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

aic_bic_table_TACC %>% 
   pivot_longer(cols=c(AIC,BIC), names_to = "test_type",values_to = "calculation") %>% 
  ggplot(.,aes(x=States,y=calculation))+
  geom_point(size=4)+
  geom_line()+
  facet_wrap(~test_type)+
  theme_bw()
  
```

The results point to 16 states yielding the best results.

I ran the models for 16 states and here is the outcome of those results:

Here is the emission table of the results:
```{r emission of 16 states}

emission_16state <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TACC_models/Chrom_model_16states_final/emissions_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

emission_16state %>% 
  mutate(`State (Emission order)`=factor(`State (Emission order)`, levels=c(1:16))) %>% 
  pivot_longer(., !`State (Emission order)`, names_to = "Histone",values_to = "emission") %>% 
  mutate(Histone=factor(Histone,levels=c("H3K27ac","H3K36me3","H3K27me3","H3K9me3"))) %>% 
  ggplot(., aes(x=Histone, y=`State (Emission order)`, fill=emission))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(expand = c(0, 0)) +
scale_y_discrete(expand = c(0, 0), limits = rev)+
  coord_fixed()+
    theme_classic()+
  ggtitle("Emission Parameters")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Here is the transition plot of the results:

```{r transition of 16 states}

transition_16state <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TACC_models/Chrom_model_16states_final/transitions_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

transition_16state %>% 
  mutate(`State (from\\to) (Emission order)`=factor(`State (from\\to) (Emission order)`, levels=c(1:16)))%>%
  dplyr::rename(state_from_to=`State (from\\to) (Emission order)`) %>% 
  pivot_longer(., !state_from_to, names_to = "States",values_to = "transition") %>% 
  mutate(States=factor(States, levels=c(1:16))) %>% 
  
  ggplot(., aes(x=States, y=state_from_to, fill=transition))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(expand = c(0, 0)) +
scale_y_discrete(expand = c(0, 0), limits = rev)+
  coord_fixed()+
    theme_classic()+
  ggtitle("Transition Parameters")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### Adding annotation to states

```{r annotation}

States_anno <- data.frame("State"=c(paste0("E",1:16)),
"Name"=c("Heterochromatin_1",
"Quiescent_Low_Coverage",
"Repressed_Heterochromatin",
"Repressed_ZFP_genes",
"Strong_Polycomb_Repressed",
 "Bivalent_Enhancer",
"Heterochromatin_2",
"Polycomb_Repressed",
 "Weak_Genic_Enhancer",
 "Bivalent_Poised_TSS",
 "Active_Enhancer",
  "Strong_Genic_Enhancer",
 "Active_TSS",
 "Very_Weak_Transcription",
 "Genic_Strong_Transcription",
 "Genic_Weak_Transcription"), "recode"=c("N","P","M","L","J","I","O","K","G","H","E","F","A","D","B","C")) %>% 
  mutate(recode=factor(recode,levels=c("A","B","C","D","E","F","G","H","I","J","K","L","M","N","O","P")))

# SOIs <- c("E4","E10","E11","E12","E13")

States_anno

```
#### reordering of the states emission:

```{r state reorder}

emission_16state %>% 
  
  mutate(`State (Emission order)`=paste0("E",`State (Emission order)`)) %>% 
left_join(.,States_anno, by = c(`State (Emission order)`="State")) %>%
   mutate(Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
  
  pivot_longer(.,  !c(`State (Emission order)`,Name,recode), names_to = "Histone",values_to = "emission") %>% 
  mutate(Histone=factor(Histone,levels=c("H3K27ac","H3K36me3","H3K27me3","H3K9me3"))) %>% 
  ggplot(., aes(x=Histone, y=Name, fill=emission))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(expand = c(0, 0)) +
scale_y_discrete(expand = c(0, 0), limits = rev)+
  coord_fixed()+
    theme_classic()+
  ggtitle("Emission Parameters")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
#### TSS neighborhood
```{r import TSS neighborhood data}
DOX_24T_TSS <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24T_DOX_16_RefSeqTSS_neighborhood.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_TSS <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24R_DOX_16_RefSeqTSS_neighborhood.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_TSS <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/144R_DOX_16_RefSeqTSS_neighborhood.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_TSS <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24T_VEH_16_RefSeqTSS_neighborhood.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_TSS <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/24R_VEH_16_RefSeqTSS_neighborhood.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_TSS <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/Chrom_model_rerun/144R_VEH_16_RefSeqTSS_neighborhood.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 

all_trt_TSS_states <- 
  VEH_144R_TSS %>% 
  bind_rows(DOX_144R_TSS) %>% 
   bind_rows(DOX_24R_TSS) %>% 
   bind_rows(VEH_24R_TSS) %>% 
   bind_rows(DOX_24T_TSS) %>% 
   bind_rows(VEH_24T_TSS) 

all_trt_TSS_states_long <- all_trt_TSS_states %>%
  pivot_longer(
    cols = -c(`State (Emission order)`, group),
    names_to = "position",
    values_to = "value"
  ) %>%
  mutate(
    position = as.numeric(position),
    state = paste0("E",`State (Emission order)`)
  ) %>% 
  left_join(States_anno,
            by = c("state" = "State")) %>% 
   mutate(Name = factor(Name,
                       levels = States_anno %>%
                         arrange(recode) %>%
                         pull(Name)))
 

```
plotting all TSS neighborhoods
```{r}
ggplot(all_trt_TSS_states_long,
       aes(position, Name, fill = value)) +
  geom_tile() +
  facet_wrap(~ group) +
  # scale_x_discrete(expand = c(0, 0)) +
scale_y_discrete(expand = c(0, 0), limits = rev)+
  # coord_fixed()+
    theme_classic()+
  scale_fill_gradient(low = "white", high = "#08306B")
```

“What is the general TSS architecture of each chromatin state, independent of condition?”
```{r}
TSS_geom_df <-  all_trt_TSS_states_long %>%
  group_by( Name, position) %>%
  summarize(
    geom_FE = exp(mean(log(value + 1e-6), na.rm = TRUE)),
    .groups = "drop")
  
  
  
  
# TSS_geom_df <-TSS_geom_df %>%
#   group_by(group) %>%
#   mutate(
#     FE_min = min(geom_FE, na.rm = TRUE),
#     FE_max = max(geom_FE, na.rm = TRUE),
#     chromhmm_scaled = (geom_FE - FE_min) / (FE_max - FE_min)
#   ) %>%
#   ungroup()

ggplot(TSS_geom_df, aes(position, Name, fill = geom_FE)) +
  geom_tile() +
  # facet_wrap(~ group) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_fill_gradient(
    low = "white",
    high = "#08306B"
  ) +
  scale_y_discrete(limits = rev) +
  theme_classic()+
  ggtitle("General TSS architecture")
```



### Enrichment across the states:
looking at enrichment of Features across the states by each group.  This helps with the classification and interpretation of what each state represents.

```{r loading other features enrichment}
DOX_24T_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_full <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/rerun_regions/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 
```

```{r loading other features enrichment2}
  DOX_24T_rpt <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/RPTmasker_run/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_rpt <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/RPTmasker_run/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_rpt <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/RPTmasker_run/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_rpt <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/RPTmasker_run/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_rpt <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/RPTmasker_run/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_rpt <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/RPTmasker_run/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 
all_rptclean_states <- 
  VEH_144R_rpt %>% 
  bind_rows(DOX_144R_rpt) %>% 
   bind_rows(DOX_24R_rpt) %>% 
   bind_rows(VEH_24R_rpt) %>% 
   bind_rows(DOX_24T_rpt) %>% 
   bind_rows(VEH_24T_rpt) 
 
 just_rpt <- all_rptclean_states[,c("State (Emission order)","repeatmasker_clean.bed","group")]
just_rpt_long <- just_rpt %>% 
   dplyr::filter(`State (Emission order)` !="Base") %>% 
   left_join(States_anno, by=c(`State (Emission order)`="State")) %>% 
   pivot_longer(., cols=!c(`State (Emission order)`,group,Name,recode), names_to = "Identity",values_to = "Fold_enrichment")  
```

```{r loading gene enrichment}
DOX_24T_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_gene <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/16-internal-enrichment/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 

 all_genes_states <- 
  VEH_144R_gene %>% 
  bind_rows(DOX_144R_gene) %>% 
   bind_rows(DOX_24R_gene) %>% 
   bind_rows(VEH_24R_gene) %>% 
   bind_rows(DOX_24T_gene) %>% 
   bind_rows(VEH_24T_gene) 
 
```

```{r loading ZFP_run data}
DOX_24T_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_ZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ZFP_run/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 all_ZFPs_states <- 
  VEH_144R_ZFP %>% 
  bind_rows(DOX_144R_ZFP) %>% 
   bind_rows(DOX_24R_ZFP) %>% 
   bind_rows(VEH_24R_ZFP) %>% 
   bind_rows(DOX_24T_ZFP) %>% 
   bind_rows(VEH_24T_ZFP) 
 
 just_ZFP <- all_ZFPs_states[,c("State (Emission order)","ZFP_proteins.bed","group")]
 just_ZFP %>% 
   dplyr::filter(`State (Emission order)` !="Base")

```


```{r loading KZFP_run data}
DOX_24T_KZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/KZFP_run/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_KZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/KZFP_run/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_KZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/KZFP_run/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_KZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/KZFP_run/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_KZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/KZFP_run/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_KZFP <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/KZFP_run/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 all_KZFPs_states <- 
  VEH_144R_KZFP %>% 
  bind_rows(DOX_144R_KZFP) %>% 
   bind_rows(DOX_24R_KZFP) %>% 
   bind_rows(VEH_24R_KZFP) %>% 
   bind_rows(DOX_24T_KZFP) %>% 
   bind_rows(VEH_24T_KZFP) 
 
 just_KZFP <- all_KZFPs_states[,c("State (Emission order)","KZFP.bed","group")]
 
```
```{r loading ATAC_run}

DOX_24T_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_ATAC <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/ATAC_run/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 
 all_ATAC_states <- 
  VEH_144R_ATAC %>% 
  bind_rows(DOX_144R_ATAC) %>% 
   bind_rows(DOX_24R_ATAC) %>% 
   bind_rows(VEH_24R_ATAC) %>% 
   bind_rows(DOX_24T_ATAC) %>% 
   bind_rows(VEH_24T_ATAC) %>% 
   left_join(.,States_anno, by = c(`State (Emission order)`="State"))
 
```

```{r loading LTR features enrichment}
DOX_24T_LTR <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTRfamily_run/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_LTR <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTRfamily_run/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_LTR <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTRfamily_run/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_LTR <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTRfamily_run/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_LTR <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTRfamily_run/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_LTR <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTRfamily_run/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 

all_LTR_states <- 
  VEH_144R_LTR %>% 
  bind_rows(DOX_144R_LTR) %>% 
   bind_rows(DOX_24R_LTR) %>% 
   bind_rows(VEH_24R_LTR) %>% 
   bind_rows(DOX_24T_LTR) %>% 
   bind_rows(VEH_24T_LTR) %>% 
   left_join(.,States_anno, by = c(`State (Emission order)`="State"))
 
```

```{r loading LTRERV1 features enrichment}
DOX_24T_LTR_ERV1 <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTR_ERV1_family_run/24t_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24T")

DOX_24R_LTR_ERV1 <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTR_ERV1_family_run/24R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="DOX_24R")

DOX_144R_LTR_ERV1 <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTR_ERV1_family_run/144R_DOX_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>% mutate(group="DOX_144R")


VEH_24T_LTR_ERV1 <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTR_ERV1_family_run/24t_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24T")

VEH_24R_LTR_ERV1 <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTR_ERV1_family_run/24R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_24R")

VEH_144R_LTR_ERV1 <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/overlap_enrichment_all/LTR_ERV1_family_run/144R_VEH_16.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)%>% mutate(group="VEH_144R") 

all_LTR_ERV1_states <- 
  VEH_144R_LTR_ERV1 %>% 
  bind_rows(DOX_144R_LTR_ERV1) %>% 
   bind_rows(DOX_24R_LTR_ERV1) %>% 
   bind_rows(VEH_24R_LTR_ERV1) %>% 
   bind_rows(DOX_24T_LTR_ERV1) %>% 
   bind_rows(VEH_24T_LTR_ERV1) %>% 
   left_join(.,States_anno, by = c(`State (Emission order)`="State"))
 
```

```{r making the large plotting dataframe}
all_files_states <- 
  VEH_144R_full %>% 
  bind_rows(DOX_144R_full) %>% 
   bind_rows(DOX_24R_full) %>% 
   bind_rows(VEH_24R_full) %>% 
   bind_rows(DOX_24T_full) %>% 
   bind_rows(VEH_24T_full) %>% 
  left_join(.,States_anno, by = c(`State (Emission order)`="State"))

long_file <- all_files_states %>% 
  dplyr::select(`State (Emission order)`,
                `Genome %`,
                `GRCh38-cCREs.bed`,
                `SCREEN_hg38_CA-CTCF.bed`:SCREEN_hg38_pELS.bed,
                Set_1_H3K27ac_ROI.bed:Set_3_H3K27ac_ROI.bed,
                all_H3K27ac_H3K27ac_ROI.bed,
                LINE_rptmasker.bed,
                SINE_rptmasker.bed,
                LTR_rptmasker.bed,
                DNA_rptmasker.bed,
                Retroposon_rptmasker.bed,
                RC_rptmasker.bed,
                Low_complexity_rptmasker.bed,
                RNA_rptmasker.bed,
                Satellite_rptmasker.bed,
                Simple_repeat_rptmasker.bed,
                Unknown_rptmasker.bed, 
                rRNA_rptmasker.bed:group,Name,recode) %>% 
  mutate(`State (Emission order)`=factor(`State (Emission order)`, levels = c(paste0("E", 1:16),"Base"))) %>%
  dplyr::filter(`State (Emission order)` != "Base") %>% 
  pivot_longer(., cols=!c(`State (Emission order)`,group,Name,recode), names_to = "Identity",values_to = "Fold_enrichment") %>% 
  mutate(Identity=str_remove(Identity,".bed")) %>% 
  mutate(Identity=factor(Identity, 
                         levels=c("Genome %",
                                  "GRCh38-cCREs",
                                  "SCREEN_hg38_CA-CTCF",
                                  "SCREEN_hg38_CA-H3K4me3",
                                  "SCREEN_hg38_CA-TF",
                                  "SCREEN_hg38_CA",
                                  "SCREEN_hg38_PLS",
                                  "SCREEN_hg38_TF",
                                  "SCREEN_hg38_pELS",
                                  "SCREEN_hg38_dELS",
                                  "all_H3K27ac_H3K27ac_ROI",
                                  "Set_1_H3K27ac_ROI",
                                  "Set_2_H3K27ac_ROI",
                                  "Set_3_H3K27ac_ROI",
                                  "LINE_rptmasker",
                                  "SINE_rptmasker",
                                  "LTR_rptmasker",
                                  "DNA_rptmasker",
                                  "Retroposon_rptmasker",
                                  "RC_rptmasker",
                                  "Low_complexity_rptmasker",
                                  "RNA_rptmasker",
                                  "Satellite_rptmasker",
                                  "Simple_repeat_rptmasker",
                                  "Unknown_rptmasker", 
                                  "rRNA_rptmasker",
                                  "scRNA_rptmasker",
                                  "snRNA_rptmasker",
                                  "srpRNA_rptmasker",
                                  "tRNA_rptmasker"))) %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) 

```

```{r plot dataframe for genes}

# columns to pivot (exclude ID columns)
id_cols <- c("State (Emission order)", "group","Name","recode")

all_genes_states_long <-
all_genes_states %>%
  mutate(`State (Emission order)`=paste0("E",`State (Emission order)`)) %>% 
  left_join(just_ZFP) %>%
  left_join(just_KZFP) %>% 
  left_join(.,States_anno, by = c(`State (Emission order)`="State")) %>% 
  
  mutate(`State (Emission order)` = factor(`State (Emission order)`,
                                           levels = c(paste0("E",1:16),"EBase"))) %>%
  filter(`State (Emission order)` != "EBase") %>%
  
  pivot_longer(
    cols = -all_of(id_cols),
    names_to = "Identity",
    values_to = "Fold_enrichment"
  ) %>%
  # preserve original order
  mutate(Identity = factor(Identity, levels =c( names(all_genes_states)[!names(all_genes_states) %in% id_cols],"ZFP_proteins.bed","KZFP.bed"))) %>% 
         # create cleaned label for plotting
         mutate(Identity_label = case_when(
      str_detect(Identity, "\\.hg38\\.bed\\.gz$") ~ str_remove(Identity, "\\.hg38\\.bed\\.gz$"),
      str_detect(Identity, "\\.bed$") ~ str_remove(Identity, "\\.bed$"),
      TRUE ~ Identity   # keep everything else unchanged
    ),
    group = factor(group,
                        levels = c("VEH_24T","VEH_24R","VEH_144R",
                                   "DOX_24T","DOX_24R","DOX_144R"))) 

  
final_order <- unique(all_genes_states_long$Identity_label)

all_genes_states_long <- all_genes_states_long %>% 
  mutate(Identity_label=factor(Identity_label,levels=final_order))

```

### Looking a genomic features:

```{r plot all genes states, fig.width=12, fig.height=6}
all_genes_states_long %>% 
  group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>%
  arrange(recode) %>%
  mutate(
    Name = factor(Name,
                  levels = unique(Name[order(recode)])) ) %>% 
 
    # ggplot(.,aes(x=Identity_label,y=`State (Emission order)`, fill=chromhmm_scaled))+
  ggplot(.,aes(x=Identity_label,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=1, ncol=6)+
    scale_y_discrete(limits=rev)+
  ggtitle("Genic region enrichment across states")+
  coord_fixed()
```



The genic enrichment across states plots shows the exact enrichment across each condition, however, for summary purposes, I wish to know the general enrichment from all conditions.  I will take the geometric mean  across  DOX and VEH conditions (and all conditions) at each timepoint for each state and identity ( in this case for the CpGIslands, Exons, genes, TESs and TSS etc..).  This treats the fold-change enrichment and depletion symmetrically about 1 because enrichment behaves multiplicatively since it is already a ratio.  I am only using this to determine which states are constitutively TSS-enriched, with are CPG-biased, which are ZFP enriched etc...
```{r geometric mean, all conditions}
geom_mean_df <- all_genes_states_long %>% 
  group_by(Identity, Identity_label, `State (Emission order)`) %>% 
  summarize(geom_FE = exp(mean(log(Fold_enrichment + 1e-6), na.rm = TRUE)),
    .groups = "drop") %>% 
  group_by(Identity) %>% 
             mutate(
    FE_min = min(geom_FE, na.rm = TRUE),
    FE_max = max(geom_FE, na.rm = TRUE),
    chromhmm_scaled = (geom_FE - FE_min) / (FE_max - FE_min)) 
  
  ggplot(geom_mean_df,aes(x=Identity_label,y=`State (Emission order)`, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  
    scale_y_discrete(limits=rev)+
  ggtitle("Genic region relative enrichment pattern across conditions\n (geometric mean)")+
  coord_fixed()

```
```{r geom_mean unscaled log2}
geom_mean_df <- geom_mean_df %>%  mutate(log2_FE = log2(geom_FE))

ggplot(geom_mean_df,
       aes(x = Identity_label,
           y = `State (Emission order)`,
           fill = log2_FE)) +

  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "darkgreen",
    mid = "white",
    high = "#08306B",
    midpoint = 0,
    name = "log2(Fold Enrichment)"
  ) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +

  ggtitle("Geometric Mean ChromHMM Enrichment Across Conditions")+
  coord_fixed()
```


#### Poster future plots

```{r poster plot data frame}
full_states_data_long <- bind_rows(all_genes_states_long,long_file) %>% 
  bind_rows(.,just_rpt_long)


Identity_label_order <- c("Genome %", "CpGIsland","RefSeqGene","RefSeqTSS2kb","RefSeqTSS","RefSeqExon","RefSeqTES",   
      "KZFP","PLS","pELS","dELS", "repeatmasker_clean")

geom_df_reorderd <-
full_states_data_long %>% 
  mutate(Identity_label = case_when(
      str_detect(Identity, "\\.hg38\\.bed\\.gz$") ~ str_remove(Identity, "\\.hg38\\.bed\\.gz$"),
      str_detect(Identity, "\\.bed$") ~ str_remove(Identity, "\\.bed$"),
      str_detect(Identity,"^SCREEN_hg38_")~ str_remove(Identity,"^SCREEN_hg38_"),
       str_detect(Identity,"_rptmasker")~ str_remove(Identity,"_rptmasker"),
      Identity =="all_H3K27ac_H3K27ac_ROI"~ "all_H3K27ac_regions",
      TRUE ~ Identity   # keep everything else unchanged
    ),
    group = factor(group,
                        levels = c("VEH_24T","VEH_24R","VEH_144R",
                                   "DOX_24T","DOX_24R","DOX_144R"))) %>% 
  dplyr::filter(Identity_label %in% Identity_label_order) %>% 
  arrange(recode) %>%
  mutate(
    Name = factor(Name,
                  levels = unique(Name[order(recode)])),
    Identity_label=factor(Identity_label,levels=Identity_label_order)) %>% 
   group_by(Identity, Identity_label, Name) %>% 
  summarize(geom_FE = exp(mean(log(Fold_enrichment + 1e-6), na.rm = TRUE)),
    .groups = "drop") %>% 
  group_by(Identity) %>% 
             mutate(
    FE_min = min(geom_FE, na.rm = TRUE),
    FE_max = max(geom_FE, na.rm = TRUE),
    chromhmm_scaled = (geom_FE - FE_min) / (FE_max - FE_min)) 


ggplot(geom_df_reorderd,aes(x=Identity_label,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  
    scale_y_discrete(limits=rev)+
  ggtitle("Genic region relative enrichment pattern across conditions\n (geometric mean)")+
  coord_fixed()

  
```

```{r plot full genes states, fig.width=12, fig.height=6}

full_states_data_long %>% 
  mutate(Identity_label = case_when(
      str_detect(Identity, "\\.hg38\\.bed\\.gz$") ~ str_remove(Identity, "\\.hg38\\.bed\\.gz$"),
      str_detect(Identity, "\\.bed$") ~ str_remove(Identity, "\\.bed$"),
      str_detect(Identity,"^SCREEN_hg38_")~ str_remove(Identity,"^SCREEN_hg38_"),
       str_detect(Identity,"_rptmasker")~ str_remove(Identity,"_rptmasker"),
      Identity =="all_H3K27ac_H3K27ac_ROI"~ "all_H3K27ac_regions",
      TRUE ~ Identity   # keep everything else unchanged
    ),
    group = factor(group,
                        levels = c("VEH_24T","VEH_24R","VEH_144R",
                                   "DOX_24T","DOX_24R","DOX_144R"))) %>% 
  dplyr::filter(Identity_label %in% Identity_label_order) %>% 
  arrange(recode) %>%
  mutate(
    Name = factor(Name,
                  levels = unique(Name[order(recode)])),
    Identity_label=factor(Identity_label,levels=Identity_label_order)) %>% 
   group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>%
 
  ggplot(.,aes(x=Identity_label,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=2, ncol=3)+
    scale_y_discrete(limits=rev)+
  ggtitle("Genic region enrichment across states")+
  coord_fixed()
```

```{r for new poster}

Identity_label_order_new <-c("Genome %", "CpGIsland","RefSeqGene","RefSeqTSS2kb","RefSeqTSS","RefSeqExon","RefSeqTES",   
      "KZFP","PLS","pELS","dELS", "repeatmasker_clean","SINE", "LINE","LTR","DNA","Retroposon")


# c("Genome %", "CpGIsland","RefSeqGene","RefSeqTSS2kb","RefSeqTSS","RefSeqExon","RefSeqTES", "KZFP","PLS","pELS","dELS", "SINE", "LINE","LTR","DNA","Retroposon")

geom_df_reorderd_new <-
full_states_data_long %>% 
  mutate(Identity_label = case_when(
      str_detect(Identity, "\\.hg38\\.bed\\.gz$") ~ str_remove(Identity, "\\.hg38\\.bed\\.gz$"),
      str_detect(Identity, "\\.bed$") ~ str_remove(Identity, "\\.bed$"),
      str_detect(Identity,"^SCREEN_hg38_")~ str_remove(Identity,"^SCREEN_hg38_"),
       str_detect(Identity,"_rptmasker")~ str_remove(Identity,"_rptmasker"),
      Identity =="all_H3K27ac_H3K27ac_ROI"~ "all_H3K27ac_regions",
      TRUE ~ Identity   # keep everything else unchanged
    ),
    group = factor(group,
                        levels = c("VEH_24T","VEH_24R","VEH_144R",
                                   "DOX_24T","DOX_24R","DOX_144R"))) %>% 
  dplyr::filter(Identity_label %in% Identity_label_order_new) %>% 
  arrange(recode) %>%
  mutate(
    Name = factor(Name,
                  levels = unique(Name[order(recode)])),
    Identity_label=factor(Identity_label,levels=Identity_label_order_new)) %>% 
   group_by(Identity, Identity_label, Name) %>% 
  summarize(geom_FE = exp(mean(log(Fold_enrichment + 1e-6), na.rm = TRUE)),
    .groups = "drop") %>% 
  group_by(Identity) %>% 
             mutate(
    FE_min = min(geom_FE, na.rm = TRUE),
    FE_max = max(geom_FE, na.rm = TRUE),
    chromhmm_scaled = (geom_FE - FE_min) / (FE_max - FE_min)) 


ggplot(geom_df_reorderd_new,aes(x=Identity_label,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  
    scale_y_discrete(limits=rev)+
  ggtitle("Genic annotations relative enrichment")+
  coord_fixed()

```


```{r suppdata for poster}
geom_df_reorderd <- geom_df_reorderd %>%  mutate(log2_FE = log2(geom_FE))

ggplot(geom_df_reorderd,
       aes(x = Identity_label,
           y = Name,
           fill = log2_FE)) +

  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "darkgreen",
    mid = "white",
    high = "#08306B",
    midpoint = 0,
    name = "log2(Fold Enrichment)"
  ) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +

  ggtitle("Geometric Mean ChromHMM Enrichment Across Conditions")+
  coord_fixed()
```
```{r}
delta_df_reorderd_new <- 
full_states_data_long %>% 
  mutate(log2_FE = log2(Fold_enrichment + 1e-6)) %>% 
        tidyr::separate(group, into=c("trt","time"), remove = FALSE) %>% 
  mutate(Identity_label = case_when(
      str_detect(Identity, "\\.hg38\\.bed\\.gz$") ~ str_remove(Identity, "\\.hg38\\.bed\\.gz$"),
      str_detect(Identity, "\\.bed$") ~ str_remove(Identity, "\\.bed$"),
      str_detect(Identity,"^SCREEN_hg38_")~ str_remove(Identity,"^SCREEN_hg38_"),
       str_detect(Identity,"_rptmasker")~ str_remove(Identity,"_rptmasker"),
      Identity =="all_H3K27ac_H3K27ac_ROI"~ "all_H3K27ac_regions",
      TRUE ~ Identity   # keep everything else unchanged
    ),
    group = factor(group,
                        levels = c("VEH_24T","VEH_24R","VEH_144R",
                                   "DOX_24T","DOX_24R","DOX_144R"))) %>%
  dplyr::filter(Identity_label %in% Identity_label_order_new) %>% 
  arrange(recode) %>%
  mutate(
    Name = factor(Name,
                  levels = unique(Name[order(recode)])),
    Identity_label=factor(Identity_label,levels=Identity_label_order_new)) %>% 
   filter(Identity != "Genome %") %>% 
  group_by(time, Identity, Identity_label,Name) %>%
  summarise(
    delta_log2FE =
      log2_FE[trt == "DOX"] -
      log2_FE[trt == "VEH"],
    .groups = "drop") %>% 
  
  mutate(time=factor(time, levels=c("24T","24R","144R")))

```

```{r full range lfe}
scaled_df_reorderd_new <- delta_df_reorderd_new %>% 
  dplyr::filter(Identity_label%in% c("repeatmasker_clean","LINE","SINE","LTR","DNA","Retroposon")) %>% 
  group_by(time) %>%
  mutate(
    scale_factor = quantile(abs(delta_log2FE), 0.9, na.rm = TRUE),
    delta_plot = delta_log2FE / scale_factor,
    delta_plot = pmax(pmin(delta_plot, 1), -1)  # clamp to [-1, 1]
  ) %>%
  ungroup()

ggplot(scaled_df_reorderd_new,
       aes(x = Identity_label,
           y = Name,
           fill = delta_plot)) +

  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "#053061",
    mid = "white",
    high = "#67001F",
    midpoint = 0,
    name = "log2(DOX / VEH)"
  ) +

  facet_wrap(~time, nrow = 1) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Delta of log2(DOX/VEH) at each timepoint, \nscaled to column min max")
```




Now I want to look at the change of log2 Fold enrichment of DOX -VEH at each timepoint


```{r delta lfe}

df_1 <- all_genes_states_long %>%
  mutate(log2_FE = log2(Fold_enrichment + 1e-6),
  Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
  separate(group, into = c("trt","time"), remove = FALSE)

delta_df_1 <- df_1 %>%
  group_by(time, Identity, Identity_label,Name) %>%
  summarise(
    delta_log2FE =
      log2_FE[trt == "DOX"] -
      log2_FE[trt == "VEH"],
    .groups = "drop") %>% 
  mutate(time=factor(time, levels=c("24T","24R","144R")))

ggplot(delta_df_1,
       aes(x = Identity_label,
           y = Name,
           fill = delta_log2FE)) +

  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "#053061",
    mid = "white",
    high = "#67001F",
    midpoint = 0,
    name = "log2(DOX / VEH)"
  ) +

  facet_wrap(~time, nrow = 1) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Delta of log2(DOX/VEH) at each timepoint")

```


#### Enhancer plot  

```{r enchacers plot, fig.height=8}
long_file %>% 
  dplyr::filter(stringr::str_detect(Identity, "SCREEN")) %>% 
   group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min),
    Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
     ggplot(.,aes(x=Identity,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=1, ncol=6)+
    scale_y_discrete(limits=rev)+
  ggtitle("All Enhancer types enrichment across states")+
  coord_fixed()
```
### Cormotif set enrichment across states  

Each Histone will have slices of their own states.  Here are all sets next to all states


```{r fig.width=12, fig.height=8}
order <- c("Genome %",
"all_H3K27ac_H3K27ac_ROI.bed",
"Set_1_H3K27ac_ROI.bed" ,
"Set_2_H3K27ac_ROI.bed",
"Set_3_H3K27ac_ROI.bed",
"all_H3K36me3_regions_H3K36me3_ROI.bed",
"Set_1_H3K36me3_ROI.bed",
"Set_2_H3K36me3_ROI.bed",
"all_H3K27me3_regions_H3K27me3_ROI.bed",
"Set_1_H3K27me3_ROI.bed",               
"Set_2_H3K27me3_ROI.bed" ,              
"all_H3K9me3_regions_H3K9me3_ROI.bed",  
"Set_1_H3K9me3_ROI.bed",               
"Set_2_H3K9me3_ROI.bed", 
"Set_3_H3K9me3_ROI.bed" ,               
"ZFP_proteins.bed")                  

all_ZFPs_states %>%
  mutate(
    `State (Emission order)` = as.character(`State (Emission order)`),
    `State (Emission order)` = case_when(
      `State (Emission order)` %in% as.character(1:16) ~ paste0("E", `State (Emission order)`),
      TRUE ~ `State (Emission order)`)) %>% 
  mutate(`State (Emission order)` = factor(`State (Emission order)`,
                                           levels = c(paste0("E",1:16), "Base"))) %>%
  filter(`State (Emission order)` != "Base") %>% 
   left_join(.,States_anno, by = c(`State (Emission order)`="State")) %>% 
  mutate(Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
 
  pivot_longer(., -c(`State (Emission order)`,group, recode, Name), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) %>% 
  mutate(Identity=factor(Identity, levels=order),
         ) %>% 
  dplyr::filter(Identity!="ZFP_proteins.bed") %>% 
  group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>% 
    ggplot(.,aes(x=Identity,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=2, ncol=3)+
    scale_y_discrete(limits=rev)+
  ggtitle("Cormotif across states")+
  coord_fixed()


  
```
```{r geometric mean of where sets are located}

all_ZFPs_states %>%
  mutate(
    `State (Emission order)` = as.character(`State (Emission order)`),
    `State (Emission order)` = case_when(
      `State (Emission order)` %in% as.character(1:16) ~ paste0("E", `State (Emission order)`),
      TRUE ~ `State (Emission order)`)) %>% 
  mutate(`State (Emission order)` = factor(`State (Emission order)`,
                                           levels = c(paste0("E",1:16), "Base"))) %>%
  filter(`State (Emission order)` != "Base") %>% 
   left_join(States_anno, by = c("State (Emission order)"="State")) %>% 
  mutate(Name = factor(Name,
                     levels = States_anno %>%
                       arrange(recode) %>%
                       pull(Name))) %>% 
 
  pivot_longer(., -c(`State (Emission order)`,group, recode, Name), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) %>% 
  mutate(Identity=factor(Identity, levels=order),
         ) %>% 
  dplyr::filter(Identity!="ZFP_proteins.bed") %>% 
#   dplyr::filter(!Identity %in% c("all_H3K36me3_regions_H3K36me3_ROI.bed",
# "Set_1_H3K36me3_ROI.bed",
# "Set_2_H3K36me3_ROI.bed",
# "all_H3K27me3_regions_H3K27me3_ROI.bed",
# "Set_1_H3K27me3_ROI.bed",               
# "Set_2_H3K27me3_ROI.bed")) %>% 
  mutate(Identity=str_remove( Identity,"_ROI.bed")) %>% 
  group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>% 
    ggplot(.,aes(x=Identity,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    scale_y_discrete(limits=rev)+
  ggtitle("Cormotif across states")+
  coord_fixed()




```

```{r delta lFE regions, fig.width=12}
plots_ZFP_long <- all_ZFPs_states %>%
  mutate(
    `State (Emission order)` = as.character(`State (Emission order)`),
    `State (Emission order)` = case_when(
      `State (Emission order)` %in% as.character(1:16) ~ paste0("E", `State (Emission order)`),
      TRUE ~ `State (Emission order)`)) %>% 
  mutate(`State (Emission order)` = factor(`State (Emission order)`,
                                           levels = c(paste0("E",1:16), "Base"))) %>%
  filter(`State (Emission order)` != "Base") %>% 
   left_join(.,States_anno, by = c(`State (Emission order)`="State")) %>% 
  mutate(Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
  pivot_longer(., -c(`State (Emission order)`,group, recode, Name), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) %>% 
  mutate(Identity=factor(Identity, levels=order),
          ID = sub("^(.*?H3K[0-9]+[a-zA-Z0-9]*)\\.?.*$", "\\1", Identity),### grabs everything up to the final H3* and removes the last bit
         ID = factor(ID,
                  levels = unique(ID[order(Identity)]))) %>%
  
  
  group_by(ID) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>% 
mutate(log2_FE = log2(Fold_enrichment + 1e-6)) %>% 
  separate(group, into = c("trt","time"), remove = FALSE) %>% 
  group_by(time, ID  ,Name) %>%
  summarise(
    delta_log2FE =
      log2_FE[trt == "DOX"] -
      log2_FE[trt == "VEH"],
    .groups = "drop") %>% 
  mutate(time=factor(time, levels=c("24T","24R","144R"))) %>% 
  dplyr::filter(!(ID %in% list("Genome %","ZFP_proteins.bed")))
  

ggplot(plots_ZFP_long,aes(x = ID,
           y = Name,
           fill = delta_log2FE)) +
  
  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "#053061",
    mid = "white",
    high = "#67001F",
    midpoint = 0,
    name = "log2(DOX / VEH)"
  ) +

  facet_wrap(~time, nrow = 1) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Delta of log2(DOX/VEH) at each timepoint")

  

```


```{r repeatplot, fig.width=10}
long_file %>% 
  dplyr::filter(stringr::str_detect(Identity, "_rptmasker")) %>% 
  mutate(rpt=case_when(Identity=="LINE_rptmasker"~"LINE",
                       Identity=="SINE_rptmasker"~"SINE",
                       Identity=="LTR_rptmasker"~"LTR",
                       Identity=="DNA_rptmasker"~"DNA",
                       Identity=="Retroposon_rptmasker"~"SVA",
                       TRUE~"Other")) %>% 
  dplyr::filter(rpt != "Other") %>% 
  group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min),
    Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>%  
    ggplot(.,aes(x=rpt,y=Name, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=2, ncol=3)+
    scale_y_discrete(limits=rev)+
  ggtitle("Transposable elements across states")+
  coord_fixed()
```
#### Looking at Delta TEs  

Now using the delta of DOX-VEH (the log2 FE of each) and plotting the states

```{r fig.height=10}

name_levels <- long_file %>%
  arrange(recode) %>%
  pull(Name) %>%
  unique()
delta_TE_df <- long_file %>% 
  dplyr::filter(stringr::str_detect(Identity, "_rptmasker")) %>% 
  mutate(rpt=case_when(Identity=="LINE_rptmasker"~"LINE",
                       Identity=="SINE_rptmasker"~"SINE",
                       Identity=="LTR_rptmasker"~"LTR",
                       Identity=="DNA_rptmasker"~"DNA",
                       Identity=="Retroposon_rptmasker"~"SVA",
                       TRUE~"Other"),
         Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>%  
  mutate(log2_FE = log2(Fold_enrichment + 1e-6)) %>% 
  separate(group, into = c("trt","time"), remove = FALSE) %>% 
  dplyr::filter(rpt != "Other") %>% 
  group_by(time,  rpt,Name) %>%
  summarise(
    delta_log2FE =
      log2_FE[trt == "DOX"] -
      log2_FE[trt == "VEH"],
    .groups = "drop") %>% 
  group_by(time) %>%
mutate(
  scale_factor = quantile(abs(delta_log2FE), 0.9, na.rm = TRUE),
  delta_plot = delta_log2FE / scale_factor,
  delta_plot = pmax(pmin(delta_plot, 1), -1)
) %>%
ungroup() %>% 
  mutate(time=factor(time, levels=c("24T","24R","144R")))



ggplot(delta_TE_df ,
       aes(x = time,
           y = Name,
           fill = delta_plot)) +
  geom_tile(color = "grey20", linewidth = 0.2) +
  
  scale_fill_gradient2(
    low =  "#2166AC", # Decrease
    mid = "white",     # No change
    high = "#B2182B",   # Increase
    midpoint = 0,
    # limits = c(-4.5, 4.5),
    name = "Δ log2(FE)\n(DOX - VEH)"
  ) +
  
  labs(
    x = "Timepoint",
    y = "ChromHMM State",
    title = "TE Enrichment Changes After DNA Damage"
  ) +
  
  facet_wrap(~rpt, ncol = 5) +  # one facet per TE type
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5)
  ) +
  coord_fixed(ratio = 1.2)+
  scale_y_discrete(limits=rev)



```


```{r scaled TE}
TE_delta_df <- delta_TE_df %>%
  # filter(rpt %in% c("LTR","LINE","SINE","DNA","SVA")) %>%
  filter(rpt %in% c("LTR")) %>% 
  mutate(
    # `State (Emission order)` = forcats::fct_rev(`State (Emission order)`),
    time = factor(time, levels = c("24T","24R","144R"))
  ) %>%
  group_by(rpt) %>%
  mutate(
    # scale relative to max abs value per TE
    delta_log2FE_scaled = delta_log2FE / max(abs(delta_log2FE), na.rm = TRUE),
    # ensure no value goes outside -1..1
    delta_log2FE_scaled = pmax(pmin(delta_log2FE_scaled, .5), -.5)
  ) %>%
  ungroup()
# Scale across the entire dataset
# max_abs_delta <- max(abs(TE_delta_df$delta_log2FE_scaled), na.rm = TRUE)

# TE_delta_df <- TE_delta_df %>%
#   mutate(delta_log2FE_scaled = delta_log2FE / max_abs_delta)

ggplot(TE_delta_df,
       aes(x = time,
           y = Name,
           fill = delta_log2FE)) +
  geom_tile(color = "grey20", linewidth = 0.2) +
  
  scale_fill_gradient2(
  low = "#2166AC",
  mid = "white",
  high = "#B2182B",
  midpoint = 0,
  limits = c(-.5, .5),
  oob = scales::squish,
  name = "Δ log2(FE)\n"
)+
  
  facet_wrap(~rpt, ncol = 5) +  # one TE type per facet
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5)
  ) +
  labs(
    x = "Timepoint",
    y = "ChromHMM State",
    title = "Scaled Δ log2(FE) per TE type after DNA damage"
  ) +
   scale_y_discrete(limits=rev)+
  coord_fixed(ratio = 1.2)
```

#### An LTR look


```{r LTR familiy}


order_LTR <- c("All_LTRs","ERV1","ERVL","ERVK","ERVL-MaLR","Gypsy","LTR")

plots_LTR_long <-all_LTR_states %>% 
  mutate(`State (Emission order)` = as.character(`State (Emission order)`),
         `State (Emission order)` = case_when(
           `State (Emission order)` %in% as.character(1:16) ~ paste0("E", `State (Emission order)`),
      TRUE ~ `State (Emission order)`)) %>% 
  mutate(`State (Emission order)` = factor(`State (Emission order)`,levels = c(paste0("E",1:16), "Base"))) %>%
  filter(`State (Emission order)` != "Base")%>% 
     mutate(Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
  pivot_longer(., -c(`State (Emission order)`,group, recode, Name), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) %>% 
  mutate(ID = case_when(
      Identity == "LTR_rptmasker.bed" ~ paste("All_LTRs"),
      str_detect(Identity, "_rptmasker") ~ str_remove(Identity, "_rptmasker.*$"),
      TRUE ~ Identity))%>%
  dplyr::filter(!(ID %in% list("Genome %","ZFP_proteins.bed"))) %>%
  mutate(ID = factor(ID,
                  levels = order_LTR))%>%
  group_by(ID) %>%   # i.e. per annotation column
  mutate(log2_FE = log2(Fold_enrichment + 1e-8)) %>% 
  separate(group, into = c("trt","time"), remove = FALSE) %>% 
  group_by(time, ID  ,Name) %>%
  summarise(
    delta_log2FE =
    mean(log2_FE[trt == "DOX"], na.rm = TRUE) -
    mean(log2_FE[trt == "VEH"], na.rm = TRUE),
    .groups = "drop") %>% 
  mutate(time=factor(time, levels=c("24T","24R","144R")))
  

ggplot(plots_LTR_long,aes(x = ID,
           y = Name,
           fill = delta_log2FE)) +
  
  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "#053061",
    mid = "white",
    high = "#67001F",
    midpoint = 0,
    name = "log2(DOX / VEH)"
  ) +

  facet_wrap(~time, nrow = 1) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Delta of log2(DOX/VEH) at each timepoint")

```

```{r}
order_LTR_new <- c("All_LTRs","ERV1","ERVL","ERVK","ERVL-MaLR","Gypsy")


plots_LTR_long_noLTR <-all_LTR_states %>% 
  mutate(`State (Emission order)` = as.character(`State (Emission order)`),
         `State (Emission order)` = case_when(
           `State (Emission order)` %in% as.character(1:16) ~ paste0("E", `State (Emission order)`),
      TRUE ~ `State (Emission order)`)) %>% 
  mutate(`State (Emission order)` = factor(`State (Emission order)`,levels = c(paste0("E",1:16), "Base"))) %>%
  filter(`State (Emission order)` != "Base")%>% 
     mutate(Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
  pivot_longer(., -c(`State (Emission order)`,group, recode, Name), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) %>% 
  mutate(ID = case_when(
      Identity == "LTR_rptmasker.bed" ~ paste("All_LTRs"),
      str_detect(Identity, "_rptmasker") ~ str_remove(Identity, "_rptmasker.*$"),
      TRUE ~ Identity))%>%
  dplyr::filter(!(ID %in% list("Genome %","ZFP_proteins.bed", "LTR"))) %>%
  mutate(ID = factor(ID,
                  levels = order_LTR_new))%>%
  group_by(ID) %>%   # i.e. per annotation column
  mutate(log2_FE = log2(Fold_enrichment + 1e-8)) %>% 
  separate(group, into = c("trt","time"), remove = FALSE) %>% 
  group_by(time, ID  ,Name) %>%
  summarise(
    delta_log2FE =
    mean(log2_FE[trt == "DOX"], na.rm = TRUE) -
    mean(log2_FE[trt == "VEH"], na.rm = TRUE),
    .groups = "drop") %>% 
  mutate(time=factor(time, levels=c("24T","24R","144R")))
  

ggplot(plots_LTR_long_noLTR,aes(x = ID,
           y = Name,
           fill = delta_log2FE)) +
  
  geom_tile(color = "grey9", linewidth = 0.1) +

  scale_fill_gradient2(
    low = "#053061",
    mid = "white",
    high = "#67001F",
    midpoint = 0,
    name = "log2(DOX / VEH)"
  ) +

  facet_wrap(~time, nrow = 1) +

  scale_y_discrete(limits = rev) +

  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ggtitle("Delta of log2(DOX/VEH) at each timepoint")
```



```{r LTR:ERV1 familiy, fig.width=13}

count_ERV1_file <- read_table("data/Bed_exports/repeatmasker_sets/ERV1/count_ERV1_file.txt", 
    col_names = FALSE) %>% dplyr::rename("count"=X1, "name"=X2) %>% 
  mutate(break_list = factor(paste0("Group", ceiling(row_number() / 20)))) %>% 
  mutate(name= str_remove(name, ".bed*$")) %>% 
  slice_head(n = length(.$name)-1)

# order_LTR_ERV1 <- c("All_LTRs","ERV1","ERVL","ERVK","ERVL-MaLR","Gypsy","LTR")

plots_LTR_ERV1_long <-all_LTR_ERV1_states %>% 
  mutate(`State (Emission order)` = as.character(`State (Emission order)`),
         `State (Emission order)` = case_when(
           `State (Emission order)` %in% as.character(1:16) ~ paste0("E", `State (Emission order)`),
      TRUE ~ `State (Emission order)`)) %>% 
  mutate(`State (Emission order)` = factor(`State (Emission order)`,levels = c(paste0("E",1:16), "Base"))) %>%
  filter(`State (Emission order)` != "Base")%>% 
     mutate(Name = factor(Name,
                  levels = unique(Name[order(recode)]))) %>% 
  pivot_longer(., -c(`State (Emission order)`,group, recode, Name), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R"))) %>% 
  mutate(ID = case_when(
      Identity == "LTR_rptmasker.bed" ~ paste("All_LTRs"),
      str_detect(Identity, ".bed") ~ str_remove(Identity, ".bed*$"),
      TRUE ~ Identity))%>%
  dplyr::filter(!(ID %in% list("Genome %","ZFP_proteins.bed"))) %>%
  # mutate(ID = factor(ID,
  #                 levels = order_LTR))%>%
  group_by(ID) %>%   # i.e. per annotation column
  mutate(log2_FE = log2(Fold_enrichment + 1e-8)) %>% 
  separate(group, into = c("trt","time"), remove = FALSE) %>% 
  group_by(time, ID  ,Name) %>%
  summarise(
    delta_log2FE =
    mean(log2_FE[trt == "DOX"], na.rm = TRUE) -
    mean(log2_FE[trt == "VEH"], na.rm = TRUE),
    .groups = "drop") %>% 
  mutate(time=factor(time, levels=c("24T","24R","144R"))) %>% 
  left_join(count_ERV1_file,by=c("ID"="name")) %>% 
  na.omit()
  
  

groups <- (unique(plots_LTR_ERV1_long$break_list)) %>% as.character()

for (g in groups) {
  p <- ggplot(
    plots_LTR_ERV1_long %>% dplyr::filter(break_list == g),
    aes(x = ID, y = Name, fill = delta_log2FE)
  ) +
    geom_tile(color = "grey9", linewidth = 0.1) +
    scale_fill_gradient2(
      low = "#053061",
      mid = "white",
      high = "#67001F",
      midpoint = 0,
      name = "log2(DOX / VEH)"
    ) +
    facet_wrap(~time, nrow = 1) +
    scale_y_discrete(limits = rev) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(paste("Delta log2(DOX/VEH) — Group", g))

  print(p)
}
 groups2 <- c("MER61E_ERV1_LTR","LTR10E_ERV1_LTR","MER61C_ERV1_LTR","MER61D_ERV1_LTR","LTR10B1_ERV1_LTR","LTR10B2_ERV1_LTR", "HERV1_LTRa_ERV1_LTR")
 ggplot(
    plots_LTR_ERV1_long %>% dplyr::filter(ID %in% groups2),
    aes(x = ID, y = Name, fill = delta_log2FE)
  ) +
    # geom_tile(color = "grey9", linewidth = 0.1) +
    # scale_fill_gradient2(
    #   low = "#053061",
    #   mid = "white",
    #   high = "#67001F",
    #   midpoint = 0,
    #   name = "log2(DOX / VEH)"
    # ) +
   
    geom_tile(color = "grey9", linewidth = 0.1) +
  scale_fill_gradient2(
    low = "#053061",
    mid = "white",
    high = "#67001F",
    midpoint = 0,
    limits = c(-5, 5),
    oob = scales::squish,
    breaks = c(-5, -2.5, 0, 2.5, 5),
    name = "log2(DOX / VEH)") +
 
    facet_wrap(~time, nrow = 1) +
    scale_y_discrete(limits = rev) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(paste("Delta log2(DOX/VEH) — Group extra"))

  
```


This is too many, so now I will filter by only those that are found in H3K27ac data.

```{r, fig.width=13}
anno_H3K27ac_summits <- readRDS("data/TE_annotation/H3K27ac_annotated_summit_TE_overlaps.RDS")

filtered_ac_ERV1s <- anno_H3K27ac_summits %>% 
  dplyr::filter(repFamily=="ERV1") %>% 
  distinct(repName)
filtered_plots <- plots_LTR_ERV1_long %>% 
  mutate(ID=str_remove(ID,"_ERV1_LTR$")) %>% 
  dplyr::filter(ID %in% filtered_ac_ERV1s$repName) 

for (g in groups) {
  p <- ggplot(
    plots_LTR_ERV1_long %>% dplyr::filter(break_list == g),
    aes(x = ID, y = Name, fill = delta_log2FE)
  ) +
    geom_tile(color = "grey9", linewidth = 0.1) +
    scale_fill_gradient2(
      low = "#053061",
      mid = "white",
      high = "#67001F",
      midpoint = 0,
      name = "log2(DOX / VEH)"
    ) +
    facet_wrap(~time, nrow = 1) +
    scale_y_discrete(limits = rev) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(paste("Delta log2(DOX/VEH) —", g,"\nFamilies found in H3K27ac summits"))

  print(p)
}
```


### ATAC and cardiac enhancer enrichment

```{r cardiac enchacers plot, fig.height=8}

ATAC_long_file <- all_ATAC_states %>% 
  filter(`State (Emission order)` != "Base") %>% 
  pivot_longer(., -c(`State (Emission order)`,group, recode, Name), names_to = "Identity", values_to="Fold_enrichment") %>% 
   mutate(group=factor(group,levels=c("VEH_24T","VEH_24R","VEH_144R","DOX_24T", "DOX_24R", "DOX_144R")),
          `State (Emission order)`=factor(`State (Emission order)`, levels = paste0("E",1:16)))


ATAC_long_file %>% 
  # dplyr::filter(stringr::str_detect(Identity, "SCREEN")) %>% 
   group_by(Identity) %>%   # i.e. per annotation column
  mutate(
    FE_min = min(Fold_enrichment, na.rm = TRUE),
    FE_max = max(Fold_enrichment, na.rm = TRUE),
    chromhmm_scaled = (Fold_enrichment - FE_min) / (FE_max - FE_min)
  ) %>% 
    ggplot(.,aes(x=Identity,y=`State (Emission order)`, fill=chromhmm_scaled))+
  geom_tile(color = "grey9",
            lwd = .1,
            linetype = 1)+
     scale_fill_gradient(
  low = "white",
  high = "#08306B",
  limits = c(0, 1),   # <- KEY
  oob = scales::squish,
   na.value = "white",   # <- this sets NAs to white
  name = "transformed scale"
)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~group, nrow=1, ncol=6)+
    scale_y_discrete(limits=rev)+
  ggtitle("CAR/DAR and cardiac Enhancer types enrichment across states")+
  coord_fixed()


```



