---
title: "Cormotif outlier removal"
author: "Renee Matthews"
date: "2025-08-19"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r packages}

library(tidyverse)
library(readr)
library(edgeR)
library(ComplexHeatmap)
library(data.table)
library(dplyr)
library(stringr)
library(ggplot2)
library(viridis)
library(DT)
library(kableExtra)
library(genomation)
library(GenomicRanges)
library(ggpubr) ## For customizing figures
library(corrplot) ## For correlation plot
library(ggpmisc)
library(gcplyr)
library(Rsubread)
library(limma)
library(ggrastr)
library(cowplot)
library(smplot2)
library(ggVennDiagram)
library(ggsignif)
library(BiocParallel)
library(gridExtra)

```

```{r data add}
sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")
```

### Cormotif code

```{r echo=TRUE, file='code/corMotifcustom.R'}

```


### Feature Counts loading  


```{r feature_cts}
H3K27ac_merged <- read_delim("data/peaks/H3K27ac_FINAL_counts.txt",
    delim = "\t", escape_double = FALSE,
    trim_ws = TRUE, skip = 1)
H3K27me3_merged <- read_delim("data/peaks/H3K27me3_FINAL_counts.txt",
    delim = "\t", escape_double = FALSE,
    trim_ws = TRUE, skip = 1)
H3K36me3_merged <- read_delim("data/peaks/H3K36me3_FINAL_counts.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE, skip = 1)
H3K9me3_merged <- read_delim("data/peaks/H3K9me3_FINAL_counts.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE, skip = 1)
rename_list <- sampleinfo %>% 
  mutate(stem= "_nobl.bam") %>% 
  mutate(prefix=paste0("/scratch/10819/styu/MW_multiQC/peaks/",Histone_Mark,"/",Treatment,"/",Timepoint,"/")) %>%
  mutate(oldname=paste0(prefix,`Library ID`,"/",`Library ID`,stem)) %>% 
  mutate(newname=paste0(Individual,"_",Treatment,"_",Timepoint)) %>% 
  dplyr::select(oldname,newname)
rename_vec <- setNames(rename_list$newname, rename_list$oldname)
names(H3K27ac_merged)[names(H3K27ac_merged) %in% names(rename_vec)] <- rename_vec[names(H3K27ac_merged)[names(H3K27ac_merged) %in% names(rename_vec)]]
names(H3K27me3_merged)[names(H3K27me3_merged) %in% names(rename_vec)] <- rename_vec[names(H3K27me3_merged)[names(H3K27me3_merged) %in% names(rename_vec)]]
names(H3K36me3_merged)[names(H3K36me3_merged) %in% names(rename_vec)] <- rename_vec[names(H3K36me3_merged)[names(H3K36me3_merged) %in% names(rename_vec)]]
names(H3K9me3_merged)[names(H3K9me3_merged) %in% names(rename_vec)] <- rename_vec[names(H3K9me3_merged)[names(H3K9me3_merged) %in% names(rename_vec)]]
```
### Removing outliers
```{r removing outliers}


H3K36me3_merged <- H3K36me3_merged %>% 
  dplyr::select(!Ind1_VEH_144R)

H3K9me3_merged<- H3K9me3_merged %>%
  dplyr::select(!Ind1_VEH_24T) %>% 
  dplyr::select(!Ind3_DOX_24T) %>%
  dplyr::select(!Ind5_DOX_144R)
```

```{r making merged raw lcpm}
H3K27ac_merged_raw <- H3K27ac_merged %>%
  dplyr::select(Geneid,contains("Ind")) %>%
  column_to_rownames("Geneid") %>%
  as.matrix()
H3K27ac_merged_lcpm <- H3K27ac_merged %>%
  dplyr::select(Geneid,contains("Ind")) %>%
  column_to_rownames("Geneid") %>%
  cpm(., log = TRUE)
H3K27ac_merged_cor <- H3K27ac_merged_lcpm %>%
  cor()


H3K27me3_merged_raw <- H3K27me3_merged %>%
  dplyr::select(Geneid,contains("Ind")) %>%
  column_to_rownames("Geneid") %>%
  as.matrix()
H3K27me3_merged_lcpm <- H3K27me3_merged %>%
  dplyr::select(Geneid,contains("Ind")) %>%
  column_to_rownames("Geneid") %>%
  cpm(., log = TRUE)
H3K27me3_merged_cor <- H3K27me3_merged_lcpm %>%
  cor()

H3K36me3_merged_raw <- H3K36me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  as.matrix()
H3K36me3_merged_lcpm <- H3K36me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  cpm(., log = TRUE)
H3K36me3_merged_cor <- H3K36me3_merged_lcpm %>% 
  cor()

H3K9me3_merged_raw <- H3K9me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  as.matrix()
H3K9me3_merged_lcpm <- H3K9me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  cpm(., log = TRUE)
H3K9me3_merged_cor <- H3K9me3_merged_lcpm %>% 
  cor()

```

Removing chrX and chrY
```{r filtering x and Y}


H3K27ac_merged_raw <- H3K27ac_merged_raw[rowMeans(H3K27ac_merged_lcpm)>0,]
H3K27ac_merged_raw <- H3K27ac_merged_raw[!grepl("chrY",rownames(H3K27ac_merged_raw)),]
H3K27ac_merged_raw <- H3K27ac_merged_raw[!grepl("chrX",rownames(H3K27ac_merged_raw)),]

H3K27me3_merged_raw <- H3K27me3_merged_raw[rowMeans(H3K27me3_merged_lcpm)>0,]
H3K27me3_merged_raw <- H3K27me3_merged_raw[!grepl("chrY",rownames(H3K27me3_merged_raw)),]
H3K27me3_merged_raw <- H3K27me3_merged_raw[!grepl("chrX",rownames(H3K27me3_merged_raw)),]


H3K36me3_merged_raw <- H3K36me3_merged_raw[rowMeans(H3K36me3_merged_lcpm)>0,]
H3K36me3_merged_raw <- H3K36me3_merged_raw[!grepl("chrY",rownames(H3K36me3_merged_raw)),]
H3K36me3_merged_raw <- H3K36me3_merged_raw[!grepl("chrX",rownames(H3K36me3_merged_raw)),]

H3K9me3_merged_raw <- H3K9me3_merged_raw[rowMeans(H3K9me3_merged_lcpm)>0,]
H3K9me3_merged_raw <- H3K9me3_merged_raw[!grepl("chrY",rownames(H3K9me3_merged_raw)),]
H3K9me3_merged_raw <- H3K9me3_merged_raw[!grepl("chrX",rownames(H3K9me3_merged_raw)),]
```

#### Setting up Matrix
```{r matrix}
H3K27ac_annomat <- data.frame(timeset=colnames(H3K27ac_merged_raw)) %>%
  mutate(sample=timeset) %>%
  separate(timeset, into = c("ind","tx","time")) %>%
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(ind = gsub("Ind", "", ind)) %>%
  mutate(txtime = paste0(tx, "_", time)) %>%
  mutate(group = txtime)
H3K27ac_annomat$group <- H3K27ac_annomat$group %>%
  gsub("DOX_24T", "1", .) %>%
  gsub("DOX_24R", "2", .) %>%
  gsub("DOX_144R", "3", .) %>%
  gsub("VEH_24T", "4", .) %>%
  gsub("VEH_24R", "5", .) %>%
  gsub("VEH_144R", "6", .)

H3K27me3_annomat <- data.frame(timeset=colnames(H3K27me3_merged_raw)) %>%
  mutate(sample=timeset) %>%
  separate(timeset, into = c("ind","tx","time")) %>%
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(ind = gsub("Ind", "", ind)) %>%
  mutate(txtime = paste0(tx, "_", time)) %>%
  mutate(group = txtime)
H3K27me3_annomat$group <- H3K27me3_annomat$group %>%
  gsub("DOX_24T", "1", .) %>%
  gsub("DOX_24R", "2", .) %>%
  gsub("DOX_144R", "3", .) %>%
  gsub("VEH_24T", "4", .) %>%
  gsub("VEH_24R", "5", .) %>%
  gsub("VEH_144R", "6", .)

H3K36me3_annomat <- data.frame(timeset=colnames(H3K36me3_merged_raw)) %>% 
  mutate(sample=timeset) %>% 
  separate(timeset, into = c("ind","tx","time")) %>% 
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(ind = gsub("Ind", "", ind)) %>%
  mutate(txtime = paste0(tx, "_", time)) %>%
  mutate(group = txtime)
H3K36me3_annomat$group <- H3K36me3_annomat$group %>%
  gsub("DOX_24T", "1", .) %>%
  gsub("DOX_24R", "2", .) %>%
  gsub("DOX_144R", "3", .) %>%
  gsub("VEH_24T", "4", .) %>%
  gsub("VEH_24R", "5", .) %>%
  gsub("VEH_144R", "6", .)

H3K9me3_annomat <- data.frame(timeset=colnames(H3K9me3_merged_raw)) %>% 
  mutate(sample=timeset) %>% 
  separate(timeset, into = c("ind","tx","time")) %>% 
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(ind = gsub("Ind", "", ind)) %>%
  mutate(txtime = paste0(tx, "_", time)) %>%
  mutate(group = txtime)
H3K9me3_annomat$group <- H3K9me3_annomat$group %>%
  gsub("DOX_24T", "1", .) %>%
  gsub("DOX_24R", "2", .) %>%
  gsub("DOX_144R", "3", .) %>%
  gsub("VEH_24T", "4", .) %>%
  gsub("VEH_24R", "5", .) %>%
  gsub("VEH_144R", "6", .)

dge_H3K27ac <- edgeR::DGEList(counts = H3K27ac_merged_raw, group = H3K27ac_annomat$group, genes = row.names(H3K27ac_merged_raw))
dge_H3K27me3 <- edgeR::DGEList(counts = H3K27me3_merged_raw, group = H3K27me3_annomat$group, genes = row.names(H3K27me3_merged_raw))
dge_H3K36me3 <- edgeR::DGEList(counts = H3K36me3_merged_raw, group = H3K36me3_annomat$group, genes = row.names(H3K36me3_merged_raw))
dge_H3K9me3 <- edgeR::DGEList(counts = H3K9me3_merged_raw, group = H3K9me3_annomat$group, genes = row.names(H3K9me3_merged_raw))

dge_H3K27ac <- edgeR::calcNormFactors(dge_H3K27ac)
dge_H3K27me3 <- edgeR::calcNormFactors(dge_H3K27me3)
dge_H3K36me3 <- edgeR::calcNormFactors(dge_H3K36me3)
dge_H3K9me3 <- edgeR::calcNormFactors(dge_H3K9me3)


```


making comparison and group matrices
```{r compid and groupid}
lcpm_dge_H3K27ac <- cpm(dge_H3K27ac, log = TRUE)
lcpm_dge_H3K27me3 <- cpm(dge_H3K27me3, log = TRUE)
lcpm_dge_H3K36me3 <- cpm(dge_H3K36me3, log = TRUE)
lcpm_dge_H3K9me3 <- cpm(dge_H3K9me3, log = TRUE)

H3K27ac_group <- H3K27ac_annomat$group
H3K27me3_group <- H3K27me3_annomat$group
H3K36me3_group <- H3K36me3_annomat$group
H3K9me3_group <- H3K9me3_annomat$group

compid <- data.frame(c1= c(1,2,3), c2 = c( 4,5,6))
compid
```

### H3K27ac cormotif  

```{r H3K27ac cormotif, eval=FALSE, include=TRUE}
set.seed(31415)
cormotif_initial_H3K27ac <- cormotiffit(exprs = lcpm_dge_H3K27ac, groupid = H3K27ac_group, compid = compid, K=1:8, max.iter = 500, runtype = "logCPM")

saveRDS(cormotif_initial_H3K27ac,"data/Cormotif_data/Cormotif_initial_H3K27ac.RDS")
```


```{r H3K27ac plots}

cormotif_initial_H3K27ac <- readRDS("data/Cormotif_data/Cormotif_initial_H3K27ac.RDS")
plotIC(cormotif_initial_H3K27ac)
plotMotif(cormotif_initial_H3K27ac)
```
### H3K27me3 cormotif  

```{r H3K27me3 cormotif, eval=FALSE, include=TRUE}
set.seed(31415)
cormotif_initial_H3K27me3 <- cormotiffit(exprs = lcpm_dge_H3K27me3, groupid = H3K27me3_group, compid = compid, K=1:8, max.iter = 500, runtype = "logCPM")

saveRDS(cormotif_initial_H3K27me3,"data/Cormotif_data/Cormotif_initial_H3K27me3.RDS")
```

```{r H3K27me3 plots}
cormotif_initial_H3K27me3 <- readRDS("data/Cormotif_data/Cormotif_initial_H3K27me3.RDS")
plotIC(cormotif_initial_H3K27me3)
plotMotif(cormotif_initial_H3K27me3)
```

### H3K36me3 cormotif  

```{r H3K36me3 cormotif, eval=FALSE, include=TRUE}
set.seed(31415)
cormotif_initial_H3K36me3 <- cormotiffit(exprs = lcpm_dge_H3K36me3, groupid = H3K36me3_group, compid = compid, K=1:8, max.iter = 500, runtype = "logCPM")

saveRDS(cormotif_initial_H3K36me3,"data/Cormotif_data/Cormotif_initial_H3K36me3_nooutlier.RDS")
```

```{r  H3K36me3 plots}
cormotif_initial_H3K36me3 <- readRDS("data/Cormotif_data/Cormotif_initial_H3K36me3_nooutlier.RDS")

myColors <-  rev(c("#FFFFFF", "#E6E6E6" ,"#CCCCCC", "#B3B3B3", "#999999", "#808080", "#666666","#4C4C4C", "#333333", "#191919","#000000"))

plot.new()
legend('center',fill=myColors, legend =rev(c("0", "0.1", "0.2", "0.3", "0.4",  "0.5", "0.6", "0.7", "0.8","0.9", "1")), box.col="white",title = "Probability\nlegend", horiz=FALSE,title.cex=.8)

plotIC(cormotif_initial_H3K36me3)
plotMotif(cormotif_initial_H3K36me3)
```

### H3K9me3 cormotif  

```{r H3K9me3 cormotif,eval=FALSE, include=TRUE}
set.seed(31415)
cormotif_initial_H3K9me3 <- cormotiffit(exprs = lcpm_dge_H3K9me3, groupid = H3K9me3_group, compid = compid, K=1:8, max.iter = 500, runtype = "logCPM")

saveRDS(cormotif_initial_H3K9me3,"data/Cormotif_data/Cormotif_initial_H3K9me3_nooutlier.RDS")
```

```{r H3K9me3 cormotifplots}

cormotif_initial_H3K9me3 <- readRDS("data/Cormotif_data/Cormotif_initial_H3K9me3_nooutlier.RDS")
plotIC(cormotif_initial_H3K9me3)
plotMotif(cormotif_initial_H3K9me3)
```
#### Breaking into motif modules
```{r motif_modules}

row.names(cormotif_initial_H3K27ac$bestmotif$clustlike) <- row.names(lcpm_dge_H3K27ac)
row.names(cormotif_initial_H3K27ac$bestmotif$p.post) <- row.names(lcpm_dge_H3K27ac)

 row.names(cormotif_initial_H3K27me3$bestmotif$clustlike) <- row.names(lcpm_dge_H3K27me3)
 row.names(cormotif_initial_H3K27me3$bestmotif$p.post) <- row.names(lcpm_dge_H3K27me3)

row.names(cormotif_initial_H3K36me3$bestmotif$clustlike) <- row.names(lcpm_dge_H3K36me3)
row.names(cormotif_initial_H3K36me3$bestmotif$p.post) <- row.names(lcpm_dge_H3K36me3)

row.names(cormotif_initial_H3K9me3$bestmotif$clustlike) <- row.names(lcpm_dge_H3K9me3)
row.names(cormotif_initial_H3K9me3$bestmotif$p.post) <- row.names(lcpm_dge_H3K9me3)


```

#### H3K27ac motifs  

```{r H3K27ac motif module}
###goal is to get the intersection between clustlike and p.post
H3K27ac_motif1 <- cormotif_initial_H3K27ac$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1>.6 & V2<.5 & V3 <.5) %>% 
  rownames_to_column("Peakid")

H3K27ac_motif1_pp <- cormotif_initial_H3K27ac$bestmotif$p.post %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.4 & V2<.4 & V3 <.4) %>% 
  rownames_to_column("Peakid")

H3K27ac_motif2 <- cormotif_initial_H3K27ac$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.5 & V2>.6 & V3 <.5) %>% 
  rownames_to_column("Peakid")

H3K27ac_motif2_pp <- cormotif_initial_H3K27ac$bestmotif$p.post %>% 
  as.data.frame() %>% 
 dplyr::filter(V1<.6 & V2>.5 & V3 >.5) %>% 
  rownames_to_column("Peakid")

H3K27ac_motif3 <- cormotif_initial_H3K27ac$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.5 & V2<.5 & V3 >.6) %>% 
  rownames_to_column("Peakid")

H3K27ac_motif3_pp <- cormotif_initial_H3K27ac$bestmotif$p.post %>% 
  as.data.frame() %>% 
  dplyr::filter(V1>.9 & V2>.5 & V3 <.5) %>% 
  rownames_to_column("Peakid")

length(intersect(H3K27ac_motif1$Peakid,H3K27ac_motif1_pp$Peakid))
H3K27ac_set1 <- intersect(H3K27ac_motif1$Peakid,H3K27ac_motif1_pp$Peakid)%>% as.data.frame()
length(intersect(H3K27ac_motif2$Peakid,H3K27ac_motif2_pp$Peakid))
H3K27ac_set2 <- intersect(H3K27ac_motif2$Peakid,H3K27ac_motif2_pp$Peakid)%>% as.data.frame()
length(intersect(H3K27ac_motif3$Peakid,H3K27ac_motif3_pp$Peakid))
H3K27ac_set3 <- intersect(H3K27ac_motif3$Peakid,H3K27ac_motif3_pp$Peakid) %>% as.data.frame()

length(intersect(H3K27ac_motif1$Peakid,H3K27ac_motif2$Peakid))

length(intersect(H3K27ac_motif3$Peakid,H3K27ac_motif2$Peakid))

length(intersect(H3K27ac_motif3$Peakid,H3K27ac_motif2$Peakid))


write_delim(H3K27ac_set1,"data/motif_lists/H3K27ac_set1.txt",delim="\t")
write_delim(H3K27ac_set2,"data/motif_lists/H3K27ac_set2.txt",delim="\t")
write_delim(H3K27ac_set3,"data/motif_lists/H3K27ac_set3.txt",delim="\t")

```

#### H3K27me3 motifs  

```{r H3K27me3 motif module}
###goal is to get the intersection between clustlike and p.post
H3K27me3_motif1 <- cormotif_initial_H3K27me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1>.6 & V2<.5) %>% 
  rownames_to_column("Peakid")

H3K27me3_motif1_pp <- cormotif_initial_H3K27me3$bestmotif$p.post %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.4 & V2<.4 & V3 <.4) %>% 
  rownames_to_column("Peakid")

H3K27me3_motif2 <- cormotif_initial_H3K27me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.5 & V2>.6 ) %>% 
  rownames_to_column("Peakid")

H3K27me3_motif2_pp <- cormotif_initial_H3K27me3$bestmotif$p.post %>% 
  as.data.frame() %>% 
 dplyr::filter(V1>.7 & V2>.7 & V3 >.7) %>% 
  rownames_to_column("Peakid")


length(intersect(H3K27me3_motif1$Peakid,H3K27me3_motif1_pp$Peakid))
H3K27me3_set1 <- intersect(H3K27me3_motif1$Peakid,H3K27me3_motif1_pp$Peakid)%>% as.data.frame()
length(intersect(H3K27me3_motif2$Peakid,H3K27me3_motif2_pp$Peakid))
H3K27me3_set2 <- intersect(H3K27me3_motif2$Peakid,H3K27me3_motif2_pp$Peakid)%>% as.data.frame()

length(intersect(H3K27me3_motif1$Peakid,H3K27me3_motif2$Peakid))


write_delim(H3K27me3_set1,"data/motif_lists/H3K27me3_set1.txt",delim="\t")
write_delim(H3K27me3_set2,"data/motif_lists/H3K27me3_set2.txt",delim="\t")


```

#### H3K36me3 motifs  

```{r H3K36me3 motif module}
###goal is to get the intersection between clustlike and p.post
H3K36me3_motif1 <- cormotif_initial_H3K36me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1>.6 & V2<.5) %>% 
  rownames_to_column("Peakid")

H3K36me3_motif1_pp <- cormotif_initial_H3K36me3$bestmotif$p.post %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.1 & V2<.1 & V3 <.1) %>% 
  rownames_to_column("Peakid")

H3K36me3_motif2 <- cormotif_initial_H3K36me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.5 & V2>.6) %>% 
  rownames_to_column("Peakid")

H3K36me3_motif2_pp <- cormotif_initial_H3K36me3$bestmotif$p.post %>% 
  as.data.frame() %>% 
 dplyr::filter(V1>.4 & V2>.4 & V3 <.1) %>% 
  rownames_to_column("Peakid")



length(intersect(H3K36me3_motif1$Peakid,H3K36me3_motif1_pp$Peakid))
H3K36me3_set1 <- intersect(H3K36me3_motif1$Peakid,H3K36me3_motif1_pp$Peakid)%>% as.data.frame()
length(intersect(H3K36me3_motif2$Peakid,H3K36me3_motif2_pp$Peakid))
H3K36me3_set2 <- intersect(H3K36me3_motif2$Peakid,H3K36me3_motif2_pp$Peakid)%>% as.data.frame()

length(intersect(H3K36me3_motif1$Peakid,H3K36me3_motif2$Peakid))


write_delim(H3K36me3_set1,"data/motif_lists/H3K36me3_set1.txt",delim="\t")
write_delim(H3K36me3_set2,"data/motif_lists/H3K36me3_set2.txt",delim="\t")



```

#### H3K9me3 motifs  

```{r H3K9me3 motif module}
###goal is to get the intersection between clustlike and p.post
H3K9me3_motif1 <- cormotif_initial_H3K9me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1>.6 & V2<.5 & V3 <.5) %>% 
  rownames_to_column("Peakid")

H3K9me3_motif1_pp <- cormotif_initial_H3K9me3$bestmotif$p.post %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.4 & V2<.4 & V3 <.4) %>% 
  rownames_to_column("Peakid")

H3K9me3_motif2 <- cormotif_initial_H3K9me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.5 & V2>.6 & V3 <.5) %>% 
  rownames_to_column("Peakid")

H3K9me3_motif2_pp <- cormotif_initial_H3K9me3$bestmotif$p.post %>% 
  as.data.frame() %>% 
 dplyr::filter(V1>.8 & V2<.5 & V3 <.5) %>% 
  rownames_to_column("Peakid")

H3K9me3_motif3 <- cormotif_initial_H3K9me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  dplyr::filter(V1<.5 & V2<.5 & V3 >.6) %>% 
  rownames_to_column("Peakid")

H3K9me3_motif3_pp <- cormotif_initial_H3K9me3$bestmotif$p.post %>% 
  as.data.frame() %>% 
  dplyr::filter(V1>.9 & V2>.9 & V3 >.3) %>% 
  rownames_to_column("Peakid")

length(intersect(H3K9me3_motif1$Peakid,H3K9me3_motif1_pp$Peakid))
H3K9me3_set1 <- intersect(H3K9me3_motif1$Peakid,H3K9me3_motif1_pp$Peakid)%>% as.data.frame()
length(intersect(H3K9me3_motif2$Peakid,H3K9me3_motif2_pp$Peakid))
H3K9me3_set2 <- intersect(H3K9me3_motif2$Peakid,H3K9me3_motif2_pp$Peakid)%>% as.data.frame()
length(intersect(H3K9me3_motif3$Peakid,H3K9me3_motif3_pp$Peakid))
H3K9me3_set3 <- intersect(H3K9me3_motif3$Peakid,H3K9me3_motif3_pp$Peakid)%>% as.data.frame()

length(intersect(H3K9me3_motif1$Peakid,H3K9me3_motif2$Peakid))

length(intersect(H3K9me3_motif3$Peakid,H3K9me3_motif2$Peakid))

length(intersect(H3K9me3_motif3$Peakid,H3K9me3_motif2$Peakid))


write_delim(H3K9me3_set1,"data/motif_lists/H3K9me3_set1.txt",delim="\t")
write_delim(H3K9me3_set2,"data/motif_lists/H3K9me3_set2.txt",delim="\t")
write_delim(H3K9me3_set3,"data/motif_lists/H3K9me3_set3.txt",delim="\t")

```

### Looking at PC contributions to variation  


```{r contribution function}



pca_plot_new <- function(df, pc_x = 1, pc_y = 2, col_var, shape_var, pca_obj, title = "") {
  # variance explained
  var_expl <- (pca_obj$sdev^2) / sum(pca_obj$sdev^2)
  
  # axis labels
  x_lab <- sprintf("PC%d (%.1f%%)", pc_x, var_expl[pc_x] * 100)
  y_lab <- sprintf("PC%d (%.1f%%)", pc_y, var_expl[pc_y] * 100)
  
  ggplot(df, aes(x = .data[[paste0("PC", pc_x)]], 
                 y = .data[[paste0("PC", pc_y)]], 
                 color = !!sym(col_var), 
                 shape = !!sym(shape_var))) +
    geom_point(size = 5) +
    labs(title = title, x = x_lab, y = y_lab) +
    scale_color_manual(values = c(
      "#8B006D","#DF707E","#F1B72B",
      "#3386DD","#707031","#41B333"
    ))
}
pca_plot <-
  function(df, col_var = NULL, shape_var = NULL, title = "") {
    ggplot(df) + geom_point(aes(
      x = PC1,
      y = PC2,
      color = col_var,
      shape = shape_var
    ),
    size = 5) +
      labs(title = title, x = "PC 1", y = "PC 2") +
      scale_color_manual(values = c(
        "#8B006D",
        "#DF707E",
        "#F1B72B",
        "#3386DD",
        "#707031",
        "#41B333"
      ))
  }
get_regr_pval <- function(mod) {
  # Returns the p-value for the Fstatistic of a linear model
  # mod: class lm
  stopifnot(class(mod) == "lm")
  fstat <- summary(mod)$fstatistic
  pval <- 1 - pf(fstat[1], fstat[2], fstat[3])
  return(pval)
}

plot_versus_pc <- function(df, pc_num, fac) {
  # df: data.frame
  # pc_num: numeric, specific PC for plotting
  # fac: column name of df for plotting against PC
  pc_char <- paste0("PC", pc_num)
  # Calculate F-statistic p-value for linear model
  pval <- get_regr_pval(lm(df[[ pc_char]] ~ df[[ fac]]))
  if (is.numeric(df[, fac])) {
    ggplot(df, aes_string(x = fac, y = pc_char)) + geom_point() +
      geom_smooth(method = "lm") + labs(title = sprintf("p-val: %.2f", pval))
  } else {
    ggplot(df, aes_string(x = fac, y = pc_char)) + geom_boxplot() +
      labs(title = sprintf("p-val: %.3f", pval))
  }
}
  

```


#### H3K27ac contribution

```{r Looking at H3K27ac}

H3K27ac_pr <- prcomp(t(lcpm_dge_H3K27ac)) 

H3K27ac_pr_anno <- H3K27ac_pr$x %>% 
  as.data.frame() %>% 
  rownames_to_column("sample") %>% 
  separate(sample, into = c("ind","tx","time")) %>%
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(txtime = paste0(tx, "_", time)) %>% 
  mutate(sex=if_else(ind %in% c("Ind1","Ind2"),"F","M"))%>% 
  mutate(sex=factor(sex, levels=c("F","M")))


facs <- c("ind", "tx", "time", "sex")
names(facs) <- c("Individual", "Treatment", "Time","Sex")

plots_H3K27ac <- lapply(names(facs), function(f_label) {
  f <- facs[f_label]
  f_v_pc1 <- plot_versus_pc(H3K27ac_pr_anno, 1, f)
  f_v_pc2 <- plot_versus_pc(H3K27ac_pr_anno, 2, f)
  f_v_pc3 <- plot_versus_pc(H3K27ac_pr_anno, 3, f)
  f_v_pc4 <- plot_versus_pc(H3K27ac_pr_anno, 4, f)

  grid.arrange(f_v_pc1, f_v_pc2, f_v_pc3,f_v_pc4, ncol = 2, top = paste0(f_label," H3K27ac"))
})
```

#### H3K27me contrib.


```{r Looking at H3K27me3}

H3K27me3_pr <- prcomp(t(lcpm_dge_H3K27me3)) 

H3K27me3_pr_anno <- H3K27me3_pr$x %>% 
  as.data.frame() %>% 
  rownames_to_column("sample") %>% 
  separate(sample, into = c("ind","tx","time")) %>%
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(txtime = paste0(tx, "_", time)) %>% 
  mutate(sex=if_else(ind %in% c("Ind1","Ind2"),"F","M"))%>% 
  mutate(sex=factor(sex, levels=c("F","M")))


facs <- c("ind", "tx", "time", "sex")
names(facs) <- c("Individual", "Treatment", "Time","Sex")

plots_H3K27me3 <- lapply(names(facs), function(f_label) {
  f <- facs[f_label]
  f_v_pc1 <- plot_versus_pc(H3K27me3_pr_anno, 1, f)
  f_v_pc2 <- plot_versus_pc(H3K27me3_pr_anno, 2, f)
  f_v_pc3 <- plot_versus_pc(H3K27me3_pr_anno, 3, f)
  f_v_pc4 <- plot_versus_pc(H3K27me3_pr_anno, 4, f)

  grid.arrange(f_v_pc1, f_v_pc2, f_v_pc3,f_v_pc4, ncol = 2, top = paste0(f_label," H3K27me3"))
})


```

#### H3K36me3 contrib.

```{r Looking at H3K36me3}

H3K36me3_pr <- prcomp(t(lcpm_dge_H3K36me3)) 

H3K36me3_pr_anno <- H3K36me3_pr$x %>% 
  as.data.frame() %>% 
  rownames_to_column("sample") %>% 
  separate(sample, into = c("ind","tx","time")) %>%
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(txtime = paste0(tx, "_", time)) %>% 
  mutate(sex=if_else(ind %in% c("Ind1","Ind2"),"F","M"))%>% 
  mutate(sex=factor(sex, levels=c("F","M")))


facs <- c("ind", "tx", "time", "sex")
names(facs) <- c("Individual", "Treatment", "Time","Sex")

plots_H3K36me3 <- lapply(names(facs), function(f_label) {
  f <- facs[f_label]
  f_v_pc1 <- plot_versus_pc(H3K36me3_pr_anno, 1, f)
  f_v_pc2 <- plot_versus_pc(H3K36me3_pr_anno, 2, f)
  f_v_pc3 <- plot_versus_pc(H3K36me3_pr_anno, 3, f)
  f_v_pc4 <- plot_versus_pc(H3K36me3_pr_anno, 4, f)

  grid.arrange(f_v_pc1, f_v_pc2, f_v_pc3, f_v_pc4, ncol = 2, top = paste0(f_label," H3K36me3"))
})
```

#### H3K9me3

```{r Looking at H3K9me3}

H3K9me3_pr <- prcomp(t(lcpm_dge_H3K9me3)) 

H3K9me3_pr_anno <- H3K9me3_pr$x %>% 
  as.data.frame() %>% 
  rownames_to_column("sample") %>% 
  separate(sample, into = c("ind","tx","time")) %>%
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(txtime = paste0(tx, "_", time)) %>% 
  mutate(sex=if_else(ind %in% c("Ind1","Ind2"),"F","M"))%>% 
  mutate(sex=factor(sex, levels=c("F","M")))


facs <- c("ind", "tx", "time", "sex")
names(facs) <- c("Individual", "Treatment", "Time","Sex")

plots_H3K9me3 <- lapply(names(facs), function(f_label) {
  f <- facs[f_label]
  f_v_pc1 <- plot_versus_pc(H3K9me3_pr_anno, 1, f)
  f_v_pc2 <- plot_versus_pc(H3K9me3_pr_anno, 2, f)
  f_v_pc3 <- plot_versus_pc(H3K9me3_pr_anno, 3, f)
  f_v_pc4 <- plot_versus_pc(H3K9me3_pr_anno, 4, f)

  grid.arrange(f_v_pc1, f_v_pc2, f_v_pc3,f_v_pc4, ncol = 2, top = paste0(f_label," H3K9me3"))
})
```
