---
title: "TEs across histones"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(ChIPseeker)
library(ggVennDiagram)
library(smplot2)
```

First steps: Pulling and overlapping each ROI set with specific families of TEs 
```{r repeatmasker loading}

repeatmasker <- read_delim("data/Other_paper_data/repeatmasker_20250911.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)


autosomes <- paste0("chr", 1:22)
repeatmasker_clean <- repeatmasker %>% mutate(
    strand = ifelse(strand == "C", "-", "+")
  ) %>% 
   mutate(
    start = genoStart + 1,
    end   = genoEnd)%>% 
  mutate(repFamily= str_remove(repFamily, "\\?$")) %>% 
  dplyr::filter(genoName %in% autosomes) %>% 
  mutate(RM_id=paste0(genoName,":",start,"-",end,":",id))


rpt_split <- split(repeatmasker_clean, repeatmasker_clean$repClass)

rpt_split_gr_list <- lapply(rpt_split, function(df) {
  GRanges(
    seqnames = df$genoName,
    ranges   = IRanges(start = df$start, end = df$end),
    strand   = df$strand,
    repName   = df$repName,
    repClass  = df$repClass,
    repFamily = df$repFamily,
    swScore   = df$swScore,
    milliDiv  = df$milliDiv,
    milliDel  = df$milliDel,
    milliIns  = df$milliIns,
    RM_id        = df$RM_id
  )
})



```

```{r subset rpt_split_gr_list}
SINE_gr <- rpt_split_gr_list$SINE
SINE_df <- SINE_gr %>% 
  as.data.frame()

SINE_split_df <- split(SINE_df, SINE_df$repFamily)

LINE_gr <- rpt_split_gr_list$LINE
LINE_df <- LINE_gr %>% 
  as.data.frame()

LINE_split_df <- split(LINE_df, LINE_df$repFamily)

LTR_gr <- rpt_split_gr_list$LTR
LTR_df <- LTR_gr %>% 
  as.data.frame()

LTR_split_df <- split(LTR_df, LTR_df$repFamily)

SVA_gr <- rpt_split_gr_list$Retroposon
SVA_df <- SVA_gr %>% 
  as.data.frame()

SVA_split_df <- split(SVA_df, SVA_df$repFamily)

DNA_gr <- rpt_split_gr_list$DNA
DNA_df <- DNA_gr %>% 
  as.data.frame()

DNA_split_df <- split(DNA_df, DNA_df$repFamily)


```





Pulling in ROI granges for H3K27ac and H3K9me3
```{r loading H3K9me3 data}

peakAnnoList_H3K9me3 <- readRDS("data/motif_lists/H3K9me3_annotated_peaks.RDS")
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

H3K9me3_lookup <- imap_dfr(peakAnnoList_H3K9me3[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)

H3K9me3_sets_gr <- lapply(peakAnnoList_H3K9me3, function(df) {
  as_granges(df)
})

H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})
```

```{r toplist data}
H3K9me3_toplist <- readRDS( "data/DER_data/H3K9me3_toplist_nooutlier.RDS")

H3K27ac_toplist <- readRDS( "data/DER_data/H3K27ac_toplist.RDS")

H3K27ac_toptable_list <- bind_rows(H3K27ac_toplist, .id = "group")

H3K9me3_toptable_list <- bind_rows(H3K9me3_toplist, .id = "group")

K9me3_lfctable <- H3K9me3_toptable_list %>% 
  dplyr::select(group,genes, logFC) %>% 
  pivot_wider(.,id_cols = genes, names_from = group, values_from = logFC)

K27ac_lfctable <- H3K27ac_toptable_list %>% 
  dplyr::select(group,genes, logFC) %>% 
  pivot_wider(.,id_cols = genes, names_from = group, values_from = logFC)
```



### SINE
Overlapping SINE family with ROIs

```{r SINE Overlapping}
sine_hits_H3K9me3 <- findOverlaps(H3K9me3_sets_gr$all_H3K9me3_regions, SINE_gr, ignore.strand = TRUE)

SINE_overlap_df_H3K9me3 <- tibble(
  peak_row = queryHits(sine_hits_H3K9me3),
  Peakid   = H3K9me3_sets_gr$all_H3K9me3$Peakid[queryHits(sine_hits_H3K9me3)],
  cluster  = H3K9me3_sets_gr$all_H3K9me3$cluster[queryHits(sine_hits_H3K9me3)],
  
  repClass = SINE_gr$repClass[subjectHits(sine_hits_H3K9me3)],
  repName  = SINE_gr$repName[subjectHits(sine_hits_H3K9me3)],
  RM_id  = SINE_gr$RM_id[subjectHits(sine_hits_H3K9me3)],
  
  TE_type  = ifelse(
    SINE_gr$repFamily[subjectHits(sine_hits_H3K9me3)] == "SVA",
    SINE_gr$repName[subjectHits(sine_hits_H3K9me3)],
    SINE_gr$repFamily[subjectHits(sine_hits_H3K9me3)]
  ),
  
  milliDiv = SINE_gr$milliDiv[subjectHits(sine_hits_H3K9me3)],
  milliDel = SINE_gr$milliDel[subjectHits(sine_hits_H3K9me3)],
  milliIns = SINE_gr$milliIns[subjectHits(sine_hits_H3K9me3)]
)

sine_hits_H3K27ac <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, SINE_gr, ignore.strand = TRUE)

SINE_overlap_df_H3K27ac <- tibble(
  peak_row = queryHits(sine_hits_H3K27ac),
  Peakid   = H3K27ac_sets_gr$all_H3K27ac$Peakid[queryHits(sine_hits_H3K27ac)],
  cluster  = H3K27ac_sets_gr$all_H3K27ac$cluster[queryHits(sine_hits_H3K27ac)],
  
  repClass = SINE_gr$repClass[subjectHits(sine_hits_H3K27ac)],
  repName  = SINE_gr$repName[subjectHits(sine_hits_H3K27ac)],
  RM_id  = SINE_gr$RM_id[subjectHits(sine_hits_H3K27ac)],
  
  TE_type  = ifelse(
    SINE_gr$repFamily[subjectHits(sine_hits_H3K27ac)] == "SVA",
    SINE_gr$repName[subjectHits(sine_hits_H3K27ac)],
    SINE_gr$repFamily[subjectHits(sine_hits_H3K27ac)]
  ),
  
  milliDiv = SINE_gr$milliDiv[subjectHits(sine_hits_H3K27ac)],
  milliDel = SINE_gr$milliDel[subjectHits(sine_hits_H3K27ac)],
  milliIns = SINE_gr$milliIns[subjectHits(sine_hits_H3K27ac)]
)

common_sine_RMs <- intersect(
  SINE_overlap_df_H3K27ac$RM_id,
  SINE_overlap_df_H3K9me3$RM_id
)

length(unique(common_sine_RMs))


```

Visualizing the overlapping SINE families

```{r}
H3K27ac_SINE_peakid <- SINE_overlap_df_H3K27ac %>% 
  dplyr::filter(RM_id %in% common_sine_RMs) %>% 
  left_join(., H3K27ac_lookup) %>% 
  distinct(Peakid)

H3K9me3_SINE_peakid <- SINE_overlap_df_H3K9me3 %>% 
  dplyr::filter(RM_id %in% common_sine_RMs) %>% 
  left_join(., H3K9me3_lookup) %>% 
  distinct(Peakid)
  
```

I struggled with figuring out whether a TE is unique to a peak.
I do have some SINE elements mapping to multiple peaks, but not a whole lot.


```{r}
SINE_overlap_df_H3K9me3 %>%
  count(RM_id, name = "n_peaks") %>%
  count(n_peaks)

SINE_overlap_df_H3K9me3 %>%
  count(RM_id) %>%
  summarise(
    min = min(n),
    median = median(n),
    mean = mean(n),
    max = max(n)
  )

SINE_overlap_df_H3K27ac %>%
  count(RM_id, name = "n_peaks") %>%
  count(n_peaks)

SINE_overlap_df_H3K27ac %>%
  count(RM_id) %>%
  summarise(
    min = min(n),
    median = median(n),
    mean = mean(n),
    max = max(n)
  )
```


Because of this many to many situation, I chose to average the log fold changes of the SINE elements overlapping multiple peaks.
```{r}
### linking shared RM_id to Peakid by histone ROI for LFC plotting
##

SINE_K27ac_rmid_lfc <- data_frame(RM_id=common_sine_RMs) %>% 
  left_join(.,(SINE_overlap_df_H3K27ac %>% 
    dplyr::filter(RM_id %in% common_sine_RMs) %>% 
    distinct(RM_id,Peakid, .keep_all = TRUE) %>% 
    mutate(K27ac_Peakid=Peakid) %>% 
    dplyr::select(RM_id, K27ac_Peakid))) %>% 
  left_join(K27ac_lfctable, by =c("K27ac_Peakid"="genes")) %>% 
  group_by(RM_id) %>% 
  summarize(K27ac_24T=mean(H3K27ac_24T, na.rm=TRUE),
             K27ac_24R=mean(H3K27ac_24R,na.rm=TRUE),
             K27ac_144R=mean(H3K27ac_144R, na.rm=TRUE),
            n_K27_peaks = dplyr::n(),
            .groups="drop")

SINE_K9me3_rmid_lfc <- data_frame(RM_id=common_sine_RMs) %>% 
  left_join(.,(SINE_overlap_df_H3K9me3 %>% 
    dplyr::filter(RM_id %in% common_sine_RMs) %>% 
    distinct(RM_id,Peakid, .keep_all = TRUE) %>% 
    mutate(K9me3_Peakid=Peakid) %>% 
    dplyr::select(RM_id, K9me3_Peakid))) %>% 
  left_join(K9me3_lfctable, by =c("K9me3_Peakid"="genes")) %>% 
  group_by(RM_id) %>% 
  summarize(K9me3_24T=mean(H3K9me3_24T, na.rm=TRUE),
             K9me3_24R=mean(H3K9me3_24R,na.rm=TRUE),
             K9me3_144R=mean(H3K9me3_144R, na.rm=TRUE),
            n_K27_peaks = dplyr::n(),
            .groups="drop")



SINE_RM_lfc_concordance <- tibble(RM_id = common_sine_RMs) %>%
  left_join(SINE_K27ac_rmid_lfc, by = "RM_id") %>%
  left_join(SINE_K9me3_rmid_lfc,  by = "RM_id")

```


Now to try to visualize this concordance.  I am specifically looking at TEs that up in one and down in another.


```{r SINE K9K27ac plot}
SINE_RM_lfc_concordance %>% 
  ggplot(., aes(x=K27ac_24T, y=K9me3_24T))+
  geom_point(, alpha=0.2, size=1)+
  geom_density_2d(color= "black")+
  sm_statCorr(corr_method="spearman")+
   geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype = "dashed", color="red") +
  labs(
    x = "H3K27ac logFC (24T)",
    y = "H3K9me3 logFC (24T)",
    title = "Concordance of chromatin changes at shared SINEs for 24T"
  ) +
  theme_classic()

SINE_RM_lfc_concordance %>% 
  ggplot(., aes(x=K27ac_24R, y=K9me3_24R))+
  geom_point(, alpha=0.2, size=1)+
  geom_density_2d(color= "black")+
   sm_statCorr(corr_method="spearman")+
   geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype = "dashed", color="red") +
  labs(
    x = "H3K27ac logFC (24R)",
    y = "H3K9me3 logFC (24R)",
    title = "Concordance of chromatin changes at shared SINEs for 24R"
  ) +
  theme_classic()


SINE_RM_lfc_concordance %>% 
  ggplot(., aes(x=K27ac_144R, y=K9me3_144R))+
  geom_point(, alpha=0.2, size=1)+
  geom_density_2d(color= "black")+
   sm_statCorr(corr_method="spearman")+
   geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype = "dashed", color="red") +
  labs(
    x = "H3K27ac logFC (144R)",
    y = "H3K9me3 logFC (144R)",
    title = "Concordance of chromatin changes at shared SINEs for 144R"
  ) +
  theme_classic()

```


### LTR
Overlapping LTR family with ROIs

```{r LTR Overlapping}
LTR_hits_H3K9me3 <- findOverlaps(H3K9me3_sets_gr$all_H3K9me3_regions, LTR_gr, ignore.strand = TRUE)

LTR_overlap_df_H3K9me3 <- tibble(
  peak_row = queryHits(LTR_hits_H3K9me3),
  Peakid   = H3K9me3_sets_gr$all_H3K9me3$Peakid[queryHits(LTR_hits_H3K9me3)],
  cluster  = H3K9me3_sets_gr$all_H3K9me3$cluster[queryHits(LTR_hits_H3K9me3)],
  
  repClass = LTR_gr$repClass[subjectHits(LTR_hits_H3K9me3)],
  repName  = LTR_gr$repName[subjectHits(LTR_hits_H3K9me3)],
  RM_id  = LTR_gr$RM_id[subjectHits(LTR_hits_H3K9me3)],
  
  TE_type  = ifelse(
    LTR_gr$repFamily[subjectHits(LTR_hits_H3K9me3)] == "SVA",
    LTR_gr$repName[subjectHits(LTR_hits_H3K9me3)],
    LTR_gr$repFamily[subjectHits(LTR_hits_H3K9me3)]
  ),
  
  milliDiv = LTR_gr$milliDiv[subjectHits(LTR_hits_H3K9me3)],
  milliDel = LTR_gr$milliDel[subjectHits(LTR_hits_H3K9me3)],
  milliIns = LTR_gr$milliIns[subjectHits(LTR_hits_H3K9me3)]
)

LTR_hits_H3K27ac <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, LTR_gr, ignore.strand = TRUE)

LTR_overlap_df_H3K27ac <- tibble(
  peak_row = queryHits(LTR_hits_H3K27ac),
  Peakid   = H3K27ac_sets_gr$all_H3K27ac$Peakid[queryHits(LTR_hits_H3K27ac)],
  cluster  = H3K27ac_sets_gr$all_H3K27ac$cluster[queryHits(LTR_hits_H3K27ac)],
  
  repClass = LTR_gr$repClass[subjectHits(LTR_hits_H3K27ac)],
  repName  = LTR_gr$repName[subjectHits(LTR_hits_H3K27ac)],
  RM_id  = LTR_gr$RM_id[subjectHits(LTR_hits_H3K27ac)],
  
  TE_type  = ifelse(
    LTR_gr$repFamily[subjectHits(LTR_hits_H3K27ac)] == "SVA",
    LTR_gr$repName[subjectHits(LTR_hits_H3K27ac)],
    LTR_gr$repFamily[subjectHits(LTR_hits_H3K27ac)]
  ),
  
  milliDiv = LTR_gr$milliDiv[subjectHits(LTR_hits_H3K27ac)],
  milliDel = LTR_gr$milliDel[subjectHits(LTR_hits_H3K27ac)],
  milliIns = LTR_gr$milliIns[subjectHits(LTR_hits_H3K27ac)]
)

common_LTR_RMs <- intersect(
  LTR_overlap_df_H3K27ac$RM_id,
  LTR_overlap_df_H3K9me3$RM_id
)

length(unique(common_LTR_RMs))


```

Visualizing the overlapping LTR families

```{r}
H3K27ac_LTR_peakid <- LTR_overlap_df_H3K27ac %>% 
  dplyr::filter(RM_id %in% common_LTR_RMs) %>% 
  left_join(., H3K27ac_lookup) %>% 
  distinct(Peakid)

H3K9me3_LTR_peakid <- LTR_overlap_df_H3K9me3 %>% 
  dplyr::filter(RM_id %in% common_LTR_RMs) %>% 
  left_join(., H3K9me3_lookup) %>% 
  distinct(Peakid)
  

```

I struggled with figuring out whether a TE is unique to a peak.
I do have some LTR elements mapping to multiple peaks, and more than SINEs most likely due to the increased TE lengths.


```{r}
LTR_overlap_df_H3K9me3 %>%
  count(RM_id, name = "n_peaks") %>%
  count(n_peaks)

LTR_overlap_df_H3K9me3 %>%
  count(RM_id) %>%
  summarise(
    min = min(n),
    median = median(n),
    mean = mean(n),
    max = max(n)
  )

LTR_overlap_df_H3K27ac %>%
  count(RM_id, name = "n_peaks") %>%
  count(n_peaks)

LTR_overlap_df_H3K27ac %>%
  count(RM_id) %>%
  summarise(
    min = min(n),
    median = median(n),
    mean = mean(n),
    max = max(n)
  )
```


Because of this many to many situation, I chose to average the log fold changes of the LTR elements overlapping multiple peaks.
```{r}
### linking shared RM_id to Peakid by histone ROI for LFC plotting
##

K27ac_rmid_lfc <- data_frame(RM_id=common_LTR_RMs) %>%
  left_join(.,(LTR_overlap_df_H3K27ac %>% 
    dplyr::filter(RM_id %in% common_LTR_RMs) %>% 
    distinct(RM_id,Peakid, .keep_all = TRUE))) %>% 
    mutate(K27ac_Peakid=Peakid) %>% 
    dplyr::select(RM_id, K27ac_Peakid) %>% 
  left_join(K27ac_lfctable, by =c("K27ac_Peakid"="genes")) %>%
  left_join(., H3K27ac_lookup, by =c("K27ac_Peakid"="Peakid")) %>% 
  group_by(RM_id) %>% 
  summarize(K27ac_24T=mean(H3K27ac_24T, na.rm=TRUE),
             K27ac_24R=mean(H3K27ac_24R,na.rm=TRUE),
             K27ac_144R=mean(H3K27ac_144R, na.rm=TRUE),
            n_K27ac_peaks = dplyr::n(),
            cluster=dplyr::first(cluster),
            .groups="drop") %>% 
   tidyr::replace_na(list(cluster = "NA"))

K9me3_rmid_lfc <- data_frame(RM_id=common_LTR_RMs) %>% 
  left_join(.,(LTR_overlap_df_H3K9me3 %>% 
    dplyr::filter(RM_id %in% common_LTR_RMs) %>% 
    distinct(RM_id,Peakid, .keep_all = TRUE) %>% 
    mutate(K9me3_Peakid=Peakid) %>% 
    dplyr::select(RM_id, K9me3_Peakid))) %>% 
  left_join(K9me3_lfctable, by =c("K9me3_Peakid"="genes")) %>% 
  group_by(RM_id) %>% 
  summarize(K9me3_24T=mean(H3K9me3_24T, na.rm=TRUE),
             K9me3_24R=mean(H3K9me3_24R,na.rm=TRUE),
             K9me3_144R=mean(H3K9me3_144R, na.rm=TRUE),
            n_K9me3_peaks = dplyr::n(),
            .groups="drop")



LTR_RM_lfc_concordance <- tibble(RM_id = common_LTR_RMs) %>%
  left_join(K27ac_rmid_lfc, by = "RM_id") %>%
  left_join(K9me3_rmid_lfc,  by = "RM_id")

```


Now to try to visualize this concordance.  I am specifically looking at TEs that up in one and down in another.


```{r LTR K9K27ac plot}
LTR_RM_lfc_concordance %>% 
  ggplot(., aes(x=K27ac_24T, y=K9me3_24T, color=cluster))+
  geom_point(, alpha=0.2, size=2)+
  geom_density_2d(color= "black")+
   # sm_statCorr(corr_method="spearman")+
   geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype = "dashed", color="red") +
  labs(
    x = "H3K27ac logFC (24T)",
    y = "H3K9me3 logFC (24T)",
    title = "Concordance of chromatin changes at shared LTRs for 24T"
  ) +
  theme_classic()

LTR_RM_lfc_concordance %>% 
  ggplot(., aes(x=K27ac_24R, y=K9me3_24R, color=cluster))+
  geom_point(, alpha=0.2, size=2)+
  geom_density_2d(color= "black")+
   # sm_statCorr(corr_method="spearman")+
   geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype = "dashed", color="red") +
  labs(
    x = "H3K27ac logFC (24R)",
    y = "H3K9me3 logFC (24R)",
    title = "Concordance of chromatin changes at shared LTRs for 24R"
  ) +
  theme_classic()


LTR_RM_lfc_concordance %>% 
  ggplot(., aes(x=K27ac_144R, y=K9me3_144R, color=cluster))+
  geom_point(, alpha=0.2, size=2)+
  geom_density_2d(color= "black")+
     # sm_statCorr(corr_method="spearman")+
  
   geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype = "dashed", color="red") +
  labs(
    x = "H3K27ac logFC (144R)",
    y = "H3K9me3 logFC (144R)",
    title = "Concordance of chromatin changes at shared LTRs for 144R"
  ) +
  theme_classic()

LTR_RM_lfc_concordance %>% 
  ggplot(., aes(x=K27ac_24T, y=K27ac_24R, color=cluster))+
  geom_point(, alpha=0.2, size=2)+
  geom_density_2d(color= "black")+
   # sm_statCorr(corr_method="spearman")+
   geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype = "dashed", color="red") +
  labs(
    x = "H3K27ac logFC (24T)",
    y = "H3K9me3 logFC (24T)",
    title = "Concordance of chromatin changes at shared LTRs for 24T v 24R H3K27ac"
  ) +
  theme_classic()

LTR_RM_lfc_concordance %>% 
  ggplot(., aes(x=K27ac_24T, y=K27ac_144R, color=cluster))+
  geom_point(, alpha=0.2, size=2)+
  geom_density_2d(color= "black")+
   # sm_statCorr(corr_method="spearman")+
   geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype = "dashed", color="red") +
  labs(
    x = "H3K27ac logFC (24T)",
    y = "H3K9me3 logFC (24T)",
    title = "Concordance of chromatin changes at shared LTRs for 24T v 144R H3K27ac"
  ) +
  theme_classic()

LTR_RM_lfc_concordance %>% 
  ggplot(., aes(x=K27ac_24R, y=K27ac_144R, color=cluster))+
  geom_point(, alpha=0.2, size=2)+
  geom_density_2d(color= "black")+
   # sm_statCorr(corr_method="spearman")+
   geom_hline(yintercept = 0, linetype = "dashed", color="red") +
  geom_vline(xintercept = 0, linetype = "dashed", color="red") +
  labs(
    x = "H3K27ac logFC (24T)",
    y = "H3K9me3 logFC (24T)",
    title = "Concordance of chromatin changes at shared LTRs for 24R v 144R H3K27ac"
  ) +
  theme_classic()
```



```{r}

K27ac_unfiltered_long <- K27ac_lfctable %>% 
  left_join(H3K27ac_lookup, by = c("genes" = "Peakid")) %>% 
  tidyr::replace_na(list(cluster = "NA")) %>% 
  dplyr::filter(cluster != "NA") %>% 
  pivot_longer(
    cols = c("H3K27ac_24T", "H3K27ac_24R", "H3K27ac_144R"),
    names_to = "group",
    values_to = "LFC") %>% 
  mutate(group = factor(group, levels = c("H3K27ac_24T", "H3K27ac_24R", "H3K27ac_144R")))

K9me3_unfiltered_long <- K9me3_lfctable %>% 
  left_join(H3K9me3_lookup, by = c("genes" = "Peakid")) %>% 
  tidyr::replace_na(list(cluster = "NA")) %>% 
  dplyr::filter(cluster != "NA") %>% 
  pivot_longer(
    cols = c("H3K9me3_24T", "H3K9me3_24R", "H3K9me3_144R"),
    names_to = "group",
    values_to = "LFC") %>% 
  mutate(group = factor(group, levels = c("H3K9me3_24T", "H3K9me3_24R", "H3K9me3_144R")))
  
K27ac_median_unfilt_lfc <- K27ac_unfiltered_long %>%
  group_by(cluster, group) %>%
  summarise(median_abs_LFC = median(abs(LFC), na.rm = TRUE),
            .groups = "drop")%>% 
  mutate(time=str_remove(group,"H3K27ac"))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R")))

K9me3_median_unfilt_lfc <- K9me3_unfiltered_long %>%
  group_by(cluster, group) %>%
  summarise(median_abs_LFC = median(abs(LFC), na.rm = TRUE),
            .groups = "drop")%>% 
  mutate(time=str_remove(group,"H3K9me3")) %>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R")))

K9me3_median_unfilt_lfc %>% 
  ggplot(., aes(x=group, y=median_abs_LFC, group=cluster, color=cluster))+
  geom_point(size=4)+
  geom_line(aes(alpha = 0.8, linewidth = 4))+
  theme_bw()+
  ggtitle("LFC across time for H3K27ac sets")


K27ac_median_LTR_lfc <- K27ac_unfiltered_long %>%
  dplyr::filter(genes %in% H3K27ac_LTR_peakid$Peakid) %>% 
  group_by(cluster, group) %>%
  summarise(median_abs_LFC = median(abs(LFC), na.rm = TRUE),
            .groups = "drop") %>% 
  mutate(time=str_remove(group,"H3K27ac"))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R")))
K9me3_median_LTR_lfc <- K9me3_unfiltered_long %>%
  dplyr::filter(genes %in% H3K9me3_LTR_peakid$Peakid) %>% 
  group_by(cluster, group) %>%
  summarise(median_abs_LFC = median(abs(LFC), na.rm = TRUE),
            .groups = "drop") %>% 
  mutate(time=str_remove(group,"H3K9me3"))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R")))

K9me3_median_unfilt_lfc %>% 
ggplot(., aes(x=time, y=median_abs_LFC, group=cluster, color=cluster))+
  geom_point(size=4)+
  geom_line(aes(alpha = 0.8, linewidth = 4))+
  
  geom_line(data=K27ac_median_unfilt_lfc, aes(x=time,y=median_abs_LFC, group=cluster,color=cluster), linetype=2, size = 1)+
   geom_point(data=K27ac_median_unfilt_lfc, aes(x=time,y=median_abs_LFC, group=cluster,color=cluster), linetype=2, size = 1)+
  theme_bw()+
  ggtitle("H3K9me3 unfiltered solid line\nH3K27ac dotted line absolute LFC All regions")

K9me3_median_LTR_lfc %>% 
ggplot(., aes(x=time, y=median_abs_LFC, group=cluster, color=cluster))+
  geom_point(size=4)+
  geom_line(aes(alpha = 0.8, linewidth = 4))+
  
  geom_line(data=K27ac_median_LTR_lfc, aes(x=time,y=median_abs_LFC, group=cluster,color=cluster), linetype=2, size = 1)+
   geom_point(data=K27ac_median_LTR_lfc, aes(x=time,y=median_abs_LFC, group=cluster,color=cluster), linetype=2, size = 1)+
  theme_bw()+
  ggtitle("H3K9me3 solid line\nH3K27ac dotted line absolute LFC overlapping LTR regions")



```  

Now I am trying to break down by LTR families:

```{r ERV1}
LTR_overlap_df_H3K27ac %>% 
  group_by(TE_type) %>% 
  tally() %>% 
   ungroup() %>% 
  DT::datatable(
    rownames = FALSE,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      "H3K27ac LTR specific overlap counts by TE familiy"),
    options = list(pageLength = 10,
      autoWidth = TRUE,
      dom = "tip"))

# now I want to breakup common_LTR_RMs and plot by family of LTR.
 


K27_family_plot <- function(lfc_df,
                            split_list, 
                            grp="K27ac",
                            repFamily,
                            repName){

  rm_ids <- split_list[[repName]]$RM_id
  
  lfc_df %>% 
  dplyr::filter(RM_id %in% rm_ids) %>%
  dplyr::filter(cluster != "NA") %>% 
  pivot_longer(.,cols=c(starts_with(grp)), names_to="group", values_to = "LFC") %>% 
  group_by(cluster, group) %>%
  summarise(median_abs_LFC = median(abs(LFC), na.rm = TRUE),
            .groups = "drop")%>% 
  mutate(time=str_remove(group,grp))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R"))) %>% 
  ggplot(., aes(x=time, y=median_abs_LFC, group=cluster, color=cluster))+
  geom_point(size=4)+
  geom_line(aes(alpha = 0.8, linewidth = 4))+
  theme_bw()+
  ggtitle(paste0(grp," absolute LFC overlapping ",repFamily,":",repName))
}
```

```{r}
names(LTR_split_df)
K27_family_plot(K27ac_rmid_lfc, LTR_split_df,"K27ac","LTR","ERV1")
K27_family_plot(K27ac_rmid_lfc, LTR_split_df,"K27ac","LTR","ERVK")
K27_family_plot(K27ac_rmid_lfc, LTR_split_df,"K27ac","LTR","ERVL")
K27_family_plot(K27ac_rmid_lfc, LTR_split_df,"K27ac","LTR","ERVL-MaLR")
K27_family_plot(K27ac_rmid_lfc, LTR_split_df,"K27ac","LTR","Gypsy")
K27_family_plot(K27ac_rmid_lfc, LTR_split_df,"K27ac","LTR","LTR")


```


```{r function to map H3K9me3 over H3K27ac }



K27_K9_family_plot <- function(lfc_df,
                            split_list, 
                            marks=c("K27ac","K9me3"),
                            repClass,
                            repFam){

  
  rm_ids <- split_list[[repFam]]$RM_id
  
  lfc_df %>% 
  dplyr::filter(RM_id %in% rm_ids,
                cluster != "NA") %>% 
  pivot_longer(.,cols=tidyselect::matches(paste0("^(",paste(marks,collapse = "|"), ")")),
               names_to="group", 
               values_to = "LFC") %>% 
     mutate(mark = stringr::str_extract(group, paste(marks, collapse="|")),
      time = stringr::str_remove(group, paste(marks, collapse="|")),
      time = factor(time, levels = c("_24T", "_24R", "_144R"))) %>% 
  group_by(cluster, mark ,time) %>%
  summarise(median_abs_LFC = median(abs(LFC), na.rm = TRUE),
            .groups = "drop")%>% 
  ggplot(., aes(x=time, y=median_abs_LFC, group=interaction(cluster,mark), color=cluster, linetype=mark, shape=mark))+
  geom_point(size=4)+
  geom_line(aes(alpha = 0.8, linewidth = 4))+
  theme_bw()+
  labs(title = paste0("Both histones, absolute LFC overlapping ",repClass,":",repFam),
      y = "Median |LFC|",
      x = "Time"
    )
}

```
```{r}
K27_K9_family_plot(LTR_RM_lfc_concordance, LTR_split_df,marks=c("K27ac","K9me3"), "LTR",repFam="ERVL")

K27_K9_family_plot(LTR_RM_lfc_concordance, LTR_split_df,marks=c("K27ac","K9me3"), "LTR",repFam="ERVK")
K27_K9_family_plot(LTR_RM_lfc_concordance, LTR_split_df,marks=c("K27ac","K9me3"), "LTR",repFam="ERV1")
K27_K9_family_plot(LTR_RM_lfc_concordance, LTR_split_df,marks=c("K27ac","K9me3"), "LTR",repFam="ERVL-MaLR")
K27_K9_family_plot(LTR_RM_lfc_concordance, LTR_split_df,marks=c("K27ac","K9me3"), "LTR",repFam="Gypsy")
K27_K9_family_plot(LTR_RM_lfc_concordance, LTR_split_df,marks=c("K27ac","K9me3"), "LTR",repFam="LTR")
```

