---
title: "GOanalysis_TFs"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```


```{r packages}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(DT)
library(gprofiler2)
library(kableExtra)
library(rrvgo)
library(org.Hs.eg.db)

```

```{r loading data H3K27ac}

H3K27ac_Set2_data <- readRDS("data/RDS_files/H3K27ac_Set2_enrmotif_data.RDS")
H3K27ac_Set3_data <- readRDS("data/RDS_files/H3K27ac_Set3_enrmotif_data.RDS")

ggVennDiagram::ggVennDiagram(list("K27ac_2"=H3K27ac_Set2_data$motif_name,"K27ac_3"=H3K27ac_Set3_data$motif_name))+
  ggtitle("Shared vs unique TF motifs enriched in Set 2 and Set 3 H3K27ac")
K27ac_acute <- setdiff(H3K27ac_Set3_data$motif_name,H3K27ac_Set2_data$motif_name)
K27ac_sus <- setdiff(H3K27ac_Set2_data$motif_name,H3K27ac_Set3_data$motif_name)
```

I am looking at GO enrichment analysis.  I will take the enriched motifs found the MEME suite and do a few strategies of analysis.

Strat 1:  taking list of enriched motifs names, breaking down pairs like FOS::JUNB into FOS and JUNB and doing GO analysis compared to all genes.
-maybe filter for expression with RNA data?

Strat 2: Using GOMO analysis.  this takes the enriched meme file of set 2 v set 1 and set 3 v set 1 and searches for the genes they regulate then does GO MF BP and CC for all the genes associated with being activated by that motif.

Strat 3:  Overlapping Set 2 and Set 3, go analysis on unique TFs
### H3K27ac  


```{r strat1}
RNA_full_toptable <-  readRDS("data/Other_paper_data/RNA_full_toptable.RDS")
RNA_lookup <- RNA_full_toptable %>% 
  dplyr::select(Entrez_ID, SYMBOL) %>% 
  distinct()

motif_ID_set2H3K27ac <- H3K27ac_Set2_data %>% 
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>% 
  pull(ID) 
  
# writeLines(motif_ID_set2H3K27ac,"data/motif_lists/motif_ID_set2H3K27ac.txt")

motif_names_set2H3K27ac <- H3K27ac_Set2_data %>%
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>%  
  dplyr::select(motif_name) %>% 
  separate_longer_delim(motif_name, delim="::") %>% 
  mutate(motif_name2=toupper(motif_name)) %>% 
  pull(motif_name2) 

motif_ID_set3H3K27ac <- H3K27ac_Set3_data %>% 
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>% 
  pull(ID) 

# writeLines(motif_ID_set3H3K27ac,"data/motif_lists/motif_ID_set3H3K27ac.txt")


motif_names_set3H3K27ac <- H3K27ac_Set3_data %>%
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>%  
  dplyr::select(motif_name) %>% 
  separate_longer_delim(motif_name, delim="::") %>% 
  mutate(motif_name2=toupper(motif_name)) %>% 
  pull(motif_name2) 

```

```{r}


Set2_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% motif_names_set2H3K27ac)

Set3_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% motif_names_set3H3K27ac)

```

```{r, eval=FALSE}
H3K27ac_Set2_gost <- gost(query = Set2_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

H3K27ac_Set3_gost <- gost(query = Set3_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

saveRDS(H3K27ac_Set2_gost,"data/RDS_files/H3K27ac_Set2_gost.RDS")
saveRDS(H3K27ac_Set3_gost,"data/RDS_files/H3K27ac_Set3_gost.RDS")
```

```{r}
H3K27ac_Set2_gost <- readRDS("data/RDS_files/H3K27ac_Set2_gost.RDS")
H3K27ac_Set3_gost <- readRDS("data/RDS_files/H3K27ac_Set3_gost.RDS")

gostplot(H3K27ac_Set2_gost)
gostplot(H3K27ac_Set3_gost)
```

```{r function to map }

GO_KEGG_analysis <- function(df,type, setp_value=1, top_n= NULL){
 a <- deparse(substitute(df)) %>% 
  gsub("_gost","",.)

# Prepare the table output
  filtered_df <-  df$result %>% 
  dplyr::select(c(source, term_id,
                  term_name,intersection_size, 
                   term_size, p_value)) %>% 
    dplyr::filter(source == type) %>%
    dplyr::filter(p_value < setp_value) %>%
    arrange(p_value)
  
  # Optional top_n slicing
  if(!is.null(top_n)){
    filtered_df <- filtered_df %>%
      slice_head(n = top_n)
  }

table <- filtered_df %>%
  mutate_at(.vars = 6, .funs = scales::scientific_format()) %>%
  kable(.,caption = paste0("Significant (p < ",setp_value," ) ",type," terms found in ",a," enriched TFs")) %>%
  kable_paper("striped", full_width = FALSE) %>%
  kable_styling(
    full_width = FALSE,
    position = "left",
    bootstrap_options = c("striped", "hover")
  ) %>%
  scroll_box(width = "100%", height = "400px")


# plot the table as ggplot object
plot <- filtered_df %>%
   # slice_head(n=10) %>% 
    dplyr::filter(p_value < setp_value) %>% 
  mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=log10_p, y=reorder(term_name, log10_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("(p < ",setp_value," ) ",type," terms found in\n ",a," enriched TFs"))

print(table)
print(plot)
}



```


```{r}
GO_KEGG_analysis(H3K27ac_Set2_gost,"GO:BP", 0.01,20)

GO_KEGG_analysis(H3K27ac_Set3_gost,"GO:BP", 0.01,20)


```

```{r}

test_df <- H3K27ac_Set2_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<1000)
go_terms <- test_df$term_id
scores <- -log10(test_df$p_value)
names(scores) <- go_terms

simMatrix <- readRDS("data/RDS_files/simMatrix_H3K27ac_set2.RDS")
# simMatrix <- calculateSimMatrix(
#   go_terms,
#   orgdb = org.Hs.eg.db,
#   ont = "BP")

reducedTerms <- reduceSimMatrix(
  simMatrix,
  scores,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms)
scatterPlot(simMatrix, reducedTerms)
head(reducedTerms)

reduced_df <- reducedTerms %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_df %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>% 
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K27ac Set 2 enriched TFs"))


```

```{r}


test_df_3 <- H3K27ac_Set3_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<1000)
go_terms_3 <- test_df_3$term_id
scores_3 <- -log10(test_df_3$p_value)
names(scores_3) <- go_terms_3

simMatrix_3 <- readRDS("data/RDS_files/simMatrix_H3K27ac_set3.RDS")
# simMatrix_3 <- calculateSimMatrix(
#   go_terms_3,
#   orgdb = org.Hs.eg.db,
#   ont = "BP")

reducedTerms_3 <- reduceSimMatrix(
  simMatrix_3,
  scores_3,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms_3)
scatterPlot(simMatrix_3, reducedTerms_3)
head(reducedTerms_3)

reduced_df_3 <- reducedTerms_3 %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_df_3 %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>% 
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K27ac Set 3 enriched TFs"))


```

#### Strat3 H3K27ac  

Separating enriched TFs by sus only; acute only; and shared.
```{r}
K27ac_acute <- setdiff(toupper(H3K27ac_Set3_data$motif_name),toupper(H3K27ac_Set2_data$motif_name))

K27ac_sus <- setdiff(toupper(H3K27ac_Set2_data$motif_name),toupper(H3K27ac_Set3_data$motif_name))
K27ac_shared <- intersect(toupper(H3K27ac_Set2_data$motif_name),toupper(H3K27ac_Set3_data$motif_name))

K27ac_acute_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% (tibble(K27ac_acute) %>% separate_longer_delim(K27ac_acute, delim="::") %>% 
                               pull(K27ac_acute)))

K27ac_sus_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% (tibble(K27ac_sus) %>% separate_longer_delim(K27ac_sus, delim="::") %>% 
                               pull(K27ac_sus)))

K27ac_shared_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% (tibble(K27ac_shared) %>% separate_longer_delim(K27ac_shared, delim="::") %>% 
                               pull(K27ac_shared)))

```

```{r gost H3K27ac strat3, eval=FALSE}
H3K27ac_acute_gost <- gost(query = K27ac_acute_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

H3K27ac_sus_gost <- gost(query = K27ac_sus_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

H3K27ac_shared_gost <- gost(query = K27ac_shared_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

saveRDS(H3K27ac_acute_gost,"data/RDS_files/H3K27ac_acute_gost.RDS")
saveRDS(H3K27ac_sus_gost,"data/RDS_files/H3K27ac_sus_gost.RDS")
saveRDS(H3K27ac_shared_gost,"data/RDS_files/H3K27ac_shared_gost.RDS")
```


```{r}
H3K27ac_acute_gost <- readRDS("data/RDS_files/H3K27ac_acute_gost.RDS")
H3K27ac_sus_gost <- readRDS("data/RDS_files/H3K27ac_sus_gost.RDS")
H3K27ac_shared_gost <- readRDS("data/RDS_files/H3K27ac_shared_gost.RDS")


GO_KEGG_analysis(H3K27ac_acute_gost,"GO:BP", 0.05,20)

GO_KEGG_analysis(H3K27ac_acute_gost,"KEGG", 0.05,20)

H3K27ac_acute_gost$result %>% 
  dplyr::filter(p_value< 0.05)

```

```{r}
GO_KEGG_analysis(H3K27ac_sus_gost,"GO:BP", 0.05,20)

GO_KEGG_analysis(H3K27ac_sus_gost,"KEGG", 0.05,20)
```
```{r}
GO_KEGG_analysis(H3K27ac_shared_gost,"GO:BP", 0.05,20)

GO_KEGG_analysis(H3K27ac_shared_gost,"KEGG", 0.05,20)
```





### H3K9me3

```{r loading data H3K9me3}

H3K9me3_Set2_data <- readRDS("data/RDS_files/H3K9me3_Set2_enrmotif_data.RDS")
H3K9me3_Set3_data <- readRDS("data/RDS_files/H3K9me3_Set3_enrmotif_data.RDS")

 ggVennDiagram::ggVennDiagram(list("K9me3_2"=H3K9me3_Set2_data$motif_name,"K9me3_3"=H3K9me3_Set3_data$motif_name))+
  ggtitle("Shared vs unique TF motifs enriched in Set 2 and Set 3 H3K9me3")


```


```{r strat1 set2}
# RNA_full_toptable <-  readRDS("data/Other_paper_data/RNA_full_toptable.RDS")
# RNA_lookup <- RNA_full_toptable %>% 
#   dplyr::select(Entrez_ID, SYMBOL) %>% 
#   distinct()

motif_ID_set2H3K9me3 <- H3K9me3_Set2_data %>% 
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>% 
  pull(ID) 
  
# writeLines(motif_ID_set2H3K9me3,"data/motif_lists/motif_ID_set2H3K9me3.txt")

motif_names_set2H3K9me3 <- H3K9me3_Set2_data %>%
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>%  
  dplyr::select(motif_name) %>% 
  separate_longer_delim(motif_name, delim="::") %>% 
  mutate(motif_name2=toupper(motif_name)) %>% 
  pull(motif_name2) 

motif_ID_set3H3K9me3 <- H3K9me3_Set3_data %>% 
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>% 
  pull(ID) 

# writeLines(motif_ID_set3H3K9me3,"data/motif_lists/motif_ID_set3H3K9me3.txt")


motif_names_set3H3K9me3 <- H3K9me3_Set3_data %>%
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>%  
  dplyr::select(motif_name) %>% 
  separate_longer_delim(motif_name, delim="::") %>% 
  mutate(motif_name2=toupper(motif_name)) %>% 
  pull(motif_name2) 

```


```{r}


Set2_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% motif_names_set2H3K9me3)

Set3_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% motif_names_set3H3K9me3)

```

```{r, eval=FALSE}
H3K9me3_Set2_gost <- gost(query = Set2_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

H3K9me3_Set3_gost <- gost(query = Set3_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

saveRDS(H3K9me3_Set2_gost,"data/RDS_files/H3K9me3_Set2_gost.RDS")
saveRDS(H3K9me3_Set3_gost,"data/RDS_files/H3K9me3_Set3_gost.RDS")
```

```{r}
H3K9me3_Set2_gost <- readRDS("data/RDS_files/H3K9me3_Set2_gost.RDS")
H3K9me3_Set3_gost <- readRDS("data/RDS_files/H3K9me3_Set3_gost.RDS")

gostplot(H3K9me3_Set2_gost)
gostplot(H3K9me3_Set3_gost)

```

```{r}
GO_KEGG_analysis(H3K9me3_Set2_gost,"GO:BP", 0.01,20)

GO_KEGG_analysis(H3K9me3_Set3_gost,"GO:BP", 0.01,20)
```

```{r H3K9me3 set 2 redux}

test_df_k9 <- H3K9me3_Set2_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<750)
go_terms_k9  <- test_df_k9 $term_id
scores_k9  <- -log10(test_df_k9 $p_value)
names(scores_k9)  <- go_terms_k9 

simMatrix_k9  <- readRDS("data/RDS_files/simMatrix_H3K9me3_set2.RDS")
  # saveRDS(simMatrix_k9,"data/RDS_files/simMatrix_H3K9me3_set2.RDS")
# simMatrix_k9  <- calculateSimMatrix(
#   go_terms_k9 ,
#   orgdb = org.Hs.eg.db,
#   ont = "BP")

reducedTerms_k9  <- reduceSimMatrix(
  simMatrix_k9 ,
  scores_k9 ,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms_k9 )
scatterPlot(simMatrix_k9 , reducedTerms_k9 )
head(reducedTerms_k9 )

reduced_df_k9  <- reducedTerms_k9  %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_df_k9  %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>% 
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K9me3 Set 2 enriched TFs"))


```

```{r H3K9me3 set3 redux}


test_df_3_k9  <- H3K9me3_Set3_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<750)
go_terms_3_k9  <- test_df_3_k9 $term_id
scores_3_k9  <- -log10(test_df_3_k9 $p_value)
names(scores_3_k9 ) <- go_terms_3_k9 

simMatrix_3_k9  <- readRDS("data/RDS_files/simMatrix_H3K9me3_set3.RDS")
# saveRDS(simMatrix_3_k9,"data/RDS_files/simMatrix_H3K9me3_set3.RDS")
# simMatrix_3_k9  <- calculateSimMatrix(
#   go_terms_3_k9 ,
#   orgdb = org.Hs.eg.db,
#   ont = "BP")

reducedTerms_3_k9  <- reduceSimMatrix(
  simMatrix_3_k9 ,
  scores_3_k9 ,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms_3_k9 )
scatterPlot(simMatrix_3_k9 , reducedTerms_3_k9 )
head(reducedTerms_3_k9 )

reduced_df_3_k9  <- reducedTerms_3_k9  %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_df_3_k9 %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>% 
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K9me3 Set 3 enriched TFs"))


```


#### Strat3 H3K9me3  

Separating enriched TFs by sus only; acute only; and shared.
```{r}
K9me3_sus <- setdiff(toupper(H3K9me3_Set3_data$motif_name),toupper(H3K9me3_Set2_data$motif_name))

K9me3_acute <- setdiff(toupper(H3K9me3_Set2_data$motif_name),toupper(H3K9me3_Set3_data$motif_name))
K9me3_shared <- intersect(toupper(H3K9me3_Set2_data$motif_name),toupper(H3K9me3_Set3_data$motif_name))

K9me3_acute_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% (tibble(K9me3_acute) %>% separate_longer_delim(K9me3_acute, delim="::") %>% 
                               pull(K9me3_acute)))

K9me3_sus_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% (tibble(K9me3_sus) %>% separate_longer_delim(K9me3_sus, delim="::") %>% 
                               pull(K9me3_sus)))

K9me3_shared_list <- RNA_lookup %>% 
  dplyr::filter(SYMBOL %in% (tibble(K9me3_shared) %>% separate_longer_delim(K9me3_shared, delim="::") %>% 
                               pull(K9me3_shared)))

```

```{r gost H3K9me3 strat3, eval=FALSE}
H3K9me3_acute_gost <- gost(query = K9me3_acute_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

H3K9me3_sus_gost <- gost(query = K9me3_sus_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

H3K9me3_shared_gost <- gost(query = K9me3_shared_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

saveRDS(H3K9me3_acute_gost,"data/RDS_files/H3K9me3_acute_gost.RDS")
saveRDS(H3K9me3_sus_gost,"data/RDS_files/H3K9me3_sus_gost.RDS")
saveRDS(H3K9me3_shared_gost,"data/RDS_files/H3K9me3_shared_gost.RDS")
```


```{r}
H3K9me3_acute_gost <- readRDS("data/RDS_files/H3K9me3_acute_gost.RDS")
H3K9me3_sus_gost <- readRDS("data/RDS_files/H3K9me3_sus_gost.RDS")
H3K9me3_shared_gost <- readRDS("data/RDS_files/H3K9me3_shared_gost.RDS")


GO_KEGG_analysis(H3K9me3_acute_gost,"GO:BP", 0.05,20)

GO_KEGG_analysis(H3K9me3_acute_gost,"KEGG", 0.05,20)

```

```{r}
GO_KEGG_analysis(H3K9me3_sus_gost,"GO:BP", 0.05,20)

GO_KEGG_analysis(H3K9me3_sus_gost,"KEGG", 0.05,20)
```
```{r}
GO_KEGG_analysis(H3K9me3_shared_gost,"GO:BP", 0.05,20)

GO_KEGG_analysis(H3K9me3_shared_gost,"KEGG", 0.05,20)
```

#### Reduction of terms



##### H3K27ac reduction
```{r}

test_acute_K27 <- H3K27ac_acute_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<750)
go_terms_acute_K27  <- test_acute_K27$term_id
scores_acute_K27  <- -log10(test_acute_K27$p_value)
names(scores_acute_K27)  <- go_terms_acute_K27 

simMatrix_acute_K27  <- readRDS("data/RDS_files/simMatrix_H3K27ac_acute.RDS")
# saveRDS(simMatrix_acute_K27,"data/RDS_files/simMatrix_H3K27ac_acute.RDS")
# simMatrix_acute_K27  <- calculateSimMatrix(
#   go_terms_acute_K27 ,
#   orgdb = org.Hs.eg.db,
#   ont = "BP")

reducedTerms_acute_K27  <- reduceSimMatrix(
  simMatrix_acute_K27 ,
  scores_acute_K27 ,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms_acute_K27 )
scatterPlot(simMatrix_acute_K27 , reducedTerms_acute_K27 )
head(reducedTerms_acute_K27 )

reduced_acute_K27  <- reducedTerms_acute_K27  %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_acute_K27  %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>% 
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K27ac acute only enriched TFs"))


```

```{r}

test_sus_K27 <- H3K27ac_sus_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<750)
go_terms_sus_K27  <- test_sus_K27$term_id
scores_sus_K27  <- -log10(test_sus_K27$p_value)
names(scores_sus_K27)  <- go_terms_sus_K27 

simMatrix_sus_K27  <- readRDS("data/RDS_files/simMatrix_H3K27ac_sus.RDS")
# saveRDS(simMatrix_sus_K27,"data/RDS_files/simMatrix_H3K27ac_sus.RDS")
# simMatrix_sus_K27  <- calculateSimMatrix(
#   go_terms_sus_K27 ,
#   orgdb = org.Hs.eg.db,
#   ont = "BP")

reducedTerms_sus_K27  <- reduceSimMatrix(
  simMatrix_sus_K27 ,
  scores_sus_K27 ,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms_sus_K27 )
scatterPlot(simMatrix_sus_K27 , reducedTerms_sus_K27 )
head(reducedTerms_sus_K27 )

reduced_sus_K27  <- reducedTerms_sus_K27  %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_sus_K27  %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>%
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K27ac sus only enriched TFs"))


```

```{r}

test_shared_K27 <- H3K27ac_shared_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<750)
go_terms_shared_K27  <- test_shared_K27$term_id
scores_shared_K27  <- -log10(test_shared_K27$p_value)
names(scores_shared_K27)  <- go_terms_shared_K27 

simMatrix_shared_K27  <- readRDS("data/RDS_files/simMatrix_H3K27ac_shared.RDS")
# saveRDS(simMatrix_shared_K27,"data/RDS_files/simMatrix_H3K27ac_shared.RDS")
# simMatrix_shared_K27  <- calculateSimMatrix(
# go_terms_shared_K27 ,
# orgdb = org.Hs.eg.db,
# ont = "BP")

reducedTerms_shared_K27  <- reduceSimMatrix(
  simMatrix_shared_K27 ,
  scores_shared_K27 ,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms_shared_K27 )
scatterPlot(simMatrix_shared_K27 , reducedTerms_shared_K27 )
head(reducedTerms_shared_K27 )

reduced_shared_K27  <- reducedTerms_shared_K27  %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_shared_K27  %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>% 
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K27ac shared only enriched TFs"))


```



##### H3K9me3 reduction  

```{r}

test_acute_k9 <- H3K9me3_acute_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<750)
go_terms_acute_k9  <- test_acute_k9$term_id
scores_acute_k9  <- -log10(test_acute_k9$p_value)
names(scores_acute_k9)  <- go_terms_acute_k9 

simMatrix_acute_k9  <- readRDS("data/RDS_files/simMatrix_H3K9me3_acute.RDS")
  # saveRDS(simMatrix_acute_k9,"data/RDS_files/simMatrix_H3K9me3_acute.RDS")
# simMatrix_acute_k9  <- calculateSimMatrix(
#   go_terms_acute_k9 ,
#   orgdb = org.Hs.eg.db,
#   ont = "BP")

reducedTerms_acute_k9  <- reduceSimMatrix(
  simMatrix_acute_k9 ,
  scores_acute_k9 ,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms_acute_k9 )
scatterPlot(simMatrix_acute_k9 , reducedTerms_acute_k9 )
head(reducedTerms_acute_k9 )

reduced_acute_k9  <- reducedTerms_acute_k9  %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_acute_k9  %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>% 
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K9me3 acute only enriched TFs"))


```

```{r fig.width=15}

test_sus_k9 <- H3K9me3_sus_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<750)
go_terms_sus_k9  <- test_sus_k9$term_id
scores_sus_k9  <- -log10(test_sus_k9$p_value)
names(scores_sus_k9)  <- go_terms_sus_k9 

simMatrix_sus_k9  <- readRDS("data/RDS_files/simMatrix_H3K9me3_sus.RDS")
# saveRDS(simMatrix_sus_k9,"data/RDS_files/simMatrix_H3K9me3_sus.RDS")
# simMatrix_sus_k9  <- calculateSimMatrix(
#   go_terms_sus_k9 ,
#   orgdb = org.Hs.eg.db,
#   ont = "BP")

reducedTerms_sus_k9  <- reduceSimMatrix(
  simMatrix_sus_k9 ,
  scores_sus_k9 ,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms_sus_k9 )
scatterPlot(simMatrix_sus_k9 , reducedTerms_sus_k9 )
head(reducedTerms_sus_k9 )

reduced_sus_k9  <- reducedTerms_sus_k9  %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_sus_k9  %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>%
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K9me3 sus only enriched TFs"))


```

```{r}

test_shared_k9 <- H3K9me3_shared_gost$result %>% 
  dplyr::filter(p_value<0.05) %>% 
  dplyr::filter(term_size<750)
go_terms_shared_k9  <- test_shared_k9$term_id
scores_shared_k9  <- -log10(test_shared_k9$p_value)
names(scores_shared_k9)  <- go_terms_shared_k9 

simMatrix_shared_k9  <- readRDS("data/RDS_files/simMatrix_H3K9me3_shared.RDS")
  # saveRDS(simMatrix_shared_k9,"data/RDS_files/simMatrix_H3K9me3_shared.RDS")
# simMatrix_shared_k9  <- calculateSimMatrix(
  # go_terms_shared_k9 ,
  # orgdb = org.Hs.eg.db,
  # ont = "BP")

reducedTerms_shared_k9  <- reduceSimMatrix(
  simMatrix_shared_k9 ,
  scores_shared_k9 ,
  threshold = 0.7,
  orgdb = org.Hs.eg.db)

treemapPlot(reducedTerms_shared_k9 )
scatterPlot(simMatrix_shared_k9 , reducedTerms_shared_k9 )
head(reducedTerms_shared_k9 )

reduced_shared_k9  <- reducedTerms_shared_k9  %>%
  group_by(parent) %>%
  summarise(
    best_p = max(score),   # if score = -log10(p)
    term = dplyr::first(term)
  ) %>%
  ungroup()
reduced_shared_k9  %>% 
  arrange( desc(best_p)) %>% 
  slice_head(n=25) %>% 
# dplyr::filter(p_value < setp_value) %>% 
  # mutate(log10_p=(-log(p_value, base=10))) %>% 
  ggplot(., aes(x=best_p, y=reorder(term, best_p)))+
    geom_col()+
  geom_vline(aes(xintercept=-log(0.05,base = 10)), col="red",linetype = 2)+
  theme_bw()+
  ylab("")+
  xlab(paste0("-log10 p-value"))+
  guides(fill="none")+
  theme(axis.text.y = element_text(color="black"))+
  ggtitle(paste0("H3K9me3 shared only enriched TFs"))


```
### TE specific GO/KEGG

```{r}
MER61E_anno_genes <- readRDS("data/RDS_files/MER61E_anno_genes.RDS")
LTR10E_anno_genes <- readRDS("data/RDS_files/LTR10E_anno_genes.RDS")

Mer_100k_3 <-MER61E_anno_genes %>% 
  dplyr::filter(abs(closest_distance)<100000) %>% 
  dplyr::filter(cluster=="Set_3") %>% 
  pull(ENTREZID)

Mer_all_3 <-MER61E_anno_genes %>% 
  # dplyr::filter(abs(closest_distance)<100000) %>% 
  dplyr::filter(cluster=="Set_3") %>% 
  pull(ENTREZID)

Ltr_100k_3 <-LTR10E_anno_genes %>% 
  dplyr::filter(abs(closest_distance)<100000) %>% 
  dplyr::filter(cluster=="Set_3") %>% 
  pull(ENTREZID)
Ltr_all_3 <-LTR10E_anno_genes %>% 
  # dplyr::filter(abs(closest_distance)<100000) %>% 
  dplyr::filter(cluster=="Set_3") %>% 
  pull(ENTREZID)

```


```{r gost TE, eval=FALSE}
MER61_all_gost <- gost(query = Mer_all_3,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

H3K27ac_Set3_gost <- gost(query = Set3_list$Entrez_ID,
                    organism = "hsapiens",
                    significant = FALSE,
                    ordered_query = FALSE,
                    domain_scope = "custom",
                    measure_underrepresentation = FALSE,
                    evcodes = FALSE,
                    user_threshold = 0.05,
                    correction_method = c("fdr"),
                    custom_bg = RNA_lookup$Entrez_ID,
                    sources=c("GO:BP","KEGG"))

saveRDS(MER61_100k_gost,"data/RDS_files/MER61_100k_gost.RDS")
saveRDS(LTR10_100k_gost,"data/RDS_files/LTR10_100k_gost.RDS")
saveRDS(MER61_all_gost,"data/RDS_files/MER61_all_gost.RDS")
saveRDS(LTR10_all_gost,"data/RDS_files/LTR10_all_gost.RDS")
```

```{r}

MER61_100k_gost <- readRDS("data/RDS_files/MER61_100k_gost.RDS")
LTR10_100k_gost <- readRDS("data/RDS_files/LTR10_100k_gost.RDS")

GO_KEGG_analysis(MER61_100k_gost,"GO:BP", 0.1,20)
GO_KEGG_analysis(MER61_100k_gost,"KEGG", 0.1,20)

GO_KEGG_analysis(LTR10_100k_gost,"GO:BP", 0.1,20)
GO_KEGG_analysis(LTR10_100k_gost,"KEGG", 0.1,20)
```

```{r}

MER61_all_gost <- readRDS("data/RDS_files/MER61_all_gost.RDS")
LTR10_all_gost <- readRDS("data/RDS_files/LTR10_all_gost.RDS")
GO_KEGG_analysis(LTR10_all_gost,"GO:BP", 0.05,20)
GO_KEGG_analysis(LTR10_all_gost,"KEGG", 0.05,20)

GO_KEGG_analysis(MER61_all_gost,"GO:BP", 0.05,20)
GO_KEGG_analysis(MER61_all_gost,"KEGG", 0.05,20)
```

