---
title: "H3K27ac_TEtype"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(data.table)
library(GenomicFeatures)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(ComplexHeatmap)
library(circlize)
library(readxl)
library(kableExtra)
```

```{r data loading}

annotated_H3K27ac_summits <- readRDS("data/TE_annotation/H3K27ac_annotated_summit_TE_overlaps.RDS")


H3K27ac_summit_ols <- readRDS("data/RDS_files/H3K27ac_summit_ROI_TE_overlaps.RDS")

MER61E_ids <- H3K27ac_summit_ols %>% 
  dplyr::filter(repName =="MER61E") %>% 
  dplyr::filter(cluster=="Set_3")

MER61E_peakid <- H3K27ac_summit_ols %>% 
  dplyr::filter(repName=="MER61E") %>% 
  distinct(Peakid)


RNA_full_toptable <-  readRDS("data/Other_paper_data/RNA_full_toptable.RDS")

RNA_filcpm_final_matrix <- read_csv("data/Other_paper_data/RNA_filcpm_final_matrix.csv")
Symbol_lookup <- RNA_full_toptable %>% 
  dplyr::select (Entrez_ID,SYMBOL) %>% 
  distinct(Entrez_ID,SYMBOL)

KZFP_targets <- read_excel("data/Other_paper_data/KZFP_primary_targests_Supplemental_Table_S3.xlsx", sheet = "Primary_targets")

KZFP_2026_list <- read_excel("data/Other_paper_data/human_protein_coding_KZFPs_Davis_natgen_2026.xlsx")

ERV1_summary_char <- read.csv(file.path("data/Other_paper_data/ERV1_file_summary.csv"))

```


### MER61E-ERV1-LTRs  


```{r MER61E elements}

# your LTRs
ltr_gr <- rtracklayer::import("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed")

# gene annotations
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
genes_gr <- genes(txdb)

# find nearest gene (distance to gene body)
hits <- distanceToNearest(ltr_gr, genes_gr)

# extract results
res <- data.frame(
  ltr = mcols(ltr_gr)$name[queryHits(hits)],
  gene = names(genes_gr)[subjectHits(hits)],
  distance = mcols(hits)$distance
)

MER61E_genes <- res %>% distinct(gene)

RNA_expressed_genes <- RNA_full_toptable %>% distinct(Entrez_ID)

summit_overlap_MER61E_genes <- res %>% 
  dplyr::filter(ltr %in% MER61E_ids$ID) %>% 
  distinct(gene)

```


```{r}


RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"), labels = c("tx + 0","tx + 24h","tx + 144h"))) %>% 
  dplyr::filter(Entrez_ID %in% MER61E_genes$gene) %>% 
  ggplot(., aes(x=timepoint, y=logFC))+
  geom_boxplot()+
  ggtitle("LFC at timepoints of genes nearest to all MER61E TE")+
  theme_bw()
  

RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"), labels = c("tx + 0","tx + 24h","tx + 144h"))) %>% 
  # dplyr::filter(Entrez_ID %in% MER61E_genes$gene) %>% 
  ggplot(., aes(x=timepoint, y=logFC))+
  geom_boxplot()+
  ggtitle("LFC at timepoints of all genes")+
  theme_bw()

RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"), labels = c("tx + 0","tx + 24h","tx + 144h"))) %>% 
  dplyr::filter(Entrez_ID %in% summit_overlap_MER61E_genes$gene) %>%
  ggplot(., aes(x=timepoint, y=logFC))+
  geom_boxplot()+
  ggtitle("LFC at timepoints of all genes that are near MER61E overlapping summits")+
  theme_bw()+
  ylim(-5,5)

  
  res %>% dplyr::filter(gene %in% RNA_expressed_genes$Entrez_ID) %>% 
    ggplot(., aes(x=distance))+
    geom_density()+
    ggtitle("density of MER61E TEs distance to expressed gene")
  
  

  res %>% dplyr::filter(gene %in%  
  RNA_expressed_genes$Entrez_ID) %>% 
    ggplot(., aes(x=distance))+
      geom_histogram(bins = 50)+
    ggtitle("density of MER61E TEs distance to expressed gene, bins=50")
```


### RNA expression of MER61E nearby genes  


```{r}
MER61E_lcpm_genes <- RNA_filcpm_final_matrix %>% 
  dplyr::filter(Entrezid %in% MER61E_genes$gene) %>% 
  pivot_longer(cols=!Entrezid, names_to = "sample",values_to = "log2cpm") %>% 
  separate_wider_delim(cols = sample, names = c("trt","time","ind"), delim = "_") %>% 
  mutate(group=paste0(trt,"_",time),
  time=factor(time,levels=c("24T","24R","144R"))) 




for (aim in MER61E_genes$gene){
  df <- MER61E_lcpm_genes %>% 
    dplyr::filter(Entrezid== aim)
  if (nrow(df) == 0) next
  symbol <- Symbol_lookup %>% 
    dplyr::filter(Entrez_ID==aim) %>% 
    dplyr::pull(SYMBOL)
  distance <- res %>%
    dplyr::filter(gene == aim) %>%
    dplyr::pull(distance)
  
plot <- ggplot(df, aes(x=time, y=log2cpm,group=group))+
  geom_boxplot(aes(fill=trt)) +
  ggtitle(paste0("gene ", aim," ",symbol," dist ",distance))+
  theme_bw()
print(plot)
}
```



#### Looking as LFC  


```{r}
RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"))) %>% 
  dplyr::filter(Entrez_ID %in% MER61E_genes$gene) %>% 
  ggplot(., aes(x=timepoint, y=logFC))+
  geom_boxplot()+
  ggtitle("LFC at timepoints of nearest gene to MER61E TE")+
  theme_bw()


RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"))) %>% 
  dplyr::filter(Entrez_ID %in% MER61E_genes$gene) %>% 
  ggplot(., aes(x=timepoint, y=logFC, group = Entrez_ID))+
  geom_point()+
  geom_line()+
  ggtitle("LFC at timepoints of nearest gene to all MER61E TE")
```


```{r}
matrix_file <- ("C:/Users/renee/Other_projects_data/DXR_data/computeMatrix_output_centered2histone/matcenter_MER61E_ERV1_LTR.gz")
bins <- 400
cm <- fread(cmd = paste("gzip -dc", matrix_file), skip = 1)
  
  sample_names <- c(
    "H3K27ac_DOX_144R_merged","H3K27ac_DOX_24R_merged","H3K27ac_DOX_24T_merged",
    "H3K27ac_VEH_144R_merged","H3K27ac_VEH_24R_merged","H3K27ac_VEH_24T_merged",
    "H3K9me3_DOX_144R_merged","H3K9me3_DOX_24R_merged","H3K9me3_DOX_24T_merged",
    "H3K9me3_VEH_144R_merged","H3K9me3_VEH_24R_merged","H3K9me3_VEH_24T_merged")
 
  signal_list <- lapply(seq_along(sample_names), function(i){
    start <- 6 + (i-1)*bins + 1
    end <- 6 + i*bins
    as.matrix(cm[, start:end])})
  
  names(signal_list) <- sample_names
  
   avg_signal <- function(mark, cond) {
    rows <- grep(paste0(mark,"_",cond), names(signal_list))
    # Average across replicates
    Reduce("+", signal_list[rows]) / length(rows)
   }
   
   bed_info <- cm[, 1:6]
  
  H3K27ac_DOX_avg <- avg_signal("H3K27ac", "DOX")
  H3K27ac_VEH_avg <- avg_signal("H3K27ac", "VEH")
  H3K9me3_DOX_avg <- avg_signal("H3K9me3", "DOX")
  H3K9me3_VEH_avg <- avg_signal("H3K9me3", "VEH")
  ### Computing log FCs
  H3K27ac_log2fc <- log2((H3K27ac_DOX_avg + 1) / (H3K27ac_VEH_avg + 1))
  H3K9me3_log2fc <- log2((H3K9me3_DOX_avg + 1) / (H3K9me3_VEH_avg + 1))
  
  H3K27ac_rowmean <- rowMeans(H3K27ac_log2fc)
  H3K9me3_rowmean <- rowMeans(H3K9me3_log2fc)
  
  switching_rows <- which(H3K27ac_rowmean > 0.01 &
                        H3K9me3_rowmean < 0)
  
  row_names <- cm[switching_rows, V4]
   prepare_mat <- function(mat, rows) {
    mat_sub <- mat[rows, , drop = FALSE]
    mat_scaled <- t(scale(t(mat_sub)))
    mat_scaled[is.na(mat_scaled)] <- 0
    mat_scaled
  }
  
  # Reference row order: H3K27ac DOX 24T
  ref_mat <- signal_list[["H3K27ac_DOX_24T_merged"]][switching_rows, ]
  # rownames(ref_mat) <- cm[switching_rows, V4][[1]]
  row_order <- order(rowMeans(ref_mat), decreasing = TRUE)
  
  # Function to make heatmaps for a given mark
  make_heatmap <- function(mark, samples) {
    ht_list <- NULL
    for(s in samples) {
      mat_scaled <- prepare_mat(signal_list[[s]], switching_rows)
      mat_scaled <- mat_scaled[row_order, ]
      short_title <- str_extract(s, "(DOX|VEH)_\\d+[RT]")
      ht <- ComplexHeatmap::Heatmap(mat_scaled,
                    col = colorRamp2(c(-2,0,2), c("blue","white","red")),
                    name = "signal",
                    show_row_names = FALSE,
                    show_column_names = FALSE,
                    cluster_rows = FALSE,
                    cluster_columns = FALSE,
                    column_title = short_title)
      ht_list <- if(is.null(ht_list)) ht else ht_list + ht
    }
    ht_list
  }
  
  # Define sample order for plotting
  H3K27ac_samples <- c("H3K27ac_DOX_24T_merged","H3K27ac_DOX_24R_merged","H3K27ac_DOX_144R_merged",
                       "H3K27ac_VEH_24T_merged","H3K27ac_VEH_24R_merged","H3K27ac_VEH_144R_merged")
  H3K9me3_samples <- c("H3K9me3_DOX_24T_merged","H3K9me3_DOX_24R_merged","H3K9me3_DOX_144R_merged",
                       "H3K9me3_VEH_24T_merged","H3K9me3_VEH_24R_merged","H3K9me3_VEH_144R_merged")
  
  ht_H3K27ac <- make_heatmap("H3K27ac", H3K27ac_samples)
  ht_H3K9me3 <- make_heatmap("H3K9me3", H3K9me3_samples)
  
  # -----------------------------
  # 7️⃣ Extract file part for master title
  # -----------------------------
  file_part <- str_extract(basename(matrix_file), "(?<=matcenter_).*?(?=_ERV1_LTR)")
  
  draw(ht_H3K27ac, heatmap_legend_side = "right", column_title = paste0("H3K27ac switching - ", file_part))
  draw(ht_H3K9me3, heatmap_legend_side = "right", column_title = paste0("H3K9me3 switching - ", file_part))
  
  row_names_MER61E <- bed_info$V4[switching_rows]
  
  
  res %>% dplyr::filter(ltr %in% row_names_MER61E)
  
```
```{r summary of rowmean}
summary(H3K27ac_rowmean)
summary(H3K9me3_rowmean)
```


### Identifying switch rows (not needed now)  


```{r identify switch rows }
MER61E_switch_genes <- res %>% 
  dplyr::filter(ltr %in% row_names_MER61E) %>% 
  distinct(gene,.keep_all = TRUE)


for (aim in MER61E_switch_genes$gene){
  df <- MER61E_lcpm_genes %>% 
    dplyr::filter(Entrezid== aim)
  if (nrow(df) == 0) next
  symbol <- Symbol_lookup %>% 
    dplyr::filter(Entrez_ID==aim) %>% 
    dplyr::pull(SYMBOL)
  distance <- MER61E_switch_genes %>% 
    dplyr::filter(gene == aim) %>% 
    dplyr::pull(distance)
  
plot <- ggplot(df, aes(x=time, y=log2cpm,group=group))+
  geom_boxplot(aes(fill=trt)) +
  ggtitle(paste0("gene ", aim," ",symbol," dist ",distance))+
  theme_bw()
print(plot)
}


```

```{r switch plots}
RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"))) %>% 
  dplyr::filter(Entrez_ID %in% MER61E_switch_genes$gene) %>% 
  ggplot(., aes(x=timepoint, y=logFC))+
  geom_boxplot()+
  ggtitle("LFC at timepoints of genes switched on MER61E TE")+
  theme_bw()


RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"))) %>% 
  dplyr::filter(Entrez_ID %in% MER61E_switch_genes$gene) %>% 
  ggplot(., aes(x=timepoint, y=logFC, group = Entrez_ID))+
  geom_point()+
  geom_line()+
  ggtitle("LFC at timepoints of genes switched on MER61E TE")


```


### Proportion tables   



```{r prop 1}
prop_table <- annotated_H3K27ac_summits %>% 
  dplyr::filter(repName=="MER61E") %>% 
  group_by(cluster) %>% 
  summarise(n = dplyr::n(), .groups = "drop") %>% 
   # optionally convert NA to a string
  mutate(cluster = if_else(is.na(cluster), "NA", cluster),
         clus_names=case_when(cluster =="Set_1"~"No response",
                              cluster=="Set_3"~"Acute response",
                              cluster=="Set_2"~"Sustained response",
                              TRUE~"not assigned")) 
  
if(!"Set_2" %in% prop_table$cluster){
  prop_table <- rbind(
    prop_table,
    data.frame(cluster = "Set_2", n = 0,clus_names="Sustained response")
  )} 
ggplot(prop_table, aes(x = clus_names, y = n, fill = clus_names)) +
  geom_col(position = "identity") +
  geom_text(aes(
      label = sprintf('%d (%.1f%%)', n, 100 * n / sum(n)),  # use n directly
      y = n + 0.05 * max(n)  # nudge above bar
    ),
    colour = 'royalblue',
    size = 5
  ) +
  theme_classic() +
  labs(
    x = "H3K27ac response group",
    y = "Proportion of MER61Es",
    fill = "Cluster",
    title = "Proportion of MER61E across each H3K27ac set"
  ) +
  theme(axis.text.x = element_text(angle = 0, hjust = .5))


```
```{r proportions of Mers}
Mer_list <- c("MER61-int","MER61A","MER61B","MER61C","MER61D","MER61E","MER61F")

for (Mer in Mer_list){

prop_table <- annotated_H3K27ac_summits %>% 
  dplyr::filter(repName==Mer) %>% 
  group_by(cluster) %>% 
  summarise(n = dplyr::n(), .groups = "drop") %>% 
  mutate(cluster = if_else(is.na(cluster), "NA", cluster),
         clus_names=case_when(cluster =="Set_1"~"No response",
                              cluster=="Set_3"~"Acute response",
                              cluster=="Set_2"~"Sustained response",
                              TRUE~"not assigned")) 
  
if(!"Set_2" %in% prop_table$cluster){
  prop_table <- rbind(
    prop_table,
    data.frame(cluster = "Set_2", n = 0,clus_names="Sustained response")
  )} 
if(!"Set_3" %in% prop_table$cluster){
  prop_table <- rbind(
    prop_table,
    data.frame(cluster = "Set_3", n = 0,clus_names="Acute response")
  )} 

plot <- ggplot(prop_table, aes(x = clus_names, y = n, fill = clus_names)) +
  geom_col(position = "identity") +
  geom_text(aes(
      label = sprintf('%d (%.1f%%)', n, 100 * n / sum(n)),  
      y = n + 0.05 * max(n)),
    colour = 'royalblue',
    size = 5) +
  theme_classic() +
  labs(x = "H3K27ac response group",
    y = paste0("Proportion of ",Mer),
    fill = "Cluster",
    title = paste0("Proportion of ",Mer," across each H3K27ac set")) +
  theme(axis.text.x = element_text(angle = 0, hjust = .5))
    
    print(plot)
}



```

### KZFP hits  

```{r}
KZFP_hits_all <- KZFP_targets %>% 
  dplyr::filter(`Target Subfamily` %in% c("HERV1_I-int","LTR10A","LTR10C", "MER61A","MER61B","MER61C","MER61D","MER61E","MER61F","MER61-int")) %>% 
  distinct(KZFP)

RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"))) %>% 
  dplyr::filter(SYMBOL %in% KZFP_hits_all$KZFP) %>% 
  ggplot(., aes(x=timepoint, y=logFC))+
  geom_boxplot()+
  geom_point(aes(color=SYMBOL))+
  ggtitle("LFC of KZFPs that target specific ERV1 subfamilies")+
  theme_bw()+
  ylim(-2,2)

 test_table <- KZFP_targets %>% 
  dplyr::filter(`Target Subfamily` %in% c("HERV1_I-int","LTR10A","LTR10C", "MER61A","MER61B","MER61C","MER61D","MER61E","MER61F","MER61-int"))  %>% 
   dplyr::select(KZFP, `Target Subfamily`)
 kable(test_table,
       caption = "ZNFs that target specific ERV1:LTRs") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE) 
```

```{r KZFP hits}

# gene_lookup <- RNA_full_toptable %>% distinct(Entrez_ID,SYMBOL)
KZFP_hits <- KZFP_targets %>% 
  dplyr::filter(`Target Subfamily` %in% c("MER61A","MER61B","MER61C","MER61D","MER61E","MER61F","MER61-int")) %>% 
  distinct(KZFP)

RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"))) %>% 
  dplyr::filter(SYMBOL %in% KZFP_hits$KZFP) %>% 
  ggplot(., aes(x=timepoint, y=logFC))+
  geom_boxplot()+
  geom_point(aes(color=SYMBOL))+
  ggtitle("LFC of KZFPs that target MER61* subfamily")+
  theme_bw()+
  ylim(-2,2)


for (aim in KZFP_hits$KZFP){
  entrez_id <- Symbol_lookup %>% 
    dplyr::filter(SYMBOL==aim) %>% 
    dplyr::pull(Entrez_ID)
  df <- RNA_filcpm_final_matrix %>% 
    dplyr::filter(Entrezid== entrez_id) %>% 
     pivot_longer(cols=!Entrezid, names_to = "sample",values_to = "log2cpm") %>% 
  separate_wider_delim(cols = sample, names = c("trt","time","ind"), delim = "_") %>% 
  mutate(group=paste0(trt,"_",time),
  time=factor(time,levels=c("24T","24R","144R"))) 
    
  if (nrow(df) == 0) next
  
plot <- ggplot(df, aes(x=time, y=log2cpm,group=group))+
  geom_boxplot(aes(fill=trt)) +
  ggtitle(paste0("gene ", aim," ",entrez_id ))+
  theme_bw()
print(plot)
}

```

```{r filtering step for ERV1}
ERV1_summary_char %>% 
  mutate(name=str_remove(filename,"_ERV1_LTR.bed")) %>% 
  dplyr::filter(name %in%Mer_list) %>% 
  dplyr::rename("count"=n_lines) %>% 
  dplyr::select( name, count:min_len)
  
                                
```


```{r MER61E_ids}
MER61E_IDs_clusters <- H3K27ac_summit_ols %>% 
  # dplyr::filter(repName %in% Mer_list) %>% 
  dplyr::filter(repName =="MER61E") %>% 
  dplyr::select(cluster,ID)

 # cm_MER61E <- cm[cm$V4 %in% MER61E_IDs_clusters$ID, ]

signal_list_filtered <- lapply(signal_list, function(mat) {
  mat[cm$V4 %in% MER61E_IDs_clusters$ID, , drop = FALSE]
})
MER61E_rows <- base::which(cm$V4 %in% MER61E_IDs_clusters$ID)
# ref_mat <- signal_list[["H3K27ac_DOX_24T_merged"]][MER61E_rows, ]
# row_order <- order(rowMeans(ref_mat), decreasing = TRUE)

cluster_df_subset <- MER61E_IDs_clusters[match(cm$V4[MER61E_rows], MER61E_IDs_clusters$ID), ]
row_order_cluster <- order(cluster_df_subset$cluster)

make_heatmap_MER61E <- function(mark, samples) {
  ht_list <- NULL
  for(s in samples) {
    mat_scaled <- prepare_mat(signal_list[[s]], MER61E_rows)
    mat_scaled <- mat_scaled[row_order_cluster, ]
    short_title <- str_extract(s, "(DOX|VEH)_\\d+[RT]")
    ht <- ComplexHeatmap::Heatmap(mat_scaled,
                  col = colorRamp2(c(-2,0,2), c("blue","white","red")),
                  name = "signal",
                  show_row_names = TRUE,  # optional
                  show_column_names = FALSE,
                  cluster_rows = FALSE,
                  cluster_columns = FALSE,
                  column_title = short_title,
                  row_split = cluster_df_subset$cluster[row_order_cluster] )
    ht_list <- if(is.null(ht_list)) ht else ht_list + ht
  }
  ht_list
}


H3K27ac_samples <- c("H3K27ac_DOX_24T_merged","H3K27ac_DOX_24R_merged","H3K27ac_DOX_144R_merged",
                       "H3K27ac_VEH_24T_merged","H3K27ac_VEH_24R_merged","H3K27ac_VEH_144R_merged")
  H3K9me3_samples <- c("H3K9me3_DOX_24T_merged","H3K9me3_DOX_24R_merged","H3K9me3_DOX_144R_merged",
                       "H3K9me3_VEH_24T_merged","H3K9me3_VEH_24R_merged","H3K9me3_VEH_144R_merged")

ht_H3K27ac_MER61E <- make_heatmap_MER61E("H3K27ac", H3K27ac_samples)
# ht_H3K27ac_MER61E  # draw heatmap


ht_H3K9me3_MER61E <- make_heatmap_MER61E("H3K9me3", H3K9me3_samples)
# ht_H3K9me3_MER61E
file_part <- str_extract(basename(matrix_file), "(?<=matcenter_).*?(?=_ERV1_LTR)")
  
  draw(ht_H3K27ac_MER61E, heatmap_legend_side = "right", column_title = paste0("H3K27ac all - ", file_part))
  draw(ht_H3K9me3_MER61E, heatmap_legend_side = "right", column_title = paste0("H3K9me3 all - ", file_part))
  

```
### MER61 subfamily intro  


here i am trying to construct a full matrices of all MER subfamilies. This involves concatenating several files
```{r MER61subfamily}
MERall_IDs_clusters <- H3K27ac_summit_ols %>% 
  dplyr::filter(repName %in% Mer_list) %>%
  # dplyr::filter(repName =="MER61E") %>% 
  dplyr::select(cluster,repName,ID)


MER61_files <- list.files(
  path = "C:/Users/renee/Other_projects_data/DXR_data/computeMatrix_output_centered2histone/",
  pattern = "MER61.*\\.gz$", full.names = TRUE
)

# Function to read each file
read_matrix_file <- function(file) {
  fread(cmd = paste("gzip -dc", file), skip = 1)
}

# Read all files and concatenate
MER61_all <- rbindlist(lapply(MER61_files, read_matrix_file))

bins <- 400  # number of bins per sample
sample_names <- c(
  "H3K27ac_DOX_144R_merged","H3K27ac_DOX_24R_merged","H3K27ac_DOX_24T_merged",
  "H3K27ac_VEH_144R_merged","H3K27ac_VEH_24R_merged","H3K27ac_VEH_24T_merged",
  "H3K9me3_DOX_144R_merged","H3K9me3_DOX_24R_merged","H3K9me3_DOX_24T_merged",
  "H3K9me3_VEH_144R_merged","H3K9me3_VEH_24R_merged","H3K9me3_VEH_24T_merged"
)

signal_list_MER61all <- lapply(seq_along(sample_names), function(i){
  start <- 6 + (i-1)*bins + 1
  end <- 6 + i*bins
  as.matrix(MER61_all[, start:end, with = FALSE])
})
names(signal_list_MER61all) <- sample_names
```

### Heatmap of MER61 subfamily signal  


```{r mappingall MER61, fig.height=10}


MER61_filtered <- MER61_all[MER61_all$V4 %in% MERall_IDs_clusters$ID, ]

MER61_filtered <- merge(MER61_filtered, MERall_IDs_clusters,
                        by.x = "V4", by.y = "ID", all.x = TRUE)
# # Optional: remove rows without a cluster assignment
# MER61_filtered <- MER61_filtered[!is.na(MER61_filtered$cluster), ]


MER61fam_rows <- base::which(MER61_all$V4 %in% MER61_filtered$V4)

# Keep filtered MER61 info in the same order
MER61_filtered <- MER61_filtered[match(MER61_all$V4[MER61fam_rows], MER61_filtered$V4), ]

row_order_MER61fam <- order(MER61_filtered$repName, MER61_filtered$cluster)

prepare_mat <- function(mat, rows) {
    mat_sub <- mat[rows, , drop = FALSE]
    mat_scaled <- t(scale(t(mat_sub)))
    mat_scaled[is.na(mat_scaled)] <- 0
    mat_scaled
  }

make_heatmap_MER61fam <- function(samples) {
  ht_list <- NULL
  for(s in samples) {
    mat_scaled <- prepare_mat(signal_list_MER61all[[s]], MER61fam_rows)
    mat_scaled <- mat_scaled[row_order_MER61fam, ]
   
    
    short_title <- str_extract(s, "(DOX|VEH)_\\d+[RT]")
    
    ht <- ComplexHeatmap::Heatmap(mat_scaled,
                  col = colorRamp2(c(-2,0,2), c("blue","white","red")),
                  name = "signal",
                  show_row_names = TRUE,
                  show_column_names = FALSE,
                  cluster_rows = FALSE,
                  cluster_columns = FALSE,
                  column_title = short_title,
                  row_split = MER61_filtered$repName[row_order_MER61fam], border=TRUE ) 
    ht_list <- if(is.null(ht_list)) ht else ht_list + ht
  }
  ht_list
}

H3K27ac_samples <- c("H3K27ac_DOX_24T_merged","H3K27ac_DOX_24R_merged","H3K27ac_DOX_144R_merged",
                       "H3K27ac_VEH_24T_merged","H3K27ac_VEH_24R_merged","H3K27ac_VEH_144R_merged")
  H3K9me3_samples <- c("H3K9me3_DOX_24T_merged","H3K9me3_DOX_24R_merged","H3K9me3_DOX_144R_merged",
                       "H3K9me3_VEH_24T_merged","H3K9me3_VEH_24R_merged","H3K9me3_VEH_144R_merged")
ht_H3K27ac_MERall <- make_heatmap_MER61fam(H3K27ac_samples)

ht_H3K9me3_MERall <- make_heatmap_MER61fam(H3K9me3_samples)


draw(ht_H3K27ac_MERall, heatmap_legend_side = "right", column_title = paste0("H3K27ac all MER61"))
  draw(ht_H3K9me3_MERall, heatmap_legend_side = "right", column_title = paste0("H3K9me3 all MER61"))
  
```
```{r}



MER61E_lcpm_genes_one <- RNA_filcpm_final_matrix %>% 
  dplyr::filter(Entrezid %in% c("374383")) %>% 
  pivot_longer(cols=!Entrezid, names_to = "sample",values_to = "log2cpm") %>% 
  separate_wider_delim(cols = sample, names = c("trt","time","ind"), delim = "_") %>% 
  mutate(group=paste0(trt,"_",time),
  time=factor(time,levels=c("24T","24R","144R"))) 


# MER61E_lcpm_genes %>% 
   
  symbol <- Symbol_lookup %>% 
    dplyr::filter(Entrez_ID=="374383") %>% 
    dplyr::pull(SYMBOL)
  
ggplot(MER61E_lcpm_genes_one, aes(x=time, y=log2cpm,group=group))+
  geom_boxplot(aes(fill=trt)) +
  ggtitle(paste0("gene 374383 ",symbol))+
  theme_bw()

```
###  Expanding the overlap +/- 200bp from summit

```{r}

annotated_H3K27ac_extended <- readRDS("data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")
# H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")
# H3K27ac_summit_gr_ext <- resize(H3K27ac_summit_gr, width(H3K27ac_summit_gr)+400, fix = "center")
# 
# extended_MER61E_subset <- subsetByOverlaps(H3K27ac_summit_gr_ext,ltr_gr)

annotated_H3K27ac_extended %>% as.data.frame() %>% 
  filter(str_detect(repName,";"))



annotated_H3K27ac_extended %>%
  
  mutate(across(repClass:repID, ~str_count(.x, ";") + 1, .names = "n_{col}")) %>%
  filter(!(n_repClass == n_repFamily & n_repFamily == n_repName & n_repName == n_repID))
  
  
  # annotated_H3K27ac_extended %>% 
  #   dplyr::filter(str_detect(repName,str_c(Mer_list, collapse="|")))
```


```{r}
for (Mer in Mer_list){

prop_table <- annotated_H3K27ac_extended %>% 
  dplyr::filter(str_detect(repName,Mer)) %>% 
  group_by(cluster) %>% 
  summarise(n = dplyr::n(), .groups = "drop") %>% 
  mutate(cluster = if_else(is.na(cluster), "NA", cluster),
         clus_names=case_when(cluster =="Set_1"~"No response",
                              cluster=="Set_3"~"Acute response",
                              cluster=="Set_2"~"Sustained response",
                              TRUE~"not assigned")) 
  
if(!"Set_2" %in% prop_table$cluster){
  prop_table <- rbind(
    prop_table,
    data.frame(cluster = "Set_2", n = 0,clus_names="Sustained response")
  )} 
if(!"Set_3" %in% prop_table$cluster){
  prop_table <- rbind(
    prop_table,
    data.frame(cluster = "Set_3", n = 0,clus_names="Acute response")
  )} 

plot <- ggplot(prop_table, aes(x = clus_names, y = n, fill = clus_names)) +
  geom_col(position = "identity") +
  geom_text(aes(
      label = sprintf('%d (%.1f%%)', n, 100 * n / sum(n)),  
      y = n + 0.05 * max(n)),
    colour = 'royalblue',
    size = 5) +
  theme_classic() +
  labs(x = "H3K27ac response group",
    y = paste0("Proportion of ",Mer),
    fill = "Cluster",
    title = paste0("Proportion of ",Mer," across each H3K27ac set extended")) +
  theme(axis.text.x = element_text(angle = 0, hjust = .5))
    
    print(plot)
}


```

```{r}
LTR_df <- readRDS("data/RDS_files/LTR_DF_hg38.RDS")

hgLTR <- LTR_df %>% 
  filter(repName %in% Mer_list) %>% 
  dplyr::select(repClass,repFamily,repName,RM_id) %>% 
  mutate(Peakid="NA",
         annotation="NA",
         geneId="NA",
         distanceToTSS=0,
         summit_pos="0",
         TE_status="TE",
         SINE_status="wnot_SINE",
         LINE_status="wnot_LINE",
         LTR_status="LTR",
         DNA_status="wnot_LTR",
         SVA_status="wnot_SVA",
         cluster="hg38")
```


#### MER subfamily  


```{r prop_extended plot}
df_clean <- annotated_H3K27ac_extended %>%
  separate_rows(repClass,repFamily,repName,repID, sep = ";")

prop_table_all <- df_clean %>%
  filter(repName %in% Mer_list) %>%
  bind_rows(hgLTR) %>% 
  mutate(
    cluster = if_else(is.na(cluster), "NA", cluster),
    clus_names = case_when(
      cluster == "Set_1" ~ "No response",
      cluster == "Set_2" ~ "Sustained response",
      cluster == "Set_3" ~ "Acute response",
      cluster == "hg38" ~ "hg38",
      TRUE ~ "Not assigned"
    )
  ) %>%
  count(clus_names, repName) %>%
  group_by(clus_names) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>% 
  mutate(clus_names=factor(clus_names, levels = c("hg38","No response","Sustained response","Acute response","Not assigned")))


ggplot(prop_table_all, aes(x = clus_names, y = prop, fill = repName)) +
  geom_col() +
  theme_classic() +
  labs(
    x = "H3K27ac response group",
    y = "Proportion",
    fill = "MER61 subfamily",
    title = "MER61 subfamily composition across H3K27ac sets"
  )
```

### ERVL-MaLR:LTRs 


```{r ERVL-MaLR initial}

ERVLMaLR_counts_H3K27ac <- df_clean %>% 
  dplyr::filter(repFamily=="ERVL-MaLR") %>% 
  group_by(repName,cluster) %>% 
  tally %>% 
  pivot_wider(id_cols = repName,values_from = n, names_from = cluster) %>% 
  replace(is.na(.), 0)

ERVLMaLR_counts_hg38 <- LTR_df %>% 
  dplyr::filter(repFamily=="ERVL-MaLR") %>% 
  group_by(repName) %>% 
  count(name="hg38")


ERVLMaLR_joined_counts <- ERVLMaLR_counts_hg38 %>% 
  left_join(ERVLMaLR_counts_H3K27ac) %>% 
  replace(is.na(.), 0)
  


calc_enrichment <- function(df, set_col) {
  total_genome <- sum(df$hg38, na.rm = TRUE)
  total_set    <- sum(df[[set_col]], na.rm = TRUE)

  df %>%
    rowwise() %>%
    mutate(
      a = .data[[set_col]],
      b = hg38 - a,
      c = total_set - a,
      d = total_genome - (a + b + c),

      odds_ratio = ((a + 1e-5) * (d + 1e-5)) /
                 ((b + 1e-5) * (c + 1e-5)),

      p_value = fisher.test(matrix(c(a, b, c, d), nrow = 2))$p.value
    ) %>%
    ungroup()
}


calc_pairwise <- function(df, setA, setB) {
  total_A <- sum(df[[setA]], na.rm = TRUE)
  total_B <- sum(df[[setB]], na.rm = TRUE)

  df %>%
    rowwise() %>%
    mutate(
      a = .data[[setA]],
      b = .data[[setB]],
      c = total_A - a,
      d = total_B - b,

       odds_ratio = ((a + 1e-5) * (d + 1e-5)) /
                 ((b + 1e-5) * (c + 1e-5)),

      p_value = fisher.test(matrix(c(a, b, c, d), nrow = 2))$p.value
    ) %>%
    ungroup() %>%
    mutate(
      comparison = paste(setA, "vs", setB),
      p_adj = p.adjust(p_value, method = "BH")
    )
}
```

```{r}
res_set1 <- calc_enrichment(ERVLMaLR_joined_counts, "Set_1")
res_set2 <- calc_enrichment(ERVLMaLR_joined_counts, "Set_2")
res_set3 <- calc_enrichment(ERVLMaLR_joined_counts, "Set_3")

res_set1 <- res_set1 %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))
res_set2 <- res_set2 %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))
res_set3 <- res_set3 %>%
  mutate(p_adj = p.adjust(p_value, method = "BH"))


res_2_v_1 <- calc_pairwise(ERVLMaLR_joined_counts,"Set_2","Set_1")

res_3_v_1 <- calc_pairwise(ERVLMaLR_joined_counts,"Set_3","Set_1")

res_2_v_3 <- calc_pairwise(ERVLMaLR_joined_counts, "Set_2","Set_3")

```


```{r}
ggplot(res_2_v_1, aes(x = (odds_ratio), y = -log10(p_adj),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  geom_text_repel(data = subset(res_2_v_1, p_adj < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  # geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "ERVL-MaLR family enrichment at extended ROI summits\n 2 v 1"
  ) +
  theme_classic()

ggplot(res_3_v_1, aes(x = (odds_ratio), y = -log10(p_adj),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  # geom_text_repel(size=3, max.overlaps = Inf)+
  geom_text_repel(data = subset(res_3_v_1, p_adj < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "ERVL-MaLR family enrichment at extended ROI summits\n 3 v 1"
  ) +
  theme_classic()


ggplot(res_2_v_3, aes(x = (odds_ratio), y = -log10(p_adj),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  # geom_text_repel(size=3, max.overlaps = Inf)+
  geom_text_repel(data = subset(res_2_v_3, p_adj < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "ERVL-MaLR family enrichment at extended ROI summits\n 2 v 3"
  ) +
  theme_classic()
```

```{r}
sum(res_3_v_1$Set_2)
sum(res_2_v_1$Set_3)
sum(res_2_v_3$not_assigned)
```

