---
title: "H3K27ac and TE overlap"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(ChIPseeker)
library(ggVennDiagram)
library(smplot2)
```

First steps: breakdown repeatmasker into groups and pull out the ones by each class I am interested in.  
```{r repeatmasker loading}

repeatmasker <- read_delim("data/Other_paper_data/repeatmasker_20250911.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
colnames(repeatmasker)

autosomes <- paste0("chr", 1:22)
repeatmasker_clean <- repeatmasker %>% mutate(
    strand = ifelse(strand == "C", "-", "+")
  ) %>% 
   mutate(
    start = genoStart + 1,
    end   = genoEnd)%>% 
  mutate(repFamily= str_remove(repFamily, "\\?$")) %>% 
  dplyr::filter(genoName %in% autosomes) %>% 
  mutate(RM_id=paste0(genoName,":",start,"-",end,":",id))


rpt_split <- split(repeatmasker_clean, repeatmasker_clean$repClass)

rpt_split_gr_list <- lapply(rpt_split, function(df) {
  GRanges(
    seqnames = df$genoName,
    ranges   = IRanges(start = df$start, end = df$end),
    strand   = df$strand,
    repName   = df$repName,
    repClass  = df$repClass,
    repFamily = df$repFamily,
    swScore   = df$swScore,
    milliDiv  = df$milliDiv,
    milliDel  = df$milliDel,
    milliIns  = df$milliIns,
    RM_id        = df$RM_id
  )
})

```

```{r subset rpt_split_gr_list}
SINE_gr <- rpt_split_gr_list$SINE
SINE_df <- SINE_gr %>% 
  as.data.frame()

SINE_split_df <- split(SINE_df, SINE_df$repFamily)

LINE_gr <- rpt_split_gr_list$LINE
LINE_df <- LINE_gr %>% 
  as.data.frame()

LINE_split_df <- split(LINE_df, LINE_df$repFamily)

LTR_gr <- rpt_split_gr_list$LTR
LTR_df <- LTR_gr %>% 
  as.data.frame()

LTR_split_df <- split(LTR_df, LTR_df$repFamily)

SVA_gr <- rpt_split_gr_list$Retroposon
SVA_df <- SVA_gr %>% 
  as.data.frame()

SVA_split_df <- split(SVA_df, SVA_df$repFamily)

DNA_gr <- rpt_split_gr_list$DNA
DNA_df <- DNA_gr %>% 
  as.data.frame()

DNA_split_df <- split(DNA_df, DNA_df$repFamily)
```

```{r loading H3K27ac data}
H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")


peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")
peakAnnoList_H3K9me3 <- readRDS("data/motif_lists/H3K9me3_annotated_peaks.RDS")

H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y))

H3K9me3_lookup <- imap_dfr(peakAnnoList_H3K9me3[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y))

H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})

H3K9me3_sets_gr <- lapply(peakAnnoList_H3K9me3, function(df) {
  as_granges(df)
})

##assigning Peakid as name of summit region
 mcols(H3K27ac_summit_gr)$name <- mcols(H3K27ac_summit_gr)$Peakid
 
 comparisons <- tibble(
  cluster2 = c("Set_2", "Set_3"),
  cluster1 = c("Set_1", "Set_1")
)

```

```{r toplist data}
H3K9me3_toplist <- readRDS( "data/DER_data/H3K9me3_toplist_nooutlier.RDS")

H3K27ac_toplist <- readRDS( "data/DER_data/H3K27ac_toplist.RDS")

H3K27ac_toptable_list <- bind_rows(H3K27ac_toplist, .id = "group")

H3K9me3_toptable_list <- bind_rows(H3K9me3_toplist, .id = "group")

K9me3_lfctable <- H3K9me3_toptable_list %>% 
  dplyr::select(group,genes, logFC) %>% 
  pivot_wider(.,id_cols = genes, names_from = group, values_from = logFC) %>%
  dplyr::left_join(H3K9me3_lookup, by =c("genes"="Peakid")) %>% 
  mutate(case= case_when(H3K9me3_24T > 0 & H3K9me3_24R >0 & H3K9me3_144R>0~"A",
                              H3K9me3_24T < 0 & H3K9me3_24R <0 & H3K9me3_144R<0~"B",
                              TRUE ~ "C")) %>% 
  mutate(set_case=paste0(cluster,"_",case))

K27ac_lfctable <- H3K27ac_toptable_list %>% 
  dplyr::select(group,genes, logFC) %>% 
  pivot_wider(.,id_cols = genes, names_from = group, values_from = logFC) %>% 
  dplyr::left_join(H3K27ac_lookup,by =c("genes"="Peakid")) %>% 
  mutate(case= case_when(H3K27ac_24T > 0 & H3K27ac_24R >0 & H3K27ac_144R>0~"A",
                              H3K27ac_24T < 0 & H3K27ac_24R <0 & H3K27ac_144R<0~"B",
                              TRUE ~ "C")) %>% 
  mutate(set_case=paste0(cluster,"_",case))

cluster_colors <- c("Set_1" = "#d93e40","Set_2" = "#1c9f50","Set_3" = "#3570b3","NA"  ="grey70")

```

```{r Fisher test information}

# Generic pairwise Fisher test
test_pair_TE_generic <- function(df_long, te_name, cluster1, cluster2) {
  
   sub_df <- df_long %>%
    filter(TE_type == te_name) %>%
    complete(
      cluster = c(cluster1, cluster2),
      status  = c("TE", "not_TE"),
      fill = list(count = 0))
   
   # enforce fixed order
 status_levels <- c("TE", "not_TE")
  
  # assume "status" column has TE vs wnot_TE automatically
  statuses <- unique(sub_df$status)
  
  if(length(statuses) != 2) {
    # ensure we have exactly two categories, fill missing with 0
    sub_df <- sub_df %>%
      complete(cluster, status, fill = list(count = 0))
    statuses <- unique(sub_df$status)
  }
  
  # extract counts for cluster1
  c1_counts <- sub_df %>%
    filter(cluster == cluster1) %>%
   arrange(factor(status, levels = status_levels)) %>%  # ensure same order
    pull(count)
  
  # extract counts for cluster2
  c2_counts <- sub_df %>%
    filter(cluster == cluster2) %>%
    arrange(factor(status, levels = status_levels)) %>%
    pull(count)
  
  # build 2x2 matrix
  mat <- matrix(
    c(c2_counts, c1_counts),
    nrow = 2,
    byrow = TRUE,
    dimnames = list(
      cluster = c(cluster2, cluster1),
      category = status_levels
    )
  )
  
  ft <- tryCatch(
    fisher.test(mat, workspace = 2e8),
    error = function(e) fisher.test(mat, simulate.p.value = TRUE, B = 1e5)
  )
  
  tibble(
    TE_type     = te_name,
    comparison  = paste(cluster2, "vs", cluster1),
    odds_ratio  = ft$estimate,
    lower_CI    = ft$conf.int[1],
    upper_CI    = ft$conf.int[2],
    p_value     = ft$p.value
  )
}
```


### SINE
Overlapping SINE family with ROIs

```{r SINE Overlapping}
sine_hits <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, SINE_gr, ignore.strand = TRUE)

SINE_overlap_df <- tibble(
  peak_row = queryHits(sine_hits),
  Peakid   = H3K27ac_sets_gr$all_H3K27ac$Peakid[queryHits(sine_hits)],
  cluster  = H3K27ac_sets_gr$all_H3K27ac$cluster[queryHits(sine_hits)],
  
  
  repClass = SINE_gr$repClass[subjectHits(sine_hits)],
  repName  = SINE_gr$repName[subjectHits(sine_hits)],
  
  TE_type  = ifelse(
    SINE_gr$repFamily[subjectHits(sine_hits)] == "SVA",
    SINE_gr$repName[subjectHits(sine_hits)],
    SINE_gr$repFamily[subjectHits(sine_hits)]
  ),
  
  milliDiv = SINE_gr$milliDiv[subjectHits(sine_hits)],
  milliDel = SINE_gr$milliDel[subjectHits(sine_hits)],
  milliIns = SINE_gr$milliIns[subjectHits(sine_hits)]
)

SINE_overlap_df %>% 
  dplyr::left_join(H3K27ac_lookup) %>% 
  dplyr::filter(cluster %in% c("Set_1","Set_2","Set_3")) %>% 
  ggplot(., aes(x = TE_type, y = milliDiv)) +
   geom_violin(trim = FALSE) +
   geom_boxplot(width = 0.15, outlier.shape = NA) +
   coord_flip() +
   labs(
     y = "Divergence from consensus (milliDiv)",
     x = "TE family",
     title = "Age distribution of SINE elements overlapping H3K27ac"
   )+
  facet_wrap(~cluster)
```


```{r SINE Overlapping 2, fig.height=10}
SINE_overlap_df %>% 
  dplyr::left_join(H3K27ac_lookup) %>% 
  dplyr::filter(cluster %in% c("Set_1","Set_2","Set_3")) %>% 
  ggplot(., aes(x = cluster, y = milliDiv)) +
  geom_violin(trim = FALSE) +
   geom_boxplot(width = 0.15, outlier.shape = NA) +
   coord_flip() +
   labs(
     y = "Divergence from consensus (milliDiv)",
     x = "TE family",
     title = "Age distribution of SINE elements overlapping H3K27ac"
   )+
  facet_wrap(~repName)
```



### LINE
Overlapping LINE family with ROIs

```{r LINE Overlapping}
LINE_hits <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, LINE_gr, ignore.strand = TRUE)

LINE_overlap_df <- tibble(
  peak_row = queryHits(LINE_hits),
  Peakid   = H3K27ac_sets_gr$all_H3K27ac$Peakid[queryHits(LINE_hits)],
  cluster  = H3K27ac_sets_gr$all_H3K27ac$cluster[queryHits(LINE_hits)],
  
  repClass = LINE_gr$repClass[subjectHits(LINE_hits)],
  repName  = LINE_gr$repName[subjectHits(LINE_hits)],
  
  TE_type  = ifelse(
    LINE_gr$repFamily[subjectHits(LINE_hits)] == "SVA",
    LINE_gr$repName[subjectHits(LINE_hits)],
    LINE_gr$repFamily[subjectHits(LINE_hits)]
  ),
  
  milliDiv = LINE_gr$milliDiv[subjectHits(LINE_hits)],
  milliDel = LINE_gr$milliDel[subjectHits(LINE_hits)],
  milliIns = LINE_gr$milliIns[subjectHits(LINE_hits)]
)

LINE_overlap_df %>% 
  dplyr::left_join(H3K27ac_lookup) %>% 
  dplyr::filter(cluster %in% c("Set_1","Set_2","Set_3")) %>% 
  ggplot(., aes(x = TE_type, y = milliDiv)) +
   geom_violin(trim = FALSE) +
   geom_boxplot(width = 0.15, outlier.shape = NA) +
   coord_flip() +
   labs(
     y = "Divergence from consensus (milliDiv)",
     x = "TE family",
     title = "Age distribution of LINE elements overlapping H3K27ac"
   )+
  facet_wrap(~cluster)
```


### LTR
Overlapping LTR family with ROIs

```{r LTR Overlapping}
LTR_hits <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, LTR_gr, ignore.strand = TRUE)

LTR_overlap_df <- tibble(
  peak_row = queryHits(LTR_hits),
  Peakid   = H3K27ac_sets_gr$all_H3K27ac$Peakid[queryHits(LTR_hits)],
  cluster  = H3K27ac_sets_gr$all_H3K27ac$cluster[queryHits(LTR_hits)],
  
  repClass = LTR_gr$repClass[subjectHits(LTR_hits)],
  repName  = LTR_gr$repName[subjectHits(LTR_hits)],
  RM_id  = LTR_gr$RM_id[subjectHits(LTR_hits)],
  TE_type  = ifelse(
    LTR_gr$repFamily[subjectHits(LTR_hits)] == "SVA",
    LTR_gr$repName[subjectHits(LTR_hits)],
    LTR_gr$repFamily[subjectHits(LTR_hits)]
  ),
  
  milliDiv = LTR_gr$milliDiv[subjectHits(LTR_hits)],
  milliDel = LTR_gr$milliDel[subjectHits(LTR_hits)],
  milliIns = LTR_gr$milliIns[subjectHits(LTR_hits)]
)

LTR_overlap_df %>% 
  dplyr::left_join(H3K27ac_lookup) %>% 
  dplyr::filter(cluster %in% c("Set_1","Set_2","Set_3")) %>% 
  ggplot(., aes(x = TE_type, y = milliDiv)) +
   geom_violin(trim = FALSE) +
   geom_boxplot(width = 0.15, outlier.shape = NA) +
   coord_flip() +
   labs(
     y = "Divergence from consensus (milliDiv)",
     x = "TE family",
     title = "Age distribution of LTR elements overlapping H3K27ac"
   )+
  facet_wrap(~cluster)
```



### DNA
Overlapping DNA family with ROIs

```{r DNA Overlapping}
DNA_hits <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, DNA_gr, ignore.strand = TRUE)

DNA_overlap_df <- tibble(
  peak_row = queryHits(DNA_hits),
  Peakid   = H3K27ac_sets_gr$all_H3K27ac$Peakid[queryHits(DNA_hits)],
  cluster  = H3K27ac_sets_gr$all_H3K27ac$cluster[queryHits(DNA_hits)],
  
  repClass = DNA_gr$repClass[subjectHits(DNA_hits)],
  repName  = DNA_gr$repName[subjectHits(DNA_hits)],
  
  TE_type  = ifelse(
    DNA_gr$repFamily[subjectHits(DNA_hits)] == "SVA",
    DNA_gr$repName[subjectHits(DNA_hits)],
    DNA_gr$repFamily[subjectHits(DNA_hits)]
  ),
  
  milliDiv = DNA_gr$milliDiv[subjectHits(DNA_hits)],
  milliDel = DNA_gr$milliDel[subjectHits(DNA_hits)],
  milliIns = DNA_gr$milliIns[subjectHits(DNA_hits)]
)

DNA_overlap_df %>% 
  dplyr::left_join(H3K27ac_lookup) %>% 
  dplyr::filter(cluster %in% c("Set_1","Set_2","Set_3")) %>% 
  ggplot(., aes(x = TE_type, y = milliDiv)) +
   geom_violin(trim = FALSE) +
   geom_boxplot(width = 0.15, outlier.shape = NA) +
   coord_flip() +
   labs(
     y = "Divergence from consensus (milliDiv)",
     x = "TE family",
     title = "Age distribution of DNA elements overlapping H3K27ac"
   )+
  facet_wrap(~cluster)
```




### SVA/Retroposon
Overlapping SVA family with ROIs

```{r SVA Overlapping}
SVA_hits <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, SVA_gr, ignore.strand = TRUE)

SVA_overlap_df <- tibble(
  peak_row = queryHits(SVA_hits),
  Peakid   = H3K27ac_sets_gr$all_H3K27ac$Peakid[queryHits(SVA_hits)],
  cluster  = H3K27ac_sets_gr$all_H3K27ac$cluster[queryHits(SVA_hits)],
  
  repClass = SVA_gr$repClass[subjectHits(SVA_hits)],
  repName  = SVA_gr$repName[subjectHits(SVA_hits)],
  
  TE_type  = ifelse(
    SVA_gr$repFamily[subjectHits(SVA_hits)] == "SVA",
    SVA_gr$repName[subjectHits(SVA_hits)],
    SVA_gr$repFamily[subjectHits(SVA_hits)]
  ),
  
  milliDiv = SVA_gr$milliDiv[subjectHits(SVA_hits)],
  milliDel = SVA_gr$milliDel[subjectHits(SVA_hits)],
  milliIns = SVA_gr$milliIns[subjectHits(SVA_hits)]
)

SVA_overlap_df %>% 
  dplyr::left_join(H3K27ac_lookup) %>% 
  dplyr::filter(cluster %in% c("Set_1","Set_2","Set_3")) %>% 
  ggplot(., aes(x = TE_type, y = milliDiv)) +
   geom_violin(trim = FALSE) +
   geom_boxplot(width = 0.15, outlier.shape = NA) +
   coord_flip() +
   labs(
     y = "Divergence from consensus (milliDiv)",
     x = "TE family",
     title = "Age distribution of SVA elements overlapping H3K27ac"
   )+
  facet_wrap(~cluster)
```



### Looking for overlapping information: LTR specific  

Here I am exploring H3K27ac ROIs which overlap LTRs and also overlap H3K9me3 ROIs, then exploring LFC of the H3K27ac ROI sets across families of LTRs.  First I wanted to know how many H3K27ac ROIs overlap LTRs, and how many of those are more than one LTR:
```{r H3K9me3 overalapping regions}

LTR_overlap_df %>% 
  count(Peakid, name = "n_peaks") %>%
  count(n_peaks)

LTR_only_K27ac_peakids <- LTR_overlap_df %>% distinct(Peakid)
```

I then overlapped ROIs with H3K9me3 ROIs to see how many unique H3K27ac ROIs are involved.
```{r}

K9me3_hits <- findOverlaps(H3K27ac_sets_gr$all_H3K27ac, H3K9me3_sets_gr$all_H3K9me3_regions, ignore.strand = TRUE)

K9me3_overlap_df <- tibble(
  peak_row = queryHits(K9me3_hits),
  Peakid   = H3K27ac_sets_gr$all_H3K27ac$Peakid[queryHits(K9me3_hits)],
  cluster  = H3K27ac_sets_gr$all_H3K27ac$cluster[queryHits(K9me3_hits)],
  ROI_K9me3 = H3K9me3_sets_gr$all_H3K9me3_regions$Peakid[subjectHits(K9me3_hits)])

K9me3_overlap_df %>%
  count(Peakid, name = "n_peaks") %>%
  count(n_peaks)

K9me3_only_K27ac_peakids <- K9me3_overlap_df %>% distinct(Peakid)

```
Now to filter and connect H3K27ac ROIs, that overlap LTRs and at least one H3K9me3
```{r}
LTR_overlap_df %>% 
  dplyr::filter(Peakid %in% K9me3_only_K27ac_peakids$Peakid) %>%  distinct(RM_id)
```

#### LFC of those regions

```{r}
LTR_overlap_df %>% 
  dplyr::filter(Peakid %in% K9me3_only_K27ac_peakids$Peakid) %>% 
  
  left_join(., K27ac_lfctable, by = c("Peakid"="genes")) %>% 
  group_by(TE_type) %>% 
  tally() %>% 
   ungroup() %>% 
  DT::datatable(
    rownames = FALSE,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      "H3K27ac LTR specific overlap counts by TE familiy which also overlap H3K9me3"),
    options = list(pageLength = 10,
      autoWidth = TRUE,
      dom = "tip"))

```

```{r}
LTR_overlap_df %>% 
  dplyr::left_join(H3K27ac_lookup) %>% 
   dplyr::filter(Peakid %in% K9me3_only_K27ac_peakids$Peakid) %>% 
  dplyr::filter(cluster %in% c("Set_1","Set_2","Set_3")) %>% 
  ggplot(., aes(x = TE_type, y = milliDiv)) +
   geom_violin(trim = FALSE) +
   geom_boxplot(width = 0.15, outlier.shape = NA) +
   coord_flip() +
   labs(
     y = "Divergence from consensus (milliDiv)",
     x = "TE family",
     title = "Age distribution of LTR elements overlapping H3K27ac and H3K9me3"
   )+
  facet_wrap(~cluster)

LTR_K9me3_H3K27ac_lfc <- LTR_overlap_df %>% 
   left_join(., K27ac_lfctable, by = c("Peakid"="genes")) %>%
   dplyr::filter(Peakid %in% K9me3_only_K27ac_peakids$Peakid)

```
```{r}
plot_K27_family <- function(lfc_df,
                            grp="H3K27ac",
                            repFamily,
                            fam_name){

  lfc_df %>% 
  dplyr::filter(Peakid %in% K9me3_only_K27ac_peakids$Peakid) %>% 
    dplyr::filter(.data$TE_type %in% fam_name) %>% 
  dplyr::filter(cluster != "NA") %>% 
  pivot_longer(.,cols=c(starts_with(grp)), names_to="group", values_to = "LFC") %>% 
  group_by(cluster, group) %>%
  summarise(median_abs_LFC = median(abs(LFC), na.rm = TRUE),
            .groups = "drop")%>% 
  mutate(time=str_remove(group,grp))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R"))) %>% 
  ggplot(., aes(x=time, y=median_abs_LFC, group=cluster, color=cluster))+
  geom_point(size=4)+
  geom_line(aes(alpha = 0.8, linewidth = 4))+
  theme_bw()+
  ggtitle(paste0(grp," absolute LFC overlapping ",repFamily,":",fam_name))
}
```

```{r LTR lfc H3K9me3}
plot_K27_family(LTR_K9me3_H3K27ac_lfc,"H3K27ac","LTR","ERV1")+
  scale_color_manual(values = cluster_colors, drop = FALSE)
plot_K27_family(LTR_K9me3_H3K27ac_lfc,"H3K27ac","LTR","ERVK")+
  scale_color_manual(values = cluster_colors, drop = FALSE)
plot_K27_family(LTR_K9me3_H3K27ac_lfc,"H3K27ac","LTR","ERVL")+
  scale_color_manual(values = cluster_colors, drop = FALSE)
plot_K27_family(LTR_K9me3_H3K27ac_lfc,"H3K27ac","LTR","Gypsy")+
  scale_color_manual(values = cluster_colors, drop = FALSE)
plot_K27_family(LTR_K9me3_H3K27ac_lfc,"H3K27ac","LTR","ERVL-MaLR")+
  scale_color_manual(values = cluster_colors, drop = FALSE)
plot_K27_family(LTR_K9me3_H3K27ac_lfc,"H3K27ac","LTR","LTR")+
  scale_color_manual(values = cluster_colors, drop = FALSE)
```


#### Raw LFC exploration


This will look at Raw LFC for all H3K27ac ROIs that overlap an LTR

```{r}
LTR_overlap_df %>% 
  
   left_join(., K27ac_lfctable, by = c("Peakid"="genes")) %>% 
  pivot_longer(.,cols=c(starts_with("H3K27ac")), names_to="group", values_to = "LFC") %>% 
  group_by(cluster, group) %>%
  # summarise(median_LFC = median(LFC), na.rm = TRUE),
  #           .groups = "drop")%>% 
  dplyr::filter(cluster=="Set_2") %>% 
   mutate(time=str_remove(group,"H3K27ac"))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R")))  %>% 
  ggplot(., aes(x=time, y = LFC, group=Peakid))+
  geom_line(alpha = 0.3) +
  geom_point(size = 1) +
  theme_classic() +
  labs(
    x = "Time point",
    y = "Log2 Fold Change",
    title = "H3K27ac LFC  for Set_2 LTR-overlapping regions")

LTR_overlap_df %>% 
   left_join(., K27ac_lfctable, by = c("Peakid"="genes")) %>% 
  pivot_longer(.,cols=c(starts_with("H3K27ac")), names_to="group", values_to = "LFC") %>% 
  group_by(cluster, group) %>%
  # summarise(median_LFC = median(LFC), na.rm = TRUE),
  #           .groups = "drop")%>% 
  # dplyr::filter(cluster=="Set_3") %>% 
   mutate(time=str_remove(group,"H3K27ac"))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R")))  %>% 
  ggplot(., aes(x=time, y = LFC, group=Peakid, color=case))+
  geom_line(alpha = 0.3) +
  geom_point(size = 1) +
  theme_classic() +
  facet_wrap(~cluster)+
  labs(
    x = "Time point",
    y = "Log2 Fold Change",
    title = "H3K27ac LFC  for all LTR-overlapping regions")
```

Getting some number about how many, etc...


```{r}

LTR_overlap_df %>% 
  dplyr::left_join(H3K27ac_lookup) %>% 
  group_by(TE_type,cluster) %>% 
  tally() %>% 
  pivot_wider(id_cols = TE_type, names_from = cluster, values_from = n) %>% 
  ungroup() %>% 
  DT::datatable(
    rownames = FALSE,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      "H3K27ac LTR specific overlap counts by TE familiy"),
    options = list(pageLength = 10,
      autoWidth = TRUE,
      dom = "tip"))

```

now breaking out by family

```{r}


LTR_overlap_df %>% 
   left_join(., K27ac_lfctable, by = c("Peakid"="genes")) %>% 
  pivot_longer(.,cols=c(starts_with("H3K27ac")), names_to="group", values_to = "LFC") %>% 
  group_by(cluster, group) %>%
  # summarise(median_LFC = median(LFC), na.rm = TRUE),
  #           .groups = "drop")%>% 
  dplyr::filter(cluster=="Set_2") %>% 
   mutate(time=str_remove(group,"H3K27ac"))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R")))  %>% 
  ggplot(., aes(x=time, y = LFC, group=Peakid, color=case))+
  geom_line(alpha = 0.3) +
  geom_point(size = 1) +
  theme_classic() +
  facet_wrap(~TE_type)+
  labs(
    x = "Time point",
    y = "Log2 Fold Change",
    title = "H3K27ac LFC  for Set_2 LTR-overlapping regions")


LTR_overlap_df %>% 
  left_join(., K27ac_lfctable, by = c("Peakid"="genes")) %>% 
  pivot_longer(.,cols=c(starts_with("H3K27ac")), names_to="group", values_to = "LFC") %>% 
  group_by(cluster, group) %>%
  # summarise(median_LFC = median(LFC), na.rm = TRUE),
  #           .groups = "drop")%>% 
  dplyr::filter(cluster=="Set_3") %>% 
   mutate(time=str_remove(group,"H3K27ac"))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R")))  %>% 
  ggplot(., aes(x=time, y = LFC, group=Peakid, color=case))+
  geom_line(alpha = 0.3) +
  geom_point(size = 1) +
  theme_classic() +
  facet_wrap(~TE_type)+
  labs(
    x = "Time point",
    y = "Log2 Fold Change",
    title = "H3K27ac LFC  for Set_3 LTR-overlapping regions")
```
There appears to be a lot of variation on LFC.  Direction is obscuring information.  I am now attempting to create a list of ROIs by set where LFC across time is always >0 (The up group), another set where LFC across time is always <0 (The down group), and the catch all mixed group. 

```{r split groups by LFC}
### already added to the K27ac_lfctable
# LTR_overlap_df %>% 
#   dplyr::left_join(H3K27ac_lookup) %>% 
#    left_join(., K27ac_lfctable, by = c("Peakid"="genes")) %>% 
#    mutate(case= case_when(H3K27ac_24T > 0 & H3K27ac_24R >0 & H3K27ac_144R>0~"A",
#                               H3K27ac_24T < 0 & H3K27ac_24R <0 & H3K27ac_144R<0~"B",
#                               TRUE ~ "C")) %>% 
#   mutate(set_case=paste0(cluster,"_",case))
median_df <- K27ac_lfctable %>% 
  pivot_longer(.,cols=c(starts_with("H3K27ac")), names_to="group", values_to = "LFC") %>% 
  mutate(
    time = stringr::str_remove(group, "H3K27ac"),
    time = factor(time, levels = c("_24T", "_24R", "_144R"))
  ) %>%
  group_by(cluster, time) %>% 
  summarise(
    median_LFC = median(LFC, na.rm = TRUE),
    n_peaks = dplyr::n(),
    .groups = "drop")
  
K27ac_lfctable %>% 
  pivot_longer(.,cols=c(starts_with("H3K27ac")), names_to="group", values_to = "LFC") %>% 
   mutate(time=str_remove(group,"H3K27ac"))%>% 
  mutate(time=factor(time, levels=c("_24T","_24R","_144R")))  %>% 
  ggplot(., aes(x=time, y = LFC, group=genes, color=case))+
  ggrastr::rasterise(geom_line(alpha = 0.15)) +
  ggrastr::geom_point_rast(size = 1) +
  geom_line(data=median_df, aes(x=time, y=median_LFC, group = cluster),inherit.aes = FALSE, linewidth =2)+
  theme_classic() +
  facet_wrap(~cluster)+
  labs(
    x = "Time point",
    y = "Log2 Fold Change",
    title = "H3K27ac LFC  for all H3K27ac regions")
```
```{r}
K27ac_lfctable %>% 
  group_by(cluster,case) %>% 
  tally() %>% pivot_wider(id_cols=case, names_from = cluster, values_from = n) %>% 
  mutate(type=case_when(case=="A"~"All LFC > 0",
                        case=="B" ~ "All LFC < 0",
                        case=="C"~"Mixed up and down LFC")) %>% 
  DT::datatable(
    rownames = FALSE,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      "H3K27ac ROI counts stratified by LFC case"),
    options = list(pageLength = 10,
      autoWidth = TRUE,
      dom = "tip"))
```


```{r export set_case list, eval=FALSE}

```

