---
title: "H3K27ac TF enrichment"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```


```{r packages}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(DT)

```

### Summit calls method 1  
This is actually summit +/- 200 bp (and 300bp for initial investigation)

```{r loading motif data Set2}

H3K27ac_Set2_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set2_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set2_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


H3K27ac_Set2_tomtom <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400/streme_tomtom_out/tomtom.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

test_bind <- bind_rows(H3K27ac_Set2_sea_disc_out, H3K27ac_Set2_sea_known) %>% 
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID, ENR_RATIO) %>% distinct()


```

```{r loading motif data Set3}

H3K27ac_Set3_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set3_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set3_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


H3K27ac_Set3_tomtom <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400/streme_tomtom_out/tomtom.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

test_bind3 <-bind_rows(H3K27ac_Set3_sea_disc_out, H3K27ac_Set3_sea_known) %>%
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID) %>% 
  distinct()


```


```{r  making data sets}
H3K27ac_Set2_data <- H3K27ac_Set2_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(test_bind) %>% 
  mutate(diff_per=`TP%` -`FP%`)

H3K27ac_Set3_data <- H3K27ac_Set3_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(test_bind3, by=c("ID"="ID","ALT_ID"="ALT_ID")) %>% 
  mutate(diff_per=`TP%` -`FP%`)

# saveRDS(H3K27ac_Set2_data,"data/RDS_files/H3K27ac_Set2_enrmotif_data.RDS")
# saveRDS(H3K27ac_Set3_data,"data/RDS_files/H3K27ac_Set3_enrmotif_data.RDS")
```


```{r}

H3K27ac_Set2_data %>% 
  dplyr::filter(ENR_RATIO>1.2) %>% 
  group_by(CLUSTER) %>% 
  slice_min(EVALUE, with_ties = FALSE) %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*2), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./2,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET2 enrichment over SET1, Enrichment ratio >1.2"))

H3K27ac_Set3_data %>% 
  dplyr::filter(ENR_RATIO>1.2) %>%
  group_by(CLUSTER) %>% 
  slice_min(EVALUE, with_ties = FALSE) %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*1), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./1,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 enrichment over SET1, Enrichment ratio > 1.2"))

```

```{r enrichment by cluster}

H3K27ac_Set2_data %>% 
  dplyr::filter(ENR_RATIO>1.2) %>% 
  group_by(CLUSTER) %>% 
  slice_min(EVALUE, with_ties = FALSE) %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*2), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./2,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET2 enrichment over SET1, Enrichment ratio >1.2"))

H3K27ac_Set3_data %>% 
  dplyr::filter(ENR_RATIO>1.1) %>%
  group_by(CLUSTER) %>% 
  slice_min(EVALUE, with_ties = FALSE) %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*1), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./1,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 enrichment over SET1, Enrichment ratio > 1.1"))

```  



```{r enrichment ratio dispersion}
H3K27ac_Set2_data %>% 
  ggplot(., aes (x= ENR_RATIO)) +
  geom_density()+
  # geom_point(aes(x=`TP%`*2), size =4)+
  # scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./2,name="Percent of peaks with motif"))+
  theme_classic()+
  # ylab("Enriched TF motif")+
  ggtitle(paste("Span of enrichment ratios for Set2"))


H3K27ac_Set3_data %>% 
  ggplot(., aes (x= ENR_RATIO)) +
  geom_density()+
  theme_classic()+
  ggtitle(paste("Span of enrichment ratios for Set3"))


H3K27ac_Set3_data %>% 
  ggplot(., aes (x= ENR_RATIO)) +
  geom_density()+
  theme_classic()+
  ggtitle(paste("Span of enrichment ratios for Set3"))+
  coord_cartesian(xlim=c(0,6))
```


```{r}

ggVennDiagram::ggVennDiagram(list("H3K27ac_Set2"=H3K27ac_Set2_data$motif_name,"H3K27ac_Set3"=H3K27ac_Set3_data$motif_name))+
  ggtitle("H3K27ac motif overlap")+
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = margin(20, 60, 20, 60)  # top, right, bottom, left
  )
  
```

```{r method1 list}
# test_bind
# xstreme_rank_bind
DT::datatable((H3K27ac_Set2_data %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 1: Set 2 H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 

DT::datatable((H3K27ac_Set3_data %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS.x)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 1: Set 3 H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 


```

```{r set 2 by facet}
H3K27ac_Set2_data %>% 
dplyr::filter(EVALUE<0.05) %>% 
   group_by(CLUSTER) %>% 
  slice_min(RANK, with_ties = FALSE) %>% 
  ungroup() %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>%
    mutate(Facets=case_when(RANK <15~"A",
                              log10Evalue >15 & log10Evalue <40~"B",
            log10Evalue>40~"C")) %>% 
  dplyr::filter(Facets=="A") %>% 
    ggplot(., aes (y= reorder(motif_name, log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
   geom_point(aes(x=ENR_RATIO*40, color=ENR_RATIO), size =4)+
  scale_x_continuous(expand = c(0, .125),
  sec.axis = sec_axis(~ . /40, name = "Enrichment Ratio"))+
 
  facet_wrap(~Facets )+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET2 enrichment over SET1\n top Enrichment Ratio by Cluster H3K27ac"))



H3K27ac_Set2_data %>% 
dplyr::filter(EVALUE<0.05) %>% 
   group_by(CLUSTER) %>% 
  slice_min(RANK, with_ties = FALSE) %>% 
  ungroup() %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>%
    mutate(Facets=case_when(RANK <15~"A",
                              log10Evalue >15 & log10Evalue <40~"B",
            log10Evalue>40~"C")) %>% 
  dplyr::filter(Facets=="B") %>% 
     ggplot(., aes (y= reorder(motif_name, log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=ENR_RATIO*10, color=ENR_RATIO), size =4)+
  
  scale_x_continuous(
  expand = c(0, .125),
  sec.axis = sec_axis(~ . / 10, name = "Enrichment Ratio")
)+
  scale_color_viridis_c(
    name = "Enrichment Ratio",
    option = "plasma") +
  tidytext::scale_y_reordered()+
  facet_wrap(~Facets )+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET2 enrichment over SET1\n top Enrichment Ratio by Cluster H3K27ac"))



H3K27ac_Set2_data %>% 
dplyr::filter(EVALUE<0.05) %>% 
   group_by(CLUSTER) %>% 
  slice_min(RANK, with_ties = FALSE) %>% 
  ungroup() %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>%
     mutate(Facets=case_when(RANK <15~"A",
                              log10Evalue >15 & log10Evalue <40~"B",
            log10Evalue>40~"C")) %>% 
  dplyr::filter(Facets=="C") %>% 
     ggplot(., aes (y= reorder(motif_name, log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=ENR_RATIO*.8, color=ENR_RATIO), size =4)+
  scale_x_continuous(
  expand = c(0, .125),
  sec.axis = sec_axis(~ . / .8, name = "Enrichment Ratio")
)+
  scale_color_viridis_c(
    name = "Enrichment Ratio",
    option = "plasma") +
  facet_wrap(~Facets )+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET2 enrichment over SET1\n top Enrichment Ratio by Cluster H3K27ac"))
```

```{r set 3 by facet}
H3K27ac_Set3_data %>% 
dplyr::filter(EVALUE<0.05) %>% 
   group_by(CLUSTER) %>% 
  slice_min(RANK, with_ties = FALSE) %>% 
  ungroup() %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>%
    mutate(Facets=case_when(log10Evalue >10~"A",
                              log10Evalue <10 & log10Evalue >5~"B",
            log10Evalue<5~"C")) %>% 
  dplyr::filter(Facets=="A") %>% 
    ggplot(., aes (y= reorder(motif_name, log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=ENR_RATIO*10, color=ENR_RATIO), size =4)+
  geom_vline(
  xintercept = 1*10,
  linetype = "dotted",
  linewidth = 1, color="red")+
  scale_x_continuous(
  expand = c(0, .125),
  sec.axis = sec_axis(~ . / 10, name = "Enrichment Ratio")
)+
   scale_color_viridis_c(
    name = "Enrichment Ratio",
    option = "plasma") +
  tidytext::scale_y_reordered()+
  facet_wrap(~Facets )+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 enrichment over SET1\n top Enrichment Ratio by Cluster H3K27ac"))



H3K27ac_Set3_data %>% 
  dplyr::filter(EVALUE<0.05) %>% 
   group_by(CLUSTER) %>% 
  slice_min(RANK, with_ties = FALSE) %>% 
  ungroup() %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>%
    mutate(Facets=case_when(log10Evalue >10~"A",
                              log10Evalue <10 & log10Evalue >5~"B",
            log10Evalue<5~"C")) %>% 
  dplyr::filter(Facets=="B") %>% 
    ggplot(., aes (y= tidytext::reorder_within(motif_name, log10Evalue, Facets))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=ENR_RATIO*5, color=ENR_RATIO), size =4)+
  geom_vline(
  xintercept = 1*5,
  linetype = "dotted",
  linewidth = 1, color="red")+
  scale_x_continuous(
  expand = c(0, .125),
  sec.axis = sec_axis(~ . / 5, name = "Enrichment Ratio")
)+ scale_color_viridis_c(
    name = "Enrichment Ratio",
    option = "plasma"
  ) +
  tidytext::scale_y_reordered()+
  facet_wrap(~Facets )+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 enrichment over SET1\n top Enrichment Ratio by Cluster H3K27ac"))



H3K27ac_Set3_data %>% 
dplyr::filter(EVALUE<0.05) %>% 
   group_by(CLUSTER) %>% 
  slice_min(RANK, with_ties = FALSE) %>% 
  ungroup() %>% 
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>%
    mutate(Facets=case_when(log10Evalue >10~"A",
                              log10Evalue <10 & log10Evalue >5~"B",
            log10Evalue<5~"C")) %>% 
  dplyr::filter(Facets=="C") %>% 
    ggplot(., aes (y= tidytext::reorder_within(motif_name, log10Evalue, Facets))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=ENR_RATIO*.05,color= pmin(ENR_RATIO, 5)), size =4)+
  scale_x_continuous(
  expand = c(0, .125),
  sec.axis = sec_axis(~ . /.05, name = "Enrichment Ratio")
)+
   scale_color_viridis_c(
    name = "Enrichment Ratio",
    option = "plasma"
  ) +
  tidytext::scale_y_reordered()+
  facet_wrap(~Facets )+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 enrichment over SET1\n top Enrichment Ratio by Cluster H3K27ac"))


```
```{r}
plot_topnum_ratio_full <- function(df, title = NULL, signif.num = 0.05,top_num=20, color_limits = NULL){
  
  ### getting the name of the data frame for plotting
 if (is.null(title)) {
    title <- deparse(substitute(df))
 }
  
   

  ### Getting the data ready
  plot_df <- df %>%
    dplyr::filter(EVALUE < signif.num) %>%
    group_by(CLUSTER) %>%
    slice_min(RANK, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(log10Evalue = -log10(EVALUE)) %>%
    arrange(desc(log10Evalue)) %>%
    mutate(motif_name_wrapped = stringr::str_wrap(motif_name, width = 10)) %>% 
     mutate(
    enr_shape = ifelse(ENR_RATIO > 1, "Enriched", "Depleted")) %>% 
    slice_head(., n=top_num)
#   
  scaler <- max(plot_df$log10Evalue) / max(plot_df$ENR_RATIO)

  ##plotting the data
   ggplot(plot_df, aes(y = reorder(motif_name_wrapped, log10Evalue))) +
    geom_col(aes(x = log10Evalue)) +
    geom_point(aes(x = ENR_RATIO* scaler, color = ENR_RATIO, shape = enr_shape), size = 4)+
    scale_x_continuous(
      expand = c(0, .125),
      sec.axis = sec_axis(~ . / scaler, name = "Enrichment Ratio")
    ) +
   scale_color_viridis_c(
  name = "Enrichment Ratio",
  option = "plasma",
  limits = color_limits,
  oob = scales::squish
)+
    scale_y_discrete(labels = function(x) stringr::str_wrap(gsub("__.*$", "", x), width = 10)) +
     scale_shape_manual(values = c("Enriched" = 16, "Depleted" = 17))+
    theme_classic() +
    ylab("Enriched TF motif") +
    ggtitle(paste(title, "-", top_num))
}

```

```{r functional full plot, fig.width=15}
plot_A <- plot_topnum_ratio_full(H3K27ac_Set3_data,signif.num = 0.05, top_num = 10)

plot_B <- plot_topnum_ratio_full(H3K27ac_Set2_data,signif.num = 0.05, top_num = 10)

aligned_plots <- cowplot::align_plots(plot_A, plot_B, align = "v", axis = "l")

cowplot::plot_grid(aligned_plots[[1]],
          aligned_plots[[2]],
          ncol = 2)
```


### RNA filtering  

```{r RNA filtering, fig.width=15}
RNA_full_toptable <-  readRDS("data/Other_paper_data/RNA_full_toptable.RDS")

RNA_expressed <- RNA_full_toptable %>% 
  dplyr::select(Entrez_ID, SYMBOL)

### had to add in a special filter to account for the :: in the names of the motifs
symbols <- str_to_upper(RNA_expressed$SYMBOL)

filtered_H3K27ac_Set3_data <- H3K27ac_Set3_data %>%
  ##create the motif-tokens column so I can keep names where there is a partial match in the RNA Symbols
  mutate(motif_tokens = str_split(str_to_upper(motif_name), "::")) %>%
  ##Because this is a list, now I need to filter using sapply, but then remove the motif token column
  filter(sapply(motif_tokens, function(x) any(x %in% symbols))) %>%
  dplyr::select(-motif_tokens)


filtered_H3K27ac_Set2_data <- H3K27ac_Set2_data %>%
  mutate(motif_tokens = str_split(str_to_upper(motif_name), "::")) %>%
  filter(sapply(motif_tokens, function(x) any(x %in% symbols))) %>%
  dplyr::select(-motif_tokens)


## binding both together to get them to set a global limit on the secondary axis using quantiles so color range is identical
global_combo <- bind_rows(filtered_H3K27ac_Set2_data, filtered_H3K27ac_Set3_data)

global_limits <-  quantile(global_combo$ENR_RATIO, probs = c(0.05, 0.95), na.rm = TRUE)
### plot seperately
  filt_A <- plot_topnum_ratio_full(filtered_H3K27ac_Set3_data,signif.num = 0.05, top_num = 10,color_limits = global_limits)
  
  filt_B <- plot_topnum_ratio_full(filtered_H3K27ac_Set2_data,signif.num = 0.05, top_num = 10,color_limits = global_limits)
 ## Good ole cowplot to align plots vertically 
  aligned_filt_plots <- cowplot::align_plots(filt_B, filt_A, align = "v", axis = "l")

cowplot::plot_grid(aligned_filt_plots[[1]],
          aligned_filt_plots[[2]],
          ncol = 2)

```

#### Just KZFPs in motifs  


```{r}
expressed_KZFPs <- readRDS("data/RDS_files/expressed_KZFPs.RDS")
filtered_H3K27ac_Set2_data %>% 
   mutate(log10Evalue = -log10(EVALUE)) %>%
    arrange(desc(log10Evalue)) %>%
  mutate( enr_shape = ifelse(ENR_RATIO > 1, "Enriched", "Depleted")) %>%
  dplyr::filter(ALT_ID%in%expressed_KZFPs) %>% 
ggplot(., aes(y = reorder(motif_name, log10Evalue))) +
    geom_col(aes(x = log10Evalue)) +
    geom_point(aes(x = ENR_RATIO* 10, color = ENR_RATIO, shape = enr_shape), size = 4)+
    scale_x_continuous(
      expand = c(0, .125),
      sec.axis = sec_axis(~ . /10, name = "Enrichment Ratio")
    ) +
   scale_color_viridis_c(
  name = "Enrichment Ratio",
  option = "plasma",
 
  oob = scales::squish
)+
  ggtitle("All expressed KZFPs in TF data H3K27ac Sustained set")


filtered_H3K27ac_Set3_data %>% 
   mutate(log10Evalue = -log10(EVALUE)) %>%
    arrange(desc(log10Evalue)) %>%
  mutate( enr_shape = ifelse(ENR_RATIO > 1, "Enriched", "Depleted")) %>%
  dplyr::filter(ALT_ID%in%expressed_KZFPs) %>% 
ggplot(., aes(y = reorder(motif_name, log10Evalue))) +
    geom_col(aes(x = log10Evalue)) +
    geom_point(aes(x = ENR_RATIO* 5, color = ENR_RATIO, shape = enr_shape), size = 4)+
    scale_x_continuous(
      expand = c(0, .125),
      sec.axis = sec_axis(~ . /5, name = "Enrichment Ratio")
    ) +
   scale_color_viridis_c(
  name = "Enrichment Ratio",
  option = "plasma",
 
  oob = scales::squish
)+
  ggtitle("All expressed KZFPs in TF data H3K27ac Acute set")
# 
# filter(motif_name=="ZNF384")
# H3K27ac_Set2_data %>% distinct(motif_name)
```
#### Marking KZFPs in motifs

```{r}
KZFP_2026_list <- read_excel("data/Other_paper_data/human_protein_coding_KZFPs_Davis_natgen_2026.xlsx")

filtered_H3K27ac_Set2_data <- filtered_H3K27ac_Set2_data %>% 
  mutate(KZFP=if_else(motif_name %in%KZFP_2026_list$Label,"KZFP","not_KZFP"))
filtered_H3K27ac_Set3_data <- filtered_H3K27ac_Set3_data %>% 
  mutate(KZFP=if_else(motif_name %in%KZFP_2026_list$Label,"KZFP","not_KZFP"))

plot_topnum_ratio_KZFP <- function(df, title = NULL, signif.num = 0.05,top_num=20, color_limits = NULL){
  
  ### getting the name of the data frame for plotting
 if (is.null(title)) {
    title <- deparse(substitute(df))
 }
  
   

  ### Getting the data ready
  plot_df <- df %>%
    dplyr::filter(EVALUE < signif.num) %>%
    group_by(CLUSTER) %>%
    slice_min(RANK, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(log10Evalue = -log10(EVALUE)) %>%
    arrange(desc(log10Evalue)) %>%
    mutate(motif_name_wrapped = stringr::str_wrap(motif_name, width = 10)) %>% 
     mutate(
    enr_shape = ifelse(ENR_RATIO > 1, "Enriched", "Depleted")) %>% 
    slice_head(., n=top_num)
#   
  scaler <- max(plot_df$log10Evalue) / max(plot_df$ENR_RATIO)

  ##plotting the data
   ggplot(plot_df, aes(y = reorder(motif_name_wrapped, log10Evalue))) +
    geom_col(aes(x = log10Evalue, fill=KZFP)) +
    geom_point(aes(x = ENR_RATIO* scaler, color = ENR_RATIO, shape = enr_shape), size = 4)+
    scale_x_continuous(
      expand = c(0, .125),
      sec.axis = sec_axis(~ . / scaler, name = "Enrichment Ratio")
    ) +
   scale_color_viridis_c(
  name = "Enrichment Ratio",
  option = "plasma",
  limits = color_limits,
  oob = scales::squish
)+
    scale_y_discrete(labels = function(x) stringr::str_wrap(gsub("__.*$", "", x), width = 10)) +
     scale_shape_manual(values = c("Enriched" = 16, "Depleted" = 17))+
    theme_classic() +
    ylab("Enriched TF motif") +
    ggtitle(paste(title, "-", top_num))
}
```


```{r, fig.width=10, fig.height=6}


## binding both together to get them to set a global limit on the secondary axis using quantiles so color range is identical
global_combo2 <- bind_rows(filtered_H3K27ac_Set2_data, filtered_H3K27ac_Set3_data)

global_limits2 <-  quantile(global_combo2$ENR_RATIO, probs = c(0.05, 0.95), na.rm = TRUE)
### plot seperately
  filt_A2 <- plot_topnum_ratio_KZFP(filtered_H3K27ac_Set3_data,signif.num = 0.05, top_num = 10,color_limits = global_limits2)
  
  filt_B2 <- plot_topnum_ratio_KZFP(filtered_H3K27ac_Set2_data,signif.num = 0.05, top_num = 10,color_limits = global_limits2)
 ## Good ole cowplot to align plots vertically 
  aligned_filt_plots <- cowplot::align_plots(filt_B2, filt_A2, align = "v", axis = "l")

cowplot::plot_grid(aligned_filt_plots[[1]],
          aligned_filt_plots[[2]],
          ncol = 2)
```




### Case and Summit calls method 1, subset A and B

This section looks at the enrichment of A and B cases of each Set.  
Case A is defined as all ROIs with an LFC > 0 at each timepoint.
Case B is defined as all ROIs with an LFC < 0 at each timepoint.
Case C is defined as all other ROIs.
The List of ROI summits was filtered for ROI that fit these cases, and enrichment was done using xstreme on TACC for each set-case compared to all of set1.


Because I have so many sets, I am now creating a function to import/visualize each section:

```{r import function}
import_function <- function(set_description, base_dir = "C:/Users/renee/Other_projects_data/DXR_data/H3K27ac") {
  
  sea_disc_out <-read_delim(
    file.path(base_dir, set_description, "sea_disc_out", "sea.tsv"),
    delim = "\t", trim_ws = TRUE
  ) %>%
    mutate(source = "disc") %>%
    slice(1:(dplyr::n() - 3))

  sea_known <-read_delim(
    file.path(base_dir, set_description, "sea_out", "sea.tsv"),
    delim = "\t", trim_ws = TRUE
  ) %>%
    mutate(source = "known") %>%
    slice(1:(dplyr::n() - 3))


  xstreme <- read_delim(
    file.path(base_dir, set_description, "xstreme.tsv"),
    delim = "\t", trim_ws = TRUE
  ) %>%
    mutate(source = "all") %>%
    slice(1:(dplyr::n() - 3))


 #### adding in later fuctions
  tomtom <- read_delim(
    file.path(base_dir, set_description, "streme_tomtom_out", "tomtom.tsv"),
    delim = "\t", trim_ws = TRUE
  )
### function to renormalize dataframes before the join
  
  clean_consensus <- function(x) {
  stringr::str_remove(x, "^\\d+-")
}
bind_df <- bind_rows(sea_disc_out, sea_known) %>% 
   mutate(CONSENSUS = clean_consensus(CONSENSUS)) %>%
  dplyr::select(DB:ENR_RATIO) %>% 
  distinct(ID, ALT_ID, ENR_RATIO, .keep_all = TRUE)



set_data <- xstreme %>% 
   mutate(CONSENSUS = clean_consensus(CONSENSUS)) %>%
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(bind_df,by = c("ALT_ID", 
                           "ID","CONSENSUS")) %>% 
  mutate(diff_per=`TP%` -`FP%`)

return(list(
  sea = bind_df,
  xstreme = set_data))

}

```

```{r looking at datatables}
H3K27ac_set2_A<- import_function("H3K27ac_set2_A_400",base_dir = "C:/Users/renee/Other_projects_data/DXR_data/H3K27ac")

H3K27ac_set2_B<- import_function("H3K27ac_set2_B_400",base_dir = "C:/Users/renee/Other_projects_data/DXR_data/H3K27ac")
H3K27ac_set3_A<- import_function("H3K27ac_set3_A_400",base_dir = "C:/Users/renee/Other_projects_data/DXR_data/H3K27ac")
H3K27ac_set3_B<- import_function("H3K27ac_set3_B_400",base_dir = "C:/Users/renee/Other_projects_data/DXR_data/H3K27ac")

DT::datatable((H3K27ac_set2_A$xstreme %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 1: Set 2A H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 

###Code below for testing issues in the function.  this helped fix the consensus .x .y issue.
# all(rest_test$xstreme$CONSENSUS.x == rest_test$xstreme$CONSENSUS.y, na.rm = TRUE)
# 
# rest_test$xstreme%>%
#   summarise(
#     n_total = dplyr::n(),
#     n_match = sum(CONSENSUS.x == CONSENSUS.y, na.rm = TRUE),
#     n_diff  = sum(CONSENSUS.x != CONSENSUS.y, na.rm = TRUE)
#   )

DT::datatable((H3K27ac_set2_B$xstreme %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 1: Set 2B H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 

DT::datatable((H3K27ac_set3_A$xstreme %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 1: Set 3A H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 

DT::datatable((H3K27ac_set3_B$xstreme %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 1: Set 3B H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 

```

```{r visualization plots function}

plot_ENR_RATIO <- function(df, title = NULL, signif.num = 0.05, page = 1){
  
  ### getting the name of the data frame for plotting
 if (is.null(title)) {
    title <- deparse(substitute(df))
  }

  ### Getting the data ready
  plot_df <- df$xstreme %>%
    dplyr::filter(EVALUE < signif.num) %>%
    group_by(CLUSTER) %>%
    slice_min(RANK, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(log10Evalue = -log10(EVALUE)) %>%
    arrange(desc(log10Evalue)) %>%
    mutate(Facets = cut(
      log10Evalue,
      breaks = quantile(log10Evalue, probs = c(0, 0.33, 0.66, 1), na.rm = TRUE),
      labels = c("Low", "Medium", "High"),
      include.lowest = TRUE
    )) %>% 
    mutate(motif_name_wrapped = stringr::str_wrap(motif_name, width = 10)) %>% 
     mutate(
    enr_shape = ifelse(ENR_RATIO > 1, "Enriched", "Depleted")
  )
  
  scaler <- max(plot_df$log10Evalue) / max(plot_df$ENR_RATIO)
  facet_levels <- levels(plot_df$Facets)
if(page > length(facet_levels)) stop("page number exceeds available facets")
facet_name <- facet_levels[page]
  ##plotting the data
   ggplot(plot_df, aes(y = tidytext::reorder_within(motif_name, log10Evalue, Facets))) +
    geom_col(aes(x = log10Evalue)) +
    geom_point(aes(x = ENR_RATIO * scaler, color = ENR_RATIO, shape = enr_shape), size = 4)+
    scale_x_continuous(
      expand = c(0, .125),
      sec.axis = sec_axis(~ . / scaler, name = "Enrichment Ratio")
    ) +
    scale_color_viridis_c(name = "Enrichment Ratio", option = "plasma") +
    scale_y_discrete(labels = function(x) stringr::str_wrap(gsub("__.*$", "", x), width = 10)) +
    ggforce::facet_wrap_paginate(
      ~Facets,
      ncol = 1,
      nrow = 1,
      page = page,
      scales = "free_y"
    ) +
     scale_shape_manual(values = c("Enriched" = 16, "Depleted" = 17))+
    theme_classic() +
    ylab("Enriched TF motif") +
    ggtitle(paste(title, "-", facet_name))
}
########################################################
plot_topnum_RATIO <- function(df, title = NULL, signif.num = 0.05,top_num=20){
  
  ### getting the name of the data frame for plotting
 if (is.null(title)) {
    title <- deparse(substitute(df))
  }

  ### Getting the data ready
  plot_df <- df$xstreme %>%
    dplyr::filter(EVALUE < signif.num) %>%
    group_by(CLUSTER) %>%
    slice_min(RANK, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(log10Evalue = -log10(EVALUE)) %>%
    arrange(desc(log10Evalue)) %>%
    mutate(motif_name_wrapped = stringr::str_wrap(motif_name, width = 10)) %>% 
     mutate(
    enr_shape = ifelse(ENR_RATIO > 1, "Enriched", "Depleted")) %>% 
    slice_head(., n=top_num)
#   
  scaler <- max(plot_df$log10Evalue) / max(plot_df$ENR_RATIO)

  ##plotting the data
   ggplot(plot_df, aes(y = reorder(motif_name, log10Evalue))) +
    geom_col(aes(x = log10Evalue)) +
    geom_point(aes(x = ENR_RATIO* scaler, color = ENR_RATIO, shape = enr_shape), size = 4)+
    scale_x_continuous(
      expand = c(0, .125),
      sec.axis = sec_axis(~ . / scaler, name = "Enrichment Ratio")
    ) +
    scale_color_viridis_c(name = "Enrichment Ratio", option = "plasma") +
    scale_y_discrete(labels = function(x) stringr::str_wrap(gsub("__.*$", "", x), width = 10)) +
     scale_shape_manual(values = c("Enriched" = 16, "Depleted" = 17))+
    theme_classic() +
    ylab("Enriched TF motif") +
    ggtitle(paste(title, "-", top_num))
}


```

```{r}
plot_ENR_RATIO(H3K27ac_set2_A, signif.num = 0.05, page = 3)
plot_ENR_RATIO(H3K27ac_set2_A, signif.num = 0.05, page = 2)
plot_ENR_RATIO(H3K27ac_set2_A, signif.num = 0.05, page = 1)

plot_ENR_RATIO(H3K27ac_set2_B, signif.num = 0.05, page = 3)
plot_ENR_RATIO(H3K27ac_set2_B, signif.num = 0.05, page = 2)
plot_ENR_RATIO(H3K27ac_set2_B, signif.num = 0.05, page = 1)

# plot_topnum_RATIO(H3K27ac_set2_A, signif.num = 0.05, top_num = 10)
# 
# plot_topnum_RATIO(H3K27ac_Set3_data,signif.num = 0.05, top_num = 10)
```
now for plotting Set 3 A and B

```{r plotting set3}
plot_ENR_RATIO(H3K27ac_set3_A, signif.num = 0.05, page = 3)
plot_ENR_RATIO(H3K27ac_set3_A, signif.num = 0.05, page = 2)
plot_ENR_RATIO(H3K27ac_set3_A, signif.num = 0.05, page = 1)

plot_ENR_RATIO(H3K27ac_set3_B, signif.num = 0.05, page = 3)
plot_ENR_RATIO(H3K27ac_set3_B, signif.num = 0.05, page = 2)
plot_ENR_RATIO(H3K27ac_set3_B, signif.num = 0.05, page = 1)
```


### Summit calls method 2  

```{r loading motif data Set2 met2}

H3K27ac_Set2_met2_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400_alt/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set2_met2_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400_alt/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set2_met2_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400_alt/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


H3K27ac_Set2_met2_tomtom <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400_alt/streme_tomtom_out/tomtom.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

test_bind_met2 <- bind_rows(H3K27ac_Set2_met2_sea_disc_out, H3K27ac_Set2_met2_sea_known) %>% 
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID, ENR_RATIO) %>% distinct()


```

```{r loading motif data Set3_met2}

H3K27ac_Set3_met2_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400_alt/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set3_met2_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400_alt/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set3_met2_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400_alt/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


H3K27ac_Set3_met2_tomtom <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400_alt/streme_tomtom_out/tomtom.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

test_bind3_met2 <-bind_rows(H3K27ac_Set3_met2_sea_disc_out, H3K27ac_Set3_met2_sea_known) %>%
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID) %>% 
  distinct()


```


```{r  making data sets met2}
H3K27ac_Set2_data_met2 <- H3K27ac_Set2_met2_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(test_bind_met2) %>% 
  mutate(diff_per=`TP%` -`FP%`)

H3K27ac_Set3_data_met2 <- H3K27ac_Set3_met2_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(test_bind3_met2, by=c("ID"="ID","ALT_ID"="ALT_ID")) %>% 
  mutate(diff_per=`TP%` -`FP%`)

```

```{r met2 list}
# test_bind
# xstreme_rank_bind
DT::datatable((H3K27ac_Set2_data_met2 %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 2: Set 2 H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 

DT::datatable((H3K27ac_Set3_data_met2 %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS.x)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 2: Set 3 H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 


```


### Summit calls method 3
```{r loading motif data Set2 met3}

H3K27ac_Set2_met3_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400_alt_concat/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set2_met3_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400_alt_concat/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set2_met3_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400_alt_concat/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


H3K27ac_Set2_met3_tomtom <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set2_400_alt_concat/streme_tomtom_out/tomtom.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

test_bind_met3 <- bind_rows(H3K27ac_Set2_met3_sea_disc_out, H3K27ac_Set2_met3_sea_known) %>% 
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID, ENR_RATIO) %>% distinct()


```

```{r loading motif data Set3_met3}

H3K27ac_Set3_met3_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400_alt_concat/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set3_met3_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400_alt_concat/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

H3K27ac_Set3_met3_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400_alt_concat/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


H3K27ac_Set3_met3_tomtom <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/H3K27ac/H3K27ac_set3_400_alt_concat/streme_tomtom_out/tomtom.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

test_bind3_met3 <-bind_rows(H3K27ac_Set3_met3_sea_disc_out, H3K27ac_Set3_met3_sea_known) %>%
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID) %>% 
  distinct()


```


```{r  making data sets met3}
H3K27ac_Set2_data_met3 <- H3K27ac_Set2_met3_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(test_bind_met3) %>% 
  mutate(diff_per=`TP%` -`FP%`)

H3K27ac_Set3_data_met3 <- H3K27ac_Set3_met3_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(test_bind3_met3, by=c("ID"="ID","ALT_ID"="ALT_ID")) %>% 
  mutate(diff_per=`TP%` -`FP%`)

```
```{r met3 list}
# test_bind
# xstreme_rank_bind
DT::datatable((H3K27ac_Set2_data_met3 %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 3: Set 2 H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 

DT::datatable((H3K27ac_Set3_data_met3 %>% 
  dplyr::select(RANK,CLUSTER,ID,ALT_ID,EVALUE, CONSENSUS.x)),
  rownames = FALSE,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Method 3: Set 3 H3K27ac"
  ),
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = TRUE)) 


```

```{r direct compare}

toplist_direct_set2 <- list("set2_met1"=H3K27ac_Set2_data$ALT_ID,
"set2_met2"=H3K27ac_Set2_data_met2$ALT_ID,
"set2_met3"=H3K27ac_Set2_data_met3$ALT_ID)

toplist_direct_set3 <- list("set3_met1"=H3K27ac_Set3_data$ALT_ID,
"set3_met2"=H3K27ac_Set3_data_met2$ALT_ID,
"set3_met3"=H3K27ac_Set3_data_met3$ALT_ID)



```

```{r}
 Shared_motifs_set2 <- Reduce(intersect, toplist_direct_set2)
ggVennDiagram::ggVennDiagram(toplist_direct_set2)+scale_fill_gradient(
    low = "white",
    high = "blue"
  )

 Shared_motifs_set3 <- Reduce(intersect, toplist_direct_set3)
ggVennDiagram::ggVennDiagram(toplist_direct_set3)+scale_fill_gradient(
    low = "white",
    high = "red"
  )
```




