---
title: "RNA and H3K27ac"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: 
  html_document:
  toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(data.table)
library(readxl)
library(ChIPseeker)
library(smplot2)
library(biomaRt)
```

adding in data files.

The current analysis should be RNA DEGs and ROIs by set expression levels.
more steps will include viewing the association of DEGs with enrichment/depletion of ROIs containing TEs by class/family with a look at age.
### Data loading  


```{r needed data files}
###Loading all ROIs, already out by Set and combined
H3K27ac_anno_ROIs <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")
H3K27ac_toplist <- readRDS( "data/DER_data/H3K27ac_toplist.RDS")
H3K27ac_toptable_list <- bind_rows(H3K27ac_toplist, .id = "group")
## Loading single summit locations and adding in 
H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")

##assigning Peakid as name of summit region
 mcols(H3K27ac_summit_gr)$name <- mcols(H3K27ac_summit_gr)$Peakid
 

## Creating a lookup file if needed
H3K27ac_lookup <- imap_dfr(H3K27ac_anno_ROIs[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)

RNA_full_toptable <-  readRDS("data/Other_paper_data/RNA_full_toptable.RDS")

RNA_cluster_lookup<- readRDS( "data/Other_paper_data/RNA_cluster_lookup.RDS")
LTR_summits_overlap_df <- readRDS("data/RDS_files/H3K27ac_summit_LTR_overlaps.RDS")

H3K27ac_set_case_lfc <- readRDS( "data/RDS_files/H3K27ac_set_case_lfc.RDS")

Symbol_lookup <- RNA_full_toptable %>% 
  dplyr::select (Entrez_ID,SYMBOL) %>% 
  distinct(Entrez_ID,SYMBOL)

  KZFP_2017_list <- read_excel("data/Other_paper_data/41586_2017_BFnature21683_MOESM103_ESM.xlsx")

  
  KZFPIDs <- KZFP_2017_list %>% 
    dplyr::filter(Species == "Homo_sapiens.GRCh38") %>%
    dplyr::select(Label,`Gene ID`, Type)
  
  mart <- useEnsembl(biomart = "genes",
                     dataset = "hsapiens_gene_ensembl",
                     mirror = "useast") ### using a different mirror due to server issues
  
  ensembl_ID <- KZFPIDs$`Gene ID`
  
  results_KZFP <- getBM(attributes = c("ensembl_gene_id", "entrezgene_id","hgnc_symbol"),
               filters = "ensembl_gene_id",
               values = ensembl_ID,
               mart = mart)
  
  results_KZFP %>% na.omit()

```

Below this code is how I combined the RNA files from Emma's paper for future reference.
```{r RNA data files, eval=FALSE}
##For my reference only
### Loading in RNA expression data
RNA_toptable_24T <- read_excel("C:/Users/renee/Other_projects_data/DXR_data/DXR_RNA_data/Pfortmiller_S2_Appendix_251213.xlsx", 
    sheet = "Table_S3", skip = 1)

RNA_toptable_24R <- read_excel("C:/Users/renee/Other_projects_data/DXR_data/DXR_RNA_data/Pfortmiller_S2_Appendix_251213.xlsx", 
    sheet = "Table_S4", skip = 1)

RNA_toptable_144R <- read_excel("C:/Users/renee/Other_projects_data/DXR_data/DXR_RNA_data/Pfortmiller_S2_Appendix_251213.xlsx", 
    sheet = "Table_S5", skip = 1)

RNA_acute_cluster <- read_excel("C:/Users/renee/Other_projects_data/DXR_data/DXR_RNA_data/Pfortmiller_S2_Appendix_251213.xlsx", 
    sheet = "Table_S6", skip = 1)

RNA_chronic_cluster <- read_excel("C:/Users/renee/Other_projects_data/DXR_data/DXR_RNA_data/Pfortmiller_S2_Appendix_251213.xlsx", 
    sheet = "Table_S7", skip = 2)

## Creating R object for my data:

RNA_full_toptable <- RNA_toptable_24T %>% 
  mutate(timepoint="24T") %>% 
  bind_rows(., (RNA_toptable_24R %>% 
                  mutate(timepoint="24R"))) %>% 
  bind_rows(., (RNA_toptable_144R %>% 
                  mutate(timepoint="144R")))
# saveRDS(RNA_full_toptable, "data/Other_paper_data/RNA_full_toptable.RDS")

RNA_cluster_lookup <- RNA_acute_cluster %>% 
    mutate(Response_Direction="NA") %>% 
    bind_rows(RNA_chronic_cluster)
# saveRDS(RNA_cluster_lookup, "data/Other_paper_data/RNA_cluster_lookup.RDS")

```

Step 1: what Entrez_IDs are connected to each set?  Are there any shared entrez_ids between sets 2 and 3?

```{r}


Joined_clusters <- H3K27ac_anno_ROIs$all_H3K27ac %>% 
  as.data.frame() %>% 
  dplyr::select(Peakid,annotation, geneId) %>% 
  left_join(., H3K27ac_lookup) %>% 
  dplyr::rename("ROI_cluster"=cluster) %>% 
  mutate(geneId=as.numeric(geneId)) %>% 
  left_join(., RNA_cluster_lookup, by=c("geneId"="Entrez_ID")) 
  

  
Joined_clusters %>%   
  group_by(ROI_cluster, Cluster) %>% 
  tally() %>% 
datatable(.,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))

```
I have associated the tables using Entrez_ID numbers.  Now I want to see if Set1, Set2, or Set3 are enriched in chronic vs acute RNA genes.
I will do a Fisher's Exact test.  Note, I am treating each ROI individually, meaning genes can have multiple ROIs.

I will first ask if Set2 is enriched in Acute or Chronic

####  Acute or Chronic Set2  

```{r  Set2 enrichment}
tab_annotate <- Joined_clusters %>%
  dplyr::filter(!is.na(ROI_cluster)) %>% 
  mutate(
    ROI_cluster= factor(ROI_cluster,levels=c("Set_1","Set_2","Set_3")),
    Cluster=factor(Cluster,levels=c("Chronic", "Acute")))
 

tab_21_acute <-  tab_annotate %>% 
  filter(ROI_cluster %in% c("Set_1", "Set_2")) %>%
  mutate(is_acute = Cluster=="Acute") %>% 
  count(ROI_cluster, is_acute) %>%
  tidyr::pivot_wider(
    names_from = is_acute,
    values_from = n,
    values_fill = 0)

tab_21_chronic <-  tab_annotate %>% 
  filter(ROI_cluster %in% c("Set_1", "Set_2")) %>%
  mutate(is_chronic=Cluster=="Chronic") %>% 
  count(ROI_cluster, is_chronic) %>%
  tidyr::pivot_wider(
    names_from = is_chronic,
    values_from = n,
    values_fill = 0)
 

mat_21_acute <- as.matrix(tab_21_acute[, c("TRUE", "FALSE")])
mat_21_chronic <- as.matrix(tab_21_chronic[, c("TRUE", "FALSE")])

fisher.test(mat_21_acute)
fisher.test(mat_21_chronic)
```


```{r  Set3 enrichment}

tab_31_acute <-  tab_annotate %>% 
  filter(ROI_cluster %in% c("Set_1", "Set_3")) %>%
  mutate(is_acute = Cluster=="Acute") %>% 
  count(ROI_cluster, is_acute) %>%
  tidyr::pivot_wider(
    names_from = is_acute,
    values_from = n,
    values_fill = 0)

tab_31_chronic <-  tab_annotate %>% 
  filter(ROI_cluster %in% c("Set_1", "Set_3")) %>%
  mutate(is_chronic=Cluster=="Chronic") %>% 
  count(ROI_cluster, is_chronic) %>%
  tidyr::pivot_wider(
    names_from = is_chronic,
    values_from = n,
    values_fill = 0)
 

mat_31_acute <- as.matrix(tab_31_acute[, c("TRUE", "FALSE")])
mat_31_chronic <- as.matrix(tab_31_chronic[, c("TRUE", "FALSE")])

fisher.test(mat_31_acute)
fisher.test(mat_31_chronic)
```

```{r  Set2 verses set3 enrichment}


tab_23_acute <-  tab_annotate %>% 
  filter(ROI_cluster %in% c("Set_2", "Set_3")) %>%
  mutate(is_acute = Cluster=="Acute") %>% 
  count(ROI_cluster, is_acute) %>%
  tidyr::pivot_wider(
    names_from = is_acute,
    values_from = n,
    values_fill = 0)

tab_23_chronic <-  tab_annotate %>% 
  filter(ROI_cluster %in% c("Set_2", "Set_3")) %>%
  mutate(is_chronic=Cluster=="Chronic") %>% 
  count(ROI_cluster, is_chronic) %>%
  tidyr::pivot_wider(
    names_from = is_chronic,
    values_from = n,
    values_fill = 0)
 

mat_23_acute <- as.matrix(tab_23_acute[, c("TRUE", "FALSE")])
mat_23_chronic <- as.matrix(tab_23_chronic[, c("TRUE", "FALSE")])

fisher.test(mat_23_acute)
fisher.test(mat_23_chronic)
```

```{r}
tab_annotate %>% 
  # dplyr::filter(Cluster=="Chronic") %>% 
  count(ROI_cluster,Cluster) %>% 
  ggplot(., aes(x=ROI_cluster,y=n, fill=Cluster))+
    geom_col(position = "fill")+
    scale_y_continuous(labels=scales::percent_format())+
  ggtitle("Proportion of Chronic and Actute ROIs by set using assigned gene")+
   ylab("Proportion")+
  xlab("Sets")
 
  
```
Now to breakdown by geneIDs and not ROIs

```{r}
gene_df <- Joined_clusters %>% 
  mutate(geneID=as.character(geneId)) %>% 
  group_by(geneId, ROI_cluster) %>% 
  summarize(ROI_ids=list(ROI_cluster),
            n_ROIs=dplyr::n(),
            cluster=paste0(unique(Cluster),collapse=","),
            SYMBOL=paste0(unique(SYMBOL),collapse = ","),
            .groups = "drop")

gene_long <- gene_df %>% 
  separate_rows(ROI_cluster,sep=",")
gene_long <- gene_long %>%
  mutate(
    is_acute = as.numeric(cluster == "Acute"),
    is_chronic = as.numeric(cluster == "Chronic")
  )
gene_long$ROI_cluster <- relevel(factor(gene_long$ROI_cluster), ref = "Set_1")

fit_acute <- glm(
  is_acute ~ ROI_cluster,
  data = gene_long,
  family = binomial
)
summary(fit_acute)
exp(cbind(OR = coef(fit_acute), confint(fit_acute)))


fit_chronic <- glm(is_chronic ~ ROI_cluster, data = gene_long, family = binomial)

summary(fit_chronic)
exp(cbind(OR = coef(fit_chronic), confint(fit_chronic)))
```

looking at expression levels of Chronic and Acute genes by set?
### RNA expression in Chronic or Acute genes  


```{r}
Set_1_nAcute <- Joined_clusters %>% 
  dplyr::filter(ROI_cluster=="Set_1") %>% 
  # dplyr::filter(Cluster=="Acute") %>% 
  distinct(geneId)

Set_2_nAcute <- Joined_clusters %>% 
  dplyr::filter(ROI_cluster=="Set_2") %>% 
  # dplyr::filter(Cluster=="Acute") %>% 
  distinct(geneId)

Set_3_nAcute <- Joined_clusters %>% 
  dplyr::filter(ROI_cluster=="Set_3") %>% 
  # dplyr::filter(Cluster=="Acute") %>% 
  distinct(geneId)

RNA_full_toptable %>% 
  dplyr::filter(Entrez_ID %in% Set_1_nAcute$geneId) %>% 
  dplyr::filter(adj.P.Val<0.05) %>% 
  mutate(timepoint=factor(timepoint,levels=c("24T","24R","144R")))%>% 
  left_join(., RNA_cluster_lookup) %>% 
  ggplot(., aes(x=timepoint, y=logFC, group=Entrez_ID))+
  geom_boxplot(aes(group=timepoint))+
  # geom_line(alpha=0.3, aes(color=Cluster))+
  ggtitle("LFC of genes in Set 1 ")


RNA_full_toptable %>% 
  dplyr::filter(Entrez_ID %in% Set_2_nAcute$geneId) %>% 
  dplyr::filter(adj.P.Val<0.05) %>% 
  mutate(timepoint=factor(timepoint,levels=c("24T","24R","144R")))%>% 
  left_join(., RNA_cluster_lookup) %>% 
  ggplot(., aes(x=timepoint, y=logFC, group=Entrez_ID))+
  geom_boxplot(aes(group=timepoint))+
  # geom_line(alpha=0.3, aes(color=Cluster))+
  ggtitle("LFC of genes in Set 2 ")

RNA_full_toptable %>% 
  dplyr::filter(Entrez_ID %in% Set_3_nAcute$geneId) %>% 
  dplyr::filter(adj.P.Val<0.05) %>% 
  mutate(timepoint=factor(timepoint,levels=c("24T","24R","144R")))%>% 
  left_join(., RNA_cluster_lookup) %>% 
  ggplot(., aes(x=timepoint, y=logFC, group=Entrez_ID))+
  geom_boxplot(aes(group=timepoint))+
  # geom_line(alpha=0.3, aes(color=Cluster))+
  ggtitle("LFC of genes in Set 3 ")

```

```{r}
H3K27ac_anno_df <- H3K27ac_anno_ROIs$all_H3K27ac %>% as.data.frame()

H3K27ac_toptable_list %>% 
  left_join(., H3K27ac_lookup,by=c("genes"="Peakid"))


summit_LTR_ROIs <- LTR_summits_overlap_df %>% 
  left_join(H3K27ac_lookup, by=c("Peakid"="Peakid")) %>% 
  left_join(H3K27ac_anno_df) %>% 
  dplyr::select(Peakid:cluster,geneId,distanceToTSS)

# summit_LTR_ROIs %>% distinct(geneId)

just_summit_LTR_ROI_peakids <- summit_LTR_ROIs %>% distinct(Peakid,.keep_all = TRUE) %>% dplyr::select(Peakid,geneId)

```

getting LFC of Peakids assigned to geneIds

### LFC_looking

The first plot is all ROI-genes regardless of cluster status or TE relationship:
```{r}

long_H3K27ac_set_case_lfc <- dplyr::bind_rows(H3K27ac_set_case_lfc, .id = "set_case")
RNA_lfc_entrez_id <- RNA_full_toptable %>% 
  dplyr::select(Entrez_ID,logFC,timepoint) %>% 
  pivot_wider(., names_from = timepoint, names_prefix = "RNA_LFC_", values_from = logFC)

LFC_joined_peakid_geneid <- long_H3K27ac_set_case_lfc %>% 
  dplyr::filter(Peakid %in% just_summit_LTR_ROI_peakids$Peakid) %>%
  left_join(just_summit_LTR_ROI_peakids) %>% 
   mutate(geneId=as.numeric(geneId)) %>%
  left_join(RNA_lfc_entrez_id, by=c("geneId"="Entrez_ID"))


H3K27ac_full_Peakid_anno <- H3K27ac_anno_ROIs$all_H3K27ac %>% 
  as.data.frame() %>% 
  dplyr::select(Peakid, geneId, distanceToTSS) %>% 
  left_join(., long_H3K27ac_set_case_lfc, by=c("Peakid" = "Peakid")) %>% 
  mutate(geneId=as.numeric(geneId)) %>%
  left_join(RNA_lfc_entrez_id, by=c("geneId"="Entrez_ID"))

ggplot(H3K27ac_full_Peakid_anno,
    aes(x = H3K27ac_24T, y = RNA_LFC_24T, color = cluster)
  ) +
    geom_point(alpha = 0.2, size = 1) +
    geom_density_2d(color = "black") +
    sm_statCorr(corr_method = "spearman") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    labs(
      x = paste("H3K27ac_24T logFC"),
      y = paste("RNA_LFC_24T logFC"),
      title = paste("All ROI LFC with each RNA LFC 24T")) +
    theme_classic()

ggplot(H3K27ac_full_Peakid_anno,
    aes(x = H3K27ac_24R, y = RNA_LFC_24R, color = cluster)
  ) +
    geom_point(alpha = 0.2, size = 1) +
    geom_density_2d(color = "black") +
    sm_statCorr(corr_method = "spearman") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    labs(
      x = paste("H3K27ac_24T logFC"),
      y = paste("RNA_LFC_24T logFC"),
      title = paste("All ROI LFC with each RNA LFC 24R")) +
    theme_classic()


ggplot(H3K27ac_full_Peakid_anno,
    aes(x = H3K27ac_144R, y = RNA_LFC_144R, color = cluster)
  ) +
    geom_point(alpha = 0.2, size = 1) +
    geom_density_2d(color = "black") +
    sm_statCorr(corr_method = "spearman") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    labs(
      x = paste("H3K27ac_24T logFC"),
      y = paste("RNA_LFC_24T logFC"),
      title = paste("All ROI LFC with each RNA LFC 144R")) +
    theme_classic()



```

```{r time to time LFC}
plot_LFC_scatter <- function(df,
                             cluster_name,
                             x_var = H3K27ac_24T,
                             y_var = RNA_LFC_24T,
                             timepoint = "24T") {

  ggplot(
    df %>% dplyr::filter(cluster == cluster_name),
    aes(x = {{ x_var }}, y = {{ y_var }}, color = set_case)
  ) +
    geom_point(alpha = 0.2, size = 1) +
    geom_density_2d(color = "black") +
    sm_statCorr(corr_method = "spearman") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    labs(
      x = paste(deparse(substitute(x_var)), "logFC"),
      y = paste(deparse(substitute(y_var)), timepoint),
      title = paste(
        "LTR_summit",
        deparse(substitute(x_var)),
        "and RNA LFC at",
        timepoint,
        tolower(cluster_name)
      )
    ) +
    theme_classic()
}


plot_LFC_boxplot <- function(df,
                             col_one = "H3K27ac_24T",
                             col_two = "RNA_LFC_24T",
                             cluster_filter = NULL,
                             peakid_filter = NULL,
                             lfc_cutoff = 1,
                             title = NULL) {

  plot_df <- df

  # optional filters
  if (!is.null(cluster_filter)) {
    plot_df <- plot_df %>% dplyr::filter(cluster %in% cluster_filter)
  }

  if (!is.null(peakid_filter)) {
    plot_df <- plot_df %>% dplyr::filter(Peakid %in% peakid_filter)
  }

  # bin based on dynamic column
  plot_df <- plot_df %>%
    dplyr::mutate(
      LFC_bin = dplyr::case_when(
        .data[[col_one]] <= -lfc_cutoff ~ "Loss",
        .data[[col_one]] >=  lfc_cutoff ~ "Gain",
        TRUE ~ "Neutral"
      ),
      LFC_bin = factor(LFC_bin, levels = c("Loss", "Neutral", "Gain"))
    )

  # default title
  if (is.null(title)) {
    title <- paste(
      col_two, "by", col_one,
      if (!is.null(cluster_filter)) paste("-", paste(cluster_filter, collapse = ", "))
    )
  }

  ggplot(plot_df,
         aes(x = LFC_bin,
             y = .data[[col_two]],
             fill = LFC_bin)) +

    geom_boxplot(outlier.shape = NA, alpha = 0.8) +
    geom_jitter(width = 0.15, alpha = 0.15, size = 0.7) +

    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +

    labs(
      x = paste0(col_one, " (binned)"),
      y = col_two,
      title = title
    ) +

    ggpubr::stat_compare_means(
      comparisons = list(c("Loss", "Gain")),
      method = "wilcox.test"
    ) +

    theme_classic() +
    theme(legend.position = "none")
}
```

  Now to look at LFC by summits that cross an LTR
```{r time to time LFC, Set1}

plot_LFC_scatter(LFC_joined_peakid_geneid,cluster_name = "Set_1",x_var = H3K27ac_24T, y_var = RNA_LFC_24T, timepoint = "24T")

plot_LFC_scatter(LFC_joined_peakid_geneid,cluster_name = "Set_1",x_var = H3K27ac_24R, y_var = RNA_LFC_24R, timepoint = "24R")

plot_LFC_scatter(LFC_joined_peakid_geneid,cluster_name = "Set_1",x_var = H3K27ac_144R, y_var = RNA_LFC_144R, timepoint = "144R")




plot_LFC_boxplot(
  LFC_joined_peakid_geneid,col_one = "H3K27ac_24T", col_two="RNA_LFC_24T",
  cluster_filter = "Set_1")

plot_LFC_boxplot(
  LFC_joined_peakid_geneid,col_one = "H3K27ac_24R", col_two="RNA_LFC_24R",
  cluster_filter = "Set_1")
  
  plot_LFC_boxplot(
  LFC_joined_peakid_geneid,col_one = "H3K27ac_144R", col_two="RNA_LFC_144R",
  cluster_filter = "Set_1")

```

```{r time to time LFC, Set2}

plot_LFC_scatter(LFC_joined_peakid_geneid,cluster_name = "Set_2",x_var = H3K27ac_24T, y_var = RNA_LFC_24T, timepoint = "24T")

plot_LFC_scatter(LFC_joined_peakid_geneid,cluster_name = "Set_2",x_var = H3K27ac_24R, y_var = RNA_LFC_24R, timepoint = "24R")

plot_LFC_scatter(LFC_joined_peakid_geneid,cluster_name = "Set_2",x_var = H3K27ac_144R, y_var = RNA_LFC_144R, timepoint = "144R")

```

```{r time to time LFC, Set3}

plot_LFC_scatter(LFC_joined_peakid_geneid,cluster_name = "Set_3",x_var = H3K27ac_24T, y_var = RNA_LFC_24T, timepoint = "24T")

plot_LFC_scatter(LFC_joined_peakid_geneid,cluster_name = "Set_3",x_var = H3K27ac_24R, y_var = RNA_LFC_24R, timepoint = "24R")

plot_LFC_scatter(LFC_joined_peakid_geneid,cluster_name = "Set_3",x_var = H3K27ac_144R, y_var = RNA_LFC_144R, timepoint = "144R")

```


maybe closer to TSS will be helpful?

filtering to only those within 2kb of a TSS for all ROIs, then by Set 1 ROIs only (whose summit crosses an LTR)
```{r 2kb set1}

promotor_region_LTR_peakid <- summit_LTR_ROIs %>% 
  dplyr::filter(distanceToTSS>-2000 & distanceToTSS <2000)

H3K27ac_2kb_LTR <- LFC_joined_peakid_geneid %>% dplyr::filter(Peakid %in% promotor_region_LTR_peakid$Peakid)

plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_24T", col_two="RNA_LFC_24T"
)

plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_24R", col_two="RNA_LFC_24R"
)

plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_144R", col_two="RNA_LFC_144R"
)



plot_LFC_scatter(H3K27ac_2kb_LTR,cluster_name = "Set_1",x_var = H3K27ac_24T, y_var = RNA_LFC_24T, timepoint = "24T")+labs(caption="within 2kb of TSS")

plot_LFC_scatter(H3K27ac_2kb_LTR,cluster_name = "Set_1",x_var = H3K27ac_24R, y_var = RNA_LFC_24R, timepoint = "24R")+labs(caption="within 2kb of TSS")

plot_LFC_scatter(H3K27ac_2kb_LTR,cluster_name = "Set_1",x_var = H3K27ac_144R, y_var = RNA_LFC_144R, timepoint = "144R")+labs(caption="within 2kb of TSS")


plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_24T", col_two="RNA_LFC_24T",
  cluster_filter = "Set_1"
)

plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_24R", col_two="RNA_LFC_24R",
  cluster_filter = "Set_1"
)

plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_144R", col_two="RNA_LFC_144R",
  cluster_filter = "Set_1"
)


```
now For set 2
```{r 2kb set2}


plot_LFC_scatter(H3K27ac_2kb_LTR,cluster_name = "Set_2",x_var = H3K27ac_24T, y_var = RNA_LFC_24T, timepoint = "24T")+labs(caption="within 2kb of TSS")

plot_LFC_scatter(H3K27ac_2kb_LTR,cluster_name = "Set_2",x_var = H3K27ac_24R, y_var = RNA_LFC_24R, timepoint = "24R")+labs(caption="within 2kb of TSS")

plot_LFC_scatter(H3K27ac_2kb_LTR,cluster_name = "Set_2",x_var = H3K27ac_144R, y_var = RNA_LFC_144R, timepoint = "144R")+labs(caption="within 2kb of TSS")
plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_24T", col_two="RNA_LFC_24T",
  cluster_filter = "Set_2"
)

plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_24R", col_two="RNA_LFC_24R",
  cluster_filter = "Set_2"
)

plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_144R", col_two="RNA_LFC_144R",
  cluster_filter = "Set_2"
)


```
now for set 3.
```{r 2kb set3}

plot_LFC_scatter(H3K27ac_2kb_LTR,cluster_name = "Set_3",x_var = H3K27ac_24T, y_var = RNA_LFC_24T, timepoint = "24T")+labs(caption="within 2kb of TSS")

plot_LFC_scatter(H3K27ac_2kb_LTR,cluster_name = "Set_3",x_var = H3K27ac_24R, y_var = RNA_LFC_24R, timepoint = "24R")+labs(caption="within 2kb of TSS")

plot_LFC_scatter(H3K27ac_2kb_LTR,cluster_name = "Set_3",x_var = H3K27ac_144R, y_var = RNA_LFC_144R, timepoint = "144R")+labs(caption="within 2kb of TSS")


plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_24T", col_two="RNA_LFC_24T",
  cluster_filter = "Set_3"
)

plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_24R", col_two="RNA_LFC_24R",
  cluster_filter = "Set_3"
)

plot_LFC_boxplot(
  H3K27ac_2kb_LTR,col_one = "H3K27ac_144R", col_two="RNA_LFC_144R",
  cluster_filter = "Set_3"
)


```





now I want to see about further away from TSS, so regions > 2kb but less than 20KB from the TSS of assigned gene. (Still summits crossing an LTR)
```{r 20kb to 2kb LTR no set}
between20kb_2kb_summit_LTR_peakids <- summit_LTR_ROIs %>%
  dplyr::mutate(abs_dist=abs(distanceToTSS)) %>% 
  dplyr::filter(abs_dist>2000 & abs_dist <20000) %>% 
  distinct(Peakid,geneId)

H3K27ac_20kb_2kb_LTR <- LFC_joined_peakid_geneid %>% dplyr::filter(Peakid %in% between20kb_2kb_summit_LTR_peakids$Peakid)

ggplot(H3K27ac_20kb_2kb_LTR,
    aes(x = H3K27ac_24T, y = RNA_LFC_24T, color = cluster)
  ) +
    geom_point(alpha = 0.2, size = 1) +
    geom_density_2d(color = "black") +
    sm_statCorr(corr_method = "spearman") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    labs(
      x = paste("H3K27ac_24T logFC"),
      y = paste("RNA_LFC_24T logFC"),
      title = paste("ROI_summits overlapping LTR with each RNA LFC 24T\n20kb>x> 2kb of TSS 24T")) +
    theme_classic()

ggplot(H3K27ac_20kb_2kb_LTR,
    aes(x = H3K27ac_24R, y = RNA_LFC_24R, color = cluster)
  ) +
    geom_point(alpha = 0.2, size = 1) +
    geom_density_2d(color = "black") +
    sm_statCorr(corr_method = "spearman") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    labs(
      x = paste("H3K27ac_24R logFC"),
      y = paste("RNA_LFC_24R logFC"),
     title = paste("ROI_summits overlapping LTR with each RNA LFC 24T\n20kb>x> 2kb of TSS 24R")) +
    theme_classic()


ggplot(H3K27ac_20kb_2kb_LTR,
    aes(x = H3K27ac_144R, y = RNA_LFC_144R, color = cluster)
  ) +
    geom_point(alpha = 0.2, size = 1) +
    geom_density_2d(color = "black") +
    sm_statCorr(corr_method = "spearman") +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    labs(
      x = paste("H3K27ac_144R logFC"),
      y = paste("RNA_LFC_144R logFC"),
     title = paste("ROI_summits overlapping LTR with each RNA LFC 24T\n20kb>x> 2kb of TSS 144R")) +
    theme_classic()





```


```{r 20kb to 2kb LTR set 3}
plot_LFC_scatter(H3K27ac_20kb_2kb_LTR,cluster_name = "Set_3",x_var = H3K27ac_24T, y_var = RNA_LFC_24T, timepoint = "24T")+labs(caption="20kb>x> 2kb of TSS")

plot_LFC_scatter(H3K27ac_20kb_2kb_LTR,cluster_name = "Set_3",x_var = H3K27ac_24R, y_var = RNA_LFC_24R, timepoint = "24R")+labs(caption="20kb>x> 2kb of TSS")

plot_LFC_scatter(H3K27ac_20kb_2kb_LTR,cluster_name = "Set_3",x_var = H3K27ac_144R, y_var = RNA_LFC_144R, timepoint = "144R")+labs(caption="20kb>x> 2kb of TSS")



plot_LFC_boxplot(
  H3K27ac_20kb_2kb_LTR,col_one = "H3K27ac_24T", col_two="RNA_LFC_24T",
  cluster_filter = "Set_3"
)

plot_LFC_boxplot(
  H3K27ac_20kb_2kb_LTR,col_one = "H3K27ac_24R", col_two="RNA_LFC_24R",
  cluster_filter = "Set_3"
)

plot_LFC_boxplot(
  H3K27ac_20kb_2kb_LTR,col_one = "H3K27ac_144R", col_two="RNA_LFC_144R",
  cluster_filter = "Set_3"
)

```



```{r 20kb to 2kb LTR set 2}
 plot_LFC_scatter(H3K27ac_20kb_2kb_LTR,cluster_name = "Set_2",x_var = H3K27ac_24T, y_var = RNA_LFC_24T, timepoint = "24T")+labs(caption="20kb>x> 2kb of TSS")

plot_LFC_scatter(H3K27ac_20kb_2kb_LTR,cluster_name = "Set_2",x_var = H3K27ac_24R, y_var = RNA_LFC_24R, timepoint = "24R")+labs(caption="20kb>x> 2kb of TSS")

plot_LFC_scatter(H3K27ac_20kb_2kb_LTR,cluster_name = "Set_2",x_var = H3K27ac_144R, y_var = RNA_LFC_144R, timepoint = "144R")+labs(caption="20kb>x> 2kb of TSS")



plot_LFC_boxplot(
  H3K27ac_20kb_2kb_LTR,col_one = "H3K27ac_24T", col_two="RNA_LFC_24T",
  cluster_filter = "Set_2"
)

plot_LFC_boxplot(
  H3K27ac_20kb_2kb_LTR,col_one = "H3K27ac_24R", col_two="RNA_LFC_24R",
  cluster_filter = "Set_2"
)

plot_LFC_boxplot(
  H3K27ac_20kb_2kb_LTR,col_one = "H3K27ac_144R", col_two="RNA_LFC_144R",
  cluster_filter = "Set_2"
)
```

```{r 20kb to 2kb LTR set 1}

plot_LFC_scatter(H3K27ac_20kb_2kb_LTR,cluster_name = "Set_1",x_var = H3K27ac_24T, y_var = RNA_LFC_24T, timepoint = "24T")+labs(caption="20kb>x> 2kb of TSS")

plot_LFC_scatter(H3K27ac_20kb_2kb_LTR,cluster_name = "Set_1",x_var = H3K27ac_24R, y_var = RNA_LFC_24R, timepoint = "24R")+labs(caption="20kb>x> 2kb of TSS")

plot_LFC_scatter(H3K27ac_20kb_2kb_LTR,cluster_name = "Set_1",x_var = H3K27ac_144R, y_var = RNA_LFC_144R, timepoint = "144R")+labs(caption="20kb>x> 2kb of TSS")
```
```{r}
summit_LTR_ROIs <- summit_LTR_ROIs %>% 
  dplyr::mutate(
    TSS_dir = ifelse(distanceToTSS < 0, "Upstream", "Downstream")
  )

# distance to TSspread(
summit_LTR_ROIs %>% 
  ggplot(., aes(x=distanceToTSS, color=cluster))+
  geom_density(linewidth=3)+
  coord_cartesian(xlim=c(-50000,50000))

summit_LTR_ROIs %>% 
  dplyr::mutate(
    TSS_dir = ifelse(distanceToTSS < 0, "Upstream", "Downstream")
  ) %>% 
  ggplot(., aes(x=cluster,y=distanceToTSS, color=cluster))+
  geom_violin(trim = FALSE, scale = "width", alpha = 0.7) +
  geom_boxplot(
    width = 0.1,
    outlier.shape = NA,
    alpha = 0.4
  ) +
    # coord_cartesian(ylim = c(-50000, 50000)) +
  theme_classic()
```

```{r}
GeomSplitViolin <- ggplot2::ggproto(
  "GeomSplitViolin", ggplot2::GeomViolin,
  draw_group = function(self, data, panel_scales, coord) {
    data <- transform(data,
      xminv = x - violinwidth * (x - xmin),
      xmaxv = x + violinwidth * (xmax - x)
    )

    grp <- data[1, "group"]
    newdata <- if (grp %% 2 == 1) {
      transform(data, x = xminv)
    } else {
      transform(data, x = xmaxv)
    }

    ggplot2::GeomPolygon$draw_panel(newdata, panel_scales, coord)
  }
)

geom_split_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                              position = "identity", ..., trim = TRUE,
                              scale = "area", show.legend = NA,
                              inherit.aes = TRUE) {
  ggplot2::layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomSplitViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}
```

```{r}
summit_LTR_ROIs %>% 
  ggplot(aes(
    x = cluster,
    y = distanceToTSS,
    fill = TSS_dir
  )) +
  geom_split_violin(trim = FALSE, scale = "width", alpha = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  # coord_cartesian(ylim = c(-50000, 50000)) +
  scale_fill_manual(
    values = c("Upstream" = "#1f78b4", "Downstream" = "#e31a1c")
  ) +
  facet_wrap(~TE_type)+
  theme_classic()+
  ggtitle("getting an idea of TSS distribution")
```
### KZFP information

#### all KZFP log2cpm expression
```{r Basic LFC}
RNA_filcpm_final_matrix <- read_csv("data/Other_paper_data/RNA_filcpm_final_matrix.csv")

just_kzfp <- results_KZFP %>% na.omit()%>%
  arrange(hgnc_symbol)


KZFP_RNA <- RNA_filcpm_final_matrix %>% 
  dplyr::filter(Entrezid %in% just_kzfp$entrezgene_id) %>% 
  pivot_longer(cols=!Entrezid, names_to = "sample",values_to = "log2cpm") %>% 
  separate_wider_delim(cols = sample, names = c("trt","time","ind"), delim = "_") %>% 
  mutate(group=paste0(trt,"_",time),
  time=factor(time,levels=c("24T","24R","144R"))) 


RNA_full_toptable %>% 
  mutate(timepoint=factor(timepoint, levels=c("24T","24R","144R"))) %>% 
  dplyr::filter(Entrez_ID %in% just_kzfp$entrezgene_id) %>% 
  ggplot(., aes(x=timepoint, y=logFC))+
  geom_boxplot()+
  ggtitle("LFC of all KZFP in RNA data")+
  theme_bw()
```

### KZFP Plots {.tabset .tabset-dropdown}
```{r Basic LFC all kzfp, results='asis'}
for (aim in just_kzfp$entrezgene_id){
  df <- KZFP_RNA %>% 
    mutate(time=factor(time, levels=c("24T","24R","144R"))) %>% 
        dplyr::filter(Entrezid== aim)
  if (nrow(df) == 0) next
  symbol <- just_kzfp %>% 
      dplyr::filter(entrezgene_id==aim) %>% 
    dplyr::pull(hgnc_symbol)
  
  # Create tab title
  cat("\n\n### ", symbol, " (", aim, ")\n\n")
  
plot <- ggplot(df, aes(x=time, y=log2cpm,group=group))+
   geom_boxplot(aes(fill=trt)) +
  ggtitle(paste0("gene ", aim," ",symbol))+
  theme_bw()
print(plot)
  }

```


Age and KZFP expression


```{r}

# results_KZFP
KZFP_age <- KZFP_2017_list %>% 
  dplyr::filter(Label %in% results_KZFP$hgnc_symbol) %>% 
  group_by(Label) %>% 
  summarise(
    max_MY = max(`Estimated evolutionary distance from human (MY)`, na.rm=TRUE),
    human_gene = `Gene ID`[`Latin name` == "Homo sapiens"][1],
    cluster = `Cluster #`[`Latin name` == "Homo sapiens"][1])

KZFP_RNA %>% 
  left_join(.,just_kzfp, by =c("Entrezid"="entrezgene_id")) %>% 
  left_join(., (KZFP_age %>% dplyr::select(Label,max_MY)), by =c("hgnc_symbol"="Label"))



RNA_full_toptable %>% 
    dplyr::filter(Entrez_ID %in%just_kzfp$entrezgene_id) %>% 
  left_join(., (KZFP_age %>% dplyr::select(Label,max_MY)), by =c("SYMBOL"="Label")) %>% 
  mutate(timepoint=factor(timepoint,levels=c("24T","24R","144R"))) %>%
  
  ggplot(.,aes(x = max_MY, y = logFC))+
  geom_point()+
  # geom_smooth(method = "lm",  se = FALSE) +
   # geom_smooth(method = "loess", color = "red", span=1) +
  facet_wrap(~timepoint)
```

```{r}
RNA_full_toptable %>% 
    dplyr::filter(Entrez_ID %in%just_kzfp$entrezgene_id) %>% 
  left_join(., (KZFP_age %>% dplyr::select(Label,max_MY)), by =c("SYMBOL"="Label")) %>% 
  na.omit() %>% 
   mutate(age_bin = case_when(
    max_MY <= 50 ~ "0-50 mya",
    max_MY <= 150 ~ "50-150 mya",
    max_MY <= 200 ~ "150-200 mya",
    max_MY > 300 ~ "> 300 mya",
    TRUE~"unknown")) %>% 
 mutate(timepoint=factor(timepoint,levels=c("24T","24R","144R")),
        age_bin=factor(age_bin,levels=c("0-50 mya","50-150 mya","150-200 mya","> 300 mya","unknown"))) %>%
   # summarise(mean_logFC = mean(logFC, na.rm=TRUE)) %>% 
  ggplot(.,aes(x = age_bin, y = logFC))+
  geom_boxplot()+
  facet_wrap(~timepoint) + 
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggtitle("RNA expression and time (mya-million years ago)")
```
```{r}
RNA_full_toptable %>% 
    dplyr::filter(Entrez_ID %in%just_kzfp$entrezgene_id) %>% 
  left_join(., (KZFP_age %>% dplyr::select(Label,max_MY)), by =c("SYMBOL"="Label")) %>% 
  # dplyr::filter(is.na(max_MY)) %>% 
  dplyr::filter(SYMBOL=="ZNF605")


RNA_full_toptable %>% 
    dplyr::filter(Entrez_ID %in%just_kzfp$entrezgene_id) %>% 
  left_join(., (KZFP_age %>% dplyr::select(Label,max_MY)), by =c("SYMBOL"="Label")) %>% 
  na.omit() %>% 
   mutate(age_bin = case_when(
    max_MY <= 50 ~ "0-50 mya",
    max_MY <= 150 ~ "50-150 mya",
    max_MY <= 200 ~ "150-200 mya",
    max_MY > 300 ~ "> 300 mya",
    TRUE~"unknown")) %>% 
 mutate(timepoint=factor(timepoint,levels=c("24T","24R","144R")),
        age_bin=factor(age_bin,levels=c("0-50 mya","50-150 mya","150-200 mya","> 300 mya","unknown"))) %>% 
  group_by(age_bin) %>% tally()
```
### H3K27ac regions crossing LTR at summit with age
