---
title: "H3K27ac Extended summits and TE overlap"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
```

First steps: breakdown repeatmasker into groups and pull out the ones by each class I am interested in.  
```{r repeatmasker loading}
repeatmasker <- read_delim("data/Other_paper_data/repeatmasker_20250911.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
colnames(repeatmasker)
##this is to remove sex chromosomes and non-canonical chromatin assignments
autosomes <- paste0("chr", 1:22)

repeatmasker_clean <- repeatmasker %>% mutate(
    strand = ifelse(strand == "C", "-", "+")
  ) %>% 
   mutate(
    start = genoStart + 1,
    end   = genoEnd)%>% 
  mutate(repFamily= str_remove(repFamily, "\\?$")) %>% 
  dplyr::filter(genoName %in% autosomes) %>% 
  mutate(RM_id=paste0(genoName,":",start,"-",end,":",id))


rpt_split <- split(repeatmasker_clean, repeatmasker_clean$repClass)

rpt_split_gr_list <- lapply(rpt_split, function(df) {
  GRanges(
    seqnames = df$genoName,
    ranges   = IRanges(start = df$start, end = df$end),
    strand   = df$strand,
    repName   = df$repName,
    repClass  = df$repClass,
    repFamily = df$repFamily,
    swScore   = df$swScore,
    milliDiv  = df$milliDiv,
    milliDel  = df$milliDel,
    milliIns  = df$milliIns,
    RM_id        = df$RM_id
  )
})
### full repeatmasker cleaned up granges frame
rmskr_std_gr <- GRanges(
  seqnames = repeatmasker_clean$genoName,
  ranges   = IRanges(
    start = repeatmasker_clean$start,
    end   = repeatmasker_clean$end
  ),
  strand   = repeatmasker_clean$strand,
  repName   = repeatmasker_clean$repName,
  repClass  = repeatmasker_clean$repClass,
  repFamily = repeatmasker_clean$repFamily,
  swScore   = repeatmasker_clean$swScore,
  milliDiv  = repeatmasker_clean$milliDiv,
  milliDel  = repeatmasker_clean$milliDel,
  milliIns  = repeatmasker_clean$milliIns,
  id        = repeatmasker_clean$RM_id
)
```

```{r subset rpt_split_gr_list}
SINE_gr <- rpt_split_gr_list$SINE
SINE_df <- SINE_gr %>% 
  as.data.frame()

SINE_split_df <- split(SINE_df, SINE_df$repFamily)

LINE_gr <- rpt_split_gr_list$LINE
LINE_df <- LINE_gr %>% 
  as.data.frame()

LINE_split_df <- split(LINE_df, LINE_df$repFamily)

LTR_gr <- rpt_split_gr_list$LTR
LTR_df <- LTR_gr %>% 
  as.data.frame()

LTR_split_df <- split(LTR_df, LTR_df$repFamily)

SVA_gr <- rpt_split_gr_list$Retroposon
SVA_df <- SVA_gr %>% 
  as.data.frame()

SVA_split_df <- split(SVA_df, SVA_df$repFamily)

DNA_gr <- rpt_split_gr_list$DNA
DNA_df <- DNA_gr %>% 
  as.data.frame()

DNA_split_df <- split(DNA_df, DNA_df$repFamily)
```

```{r loading H3K27ac data}
H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")


peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)

##assigning Peakid as name of summit region
 mcols(H3K27ac_summit_gr)$name <- mcols(H3K27ac_summit_gr)$Peakid
 
 comparisons <- tibble(
  cluster2 = c("Set_2", "Set_3"),
  cluster1 = c("Set_1", "Set_1")
)

 H3K27ac_summit_gr_ext <- resize(H3K27ac_summit_gr, width(H3K27ac_summit_gr)+400, fix = "center")
 
```

```{r toplist data}

H3K27ac_toplist <- readRDS( "data/DER_data/H3K27ac_toplist.RDS")

H3K27ac_toptable_list <- bind_rows(H3K27ac_toplist, .id = "group")


cluster_colors <- c("Set_1" = "#d93e40","Set_2" = "#1c9f50","Set_3" = "#3570b3","NA"  ="grey70")

```

### making extened dataframe

```{r overlapping all, eval=FALSE}
dataf <- H3K27ac_ext200_gr

  
   # 1️⃣ find overlaps
  hits <- findOverlaps(dataf, rmskr_std_gr)
  
  if (length(hits) == 0) {
  H3K27ac_ext200_ols <- tibble()
} else {

  H3K27ac_ext200_ols <- tibble(
    Peakid     = dataf$Peakid[queryHits(hits)],
    cluster    = dataf$cluster[queryHits(hits)],
    seqnames   = as.character(seqnames(dataf))[queryHits(hits)],
    summit_pos = start(dataf)[queryHits(hits)],

    repName   = rmskr_std_gr$repName[subjectHits(hits)],
    repFamily = rmskr_std_gr$repFamily[subjectHits(hits)],
    repClass  = rmskr_std_gr$repClass[subjectHits(hits)],
    milliDiv  = rmskr_std_gr$milliDiv[subjectHits(hits)],
    milliDel  = rmskr_std_gr$milliDel[subjectHits(hits)],
    milliIns  = rmskr_std_gr$milliIns[subjectHits(hits)],
    ID  = rmskr_std_gr$id[subjectHits(hits)]
  )
}
  
  All_H3K27ac <-H3K27ac_sets_gr$all_H3K27ac
annotated_H3K27ac_ext200 <-
  All_H3K27ac %>% 
      as.data.frame() %>% 
    dplyr::select(Peakid,annotation,geneId:distanceToTSS) %>% 
     left_join(., (H3K27ac_ext200_ols %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(repClass, collapse=";"),
                            repFamily=paste0(repFamily, collapse=";"),
                            repName=paste0(repName,collapse=";"),
                            repID=paste0(ID, collapse = ";"),
                            summit_pos=paste0(unique(summit_pos),collapse = ";"),
                            .groups = "drop")))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
    left_join(H3K27ac_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))
  
# saveRDS(annotated_H3K27ac_ext200,"data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")
```

```{r}
test_pair_TE_repName <- function(df_long, rep_name, cluster1, cluster2) {
  
  # Subset for the specific repName
  sub_df <- df_long %>%
    filter(repName == rep_name) %>%
    complete(
      cluster = c(cluster1, cluster2),
      status  = c("TE", "not_TE"),
      fill = list(count = 0)
    )
  
  # fixed order of statuses
  status_levels <- c("TE", "not_TE")
  
  # make sure both statuses exist
  statuses <- unique(sub_df$status)
  if(length(statuses) != 2) {
    sub_df <- sub_df %>%
      complete(cluster, status, fill = list(count = 0))
    statuses <- unique(sub_df$status)
  }
  
  # counts for cluster1
  c1_counts <- sub_df %>%
    filter(cluster == cluster1) %>%
    arrange(factor(status, levels = status_levels)) %>%
    pull(count)
  
  # counts for cluster2
  c2_counts <- sub_df %>%
    filter(cluster == cluster2) %>%
    arrange(factor(status, levels = status_levels)) %>%
    pull(count)
  
  # 2x2 matrix for Fisher test
  mat <- matrix(
    c(c2_counts, c1_counts),
    nrow = 2,
    byrow = TRUE,
    dimnames = list(
      cluster = c(cluster2, cluster1),
      category = status_levels
    )
  )
  
  ft <- tryCatch(
    fisher.test(mat, workspace = 2e8),
    error = function(e) fisher.test(mat, simulate.p.value = TRUE, B = 1e5)
  )
  
  tibble(
    repName     = rep_name,
    comparison  = paste(cluster2, "vs", cluster1),
    odds_ratio  = ft$estimate,
    lower_CI    = ft$conf.int[1],
    upper_CI    = ft$conf.int[2],
    p_value     = ft$p.value
  )
}

```

```{r Fisher test information}

# Generic pairwise Fisher test
test_pair_TE_generic <- function(df_long, te_name, cluster1, cluster2) {
  
   sub_df <- df_long %>%
    filter(TE_type == te_name) %>%
    complete(
      cluster = c(cluster1, cluster2),
      status  = c("TE", "not_TE"),
      fill = list(count = 0))
   
   # enforce fixed order
 status_levels <- c("TE", "not_TE")
  
  # assume "status" column has TE vs wnot_TE automatically
  statuses <- unique(sub_df$status)
  
  if(length(statuses) != 2) {
    # ensure we have exactly two categories, fill missing with 0
    sub_df <- sub_df %>%
      complete(cluster, status, fill = list(count = 0))
    statuses <- unique(sub_df$status)
  }
  
  # extract counts for cluster1
  c1_counts <- sub_df %>%
    filter(cluster == cluster1) %>%
   arrange(factor(status, levels = status_levels)) %>%  # ensure same order
    pull(count)
  
  # extract counts for cluster2
  c2_counts <- sub_df %>%
    filter(cluster == cluster2) %>%
    arrange(factor(status, levels = status_levels)) %>%
    pull(count)
  
  # build 2x2 matrix
  mat <- matrix(
    c(c2_counts, c1_counts),
    nrow = 2,
    byrow = TRUE,
    dimnames = list(
      cluster = c(cluster2, cluster1),
      category = status_levels
    )
  )
  
  ft <- tryCatch(
    fisher.test(mat, workspace = 2e8),
    error = function(e) fisher.test(mat, simulate.p.value = TRUE, B = 1e5)
  )
  
  tibble(
    TE_type     = te_name,
    comparison  = paste(cluster2, "vs", cluster1),
    odds_ratio  = ft$estimate,
    lower_CI    = ft$conf.int[1],
    upper_CI    = ft$conf.int[2],
    p_value     = ft$p.value
  )
}
```


### SINE
Overlapping SINE family with summits to get a count

```{r SINE Overlapping}
sine_hits <- findOverlaps(H3K27ac_summit_gr_ext, SINE_gr, ignore.strand = TRUE)
SINE_overlap_df <- tibble(
  summit_id = queryHits(sine_hits),
  cluster   = mcols(H3K27ac_summit_gr)$cluster[queryHits(sine_hits)],
  TE_type   = mcols(SINE_gr)$repFamily[subjectHits(sine_hits)])

SINE_counts <- SINE_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_SINE_summits <- tibble(
  cluster = mcols(H3K27ac_summit_gr)$cluster
) %>% count(cluster, name = "total")

not_SINE_counts <- SINE_counts %>%
  left_join(total_SINE_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  select(cluster, TE_type, status, count)

SINE_df_long <- bind_rows(SINE_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_SINE_counts) %>% 
  filter(!is.na(cluster))

SINE_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      SINE_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            SINE_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))
```

```{r}
datatable(SINE_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_SINE_summits
```

```{r SINE datatable}
# ---- Prepare the table ----
SINE_counts_display <- SINE_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  SINE_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```

```{r Plot SINE}
ggplot(SINE_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "SINE family enrichment at extended ROI summits"
  ) +
  theme_classic() +
  facet_wrap(~comparison)
```



### LINE
Overlapping LINE family with summits to get a count
```{r LINE Overlapping}
LINE_hits <- findOverlaps(H3K27ac_summit_gr_ext, LINE_gr, ignore.strand = TRUE)
LINE_overlap_df <- tibble(
  summit_id = queryHits(LINE_hits),
  cluster   = mcols(H3K27ac_summit_gr)$cluster[queryHits(LINE_hits)],
  TE_type   = mcols(LINE_gr)$repFamily[subjectHits(LINE_hits)])

LINE_counts <- LINE_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_LINE_summits <- tibble(
  cluster = mcols(H3K27ac_summit_gr)$cluster
) %>% count(cluster, name = "total")

not_LINE_counts <- LINE_counts %>%
  left_join(total_LINE_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  select(cluster, TE_type, status, count)

LINE_df_long <- bind_rows(LINE_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_LINE_counts) %>% 
  filter(!is.na(cluster))

LINE_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      LINE_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            LINE_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

```

```{r}
datatable(LINE_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_LINE_summits
```

```{r LINE datatable}
# ---- Prepare the table ----
LINE_counts_display <- LINE_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  LINE_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```
```{r Plot LINE}
ggplot(LINE_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LINE family enrichment at extended ROI summits"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```


### LTR  

Overlapping LTR family summits to get a count

```{r LTR Overlapping}
LTR_hits <- findOverlaps(H3K27ac_summit_gr_ext, LTR_gr, ignore.strand = TRUE)
LTR_overlap_df <- tibble(
  summit_id = queryHits(LTR_hits),
  cluster   = mcols(H3K27ac_summit_gr)$cluster[queryHits(LTR_hits)],
  TE_type   = mcols(LTR_gr)$repFamily[subjectHits(LTR_hits)])

LTR_counts <- LTR_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_LTR_summits <- tibble(
  cluster = mcols(H3K27ac_summit_gr)$cluster
) %>% count(cluster, name = "total")

not_LTR_counts <- LTR_counts %>%
  left_join(total_LTR_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  select(cluster, TE_type, status, count)

LTR_df_long <- bind_rows(LTR_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_LTR_counts) %>% 
  filter(!is.na(cluster))

LTR_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      LTR_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            LTR_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

```

```{r}
datatable(LTR_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_LTR_summits
```

```{r LTR datatable}
# ---- Prepare the table ----
LTR_counts_display <- LTR_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  LTR_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```

```{r Plot LTR}
ggplot(LTR_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LTR family enrichment at extended ROI summits"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```


```{r}
LTR_hits <- findOverlaps(H3K27ac_summit_gr_ext, LTR_gr, ignore.strand = TRUE)
LTR_name_overlap_df <- tibble(
  summit_id = queryHits(LTR_hits),
  cluster   = mcols(H3K27ac_summit_gr)$cluster[queryHits(LTR_hits)],
  TE_type   = mcols(LTR_gr)$repFamily[subjectHits(LTR_hits)],
  repName= mcols(LTR_gr)$repName[subjectHits(LTR_hits)])

LTR_ERV1_counts <-
  LTR_name_overlap_df %>%
    dplyr::filter(TE_type=="ERV1") %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_LTR_summits <- tibble(
  cluster = mcols(H3K27ac_summit_gr)$cluster
) %>% count(cluster, name = "total")

not_LTR_ERV1_counts <- LTR_ERV1_counts %>%
  left_join(total_LTR_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  select(cluster, repName, status, count)

LTR_ERV1_df_long <- bind_rows(LTR_ERV1_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_LTR_ERV1_counts) %>% 
  filter(!is.na(cluster))

ERV1_LTR_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

       LTR_ERV1_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(rep) {
          test_pair_TE_repName(
            LTR_ERV1_df_long,
            rep_name  = rep,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

ggplot(ERV1_LTR_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(data = subset(ERV1_LTR_results, FDR < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LTR Name (type) enrichment at ROI extended-summits\nH3K27ac"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```






### SVA  

Overlapping SVA family with summits to get a count

```{r SVA Overlapping}
SVA_hits <- findOverlaps(H3K27ac_summit_gr_ext, SVA_gr, ignore.strand = TRUE)
SVA_overlap_df <- tibble(
  summit_id = queryHits(SVA_hits),
  cluster   = mcols(H3K27ac_summit_gr)$cluster[queryHits(SVA_hits)],
  TE_type   = mcols(SVA_gr)$repName[subjectHits(SVA_hits)])

SVA_counts <- SVA_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_SVA_summits <- tibble(
  cluster = mcols(H3K27ac_summit_gr)$cluster
) %>% count(cluster, name = "total")

not_SVA_counts <- SVA_counts %>%
  left_join(total_SVA_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  select(cluster, TE_type, status, count)

SVA_df_long <- bind_rows(SVA_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_SVA_counts) %>% 
  filter(!is.na(cluster))

SVA_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      SVA_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            SVA_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))
```

```{r}
datatable(SVA_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_SVA_summits
```

```{r SVA datatable}
# ---- Prepare the table ----
SVA_counts_display <- SVA_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  SVA_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```

```{r Plot SVA}
ggplot(SVA_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "SVA family enrichment at extended ROI summits"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```
### DNA  

Overlapping DNA family with summits to get a count


```{r DNA Overlapping}
DNA_hits <- findOverlaps(H3K27ac_summit_gr_ext, DNA_gr, ignore.strand = TRUE)
DNA_overlap_df <- tibble(
  summit_id = queryHits(DNA_hits),
  cluster   = mcols(H3K27ac_summit_gr)$cluster[queryHits(DNA_hits)],
  TE_type   = mcols(DNA_gr)$repFamily[subjectHits(DNA_hits)])

DNA_counts <- DNA_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_DNA_summits <- tibble(
  cluster = mcols(H3K27ac_summit_gr)$cluster
) %>% count(cluster, name = "total")

not_DNA_counts <- DNA_counts %>%
  left_join(total_DNA_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  select(cluster, TE_type, status, count)

DNA_df_long <- bind_rows(DNA_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_DNA_counts) %>% 
  filter(!is.na(cluster))

DNA_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      DNA_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            DNA_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))
```

```{r}
datatable(DNA_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_DNA_summits
```

```{r DNA datatable}
# ---- Prepare the table ----
DNA_counts_display <- DNA_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  DNA_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```

```{r Plot DNA}
ggplot(DNA_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "DNA family enrichment at extended ROI summits"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```

```{r making named bedfiles, eval=FALSE}
names(rpt_split_gr_list)

out_dir <- "data/Bed_exports/repeatmasker_sets"
set_list <- names(rpt_split_gr_list)

for (group_name in set_list) {
  gr <- rpt_split_gr_list[[group_name]]
  
  
  # Set BED name column
  mcols(gr)$name <- mcols(gr)$repName
  # 
  # # Optional: if you want a score column
  # if(!"score" %in% colnames(mcols(gr))) {
  #   mcols(gr)$score <- 0
  # }
  # 
  # Export to BED
 bed_file <- file.path(out_dir, paste0(group_name, "_rptmasker.bed"))
  
  # Export
  rtracklayer::export(gr, bed_file, format = "BED")
  
  cat("Exported:", bed_file, "\n")
}

```


```{r making named bedfiles 1, eval=FALSE}
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

out_dir <- "data/Bed_exports/repeatmasker_sets"
set_list <- names(peakAnnoList_H3K27ac)

for (group_name in set_list) {
  cs <- peakAnnoList_H3K27ac[[group_name]]
  gr <- cs@anno
  
  # Set BED name column
  mcols(gr)$name <- mcols(gr)$Peakid
  
  # Optional: if you want a score column
  if(!"score" %in% colnames(mcols(gr))) {
    mcols(gr)$score <- 0
  }
  
  # Export to BED
 bed_file <- file.path(out_dir, paste0(group_name, "_H3K27ac.bed"))
  
  # Export
  export(gr, bed_file, format = "BED")
  
  cat("Exported:", bed_file, "\n")
}

```


