---
title: "Using only topq hit of summits only"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```


```{r packages}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(BiocParallel)
library(parallel)
library(future.apply)
```

```{r calling summit files}
sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")
##Path to histone summit files
H3K27ac_dir <- "C:/Users/renee/Other_projects_data/DXR_data/final_data/summit_files/H3K27ac" 

##pull all histone files together
H3K27ac_summit_files <- list.files(
  path = H3K27ac_dir,
  pattern = "\\.bed$",
  recursive = TRUE,
  full.names = TRUE
)

```


### TOP qval individual summits  
First steps:  pulling in the information of Sets and locations
Pulling in the summit files,
concatenation of all summits into one large file

This is to pull the top qval summit for each ROI, then exxpand +/- 200 bp to check for enrichment of TFs.

```{r pulling in other data}
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")
H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})

read_summit <- function(file){
  peaks <- read.table(file,header = FALSE)
  colnames(peaks) <- c("chr","start","end","name","score")
  GRanges(
    seqnames = peaks$chr,
    ranges = IRanges(start=peaks$start, end = peaks$start),
    score=peaks$score,
    file=basename(file),
    Library_ID = stringr::str_remove(basename(file), "_FINAL_summits\\.bed$")
  )
}

all_H3K27ac_summits_list<- lapply(H3K27ac_summit_files, read_summit)
all_H3K27ac_summits_gr <- do.call(c, all_H3K27ac_summits_list)  # combine into one GRanges object
H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
```

```{r pick summit per roi}
pick_best_summit_per_roi <- function(roi_gr, summit_gr, score_col = "signalValue") {

  hits <- findOverlaps(roi_gr, summit_gr)

  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    score = mcols(summit_gr)[[score_col]][subjectHits(hits)]
  )

  # Pick summit with highest score for each ROI
  best <- df %>%
    group_by(roi_idx) %>%
    slice_max(order_by = score, n = 1, with_ties = FALSE) %>%
    ungroup()

  # Build the final GRanges object
  best_summits <- summit_gr[best$summit_idx]
  best_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]

  best_summits
}
# 
# best_summits <- pick_best_summit_per_roi(
#   ROIs, highest_summits_long_gr, score_col = "peakHeight"
# )
```


```{r apply summit and ROI single overlap}
ROIs <- H3K27ac_sets_gr$all_H3K27ac 
test <- pick_best_summit_per_roi(ROIs,all_H3K27ac_summits_gr, score_col = "score")
test %>% as.data.frame() %>% 
  group_by(ROI_name) #%>% tally() %>% dplyr::filter(n>1)



```


```{r}
idx <- match(test$ROI_name, H3K27ac_lookup$Peakid)
test$cluster <- H3K27ac_lookup$cluster[idx]
  test %>% as.data.frame()
```



```{r making set files, eval=FALSE}
### adding in cluster membership for export

library(BSgenome.Hsapiens.UCSC.hg38)
genome <- BSgenome.Hsapiens.UCSC.hg38
seqlengths(test) <- seqlengths(genome)[names(seqlengths(test))]

SET_1_gr <- test[
  !is.na(mcols(test)$cluster) &
  mcols(test)$cluster == "Set_1"]

resize_and_trim <- function(gr, flank = 300, genome = BSgenome.Hsapiens.UCSC.hg38) {
  gr <- resize(gr, width = 1 + 2*flank, fix = "center")
  seqlengths(gr) <- seqlengths(genome)[seqlevels(gr)]
  gr <- trim(gr)
  gr <- gr[width(gr) > 0]
  return(gr)
}

H3K27ac_set1_600 <- resize_and_trim(SET_1_gr,flank=300)
H3K27ac_set1_400 <- resize_and_trim(SET_1_gr,flank=200)
H3K27ac_set1_600<- trim(H3K27ac_set1_600)
H3K27ac_set1_600 <- H3K27ac_set1_600[width(H3K27ac_set1_600) > 0]

H3K27ac_set1_400 <- resize(SET_1_gr, width = 1 + 200*2, fix = "center")
H3K27ac_set1_400<- trim(H3K27ac_set1_400)
H3K27ac_set1_400 <- H3K27ac_set1_400[width(H3K27ac_set1_400) > 0]

rtracklayer::export(SET_1_gr, "data/Bed_exports/H3K27ac_Set_1_summits_alt.bed")
# rtracklayer::export(H3K27ac_set1_600, "data/Bed_exports/H3K27ac_Set_1_600_alt.bed")
rtracklayer::export(H3K27ac_set1_400, "data/Bed_exports/H3K27ac_Set_1_400_alt.bed")

SET_2_gr <- test[
  !is.na(mcols(test)$cluster) &
  mcols(test)$cluster == "Set_2"]

H3K27ac_set2_600 <- resize_and_trim(SET_2_gr,flank=300)
H3K27ac_set2_400 <- resize_and_trim(SET_2_gr,flank=400)

H3K27ac_set2_600 <- resize(SET_2_gr, width = 1 + 300*2, fix = "center")
H3K27ac_set2_600 <- trim(H3K27ac_set2_600)
H3K27ac_set2_600 <- H3K27ac_set2_600[width(H3K27ac_set2_600) > 0]

H3K27ac_set2_400 <- resize(SET_2_gr, width = 1 + 200*2, fix = "center")
H3K27ac_set2_400 <- trim(H3K27ac_set2_400)
H3K27ac_set2_400 <- H3K27ac_set2_400[width(H3K27ac_set2_400) > 0]
rtracklayer::export(SET_2_gr, "data/Bed_exports/H3K27ac_Set_2_summits_alt.bed")
# rtracklayer::export(H3K27ac_set2_600, "data/Bed_exports/H3K27ac_Set_2_600_alt.bed")
rtracklayer::export(H3K27ac_set2_400, "data/Bed_exports/H3K27ac_Set_2_400_alt.bed")

SET_3_gr <- test[
  !is.na(mcols(test)$cluster) &
  mcols(test)$cluster == "Set_3"]

H3K27ac_set3_600 <- resize_and_trim(SET_3_gr,flank=300)
H3K27ac_set3_400 <- resize_and_trim(SET_3_gr,flank=200)

H3K27ac_set3_600 <- resize(SET_3_gr, width = 1 + 300*2, fix = "center")
H3K27ac_set3_600 <- trim(H3K27ac_set3_600)
H3K27ac_set3_600 <- H3K27ac_set3_600[width(H3K27ac_set3_600) > 0]

H3K27ac_set3_400 <- resize(SET_3_gr, width = 1 + 200*2, fix = "center")
H3K27ac_set3_400 <- trim(H3K27ac_set3_400)
H3K27ac_set3_400 <- H3K27ac_set3_400[width(H3K27ac_set3_400) > 0]

rtracklayer::export(SET_3_gr, "data/Bed_exports/H3K27ac_Set_3_summits_alt.bed")
# rtracklayer::export(H3K27ac_set3_600, "data/Bed_exports/H3K27ac_Set_3_600.bed")
rtracklayer::export(H3K27ac_set3_400, "data/Bed_exports/H3K27ac_Set_3_400_alt.bed")
  

rtracklayer::export(test, "data/Bed_exports/H3K27ac_complete_final_summits_alt.bed")



```

### Top qval for concat summits


Here I first concatenated all trt-time (conditions) into one bam file, then ran peak calls on just those conditions.  I then used the summits from those sets to call on one summit per ROI, regardless of condition.  

```{r}
H3K27ac_alt_dir <- "C:/Users/renee/Other_projects_data/DXR_data/TACC_merged_peaks" 

##pull all histone files together
H3K27ac_alt_summit_files <- list.files(
  path = H3K27ac_dir,
  pattern = "\\.bed$",
  recursive = TRUE,
  full.names = TRUE
)

all_H3K27ac_summits_alt_list<- lapply(H3K27ac_alt_summit_files, read_summit)
all_H3K27ac_summits_alt_gr <- do.call(c, all_H3K27ac_summits_alt_list)  # combine into one GRanges object

```



```{r concat alt}
test_alt <- pick_best_summit_per_roi(ROIs,all_H3K27ac_summits_alt_gr, score_col = "score")
test_alt %>% as.data.frame() %>% 
  group_by(ROI_name) #%>% tally() %>% dplyr::filter(n>1)
idx_alt <- match(test_alt$ROI_name, H3K27ac_lookup$Peakid)
test_alt$cluster <- H3K27ac_lookup$cluster[idx_alt]
test_alt %>% as.data.frame()
```
Here is where and how I exported those files:

```{r, eval=FALSE}
library(BSgenome.Hsapiens.UCSC.hg38)
genome <- BSgenome.Hsapiens.UCSC.hg38
seqlengths(test_alt) <- seqlengths(genome)[names(seqlengths(test_alt))]

SET_1_gr <- test_alt[
  !is.na(mcols(test_alt)$cluster) &
  mcols(test_alt)$cluster == "Set_1"]

resize_and_trim <- function(gr, flank = 300, genome = BSgenome.Hsapiens.UCSC.hg38) {
  gr <- resize(gr, width = 1 + 2*flank, fix = "center")
  seqlengths(gr) <- seqlengths(genome)[seqlevels(gr)]
  gr <- trim(gr)
  gr <- gr[width(gr) > 0]
  return(gr)
}

# H3K27ac_set1_600 <- resize_and_trim(SET_1_gr,flank=300)
H3K27ac_set1_400 <- resize_and_trim(SET_1_gr,flank=200)
# H3K27ac_set1_600<- trim(H3K27ac_set1_600)
# H3K27ac_set1_600 <- H3K27ac_set1_600[width(H3K27ac_set1_600) > 0]

H3K27ac_set1_400 <- resize(SET_1_gr, width = 1 + 200*2, fix = "center")
H3K27ac_set1_400<- trim(H3K27ac_set1_400)
H3K27ac_set1_400 <- H3K27ac_set1_400[width(H3K27ac_set1_400) > 0]

rtracklayer::export(SET_1_gr, "data/Bed_exports/H3K27ac_Set_1_concat_summits_alt.bed")
# rtracklayer::export(H3K27ac_set1_600, "data/Bed_exports/H3K27ac_Set_1_600_alt.bed")
rtracklayer::export(H3K27ac_set1_400, "data/Bed_exports/H3K27ac_Set_1_400_alt_concat.bed")

SET_2_gr <- test_alt[
  !is.na(mcols(test_alt)$cluster) &
  mcols(test_alt)$cluster == "Set_2"]

# H3K27ac_set2_600 <- resize_and_trim(SET_2_gr,flank=300)
H3K27ac_set2_400 <- resize_and_trim(SET_2_gr,flank=400)

# H3K27ac_set2_600 <- resize(SET_2_gr, width = 1 + 300*2, fix = "center")
# H3K27ac_set2_600 <- trim(H3K27ac_set2_600)
# H3K27ac_set2_600 <- H3K27ac_set2_600[width(H3K27ac_set2_600) > 0]

H3K27ac_set2_400 <- resize(SET_2_gr, width = 1 + 200*2, fix = "center")
H3K27ac_set2_400 <- trim(H3K27ac_set2_400)
H3K27ac_set2_400 <- H3K27ac_set2_400[width(H3K27ac_set2_400) > 0]
rtracklayer::export(SET_2_gr, "data/Bed_exports/H3K27ac_Set_2_concat_summits_alt.bed")
# rtracklayer::export(H3K27ac_set2_600, "data/Bed_exports/H3K27ac_Set_2_600_alt.bed")
rtracklayer::export(H3K27ac_set2_400, "data/Bed_exports/H3K27ac_Set_2_400_alt_concat.bed")

SET_3_gr <- test_alt[
  !is.na(mcols(test_alt)$cluster) &
  mcols(test_alt)$cluster == "Set_3"]

# H3K27ac_set3_600 <- resize_and_trim(SET_3_gr,flank=300)
H3K27ac_set3_400 <- resize_and_trim(SET_3_gr,flank=200)

# H3K27ac_set3_600 <- resize(SET_3_gr, width = 1 + 300*2, fix = "center")
# H3K27ac_set3_600 <- trim(H3K27ac_set3_600)
# H3K27ac_set3_600 <- H3K27ac_set3_600[width(H3K27ac_set3_600) > 0]

H3K27ac_set3_400 <- resize(SET_3_gr, width = 1 + 200*2, fix = "center")
H3K27ac_set3_400 <- trim(H3K27ac_set3_400)
H3K27ac_set3_400 <- H3K27ac_set3_400[width(H3K27ac_set3_400) > 0]

rtracklayer::export(SET_3_gr, "data/Bed_exports/H3K27ac_Set_3_concat_summits_alt.bed")
# rtracklayer::export(H3K27ac_set3_600, "data/Bed_exports/H3K27ac_Set_3_600.bed")
rtracklayer::export(H3K27ac_set3_400, "data/Bed_exports/H3K27ac_Set_3_400_alt_concat.bed")
  

rtracklayer::export(test_alt, "data/Bed_exports/H3K27ac_complete_concat_summits_alt.bed")


```

