---
title: "H3K9me3 Extended ext_summits and TE overlap"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(data.table)
library(GenomicFeatures)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(ComplexHeatmap)
library(circlize)
library(readxl)
library(kableExtra)
```

First steps: breakdown repeatmasker into groups and pull out the ones by each class I am interested in.  
```{r repeatmasker loading}

repeatmasker <- read_delim("data/Other_paper_data/repeatmasker_20250911.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
colnames(repeatmasker)
##this is to remove sex chromosomes and non-canonical chromatin assignments
autosomes <- paste0("chr", 1:22)

repeatmasker_clean <- repeatmasker %>% mutate(
    strand = ifelse(strand == "C", "-", "+")
  ) %>% 
   mutate(
    start = genoStart + 1,
    end   = genoEnd)%>% 
  mutate(repFamily= str_remove(repFamily, "\\?$")) %>% 
  dplyr::filter(genoName %in% autosomes) %>% 
  mutate(RM_id=paste0(genoName,":",start,"-",end,":",id))


rpt_split <- split(repeatmasker_clean, repeatmasker_clean$repClass)

rpt_split_gr_list <- lapply(rpt_split, function(df) {
  GRanges(
    seqnames = df$genoName,
    ranges   = IRanges(start = df$start, end = df$end),
    strand   = df$strand,
    repName   = df$repName,
    repClass  = df$repClass,
    repFamily = df$repFamily,
    swScore   = df$swScore,
    milliDiv  = df$milliDiv,
    milliDel  = df$milliDel,
    milliIns  = df$milliIns,
    RM_id        = df$RM_id
  )
})
### full repeatmasker cleaned up granges frame
rmskr_std_gr <- GRanges(
  seqnames = repeatmasker_clean$genoName,
  ranges   = IRanges(
    start = repeatmasker_clean$start,
    end   = repeatmasker_clean$end
  ),
  strand   = repeatmasker_clean$strand,
  repName   = repeatmasker_clean$repName,
  repClass  = repeatmasker_clean$repClass,
  repFamily = repeatmasker_clean$repFamily,
  swScore   = repeatmasker_clean$swScore,
  milliDiv  = repeatmasker_clean$milliDiv,
  milliDel  = repeatmasker_clean$milliDel,
  milliIns  = repeatmasker_clean$milliIns,
  id        = repeatmasker_clean$RM_id
)

```

```{r subset rpt_split_gr_list}
SINE_gr <- rpt_split_gr_list$SINE
SINE_df <- SINE_gr %>% 
  as.data.frame()

SINE_split_df <- split(SINE_df, SINE_df$repFamily)

LINE_gr <- rpt_split_gr_list$LINE
LINE_df <- LINE_gr %>% 
  as.data.frame()

LINE_split_df <- split(LINE_df, LINE_df$repFamily)

LTR_gr <- rpt_split_gr_list$LTR
LTR_df <- LTR_gr %>% 
  as.data.frame()

LTR_split_df <- split(LTR_df, LTR_df$repFamily)

SVA_gr <- rpt_split_gr_list$Retroposon
SVA_df <- SVA_gr %>% 
  as.data.frame()

SVA_split_df <- split(SVA_df, SVA_df$repFamily)

DNA_gr <- rpt_split_gr_list$DNA
DNA_df <- DNA_gr %>% 
  as.data.frame()

DNA_split_df <- split(DNA_df, DNA_df$repFamily)
```

```{r loading H3K9me3 data}
H3K9me3_summit_gr <- readRDS("data/RDS_files/H3K9me3_complete_summit_gr.RDS")


peakAnnoList_H3K9me3 <- readRDS("data/motif_lists/H3K9me3_annotated_peaks.RDS")

H3K9me3_lookup <- imap_dfr(peakAnnoList_H3K9me3[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
### making these object as granges objects
H3K9me3_sets_gr <- lapply(peakAnnoList_H3K9me3, function(df) {
  as_granges(df)
})

##assigning Peakid as name of summit region
 mcols(H3K9me3_summit_gr)$name <- mcols(H3K9me3_summit_gr)$Peakid
 
 comparisons <- tibble(
  cluster2 = c("Set_2", "Set_3"),
  cluster1 = c("Set_1", "Set_1")
)

 H3K9me3_summit_gr_ext <- resize(H3K9me3_summit_gr, width(H3K9me3_summit_gr)+400, fix = "center")
```

```{r Fisher test information}

# Generic pairwise Fisher test
test_pair_TE_generic <- function(df_long, te_name, cluster1, cluster2) {
  
   sub_df <- df_long %>%
    filter(TE_type == te_name) %>%
    complete(
      cluster = c(cluster1, cluster2),
      status  = c("TE", "not_TE"),
      fill = list(count = 0))
   
   # enforce fixed order
 status_levels <- c("TE", "not_TE")
  
  # assume "status" column has TE vs wnot_TE automatically
  statuses <- unique(sub_df$status)
  
  if(length(statuses) != 2) {
    # ensure we have exactly two categories, fill missing with 0
    sub_df <- sub_df %>%
      complete(cluster, status, fill = list(count = 0))
    statuses <- unique(sub_df$status)
  }
  
  # extract counts for cluster1
  c1_counts <- sub_df %>%
    filter(cluster == cluster1) %>%
   arrange(factor(status, levels = status_levels)) %>%  # ensure same order
    pull(count)
  
  # extract counts for cluster2
  c2_counts <- sub_df %>%
    filter(cluster == cluster2) %>%
    arrange(factor(status, levels = status_levels)) %>%
    pull(count)
  
  # build 2x2 matrix
  mat <- matrix(
    c(c2_counts, c1_counts),
    nrow = 2,
    byrow = TRUE,
    dimnames = list(
      cluster = c(cluster2, cluster1),
      category = status_levels
    )
  )
  
  ft <- tryCatch(
    fisher.test(mat, workspace = 2e8),
    error = function(e) fisher.test(mat, simulate.p.value = TRUE, B = 1e5)
  )
  
  tibble(
    TE_type     = te_name,
    comparison  = paste(cluster2, "vs", cluster1),
    odds_ratio  = ft$estimate,
    lower_CI    = ft$conf.int[1],
    upper_CI    = ft$conf.int[2],
    p_value     = ft$p.value
  )
}
```

```{r}
test_pair_TE_repName <- function(df_long, rep_name, cluster1, cluster2) {
  
  # Subset for the specific repName
  sub_df <- df_long %>%
    filter(repName == rep_name) %>%
    complete(
      cluster = c(cluster1, cluster2),
      status  = c("TE", "not_TE"),
      fill = list(count = 0)
    )
  
  # fixed order of statuses
  status_levels <- c("TE", "not_TE")
  
  # make sure both statuses exist
  statuses <- unique(sub_df$status)
  if(length(statuses) != 2) {
    sub_df <- sub_df %>%
      complete(cluster, status, fill = list(count = 0))
    statuses <- unique(sub_df$status)
  }
  
  # counts for cluster1
  c1_counts <- sub_df %>%
    filter(cluster == cluster1) %>%
    arrange(factor(status, levels = status_levels)) %>%
    pull(count)
  
  # counts for cluster2
  c2_counts <- sub_df %>%
    filter(cluster == cluster2) %>%
    arrange(factor(status, levels = status_levels)) %>%
    pull(count)
  
  # 2x2 matrix for Fisher test
  mat <- matrix(
    c(c2_counts, c1_counts),
    nrow = 2,
    byrow = TRUE,
    dimnames = list(
      cluster = c(cluster2, cluster1),
      category = status_levels
    )
  )
  
  ft <- tryCatch(
    fisher.test(mat, workspace = 2e8),
    error = function(e) fisher.test(mat, simulate.p.value = TRUE, B = 1e5)
  )
  
  tibble(
    repName     = rep_name,
    comparison  = paste(cluster2, "vs", cluster1),
    odds_ratio  = ft$estimate,
    lower_CI    = ft$conf.int[1],
    upper_CI    = ft$conf.int[2],
    p_value     = ft$p.value
  )
}

```
### Overlapping all repeatframes 200bp

```{r overlapping all, eval=FALSE}
dataf <- H3K9me3_summit_gr_ext

  
   # 1️⃣ find overlaps
  hits <- findOverlaps(dataf, rmskr_std_gr)
  
  if (length(hits) == 0) {
  H3K9me3_ext200_ols <- tibble()
} else {

  H3K9me3_ext200_ols <- tibble(
    Peakid     = dataf$Peakid[queryHits(hits)],
    cluster    = dataf$cluster[queryHits(hits)],
    seqnames   = as.character(seqnames(dataf))[queryHits(hits)],
    summit_pos = start(dataf)[queryHits(hits)],

    repName   = rmskr_std_gr$repName[subjectHits(hits)],
    repFamily = rmskr_std_gr$repFamily[subjectHits(hits)],
    repClass  = rmskr_std_gr$repClass[subjectHits(hits)],
    milliDiv  = rmskr_std_gr$milliDiv[subjectHits(hits)],
    milliDel  = rmskr_std_gr$milliDel[subjectHits(hits)],
    milliIns  = rmskr_std_gr$milliIns[subjectHits(hits)],
    ID  = rmskr_std_gr$id[subjectHits(hits)]
  )
}
  
  All_H3K9me3 <-H3K9me3_sets_gr$all_H3K9me3_regions
annotated_H3K9me3_ext200 <-
  All_H3K9me3 %>% 
      as.data.frame() %>% 
    dplyr::select(Peakid,annotation,geneId:distanceToTSS) %>% 
     left_join(., (H3K9me3_ext200_ols %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(repClass, collapse=";"),
                            repFamily=paste0(repFamily,collapse=";"),
                            repName=paste0(repName,collapse=";"),
                            repID=paste0(ID, collapse = ";"),
                            summit_pos=paste0(unique(summit_pos),collapse = ";"),
                            .groups = "drop")))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
    left_join(H3K9me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))
  
# saveRDS(annotated_H3K9me3_ext200,"data/TE_annotation/annotated_H3K9me3_ext200_TE_overlaps.RDS")
```


```{r anddin in anno-summits only}
te_cols <- c("TE_status","SINE_status","LINE_status","DNA_status","LTR_status","SVA_status")
annotated_H3K9me3_summits <- readRDS("data/TE_annotation/H3K9me3_annotated_summit_TE_overlaps.RDS")

annotated_H3K9me3_ext200 <- readRDS("data/TE_annotation/annotated_H3K9me3_ext200_TE_overlaps.RDS")
```

### Summits only  


```{r running summit cluster analysis}

H3K9me3_counts_summit <- map(te_cols, ~ {
  annotated_H3K9me3_summits %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)

cluster_pairs_3 <- list(
  c("Set_2","Set_1"),
  c("Set_3","Set_1")
  # add more pairs if needed or
)
# cluster_pairs_2 <- list(
#   c("Set_2","Set_1"))

H3K9me3_counts_summit_long <- map2_dfr(
  H3K9me3_counts_summit,      # the named list of wide tables
  names(H3K9me3_counts_summit), # names of TE columns (same names of TE columns)
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / wnot_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)
H3K9me3_counts_summit_long <-
  H3K9me3_counts_summit_long %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()

H3K9me3_count_summit_long_fixed <- 
H3K9me3_counts_summit_long %>% 
  mutate(status=case_when(str_detect(status,"^wnot_")~"not_TE",
                          TRUE~"TE"))
TE_types <- te_cols

H3K9me3_summit_results<- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_3, function(pair){
    test_pair_TE_generic(H3K9me3_count_summit_long_fixed, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})
H3K9me3_summit_results%>% 
  mutate(TE_label=str_remove(TE_type,"_status$")) %>% 
  ggplot(., aes(x = (odds_ratio), y = -log10(p_value),label = TE_label)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "TE class enrichment at ROIs\nH3K9me3 summits"
  ) +
  theme_classic() +
  facet_wrap(~comparison)

```


### Extended region analysis  


```{r running ext set cluster analysis}

H3K9me3_counts_ext200 <- map(te_cols, ~ {
  annotated_H3K9me3_ext200 %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)

cluster_pairs_3 <- list(
  c("Set_2","Set_1"),
  c("Set_3","Set_1")
  # add more pairs if needed or
)
# cluster_pairs_2 <- list(
#   c("Set_2","Set_1"))

H3K9me3_counts_ext200_long <- map2_dfr(
  H3K9me3_counts_ext200,      # the named list of wide tables
  names(H3K9me3_counts_ext200), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / wnot_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)
H3K9me3_counts_ext200_long <-
  H3K9me3_counts_ext200_long %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()
H3K9me3_count_ext200_long_fixed <- 
H3K9me3_counts_ext200_long %>% 
  mutate(status=case_when(str_detect(status,"^wnot_")~"not_TE",
                          TRUE~"TE"))
```

```{r}
TE_types <- te_cols


H3K9me3_ext200_results<- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_3, function(pair){
    test_pair_TE_generic(H3K9me3_count_ext200_long_fixed, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})
H3K9me3_ext200_results%>% 
  mutate(TE_label=str_remove(TE_type,"_status$")) %>% 
  ggplot(., aes(x = (odds_ratio), y = -log10(p_value),label = TE_label)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "TE class enrichment at ROIs\nH3K9me3 200bp ext"
  ) +
  theme_classic() +
  facet_wrap(~comparison)
```

### SINE
Overlapping SINE family with ext_summits to get a count

```{r SINE Overlapping}
sine_hits <- findOverlaps(H3K9me3_summit_gr_ext, SINE_gr, ignore.strand = TRUE)
SINE_overlap_df <- tibble(
  summit_id = queryHits(sine_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(sine_hits)],
  TE_type   = mcols(SINE_gr)$repFamily[subjectHits(sine_hits)])

SINE_counts <- SINE_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_SINE_ext_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_SINE_counts <- SINE_counts %>%
  left_join(total_SINE_ext_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, TE_type, status, count)

SINE_df_long <- bind_rows(SINE_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_SINE_counts) %>% 
  filter(!is.na(cluster))

SINE_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      SINE_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            SINE_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))
```

```{r}
datatable(SINE_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_SINE_ext_summits
```

```{r SINE datatable}
# ---- Prepare the table ----
SINE_counts_display <- SINE_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  SINE_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```

```{r Plot SINE}
ggplot(SINE_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "SINE family enrichment at extended ROI ext_summits \nH3K9me3"
  ) +
  theme_classic() +
  facet_wrap(~comparison)
```

#### SINE Alu  


```{r SINE Alu Overlapping}
SINE_hits <- findOverlaps(H3K9me3_summit_gr_ext, SINE_gr, ignore.strand = TRUE)
SINE_name_overlap_df <- tibble(
  summit_id = queryHits(SINE_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(SINE_hits)],
  TE_type   = mcols(SINE_gr)$repFamily[subjectHits(SINE_hits)],
  repName= mcols(SINE_gr)$repName[subjectHits(SINE_hits)])

SINE_Alu_counts <-
  SINE_name_overlap_df %>%
    dplyr::filter(TE_type=="Alu") %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_SINE_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_SINE_Alu_counts <- SINE_Alu_counts %>%
  left_join(total_SINE_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

SINE_Alu_df_long <- bind_rows(SINE_Alu_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_SINE_Alu_counts) %>% 
  filter(!is.na(cluster))

Alu_SINE_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

       SINE_Alu_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(rep) {
          test_pair_TE_repName(
            SINE_Alu_df_long,
            rep_name  = rep,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

ggplot(Alu_SINE_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(data = subset(Alu_SINE_results, FDR < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "SINE Name (Alu) enrichment at ROI extended-summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```

#### SINE MIR  


```{r SINE MIR Overlapping}
# SINE_hits <- findOverlaps(H3K9me3_summit_gr_ext, SINE_gr, ignore.strand = TRUE)
# SINE_name_overlap_df <- tibble(
#   summit_id = queryHits(SINE_hits),
#   cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(SINE_hits)],
#   TE_type   = mcols(SINE_gr)$repFamily[subjectHits(SINE_hits)],
#   repName= mcols(SINE_gr)$repName[subjectHits(SINE_hits)])

SINE_MIR_counts <-
  SINE_name_overlap_df %>%
    dplyr::filter(TE_type=="MIR") %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_SINE_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_SINE_MIR_counts <- SINE_MIR_counts %>%
  left_join(total_SINE_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

SINE_MIR_df_long <- bind_rows(SINE_MIR_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_SINE_MIR_counts) %>% 
  filter(!is.na(cluster))

MIR_SINE_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

       SINE_MIR_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(rep) {
          test_pair_TE_repName(
            SINE_MIR_df_long,
            rep_name  = rep,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

ggplot(MIR_SINE_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(data = subset(MIR_SINE_results, FDR < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "SINE Name (MIR) enrichment at ROI extended-summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```
### LINE
Overlapping LINE family with ext_summits to get a count
```{r LINE Overlapping}
LINE_hits <- findOverlaps(H3K9me3_summit_gr_ext, LINE_gr, ignore.strand = TRUE)
LINE_overlap_df <- tibble(
  summit_id = queryHits(LINE_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(LINE_hits)],
  TE_type   = mcols(LINE_gr)$repFamily[subjectHits(LINE_hits)])

LINE_counts <- LINE_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_LINE_ext_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_LINE_counts <- LINE_counts %>%
  left_join(total_LINE_ext_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, TE_type, status, count)

LINE_df_long <- bind_rows(LINE_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_LINE_counts) %>% 
  filter(!is.na(cluster))

LINE_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      LINE_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            LINE_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

```

```{r}
datatable(LINE_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_LINE_ext_summits
```

```{r LINE datatable}
# ---- Prepare the table ----
LINE_counts_display <- LINE_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  LINE_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```

```{r Plot LINE}
ggplot(LINE_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LINE family enrichment at extended ROI ext_summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```

#### LINE L1  

```{r LINE L1 Overlapping}
# LINE_hits <- findOverlaps(H3K9me3_summit_gr_ext, LINE_gr, ignore.strand = TRUE)
LINE_name_overlap_df <- tibble(
  summit_id = queryHits(LINE_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(LINE_hits)],
  TE_type   = mcols(LINE_gr)$repFamily[subjectHits(LINE_hits)],
  repName= mcols(LINE_gr)$repName[subjectHits(LINE_hits)])

LINE_L1_counts <-
  LINE_name_overlap_df %>%
    dplyr::filter(TE_type=="L1") %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_LINE_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_LINE_L1_counts <- LINE_L1_counts %>%
  left_join(total_LINE_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

LINE_L1_df_long <- bind_rows(LINE_L1_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_LINE_L1_counts) %>% 
  filter(!is.na(cluster))

L1_LINE_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

       LINE_L1_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(rep) {
          test_pair_TE_repName(
            LINE_L1_df_long,
            rep_name  = rep,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

ggplot(L1_LINE_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(data = subset(L1_LINE_results, FDR < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LINE Name (L1) enrichment at ROI extended-summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```

### LTR  

Overlapping LTR family ext_summits to get a count

```{r LTR Overlapping}
LTR_hits <- findOverlaps(H3K9me3_summit_gr_ext, LTR_gr, ignore.strand = TRUE)
LTR_overlap_df <- tibble(
  summit_id = queryHits(LTR_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(LTR_hits)],
  TE_type   = mcols(LTR_gr)$repFamily[subjectHits(LTR_hits)])

LTR_counts <- LTR_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_LTR_ext_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr)$cluster
) %>% count(cluster, name = "total")

not_LTR_counts <- LTR_counts %>%
  left_join(total_LTR_ext_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, TE_type, status, count)

LTR_df_long <- bind_rows(LTR_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_LTR_counts) %>% 
  filter(!is.na(cluster))

LTR_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      LTR_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            LTR_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

```

```{r}
datatable(LTR_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_LTR_ext_summits
```

```{r LTR datatable}
# ---- Prepare the table ----
LTR_counts_display <- LTR_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  LTR_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```

```{r Plot LTR}
ggplot(LTR_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LTR family enrichment at extended ROI ext_summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```

#### LTR ERV1  


```{r LTR ERV1}
LTR_hits <- findOverlaps(H3K9me3_summit_gr_ext, LTR_gr, ignore.strand = TRUE)
LTR_name_overlap_df <- tibble(
  summit_id = queryHits(LTR_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(LTR_hits)],
  TE_type   = mcols(LTR_gr)$repFamily[subjectHits(LTR_hits)],
  repName= mcols(LTR_gr)$repName[subjectHits(LTR_hits)])

LTR_ERV1_counts <-
  LTR_name_overlap_df %>%
    dplyr::filter(TE_type=="ERV1") %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_LTR_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_LTR_ERV1_counts <- LTR_ERV1_counts %>%
  left_join(total_LTR_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

LTR_ERV1_df_long <- bind_rows(LTR_ERV1_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_LTR_ERV1_counts) %>% 
  filter(!is.na(cluster))

ERV1_LTR_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

       LTR_ERV1_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(rep) {
          test_pair_TE_repName(
            LTR_ERV1_df_long,
            rep_name  = rep,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

ggplot(ERV1_LTR_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(data = subset(ERV1_LTR_results, FDR < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LTR Name (type) enrichment at ROI extended-summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```


#### LTR ERVK  


```{r LTR ERVK}
LTR_hits <- findOverlaps(H3K9me3_summit_gr_ext, LTR_gr, ignore.strand = TRUE)
LTR_name_overlap_df <- tibble(
  summit_id = queryHits(LTR_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(LTR_hits)],
  TE_type   = mcols(LTR_gr)$repFamily[subjectHits(LTR_hits)],
  repName= mcols(LTR_gr)$repName[subjectHits(LTR_hits)])

LTR_ERVK_counts <-
  LTR_name_overlap_df %>%
    dplyr::filter(TE_type=="ERVK") %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_LTR_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_LTR_ERVK_counts <- LTR_ERVK_counts %>%
  left_join(total_LTR_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

LTR_ERVK_df_long <- bind_rows(LTR_ERVK_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_LTR_ERVK_counts) %>% 
  filter(!is.na(cluster))

ERVK_LTR_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

       LTR_ERVK_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(rep) {
          test_pair_TE_repName(
            LTR_ERVK_df_long,
            rep_name  = rep,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

ggplot(ERVK_LTR_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(data = subset(ERVK_LTR_results, FDR < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LTR Name (ERVK) enrichment at ROI extended-summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```

#### LTR ERVL  


```{r LTR ERVL}
LTR_hits <- findOverlaps(H3K9me3_summit_gr_ext, LTR_gr, ignore.strand = TRUE)
LTR_name_overlap_df <- tibble(
  summit_id = queryHits(LTR_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(LTR_hits)],
  TE_type   = mcols(LTR_gr)$repFamily[subjectHits(LTR_hits)],
  repName= mcols(LTR_gr)$repName[subjectHits(LTR_hits)])

LTR_ERVL_counts <-
  LTR_name_overlap_df %>%
    dplyr::filter(TE_type=="ERVL") %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_LTR_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_LTR_ERVL_counts <- LTR_ERVL_counts %>%
  left_join(total_LTR_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

LTR_ERVL_df_long <- bind_rows(LTR_ERVL_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_LTR_ERVL_counts) %>% 
  filter(!is.na(cluster))

ERVL_LTR_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

       LTR_ERVL_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(rep) {
          test_pair_TE_repName(
            LTR_ERVL_df_long,
            rep_name  = rep,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

ggplot(ERVL_LTR_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(data = subset(ERVL_LTR_results, FDR < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LTR Name (ERVL) enrichment at ROI extended-summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```

#### LTR ERVL-MaLR  


```{r LTR ERVL-MaLR}
LTR_hits <- findOverlaps(H3K9me3_summit_gr_ext, LTR_gr, ignore.strand = TRUE)
LTR_name_overlap_df <- tibble(
  summit_id = queryHits(LTR_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(LTR_hits)],
  TE_type   = mcols(LTR_gr)$repFamily[subjectHits(LTR_hits)],
  repName= mcols(LTR_gr)$repName[subjectHits(LTR_hits)])

LTR_ERVLMaLR_counts <-
  LTR_name_overlap_df %>%
    dplyr::filter(TE_type=="ERVL-MaLR") %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_LTR_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_LTR_ERVLMaLR_counts <- LTR_ERVLMaLR_counts %>%
  left_join(total_LTR_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

LTR_ERVLMaLR_df_long <- bind_rows(LTR_ERVLMaLR_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_LTR_ERVLMaLR_counts) %>% 
  filter(!is.na(cluster))

ERVLMaLR_LTR_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

       LTR_ERVLMaLR_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(rep) {
          test_pair_TE_repName(
            LTR_ERVLMaLR_df_long,
            rep_name  = rep,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

ggplot(ERVLMaLR_LTR_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(data = subset(ERVLMaLR_LTR_results, FDR < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "LTR Name (ERVL-MaLR) enrichment at ROI extended-summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```






### SVA  

Overlapping SVA family with ext_summits to get a count

```{r SVA Overlapping}
SVA_hits <- findOverlaps(H3K9me3_summit_gr_ext, SVA_gr, ignore.strand = TRUE)
SVA_overlap_df <- tibble(
  summit_id = queryHits(SVA_hits),
  cluster   = mcols(H3K9me3_summit_gr)$cluster[queryHits(SVA_hits)],
  TE_type   = mcols(SVA_gr)$repName[subjectHits(SVA_hits)])

SVA_counts <- SVA_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_SVA_ext_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr)$cluster
) %>% count(cluster, name = "total")

not_SVA_counts <- SVA_counts %>%
  left_join(total_SVA_ext_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, TE_type, status, count)

SVA_df_long <- bind_rows(SVA_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_SVA_counts) %>% 
  filter(!is.na(cluster))

SVA_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      SVA_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            SVA_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))
```

```{r}
datatable(SVA_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_SVA_ext_summits
```

```{r SVA datatable}
# ---- Prepare the table ----
SVA_counts_display <- SVA_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  SVA_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```

```{r Plot SVA}
ggplot(SVA_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "SVA family enrichment at extended ROI ext_summits"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```
### DNA  

Overlapping DNA family with ext_summits to get a count


```{r DNA Overlapping}
DNA_hits <- findOverlaps(H3K9me3_summit_gr_ext, DNA_gr, ignore.strand = TRUE)
DNA_overlap_df <- tibble(
  summit_id = queryHits(DNA_hits),
  cluster   = mcols(H3K9me3_summit_gr)$cluster[queryHits(DNA_hits)],
  TE_type   = mcols(DNA_gr)$repFamily[subjectHits(DNA_hits)])

DNA_counts <- DNA_overlap_df %>%
  count(cluster, TE_type, name = "count") %>%
  mutate(status = "TE")

total_DNA_ext_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr)$cluster
) %>% count(cluster, name = "total")

not_DNA_counts <- DNA_counts %>%
  left_join(total_DNA_ext_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, TE_type, status, count)

DNA_df_long <- bind_rows(DNA_counts %>%
                            dplyr::select(cluster, TE_type, status, count),
                          not_DNA_counts) %>% 
  filter(!is.na(cluster))

DNA_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      DNA_df_long %>%
        distinct(TE_type) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_generic(
            DNA_df_long,
            te_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))
```

```{r}
datatable(DNA_counts,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE))
total_DNA_ext_summits
```

```{r DNA datatable}
# ---- Prepare the table ----
DNA_counts_display <- DNA_results %>%
  # 1. Split comparison
  separate(comparison, into = c("cluster2", "cluster1"), sep = " vs ", remove = FALSE) %>%
  
  # 2. Add log2 odds ratio
  mutate(log2OR = log2(odds_ratio)) %>%
  
  # 3. Add enrichment/depletion direction
  mutate(direction = case_when(
    odds_ratio > 1 ~ "enriched",
    odds_ratio < 1 ~ "depleted",
    TRUE           ~ "neutral"
  )) %>%
  
  # 4. Flag significant
  mutate(significant = FDR < 0.05) %>%
  
  # Optional: arrange for readability
  arrange(cluster2, direction, desc(log2OR))

# ---- Create interactive datatable ----
datatable(
  DNA_counts_display,
  rownames = FALSE,
  filter = 'top',       # add filter/search boxes
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  )
) %>%
  # Conditional coloring by direction
  formatStyle(
    'direction',
    target = 'row',
    backgroundColor = styleEqual(
      c('enriched', 'depleted'),
      c('#FFDD99', '#99CCFF')  # enriched = light orange, depleted = light blue
    )
  ) %>%
  # Bold significant rows
  formatStyle(
    'significant',
    fontWeight = styleEqual(TRUE, 'bold')
  ) %>%
  # Round numeric columns for readability
  formatRound(columns = c('odds_ratio','log2OR','FDR'), digits = 2)
```

```{r Plot DNA}
ggplot(DNA_results, aes(x = (odds_ratio), y = -log10(FDR),label = TE_type)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "DNA family enrichment at extended ROI ext_summits"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```
#### hAT-Charlie-DNA

```{r DNA hATCharlie Overlapping}
DNA_hits <- findOverlaps(H3K9me3_summit_gr_ext, DNA_gr, ignore.strand = TRUE)
DNA_name_overlap_df <- tibble(
  summit_id = queryHits(DNA_hits),
  cluster   = mcols(H3K9me3_summit_gr_ext)$cluster[queryHits(DNA_hits)],
  TE_type   = mcols(DNA_gr)$repFamily[subjectHits(DNA_hits)],
  repName= mcols(DNA_gr)$repName[subjectHits(DNA_hits)])

DNA_hATCharlie_counts <-
  DNA_name_overlap_df %>%
    dplyr::filter(TE_type=="hAT-Charlie") %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_DNA_summits <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_DNA_hATCharlie_counts <- DNA_hATCharlie_counts %>%
  left_join(total_DNA_summits, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

DNA_hATCharlie_df_long <- bind_rows(DNA_hATCharlie_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_DNA_hATCharlie_counts) %>% 
  filter(!is.na(cluster))

hATCharlie_DNA_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

       DNA_hATCharlie_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(rep) {
          test_pair_TE_repName(
            DNA_hATCharlie_df_long,
            rep_name  = rep,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

ggplot(hATCharlie_DNA_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(data = subset(hATCharlie_DNA_results, FDR < 0.05),  # only significant points
    size = 3,
    max.overlaps = 100
  ) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "DNA Name (hAT-Charlie) enrichment at ROI extended-summits\nH3K9me3"
  ) +
  theme_classic()+
  facet_wrap(~comparison)
```

### Investigating proportions


```{r}

anno_H3K9me3_ext200 <- readRDS("data/TE_annotation/annotated_H3K9me3_ext200_TE_overlaps.RDS")
anno_H3K9me3_ext200 %>% 
  
  group_by(cluster) %>% 
  count() %>%
  ungroup() %>% 
  mutate(percent = n / sum(n)) %>% 
  
  ggplot(.,aes(x=cluster, y=percent, fill=cluster)) +
  geom_col() +
    geom_text(
    aes(label = sprintf("%.3f", percent)),
    vjust = -0.3,
    size = 3
  ) +
  ylab("proportion")+
  ggtitle("Proportion of ROIs by cluster")+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )

```
```{r pie overlap of number of elements}
H3K9me3_te_summary <- anno_H3K9me3_ext200 %>% 
   mutate(repClass = na_if(repClass, ""),   # treat empty as NA
    n_TE_class = case_when(
      is.na(repClass) ~ 0L,
      TRUE ~ lengths(strsplit(repClass, ";")))) %>% 
  mutate( category = case_when(n_TE_class == 0 ~ "No TE overlap",
                               n_TE_class == 1 ~ "Single TE class",
                               n_TE_class > 1  ~ "Multiple TE classes"))


pie_df <- H3K9me3_te_summary %>% 
  count(category) %>% 
  mutate(
    percent = n / sum(n) * 100
  )

ggplot(pie_df, aes(x = "", y = percent, fill = category)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  geom_text(
    aes(label = sprintf("%.1f%%", percent)),
    position = position_stack(vjust = 0.5),
    size = 4
  ) +
  labs(
    title = "ROI summit-ext overlap with transposable elements"
  )
```
### TE class distribution  

of  ROIs that overlap TEs, what is the distribution across TE families?
```{r classROI}
H3K9me3_te_family_long <- H3K9me3_te_summary %>% 
  filter(!is.na(repClass), repClass != "") %>%   # TE-positive only
  separate_rows(repClass,repFamily,repName, repID, sep = ";") %>% 
  distinct(Peakid, repClass,cluster)


fam_pie_df <- H3K9me3_te_family_long %>% 
  count(repClass, name = "n") %>% 
  mutate(percent = n / sum(n)*100) %>% 
  mutate(
    repClass = if_else(percent < 2, "Other", repClass)
  ) %>% 
  group_by(repClass) %>% 
  summarise(n = sum(n), .groups = "drop") %>% 
  mutate(percent = n / sum(n) * 100)
  

ggplot(fam_pie_df, aes(x = "", y = percent, fill = repClass)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  geom_text(
    aes(label = sprintf("%s\n%.1f%%", repClass, percent)),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  labs(
    title = "TE Class composition among summit-ext TE-overlapping H3K9me3 ROIs",
    fill = "TE Class")
# unique(rpt_split_gr_list$LTR$repFamily)
```
```{r bar overlap data}
rpt_genome <- repeatmasker_clean %>% 
 mutate(cluster = "hg38") %>% 
  group_by(cluster, repClass) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  mutate(percent = n / sum(n) * 100) %>% 
  mutate(
    repClass = if_else(percent < 1.2, "Other", repClass)
  ) %>% 
  group_by(repClass) %>% 
  summarise(n = sum(n), .groups = "drop") %>% 
  mutate(percent = n / sum(n) * 100)


H3K9me3_te_family_clust <- H3K9me3_te_family_long %>% 
  dplyr::filter(cluster != "not_assigned") %>%
   distinct(Peakid, cluster, repClass) %>%
    count(cluster, repClass, name = "n") %>% 
  group_by(cluster) %>% 
  mutate(percent = n / sum(n)*100) %>% 
  mutate(
    repClass = if_else(percent < 2, "Other", repClass)
  ) %>% 
  group_by(cluster,repClass) %>% 
  summarise(n = sum(n), .groups = "drop") %>% 
  group_by(cluster) %>% 
  mutate(percent = n / sum(n) * 100) %>% 
  rbind((rpt_genome %>% mutate(cluster="hg38")))

# ggplot(H3K9me3_te_family_clust, aes(x = "", y=percent, fill=repClass))+
#   geom_col(width =1, color="white")+
#   coord_polar(theta = "y")+
#   geom_text(
#     aes(x= 1.5,label = sprintf("%s\n%.1f%%", repClass, percent)),
#     position = position_stack(vjust = 0.5),
#     size = 3
#   ) +
#   facet_wrap(~cluster)+
#   theme_void()+
#   labs(
#     title = "TE Class composition among TE-overlapping ROIs per set",
#     fill = "TE family"
#   )

ggplot(H3K9me3_te_family_clust, aes(x = cluster, y=percent, fill=repClass))+
   geom_col(width = 0.8, color = "white") +
  geom_text(
    aes(label = sprintf("%s\n%.1f%%", repClass, percent)),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme_bw() +
  labs(
    title = "TE Class composition among H3K9me3 summit-exit TEs per set",
    fill = "TE family"
  )
  ggplot(H3K9me3_te_family_clust, aes(x = cluster, y=percent, fill=repClass))+
   geom_col(width = 0.8, color = "white") +
  # geom_text(
  #   aes(label = sprintf("%s\n%.1f%%", repClass, percent)),
  #   position = position_stack(vjust = 0.5),
  #   size = 3
  # ) +
  theme_bw() +
  labs(
    title = "TE Class composition among TE-overlapping ROIs per set",
    fill = "TE family"
  )
```

#### relative LTR abundance
```{r rel abundance LTR families}
H3K9me3_long_repFamily <-
H3K9me3_te_summary %>% 
   dplyr::filter(cluster != "not_assigned") %>%
  filter(!is.na(repFamily), repFamily != "") %>%   # TE-positive only
  separate_rows(repFamily, sep = ";") %>% 
  distinct(cluster,Peakid, repFamily) %>% 
  dplyr::filter(repFamily %in% c("ERV1", "ERVK","ERVL", "ERVL-MaLR","Gypsy","LTR")) %>%
  count(cluster,repFamily, name = "n") %>% 
  group_by(cluster) %>% 
  mutate(percent = n / sum(n)*100) %>% 
  # mutate(
  #   repFamily = if_else(percent < 2, "Other", repFamily)
  # ) %>% 
  group_by(cluster,repFamily) %>% 
  summarise(n = sum(n), .groups = "drop") %>% 
  group_by(cluster) %>% 
  mutate(percent = n / sum(n) * 100)


# ggplot(H3K9me3_long_repFamily, aes(x = "", y=percent, fill=repFamily))+
#   geom_col(width =1, color="white")+
#   coord_polar(theta = "y")+
#     geom_text_repel(
#     aes(label = sprintf("%s\n%.1f%%", repFamily, percent)),
#     position = position_stack(vjust = 0.5),
#     size = 3,
#     show.legend = FALSE
#   ) +
#   facet_wrap(~cluster)+
#   theme_void()+
#   labs(
#     title = "TE family relative abundance among ROIs by set",
#     fill = "TE family"
#   )


```

```{r rel abundance summit LTR families}
LTR_genome <-
repeatmasker_clean %>% 
  dplyr::filter(repFamily %in% c("ERV1", "ERVK","ERVL", "ERVL-MaLR","Gypsy","LTR"))%>%
  group_by(repFamily) %>% 
  summarise(n = dplyr::n(), .groups = "drop") %>% 
  mutate(percent = n / sum(n) * 100) %>% 
   mutate(cluster = "hg38") 
 

H3K9me3_long_repFamily_ext <-H3K9me3_long_repFamily %>% 
  rbind((LTR_genome %>% mutate(cluster="hg38")))

H3K9me3_long_repFamily_ext
```

```{r LTR barplot}

LTR_bar_df <-H3K9me3_te_summary %>%
  filter(!is.na(repFamily), repFamily != "")  %>%   # TE-positive only
  separate_rows(repFamily, sep = ";") %>% 
  distinct(cluster,Peakid, repFamily) %>% 
  dplyr::filter(repFamily %in% c("ERV1", "ERVK","ERVL", "ERVL-MaLR","Gypsy","LTR")) %>%
  group_by(repFamily, cluster) %>% 
  summarise(n = dplyr::n(), .groups = "drop") %>% 
  # optionally convert NA to a string
  mutate(cluster = if_else(is.na(cluster), "NA", cluster))

ggplot(LTR_bar_df, aes(x = repFamily, y = n, fill = cluster)) +
  geom_col(position = "fill") +
  theme_classic() +
  labs(
    x = "LTR Family",
    y = "Proportion of ROIs",
    fill = "Cluster",
    title = "Proportion of ROI clusters for each LTR family"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### calculating all subfamilies

```{r}
annotated_H3K9me3_ext200 <- readRDS("data/TE_annotation/annotated_H3K9me3_ext200_TE_overlaps.RDS")

repName_counts <- annotated_H3K9me3_ext200 %>%
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_cluster_counts <- tibble(
  cluster = mcols(H3K9me3_summit_gr_ext)$cluster
) %>% count(cluster, name = "total")

not_repName_counts <- repName_counts %>%
  left_join(total_cluster_counts, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

repName_df_long <- bind_rows(repName_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_repName_counts) %>% 
  filter(!is.na(cluster))

repName_totals <- repName_df_long %>%
  filter(status == "TE") %>%
  group_by(repName) %>%
  summarize(total_TE = sum(count), .groups = "drop")

repName_df_long <- repName_df_long %>%
  filter(!is.na(repName)) %>%                      # remove NA repNames
  group_by(repName) %>%
  filter(sum(count[status == "TE"]) >= 10) %>%    # keep only sufficiently abundant TEs
  ungroup()

repName_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      repName_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_repName(
            repName_df_long,
            rep_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))
```


```{r}
ggplot(repName_results, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   geom_text_repel(data = subset(repName_results, FDR < 0.05)) +
  # geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "all repName enrichment at extended ROI summits, H3K9me3"
  ) +
  theme_classic() +
  facet_wrap(~comparison)

annotated_H3K9me3_ext200 %>% 
  dplyr::filter(repName=="MLT1H")
```


```{r making named bedfiles, eval=FALSE}
names(rpt_split_gr_list)

out_dir <- "data/Bed_exports/repeatmasker_sets"
set_list <- names(rpt_split_gr_list)

for (group_name in set_list) {
  gr <- rpt_split_gr_list[[group_name]]
  
  
  # Set BED name column
  mcols(gr)$name <- mcols(gr)$repName
  # 
  # # Optional: if you want a score column
  # if(!"score" %in% colnames(mcols(gr))) {
  #   mcols(gr)$score <- 0
  # }
  # 
  # Export to BED
 bed_file <- file.path(out_dir, paste0(group_name, "_rptmasker.bed"))
  
  # Export
  rtracklayer::export(gr, bed_file, format = "BED")
  
  cat("Exported:", bed_file, "\n")
}

```


```{r making named bedfiles 1, eval=FALSE}
peakAnnoList_H3K9me3 <- readRDS("data/motif_lists/H3K9me3_annotated_peaks.RDS")

out_dir <- "data/Bed_exports/repeatmasker_sets"
set_list <- names(peakAnnoList_H3K9me3)

for (group_name in set_list) {
  cs <- peakAnnoList_H3K9me3[[group_name]]
  gr <- cs@anno
  
  # Set BED name column
  mcols(gr)$name <- mcols(gr)$Peakid
  
  # Optional: if you want a score column
  if(!"score" %in% colnames(mcols(gr))) {
    mcols(gr)$score <- 0
  }
  
  # Export to BED
 bed_file <- file.path(out_dir, paste0(group_name, "_H3K9me3.bed"))
  
  # Export
  export(gr, bed_file, format = "BED")
  
  cat("Exported:", bed_file, "\n")
}

```


