---
title: "MER61_building dataframe"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(ComplexHeatmap)
library(circlize)
library(ggalluvial)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(janitor)
library(MASS)
```

The purpose of this page is to build a large dataframe for MER61E annotations.

```{r annotation data frames}
# ZNF680 narrowpeak binding Hek293 ENCFF935XZI
ZNF680_regions <- import("data/Other_paper_data/ZNF680_CHIP_ENCFF935XZI.bed",format = "narrowPeak")

ZNF780A_regions <- import("data/Other_paper_data/GSM2466657_lifted_hg38_peaks_ZNF780A.bed")
MER61E_regions <- import("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed")

H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")
 H3K27ac_summit_gr_ext <- GenomicRanges::resize(H3K27ac_summit_gr, width(H3K27ac_summit_gr)+400, fix = "center")
# export(H3K27ac_summit_gr_ext,"data/Bed_exports/H3K27ac_summit_gr_ext.bed")
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y))

##assigning Peakid as name of ext summit region
 mcols(H3K27ac_summit_gr_ext)$name <- mcols(H3K27ac_summit_gr_ext)$Peakid
 
 H3K27ac_ext200_anno <- readRDS("data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")
 
 MER61E_anno_genes <- readRDS("data/RDS_files/MER61E_anno_genes.RDS")
 MER61E_anno_chipseeker <- readRDS("data/RDS_files/MER61E_anno_chipseeker.RDS")
 
chipseeker_anno_MER61E <-  MER61E_anno_chipseeker %>% 
  as.data.frame() %>% 
  dplyr::select(name,annotation,geneId,distanceToTSS) %>% 
  dplyr::rename("repID"=name,"CS_MER61E_gene"=geneId,"CS_distanceToTSS"=distanceToTSS)
 MER61E_intermediat_protein <- readRDS("data/MER61intermediatemarking_protein.RDS")
 
 full_p53_canonical  <- readRDS("data/full_p53_canonical.RDS")
  p53halfsite_TE_ids <- readRDS("data/p53halfsite_TE_ids.RDS")

  LTR_df <- readRDS("data/RDS_files/LTR_DF_hg38.RDS")

  H3K27ac_toplist <- readRDS("data/DER_data/H3K27ac_toplist.RDS")

  H3K27ac_toptable_list <- bind_rows(H3K27ac_toplist, .id = "group")


set_case_list_H3K27ac <-  H3K27ac_toptable_list %>% 
  dplyr::select(group,genes, logFC) %>% 
  pivot_wider(.,id_cols = genes, names_from = group, values_from = logFC) %>% 
  dplyr::left_join(H3K27ac_lookup,by =c("genes"="Peakid")) %>% 
  mutate(cluster= case_when (is.na(cluster)~"not_assigned",
                           TRUE ~ cluster),
         case= case_when(H3K27ac_24T > 0 & H3K27ac_24R >0 & H3K27ac_144R>0~"A",
                              H3K27ac_24T < 0 & H3K27ac_24R <0 & H3K27ac_144R<0~"B",
                              TRUE ~ "C"),
         alt_case= case_when(H3K27ac_24T > 0~"D",
                              H3K27ac_24T < 0~"E",
                              TRUE ~ "F")) %>% 
  mutate(set_case=paste0(cluster,"_",case),
         set_alt_case=paste0(cluster,"_",alt_case)) %>% 
  dplyr::select(genes, cluster,case,set_case,alt_case,set_alt_case)

LTR10E_trajectory_broad <- readRDS("data/RDS_files/LTR10E_trajectory_broad.RDS")
# LTR10E_trajectory_detailed <- readRDS("data/RDS_files/LTR10E_trajectory_detailed.RDS")
MER61E_trajectory_broad <- readRDS("data/RDS_files/MER61E_trajectory_broad.RDS")
# MER61E_trajectory_detailed <- readRDS("data/RDS_files/MER61E_trajectory_detailed.RDS")

HALLMARK_DNA_REPAIR <- read_csv("data/Other_paper_data/HALLMARK_DNA_REPAIR.v2026.1.Hs.tsv", 
    skip = 16) %>% 
  pivot_longer(everything()) %>% 
 dplyr::filter(name !="SOURCE_MEMBERS\tAAAS") %>% 
  pull(value)
 
 DNA_repair <- 
 AnnotationDbi::select(org.Hs.eg.db,
  keys = HALLMARK_DNA_REPAIR,   # your UniProt IDs
  keytype = "SYMBOL",
  columns = c("SYMBOL", "ENTREZID"))
 
 DOX_24_ATAC <- readRDS("data/Other_paper_data/Matthews_2025_DOX_24h_ATAC.RDS")
 DOX_24_DAR_gr <- DOX_24_ATAC %>% 
  dplyr::filter(adj.P.Val<0.05) %>% 
  dplyr::select(peak) %>% 
  separate(peak, into = c("seqnames","start","stop"),remove = FALSE) %>% 
  mutate(start=as.numeric(start),
         stop=as.numeric(stop)) %>% 
  as_granges(seqnames = seqnames,
             start=start,
             end=stop,
             strand="*")
DOX_24_CAR_gr <- DOX_24_ATAC %>% 
  dplyr::filter(adj.P.Val>0.05) %>% 
  dplyr::select(peak) %>% 
  separate(peak, into = c("seqnames","start","stop"),remove = FALSE) %>% 
  mutate(start=as.numeric(start),
         stop=as.numeric(stop)) %>% 
  as_granges(seqnames = seqnames,
             start=start,
             end=stop,
             strand="*")
mcols(DOX_24_CAR_gr)$name <- mcols(DOX_24_CAR_gr)$peak
mcols(DOX_24_DAR_gr)$name <- mcols(DOX_24_DAR_gr)$peak

H3K27ac_summit_gr_ext <- keepStandardChromosomes(
  H3K27ac_summit_gr_ext,
  pruning.mode = "coarse"
)
phyloP_df <- read.delim(
  "C:/Users/renee/Other_projects_data/genome_files/H3K27ac_summit_phyloP.tsv",
  na.strings = "",
  stringsAsFactors = FALSE)

phyloP_MER61E_df <- read.delim(
  "C:/Users/renee/Other_projects_data/genome_files/H3K27ac_MER61E_phyloP.tsv",
  na.strings = "",
  stringsAsFactors = FALSE)
### because some (6,691) regions have more than one summit ext region:

phyloP_df$Peakid <- H3K27ac_summit_gr_ext$Peakid
phyloP_df$width <- width(H3K27ac_summit_gr_ext)

phyloP_MER61E_df$repID <- MER61E_regions$name
phyloP_MER61E_df <- phyloP_MER61E_df %>% 
  dplyr::select(repID,phyloP_mean) %>% 
  dplyr::rename("MER61E_phyloP_mean"=phyloP_mean)
peak_phyloP_by_id <- phyloP_df %>%
  group_by(Peakid) %>%
  summarise(
    n_summit_regions = n(),
    total_summit_width = sum(width, na.rm = TRUE),
    peak_phyloP_mean = weighted.mean(phyloP_mean, w = width, na.rm = TRUE),
    peak_phyloP_max = ifelse(all(is.na(phyloP_max)),NA_real_,
      max(phyloP_max, na.rm = TRUE)),  .groups = "drop")
idx <-match(H3K27ac_summit_gr_ext$Peakid,peak_phyloP_by_id$Peakid)


mcols(H3K27ac_summit_gr_ext)$phyloP_mean <- phyloP_df$phyloP_mean
mcols(H3K27ac_summit_gr_ext)$phyloP_max  <- phyloP_df$phyloP_max
mcols(H3K27ac_summit_gr_ext)$peak_phyloP_mean <- peak_phyloP_by_id$peak_phyloP_mean[idx]
mcols(H3K27ac_summit_gr_ext)$peak_phyloP_max <- peak_phyloP_by_id$peak_phyloP_max[idx]
mcols(H3K27ac_summit_gr_ext)$n_summit_regions <- peak_phyloP_by_id$n_summit_regions[idx]


p53_sp_ChIP <- rtracklayer::import("C:/Users/renee/Other_projects_data/DOX_CHIP/macs3_out/MCW_SP_ChIP57_VEH_p53_Ind6/MCW_SP_ChIP57_VEH_p53_Ind6_q0.05_peaks.narrowPeak",format = "narrowPeak")

CTCF_sp_ChIP <-  rtracklayer::import("C:/Users/renee/Other_projects_data/DOX_CHIP/macs3_out/MCW_SP_ChIP63_VEH_CTCF_Ind1/MCW_SP_ChIP63_VEH_CTCF_Ind1_q0.05_peaks.narrowPeak",format = "narrowPeak")

RAD21_sp_ChIP <-  rtracklayer::import("C:/Users/renee/Other_projects_data/DOX_CHIP/macs3_out/MCW_SP_ChIP69_VEH_RAD21_Ind1/MCW_SP_ChIP69_VEH_RAD21_Ind1_q0.05_peaks.narrowPeak",format = "narrowPeak")

TOP2B_sp_ChIP <-  rtracklayer::import("data/Other_paper_data/ChIP3_TOP2B_CM_87-1.bed")

```


canonical assignments
```{r function p53 canonical motifs}
p53_canonical_motifs <- function(fimo_root_dir, te_bed_base_dir) {
 
  
  # Initialize results list
  results_list <- list()
  hit_list <- list()
  
  # Find all directories ending with "_fimo"
  fimo_dirs <- list.dirs(fimo_root_dir, recursive = FALSE)
  fimo_dirs <- fimo_dirs[grepl("_sea_jaspar$", basename(fimo_dirs))]
  
  for (fimo_dir in fimo_dirs) {
    base_name <- sub("_sea_jaspar$", "", basename(fimo_dir))
    
    # Find FIMO file
    sites_file <- file.path(fimo_dir, "sites.tsv")
    if (!file.exists(sites_file)) {
      message("Skipping ", fimo_dir, ": sites.tsv not found")
      next
    }
  
  
  te_file <- file.path(te_bed_base_dir, paste0(base_name, ".bed"))
    if (!file.exists(te_file)) {
      message("Skipping ", base_name, ": TE BED file not found")
      next
    }
  
    
    # ------------------------------
    # Load TE BED file
    # ------------------------------
    te_df <- read.table(te_file, header = FALSE, stringsAsFactors = FALSE)
    colnames(te_df) <- c("chr","start","end","name","score","strand")
    te_df <- te_df[!is.na(te_df$start) & !is.na(te_df$end) & (te_df$start <= te_df$end), ]
    te_gr <- GRanges(
    seqnames = te_df$chr,
    ranges = IRanges(start = te_df$start, end = te_df$end),
    strand = te_df$strand,
    TE_id = te_df$name,
    TE_score = te_df$score
    )
    # ------------------------------
    # Load TP53 sites
    # ------------------------------
    cano_sites <- read.delim(sites_file, stringsAsFactors = FALSE)
    
    # Filter TP53 motifs
    if (!"motif_ALT_ID" %in% colnames(cano_sites)) {
      message("No motif_ALT_ID column in ", sites_file)
      next
    }
    cano_sites <- cano_sites[
      grepl("TP53", cano_sites$motif_ALT_ID, ignore.case = TRUE),
    ]
    # Skip if no motifs found
    if (nrow(cano_sites) == 0) {
      message("No TP53 motifs in ", base_name)
      results_list[[paste0(base_name, "_no_hits")]] <- data.frame(
        TE_name = base_name,
        motif_name = "TP53",
        total_TEs = length(te_gr),
        TEs_with_TP53 = 0,
        total_TP53_hits = 0,
        occupancy = 0
      )
      next
    }
    
    # Clean FIMO entries
    cano_sites <- cano_sites[!is.na(cano_sites$site_Start) & !is.na(cano_sites$site_End) & (cano_sites$site_Start <= cano_sites$site_End), ]
    
    motif_name <- unique(cano_sites$motif_ALT_ID)
    
    # Loop over each motif separately (in case multiple motifs exist)
    for (mn in motif_name) {
      fimo_motif <- cano_sites[cano_sites$motif_ALT_ID == mn, ]
      
      fimo_gr <- GRanges(
        seqnames = fimo_motif$seq_ID,
        ranges = IRanges(start = fimo_motif$site_Start, end = fimo_motif$site_End),
        strand = fimo_motif$site_Strand,
        score = fimo_motif$site_Score
      )
      
      # ------------------------------
      # Overlaps
      # ------------------------------
      hits <- findOverlaps(fimo_gr, te_gr)
      n_te_total <- length(te_gr)
      
      if (length(hits) == 0) {
        message("No overlaps for motif ", mn, " in ", base_name)
        results_list[[paste0(base_name, "_", mn)]] <- data.frame(
          TE_name = base_name,
          motif_name = mn,
          total_TEs = n_te_total,
          TEs_with_TP53 = 0,
          total_TP53_hits = 0,
          occupancy = 0
        )
        next
      }
      
      n_te_with_hits <- length(unique(subjectHits(hits)))
      n_total_hits <- length(hits)
      occupancy <- ifelse(n_te_total > 0, n_te_with_hits / n_te_total, NA)
      
    
    # ------------------------------
      # Compute relative positions
      # ------------------------------
      rel_pos <- (start(fimo_gr)[queryHits(hits)] - start(te_gr)[subjectHits(hits)]) /
                 width(te_gr)[subjectHits(hits)]
      rel_pos <- pmin(pmax(rel_pos, 0), 1)
      df <- data.frame(rel_pos = rel_pos)
      hit_df <- data.frame(
        TE_name = base_name,
       motif_name = mn,

    # which FIMO motif hit
       motif_chr = as.character(seqnames(fimo_gr)[queryHits(hits)]),
        motif_start = start(fimo_gr)[queryHits(hits)],
        motif_end = end(fimo_gr)[queryHits(hits)],
        motif_strand = as.character(strand(fimo_gr)[queryHits(hits)]),
        motif_score = mcols(fimo_gr)$score[queryHits(hits)],

    # which TE it overlaps
        TE_id = mcols(te_gr)$TE_id[subjectHits(hits)],
        TE_chr = as.character(seqnames(te_gr)[subjectHits(hits)]),
        TE_start = start(te_gr)[subjectHits(hits)],
        TE_end = end(te_gr)[subjectHits(hits)],
        TE_strand = as.character(strand(te_gr)[subjectHits(hits)]),

          rel_pos = rel_pos
        )

hit_list[[paste0(base_name, "_", mn)]] <- hit_df
      
      # ------------------------------
      # Store results
      # ------------------------------
      results_list[[paste0(base_name, "_", mn)]] <- data.frame(
        TE_name = base_name,
        motif_name = mn,
        total_TEs = n_te_total,
        TEs_with_TP53 = n_te_with_hits,
        total_TP53_hits = n_total_hits,
        occupancy = occupancy
      )
      
      # ------------------------------
      # Plot density
      # ------------------------------
      cat(paste0("### ", base_name, " ", mn, "\n"))
      cat(base_name, "| total TEs:", n_te_total, "| with cTP53:", n_te_with_hits, 
          "| hits:", n_total_hits, "\n")
      cat("Sequences with full TP53 motif: ", round(occupancy * 100, 2), "%\n\n")
      
      p <- ggplot(df, aes(x = rel_pos)) +
        geom_density(fill = "steelblue", alpha = 0.5) +
        labs(x = "Relative position in TE", y = "Motif density") +
        theme_classic() +
        ggtitle(paste0("canonical TP53 motif distribution: ", base_name,
                       "\nSequences with motif: ", round(occupancy * 100, 1), "%"))
      print(p)
      cat("\n\n")
    }
  } # end fimo_dirs loop
  
  # Combine all results into one data frame
  results_df <- do.call(rbind, results_list)
  
  # print(results_df)
  hits_df <- do.call(rbind, hit_list)

return(list(
  summary = results_df,
  hits = hits_df
))
}
    

```

```{r}
full_p53 <- p53_canonical_motifs("C:/Users/renee/Other_projects_data/DXR_data/TE_sea_jaspar_results","data/Bed_exports/repeatmasker_sets/ERV1")

p53_canonical <- full_p53$hits
```

```{r pulling if TE is in my data}
MER61E_mydata <- 
H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(repName=="MER61E")%>% 
  pull(repID) %>% 
  unique()
protein_info <- MER61E_intermediat_protein %>% 
  dplyr::select(X1,my_MER,is_DAP,DOX_cor)
Peakid_lookup_repID <-
  H3K27ac_ext200_anno %>% 
  separate_longer_delim(.,cols=c(repName,repID),delim=";") %>% 
    dplyr::filter(repName=="MER61E") %>% 
    dplyr::select(repID,Peakid,geneId,distanceToTSS)
```

```{r determining which repIDs have binding}
ZNF780A_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,ZNF780A_regions) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()
ZNF680_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,ZNF680_regions) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

p53_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,p53_sp_ChIP) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

CTCF_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,CTCF_sp_ChIP) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

RAD21_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,RAD21_sp_ChIP) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

TOP2B_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,TOP2B_sp_ChIP) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()
```

```{r overlappig DARs and CARs}

DAR_set <- subsetByOverlaps(MER61E_regions,DOX_24_DAR_gr)
DAR_MER61E_set <- DAR_set %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

CAR_set <- subsetByOverlaps(MER61E_regions,DOX_24_CAR_gr)
CAR_MER61E_set <- CAR_set %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

```

getting phyloP added per region:
```{r}
phyloP_calc_df <- H3K27ac_summit_gr_ext %>% 
  as.data.frame() %>% 
  dplyr::select(Peakid,cluster, name,peak_phyloP_max,peak_phyloP_mean) %>% distinct()
    
```


### MER61E making of data frame  

```{r}
mintermediate_data <-
  MER61E_anno_genes %>% 
  mutate(cluster= case_when(is.na(cluster)~"NA",
                            TRUE~ cluster)) %>% 
  left_join(protein_info,by=c("repID"="X1")) %>% 
  left_join(Peakid_lookup_repID) %>% 
  dplyr::rename("GREAT_gene_assign"=2,
                "H3K27ac_region"=Peakid,
                "H3K27ac_region_geneId"=geneId,
                "H3K27ac_region_distanceToTSS"=distanceToTSS) %>% 
  left_join(LTR_df, by=c("repID"="RM_id")) %>% 
  mutate(my_MER=if_else(repID %in% MER61E_mydata, "yes","no"),
         p53_halfsite =if_else(repID %in% p53halfsite_TE_ids$Peakid, "yes","no"),
         p53_fullsite=if_else(repID %in% full_p53_canonical$TE_id,"yes","no")) %>% 
  left_join(set_case_list_H3K27ac,by=c("H3K27ac_region"="genes", "cluster"="cluster")) %>% 
  left_join(MER61E_trajectory_broad,by=c("repID"="te_id", "repName"="te_name")) %>% 
  mutate(ZNF780A_binding=if_else(repID %in%ZNF780A_overlaps_MER61E,"yes","no"),
         ZNF680_binding=if_else(repID %in% ZNF680_overlaps_MER61E, "yes","no"),
         sp_p53_binding=if_else(repID %in% p53_overlaps_MER61E, "yes","no"),
              sp_CTCF_binding=if_else(repID %in% CTCF_overlaps_MER61E, "yes","no"),
         sp_RAD21_binding=if_else(repID %in% RAD21_overlaps_MER61E, "yes","no"),
         sp_TOP2B_binding=if_else(repID %in% TOP2B_overlaps_MER61E, "yes","no"),) %>% 
  dplyr::mutate(
    state_24T_compare = dplyr::if_else(DOX_24T == VEH_24T, "same", "different"),
    state_24R_compare = dplyr::if_else(DOX_24R == VEH_24R, "same", "different"),
    state_144R_compare = dplyr::if_else(DOX_144R == VEH_144R, "same", "different")) %>% 
  dplyr::mutate(overall_compare = dplyr::case_when(DOX_24T == VEH_24T & 
                                                     DOX_24R == VEH_24R &
                                                     DOX_144R == VEH_144R ~"same_all_timepoints",
                                                   TRUE ~ "different_at_one_or_more_timepoints")) %>% 
  mutate(is_DNA_repair=ENTREZID%in% DNA_repair$ENTREZID,
         overlap_DAR=if_else(repID %in% DAR_MER61E_set,"yes","no"),
         overlap_CAR=if_else(repID %in% CAR_MER61E_set,"yes","no")) %>% 
  left_join(phyloP_calc_df, by=c("H3K27ac_region"="Peakid","cluster"="cluster")) %>% 
  left_join(chipseeker_anno_MER61E) %>% 
  left_join(phyloP_MER61E_df,by="repID")
  

DT::datatable(mintermediate_data,
              options=list(scrollX = TRUE,
                           scrollY="500px",
                           paging=TRUE,
                           autoWidth=TRUE))

# saveRDS(mintermediate_data, "data/RDS_files/MER61E_intermediate_dataframe.RDS")
```

###  Specs about MER61E

```{r chrom number}
mintermediate_data %>% 
  group_by (seqnames, my_MER) %>% 
  tally() %>% 
  ggplot(aes(x=seqnames, y=n, fill=my_MER))+
  geom_col()+
  geom_text(aes(label=n), position = position_stack(vjust = 0.5))+
  ggtitle("MER61E distribution across chromosomes")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=90))+
  labs(fill="overlaps an\nH3K27ac region")+
  scale_fill_brewer(palette = "Paired")
```
```{r}
mintermediate_data %>% 
  group_by (seqnames, p53_halfsite) %>% 
  tally() %>% 
  ggplot(aes(x=seqnames, y=n, fill=p53_halfsite))+
  geom_col()+
  geom_text(aes(label=n), position = position_stack(vjust = 0.5))+
  ggtitle("MER61E distribution across chromosomes and p53halfsite containing")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=90))+
  labs(fill="Has a p53\nhalfsite")
```


```{r}
mintermediate_data %>% 
  dplyr::filter(my_MER=="yes") %>% 
  group_by (seqnames, p53_halfsite) %>% 
  tally() %>% 
  ggplot(aes(x=seqnames, y=n, fill=p53_halfsite))+
  geom_col()+
  geom_text(aes(label=n), position = position_stack(vjust = 0.5))+
  ggtitle("MER61E in my data distribution across chromosomes and p53halfsite containing")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=90))+
  labs(fill="Has a p53\nhalfsite")
```


distance from neargene
```{r}
mintermediate_data %>% 
  mutate(distance_to_closest=abs(closest_distance)) %>% 
  ggplot(., aes(x=my_MER, y=distance_to_closest))+
  geom_boxplot(aes(color=p53_halfsite))+
  # geom_point()
  ggpubr::stat_compare_means(
    aes(group = p53_halfsite),
    method = "wilcox.test",
    label = "p.signif"
  ) +
  theme_bw()+
  ggtitle("nearest gene distance(GREAT) to MER61E")
```
Below I tested if distance to closes assigned gene was different between those that have a p53halfsite and those that don't
```{r}
mer_tests <- mintermediate_data %>%
  mutate(distance_to_closest=abs(closest_distance)) %>% 
  dplyr::filter(
    my_MER %in% c("yes", "no"),
    p53_halfsite %in% c("yes", "no"),
    !is.na(distance_to_closest)
  ) %>%
  dplyr::group_by(p53_halfsite) %>%
  rstatix::wilcox_test(
    distance_to_closest ~ my_MER
  ) %>%
  rstatix::adjust_pvalue(method = "BH") %>%
  rstatix::add_significance("p.adj")

mer_tests
```


```{r}
distance_model <- lm(
  log1p(distance_to_closest) ~ my_MER * p53_halfsite,
  data = (mintermediate_data %>%
  mutate(distance_to_closest=abs(closest_distance)))
)

anova(distance_model)
summary(distance_model)
```
So the conclusion is that the distances to the nearest gene are not associated with being in H3K27ac data (my_MER) or having a p53halfsite or the interaction of p53halfsite and my_MER.

### Age between MER61E elements in my data vs not in my H3K27ac data
```{r}


mintermediate_data %>% 
  dplyr::group_by(my_MER,cluster) %>% 
  summarize("total"=dplyr::n(),avg_milliDiv=mean(milliDiv))

mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=cluster,y=milliDiv))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","not_assigned")
),
    method = "wilcox.test"
  ) +
  labs(title = "MER61E milliDiv dispersion in groups",
       subtitle="Pairwise Wilcoxon rank-sum tests")


```
```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1","Set_2", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=my_MER,y=milliDiv))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no"))
#   # c("Set_1", "not_assigned"),
#   # c("Set_1", "NA"),
#   # c("Set_3","NA"),
#   # c("Set_3","not_assigned")
# ),
#     method = "wilcox.test"
  ) +
  labs(title = "MER61E milliDiv dispersion in vs out of my data",
       subtitle="Pairwise Wilcoxon rank-sum tests")
```

```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=cluster,y=milliDel))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","not_assigned")
),
    method = "wilcox.test"
  ) +
  labs(title = "MER61E milliDel dispersion in groups",
       subtitle="Pairwise Wilcoxon rank-sum tests")


mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=cluster,y=milliIns))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","not_assigned")
),
    method = "wilcox.test"
  ) +
  labs(title = "MER61E milliIns dispersion in groups",
       subtitle="Pairwise Wilcoxon rank-sum tests")
```



```{r}
mintermediate_data %>% 
  # group_by(p53_halfsite) %>% #, p53_fullsite,cluster) %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  # ggplot(aes(x=cluster,y=milliDiv, color=case))+
  ggplot(aes(x=cluster,y=milliDiv))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","not_assigned")
),
    method = "wilcox.test"
  )+
  facet_wrap(~p53_halfsite)+
  ggtitle("MER61E with and without p53 halfsites")

# saveRDS(intermediate_data,"data/H3K27ac_MER61E_integrant_anno.RDS")


mintermediate_data %>% 
  # group_by(p53_halfsite) %>% #, p53_fullsite,cluster) %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  # ggplot(aes(x=cluster,y=milliDiv, color=case))+
  ggplot(aes(x=cluster,y=milliDiv))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  # c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA")
  # c("Set_3","not_assigned")
),
    method = "wilcox.test",
 method.args = list(exact = FALSE))+
  facet_wrap(~p53_fullsite)+
  ggtitle("MER61E with and without p53 full sites")

# saveRDS(intermediate_data,"data/H3K27ac_MER61E_integrant_anno.RDS")
```
### Looking at p53 new data

Look at relationships between TOP2B and MER61E; RAD21 and MER61E, p53 and MER61E ; CTCF and MER61E;

NOTE, no overlaps were observed using TOP2B regions and CTCF regions.  for RAD21- 3 elements overlapped a RAD21 peak, P53: 20 p53 regions overlap a MER61E.


```{r}
mintermediate_data %>% 
  group_by(my_MER, sp_p53_binding, p53_halfsite) %>% 
  count

mintermediate_data %>% 
  group_by(my_MER, sp_RAD21_binding, p53_halfsite) %>% 
  count
```


```{r}
mintermediate_data %>% 
  # group_by(p53_halfsite) %>% #, p53_fullsite,cluster) %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=sp_p53_binding,y=peak_phyloP_mean, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  # facet_wrap(~sp_p53_binding)+
  ggtitle("H3K27ac region overlapping mean phyloP which overlap p53 ChIP data")

```
lets look at the actual MER61E phyloP mean scores across the TE integrant

```{r}
mintermediate_data %>% 
  
  # group_by(p53_halfsite) %>% #, p53_fullsite,cluster) %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=sp_p53_binding,y=MER61E_phyloP_mean, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  facet_wrap(~my_MER)+
  ggtitle("mean MER61E phyloP which overlap p53 ChIP data")
```

Is there a difference between the phyloP calculated across integrants for those overlapping an H3K27ac region vs. not overlapping an H3K27ac region?
```{r}
mintermediate_data %>% 
  # group_by(p53_halfsite) %>% #, p53_fullsite,cluster) %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=my_MER,y=MER61E_phyloP_mean, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  # facet_wrap(~sp_p53_binding)+
  ggtitle("MER61E mean phyloP of integrants in my H3K27ac data")
```
 What about the phyloP mean across integrants both within my H3K27ac data (overlapping a region) vs not in my data and when I stratify whether they have a p53 halfsite or not?
```{r}
mintermediate_data %>% 
  # group_by(p53_halfsite) %>% #, p53_fullsite,cluster) %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=my_MER,y=MER61E_phyloP_mean, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  facet_wrap(~p53_halfsite)+
  ggtitle("MER61E mean phyloP of integrants overlapping p53 halfsites\nnot overlapping and overlapping my H3K27ac data")
```
Yes, for those integrants with a p53 halfsite and in my data, there is a difference in phyloP mean than those integrants with a p53 halfsite but not in my data.


what about the milliDiv?

```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=my_MER,y=milliDiv, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  facet_wrap(~p53_halfsite)+
  ggtitle("MER61E milliDiv of integrants overlapping p53 halfsites\nnot overlapping and overlapping my H3K27ac data")
```
So there are significantly newer (lower milliDiv) MER61E integrant containing a p53 site in my data than not in my data. There is also a significantly higher milliDiv in integrants not containing a p53 site in my data than integrants not in my data (in my data= overlaps an H3K27ac region. not in my data= does not overlap and H3K27ac region)

### looking at length and cluster  
```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=my_MER,y=width))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")
  # c("Set_1", "not_assigned"),
  # c("Set_1", "NA"),
  # c("Set_3","NA"),
  # c("Set_3","not_assigned")
),
    method = "wilcox.test"
  )+
  # facet_wrap(~p53_halfsite)+
  ggtitle("MER61E width of TEs")
```
```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=cluster,y=width))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","not_assigned")
),
    method = "wilcox.test"
  )+
  # facet_wrap(~p53_halfsite)+
  ggtitle("MER61E width of TEs")
```

Width of TE is different between TEs overlapping H3K27ac regions vs those not overlapping H3K27ac regions, but not between clusters.  
```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=cluster,y=width))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  # c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA")  # c("Set_3","not_assigned")
),
    method = "wilcox.test"
  )+
  facet_wrap(~p53_halfsite)+
  ggtitle("MER61E width of TEs by halfsite of p53")

```
This width difference does not exist when considering only those containing a p53 halfsite.





### Investigating DAR/CAR

```{r}
mintermediate_data %>% 
  group_by(my_MER,overlap_DAR) %>% 
  tally()
DAR_MER61E_mat <- matrix(c(44,23,17,183),nrow=2, byrow = TRUE)
fisher.test(DAR_MER61E_mat)
```
  tested if MER61E integrants in my data are enriched in DARs vs MER61E integrants not in my data:
  Found Odds Ratio 20.21 pval 2.2E-16 that  MER61Es tha overlap my H3K27ac regions are enriched in DARs compared to MER61E not overlapping my H3K27ac regions.
  
```{r}
mintermediate_data %>% 
  group_by(my_MER,overlap_CAR) %>% 
  tally()
CAR_MER61E_mat <- matrix(c(16,51,8,192),nrow=2, byrow = TRUE)
fisher.test(CAR_MER61E_mat)


```

tested if MER61E integrants in my data are enriched in CARs vs MER61E integrants not in my data:
  Found Odds Ratio 7.45 pval 7.363e-6   MER61Es that overlap my H3K27ac regions are enriched in CARs compared to MER61E not overlapping my H3K27ac regions.
```{r}
DARvCAR_MER61E_mat <- matrix(c(44,23,16,51),nrow=2, byrow = TRUE)
fisher.test(DARvCAR_MER61E_mat)

```
 
 
 tested if MER61E integrants in my data are enriched in DARs vs CARs:
  Found Odds Ratio 6.01 pval 1.946e-6.   MER61Es that overlap my H3K27ac regions are enriched in DARs vs. CARs from DOX24 ATAC data.
  
```{r}
mintermediate_data %>% 
  group_by(my_MER,cluster,overlap_DAR) %>% 
  tally()

mintermediate_data %>% 
  group_by(my_MER,cluster,overlap_CAR) %>% 
  tally()

Set3_MER61E_DARvCAR_mat <- matrix(c(30,10,6,34),nrow=2,byrow = TRUE)

fisher.test(Set3_MER61E_DARvCAR_mat)
```
Now I am testing if MER61E in Set3 are enriched in DARs v CARs, and yes with an OR of 16.21, p value= 9.508e-8, MER61E in set 3 are enriched in DARs.

```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=overlap_CAR,y=MER61E_phyloP_mean, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  facet_wrap(~p53_halfsite)+
  ggtitle("MER61E mean phyloP of integrants overlapping p53 halfsites\nnot overlapping and overlapping my a CAR")
```
Looking at mean phyloP of integrants that overlap a CAR, with a p53 halfsite and without a p53 halfsite
```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  dplyr::filter(p53_halfsite =="yes") %>% 
  ggplot(aes(x=overlap_CAR,y=MER61E_phyloP_mean, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  facet_wrap(~p53_halfsite)+
  ggtitle("MER61E mean phyloP of integrants overlapping p53 halfsites\nnot overlapping and overlapping my a CAR")
```
```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  dplyr::filter(p53_halfsite =="yes") %>% 
  ggplot(aes(x=overlap_DAR,y=MER61E_phyloP_mean, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  # facet_wrap(~p53_halfsite)+
  ggtitle("of MER61E integrants containing at least one p53 halfsite: What is the phyloP mean for those that also overlap a DAR vs not overlap a DAR")
```


```{r}
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
  dplyr::filter(my_MER=="yes") %>% 
  ggplot(aes(x=p53_halfsite,y=MYA_bourque, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  ylab("Million Years Ago (Bourque method)")+
  xlab("has a p53 halfsite?") +
    # facet_wrap(~p53_halfsite, labeller=as_labeller(p53_label))+
  ggtitle("MER61E MYA (Bourque method) of integrants overlapping my\n H3K27ac data and looking at p53 halfsites")
```



### other TF binding enrichment


  
```{r function plot trajectory}
plot_trajectory <- function(input_df, TE) {
  
  state_colors <- c(
  "Transcription" = "#D73027",
  "Enhancer" = "#FC8D59",
  "Poised" = "#FEE08B",
  "Repressed" = "#4575B4",
  "Quiescent" = "#D9D9D9")
  
  trajectory_cols <- c(
    "VEH_24T",
  "VEH_24R",
  "VEH_144R",
  "DOX_24T",
  "DOX_24R",
  "DOX_144R"
)

trajectory_complete <- input_df %>%
  tidyr::pivot_longer(
    cols = c( VEH_24T, VEH_24R, VEH_144R,DOX_24T, DOX_24R, DOX_144R),
    names_to = c("Condition", "Time"),
    names_sep = "_",
    values_to = "State"
  ) %>%
  tidyr::pivot_wider(
    id_cols = c(te_id, te_name, Condition),
    names_from = Time,
    values_from = State
  ) %>%
  mutate(
    Condition = factor(Condition, levels = c("VEH", "DOX"))
  )

plot_input <-  ggplot(trajectory_complete, aes(axis1 = `24T`,
                                               axis2 = `24R`,
                                               axis3 = `144R`,
                                               y = 1)) +
     geom_alluvium(aes(fill = after_stat(stratum)),
                   alpha = 0.4) +
     geom_stratum(aes(fill = after_stat(stratum)),
                  color = "black") +
     geom_text(stat = "stratum",aes(label = after_stat(stratum)),
               size = 3) +
     facet_wrap(~ Condition) +
     scale_x_discrete(limits = c("24T", "24R", "144R"),
                      expand = c(0.1, 0.05)) +
     scale_fill_manual(values = state_colors,
                       na.value = "grey80") +
     labs(title = paste0("Complete ChromHMM state trajectories for ", TE),
          x = NULL,
          y = paste0("Number of ", TE, " elements"),
          fill = "State") +
     theme_minimal()

 return(plot_input)

}

```

```{r}
plot_trajectory_flow <- function(input_df, TE) {
  
  state_colors <- c(
    "Transcription" = "#D73027",
    "Enhancer" = "#FC8D59",
    "Poised" = "#FEE08B",
    "Repressed" = "#4575B4",
    "Quiescent" = "#D9D9D9"
  )
  
  axis_order <- c(
    "VEH_24T", "VEH_24R", "VEH_144R",
    "DOX_24T", "DOX_24R", "DOX_144R"
  )
  
  # ---- DOX trajectories only ----
  dox_summary <- input_df %>%
    count(DOX_24T, DOX_24R, DOX_144R, name = "n") %>%
    mutate(traj_id = paste0("DOX_", row_number()))
  
  dox_lodes <- dox_summary %>%
    pivot_longer(
      cols = c(DOX_24T, DOX_24R, DOX_144R),
      names_to = "axis",
      values_to = "state"
    ) %>%
    mutate(axis = factor(axis, levels = axis_order))
  
  # ---- VEH trajectories only ----
  veh_summary <- input_df %>%
    count(VEH_24T, VEH_24R, VEH_144R, name = "n") %>%
    mutate(traj_id = paste0("VEH_", row_number()))
  
  veh_lodes <- veh_summary %>%
    pivot_longer(
      cols = c(VEH_24T, VEH_24R, VEH_144R),
      names_to = "axis",
      values_to = "state"
    ) %>%
    mutate(axis = factor(axis, levels = axis_order))
  
  # combine for strata/text layers
  strata_df <- bind_rows(dox_lodes, veh_lodes)
  
  p <- ggplot() +
    # DOX flows: color each segment by the state it comes from
    geom_flow(
      data = dox_lodes,
      aes(
        x = axis,
        stratum = state,
        alluvium = traj_id,
        y = n,
        fill = state
      ),
      stat = "alluvium",
      aes.flow = "forward",
      alpha = 0.8
    ) +
    # VEH flows: color each segment by the state it comes from
    geom_flow(
      data = veh_lodes,
      aes(
        x = axis,
        stratum = state,
        alluvium = traj_id,
        y = n,
        fill = state
      ),
      stat = "alluvium",
      aes.flow = "forward",
      alpha = 0.8
    ) +
    geom_stratum(
      data = strata_df,
      aes(
        x = axis,
        stratum = state,
        alluvium = traj_id,
        y = n,
        fill = state
      ),
      color = "black"
    ) +
    geom_text(
      data = strata_df,
      stat = "stratum",
      aes(
        x = axis,
        stratum = state,
        y = n,
        label = after_stat(stratum)
      ),
      size = 3
    ) +
    scale_x_discrete(
      limits = axis_order,
      labels = c(
        "VEH 24T", "VEH 24R", "VEH 144R",
        "DOX 24T", "DOX 24R", "DOX 144R"
      ),
      expand = c(0.08, 0.05)
    ) +
    scale_fill_manual(
      values = state_colors,
      na.value = "grey80"
    ) +
    labs(
      title = paste0(TE, " regulatory-state trajectories"),
      x = NULL,
      y = paste0("Number of ", TE, " elements"),
      fill = "State"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  
  return(p)
}
```

```{r p53half site traj}
mfiltered_trajectory_test <- MER61E_trajectory_broad %>% 
  dplyr::filter(te_id %in% Peakid_lookup_repID$repID) %>% 
  dplyr::filter(te_id %in% (p53halfsite_TE_ids %>% 
                              dplyr::filter(TE_name=="MER61E_ERV1_LTR") %>%
                              pull(Peakid) %>% 
                              unique()))

plot_trajectory(mfiltered_trajectory_test,"MER61E with p53 halfsite")
plot_trajectory_flow(mfiltered_trajectory_test,"MER61E with p53 halfsite")
```

```{r p53full site traj}
mfilteredfull_trajectory_test <- MER61E_trajectory_broad %>% 
  dplyr::filter(te_id %in% Peakid_lookup_repID$repID) %>% 
  dplyr::filter(te_id %in% (full_p53_canonical %>% 
                              dplyr::filter(TE_name=="MER61E_ERV1_LTR") %>%
                              pull(TE_id) %>% 
                              unique()))

plot_trajectory(mfilteredfull_trajectory_test,"MER61E with p53 fullsite")
plot_trajectory_flow(mfilteredfull_trajectory_test,"MER61E with p53 fullsite")
```


```{r}
plot_trajectory_flow(MER61E_trajectory_broad,"all MER61E elements")
```
```{r}

rep_ID_set3_MER61E <- mintermediate_data %>% 
  dplyr::filter(cluster=="Set_3")

acute_cluster_only_broad <- MER61E_trajectory_broad %>% 
  dplyr::filter(te_id %in% Peakid_lookup_repID$repID) %>% 
  dplyr::filter(te_id %in% rep_ID_set3_MER61E$repID)
plot_trajectory_flow(acute_cluster_only_broad,"Acute_cluster only MER61E elements")
```


```{r 780a plot}

filtered_trajectory_780A <- MER61E_trajectory_broad %>% 
  # dplyr::filter(te_id %in% Peakid_lookup_repID$repID) %>% 
  dplyr::filter(te_id %in% ZNF780A_overlaps_MER61E)

plot_trajectory_flow(filtered_trajectory_780A,"any MER61E with ZNF780A binding")
```
```{r}
filtered_trajectory_680 <- MER61E_trajectory_broad %>% 
  # dplyr::filter(te_id %in% Peakid_lookup_repID$repID) %>% 
  dplyr::filter(te_id %in% ZNF680_overlaps_MER61E)

plot_trajectory_flow(filtered_trajectory_680,"any MER61E with ZNF680 binding")
```







```{r}

MER61Ep53half <- p53halfsite_TE_ids %>% dplyr::filter(TE_name =="MER61E_ERV1_LTR") %>% pull(Peakid) %>% unique()
  MER61Ep53full <- full_p53_canonical %>% dplyr::filter(TE_name =="MER61E_ERV1_LTR") %>% pull(TE_id)%>% unique()
  
  MER61p53all <- union(MER61Ep53half, MER61Ep53full)

filtered_comboP53 <- MER61E_trajectory_broad %>% 
  # dplyr::filter(te_id %in% Peakid_lookup_repID$repID) %>% 
  dplyr::filter(te_id %in% MER61p53all)

plot_trajectory_flow(filtered_comboP53,"All MER61E with p53 binding")
```

```{r}
filtered_comboP53a <- MER61E_trajectory_broad %>% 
  dplyr::filter(te_id %in% Peakid_lookup_repID$repID) %>%
  dplyr::filter(te_id %in% MER61p53all) %>% distinct()

filtered_comboP53b <- MER61E_trajectory_broad %>% 
  dplyr::filter(!te_id %in% Peakid_lookup_repID$repID) %>%
  dplyr::filter(te_id %in% MER61p53all)

plot_trajectory_flow(filtered_comboP53a,"MER61E any p53 binding in my H3K27ac regions")
plot_trajectory_flow(filtered_comboP53b,"MER61E any p53 binding NOT in my H3K27ac regions")


```

```{r}
filtered_comboP53c <- MER61E_trajectory_broad %>% 
  # dplyr::filter(te_id %in% Peakid_lookup_repID$repID) %>%
  dplyr::filter(!te_id %in% MER61p53all)

filtered_comboP53d <- MER61E_trajectory_broad %>% 
  dplyr::filter(te_id %in% Peakid_lookup_repID$repID) %>%
  dplyr::filter(!te_id %in% MER61p53all)

filtered_comboP53e <- MER61E_trajectory_broad %>% 
  dplyr::filter(!te_id %in% Peakid_lookup_repID$repID) %>%
  dplyr::filter(!te_id %in% MER61p53all)
plot_trajectory_flow(filtered_comboP53c,"MER61E no p53 binding ")
plot_trajectory_flow(filtered_comboP53d,"MER61E no p53 binding in my H3K27ac regions")

plot_trajectory_flow(filtered_comboP53e,"MER61E no p53 binding not \nin my H3K27ac regions")
```


```{r}
mintermediate_data %>% 
  group_by(cluster,my_MER,overall_compare) %>%
  tally()
matdata <- matrix(c(61,6, 49, 151), nrow = 2, byrow = FALSE)
matdata
fisher.test(matdata)

```


```{r}
mintermediate_data %>% 
  group_by(my_MER,ZNF780A_binding) %>% 
  tally()

ZNF780A_data <- matrix(c(27,40, 64, 136), nrow = 2, byrow = FALSE)
ZNF780A_data
fisher.test(ZNF780A_data)
```

Integrants in my data not significantly enriched in ZNF780A binding sites (exo-ChIP) than integrants not in my data.

```{r}
mintermediate_data %>% 
  group_by(cluster,my_MER,ZNF780A_binding) %>% 
  tally()

ZNF780A_data_3v1 <- matrix(c(19,21,2, 9), nrow = 2, byrow = FALSE)
ZNF780A_data_3v1
fisher.test(ZNF780A_data_3v1)
```
Integrants in set3 not significantly enriched in ZNF780A binding sites (exo-ChIP) than integrants in set1.



```{r}
mintermediate_data %>% 
  group_by(my_MER,ZNF680_binding) %>% 
  tally()

ZNF680_data <- matrix(c(25,42, 37, 163), nrow = 2, byrow = FALSE)
ZNF680_data
fisher.test(ZNF680_data)
```
Integrants in my data are more likely to be enriched in ZNF680 binding sites (ChIP) than those not in my data.
```{r}
mintermediate_data %>% 
  group_by(cluster,my_MER,ZNF680_binding) %>% 
  tally()

ZNF680_3v1_data <- matrix(c(15,25, 4, 7), nrow = 2, byrow = FALSE)
ZNF680_3v1_data
fisher.test(ZNF680_3v1_data)
```
Integrants in set3 not significantly enriched in ZNF680 binding sites (exo-ChIP) than integrants in set1.


```{r}
mintermediate_data %>% 
  group_by(my_MER,p53_fullsite) %>% 
  tally()

p53full_data <- matrix(c(12,55, 10, 190), nrow = 2, byrow = FALSE)
p53full_data
fisher.test(p53full_data)

```
MER61E integrants in my data are more likely to contain a full p53 site than integrants not in my data.
```{r}
mintermediate_data %>% 
  group_by(my_MER,p53_halfsite) %>% 
  tally()

p53half_data <- matrix(c(63,4, 93, 107), nrow = 2, byrow = FALSE)
p53half_data
fisher.test(p53half_data)

```
MER61E integrants in my data are more likely to contain a half p53 site than integrants not in my data.

```{r}


mintermediate_data %>%
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  ggplot(aes(x=cluster,y=closest_distance, color=case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","not_assigned")),
    method = "wilcox.test")+
  ggtitle("distance to nearest gene?")
  
mintermediate_data %>%
  mutate(my_MER = factor(my_MER,levels= c("yes","no"))) %>%
  ggplot(aes(x=my_MER,y=closest_distance, color=case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(comparisons = list(c("yes", "no")),
                                                method = "wilcox.test")+
  ggtitle("distance to nearest gene not broken up by set")
    


```

```{r}
mintermediate_data %>%
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
    ggplot(aes(x=cluster,y=MER61E_phyloP_mean, color=case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","not_assigned")),
    method = "wilcox.test")+
  ggtitle("mean phylop score of region overlapping MER61E element")
```
```{r}
mintermediate_data %>%
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1","Set_2","Set_3", "not_assigned", "NA"))) %>% 
    ggplot(aes(x=cluster,y=peak_phyloP_mean, color=case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  # c("Set_1", "not_assigned"),
  # c("Set_1", "NA"),
  # c("Set_3","NA"),
  c("Set_3","Set_1")),
    method = "wilcox.test")+
  ggtitle("Mean phylop score of region overlapping MER61E element")
```
```{r}
mintermediate_data %>%
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1","Set_2","Set_3", "not_assigned", "NA"))) %>% 
    ggplot(aes(x=cluster,y=MER61E_phyloP_mean, color=case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  # c("Set_1", "not_assigned"),
  # c("Set_1", "NA"),
  c("Set_3","NA"),
  c("NA","Set_1")),
    method = "wilcox.test")+
  ggtitle("Mean phylop overlapping MER61E element")
```




```{r}
peak_phyloP_by_id %>% 
  left_join(set_case_list_H3K27ac,by=c("Peakid"="genes")) %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1","Set_2","Set_3", "not_assigned", "NA"))) %>% 
    ggplot(aes(x=cluster,y=peak_phyloP_mean))+
  geom_violin()+
  # smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "Set_2"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","Set_2")),
    method = "wilcox.test")+
  ggtitle("mean phylop score of all H3K27ac regions")
```
```{r}
mintermediate_data %>%
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_2","Set_3", "not_assigned", "NA"))) %>% 
  ggplot(.,aes(x=closest_distance,MER61E_phyloP_mean)) +
  geom_point( aes(color=cluster))


```
### looking at other data dimensions

```{r}

plotAnnoBar(MER61E_anno_chipseeker,title = "Feature Distribution of all MER61E integrants")

mintermediate_data  %>% 
  mutate(anno_alt=case_when(str_detect(annotation,"Intron")~"Intron",
         str_detect(annotation, "Exon") ~ "Exon",
      str_detect(annotation, "Promoter") ~ "Promoter",
      str_detect(annotation, "Distal Intergenic") ~ "Distal Intergenic",
      str_detect(annotation, "Downstream") ~ "Downstream",
      TRUE ~ annotation)) %>% 
  # group_by(my_MER,cluster, anno_alt) %>% 
  ggplot(., aes(x=anno_alt,y=peak_phyloP_mean))+
  geom_boxplot()+
  geom_point(aes(color=set_case))
  
  peakAnno_my_MER <- subset(
  MER61E_anno_chipseeker,
  name %in% MER61E_mydata)
  peakAnno_not_my_MER <- subset(
  MER61E_anno_chipseeker,
  !name %in% MER61E_mydata)
  
  
  plotAnnoBar(peakAnno_my_MER,title = "Feature Distribution of all MER61E overlapping an H3K27ac region")
  
  plotAnnoBar(peakAnno_not_my_MER,title = "Feature Distribution of all MER61E NOT overlapping an H3K27ac region")
 
```

```{r}
mintermediate_data %>% 
  dplyr::filter(my_MER=="yes") %>% 
  ggplot(., aes(x=CS_distanceToTSS, y = p53_halfsite,color=cluster))+
  geom_point()


mintermediate_data %>% 
  # dplyr::filter(my_MER=="yes") %>% 
  # dplyr::filter(p53_halfsite=="yes") %>% 
  ggplot(., aes(y=CS_distanceToTSS, x = cluster))+
  geom_boxplot() +
  # geom_point()+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1","not_assigned")),
      method = "wilcox.test")+
  labs(
    title = "Distance to TSS for MER61E elements",
    x = "Cluster",
    y = "Distance to TSS"
  ) +
  theme_classic()
  
```
```{r}
p53_label <- c(yes="p53 overlap",no="no p53 overlap")
mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9)) %>% 
  ggplot(aes(x=my_MER,y=milliDiv, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  facet_wrap(~p53_halfsite, labeller=as_labeller(p53_label))+
  ggtitle("MER61E milliDiv of integrants overlapping p53 halfsites\nnot overlapping and overlapping my H3K27ac data")


mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2E-9) %>% 
  ggplot(aes(x=my_MER,y=MYA_jukes_cantor, color=alt_case))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  ylab("Million Years Ago (Jukes/Cantor method)")+
    facet_wrap(~p53_halfsite, labeller=as_labeller(p53_label))+
  ggtitle("MER61E MYA of integrants overlapping p53 halfsites\nnot overlapping and overlapping my H3K27ac data")

mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
  ggplot(aes(x=my_MER,y=MYA_bourque, color=cluster))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test"
  )+
  ylab("Million Years Ago (Bourque method)")+
    facet_wrap(~p53_halfsite, labeller=as_labeller(p53_label)) +
  ggtitle("MER61E MYA (Bourque method) of integrants overlapping p53 halfsites\nnot overlapping and overlapping my H3K27ac data")


mintermediate_data %>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
  ggplot(aes(x=cluster,y=MYA_bourque, color=my_MER))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1","NA"),
  c("Set_3","NA")),
    method = "wilcox.test"
  )+
  ylab("Million Years Ago (Bourque method)")+
    # facet_wrap(~p53_halfsite, labeller=as_labeller(p53_label)) +
  ggtitle("MER61E MYA (Bourque method) of integrants overlapping p53 halfsites\nnot overlapping and overlapping my H3K27ac data")

```
```{r}
mintermediate_data %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
  group_by(my_MER,p53_halfsite) %>% 
  summarise(milliDiv_avg=(mean(milliDiv)),
            MYA_jukes_cantor_avg=mean(MYA_jukes_cantor),
            MYA_bourque_avg=mean(MYA_bourque)) %>% 
  datatable(.) %>% 
  DT::formatRound(
    columns = c("milliDiv_avg", "MYA_jukes_cantor_avg", "MYA_bourque_avg"),
    digits = 2)
```



```{r}

mintermediate_data %>% 
  group_by(my_MER,cluster, is_DNA_repair) %>% 
  tally()
```


```{r}
mintermediate_data %>% 
mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
  dplyr::filter(my_MER=="yes") %>%
  dplyr::filter(cluster=="Set_3") %>% 
  ggplot(aes(x=p53_fullsite,y=MYA_bourque, color=cluster))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test",
   method.args = list(exact = FALSE)
  )+
  ylab("Million Years Ago (Bourque method)")+
  xlab("has a p53 halfsite?")+ 
    # facet_wrap(~p53_halfsite, labeller=as_labeller(p53_label))+
  ggtitle("MER61E MYA (Bourque method) of integrants overlapping my H3K27ac data \nand looking at p53 halfsites")

```
```{r}
mintermediate_data %>% 
mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
  dplyr::filter(my_MER=="yes") %>%
  dplyr::filter(cluster=="Set_3") %>% 
  ggplot(aes(x=p53_fullsite,y=MYA_bourque, color=cluster))+
  smplot2::sm_boxplot(legends = TRUE)+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("yes", "no")),
    method = "wilcox.test",
   method.args = list(exact = FALSE)
  )+
  ylab("Million Years Ago (Bourque method)")+
  xlab("has a p53 halfsite?")+ 
    # facet_wrap(~p53_halfsite, labeller=as_labeller(p53_label))+
  ggtitle("MER61E MYA (Bourque method) of integrants overlapping my H3K27ac data \nand looking at p53 fullsites")
```



### TF enrichment:

```{r MER61E TF enrichment, eval=FALSE, include=FALSE}
MER61E_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_results_2024/MER61E_ERV1_LTR_xstreme/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

MER61E_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_results_2024/MER61E_ERV1_LTR_xstreme/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

MER61E_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_results_2024/MER61E_ERV1_LTR_xstreme/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


test_bind <- bind_rows(MER61E_sea_disc_out, MER61E_sea_known) %>% 
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID, ENR_RATIO) %>% distinct() %>% 
  dplyr::rename("consensus"="CONSENSUS")

MER61E_set_data <- MER61E_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(test_bind, by=c("ALT_ID"="ALT_ID","ID"="ID")) %>% 
  mutate(diff_per=`TP%` -`FP%`) %>% 
  dplyr::select(CLUSTER,ID,consensus,ALT_ID,EVALUE,SIM_MOTIF:motif_name,TP:diff_per)

novel_MER61E <- MER61E_set_data %>% dplyr::filter(is.na(MOTIF_URL))
found_MER61E <- MER61E_set_data %>% dplyr::filter(!is.na(MOTIF_URL))
```


```{r MER61E TF enrichment2024}
MER61E_sea_only <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_sea_jaspar_results_2024/MER61E_ERV1_LTR_sea_jaspar/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)
```


```{r,fig.height=10, fig.width=8}
MER61E_sea_only %>% 
  dplyr::filter(QVALUE<0.05) %>% 
  arrange(QVALUE) %>% 
  mutate(log10Qvalue= log(QVALUE, base = 10)*(-1),
         facet_group=ceiling(row_number() / 30)) %>% 
  ggplot(., aes (y= reorder(ALT_ID,log10Qvalue))) +
  geom_col(aes(x=log10Qvalue))+
  geom_point(aes(x=ENR_RATIO/3), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~.*3,name="Enrichment Ratio of Motifs"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("MER61E enriched motifs compared to all ERV1"))+
 facet_wrap(~ facet_group, scales = "free_y")+
  scale_y_discrete(drop = TRUE)
  
  
  
```
Not really helpful


```{r, fig.height=6, fig.width=5, eval=FALSE}
found_MER61E %>% 
  dplyr::filter(EVALUE<0.05) %>% 
  group_by(motif_name) %>%
  slice_min(EVALUE, with_ties = FALSE) %>%
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=ENR_RATIO*.25), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./.25,name="Enrichment Ratio of regions"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("MER61E enriched motifs compared to all ERV1"))
```


### DERs

```{r venn up down}
regions_24T <- H3K27ac_toplist$H3K27ac_24T %>%
  dplyr::filter(adj.P.Val<0.05) %>% 
  dplyr::select(genes,logFC,adj.P.Val) %>% 
  left_join(H3K27ac_lookup, by=c("genes"="Peakid")) %>% 
  mutate(direction=case_when(logFC>0~"up",
                             logFC<0~"down",
                             TRUE~"even"))

regions_24R <- H3K27ac_toplist$H3K27ac_24R %>%
  dplyr::filter(adj.P.Val<0.05) %>% 
  dplyr::select(genes,logFC,adj.P.Val) %>%
  left_join(H3K27ac_lookup, by=c("genes"="Peakid")) %>% 
  mutate(direction=case_when(logFC>0~"up",
                             logFC<0~"down",
                             TRUE~"even"))

regions_144R <- H3K27ac_toplist$H3K27ac_144R %>%
  dplyr::filter(adj.P.Val<0.05)%>% 
  dplyr::select(genes,logFC,adj.P.Val) %>%
  left_join(H3K27ac_lookup, by=c("genes"="Peakid")) %>% 
  mutate(direction=case_when(logFC>0~"up",
                             logFC<0~"down",
                             TRUE~"even")) 

ggVennDiagram::ggVennDiagram(list("24T_up"=regions_24T[regions_24T$direction=="up",]$genes,"24R_up"=regions_24R[regions_24R$direction=="up",]$genes,"144R_up"=regions_144R[regions_144R$direction=="up",]$genes))+
  ggtitle("H3K27ac DERs up-regulated by time")+
  scale_fill_viridis_c()


ggVennDiagram::ggVennDiagram(list("24T_down"=regions_24T[regions_24T$direction=="down",]$genes,"24R_down"=regions_24R[regions_24R$direction=="down",]$genes,"144R_down"=regions_144R[regions_144R$direction=="down",]$genes))+
  ggtitle("H3K27ac DERs down-regulated by time")+
  scale_fill_viridis_c()

```

```{r}

mintermediate_data
ggVennDiagram::ggVennDiagram(list("24T_up"=regions_24T[regions_24T$direction=="up",]$genes,"24R_up"=regions_24R[regions_24R$direction=="up",]$genes,"144R_up"=regions_144R[regions_144R$direction=="up",]$genes))+
  ggtitle("H3K27ac DERs up-regulated by time")+
  scale_fill_viridis_c()

```

```{r LFC}
add_on_mintermediate_data <- mintermediate_data %>% 
  dplyr::select(repID,cluster, my_MER, H3K27ac_region,milliDiv:p53_fullsite,overlap_DAR, overlap_CAR, MER61E_phyloP_mean,CS_distanceToTSS) %>% 
  mutate(DER_24T=if_else(H3K27ac_region %in% regions_24T$genes, "DER_24T","not_DER_24T"),
         DER_24R=if_else(H3K27ac_region%in% regions_24R$genes,"DER_24R","not_DER_24R"),
         DER_144R=if_else(H3K27ac_region%in% regions_144R$genes,"DER_144R","not_DER_144R"))


add_on_mintermediate_data %>% 
  group_by(my_MER,p53_halfsite,DER_24T) %>% 
  tally()
add_on_mintermediate_data %>% 
  group_by(my_MER,cluster,DER_24T) %>% 
  tally(name="n") %>% 
  ggplot(aes(x = my_MER, y = n, fill = DER_24T)) +
  geom_col(position = "fill") +
  # facet_wrap(~ cluster) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "MER61E in my data",
    y = "Proportion",
    fill = "DER_24T"
  ) +
  theme_minimal()
```


```{r}
add_on_mintermediate_data %>% 
  group_by(my_MER,cluster,DER_24R) %>% 
  tally(name="n") %>% 
  ggplot(aes(x = my_MER, y = n, fill = DER_24R)) +
  geom_col(position = "fill") +
  # facet_wrap(~ cluster) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "MER61E in my data",
    y = "Proportion",
    fill = "DER_24R"
  ) +
  theme_minimal()
```

```{r}
add_on_mintermediate_data %>% 
  group_by(my_MER,cluster,DER_144R) %>% 
  tally(name="n") %>% 
  ggplot(aes(x = my_MER, y = n, fill = DER_144R)) +
  geom_col(position = "fill") +
  # facet_wrap(~ cluster) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "MER61E in my data",
    y = "Proportion",
    fill = "DER_144R"
  ) +
  theme_minimal()
```

```{r}
add_on_mintermediate_data %>% 
  group_by(my_MER,DER_24T,p53_halfsite, overlap_DAR) %>% 
  tally()

add_on_mintermediate_data %>% 
  group_by(my_MER,DER_24R,p53_halfsite, overlap_DAR) %>% 
  tally()

add_on_mintermediate_data %>% 
  group_by(my_MER,DER_144R,p53_halfsite, overlap_DAR) %>% 
  tally()
```
```{r}
add_on_mintermediate_data %>% 
  group_by(my_MER,DER_24T,cluster, p53_halfsite) %>% 
  tally()
```
```{r}
add_on_mintermediate_data %>% 
  dplyr::filter(my_MER=="yes") %>% 
   left_join(H3K27ac_toplist$H3K27ac_24T,by =c("H3K27ac_region"="genes")) %>% 
   ggplot(.,aes(x=CS_distanceToTSS,y=logFC)) +
  geom_point(aes(color=p53_halfsite))+
  ggtitle("24T logFC vs mean phyloP score")+
  smplot2::sm_statCorr()


add_on_mintermediate_data %>% 
  dplyr::filter(my_MER=="yes") %>% 
   left_join(H3K27ac_toplist$H3K27ac_24T,by =c("H3K27ac_region"="genes")) %>% 
   ggplot(.,aes(x=p53_halfsite,y=logFC)) +
  geom_boxplot(aes(color=p53_halfsite))+
   geom_point()+
  ggtitle("24T logFC across p53 halfsite content")+
  ggpubr::stat_compare_means(comparisons = list(
  c("yes", "no")),
  # c("Set_1", "not_assigned"),
  # c("Set_1", "NA"),
      method = "wilcox.test")
  

add_on_mintermediate_data %>% 
  dplyr::filter(my_MER=="yes") %>% 
   left_join(H3K27ac_toplist$H3K27ac_24T,by =c("H3K27ac_region"="genes")) %>% 
   ggplot(.,aes(x=p53_halfsite,y=MER61E_phyloP_mean)) +
  geom_boxplot()+
  ggtitle("24T mean PhyloP score across my_MER elements with/without p53 halfsite")+
  ggpubr::stat_compare_means(comparisons = list(
  c("yes", "no")),
  # c("Set_1", "not_assigned"),
  # c("Set_1", "NA"),
      method = "wilcox.test")


  
```
```{r}
add_on_mintermediate_data %>% 
  dplyr::filter(my_MER=="yes") %>% 
   left_join(H3K27ac_toplist$H3K27ac_24R,by =c("H3K27ac_region"="genes")) %>% 
   ggplot(.,aes(x=p53_halfsite,y=logFC)) +
  geom_boxplot(aes(color=p53_halfsite))+
  geom_point()+
  ggtitle("24R logFC across p53 halfsite content")+
  ggpubr::stat_compare_means(comparisons = list(
  c("yes", "no")),
  # c("Set_1", "not_assigned"),
  # c("Set_1", "NA"),
      method = "wilcox.test")

add_on_mintermediate_data %>% 
  dplyr::filter(my_MER=="yes") %>% 
   left_join(H3K27ac_toplist$H3K27ac_24R,by =c("H3K27ac_region"="genes")) %>% 
   ggplot(.,aes(x=p53_fullsite,y=logFC)) +
  geom_boxplot(aes(color=p53_fullsite))+
  geom_point()+
  ggtitle("24R logFC across p53 halfsite content")+
  ggpubr::stat_compare_means(comparisons = list(
  c("yes", "no")),
  # c("Set_1", "not_assigned"),
  # c("Set_1", "NA"),
      method = "wilcox.test")
```


#### Creating activated bed table
```{r, eval=FALSE}

Set3_mer61e <- mintermediate_data %>% 
  dplyr::filter(my_MER=="yes" & cluster == "Set_3") %>%
  dplyr::filter(alt_case=="D") %>% 
  distinct(repID)



test <- MER61E_regions[mcols(MER61E_regions)$name %in% Set3_mer61e$repID]

export(test,"data/Bed_exports/MER61E_Set3_active.bed")
```

### Looking at TF enrichment in active TEs

```{r}

H3K27ac_aSet3_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3_results/MER61E_Set3_active_xstreme/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

H3K27ac_aSet3_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3_results/MER61E_Set3_active_xstreme/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

H3K27ac_aSet3_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3_results/MER61E_Set3_active_xstreme/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


test_bind_mer <-   H3K27ac_aSet3_sea_disc_out %>%
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID) %>% 
  distinct()


```

```{r}
 mer_set3_table <- 
H3K27ac_aSet3_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(test_bind_mer) %>% 
  mutate(diff_per=`TP%` -`FP%`)

```


```{r}

 mer_set3_table %>% 
  group_by(motif_name) %>% 
slice_min(EVALUE, with_ties = FALSE) %>%
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*2), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./2,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("Motifs enriched in Set 3 active MER61E vs all others MER61E"))

novel_mer_set3 <- mer_set3_table %>% 
  dplyr::filter(is.na(MOTIF_URL))

known_mer_set3 <- mer_set3_table %>% 
  dplyr::filter(!is.na(MOTIF_URL))
```


```{r}
novel_mer_set3 %>% 
group_by(motif_name) %>% 
slice_min(EVALUE, with_ties = FALSE) %>%
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=ENR_RATIO*50), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./50,name="Enrichment Ratio"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("Novel Motifs enriched in Set 3 active \n MER61E vs all others MER61E n = 13"))


known_mer_set3 %>% 
group_by(motif_name) %>% 
slice_min(EVALUE, with_ties = FALSE) %>%
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=ENR_RATIO*20), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./20,name="Enrichment Ratio"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("Known Motifs enriched in Set 3 active \n MER61E vs all others MER61E n = 26"))
# saveRDS(known_mer_set3, "data/RDS_files/MER61E_set3active_enr.RDS")
```
Filtering for expressed TFs
```{r}
RNA_full_toptable <-  readRDS("data/Other_paper_data/RNA_full_toptable.RDS")

RNA_expressed <- RNA_full_toptable %>% 
  dplyr::select(Entrez_ID, SYMBOL)

### had to add in a special filter to account for the :: in the names of the motifs
symbols <- str_to_upper(RNA_expressed$SYMBOL)

filtered_known_mer_set3 <- known_mer_set3 %>%
  ##create the motif-tokens column so I can keep names where there is a partial match in the RNA Symbols
  mutate(motif_tokens = str_split(str_to_upper(motif_name), "::")) %>%
  ##Because this is a list, now I need to filter using sapply, but then remove the motif token column
  filter(sapply(motif_tokens, function(x) any(x %in% symbols))) %>%
  dplyr::select(-motif_tokens)


filtered_known_mer_set3 %>% 
group_by(motif_name) %>% 
slice_min(EVALUE, with_ties = FALSE) %>%
   mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=ENR_RATIO*20), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./20,name="Enrichment Ratio"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("Expressed TFs Known Motifs enriched in Set 3 active \n MER61E vs all others MER61E n = 16"))

```

### Persistantly Activated MER61E

```{r}

what_mer <-
mintermediate_data %>% 
  mutate(active_DOX= case_when(
    DOX_144R %in% c("Enhancer","Transcription")&
      DOX_24R %in% c("Enhancer","Transcription") &
      DOX_24T %in% c("Enhancer","Transcription")~"Active_DOX",
    DOX_144R %in% c("Repressed" ,  "Quiescent")&
      DOX_24T %in% c("Enhancer","Transcription")~"Transition_DOX",
    TRUE~"neither"),
    active_VEH= case_when(
    VEH_144R %in% c("Enhancer","Transcription")&
      VEH_24R %in% c("Enhancer","Transcription") &
      VEH_24T %in% c("Enhancer","Transcription")~"Active_VEH",
    VEH_144R %in% c("Repressed" ,  "Quiescent")&
      VEH_24T %in% c("Enhancer","Transcription")~"Transition_VEH",
    TRUE~"neither")) %>% 
  mutate(trt_specific= case_when(active_DOX=="Active_DOX" & active_VEH !="Active_VEH"~"active_DOX_only",
         active_DOX=="Active_DOX" & active_VEH =="Active_VEH"~"both_active",
active_DOX=="Transition_DOX" & active_VEH !="Transition_VEH"~"transition_DOX_only",
active_DOX=="Transition_DOX" & active_VEH =="Transition_VEH"~"both_transition",
TRUE ~active_DOX))


what_mer %>% 
  group_by(my_MER,active_DOX, active_VEH ) %>% 
  tally()

what_mer %>% 
  group_by(active_DOX, my_MER, p53_halfsite) %>% 
  tally()

what_mer %>% 
  group_by(my_MER,trt_specific) %>% tally %>% 
  pivot_wider(id_cols=trt_specific, names_from = my_MER, values_from = n) %>% 
  dplyr::rename("H3K27ac overlap"=yes,"no H3K27ac overlap"=no) %>% 
  DT::datatable()
```

```{r}
what_mer%>% 
 
  # dplyr::filter(active_DOX=="Active_DOX") %>% 
   mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=cluster,y=MER61E_phyloP_mean, color=cluster))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","not_assigned")
),
    method = "wilcox.test"
  ) +
    facet_wrap(~active_DOX)+
  labs(title = " Active MER61E mean phyloP in groups",
       subtitle="Pairwise Wilcoxon rank-sum tests")
```



```{r}
what_mer%>% 
 
  #dplyr::filter(active_DOX=="Active_DOX") %>% 
   mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=cluster,y=milliDiv, color=p53_halfsite))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
  c("Set_1", "Set_3"),
  c("Set_1", "not_assigned"),
  c("Set_1", "NA"),
  c("Set_3","NA"),
  c("Set_3","not_assigned")
),
    method = "wilcox.test"
  ) +
  facet_wrap(~active_DOX)+
  labs(title = " Active MER61E milliDiv in groups",
       subtitle="Pairwise Wilcoxon rank-sum tests")
```

Okay, I am going to split out the MER61Es that are persistently active compared to those that transition to repressed state.  Then I will run SEA enrichment instead of xstreme due to only using known motifs (Jaspar2024, core vertebrate non-redundant) between these elements.  I will also only focus on 100+/- bp around the center of the integrant.
Below is how I made the bedfiles

```{r, eval=FALSE}
active_df <- what_mer[what_mer$active_DOX == "Active_DOX", ]
not_active_df <- what_mer[what_mer$active_DOX == "not_active", ]

active_names <- active_df$repID
not_active_names <- not_active_df$repID

gr_names <- mcols(MER61E_regions)$name

missing_active <- setdiff(active_names, gr_names)
missing_not_active <- setdiff(not_active_names, gr_names)

if (length(missing_active) > 0) {
  warning("These Active_DOX regions were not found in MER61E_regions:")
  print(missing_active)
}

if (length(missing_not_active) > 0) {
  warning("These not_active regions were not found in MER61E_regions:")
  print(missing_not_active)
}

active_gr <- MER61E_regions[gr_names %in% active_names]
not_active_gr <- MER61E_regions[gr_names %in% not_active_names]


# Center and resize to 100 bp
active_gr_200bp <- resize(active_gr, width = 200, fix = "center")
not_active_gr_200bp <- resize(not_active_gr, width = 200, fix = "center")

# Keep original MER61E region names as BED column 4
mcols(active_gr_200bp)$name <- mcols(active_gr)$name
mcols(not_active_gr_200bp)$name <- mcols(not_active_gr)$name

# Keep score if available
if ("score" %in% colnames(mcols(active_gr))) {
  mcols(active_gr_200bp)$score <- mcols(active_gr)$score
}

if ("score" %in% colnames(mcols(not_active_gr))) {
  mcols(not_active_gr_200bp)$score <- mcols(not_active_gr)$score
}

# Trim to chromosome bounds if seqinfo has chromosome lengths
active_gr_200bp <- trim(active_gr_200bp)
not_active_gr_200bp <- trim(not_active_gr_200bp)

# Keep only full 100 bp regions
active_gr_200bp <- active_gr_200bp[width(active_gr_200bp) == 200]
not_active_gr_200bp <- not_active_gr_200bp[width(not_active_gr_200bp) == 200]


export(active_gr_200bp, "data/Bed_exports/repeatmasker_sets/ERV1/MER61E_Active_DOX_200bp.bed", format = "BED")
export(not_active_gr_200bp, "data/Bed_exports/repeatmasker_sets/ERV1/MER61E_not_active_DOX_200bp.bed", format = "BED")


H3K27ac_toptable_list %>% 
 dplyr::filter(group=="H3K27ac_24T") %>% 
 left_join(H3K27ac_lookup,by=c("genes"="Peakid")) %>% 
 write_delim(., file="data/lookup_table_andstuff.txt")

```



### treatment focused  


```{r}
what_mer%>% 
  mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=my_MER,y=width))+
  geom_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
   c("yes","no")),
    method = "wilcox.test") +
  facet_wrap(~trt_specific, nrow=1, ncol=5)+
  theme_bw()+
  labs(title = "All MER61E width of integrant",
       subtitle="Pairwise Wilcoxon rank-sum tests")


```


```{r}
what_mer%>% 
    mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=my_MER,y=width, color=cluster))+
  smplot2::sm_boxplot(aes(legend=TRUE))+
  ggpubr::stat_compare_means(
    comparisons = list(
   c("yes","no")
),
    method = "wilcox.test"
  ) +
   facet_wrap(~trt_specific, nrow=1, ncol=5)+
  labs(title = "All MER61E width of integrant",
       subtitle="Pairwise Wilcoxon rank-sum tests")+
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

what_mer%>% 
   # dplyr::filter(cluster=="Set_3") %>% 
  ggplot(aes(x=trt_specific,y=width))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
   c("active_DOX_only","both_active"),
   # c("active_DOX_only","both_transition"),
   c("active_DOX_only","neither"),
   c("active_DOX_only","transition_DOX_only")
   # c("transition_DOX_only","both_transition")
),
    method = "wilcox.test"
  ) +
   facet_wrap(~cluster, nrow=1, ncol=5)+
  labs(title = "Acute-only cluster MER61E width of integrant",
       subtitle="Pairwise Wilcoxon rank-sum tests")+
  theme(axis.text.x = element_text(angle = 90))

what_mer%>% 
   dplyr::filter(cluster=="Set_3") %>% 
  ggplot(aes(x=trt_specific,y=width))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
   c("active_DOX_only","both_active"),
   c("active_DOX_only","both_transition"),
   c("active_DOX_only","neither"),
   c("active_DOX_only","transition_DOX_only"),
   c("transition_DOX_only","both_transition")
),
    method = "wilcox.test"
  ) +
   facet_wrap(~cluster, nrow=1, ncol=5)+
  labs(title = "Acute-only cluster MER61E width of integrant",
       subtitle="Pairwise Wilcoxon rank-sum tests")

```




```{r}
what_mer%>% 
  dplyr::filter(my_MER=="yes") %>%
   mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=trt_specific,y=width, color=p53_halfsite))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
   c("active_DOX_only","both_active"),
   c("active_DOX_only","both_transition"),
   c("active_DOX_only","neither"),
   c("active_DOX_only","transition_DOX_only"),
   c("transition_DOX_only","both_transition")
),
    method = "wilcox.test"
  ) +
  facet_wrap(~cluster)+
  labs(title = "MER61E width of integrant that overlap H3K27ac ROI",
       subtitle="Pairwise Wilcoxon rank-sum tests")
```

```{r}
what_mer%>% 
  dplyr::filter(my_MER=="yes" &
                  active_VEH != "Active_VEH") %>%
   mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=active_DOX,y=width, color=p53_halfsite))+
  smplot2::sm_boxplot()+
  ggpubr::stat_compare_means(
    comparisons = list(
   c("Active_DOX","neither"),
   c("Active_DOX","Transition_DOX")
),
    method = "wilcox.test"
  ) +
  # facet_wrap(~active_DOX)+
  labs(title = "MER61E width of integrant that overlap H3K27ac ROI",
       subtitle="Pairwise Wilcoxon rank-sum tests, removed sets where integrants are not Active in VEH")


what_mer%>% 
  dplyr::filter(my_MER=="yes" &
                  active_VEH != "Active_VEH") %>%
  group_by(active_DOX, active_VEH) %>% tally
```


```{r}
what_mer%>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
    mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=width,y=MYA_jukes_cantor, color=p53_halfsite))+
  geom_point()+
  labs(title = "All MER61E width of integrant")


what_mer%>% 
  dplyr::filter(active_VEH != "Active_VEH" ) %>% 
  dplyr::filter(active_DOX=="Active_DOX"&my_MER=="yes") %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
    mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=width,y=MYA_jukes_cantor, color=p53_halfsite))+
  geom_point()+
  labs(title = "Active_DOX MER61E integrants")


what_mer%>% 
  dplyr::filter(active_VEH != "Transition_VEH" ) %>% 
  dplyr::filter(active_DOX=="Transition_DOX"&my_MER=="yes") %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
    mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=width,y=MYA_jukes_cantor, color=p53_halfsite))+
  geom_point()+
  labs(title = "Transition_DOX MER61E integrants")


what_mer%>% 
  dplyr::filter(active_VEH != "Transition_VEH" ) %>% 
  dplyr::filter(trt_specific%in%c("transition_DOX_only","active_DOX_only","neither")&my_MER=="yes") %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
    mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))
  ) %>% 
  ggplot(aes(x=width,y=MYA_jukes_cantor, color=trt_specific))+
  geom_point()+
  labs(title = "Active and transition MER61E integrants")
```

This leads me to asking if there is a relationship between length and p53 halfsites for all of MER61E integrants.

because there is a higher probability of randomly containing a p53 half site because the TE is longer, i want to control for it. I need to quantify number of half site/width.

```{r}
what_mer%>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
  mutate(age_bin=cut(MYA_jukes_cantor, breaks = 10),
         width_bin= cut(width, breaks=10),
         p53_binary=if_else(p53_halfsite=="yes",1,0)) %>% 
  group_by(age_bin, width_bin) %>% 
  summarise(
    n = dplyr::n(),
    prop_p53 = mean(p53_binary),
    .groups = "drop"
  ) %>%
  filter(n >= 5) %>%
  ggplot(aes(x = age_bin, y = width_bin, fill = prop_p53)) +
  geom_tile(color = "white") +
   geom_text(aes(label = n), size = 3) +
  scale_fill_viridis_c(labels = scales::percent_format()) +
  theme_bw() +
  labs(
    x = "MER61E age bin",
    y = "MER61E width bin",
    fill = "p53 halfsite %",
    title = "p53 halfsite enrichment across MER61E age and width"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
what_mer%>% 
  dplyr::filter(active_VEH != "Active_VEH" ) %>% 
  dplyr::filter(active_DOX=="Active_DOX"&my_MER=="yes") %>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
  mutate(age_bin=cut(MYA_jukes_cantor, breaks = 10),
         width_bin= cut(width, breaks=10),
         p53_binary=if_else(p53_halfsite=="yes",1,0)) %>% 
  group_by(age_bin, width_bin) %>% 
  summarise(
    n = dplyr::n(),
    prop_p53 = mean(p53_binary),
    .groups = "drop"
  ) %>%
  # filter(n >= 5) %>%
  ggplot(aes(x = age_bin, y = width_bin, fill = prop_p53)) +
  geom_tile(color = "white") +
   geom_text(aes(label = n), size = 3) +
  scale_fill_viridis_c(labels = scales::percent_format()) +
  theme_bw() +
  labs(
    x = "MER61E age bin",
    y = "MER61E width bin",
    fill = "p53 halfsite %",
    title = "Active_DOX only MER61E p53 breakdown"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
c
```
Summary so far, there appears to be a p53 relationship with width and MER61E, however that appearance is not seen in my data because the MER61E integrants in my H3K27ac data because all but one has a p53 halfsite.

```{r}


p53_halfsite_count <- p53halfsite_TE_ids %>% 
  dplyr::filter(TE_name=="MER61E_ERV1_LTR") %>% 
  group_by(Peakid) %>% 
  count()  %>% 
  dplyr::rename("p53_hscount"=n)

what_mer_width <- what_mer %>% 
  dplyr::select(repID,width,my_MER,milliDiv:p53_halfsite, MER61E_phyloP_mean:trt_specific,cluster)%>% 
  left_join(p53_halfsite_count, by=c("repID"="Peakid")) %>% 
  mutate(p53_halfsite_count=if_else(is.na(p53_hscount),0,p53_hscount),
         p53_site_per_width=p53_halfsite_count/width)


  what_mer_width%>% 
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
    ggplot(aes(x=p53_site_per_width,y=MYA_jukes_cantor, color=my_MER))+
  geom_point()+
  labs(title = "All MER61E width of integrant")
  
```
```{r}

what_mer_width %>%
  dplyr::filter(my_MER == "yes") %>% # & trt_specific=="active_DOX_only") %>%
  mutate(
    p53_density = p53_halfsite_count / width
  ) %>%
  ggplot(aes(x = width, y = p53_density)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm") +
  ggpubr::stat_cor(
    method = "spearman"
      ) +
  theme_bw() +
  labs(
    x = "MER61E width",
    y = "p53 half-site density",
    title = "p53 half-site density across MER61E width"
  )

what_mer_width %>%
  dplyr::filter(my_MER == "yes") %>%
  ggplot(aes(x = width, y = p53_halfsite_count)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "glm", method.args = list(family = "poisson")) +
  theme_bw() +
  labs(
    x = "MER61E width",
    y = "p53 half-site count",
    title = "Across all MER61E, does p53 half-site count increases with available sequence length?"
  )
```
Wait, does the active_DOX_only MERs (those not including the active_VEH MERs) have more half_sites or whole sites?

```{r}
what_mer_width %>%
  dplyr::filter(my_MER == "yes") %>% 
  # dplyr::filter(active_VEH != "Active_VEH") %>% 
  ggplot(aes(x=trt_specific, y= p53_halfsite_count))+
  geom_violin()+
  theme_bw()+
  ggtitle("Looking at MER61E number of halfsites")


# p53_model <- glm(
#   p53_halfsite_count ~ trt_specific + offset(log(width)),
#   family = poisson(),
#   data = what_mer_width %>%
#     dplyr::filter(my_MER == "yes") %>%
#     dplyr::filter(active_VEH != "Active_VEH")
#   
# )
# 
# summary(p53_model)
```

```{r}
plot_df_all <- what_mer_width %>%
  dplyr::filter(my_MER == "yes") %>% 
  # dplyr::filter(cluster=="Set_3") %>% 
  dplyr::filter(active_VEH != "Active_VEH") %>%
  mutate(
    p53_per_kb = p53_halfsite_count / width * 1000)

n_labs_all <- plot_df_all %>%
  count(trt_specific) %>%
  mutate(label = paste0(trt_specific, "\n(n=", n, ")"))

label_vec_all <- n_labs_all$label
names(label_vec_all) <- n_labs_all$trt_specfic

plot_df_all %>%
  ggplot(aes(x = trt_specific, y = p53_per_kb)) +
  geom_violin(trim = FALSE) +
  geom_jitter(width = 0.15, alpha = 0.5) +
  ggpubr::stat_compare_means(
    comparisons = list(
      c("active_DOX_only", "transition_DOX_only")
    ),
     method = "wilcox.test",
  method.args = list(exact = FALSE))+
  scale_x_discrete(labels = label_vec_all) +
  theme_bw() +
  labs(
    x = "DOX activity state",
    y = "p53 halfsites per kb",
    title = "Width-normalized MER61E p53 halfsite density across DOX states H3K27ac overlapping")
```
```{r}
plot_df <- what_mer_width %>%
  dplyr::filter(my_MER == "yes") %>% 
  dplyr::filter(cluster=="Set_3") %>% 
  dplyr::filter(active_VEH != "Active_VEH") %>%
  mutate(
    p53_per_kb = p53_halfsite_count / width * 1000)

n_labs <- plot_df %>%
  count(trt_specific) %>%
  mutate(label = paste0(trt_specific, "\n(n=", n, ")"))

label_vec <- n_labs$label
names(label_vec) <- n_labs$trt_specfic

plot_df %>%
  ggplot(aes(x = trt_specific, y = p53_per_kb)) +
  geom_violin(trim = FALSE) +
  geom_jitter(width = 0.15, alpha = 0.5) +
  ggpubr::stat_compare_means(
    comparisons = list(
      c("active_DOX_only", "transition_DOX_only")
    ),
     method = "wilcox.test",
  method.args = list(exact = FALSE))+
  scale_x_discrete(labels = label_vec) +
  theme_bw() +
  labs(
    x = "DOX activity state",
    y = "p53 halfsites per kb",
    title = "Width-normalized p53 halfsite density across DOX states Set_3 only")
```


```{r}

test_df <- what_mer_width %>% 
  mutate(p53_site_group = case_when(
    p53_halfsite_count == 0 ~ "0 sites",
    p53_halfsite_count == 1 ~ "1 site",
    p53_halfsite_count == 2 ~ "2 sites",
    p53_halfsite_count == 3 ~ "3 sites",
      p53_halfsite_count >= 4 ~ "4+ sites"),
    p53_site_group = factor(p53_site_group,
                            levels = c("0 sites", "1 site", "2 sites", "3 sites", "4+ sites")))


ggplot(test_df, aes(x = milliDiv, y = width, color = p53_site_group)) +
  geom_point(alpha = 0.7) +
  theme_bw() +
  labs(
    x = "MER61E age / divergence",
    y = "MER61E width",
    color = "p53 halfsite count",
    title = "MER61E width and age by p53 halfsite count"
  )



```
```{r}
ggplot(test_df, aes(x = milliDiv, y = width, color = trt_specific)) +
  geom_point(alpha = 0.7) +
  theme_bw() +
  labs(
    x = "MER61E age / divergence",
    y = "MER61E width",
    color = "ChromHMM type",
    title = "MER61E width and age color by ChromHMM"
  )

```



```{r}
ggplot(test_df, aes(x = milliDiv, y = width, color = p53_site_per_width)) +
  geom_point(alpha = 0.7) +
  scale_color_viridis_c() +
  theme_bw() +
  labs(
    x = "MER61E age / divergence",
    y = "MER61E width",
    color = "p53 sites / width",
    title = "MER61E p53 halfsite density by age and width"
  )
```
```{r}


pois_model <- glm(
  p53_halfsite_count ~ milliDiv + offset(log(width)),
  data = test_df,
  family = poisson()
)

summary(pois_model)
```

Using a Poisson model with TE width included as an offset, MER61E p53-halfsite count was significantly associated with TE divergence/age. Higher milliDiv was associated with lower p53-halfsite density. This could mean that less-diverged MER61E integrants have more p53 halfsites per unit length.

I used a session with ChatGPT to figure out how to test this.


```{r}
what_mer%>%
  # dplyr::filter(my_MER=="yes") %>%
  ggplot(., aes(x=active_DOX,y=milliDiv))+
  geom_boxplot()+
   ggpubr::stat_compare_means(
    comparisons = list(
   c("Active_DOX","Transition_DOX"),
   c("not_active","Active_DOX")),
    method = "wilcox.test" )+
  facet_wrap(~my_MER)

what_mer%>%
  group_by(my_MER, active_DOX) %>% tally

```

```{r}
what_mer%>%
  dplyr::filter(my_MER=="yes") %>%
  ggplot(., aes(x=trt_specific,y=milliDiv))+
  geom_boxplot()+
   ggpubr::stat_compare_means(
    comparisons = list(
   c("active_DOX_only","transition_DOX_only")),
   # c("neither","Active_DOX")),
    method = "wilcox.test" )+
  facet_wrap(~cluster)+
  theme_bw()+
  ggtitle("Looking at milliDiv across ChromHMM lables")+
  theme(axis.text.x=element_text(angle=90))


what_mer%>%
  dplyr::filter(my_MER=="yes") %>%
  ggplot(., aes(x=trt_specific,y=milliDiv))+
  geom_boxplot()+
   ggpubr::stat_compare_means(
    comparisons = list(
   c("active_DOX_only","transition_DOX_only")),
   # c("neither","Active_DOX")),
    method = "wilcox.test" )+
  # facet_wrap(~cluster)+
  theme_bw()+
  ggtitle("Looking at milliDiv across ChromHMM lables")+
  theme(axis.text.x=element_text(angle=90))
  
  
what_mer%>%
  dplyr::filter(my_MER=="yes") %>%
  group_by(trt_specific, cluster) %>% tally

```
### Transisiton only DOX from Acute cluster
 
 I wanna look at TF enrichment of this group.  


```{r, eval=FALSE}
transisition_df <- what_mer[what_mer$trt_specific == "transition_DOX_only" & what_mer$cluster=="Set_3", ]

newactive_df <- what_mer[what_mer$trt_specific == "active_DOX_only" & what_mer$cluster=="Set_3", ]
other_mer_df <- what_mer[what_mer$trt_specific %in% c("both_active","both_transition","neither"), ]

trans_names <- transisition_df$repID
not_trans_names <- other_mer_df$repID
newactive_names <- newactive_df$repID

gr_names <- mcols(MER61E_regions)$name

missing_trans <- setdiff(trans_names, gr_names)
missing_not_trans <- setdiff(not_trans_names, gr_names)

if (length(missing_trans) > 0) {
  warning("These Active_DOX regions were not found in MER61E_regions:")
  print(missing_trans)
}

if (length(missing_not_trans) > 0) {
  warning("These not_active regions were not found in MER61E_regions:")
  print(missing_not_trans)
}

trans_gr <- MER61E_regions[gr_names %in% trans_names]
not_trans_gr <- MER61E_regions[gr_names %in% not_trans_names]

newactive_gr <- MER61E_regions[gr_names %in% newactive_names]
# Center and resize to 100 bp
trans_gr_200bp <- resize(trans_gr, width = 200, fix = "center")
not_trans_gr_200bp <- resize(not_trans_gr, width = 200, fix = "center")
newactive_gr_200bp <- resize(newactive_gr, width=200, fix="center")
# Keep original MER61E region names as BED column 4
mcols(trans_gr_200bp)$name <- mcols(trans_gr)$name
mcols(not_trans_gr_200bp)$name <- mcols(not_trans_gr)$name
mcols(newactive_gr_200bp)$name <- mcols(newactive_gr)$name
# Keep score if available
if ("score" %in% colnames(mcols(trans_gr))) {
  mcols(trans_gr_200bp)$score <- mcols(trans_gr)$score
}

if ("score" %in% colnames(mcols(not_trans_gr))) {
  mcols(not_trans_gr_200bp)$score <- mcols(not_trans_gr)$score
}

# Trim to chromosome bounds if seqinfo has chromosome lengths
trans_gr_200bp <- trim(trans_gr_200bp)
not_trans_gr_200bp <- trim(not_trans_gr_200bp)
newactive_gr_200bp <- trim(newactive_gr_200bp)
# Keep only full 100 bp regions
trans_gr_200bp <- trans_gr_200bp[width(trans_gr_200bp) == 200]
not_trans_gr_200bp <- not_trans_gr_200bp[width(not_trans_gr_200bp) == 200]
newactive_gr_200bp <- newactive_gr_200bp[width(newactive_gr_200bp) == 200]

export(trans_gr_200bp, "data/Bed_exports/repeatmasker_sets/ERV1/MER61E_trans_DOX_200bp.bed", format = "BED")
export(not_trans_gr_200bp, "data/Bed_exports/repeatmasker_sets/ERV1/MER61E_not_trans_DOX_200bp.bed", format = "BED")

export(newactive_gr_200bp, "data/Bed_exports/repeatmasker_sets/ERV1/MER61E_newactive_DOX_200bp.bed", format = "BED")
```

### TF for transition MER61Es

```{r, fig.width=8}
trans_MER61 <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/SEA_MER61E_trans_output/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

trans_MER61 %>% 
  clean_names() %>% 
  mutate(motif_names=toupper(alt_id)) %>% 
  dplyr::select (tp_percent,fp_percent, motif_names,qvalue) %>%
  dplyr::filter(qvalue<0.05) %>% 
  dplyr::rename("transition_MER61E"=tp_percent,"not_transition_MER61E"=fp_percent) %>% 
  pivot_longer(cols=c(transition_MER61E, not_transition_MER61E), names_to = "type", values_to="percent") %>% 
  ggplot(., aes(x=motif_names, y=percent, fill=(type)))+
  geom_col(stat = "identity", position="dodge")+
  # geom_text(
  #   aes(label = signif(qvalue, 2), y = percent + 2),
  #   position = position_dodge(width = 0.9),
  #   size = 3
  # ) +
  theme_bw()+
  ggtitle("Significant TF enrichment in transition MER61Es\nq value < 0.05")+
  theme(strip.text = element_text(face = "bold",  hjust = 0, size = 8),
        strip.background = element_rect(fill = "white", linetype = "solid",
                                        color = "black", linewidth = 1),
        panel.spacing = unit(1, 'points'),
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1))

```
Some interesting finds


```{r fig.height=7}
trans_MER_motifs <- trans_MER61 %>% 
  clean_names() %>% 
  mutate(motif_names=toupper(alt_id)) %>% 
  dplyr::filter(qvalue<0.05) 
 
### had to add in a special filter to account for the :: in the names of the motifs
symbols <- str_to_upper(RNA_expressed$SYMBOL)

filtered_trans_MER <- trans_MER_motifs %>%
  ##create the motif-tokens column so I can keep names where there is a partial match in the RNA Symbols
  mutate(motif_tokens = str_split(str_to_upper(motif_names), "::")) %>%
  ##Because this is a list, now I need to filter using sapply, but then remove the motif token column
  filter(sapply(motif_tokens, function(x) any(x %in% symbols))) %>%
  dplyr::select(-motif_tokens)


filtered_trans_MER %>% 
group_by(motif_names) %>%
slice_min(qvalue, with_ties = FALSE) %>%
   mutate(log10qvalue= log(qvalue, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_names,log10qvalue))) +
  geom_col(aes(x=log10qvalue))+
  geom_point(aes(x=enr_ratio*.07), size =4)+
  scale_x_continuous(expand=expansion(mult=c(0,.125)),sec.axis = sec_axis(transform= ~./.07,name="Enrichment Ratio"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("Expressed TFs Motifs enriched in Set 3 DOX transisition MER61E\nMER61E n = 24"))

# saveRDS(filtered_trans_MER,"data/RDS_files/MER61E_filtered_transition_DOX_only_TF_Enrichment.RDS")

```

```{r}
filtered_trans_MER %>% 
group_by(motif_names) %>%
slice_min(qvalue, with_ties = FALSE) %>%
   mutate(log10qvalue= log(qvalue, base = 10)*(-1)) %>% 
  summarise(log10qvalue)
```

#### P53 information relationships:

```{r}
what_mer %>% 
  dplyr::select(repID,my_MER,cluster, width,milliDiv:p53_fullsite, DOX_144R:VEH_24T, trt_specific) %>% 
  group_by(my_MER,trt_specific,p53_halfsite) %>% 
  summarize(total=dplyr::n(),.groups = "drop") %>% 
  group_by(my_MER,trt_specific) %>% 
  mutate( sum_total = sum(total),
          prop = total / sum_total) %>% 
  ungroup() %>% 
  ggplot(., aes(x=trt_specific, y=prop,fill=p53_halfsite))+
   geom_col(position = "fill") +
  geom_text(
    aes(label =total),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  facet_wrap(~ my_MER) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title ="MER61E p53 halfsite enrichment, in my MER vs not my MER",
    x = "Treatment specificity",
    y = "Proportion",
    fill = "p53 half-site"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
```{r}
what_mer %>% 
  dplyr::select(repID,my_MER,cluster, width,milliDiv:p53_fullsite, DOX_144R:VEH_24T, trt_specific) %>% 
  dplyr::filter(my_MER=="yes") %>% 
  group_by(cluster,trt_specific,p53_halfsite) %>% 
  summarize(total=dplyr::n(),.groups = "drop") %>% 
  group_by(cluster,trt_specific) %>% 
  mutate( sum_total = sum(total),
          prop = total / sum_total) %>% 
  ungroup() %>% 
  ggplot(., aes(x=trt_specific, y=prop,fill=p53_halfsite))+
   geom_col(position = "fill") +
  geom_text(
    aes(label =total),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  facet_wrap(~ cluster) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title="looking at MER61E within clusters p53 halfsites",
    x = "Treatment specificity",
    y = "Proportion",
    fill = "p53 half-site"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
```{r}
what_mer %>% 
  dplyr::select(repID,my_MER,cluster, width,milliDiv:p53_fullsite, DOX_144R:VEH_24T, trt_specific) %>% 
  dplyr::filter(my_MER=="yes") %>% 
  group_by(cluster,trt_specific,p53_fullsite) %>% 
  summarize(total=dplyr::n(),.groups = "drop") %>% 
  group_by(cluster,trt_specific) %>% 
  mutate( sum_total = sum(total),
          prop = total / sum_total) %>% 
  ungroup() %>% 
  ggplot(., aes(x=trt_specific, y=prop,fill=p53_fullsite))+
   geom_col(position = "fill") +
  geom_text(
    aes(label =total),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  facet_wrap(~ cluster) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title="looking at MER61E within clusters p53 fullsites",
    x = "Treatment specificity",
    y = "Proportion",
    fill = "p53 full-site"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


```{r}
halfsite_table <- what_mer %>% 
  dplyr::filter(my_MER=="yes") %>% 
  dplyr::filter(!is.na(cluster), !is.na(p53_halfsite)) %>% 
  count(cluster, p53_halfsite) %>% 
  tidyr::pivot_wider(
    names_from = p53_halfsite,
    values_from = n,
    values_fill = 0
  ) %>% 
  tibble::column_to_rownames("cluster") %>% 
  as.matrix()

halfsite_table

chisq.test(halfsite_table)
```
```{r}
pairwise_halfsite_tests <- what_mer %>% 
  dplyr::filter(my_MER=="yes") %>% 
  dplyr::filter(!is.na(cluster), p53_halfsite %in% c("yes", "no")) %>% 
  group_by(cluster) %>% 
  summarise(
    yes = sum(p53_halfsite == "yes"),
    total = dplyr::n(),
    prop_yes = yes / total,
    .groups = "drop"
  )

pairwise_halfsite_tests
yes_counts <- pairwise_halfsite_tests$yes
total_counts <- pairwise_halfsite_tests$total

names(yes_counts) <- pairwise_halfsite_tests$cluster
names(total_counts) <- pairwise_halfsite_tests$cluster

pairwise.prop.test(
  x = yes_counts,
  n = total_counts,
  p.adjust.method = "BH"
)


```


```{r}
what_mer%>%
  dplyr::filter(my_MER=="yes") %>%
mutate(anno_alt=case_when(str_detect(annotation,"Intron")~"Intron",
         str_detect(annotation, "Exon") ~ "Exon",
      str_detect(annotation, "Promoter") ~ "Promoter",
      str_detect(annotation, "Distal Intergenic") ~ "Distal Intergenic",
      str_detect(annotation, "Downstream") ~ "Downstream",
      TRUE ~ annotation),
      active_DOX = factor(
      active_DOX,
      levels = c("neither", "Transition_DOX", "Active_DOX")),
      trt_specific = factor(trt_specific,
                            levels = c("neither", "both_transition", "both_active", "transition_DOX_only", "active_DOX_only"))) %>% 
  count(anno_alt, trt_specific, name = "n") %>%
  ggplot(aes(x = anno_alt, y = n, fill = trt_specific)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  xlab("Annotation") +
  ylab("Proportion") +
  labs(fill = "DOX activity") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )






what_mer%>%
  dplyr::filter(my_MER=="yes") %>%
  group_by(active_DOX, cluster) %>% tally
```
The plot above is based on the chipseeker annotation.  This annotation set includes long non-coding RNA as genes, unlike the GREAT nearest gene annotation, where only protein coding genes are annotated by distance.


```{r}
what_mer%>%
  dplyr::filter(my_MER=="yes") %>%
mutate(anno_alt=case_when(str_detect(annotation,"Intron")~"Intron",
         str_detect(annotation, "Exon") ~ "Exon",
      str_detect(annotation, "Promoter") ~ "Promoter",
      str_detect(annotation, "Distal Intergenic") ~ "Distal Intergenic",
      str_detect(annotation, "Downstream") ~ "Downstream",
      TRUE ~ annotation),
      active_DOX = factor(
      active_DOX,
      levels = c("neither", "Transition_DOX", "Active_DOX")),
      trt_specific = factor(trt_specific,
                            levels = c("neither", "both_transition", "both_active", "transition_DOX_only", "active_DOX_only"))) %>% 
   count(anno_alt, trt_specific, name = "n") %>%
  ggplot(aes(x = trt_specific, y = n, fill = anno_alt)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  xlab("Annotation") +
  ylab("Proportion") +
  labs(fill = "genomic annotation") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

what_mer%>%
  dplyr::filter(my_MER=="yes") %>%
mutate(anno_alt=case_when(str_detect(annotation,"Intron")~"Intron",
         str_detect(annotation, "Exon") ~ "Exon",
      str_detect(annotation, "Promoter") ~ "Promoter",
      str_detect(annotation, "Distal Intergenic") ~ "Distal Intergenic",
      str_detect(annotation, "Downstream") ~ "Downstream",
      TRUE ~ annotation),
      active_DOX = factor(
      active_DOX,
      levels = c("neither", "Transition_DOX", "Active_DOX")),
      trt_specific = factor(trt_specific,
                            levels = c("neither", "both_transition", "both_active", "transition_DOX_only", "active_DOX_only"))) %>% 
  dplyr::filter(my_MER=="yes") %>%
  group_by( p53_halfsite,anno_alt) %>% 
  tally()

```
```{r}
 what_mer%>%
  dplyr::filter(my_MER=="yes") %>%
mutate(anno_alt=case_when(str_detect(annotation,"Intron")~"Intron",
         str_detect(annotation, "Exon") ~ "Exon",
      str_detect(annotation, "Promoter") ~ "Promoter",
      str_detect(annotation, "Distal Intergenic") ~ "Distal Intergenic",
      str_detect(annotation, "Downstream") ~ "Downstream",
      TRUE ~ annotation),
      active_DOX = factor(
      active_DOX,
      levels = c("neither", "Transition_DOX", "Active_DOX")),
      trt_specific = factor(trt_specific,
                            levels = c("neither", "both_transition", "both_active", "transition_DOX_only", "active_DOX_only"))) %>% 
  count(anno_alt, trt_specific, name = "n") %>%
  ggplot(aes(x = trt_specific, y = n, fill = anno_alt)) +
  geom_col(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  xlab("Annotation") +
  ylab("Proportion") +
  labs(fill = "genomic annotation") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


```
There are two mer61Es in my data that are Transitions and exist in a promotor. These are not the same promotors as the GREAT genes assigned.

```{r}
what_mer%>% 
  dplyr::filter(my_MER=="yes" &
                  active_VEH != "Active_VEH") %>%
  mutate(anno_alt=case_when(str_detect(annotation,"Intron")~"Intron",
         str_detect(annotation, "Exon") ~ "Exon",
      str_detect(annotation, "Promoter") ~ "Promoter",
      str_detect(annotation, "Distal Intergenic") ~ "Distal Intergenic",
      str_detect(annotation, "Downstream") ~ "Downstream",
      TRUE ~ annotation),
      active_DOX = factor(
      active_DOX,
      levels = c("not_active", "Transition_DOX", "Active_DOX")
    )) %>% 
  dplyr::filter(active_DOX=="Active_DOX")

  
```


```{r}
interested_genes <- c("LRRC7","TRIM24","LIN7C","DEPDC1","AKR1D1","BDNF")
# other_interested_genes <- c("DEPDC1","AKR1D1","BDNF")
RNA_lfc_SYMBOL <-RNA_full_toptable %>% 
   dplyr::filter(SYMBOL != "NA")%>%
  dplyr::select(SYMBOL,logFC,timepoint, adj.P.Val) %>% 
  pivot_wider(., names_from = timepoint,
    values_from = c(logFC, adj.P.Val),
    names_glue = "RNA_{.value}_{timepoint}")


# tibble(interested_genes) %>% 
#   mutate(RNA_exp=if_else(interested_genes %in% RNA_expressed$SYMBOL,"exp","not_exp")) %>% 
#   left_join(spec_lfc_SYMBOL, by=c("interested_genes"="SYMBOL")) %>% 
#   dplyr::filter(RNA_exp!="not_exp") %>% 
#   dplyr::select(interested_genes, RNA_LFC_24T:RNA_LFC_144R)




splitting_dox_mer_df <- what_mer%>% 
  dplyr::filter(my_MER=="yes" ) %>%
   separate_rows(GREAT_gene_assign, sep = ",\\s*") %>%
  mutate(
    gene = str_extract(GREAT_gene_assign, "^[A-Za-z0-9_.-]+"),
    distance = str_extract(GREAT_gene_assign,"(?<=\\()[+-]?\\d+(?=\\))") %>% as.numeric(),
    direction = case_when(
      distance < 0 ~ "upstream",
      distance > 0 ~ "downstream",
      distance == 0 ~ "overlapping",
      TRUE ~ NA_character_)) %>% 
  dplyr::select(repID,my_MER,cluster, H3K27ac_region, H3K27ac_region_geneId, H3K27ac_region_distanceToTSS, active_DOX:direction,p53_halfsite) %>% 
  mutate(direction_range=case_when(is.na(direction)~"no_gene",
                                   abs(distance)>100000~">100kb",
                                   abs(distance)<100000 & abs(distance)>50000~"100-50kb",
                                   abs(distance)<50000 & abs(distance)>10000~ "50-10kb",
                                   abs(distance)<10000~"<10kb",
                                       TRUE~"messedup")) %>% 
  mutate(RNA_exp=if_else(gene %in% RNA_expressed$SYMBOL,"exp","not_exp"))
active_DOX_df <- 
splitting_dox_mer_df %>% dplyr::filter(trt_specific=="active_DOX_only") %>% 
  dplyr::filter(RNA_exp=="exp")

transition_DOX_df <- 
splitting_dox_mer_df %>% dplyr::filter(trt_specific=="transition_DOX_only") %>% 
  dplyr::filter(RNA_exp=="exp")

H3K27ac_lfc_table <- H3K27ac_toptable_list %>%
  dplyr::select(genes,logFC,group,adj.P.Val) %>% 
  pivot_wider(., names_from = group,
    values_from = c(logFC, adj.P.Val),
    names_glue = "{group}_{.value}" )
  
```

### LFC RNA and H3K27ac LFC heatmap

```{r}
RNA_H3K27ac_LFC_dataframe <- active_DOX_df %>% 
  left_join(RNA_lfc_SYMBOL, by=c("gene"="SYMBOL")) %>% 
  left_join(H3K27ac_lfc_table, by=c("H3K27ac_region"="genes"))

  
gene_LFC_mat <- 
  RNA_H3K27ac_LFC_dataframe %>% 
    dplyr::select(gene, RNA_logFC_24T,RNA_logFC_24R,RNA_logFC_144R,H3K27ac_24T_logFC,H3K27ac_24R_logFC,H3K27ac_144R_logFC) %>% 
    column_to_rownames("gene") %>% 
    as.matrix()

sig_mat <- RNA_H3K27ac_LFC_dataframe %>% 
  dplyr::select(
    gene,
    RNA_adj.P.Val_24T, RNA_adj.P.Val_24R, RNA_adj.P.Val_144R,
    H3K27ac_24T_adj.P.Val, H3K27ac_24R_adj.P.Val, H3K27ac_144R_adj.P.Val
  ) %>% 
  column_to_rownames("gene") %>% 
  as.matrix()



gene_annotations_ordered <- RNA_H3K27ac_LFC_dataframe %>%
  dplyr::filter(gene %in% rownames(gene_LFC_mat)) %>%
  dplyr::slice(match(rownames(gene_LFC_mat), gene))

  
```

```{r}
row_anno <- rowAnnotation(
  cluster = gene_annotations_ordered$cluster,
  distance = gene_annotations_ordered$direction_range,
  direction = gene_annotations_ordered$direction)

Heatmap(
  gene_LFC_mat,
  column_title = "active_DOX_only",
  name = "LFC",
  right_annotation = row_anno, 
  cluster_rows=FALSE,
  cluster_columns = FALSE,
   cell_fun = function(j, i, x, y, width, height, fill) {
    if (!is.na(sig_mat[i, j]) && sig_mat[i, j] < 0.05) {
      grid::grid.text("*", x, y, gp = grid::gpar(fontsize = 10))}})
```

```{r}
RNA_H3K27ac_LFC_dataframe_trans <- transition_DOX_df %>% 
  left_join(RNA_lfc_SYMBOL, by=c("gene"="SYMBOL")) %>% 
  left_join(H3K27ac_lfc_table, by=c("H3K27ac_region"="genes"))

  
gene_LFC_mat_trans <- 
  RNA_H3K27ac_LFC_dataframe_trans %>% 
  mutate(gene_unique=make.unique(gene)) %>% 
    dplyr::select(gene_unique, RNA_logFC_24T,RNA_logFC_24R,RNA_logFC_144R,H3K27ac_24T_logFC,H3K27ac_24R_logFC,H3K27ac_144R_logFC) %>% 
  
    column_to_rownames("gene_unique") %>% 
    as.matrix()

sig_mat_trans <- RNA_H3K27ac_LFC_dataframe_trans %>% 
  mutate(gene_unique=make.unique(gene)) %>% 
    dplyr::select(gene_unique,  RNA_adj.P.Val_24T, RNA_adj.P.Val_24R, RNA_adj.P.Val_144R,
    H3K27ac_24T_adj.P.Val, H3K27ac_24R_adj.P.Val, H3K27ac_144R_adj.P.Val
  ) %>% 
  column_to_rownames("gene_unique") %>% 
  as.matrix()



gene_annotations_ordered_trans <- RNA_H3K27ac_LFC_dataframe_trans %>%
   mutate(gene_unique=make.unique(gene)) %>% 
  dplyr::filter(gene_unique %in% rownames(gene_LFC_mat_trans)) %>%
  dplyr::slice(match(rownames(gene_LFC_mat_trans), gene_unique))

  
```




```{r, fig.height=12}
row_anno_trans <- rowAnnotation(
  cluster = gene_annotations_ordered_trans$cluster,
  distance = gene_annotations_ordered_trans$direction_range,
  direction = gene_annotations_ordered_trans$direction)

Heatmap(
  gene_LFC_mat_trans,
  column_title = "transition_DOX_only",
  name = "LFC",
  right_annotation = row_anno_trans, 
  cluster_rows=FALSE,
  cluster_columns = FALSE,
   cell_fun = function(j, i, x, y, width, height, fill) {
    if (!is.na(sig_mat_trans[i, j]) && sig_mat_trans[i, j] < 0.05) {
      grid::grid.text("*", x, y, gp = grid::gpar(fontsize = 10))}})
```

```{r}
trans_DOX_alt_dataframe <- transition_DOX_df %>% 
  dplyr::select(repID,cluster,H3K27ac_region:trt_specific,p53_halfsite) %>% 
  mutate(direction_range=case_when(                              abs(H3K27ac_region_distanceToTSS)>100000~">100kb",
                                   abs(H3K27ac_region_distanceToTSS)<100000 & abs(H3K27ac_region_distanceToTSS)>50000~"100-50kb",
                                   abs(H3K27ac_region_distanceToTSS)<50000 & abs(H3K27ac_region_distanceToTSS)>10000~ "50-10kb",
                                   abs(H3K27ac_region_distanceToTSS)<10000~"<10kb",
                                       TRUE~"messedup")) %>% 
  mutate(H3K27ac_region_geneId=as.numeric(H3K27ac_region_geneId)) %>% 
  left_join(RNA_expressed, by=c("H3K27ac_region_geneId"="Entrez_ID")) %>% 
  distinct() %>% 
   left_join(RNA_lfc_SYMBOL, by=c("SYMBOL"="SYMBOL")) %>% 
  left_join(H3K27ac_lfc_table, by=c("H3K27ac_region"="genes")) %>% 
  na.omit(SYMBOL)


trans_gene_LFC_mat <- 
  trans_DOX_alt_dataframe %>% 
  mutate(gene_unique=make.unique(SYMBOL)) %>% 
    dplyr::select(gene_unique, RNA_logFC_24T,RNA_logFC_24R,RNA_logFC_144R,H3K27ac_24T_logFC,H3K27ac_24R_logFC,H3K27ac_144R_logFC) %>% 
  
    column_to_rownames("gene_unique") %>% 
    as.matrix()

trans_sig_mat <- trans_DOX_alt_dataframe %>% 
  mutate(gene_unique=make.unique(SYMBOL)) %>% 
    dplyr::select(gene_unique,  RNA_adj.P.Val_24T, RNA_adj.P.Val_24R, RNA_adj.P.Val_144R,
    H3K27ac_24T_adj.P.Val, H3K27ac_24R_adj.P.Val, H3K27ac_144R_adj.P.Val
  ) %>% 
  column_to_rownames("gene_unique") %>% 
  as.matrix()



trans_gene_annotations_ordered <- trans_DOX_alt_dataframe %>%
   mutate(gene_unique=make.unique(SYMBOL)) %>% 
  dplyr::filter(gene_unique %in% rownames(trans_gene_LFC_mat)) %>%
  dplyr::slice(match(rownames(trans_gene_LFC_mat), gene_unique))

```


```{r, fig.height=6}
row_anno_trans_again <- rowAnnotation(
  cluster = trans_gene_annotations_ordered$cluster,
  distance = trans_gene_annotations_ordered$direction_range,
  p53 = trans_gene_annotations_ordered$p53_halfsite)

Heatmap(
  trans_gene_LFC_mat,
  column_title = "transition_DOX_only",
  name = "LFC",
  right_annotation = row_anno_trans_again, 
  cluster_rows=FALSE,
  cluster_columns = FALSE,
   cell_fun = function(j, i, x, y, width, height, fill) {
    if (!is.na(trans_sig_mat[i, j]) && trans_sig_mat[i, j] < 0.05) {
      grid::grid.text("*", x, y, gp = grid::gpar(fontsize = 10))}})
```
### Persistantly activated TEs TF enrichment

```{r , fig.width=8}
PA_MER61 <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/SEA_MER61E_active_output/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

PA_MER61 %>% 
  clean_names() %>% 
  mutate(motif_names=toupper(alt_id)) %>% 
  dplyr::select (tp_percent,fp_percent, motif_names,qvalue) %>% 
  dplyr::rename("active_MER61E"=tp_percent,"not_persistantly_active_MER61E"=fp_percent) %>% 
  pivot_longer(cols=c(active_MER61E, not_persistantly_active_MER61E), names_to = "type", values_to="percent") %>% 
  ggplot(., aes(x=motif_names, y=percent, fill=(type)))+
  geom_col(stat = "identity", position="dodge")+
  # geom_text(
  #   aes(label = signif(qvalue, 2), y = percent + 2),
  #   position = position_dodge(width = 0.9),
  #   size = 3
  # ) +
  theme_bw()+
  ggtitle("TF enrichment in persistantly activated TEs")+
  theme(strip.text = element_text(face = "bold",  hjust = 0, size = 8),
        strip.background = element_rect(fill = "white", linetype = "solid",
                                        color = "black", linewidth = 1),
        panel.spacing = unit(1, 'points'),
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1))
```


```{r , fig.width=8}

PA_MER61 %>% 
  clean_names() %>% 
  mutate(motif_names=toupper(alt_id)) %>% 
  dplyr::filter(motif_names %in% RNA_expressed$SYMBOL) %>% 
  dplyr::select (tp_percent,fp_percent, motif_names,qvalue) %>% 
  dplyr::rename("active_MER61E"=tp_percent,"not_persistantly_active_MER61E"=fp_percent) %>% 
  pivot_longer(cols=c(active_MER61E, not_persistantly_active_MER61E), names_to = "type", values_to="percent") %>% 
  ggplot(., aes(x=motif_names, y=percent, fill=(type)))+
  geom_col(stat = "identity", position="dodge")+
  # geom_text(
  #   aes(label = signif(qvalue, 2), y = percent + 2),
  #   position = position_dodge(width = 0.9),
  #   size = 3
  # ) +
  theme_bw()+
  ggtitle("expressed TF enrichment in persistantly activated TEs")+
  theme(strip.text = element_text(face = "bold",  hjust = 0, size = 8),
        strip.background = element_rect(fill = "white", linetype = "solid",
                                        color = "black", linewidth = 1),
        panel.spacing = unit(1, 'points'),
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1))
```
```{r,eval=FALSE}
pa_tf_names <- PA_MER61 %>% 
  clean_names() %>% 
  mutate(motif_names=toupper(alt_id)) %>%
  pull(motif_names)

saveRDS(pa_tf_names,"data/RDS_files/pa_tf_names.RDS")
```

### TAD info

```{r}
Left_ventricle_TAD <- import(
  con = "C://Users/renee/Downloads/hg38.TADs/hg38/VentricleLeft_STL003_Leung_2015-raw_TADs.txt",
  format = "bed",
  genome = "hg38"
)

mcols(Left_ventricle_TAD)$TAD_id <- paste0("TAD_", seq_along(Left_ventricle_TAD))

SCREEN_heart_hg38_cCREs <- import(
  "data/Other_paper_data/SCREEN_LVHEART_F_53_ENCFF417JSF_ENCFF651XRK_ENCFF320IPT_ENCFF440RUS.bed",
  format = "BED",
  extraCols = c(
    DHS_ID = "character",
    cCRE_ID = "character",
    cCRE_class = "character"))

SCREEN_hg38_cCREs <- import(
  "data/Bed_exports/cisRE/GRCh38-cCREs.bed",
  format = "BED",
  extraCols = c(
    DHS_ID = "character",
    cCRE_ID = "character",
    cCRE_class = "character"))
```

```{r}
MER61E_tad <- mergeByOverlaps(MER61E_regions,Left_ventricle_TAD)

MER61E_tad %>%as.data.frame() %>% 
  group_by(TAD_id) %>%tally() 
  
  
```
```{r}
MER61E_cre <- mergeByOverlaps(MER61E_regions,SCREEN_heart_hg38_cCREs)

MER61E_cre %>% 
  as.data.frame() %>% 
  group_by(MER61E_regions.name) %>% 
  dplyr::filter(MER61E_regions.name %in%MER61E_mydata) %>% 
  left_join((SCREEN_hg38_cCREs %>% as.data.frame()), by=c("SCREEN_heart_hg38_cCREs.name" = "cCRE_ID")) %>% 
  dplyr::select(name,name.1,cCRE_class.y) %>% 
  group_by(name,cCRE_class.y) %>% tally() %>% 
  left_join(mintermediate_data,by=c("name"="repID"))

```

