---
title: "Additional Analysis"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(ComplexHeatmap)
library(circlize)
library(ggalluvial)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(janitor)
library(MASS)
library(readxl)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(GenomicFeatures)
library(edgeR)
```

The purpose of this page is to build a large dataframe for MER61E annotations.

```{r annotation data frames}
# ZNF680 narrowpeak binding Hek293 ENCFF935XZI
ZNF680_regions <- import("data/Other_paper_data/ZNF680_CHIP_ENCFF935XZI.bed",format = "narrowPeak")

ZNF780A_regions <- import("data/Other_paper_data/GSM2466657_lifted_hg38_peaks_ZNF780A.bed")
MER61E_regions <- import("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed")

LTR10E_regions <- import("data/Bed_exports/repeatmasker_sets/ERV1/LTR10E_ERV1_LTR.bed")


MER61E_p53_best_site_score <- import("C:/Users/renee/Other_projects_data/DXR_data/fimo_results/MER61E_ERV1_LTR_fimo/best_site.narrowPeak")

LTR10E_p53_best_site_score <- import("C:/Users/renee/Other_projects_data/DXR_data/fimo_results/LTR10E_ERV1_LTR_fimo/best_site.narrowPeak")

H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")

H3K27ac_summit_gr_ext <- GenomicRanges::resize(H3K27ac_summit_gr, width(H3K27ac_summit_gr)+400, fix = "center")
# export(H3K27ac_summit_gr_ext,"data/Bed_exports/H3K27ac_summit_gr_ext.bed")
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

peakAnnoList_H3K9me3 <- readRDS("data/motif_lists/H3K9me3_annotated_peaks.RDS")

H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y))

##assigning Peakid as name of ext summit region
 mcols(H3K27ac_summit_gr_ext)$name <- mcols(H3K27ac_summit_gr_ext)$Peakid
 
 H3K27ac_ext200_anno <- readRDS("data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")
 
 MER61E_anno_genes <- readRDS("data/RDS_files/MER61E_anno_genes.RDS")
 MER61E_anno_chipseeker <- readRDS("data/RDS_files/MER61E_anno_chipseeker.RDS")
 
 LTR10E_anno_genes <- readRDS("data/RDS_files/LTR10E_anno_genes.RDS")
 LTR10E_anno_chipseeker <- readRDS("data/RDS_files/LTR10E_anno_chipseeker.RDS")
 
chipseeker_anno_MER61E <-  MER61E_anno_chipseeker %>% 
  as.data.frame() %>% 
  dplyr::select(name,annotation,geneId,distanceToTSS) %>% 
  dplyr::rename("repID"=name,"CS_MER61E_gene"=geneId,"CS_distanceToTSS"=distanceToTSS)

chipseeker_anno_LTR10E <-  LTR10E_anno_chipseeker %>% 
  as.data.frame() %>% 
  dplyr::select(name,annotation,geneId,distanceToTSS) %>% 
  dplyr::rename("repID"=name,"CS_MER61E_gene"=geneId,"CS_distanceToTSS"=distanceToTSS)

 # MER61E_intermediat_protein <- readRDS("data/MER61intermediatemarking_protein.RDS")
 
full_p53_canonical  <- readRDS("data/full_p53_canonical.RDS")

p53halfsite_TE_ids <- readRDS("data/p53halfsite_TE_ids.RDS")

# LTR_df <- readRDS("data/RDS_files/LTR_DF_hg38.RDS")

H3K27ac_toplist <- readRDS("data/DER_data/H3K27ac_toplist.RDS")

H3K27ac_toptable_list <- bind_rows(H3K27ac_toplist, .id = "group")

set_case_list_H3K27ac <-  H3K27ac_toptable_list %>% 
  dplyr::select(group,genes, logFC) %>% 
  pivot_wider(.,id_cols = genes, names_from = group, values_from = logFC) %>% 
  dplyr::left_join(H3K27ac_lookup,by =c("genes"="Peakid")) %>% 
  mutate(cluster= case_when (is.na(cluster)~"not_assigned",
                           TRUE ~ cluster),
         case= case_when(H3K27ac_24T > 0 & H3K27ac_24R >0 & H3K27ac_144R>0~"A",
                              H3K27ac_24T < 0 & H3K27ac_24R <0 & H3K27ac_144R<0~"B",
                              TRUE ~ "C"),
         alt_case= case_when(H3K27ac_24T > 0~"D",
                              H3K27ac_24T < 0~"E",
                              TRUE ~ "F")) %>% 
  mutate(set_case=paste0(cluster,"_",case),
         set_alt_case=paste0(cluster,"_",alt_case)) %>% 
  dplyr::select(genes, cluster,case,set_case,alt_case,set_alt_case)

LTR10E_trajectory_broad <- readRDS("data/RDS_files/LTR10E_trajectory_broad.RDS")

MER61E_trajectory_broad <- readRDS("data/RDS_files/MER61E_trajectory_broad.RDS")



DOX_24_ATAC <- readRDS("data/Other_paper_data/Matthews_2025_DOX_24h_ATAC.RDS")

DOX_24_DAR_gr <- DOX_24_ATAC %>% 
  dplyr::filter(adj.P.Val<0.05) %>% 
  dplyr::select(peak) %>%
  separate(peak, into = c("seqnames","start","stop"),remove = FALSE) %>% 
  mutate(start=as.numeric(start),
         stop=as.numeric(stop)) %>% 
  as_granges(seqnames = seqnames,
             start=start,
             end=stop,
             strand="*")


upDOX_24_DAR_gr <- DOX_24_ATAC %>% 
  dplyr::filter(adj.P.Val<0.05) %>% 
  dplyr::filter(logFC>0) %>% 
  dplyr::select(peak) %>%
  separate(peak, into = c("seqnames","start","stop"),remove = FALSE) %>% 
  mutate(start=as.numeric(start),
         stop=as.numeric(stop)) %>% 
  as_granges(seqnames = seqnames,
             start=start,
             end=stop,
             strand="*")

DOX_24_CAR_gr <- DOX_24_ATAC %>% 
  dplyr::filter(adj.P.Val>0.05) %>% 
  dplyr::select(peak) %>% 
  separate(peak, into = c("seqnames","start","stop"),remove = FALSE) %>% 
  mutate(start=as.numeric(start),
         stop=as.numeric(stop)) %>% 
  as_granges(seqnames = seqnames,
             start=start,
             end=stop,
             strand="*")

mcols(DOX_24_CAR_gr)$name <- mcols(DOX_24_CAR_gr)$peak

mcols(DOX_24_DAR_gr)$name <- mcols(DOX_24_DAR_gr)$peak

H3K27ac_summit_gr_ext <- keepStandardChromosomes(
  H3K27ac_summit_gr_ext,
  pruning.mode = "coarse")

p53_sp_ChIP <- rtracklayer::import("C:/Users/renee/Other_projects_data/DOX_CHIP/macs3_out/MCW_SP_ChIP57_VEH_p53_Ind6/MCW_SP_ChIP57_VEH_p53_Ind6_q0.05_peaks.narrowPeak",format = "narrowPeak")

CTCF_sp_ChIP <- rtracklayer::import("C:/Users/renee/Other_projects_data/DOX_CHIP/macs3_out/MCW_SP_ChIP63_VEH_CTCF_Ind1/MCW_SP_ChIP63_VEH_CTCF_Ind1_q0.05_peaks.narrowPeak",format = "narrowPeak")

RAD21_sp_ChIP <-  rtracklayer::import("C:/Users/renee/Other_projects_data/DOX_CHIP/macs3_out/MCW_SP_ChIP69_VEH_RAD21_Ind1/MCW_SP_ChIP69_VEH_RAD21_Ind1_q0.05_peaks.narrowPeak",format = "narrowPeak")

TOP2B_sp_ChIP <-  rtracklayer::import("data/Other_paper_data/ChIP3_TOP2B_CM_87-1.bed")

ATAC_regions_Ind2 <- rtracklayer::import("C:/Users/renee/Other_projects_data/DXR_data/ATAC_peak_files/trimmed_Ind2_87V24h_S12_peaks.narrowPeak",format = "narrowPeak")

mint_anno <- readRDS("data/RDS_files/MER61E_intermediate_dataframe.RDS")
lint_anno <- readRDS("data/RDS_files/LTR10E_intermediate_dataframe.RDS")


K27ac_VEH_Ind4 <- rtracklayer::import("data/peaks/H3K27ac/MCW_CT_ROA_106_FINAL_peaks.narrowPeak",format = "narrowPeak")


MER61E_site <- mergeByOverlaps(MER61E_regions,MER61E_p53_best_site_score)

MER61E_best_site_signal <- MER61E_site %>% 
  as.data.frame() %>% 
  dplyr::select(MER61E_regions.name,MER61E_p53_best_site_score.signalValue,MER61E_p53_best_site_score.qValue) %>% 
  dplyr::rename("repID"=1,"MER61E.signal"=2,"MER61E.qvalue"=3)


LTR10E_site <- mergeByOverlaps(LTR10E_regions,LTR10E_p53_best_site_score)

LTR10E_best_site_signal <- LTR10E_site %>% 
  as.data.frame() %>% 
  dplyr::select(LTR10E_regions.name,LTR10E_p53_best_site_score.signalValue,LTR10E_p53_best_site_score.qValue) %>% 
  dplyr::rename("repID"=1,"LTR10E.signal"=2,"LTR10E.qvalue"=3)

SOX11_hg38_gr <- readRDS("data/Other_paper_data/SOX11_hg38_NIHMS686659_gr.RDS")
```

```{r}
sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")

sampleinfo %>% 
  dplyr::filter(Histone_Mark=="H3K27ac" & Treatment=="VEH" & Timepoint=="24T")
```

canonical assignments
```{r function p53 canonical motifs, eval=FALSE}
p53_canonical_motifs <- function(fimo_root_dir, te_bed_base_dir) {
 
  
  # Initialize results list
  results_list <- list()
  hit_list <- list()
  
  # Find all directories ending with "_fimo"
  fimo_dirs <- list.dirs(fimo_root_dir, recursive = FALSE)
  fimo_dirs <- fimo_dirs[grepl("_sea_jaspar$", basename(fimo_dirs))]
  
  for (fimo_dir in fimo_dirs) {
    base_name <- sub("_sea_jaspar$", "", basename(fimo_dir))
    
    # Find FIMO file
    sites_file <- file.path(fimo_dir, "sites.tsv")
    if (!file.exists(sites_file)) {
      message("Skipping ", fimo_dir, ": sites.tsv not found")
      next
    }
  
  
  te_file <- file.path(te_bed_base_dir, paste0(base_name, ".bed"))
    if (!file.exists(te_file)) {
      message("Skipping ", base_name, ": TE BED file not found")
      next
    }
  
    
    # ------------------------------
    # Load TE BED file
    # ------------------------------
    te_df <- read.table(te_file, header = FALSE, stringsAsFactors = FALSE)
    colnames(te_df) <- c("chr","start","end","name","score","strand")
    te_df <- te_df[!is.na(te_df$start) & !is.na(te_df$end) & (te_df$start <= te_df$end), ]
    te_gr <- GRanges(
    seqnames = te_df$chr,
    ranges = IRanges(start = te_df$start, end = te_df$end),
    strand = te_df$strand,
    TE_id = te_df$name,
    TE_score = te_df$score
    )
    # ------------------------------
    # Load TP53 sites
    # ------------------------------
    cano_sites <- read.delim(sites_file, stringsAsFactors = FALSE)
    
    # Filter TP53 motifs
    if (!"motif_ALT_ID" %in% colnames(cano_sites)) {
      message("No motif_ALT_ID column in ", sites_file)
      next
    }
    cano_sites <- cano_sites[
      grepl("TP53", cano_sites$motif_ALT_ID, ignore.case = TRUE),
    ]
    # Skip if no motifs found
    if (nrow(cano_sites) == 0) {
      message("No TP53 motifs in ", base_name)
      results_list[[paste0(base_name, "_no_hits")]] <- data.frame(
        TE_name = base_name,
        motif_name = "TP53",
        total_TEs = length(te_gr),
        TEs_with_TP53 = 0,
        total_TP53_hits = 0,
        occupancy = 0
      )
      next
    }
    
    # Clean FIMO entries
    cano_sites <- cano_sites[!is.na(cano_sites$site_Start) & !is.na(cano_sites$site_End) & (cano_sites$site_Start <= cano_sites$site_End), ]
    
    motif_name <- unique(cano_sites$motif_ALT_ID)
    
    # Loop over each motif separately (in case multiple motifs exist)
    for (mn in motif_name) {
      fimo_motif <- cano_sites[cano_sites$motif_ALT_ID == mn, ]
      
      fimo_gr <- GRanges(
        seqnames = fimo_motif$seq_ID,
        ranges = IRanges(start = fimo_motif$site_Start, end = fimo_motif$site_End),
        strand = fimo_motif$site_Strand,
        score = fimo_motif$site_Score
      )
      
      # ------------------------------
      # Overlaps
      # ------------------------------
      hits <- findOverlaps(fimo_gr, te_gr)
      n_te_total <- length(te_gr)
      
      if (length(hits) == 0) {
        message("No overlaps for motif ", mn, " in ", base_name)
        results_list[[paste0(base_name, "_", mn)]] <- data.frame(
          TE_name = base_name,
          motif_name = mn,
          total_TEs = n_te_total,
          TEs_with_TP53 = 0,
          total_TP53_hits = 0,
          occupancy = 0
        )
        next
      }
      
      n_te_with_hits <- length(unique(subjectHits(hits)))
      n_total_hits <- length(hits)
      occupancy <- ifelse(n_te_total > 0, n_te_with_hits / n_te_total, NA)
      
    
    # ------------------------------
      # Compute relative positions
      # ------------------------------
      rel_pos <- (start(fimo_gr)[queryHits(hits)] - start(te_gr)[subjectHits(hits)]) /
                 width(te_gr)[subjectHits(hits)]
      rel_pos <- pmin(pmax(rel_pos, 0), 1)
      df <- data.frame(rel_pos = rel_pos)
      hit_df <- data.frame(
        TE_name = base_name,
       motif_name = mn,

    # which FIMO motif hit
       motif_chr = as.character(seqnames(fimo_gr)[queryHits(hits)]),
        motif_start = start(fimo_gr)[queryHits(hits)],
        motif_end = end(fimo_gr)[queryHits(hits)],
        motif_strand = as.character(strand(fimo_gr)[queryHits(hits)]),
        motif_score = mcols(fimo_gr)$score[queryHits(hits)],

    # which TE it overlaps
        TE_id = mcols(te_gr)$TE_id[subjectHits(hits)],
        TE_chr = as.character(seqnames(te_gr)[subjectHits(hits)]),
        TE_start = start(te_gr)[subjectHits(hits)],
        TE_end = end(te_gr)[subjectHits(hits)],
        TE_strand = as.character(strand(te_gr)[subjectHits(hits)]),

          rel_pos = rel_pos
        )

hit_list[[paste0(base_name, "_", mn)]] <- hit_df
      
      # ------------------------------
      # Store results
      # ------------------------------
      results_list[[paste0(base_name, "_", mn)]] <- data.frame(
        TE_name = base_name,
        motif_name = mn,
        total_TEs = n_te_total,
        TEs_with_TP53 = n_te_with_hits,
        total_TP53_hits = n_total_hits,
        occupancy = occupancy
      )
      
      # ------------------------------
      # Plot density
      # ------------------------------
      cat(paste0("### ", base_name, " ", mn, "\n"))
      cat(base_name, "| total TEs:", n_te_total, "| with cTP53:", n_te_with_hits, 
          "| hits:", n_total_hits, "\n")
      cat("Sequences with full TP53 motif: ", round(occupancy * 100, 2), "%\n\n")
      
      p <- ggplot(df, aes(x = rel_pos)) +
        geom_density(fill = "steelblue", alpha = 0.5) +
        labs(x = "Relative position in TE", y = "Motif density") +
        theme_classic() +
        ggtitle(paste0("canonical TP53 motif distribution: ", base_name,
                       "\nSequences with motif: ", round(occupancy * 100, 1), "%"))
      print(p)
      cat("\n\n")
    }
  } # end fimo_dirs loop
  
  # Combine all results into one data frame
  results_df <- do.call(rbind, results_list)
  
  # print(results_df)
  hits_df <- do.call(rbind, hit_list)

return(list(
  summary = results_df,
  hits = hits_df
))
}
    

```

```{r pulling canonical assignments, eval=FALSE}
full_p53 <- p53_canonical_motifs("C:/Users/renee/Other_projects_data/DXR_data/TE_sea_jaspar_results","data/Bed_exports/repeatmasker_sets/ERV1")

p53_canonical <- full_p53$hits
```


### Overlap DARs v CARs TEs

```{r fisher repname test}
test_pair_TE_repName <- function(df_long, rep_name, cluster1, cluster2) {
  
  # Subset for the specific repName
  sub_df <- df_long %>%
    filter(repName == rep_name) %>%
    complete(
      cluster = c(cluster1, cluster2),
      status  = c("TE", "not_TE"),
      fill = list(count = 0)
    )
  
  # fixed order of statuses
  status_levels <- c("TE", "not_TE")
  
  # make sure both statuses exist
  statuses <- unique(sub_df$status)
  if(length(statuses) != 2) {
    sub_df <- sub_df %>%
      complete(cluster, status, fill = list(count = 0))
    statuses <- unique(sub_df$status)
  }
  
  # counts for cluster1
  c1_counts <- sub_df %>%
    filter(cluster == cluster1) %>%
    arrange(factor(status, levels = status_levels)) %>%
    pull(count)
  
  # counts for cluster2
  c2_counts <- sub_df %>%
    filter(cluster == cluster2) %>%
    arrange(factor(status, levels = status_levels)) %>%
    pull(count)
  
  # 2x2 matrix for Fisher test
  mat <- matrix(
    c(c2_counts, c1_counts),
    nrow = 2,
    byrow = TRUE,
    dimnames = list(
      cluster = c(cluster2, cluster1),
      category = status_levels
    )
  )
  
  ft <- tryCatch(
    fisher.test(mat, workspace = 2e8),
    error = function(e) fisher.test(mat, simulate.p.value = TRUE, B = 1e5)
  )
  
  tibble(
    repName     = rep_name,
    comparison  = paste(cluster2, "vs", cluster1),
    odds_ratio  = ft$estimate,
    lower_CI    = ft$conf.int[1],
    upper_CI    = ft$conf.int[2],
    p_value     = ft$p.value
  )
}

```

Evaluate DAR and CAR overlap of TEs in ATAC data
```{r enrichment plot}
repeatmasker_clean <- readRDS("data/RDS_files/repeatmasker_clean.RDS")

sat_lowc_simrpt_repNames <- readRDS("data/RDS_files/sat_lowc_simrpt_repNames.RDS")

rmskr_std_gr <- GRanges(
  seqnames = repeatmasker_clean$genoName,
  ranges   = IRanges(
  start = repeatmasker_clean$start,
  end   = repeatmasker_clean$end),
  strand   = repeatmasker_clean$strand,
  repName   = repeatmasker_clean$repName,
  repClass  = repeatmasker_clean$repClass,
  repFamily = repeatmasker_clean$repFamily,
  swScore   = repeatmasker_clean$swScore,
  milliDiv  = repeatmasker_clean$milliDiv,
  milliDel  = repeatmasker_clean$milliDel,
  milliIns  = repeatmasker_clean$milliIns,
  id        = repeatmasker_clean$RM_id)

mcols(DOX_24_CAR_gr)$cluster_type <- "CAR"
mcols(DOX_24_DAR_gr)$cluster_type <- "DAR"
ATAC_gr <- c(DOX_24_DAR_gr,DOX_24_CAR_gr)


  
   # 1️⃣ find overlaps
  hits <- findOverlaps(ATAC_gr, rmskr_std_gr)
  
  if (length(hits) == 0) {
  ATAC_range_ols <- tibble()
} else {

  ATAC_range_ols <- tibble(
    Peakid     = ATAC_gr$peak[queryHits(hits)],
    seqnames   = as.character(seqnames(ATAC_gr))[queryHits(hits)],
    cluster = ATAC_gr$cluster_type[queryHits(hits)],
    repName   = rmskr_std_gr$repName[subjectHits(hits)],
    repFamily = rmskr_std_gr$repFamily[subjectHits(hits)],
    repClass  = rmskr_std_gr$repClass[subjectHits(hits)],
    milliDiv  = rmskr_std_gr$milliDiv[subjectHits(hits)],
    milliDel  = rmskr_std_gr$milliDel[subjectHits(hits)],
    milliIns  = rmskr_std_gr$milliIns[subjectHits(hits)],
    ID  = rmskr_std_gr$id[subjectHits(hits)]
  )
}
  
comparisons <- tibble(
  cluster2 = c("DAR"),
  cluster1 = c("CAR")
)


repName_counts <- ATAC_range_ols %>%
   dplyr::filter(!repName %in% sat_lowc_simrpt_repNames) %>% 
  count(cluster, repName, name = "count") %>%
  mutate(status = "TE")

total_cluster_counts <- tibble(
  cluster = mcols(ATAC_gr)$cluster_type
) %>% count(cluster, name = "total")

not_repName_counts <- repName_counts %>%
  left_join(total_cluster_counts, by = "cluster") %>%
  mutate(count = total - count,
         status = "not_TE") %>%
  dplyr::select(cluster, repName, status, count)

repName_df_long <- bind_rows(repName_counts %>%
                            dplyr::select(cluster, repName, status, count),
                          not_repName_counts) %>% 
  filter(!is.na(cluster))

repName_totals <- repName_df_long %>%
  filter(status == "TE") %>%
  group_by(repName) %>%
  summarize(total_TE = sum(count), .groups = "drop")

repName_df_long <- repName_df_long %>%
  filter(!is.na(repName)) %>%                      # remove NA repNames
  group_by(repName) %>%
  filter(sum(count[status == "TE"]) >= 10) %>%    # keep only sufficiently abundant TEs
  ungroup()

repName_results <- comparisons %>%
  mutate(results = map2(cluster2, cluster1, function(c2, c1) {

      repName_df_long %>%
        distinct(repName) %>%
        pull() %>%
        map_dfr(function(te) {
          test_pair_TE_repName(
            repName_df_long,
            rep_name  = te,
            cluster1 = c1,
            cluster2 = c2
          )
        })

    })
  ) %>%
  unnest(results) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))
repName_results_final <- repName_results %>% 
  dplyr::filter(!repName %in% sat_lowc_simrpt_repNames) 

ggplot(repName_results_final, aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  geom_text_repel(data = subset(repName_results, FDR < 0.01 & odds_ratio >1),
    size = 3,
    force = 3,
    box.padding = 0.6,
    point.padding = 0.4,
    max.overlaps = Inf,
    segment.color = "grey50",
    segment.size = 0.3
  ) +
  # geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.01), linetype = "solid") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "all repName enrichment in DAR v CAR ATAC regions"
  ) +
  theme_classic() 
```

```{r just enriched}
repName_results_final %>% 
  dplyr::filter(odds_ratio>1) %>% 
  ggplot(., aes(x = (odds_ratio), y = -log10(FDR),label = repName)) +
  geom_point(aes(color = odds_ratio > 1), size = 3) +
  geom_text_repel(data = subset(repName_results, FDR < 0.01 & odds_ratio >1),
    size = 3,
    force = 3,
    box.padding = 0.6,
    point.padding = 0.4,
    max.overlaps = Inf,
    segment.color = "grey50",
    segment.size = 0.3
  ) +
  # geom_text_repel(size=3, max.overlaps = Inf)+
  # geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.01), linetype = "dotted") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "all onlye enriched repName in DAR v CAR ATAC regions"
  ) +
  theme_bw() 
```


### LFC ATAC v LFC H3K27ac
#### MER61E

```{r LFC ATAC v K27ac}
H3K27ac_MER61E_LFC <-H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(repName=="MER61E") %>% 
  dplyr::select(Peakid,repID) %>% 
  dplyr::rename("H3K27ac_peakid"=Peakid) %>% 
  left_join((H3K27ac_toplist$H3K27ac_24T %>% dplyr::select(genes,logFC)), by=c("H3K27ac_peakid"="genes")) %>% 
  dplyr::rename("H3K27ac_LFC"=logFC)

H3K27ac_acuteMER61E_LFC <-H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(repName=="MER61E"&cluster=="Set_3") %>% 
  dplyr::select(Peakid,repID) %>% 
  dplyr::rename("H3K27ac_peakid"=Peakid) %>% 
  left_join((H3K27ac_toplist$H3K27ac_24T %>% dplyr::select(genes,logFC)), by=c("H3K27ac_peakid"="genes")) %>% 
  dplyr::rename("H3K27ac_LFC"=logFC)



ATAC_MER61E_LFC <- ATAC_range_ols %>% 
  dplyr::filter(repName=="MER61E") %>% 
  group_by(ID) %>% 
  dplyr::select(Peakid,ID) %>% 
  dplyr::rename("ATAC_peakid"=Peakid,"repID"=ID) %>% 
  left_join((DOX_24_ATAC %>% dplyr::select(peak,logFC)), by=c("ATAC_peakid"="peak")) %>% 
  dplyr::rename("ATAC_24_LFC"=logFC)

### Joining by repID

Joined_df <- H3K27ac_MER61E_LFC %>% 
  left_join(ATAC_MER61E_LFC) 

fit_lm <- lm(ATAC_24_LFC~H3K27ac_LFC, data=Joined_df)

r2 <- summary(fit_lm)$r.squared
p <- summary(fit_lm)$coefficients["H3K27ac_LFC", "Pr(>|t|)"]

  ggplot(Joined_df,aes(x=H3K27ac_LFC,y=ATAC_24_LFC))+
  geom_point()+
  geom_smooth(method="lm")+
  #  ggpubr::stat_cor(
  #   method = "spearman",
  #   aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))
  # )
    annotate("text",x = Inf, y = Inf,
             hjust = 1.1, 
             vjust = 1.2,
             label = paste0("R² = ", round(r2, 3),
      "\np = ", signif(p, 3)))+
    theme_bw()+
    ggtitle("Scatterplot with linear model fit \n of regions overlapping the same MER61Es")+
    labs(x="H3K27ac tx+0 log fold change",
         y="ATAC 24 hour log fold change")

```
with correlation 
```{r}
ggplot(Joined_df,aes(x=H3K27ac_LFC,y=ATAC_24_LFC))+
  geom_point()+
  geom_smooth(method="lm", se=FALSE)+
   ggpubr::stat_cor(
    method = "spearman",
    aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))
  )+
    # annotate("text",x = Inf, y = Inf,
    #          hjust = 1.1, 
    #          vjust = 1.2,
    #          label = paste0("R² = ", round(r2, 3),
    #   "\np = ", signif(p, 3)))+
    theme_bw()+
    ggtitle("Scatterplot with Spearmann correlation\nof regions overlapping the same MER61Es")+
    labs(x="H3K27ac tx+0 log fold change",
         y="ATAC 24 hour log fold change")

```

#### LTR10E   


```{r LFC ATAC v K27ac LTR10e}
H3K27ac_LTR10E_LFC <-H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(repName=="LTR10E") %>% 
  dplyr::select(Peakid,repID) %>% 
  dplyr::rename("H3K27ac_peakid"=Peakid) %>% 
  left_join((H3K27ac_toplist$H3K27ac_24T %>% dplyr::select(genes,logFC)), by=c("H3K27ac_peakid"="genes")) %>% 
  dplyr::rename("H3K27ac_LFC"=logFC)

H3K27ac_acuteLTR10E_LFC <-H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(repName=="LTR10E"&cluster=="Set_3") %>% 
  dplyr::select(Peakid,repID) %>% 
  dplyr::rename("H3K27ac_peakid"=Peakid) %>% 
  left_join((H3K27ac_toplist$H3K27ac_24T %>% dplyr::select(genes,logFC)), by=c("H3K27ac_peakid"="genes")) %>% 
  dplyr::rename("H3K27ac_LFC"=logFC)



ATAC_LTR10E_LFC <- ATAC_range_ols %>% 
  dplyr::filter(repName=="LTR10E") %>% 
  group_by(ID) %>% 
  dplyr::select(Peakid,ID) %>% 
  dplyr::rename("ATAC_peakid"=Peakid,"repID"=ID) %>% 
  left_join((DOX_24_ATAC %>% dplyr::select(peak,logFC)), by=c("ATAC_peakid"="peak")) %>% 
  dplyr::rename("ATAC_24_LFC"=logFC)

### Joining by repID

Joined_df_ltr10e <- H3K27ac_LTR10E_LFC %>% 
  left_join(ATAC_LTR10E_LFC) 

fit_lm_ltr10e <- lm(ATAC_24_LFC~H3K27ac_LFC, data=Joined_df_ltr10e)

r2_ltr10e <- summary(fit_lm_ltr10e)$r.squared
p_ltr10e <- summary(fit_lm_ltr10e)$coefficients["H3K27ac_LFC", "Pr(>|t|)"]

  ggplot(Joined_df_ltr10e,aes(x=H3K27ac_LFC,y=ATAC_24_LFC))+
  geom_point()+
  geom_smooth(method="lm")+
  #  ggpubr::stat_cor(
  #   method = "spearman",
  #   aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))
  # )
    annotate("text",x = Inf, y = Inf,
             hjust = 1.1, 
             vjust = 1.2,
             label = paste0("R² = ", round(r2_ltr10e, 3),
      "\np = ", signif(p_ltr10e, 3)))+
    theme_bw()+
    ggtitle("Scatterplot with linear model fit \nof regions overlapping the same LTR10Es")+
    labs(x="H3K27ac tx+0 log fold change",
         y="ATAC 24 hour log fold change")

```

with correlation 
```{r}
ggplot(Joined_df_ltr10e,aes(x=H3K27ac_LFC,y=ATAC_24_LFC))+
  geom_point()+
  geom_smooth(method="lm", se=FALSE)+
   ggpubr::stat_cor(
    method = "spearman",
    aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))
  )+
    # annotate("text",x = Inf, y = Inf,
    #          hjust = 1.1, 
    #          vjust = 1.2,
    #          label = paste0("R² = ", round(r2, 3),
    #   "\np = ", signif(p, 3)))+
    theme_bw()+
    ggtitle("Scatterplot with Spearmann correlation\nof regions overlapping the same LTR10Es")+
    labs(x="H3K27ac tx+0 log fold change",
         y="ATAC 24 hour log fold change")

```

#### Acute-only MER61Es
```{r acute LFC atac v K27ac}
acute_joined_df_MER61E <- H3K27ac_acuteMER61E_LFC %>% 
  left_join(ATAC_MER61E_LFC) 

fit_lm_acute_MER61E <- lm(ATAC_24_LFC~H3K27ac_LFC, data=acute_joined_df_MER61E)

r2_acute_MER61E <- summary(fit_lm_acute_MER61E)$r.squared
p_acute_MER61E <- summary(fit_lm_acute_MER61E)$coefficients["H3K27ac_LFC", "Pr(>|t|)"]

  ggplot(acute_joined_df_MER61E,aes(x=H3K27ac_LFC,y=ATAC_24_LFC))+
  geom_point()+
  geom_smooth(method="lm")+
  #  ggpubr::stat_cor(
  #   method = "spearman",
  #   aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))
  # )
    annotate("text",x = Inf, y = Inf,
             hjust = 1.1, 
             vjust = 1.2,
             label = paste0("R² = ", round(r2_acute_MER61E, 3),
      "\np = ", signif(p_acute_MER61E, 3)))+
    theme_bw()+
    ggtitle("Scatterplot with linear model fit \n(not correlation) of regions overlapping Acute MER61Es")+
    labs(x="H3K27ac tx+0 log fold change",
         y="ATAC 24 hour log fold change")

```

```{r}
ggplot(acute_joined_df_MER61E,aes(x=H3K27ac_LFC,y=ATAC_24_LFC))+
  geom_point()+
  geom_smooth(method="lm", se=FALSE)+
   ggpubr::stat_cor(
    method = "spearman",
    aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))
  )+
    # annotate("text",x = Inf, y = Inf,
    #          hjust = 1.1, 
    #          vjust = 1.2,
    #          label = paste0("R² = ", round(r2_acute, 3),
    #   "\np = ", signif(p_acute, 3)))+
    theme_bw()+
    ggtitle("Scatterplot correlation (Spearmann) of regions overlapping \nAcute MER61Es")+
    labs(x="H3K27ac tx+0 log fold change",
         y="ATAC 24 hour log fold change")

```

#### Acute-only LTR10Es

```{r acute LFC atac v K27ac ltr10e}
acute_joined_df_ltr10e <- H3K27ac_acuteLTR10E_LFC %>% 
  left_join(ATAC_LTR10E_LFC) 

fit_lm_acute_ltr10e <- lm(ATAC_24_LFC~H3K27ac_LFC, data=acute_joined_df_ltr10e)

r2_acute_ltr10e <- summary(fit_lm_acute_ltr10e)$r.squared
p_acute_ltr10e <- summary(fit_lm_acute_ltr10e)$coefficients["H3K27ac_LFC", "Pr(>|t|)"]

  ggplot(acute_joined_df_ltr10e,aes(x=H3K27ac_LFC,y=ATAC_24_LFC))+
  geom_point()+
  geom_smooth(method="lm")+
  #  ggpubr::stat_cor(
  #   method = "spearman",
  #   aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))
  # )
    annotate("text",x = Inf, y = Inf,
             hjust = 1.1, 
             vjust = 1.2,
             label = paste0("R² = ", round(r2_acute_ltr10e, 3),
      "\np = ", signif(p_acute_ltr10e, 3)))+
    theme_bw()+
    ggtitle("Scatterplot with linear model fit \n(not correlation) of regions overlapping Acute LTR10Es")+
    labs(x="H3K27ac tx+0 log fold change",
         y="ATAC 24 hour log fold change")

```
```{r}
ggplot(acute_joined_df_ltr10e,aes(x=H3K27ac_LFC,y=ATAC_24_LFC))+
  geom_point()+
  geom_smooth(method="lm", se=FALSE)+
   ggpubr::stat_cor(
    method = "spearman",
    aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))
  )+
    # annotate("text",x = Inf, y = Inf,
    #          hjust = 1.1, 
    #          vjust = 1.2,
    #          label = paste0("R² = ", round(r2_acute, 3),
    #   "\np = ", signif(p_acute, 3)))+
    theme_bw()+
    ggtitle("Scatterplot correlation (Spearmann) of regions overlapping \nAcute LTR10Es")+
    labs(x="H3K27ac tx+0 log fold change",
         y="ATAC 24 hour log fold change")

```

### Age and LFC
#### MER61E  


```{r age and LFC}
MER61E_K27acLFC_age <-H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(repName=="MER61E") %>% 
  dplyr::select(Peakid,repID, cluster) %>%
  left_join(repeatmasker_clean, by=c("repID"="RM_id")) %>% 
  dplyr::rename("H3K27ac_peakid"=Peakid) %>% 
  left_join((H3K27ac_toplist$H3K27ac_24T %>% dplyr::select(genes, logFC)), by=c("H3K27ac_peakid"="genes")) %>% 
  left_join((mint_anno %>% dplyr::select(repID, p53_halfsite)), by="repID")


ggplot(MER61E_K27acLFC_age, aes(x=logFC,y=milliDiv))+
  geom_point(aes(color=p53_halfsite))+
  ggtitle("H3K27ac logFC at 0R which overlap a MER61E \nand the milliDiv of the MER61E, n=67")+
  theme_bw()
```


```{r acute set age and LFC}

acute_MER61E_K27acLFC_age <- MER61E_K27acLFC_age %>% 
  dplyr::filter(cluster=="Set_3")

ggplot(acute_MER61E_K27acLFC_age, aes(x=logFC,y=milliDiv))+
  geom_point(aes(color=p53_halfsite))+
  ggtitle("H3K27ac logFC at 0R which overlap an acute MER61E \nand the milliDiv of the MER61E, n=40")+
  theme_bw()
```


#### LTR10E

```{r age and LFC LTR10E}
LTR10E_K27acLFC_age <-H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(repName=="LTR10E") %>% 
  dplyr::select(Peakid,repID, cluster) %>%
  left_join(repeatmasker_clean, by=c("repID"="RM_id")) %>% 
  dplyr::rename("H3K27ac_peakid"=Peakid) %>% 
  left_join((H3K27ac_toplist$H3K27ac_24T %>% dplyr::select(genes, logFC)), by=c("H3K27ac_peakid"="genes")) %>% 
  left_join((lint_anno %>% dplyr::select(repID, p53_halfsite)), by="repID")


ggplot(LTR10E_K27acLFC_age, aes(x=logFC,y=milliDiv))+
  geom_point(aes(color=p53_halfsite))+
  ggtitle("H3K27ac logFC at 0R which overlap a LTR10E \nand the milliDiv of the LTR10E, n=35")+
  theme_bw()
```


```{r acute set age and LFC LTR10E}

acute_LTR10E_K27acLFC_age <- LTR10E_K27acLFC_age %>% 
  dplyr::filter(cluster=="Set_3")

ggplot(acute_LTR10E_K27acLFC_age, aes(x=logFC,y=milliDiv))+
  geom_point(aes(color=p53_halfsite))+
  ggtitle("H3K27ac logFC at 0R which overlap an acute LTR10E \nand the milliDiv of the LTR10E, n=28")+
  theme_bw()+
  coord_cartesian(xlim=c(-2,5))
```

mer61e

```{r LFC and motifscore mer61e}

MER61E_K27acLFC_age %>% 
  left_join(MER61E_best_site_signal,by="repID") %>% 
  left_join((mint_anno %>% dplyr::select(repID,cluster)), by=c("repID","cluster")) %>% 
   mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>% 
  ggplot(., aes(x=logFC, y=MER61E.signal))+
# geom_point(aes(color=cluster))+
  geom_point()+
  ggtitle("H3K27ac logFC at 0R against p53half motif score ")+
  theme_bw()


MER61E_K27acLFC_age %>% 
  left_join(MER61E_best_site_signal,by="repID") %>% 
  left_join((mint_anno %>% dplyr::select(repID,cluster)), by=c("repID","cluster")) %>% 
   mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>% 
  ggplot(., aes(y=milliDiv, x=MER61E.signal))+
# geom_point(aes(color=cluster))+
  geom_point()+
  ggtitle("MilliDiv against p53half motif score ")+
  theme_bw()
```

ltr10e
```{r LFC and motifscore ltr10e}

LTR10E_K27acLFC_age %>% 
  left_join(LTR10E_best_site_signal,by="repID") %>% 
  left_join((lint_anno %>% dplyr::select(repID,cluster)), by=c("repID","cluster")) %>% 
   mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>% 
  ggplot(., aes(x=logFC, y=LTR10E.signal))+
# geom_point(aes(color=cluster))+
  geom_point()+
  ggtitle("H3K27ac logFC at 0R against p53half motif score ")+
  theme_bw()


LTR10E_K27acLFC_age %>% 
  left_join(LTR10E_best_site_signal,by="repID") %>% 
  left_join((lint_anno %>% dplyr::select(repID,cluster)), by=c("repID","cluster")) %>% 
   mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>% 
  ggplot(., aes(y=milliDiv, x=LTR10E.signal))+
# geom_point(aes(color=cluster))+
  geom_point()+
  ggtitle("MilliDiv against p53half motif score ")+
  theme_bw()
```
### More data and proportion plots
```{r TE peak id info}
MER61E_mydata <- H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(repName=="MER61E")%>% 
  pull(repID) %>% 
  unique()

Peakid_lookup_repID <-
  H3K27ac_ext200_anno %>% 
  separate_longer_delim(.,cols=c(repName,repID),delim=";") %>% 
    dplyr::filter(repName=="MER61E") %>% 
    dplyr::select(repID,Peakid,geneId,distanceToTSS)
```

```{r determining which repIDs have binding}
ZNF780A_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,ZNF780A_regions) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()
ZNF680_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,ZNF680_regions) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

p53_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,p53_sp_ChIP) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

CTCF_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,CTCF_sp_ChIP) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

RAD21_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,RAD21_sp_ChIP) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

TOP2B_overlaps_MER61E <- subsetByOverlaps(MER61E_regions,TOP2B_sp_ChIP) %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()
```

```{r overlappig DARs and CARs}

DAR_set <- subsetByOverlaps(MER61E_regions,DOX_24_DAR_gr)
DAR_MER61E_set <- DAR_set %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

CAR_set <- subsetByOverlaps(MER61E_regions,DOX_24_CAR_gr)
CAR_MER61E_set <- CAR_set %>% 
  as.data.frame() %>% 
  pull(name) %>% 
  unique()

```



```{r addingin other data}
Left_ventricle_TAD <- import(
  con = "C://Users/renee/Downloads/hg38.TADs/hg38/VentricleLeft_STL003_Leung_2015-raw_TADs.txt",
  format = "bed",
  genome = "hg38"
)

mcols(Left_ventricle_TAD)$TAD_id <- paste0("TAD_", seq_along(Left_ventricle_TAD))

SCREEN_heart_hg38_cCREs <- import(
  "data/Other_paper_data/SCREEN_LVHEART_F_53_ENCFF417JSF_ENCFF651XRK_ENCFF320IPT_ENCFF440RUS.bed",
  format = "BED",
  extraCols = c(
    DHS_ID = "character",
    cCRE_ID = "character",
    cCRE_class = "character"))

SCREEN_hg38_cCREs <- import(
  "data/Bed_exports/cisRE/GRCh38-cCREs.bed",
  format = "BED",
  extraCols = c(
    DHS_ID = "character",
    cCRE_ID = "character",
    cCRE_class = "character"))

# TP53_A549_regions<- import("data/Other_paper_data/ENCFF229ULU_A549_exp_ENCSR112XUO_TP53.bigBed", format="BigBED")

TP53_HEPG2_regions<- import("data/Other_paper_data/ENCFF687JDU_HepG2_exp_ENCSR980EGJ.bigBed", format="BigBED")

H3K9me3_toplist <- readRDS( "data/DER_data/H3K9me3_toplist.RDS")
K9me3_24T <- H3K9me3_toplist$H3K9me3_24T 
K9me3_24T <- K9me3_24T %>% 
  mutate(LFC_split=case_when(logFC>0&adj.P.Val<0.05~"K9_DER_up",
                             logFC<0&adj.P.Val<0.05~"K9_DER_down",
                             TRUE~"not_DER"))  



```

```{r K9me3 overlap}
K9me3_gr <- peakAnnoList_H3K9me3$all_H3K9me3_regions %>% as.GRanges()


mcols(K9me3_gr) <-
mcols(K9me3_gr) %>%
  as.data.frame() %>%
  left_join(
    K9me3_24T %>% dplyr::select(genes, LFC_split),
    by = c("Peakid"="genes"))


K9me3_split_gr <- split(K9me3_gr, K9me3_gr$LFC_split)


```

```{r merging tads}
MER61E_tad <- mergeByOverlaps(MER61E_regions,Left_ventricle_TAD)

MER61E_tad_df <- MER61E_tad %>%as.data.frame() %>% 
  group_by(MER61E_regions.name) %>%tally() 
  

# MER61E_p53_pub <- mergeByOverlaps(MER61E_regions,TP53_A549_regions)

# MER61E_tp53_df <- MER61E_p53_pub %>% as.data.frame() %>% 
#   group_by(MER61E_regions.name) %>% tally()

MER61E_p53_hepg2 <- mergeByOverlaps(MER61E_regions,TP53_HEPG2_regions)

MER61E_tp53_hepg2_df <- MER61E_p53_hepg2 %>% as.data.frame() %>% 
  group_by(MER61E_regions.name) %>% tally()

MER61E_k9upder <- mergeByOverlaps(MER61E_regions,K9me3_split_gr$K9_DER_up)

MER61E_k9downder <- mergeByOverlaps(MER61E_regions,K9me3_split_gr$K9_DER_down)

MER61E_k9notder <- mergeByOverlaps(MER61E_regions,K9me3_split_gr$not_DER)

```



```{r cre overlaps}
MER61E_cre <- mergeByOverlaps(MER61E_regions,SCREEN_heart_hg38_cCREs)

MER61E_cre_df <- MER61E_cre %>% 
  as.data.frame() %>% 
  group_by(MER61E_regions.name) %>% 
  # dplyr::filter(MER61E_regions.name %in%MER61E_mydata) %>% 
  left_join((SCREEN_hg38_cCREs %>% as.data.frame()), by=c("SCREEN_heart_hg38_cCREs.name" = "cCRE_ID")) %>% 
  dplyr::select(name,name.1,cCRE_class.y) %>% 
  group_by(name,cCRE_class.y) %>% tally() %>% 
  left_join(mint_anno,by=c("name"="repID"))


```


```{r adding onto mintanno}

mint_anno <- mint_anno %>% 
  mutate(heart_cRE=if_else(repID %in% MER61E_cre_df$name,"yes","no"),
          LV_TAD_overlap=if_else(repID %in% MER61E_tad_df$MER61E_regions.name,"yes","no"),
         # TP53_A549_overlap=if_else(repID %in% MER61E_tp53_df$MER61E_regions.name,"yes","no"),
         TP53_HEPG2_overlap=if_else(repID %in% MER61E_tp53_hepg2_df$MER61E_regions.name,"yes","no"),
         MER61E_k9notder=if_else( repID %in% MER61E_k9notder[["name"]],"yes","no"),
         MER61E_k9downder=if_else( repID %in% MER61E_k9downder[["name"]],"yes","no"),
         MER61E_k9upder=if_else( repID %in% MER61E_k9upder[["name"]],"yes","no"))  


  
```



```{r function prop_plot}
prop_plot <- function(data, group_col, yes_no_col) {

  # Prepare data
  plot_data <- data %>%
    mutate(cluster = tidyr::replace_na(cluster, "NA"),
      cluster = factor(cluster,
        levels = c("Set_3", "Set_1", "not_assigned", "NA"))) %>%
    filter(cluster %in% c("Set_3", "Set_1", "not_assigned", "NA"))

  
  # Pairwise comparisons against Set_3
  comparison_groups <- c("Set_1", "not_assigned", "NA")

  test_results <- lapply(comparison_groups, function(g) {

    test_df <- plot_data %>%
      filter(cluster %in% c("Set_3", g))

    tab <- table(test_df$cluster,
              test_df[[rlang::as_name(rlang::ensym(yes_no_col))]])

    # Check expected counts
    chi <- suppressWarnings(chisq.test(tab))

    if (any(chi$expected < 5)) {
      test <- fisher.test(tab)
      method <- "Fisher"
    } else {
      test <- chi
      method <- "Chi-square"
    }

    data.frame(comparison = paste("Set_3 vs", g),
               group = g,
               p = test$p.value,
               method = method)}) %>%
    bind_rows() %>%
    mutate(p_adj = p.adjust(p, method = "BH"),
           label = paste0("p=", format.pval(p_adj, digits = 2, eps = 0.001)),
           y = c(1.10, 1.18, 1.26))


  # Proportions
  count_data <- plot_data %>%
    count({{ group_col }}, {{ yes_no_col }}) %>%
    group_by({{ group_col }}) %>%
    mutate(prop = n / sum(n)) %>%
    ungroup()


  # Plot
  ggplot(count_data,aes(x = {{ group_col }},y = prop,
      fill = {{ yes_no_col }})) +
    geom_col() +
    geom_text(aes(label = paste0(scales::percent(prop, accuracy = 1), "\n(n=", n, ")")),
              position = position_stack(vjust = 0.5)) +

    # p-values over comparison groups
    geom_text(data = test_results,aes(x = group,y = y, label = label),
              inherit.aes = FALSE,
              size = 3.5) +
    scale_y_continuous(labels = scales::percent,
                       limits = c(0, 1.3),
                       breaks = seq(0, 1, 0.25)) +
    scale_fill_manual(values =RColorBrewer::brewer.pal(12, "Paired")[1:2]) +

    labs(x = NULL,y = "Proportion") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90,
                                     hjust = 1))
}


```

```{r prop set 1}

prop_plot(mint_anno, cluster,p53_halfsite)+
  ggtitle("p53 halfsite proportion across MER61E")


prop_plot(mint_anno, cluster,p53_fullsite)+
  ggtitle("p53 fullsite proportion across MER61E")


prop_plot(mint_anno, cluster,ZNF780A_binding)+
  ggtitle("ZNF780A binding proportion across MER61E")


prop_plot(mint_anno, cluster,ZNF680_binding)+
  ggtitle("ZNF680 binding proportion across MER61E")


prop_plot(mint_anno, cluster,sp_p53_binding)+
  ggtitle("DOX_ChIP p53_binding proportion across MER61E")


prop_plot(mint_anno, cluster,overlap_DAR)+
  ggtitle("DAR overlap proportion across MER61E")
```

```{r prop set 2}
prop_plot(mint_anno,cluster, LV_TAD_overlap)+
  ggtitle("LV TAD overlap of MER61Es")

prop_plot(mint_anno,cluster, heart_cRE)+
  ggtitle("Heart cRE overlap of MER61Es")

# prop_plot(mint_anno,cluster, TP53_A549_overlap)+
#   ggtitle("Published TP53 regions from A549 cells and MER61E overlaps")


prop_plot(mint_anno,cluster, TP53_HEPG2_overlap)+
  ggtitle("Published TP53 regions from HepG2 cells and MER61E overlaps")
```

```{r prop set 3}

prop_plot(mint_anno,cluster, MER61E_k9upder)+
  ggtitle("K9me3 overlapped MER61Es by cluster up DER ")

  prop_plot(mint_anno,cluster, MER61E_k9downder)+
  ggtitle("K9me3 overlapped MER61Es by cluster down DER ")

prop_plot(mint_anno,cluster,MER61E_k9notder)+
  ggtitle("K9me3 overlapped MER61Es by cluster not DER ")

```

```{r RNA}
RNA_acute_cluster <- read_excel("C:/Users/renee/Other_projects_data/DXR_data/DXR_RNA_data/Pfortmiller_S2_Appendix_251213.xlsx", 
    sheet = "Table_S6", skip = 1)

RNA_chronic_cluster <- read_excel("C:/Users/renee/Other_projects_data/DXR_data/DXR_RNA_data/Pfortmiller_S2_Appendix_251213.xlsx", 
    sheet = "Table_S7", skip = 2)

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

all_gene_gr <- genes(txdb)

acute_gr <-  all_gene_gr[
  names(all_gene_gr) %in% RNA_acute_cluster$Entrez_ID]
chronic_gr <-  all_gene_gr[
  names(all_gene_gr) %in% RNA_chronic_cluster$Entrez_ID]

acute_overlap <- mergeByOverlaps(acute_gr,Left_ventricle_TAD)

chronic_overlap <- mergeByOverlaps(chronic_gr,Left_ventricle_TAD)


acute_chron_TADS <-c(acute_overlap[["TAD_id"]],chronic_overlap[["TAD_id"]]) %>% 
  unique()


MER61E_acute_chron_TAD <- MER61E_tad %>% as.data.frame() %>% 
  mutate(acuteChron= if_else(Left_ventricle_TAD.TAD_id %in% acute_chron_TADS,"yes","no")) %>% 
  dplyr::filter(acuteChron=="yes") %>% 
  pull(name)

mint_anno <- mint_anno %>% 
  mutate(AC_TADs=if_else(repID %in% MER61E_acute_chron_TAD, "yes","no"))

prop_plot(mint_anno,cluster,AC_TADs)+
  ggtitle("MER61E that overlap a TAD with a chronic/acute gene")

H3K27ac_ext200_anno %>% group_by(cluster) %>% tally()
```

### Boxplot alley

```{r canonical p53 boxplot}

full_p53_canonical %>% 
  dplyr::filter(TE_name=="MER61E_ERV1_LTR") %>% 
  dplyr::select(TE_id, motif_score) %>% 
  left_join((mint_anno %>%
               dplyr::select(repID,cluster)),
            by=c("TE_id"="repID")) %>% 
  mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>%
  # group_by(cluster) %>% tally()
  ggplot(., aes(x=cluster,y=motif_score))+
  geom_boxplot()+
  geom_point(position="jitter")+
  ggpubr::stat_compare_means(
    comparisons = list(
      c("Set_3", "Set_1"),
      c("Set_3", "NA")
    ),
    method = "wilcox.test"
  )+
  theme_bw()+
  ggtitle("n=22 MER61Es and motif scores for canonical p53 motif")

full_p53_canonical %>% 
  dplyr::filter(TE_name=="LTR10E_ERV1_LTR") %>% 
  dplyr::select(TE_id, motif_score) %>% 
  left_join((lint_anno %>%
               dplyr::select(repID,cluster)),
            by=c("TE_id"="repID")) %>% 
  mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>%
  # group_by(cluster) %>% tally()
  ggplot(., aes(x=cluster,y=motif_score))+
  geom_boxplot()+
  geom_point(position="jitter")+
  ggpubr::stat_compare_means(
    comparisons = list(
      c("Set_3", "Set_1"),
      c("Set_3", "NA")
    ),
    method = "wilcox.test"
  )+
  theme_bw()+
  ggtitle("n=54 LTR10Es and motif scores for canonical p53 motif")
```

```{r p53halfsite boxplot}
mint_anno %>% 
  dplyr::select(repID,cluster,p53_halfsite,p53_fullsite) %>% 
  left_join(MER61E_best_site_signal) %>% 
  mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>%
  dplyr::filter(!is.na(MER61E.signal)) %>% 
  # group_by(cluster) %>% tally()
  ggplot(., aes(x=cluster,y=MER61E.signal))+
  geom_boxplot()+
  geom_point(position="jitter")+
  ggpubr::stat_compare_means(
    comparisons = list(
      c("Set_3", "Set_1"),
      c("Set_3","not_assigned"),
      c("Set_3", "NA")
    ),
    method = "wilcox.test"
  )+
  theme_bw()+
  ggtitle("n=156 MER61Es and motif scores for halfsite p53 motif")


lint_anno %>% 
  dplyr::select(repID,cluster,p53_halfsite,p53_fullsite) %>% 
  left_join(LTR10E_best_site_signal) %>% 
  mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>%
  dplyr::filter(!is.na(LTR10E.signal)) %>% 
  # group_by(cluster) %>% tally()
  ggplot(., aes(x=cluster,y=LTR10E.signal))+
  geom_boxplot()+
  geom_point(position="jitter")+
  ggpubr::stat_compare_means(
    comparisons = list(
      # c("Set_3", "Set_1"),
      # c("Set_3","not_assigned"),
      c("Set_3", "NA")
    ),
    method = "wilcox.test"
  )+
  theme_bw()+
  ggtitle("n=152 LTR10Es and motif scores for halfsite p53 motif")
```

```{r ATAC LFC Boxplot}
MER61E_rep_cluster <- mint_anno %>% 
  dplyr::select(repID,cluster,p53_halfsite)

ATAC_MER61E_LFC %>% 
  left_join(MER61E_rep_cluster) %>% 
  mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>%
   # group_by(cluster) %>% tally()
  ggplot(., aes(x=cluster,y=ATAC_24_LFC))+
  geom_boxplot()+
  geom_point(aes(color=p53_halfsite),position="jitter")+
  ggpubr::stat_compare_means(
    comparisons = list(
      c("Set_3", "Set_1"),
      c("Set_3","not_assigned"),
      c("Set_3", "NA")
    ),
    method = "wilcox.test"
  )+
  theme_bw()+
  ggtitle("n=85 MER61Es which overlap an ATAC 24hour region")
```
This is a weird plot to gather data for.  What I did:  I pulled the repID and cluster association from the K27ac data.  Then I used the ATAC LFC, and filtered the peakid by the ATAC regions that overlap a MER61E/ and joined by cluster.  

```{r LTR10e version}
LTR10E_rep_cluster <- lint_anno %>% 
  dplyr::select(repID,cluster,p53_halfsite)

ATAC_LTR10E_LFC %>% 
  left_join(LTR10E_rep_cluster) %>% 
  mutate(cluster=factor(cluster, levels=c("Set_3","Set_2","Set_1","not_assigned","NA"))) %>%
   # group_by(cluster) %>% tally()
  ggplot(., aes(x=cluster,y=ATAC_24_LFC))+
  geom_boxplot()+
  geom_point(aes(color=p53_halfsite),position="jitter")+
  ggpubr::stat_compare_means(
    comparisons = list(
      c("Set_3", "Set_1"),
      c("Set_3","not_assigned"),
      c("Set_3", "NA")
    ),
    method = "wilcox.test"
  )+
  theme_bw()+
  ggtitle("n=21 LTR10Es which overlap an ATAC 24hour region")

```


### More proportion plots

```{r overlapping sets 2 mer61e}

upDAR_ols <- mergeByOverlaps(MER61E_regions,upDOX_24_DAR_gr)

p53_sp_ChIP_ols <- mergeByOverlaps(MER61E_regions,p53_sp_ChIP)

VEH_ATAC_ols <- mergeByOverlaps(MER61E_regions, ATAC_regions_Ind2)

K27ac_VEH_Ind4_ols <- mergeByOverlaps(MER61E_regions, K27ac_VEH_Ind4)

SOX11_peak_ols <- mergeByOverlaps(MER61E_regions,SOX11_hg38_gr)

#### ENCODE TRIM28 ChIP data (K562 cells):experiment: ENCSR474CVP, accession file processed data:ENCFF770JQM
TRIM28_K562 <- rtracklayer::import("data/Other_paper_data/ENCFF770JQMTRIM28_ENCSR474CVP.bed",format = "narrowPeak")
TRIM28_K562_ols <- mergeByOverlaps(MER61E_regions, TRIM28_K562)
#### Encode tried two, one from HEPG2 ENCSR193ADW:ENCFF615OJH and one from K562 ENCSR000AUT:ENCFF267QRI",
SETDB1_HEPG2 <- rtracklayer::import("data/Other_paper_data/ENCFF615OJH_SET1_ENCSR193ADW.bed",format = "narrowPeak")
SETDB1_HEPG2_ols <- mergeByOverlaps(MER61E_regions, SETDB1_HEPG2)
SETDB1_K562 <- rtracklayer::import("data/Other_paper_data/ENCFF267QRI_SETDB1_K562_ENCSR000AUT.bed",format = "narrowPeak")
SETDB1_K562_ols <- mergeByOverlaps(MER61E_regions, SETDB1_K562)

mint_anno_new <- mint_anno %>%
  dplyr::select(repID,cluster) %>% 
  mutate(upDAR_overlap=if_else(repID %in% upDAR_ols[["name"]],"yes","no"),
         p53_sp_ChIP_overlap=if_else(repID %in% p53_sp_ChIP_ols[["name"]],"yes","no"),
         VEH_ATAC_overlap=if_else(repID %in% VEH_ATAC_ols[["name"]],"yes","no"),
         K27_VEH_overlap=if_else( repID %in% K27ac_VEH_Ind4_ols[["name"]],"yes","no"),
         TRIM28_K562_overlap=if_else( repID %in% TRIM28_K562_ols[["name"]],"yes","no"),
         SETDB1_HEPG2_overlap=if_else( repID %in% SETDB1_HEPG2_ols[["name"]],"yes","no"),
         SETDB1_K562_overlap=if_else( repID %in% SETDB1_K562_ols[["name"]],"yes","no"),
         SOX11_peak_overlap=if_else(repID %in% SOX11_peak_ols[["name"]],"yes","no"))

mint_anno_new %>% 
  group_by(SOX11_peak_overlap) %>% 
  tally()
```


```{r overlapping sets 2 LTR10E}

upDAR_ols_ltr <- mergeByOverlaps(LTR10E_regions,upDOX_24_DAR_gr)

p53_sp_ChIP_ols_ltr <- mergeByOverlaps(LTR10E_regions,p53_sp_ChIP)

VEH_ATAC_ols_ltr <- mergeByOverlaps(LTR10E_regions, ATAC_regions_Ind2)

K27ac_VEH_Ind4_ols_ltr <- mergeByOverlaps(LTR10E_regions, K27ac_VEH_Ind4)

SOX11_peak_ols_ltr <- mergeByOverlaps(LTR10E_regions,SOX11_hg38_gr)
#### ENCODE TRIM28 ChIP data (K562 cells):experiment: ENCSR474CVP, accession file processed data:ENCFF770JQM
# TRIM28_K562 <- rtracklayer::import("data/Other_paper_data/ENCFF770JQMTRIM28_ENCSR474CVP.bed",format = "narrowPeak")
TRIM28_K562_ols_ltr <- mergeByOverlaps(LTR10E_regions, TRIM28_K562)
#### Encode tried two, one from HEPG2 ENCSR193ADW:ENCFF615OJH and one from K562 ENCSR000AUT:ENCFF267QRI",
# SETDB1_HEPG2 <- rtracklayer::import("data/Other_paper_data/ENCFF615OJH_SET1_ENCSR193ADW.bed",format = "narrowPeak")
SETDB1_HEPG2_ols_ltr <- mergeByOverlaps(LTR10E_regions, SETDB1_HEPG2)
# SETDB1_K562 <- rtracklayer::import("data/Other_paper_data/ENCFF267QRI_SETDB1_K562_ENCSR000AUT.bed",format = "narrowPeak")
SETDB1_K562_ols_ltr <- mergeByOverlaps(LTR10E_regions, SETDB1_K562)

lint_anno_new <- lint_anno %>%
  dplyr::select(repID,cluster) %>% 
  mutate(upDAR_overlap=if_else(repID %in% upDAR_ols_ltr[["name"]],"yes","no"),
         p53_sp_ChIP_overlap=if_else(repID %in% p53_sp_ChIP_ols[["name"]],"yes","no"),
         VEH_ATAC_overlap=if_else(repID %in% VEH_ATAC_ols_ltr[["name"]],"yes","no"),
         K27_VEH_overlap=if_else( repID %in% K27ac_VEH_Ind4_ols_ltr[["name"]],"yes","no"),
         TRIM28_K562_overlap=if_else( repID %in% TRIM28_K562_ols_ltr[["name"]],"yes","no"),
         SETDB1_HEPG2_overlap=if_else( repID %in% SETDB1_HEPG2_ols_ltr[["name"]],"yes","no"),
         SETDB1_K562_overlap=if_else( repID %in% SETDB1_K562_ols_ltr[["name"]],"yes","no"),
         SOX11_peak_overlap=if_else(repID %in% SOX11_peak_ols_ltr[["name"]],"yes","no"))

lint_anno_new %>% 
  group_by(SOX11_peak_overlap) %>% 
  tally()
```
### MER61E prop plots
```{r prop set4 MER61E}
prop_plot(mint_anno_new, cluster,upDAR_overlap)+
  ggtitle("upDAR overlaps proportion across MER61E")

prop_plot(mint_anno_new, cluster,p53_sp_ChIP_overlap)+
  ggtitle("sp_VEH p53 overlap proportions across MER61E")

prop_plot(mint_anno_new, cluster,VEH_ATAC_overlap)+
  ggtitle("ATAC 24hr VEH proportions across MER61E")

prop_plot(mint_anno_new, cluster,K27_VEH_overlap)+
  ggtitle("H3K27ac VEH tx+0 proportion across MER61E")

prop_plot(mint_anno_new, cluster,TRIM28_K562_overlap)+
  ggtitle("TRIM28 in K562 cells overlap proportion across MER61E")

prop_plot(mint_anno_new, cluster,SETDB1_HEPG2_overlap)+
  ggtitle("SETDB1 in HepG2 cells overlap proportion across MER61E")
prop_plot(mint_anno_new, cluster,SOX11_peak_overlap)+
  ggtitle("SOX11 peaks in Granta-519 MCL cells overlap proportion across MER61E")

```


### LTR10E prop plots

```{r prop set5 LTR10E}
prop_plot(lint_anno_new, cluster,upDAR_overlap)+
  ggtitle("upDAR overlaps proportion across LTR10E")

prop_plot(lint_anno_new, cluster,p53_sp_ChIP_overlap)+
  ggtitle("sp_VEH p53 overlap proportions across LTR10E")

prop_plot(lint_anno_new, cluster,VEH_ATAC_overlap)+
  ggtitle("ATAC 24hr VEH proportions across LTR10E")

prop_plot(lint_anno_new, cluster,K27_VEH_overlap)+
  ggtitle("H3K27ac VEH tx+0 proportion across LTR10E")

prop_plot(lint_anno_new, cluster,TRIM28_K562_overlap)+
  ggtitle("TRIM28 in K562 cells overlap proportion across LTR10E")

prop_plot(lint_anno_new, cluster,SETDB1_HEPG2_overlap)+
  ggtitle("SETDB1 in HepG2 cells overlap proportion across LTR10E")
prop_plot(lint_anno_new, cluster,SOX11_peak_overlap)+
  ggtitle("SOX11 peaks in Granta-519 MCL cells overlap proportion across LTR10E")

```




### SOX11 RNA expression

```{r SOX11 expression}
##6664 NCBI number for SOX11 uni-prot P35716

RNA_count_table <- read_csv("C:/Users/renee/Other_projects_data/DXR_data/DXR_RNA_data/filcpm_final_matrix.csv") %>% dplyr::rename(ENTREZID=1)%>% column_to_rownames("ENTREZID")
tab <- read_csv("C:/Users/renee/Other_projects_data/DXR_data/DXR_RNA_data/Pfortmiller_S2_Appendix_251213.xlsx")

```







```{r SOX11 expression RNA}
RNA_count_table %>% dplyr::filter(row.names(.)=="6664") %>% 
  pivot_longer(cols = colnames(.), names_to = "Samples",values_to = "lcpm") %>% 
  separate_wider_delim(cols=Samples,names=c("trt","time","ind"),delim="_",cols_remove = FALSE) %>% 
  mutate(time=factor(time,levels=c("24T","24R","144R")),
         trt=factor(trt,levels=c("DMSO","DOX"))) %>% 
  ggplot(., aes(x=interaction(trt,time),y=lcpm))+
  geom_boxplot(aes(fill=trt))+
  geom_point(aes(color=ind))+
  ggtitle("SOX11 RNA expression over time")+
  theme_bw()
```

```{r SOX10 expression}
RNA_count_table %>% dplyr::filter(row.names(.)=="6663") %>%
  pivot_longer(cols = colnames(.), names_to = "Samples",values_to = "lcpm") %>% 
  separate_wider_delim(cols=Samples,names=c("trt","time","ind"),delim="_",cols_remove = FALSE) %>% 
  mutate(time=factor(time,levels=c("24T","24R","144R")),
         trt=factor(trt,levels=c("DMSO","DOX"))) %>% 
  ggplot(., aes(x=interaction(trt,time),y=lcpm))+
  geom_boxplot(aes(fill=trt))+
  geom_point(aes(color=ind))+
  ggtitle("SOX10 RNA expression over time")+
  theme_bw()
```






### Extra code bits
```{r misc code, eval=FALSE}

mutate(
    cluster = as.character(cluster),
    cluster = if_else(is.na(cluster), "NA", cluster),
    cluster = factor(cluster, levels = c("Set_1", "Set_3", "not_assigned", "NA"))) %>% 
  mutate(anno_alt=case_when(str_detect(annotation,"Intron")~"Intron",
         str_detect(annotation, "Exon") ~ "Exon",
      str_detect(annotation, "Promoter") ~ "Promoter",
      str_detect(annotation, "Distal Intergenic") ~ "Distal Intergenic",
      str_detect(annotation, "Downstream") ~ "Downstream",
      TRUE ~ annotation)) %>% 
  
  p53_label <- c(yes="p53 overlap",no="no p53 overlap")

facet_wrap(~p53_halfsite, labeller=as_labeller(p53_label))+
  mutate(MYA_jukes_cantor=(-.75*log(1-4*(milliDiv/1000)/3))/(1000000*2*2.2E-9),
         MYA_bourque=milliDiv/2.2) %>% 
  
  add_on_mintermediate_data <- mintermediate_data %>% 
  dplyr::select(repID,cluster, my_MER, H3K27ac_region,milliDiv:p53_fullsite,overlap_DAR, overlap_CAR, MER61E_phyloP_mean,CS_distanceToTSS) %>% 
  mutate(DER_24T=if_else(H3K27ac_region %in% regions_24T$genes, "DER_24T","not_DER_24T"),
         DER_24R=if_else(H3K27ac_region%in% regions_24R$genes,"DER_24R","not_DER_24R"),
         DER_144R=if_else(H3K27ac_region%in% regions_144R$genes,"DER_144R","not_DER_144R"))


separate_rows(GREAT_gene_assign, sep = ",\\s*") %>%
  mutate(
    gene = str_extract(GREAT_gene_assign, "^[A-Za-z0-9_.-]+"),
    distance = str_extract(GREAT_gene_assign,"(?<=\\()[+-]?\\d+(?=\\))") %>% as.numeric(),
    direction = case_when(
      distance < 0 ~ "upstream",
      distance > 0 ~ "downstream",
      distance == 0 ~ "overlapping",
      TRUE ~ NA_character_)) %>% 
```





```{r RNA_lfc_SYMBOL, eval=FALSE}

RNA_lfc_SYMBOL <-RNA_full_toptable %>% 
   dplyr::filter(SYMBOL != "NA")%>%
  dplyr::select(SYMBOL,logFC,timepoint, adj.P.Val) %>% 
  pivot_wider(., names_from = timepoint,
    values_from = c(logFC, adj.P.Val),
    names_glue = "RNA_{.value}_{timepoint}")

```


```{r bed files, eval=FALSE}
head(mcols(MER61E_regions)$name)

head(mint_anno)

set3_names <- mint_anno$repID[mint_anno$cluster=="Set_3"]

mset3_gr <- MER61E_regions[mcols(MER61E_regions)$name %in% set3_names]
length(mset3_gr)
mnotset3_gr <- MER61E_regions[!mcols(MER61E_regions)$name %in% set3_names]
length(mnotset3_gr)


export.bed(mset3_gr, con = "data/Bed_exports/MER61E_Set3.bed")
export.bed(mnotset3_gr, con = "data/Bed_exports/MER61E_notSet3.bed")


lset3_names <- lint_anno$repID[lint_anno$cluster %in% "Set_3"]

lset3_gr <- LTR10E_regions[mcols(LTR10E_regions)$name %in% lset3_names]
length(lset3_gr)
lnotset3_gr <- LTR10E_regions[!mcols(LTR10E_regions)$name %in% lset3_names]
length(lnotset3_gr)


export.bed(lset3_gr, con = "data/Bed_exports/LTR10E_Set3.bed")
export.bed(lnotset3_gr, con = "data/Bed_exports/LTR10E_notSet3.bed")
# lint_anno %>% group_by(cluster) %>% tally

both_set3_gr<- c(lset3_gr,mset3_gr)

both_notset3_gr <- c(lnotset3_gr,mnotset3_gr)

export.bed(both_set3_gr, con = "data/Bed_exports/twoTE_Set3.bed")
```

### Motif enrichment


This is utilizing xstreme to look for enrichment of Set3 TEs in MER61E vs the rest of MER61E TEs


#### MER61E  


```{r}
MER61E_Set3_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3v_therest_results/MER61E_Set3_xstreme/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

MER61E_Set3_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3v_therest_results/MER61E_Set3_xstreme/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

MER61E_Set3_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3v_therest_results/MER61E_Set3_xstreme/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


MER61E_bind <- bind_rows(MER61E_Set3_sea_disc_out, MER61E_Set3_sea_known) %>%
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID) %>% 
  distinct()

```

```{r}
MER61E_Set3_data <- MER61E_Set3_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(MER61E_bind, by=c("ID"="ID","ALT_ID"="ALT_ID", "CONSENSUS"="CONSENSUS")) %>% 
  mutate(diff_per=`TP%` -`FP%`)
```


```{r}
MER61E_Set3_data %>% 
  dplyr::filter(!is.na(MOTIF_URL)) %>% 
  group_by(CLUSTER) %>% 
  slice_min(EVALUE, with_ties = FALSE) %>% 
  mutate(order= row_number()) %>% 
  mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*1.5), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./1.5,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 enrichment of MER61E vs all other MER61E"))

```
```{r Filtering for expressed TFs, eval=TRUE}
RNA_full_toptable <-  readRDS("data/Other_paper_data/RNA_full_toptable.RDS")

RNA_expressed <- RNA_full_toptable %>% 
  dplyr::select(Entrez_ID, SYMBOL)

### had to add in a special filter to account for the :: in the names of the motifs
symbols <- str_to_upper(RNA_expressed$SYMBOL)

filtered_MER61E_set3_data <- MER61E_Set3_data %>%
  ##create the motif-tokens column so I can keep names where there is a partial match in the RNA Symbols
  mutate(motif_tokens = str_split(str_to_upper(motif_name), "::")) %>%
  ##Because this is a list, now I need to filter using sapply, but then remove the motif token column
  filter(sapply(motif_tokens, function(x) any(x %in% symbols))) %>%
  dplyr::select(-motif_tokens)

filtered_MER61E_set3_data %>% 
  dplyr::filter(!is.na(MOTIF_URL)) %>% 
  group_by(CLUSTER) %>% 
  slice_min(EVALUE, with_ties = FALSE) %>% 
  mutate(order= row_number()) %>% 
  mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*1.5), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./1.5,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 filtered motif enrichment of MER61E vs all other MER61E"))
```

#### LTR10E  


```{r}
LTR10E_Set3_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3v_therest_results/LTR10E_Set3_xstreme/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

LTR10E_Set3_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3v_therest_results/LTR10E_Set3_xstreme/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

LTR10E_Set3_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3v_therest_results/LTR10E_Set3_xstreme/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


LTR10E_bind <- bind_rows(LTR10E_Set3_sea_disc_out, LTR10E_Set3_sea_known) %>%
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID) %>% 
  distinct()

```

```{r}
LTR10E_Set3_data <- LTR10E_Set3_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(LTR10E_bind, by=c("ID"="ID","ALT_ID"="ALT_ID", "CONSENSUS"="CONSENSUS")) %>% 
  mutate(diff_per=`TP%` -`FP%`)
```


```{r}
LTR10E_Set3_data %>% 
  dplyr::filter(!is.na(MOTIF_URL)) %>% 
  group_by(CLUSTER) %>% 
  slice_min(EVALUE, with_ties = FALSE) %>% 
  ungroup() %>% 
  mutate(order= row_number()) %>%
  mutate(motif_name=if_else(motif_name%in% c("SOX10","Sox11","IKZF2"),paste0(motif_name,"_",order),motif_name)) %>% 
  mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*1), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./1,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 enrichment of LTR10E vs all other LTR10E"))

```
```{r Filtering for expressed TFs LTR10E, eval=TRUE}

filtered_LTR10E_set3_data <- LTR10E_Set3_data %>%
  ##create the motif-tokens column so I can keep names where there is a partial match in the RNA Symbols
  mutate(motif_tokens = str_split(str_to_upper(motif_name), "::")) %>%
  ##Because this is a list, now I need to filter using sapply, but then remove the motif token column
  filter(sapply(motif_tokens, function(x) any(x %in% symbols))) %>%
  dplyr::select(-motif_tokens)

filtered_LTR10E_set3_data %>% 
  dplyr::filter(!is.na(MOTIF_URL)) %>% 
  group_by(CLUSTER) %>% 
  slice_min(EVALUE, with_ties = FALSE) %>% 
  ungroup() %>% 
  mutate(order= row_number()) %>% 
  mutate(motif_name=if_else(motif_name%in% c("SOX10","Sox11","IKZF2"),paste0(motif_name,"_",order),motif_name)) %>%
  mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*1), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./1,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 filtered motif enrichment of LTR10E vs all other LTR10E"))
```

MER61E



#### LTR10E  


```{r}
twoTE_Set3_sea_disc_out <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3v_therest_results/twoTE_Set3_xstreme/sea_disc_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="disc") %>% slice_head(n = length(.$ID)-3)

twoTE_Set3_sea_known <-read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3v_therest_results/twoTE_Set3_xstreme/sea_out/sea.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="known") %>% slice_head(n = length(.$ID)-3)

twoTE_Set3_xstreme <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/TE_xstreme_TEset3v_therest_results/twoTE_Set3_xstreme/xstreme.tsv", delim = "\t", escape_double = FALSE,     trim_ws = TRUE) %>% mutate(source="all") %>% slice_head(n = length(.$ID)-3)


twoTE_bind <- bind_rows(twoTE_Set3_sea_disc_out, twoTE_Set3_sea_known) %>%
  dplyr::select(DB:ENR_RATIO) %>% 
  group_by(ID,ALT_ID) %>% 
  distinct()

```

```{r}
twoTE_Set3_data <- twoTE_Set3_xstreme %>% 
  dplyr::select(RANK:CLUSTER,ID:CONSENSUS,EVALUE,SIM_MOTIF,MOTIF_URL) %>% 
  mutate(MOTIF_URL=str_replace(MOTIF_URL,"https://jaspar2024.elixir.no/matrix/","")) %>% 
  mutate(motif_name=case_when(
    str_detect(SIM_MOTIF, "\\(") ~ str_extract(SIM_MOTIF, "(?<=\\().+?(?=\\))"),
    str_detect(SIM_MOTIF, "^MA\\d+\\.\\d+") ~ ALT_ID,
    str_detect(SIM_MOTIF, "^\\d+-") ~ str_replace(SIM_MOTIF, "^\\d+-", ""),
  TRUE ~ SIM_MOTIF)) %>% 
  left_join(twoTE_bind, by=c("ID"="ID","ALT_ID"="ALT_ID", "CONSENSUS"="CONSENSUS")) %>% 
  mutate(diff_per=`TP%` -`FP%`)
```


```{r}
twoTE_Set3_data %>% 
  dplyr::filter(!is.na(MOTIF_URL)) %>% 
  group_by(CLUSTER) %>% 
  slice_min(EVALUE, with_ties = FALSE) %>% 
  ungroup() %>% 
  mutate(order= row_number()) %>%
  mutate(motif_name=if_else(motif_name%in% c("SOX10","Sox11","IKZF2"),paste0(motif_name,"_",order),motif_name)) %>% 
    mutate(log10Evalue= log(EVALUE, base = 10)*(-1)) %>% 
  ggplot(., aes (y= reorder(motif_name,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`TP%`*1.5), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~./1.5,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched TF motif")+
  ggtitle(paste("SET3 enrichment of twoTE vs all other twoTE"))

```


```{r}
packageVersion("dplyr")
packageVersion("tidyr")
packageVersion("tidyselect")
.libPaths()
```
