---
title: "Examining motif clusters"
author: "Renee Matthews"
date: "2025-08-25"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r packages}

library(tidyverse)
library(readr)
library(edgeR)
library(ComplexHeatmap)
library(data.table)
library(dplyr)
library(stringr)
library(ggplot2)
library(viridis)
library(DT)
library(kableExtra)
library(genomation)
library(GenomicRanges)
library(ggpubr) ## For customizing figures
library(corrplot) ## For correlation plot
library(ggpmisc)
library(gcplyr)
library(Rsubread)
library(limma)
library(ggrastr)
library(cowplot)
library(smplot2)
library(ggVennDiagram)
library(ggsignif)
library(BiocParallel)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(ChIPseeker)

```


```{r data add}
# sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")
H3K27ac_set1 <- read_delim("data/motif_lists/H3K27ac_set1.txt",delim="\t")
H3K27ac_set2 <- read_delim("data/motif_lists/H3K27ac_set2.txt",delim="\t")
H3K27ac_set3 <- read_delim("data/motif_lists/H3K27ac_set3.txt",delim="\t")

H3K27me3_set1 <- read_delim("data/motif_lists/H3K27me3_set1.txt",delim="\t")
H3K27me3_set2 <- read_delim("data/motif_lists/H3K27me3_set2.txt",delim="\t")

H3K36me3_set1 <- read_delim("data/motif_lists/H3K36me3_set1.txt",delim="\t")
H3K36me3_set2 <- read_delim("data/motif_lists/H3K36me3_set2.txt",delim="\t")

H3K9me3_set1 <- read_delim("data/motif_lists/H3K9me3_set1.txt",delim="\t")
H3K9me3_set2 <- read_delim("data/motif_lists/H3K9me3_set2.txt",delim="\t")
H3K9me3_set3 <- read_delim("data/motif_lists/H3K9me3_set3.txt",delim="\t")

### loading filtered raw counts for visualization
H3K27ac_filt_raw_counts <- read_delim("data/DER_data/H3K27ac_filtered_raw_counts.txt",delim = "\t") %>% column_to_rownames("Peakid")
H3K27me3_filt_raw_counts <- read_delim("data/DER_data/H3K27me3_filtered_raw_counts.txt",delim = "\t")%>% column_to_rownames("Peakid")

H3K36me3_filt_raw_counts <- read_delim("data/DER_data/H3K36me3_filtered_raw_counts.txt",delim = "\t")%>% column_to_rownames("Peakid")
H3K9me3_filt_raw_counts <- read_delim( "data/DER_data/H3K9me3_filtered_raw_counts.txt",delim = "\t")%>% column_to_rownames("Peakid")

all_H3K27ac_regions <- H3K27ac_filt_raw_counts %>% rownames_to_column("Peakid") %>% distinct(Peakid)
all_H3K27me3_regions <- H3K27me3_filt_raw_counts %>% rownames_to_column("Peakid") %>% distinct(Peakid)
all_H3K36me3_regions <- H3K36me3_filt_raw_counts %>% rownames_to_column("Peakid") %>% distinct(Peakid)
all_H3K9me3_regions <- H3K9me3_filt_raw_counts %>% rownames_to_column("Peakid") %>% distinct(Peakid)

H3K27ac_filt_lcpm <- cpm(H3K27ac_filt_raw_counts, log = TRUE)
H3K27me3_filt_lcpm <- cpm(H3K27me3_filt_raw_counts, log = TRUE)
H3K36me3_filt_lcpm <- cpm(H3K36me3_filt_raw_counts, log = TRUE)
H3K9me3_filt_lcpm <- cpm(H3K9me3_filt_raw_counts, log = TRUE)

H3K9me3_toplist <- readRDS( "data/DER_data/H3K9me3_toplist_nooutlier.RDS")
H3K36me3_toplist <- readRDS( "data/DER_data/H3K36me3_toplist_nooutlier.RDS")
H3K27me3_toplist <- readRDS( "data/DER_data/H3K27me3_toplist.RDS")
H3K27ac_toplist <- readRDS( "data/DER_data/H3K27ac_toplist.RDS")

H3K27ac_toptable_list <- bind_rows(H3K27ac_toplist, .id = "group")
H3K27me3_toptable_list <- bind_rows(H3K27me3_toplist, .id = "group")
H3K36me3_toptable_list <- bind_rows(H3K36me3_toplist, .id = "group")
H3K9me3_toptable_list <- bind_rows(H3K9me3_toplist, .id = "group")

```

```{r function code}
standardizeAnnoList <- function(annoList, categories, color_pal = NULL) {
  if (is.null(color_pal)) {
    if (!requireNamespace("RColorBrewer", quietly = TRUE)) {
      stop("Please install RColorBrewer to use the default palette")
    }
    color_pal <- RColorBrewer::brewer.pal(length(categories), "Set3")
    names(color_pal) <- categories
  }

  for (nm in names(annoList)) {
    annoStat <- annoList[[nm]]@annoStat
    
    # pad missing categories
    missing_cats <- setdiff(categories, as.character(annoStat$Feature))
    if (length(missing_cats) > 0) {
      pad <- data.frame(Feature = missing_cats, Frequency = 0)
      annoStat <- rbind(annoStat, pad)
    }
    
    # factor + reorder rows
    annoStat$Feature <- factor(annoStat$Feature, levels = categories)
    annoStat <- annoStat[order(match(annoStat$Feature, categories)), ]
    
    annoList[[nm]]@annoStat <- annoStat
  }
  
  return(list(
    annoList = annoList,
    color_pal = color_pal
  ))
}

```

###  H3K27ac motifs samples

#### Set 1

```{r H3K27ac set1}

examp_H3K27ac_1 <-#slice_sample(H3K27ac_set1,n = 3, replace=FALSE)
    c("chr1:77923740-77924468","chr20:25484622-25485357","chr1:229148017-229149226")
examp_H3K27ac_2 <-#slice_sample(H3K27ac_set2,n = 3, replace=FALSE)
  c("chr5:85451856-85453893","chr10:114271262-114272616", "chr3:63343167-63344341" )
examp_H3K27ac_3 <- #slice_sample(H3K27ac_set3,n = 3, replace=FALSE)
  c("chr8:8664034-8666371","chr3:141152796-141154087","chr21:43701994-43704465")


H3K27ac_filt_lcpm %>% 
  as.data.frame() %>% 
   rownames_to_column("Peakid") %>%
  dplyr::filter(Peakid %in% examp_H3K27ac_1) %>% 
  ###pivot and add additional information from dataframe
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K27ac_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 1 H3K27ac samples")
```

#### Set 2

```{r H3K27ac set2}
H3K27ac_filt_lcpm %>% 
  as.data.frame() %>% 
  dplyr::filter(row.names(H3K27ac_filt_lcpm) %in% examp_H3K27ac_2) %>% 
  ###pivot and add additional information from dataframe
  rownames_to_column("Peakid") %>% 
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K27ac_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 2 H3K27ac samples")
```

#### Set 3

```{r H3K27ac set 3}
H3K27ac_filt_lcpm %>% 
  as.data.frame() %>% 
  dplyr::filter(row.names(H3K27ac_filt_lcpm) %in% examp_H3K27ac_3) %>% 
  ###pivot and add additional information from dataframe
  rownames_to_column("Peakid") %>% 
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K27ac_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 3 H3K27ac samples")
```

Genomic regions

```{r H3K27ac genomic regions}

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

all_H3K27ac_regions_gr <- all_H3K27ac_regions %>% 
    tidyr::separate(., col = "Peakid", into=c("seqnames","range"), sep = ":", remove=FALSE) %>%
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>%
  GRanges()


H3K27ac_set1_gr <- H3K27ac_set1 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>%
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>%
  dplyr::rename("Peakid"=".") %>%
  GRanges()

H3K27ac_set2_gr <- H3K27ac_set2 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>% 
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>% 
  dplyr::rename("Peakid"=".") %>% 
  GRanges()

H3K27ac_set3_gr <- H3K27ac_set3 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>% 
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>% 
  dplyr::rename("Peakid"=".") %>% 
  GRanges()
H3K27ac_list <- list("Set_1"=H3K27ac_set1_gr,"Set_2"=H3K27ac_set2_gr,"Set_3"=H3K27ac_set3_gr,"all_H3K27ac"=all_H3K27ac_regions_gr)


# peakAnnoList_H3K27ac <- lapply(H3K27ac_list, annotatePeak, tssRegion =c(-2000,2000), TxDb=txdb)

# saveRDS(peakAnnoList_H3K27ac, "data/motif_lists/H3K27ac_annotated_peaks.RDS")
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

categories <- c(
  "Distal Intergenic", "Downstream (<=300)", 
  "Other Intron", "1st Intron", 
  "Other Exon", "1st Exon", 
  "3' UTR", "5' UTR", 
  "Promoter (1-2kb)", "Promoter (<=1kb)"
)

# Optional: custom color palette
color_pal <- c(
  "#6a3d9a", "#cab2d6",
  "#ff7f00", "#fdbf6f",
  "#e31a1c", "#fb9a99",
  "#33a02c", "#b2df8a",
  "#1f78b4", "#a6cee3"
)
names(color_pal) <- categories

res <- standardizeAnnoList(peakAnnoList_H3K27ac, categories, color_pal)
peakAnnoList_H3K27ac_std <- res$annoList
color_pal_std <- res$color_pal

# bar plot
plotAnnoBar(peakAnnoList_H3K27ac_std[c(4,1,2,3)]) +
  scale_fill_manual(values=color_pal_std) +
  ggtitle("H3K27ac Feature Distribution")

# pie plots
for (nm in names(peakAnnoList_H3K27ac_std)) {
  plotAnnoPie(peakAnnoList_H3K27ac_std[[nm]], col=color_pal_std)
  title(main = nm, line = -2)
}




```

#### LFC H3K27ac
  Set1
```{r H3K27ac LFC set 1}
# H3K27ac_toptable_list %>%
#   dplyr::filter(genes %in% H3K27ac_set1$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K27ac_24T", "H3K27ac_144R"),
#                                  c("H3K27ac_24T","H3K27ac_24R"),
#                                  c("H3K27ac_24R", "H3K27ac_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K27ac set1")


H3K27ac_toptable_list %>%
  dplyr::filter(genes %in% H3K27ac_set1$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K27ac_24T", "H3K27ac_144R"),
                                 c("H3K27ac_24T","H3K27ac_24R"),
                                 c("H3K27ac_24R", "H3K27ac_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K27ac set1")
```


```{r H3K27ac line plot LFC set 1}
   
  set1_27ac <- H3K27ac_toptable_list %>%
  dplyr::filter(genes %in% H3K27ac_set1$.) %>% 
  dplyr::select(group,genes,logFC) %>%
   mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set1=median(logFC),.groups="drop")



```

set2
```{r H3K27ac LFC set 2}

# H3K27ac_toptable_list %>%
#   dplyr::filter(genes %in% H3K27ac_set2$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K27ac_24T", "H3K27ac_144R"),
#                                  c("H3K27ac_24T","H3K27ac_24R"),
#                                  c("H3K27ac_24R", "H3K27ac_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K27ac set2")


H3K27ac_toptable_list %>%
  dplyr::filter(genes %in% H3K27ac_set2$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K27ac_24T", "H3K27ac_144R"),
                                 c("H3K27ac_24T","H3K27ac_24R"),
                                 c("H3K27ac_24R", "H3K27ac_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K27ac set2")


```

```{r H3K27ac line plot LFC set 2}
   
  set2_27ac <- H3K27ac_toptable_list %>%
  dplyr::filter(genes %in% H3K27ac_set2$.) %>% 
  dplyr::select(group,genes,logFC) %>%
   mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set2=median(logFC),.groups="drop")


```



set3
```{r H3K27ac LFC set 3}

# H3K27ac_toptable_list %>%
#   dplyr::filter(genes %in% H3K27ac_set3$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K27ac_24T", "H3K27ac_144R"),
#                                  c("H3K27ac_24T","H3K27ac_24R"),
#                                  c("H3K27ac_24R", "H3K27ac_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K27ac set3")


H3K27ac_toptable_list %>%
  dplyr::filter(genes %in% H3K27ac_set3$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K27ac_24T", "H3K27ac_144R"),
                                 c("H3K27ac_24T","H3K27ac_24R"),
                                 c("H3K27ac_24R", "H3K27ac_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K27ac set3")


```

```{r H3K27ac line plot LFC set 3}
   
  H3K27ac_toptable_list %>%
  dplyr::filter(genes %in% H3K27ac_set3$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set3=median(logFC),.groups="drop") %>% 
  left_join(set1_27ac, by="group") %>% 
  left_join(set2_27ac, by="group") %>% 
  pivot_longer(cols=!group, values_to = "median_LFC", names_to = "set") %>% 
  ggplot(., aes(x=group, y=median_LFC, group=set, color=set))+
  geom_point()+
  geom_line(aes(alpha = 0.8))+
  theme_bw()+
  ggtitle("LFC across time for H3K27ac sets")


```

##### Top 100

This is the calculation of Top 100 in each category, then united
```{r H3K27ac top100}
cormotif_initial_H3K27ac <- readRDS("data/Cormotif_data/Cormotif_initial_H3K27ac.RDS")
row.names(cormotif_initial_H3K27ac$bestmotif$clustlike) <- row.names(H3K27ac_filt_lcpm)
H3K27ac_clustlike <- cormotif_initial_H3K27ac$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  rownames_to_column("Peakid")

H3K27ac_clusters <- list(
  Set1="V1",
  Set2="V2",
  Set3="V3")


Sets_H3K27ac <- H3K27ac_toptable_list %>% 
  mutate(cluster=case_when(genes %in% H3K27ac_set1$. ~ "Set1",
                           genes %in% H3K27ac_set2$. ~ "Set2",
                           genes %in% H3K27ac_set3$. ~ "Set3",
                           TRUE ~ "not_assigned")) %>% 
  left_join(., H3K27ac_clustlike,by=c("genes"="Peakid")) %>% 
  dplyr::filter(cluster!="not_assigned")

###This code uses the H3K27ac_clusters list and filters through Sets_H3K27ac data frame by selecting
Sets_H3K27ac_top <- map_dfr(names(H3K27ac_clusters), function(cl) {
  col <- H3K27ac_clusters[[cl]]
  Sets_H3K27ac %>%
    filter(cluster == cl) %>%
    # group_by(group) %>% 
    slice_max(order_by = .data[[col]], n = 100)
}) %>%
  ungroup()

## now plot

Sets_H3K27ac_top %>% 
  mutate(group=factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R")),
         cluster=factor(cluster, levels=c("Set1","Set2","Set3"))) %>% 
  ggplot(., aes(x= group,y = logFC, group = interaction(cluster, genes), color=cluster))+
  geom_point()+
  geom_line(aes(alpha = 0.8))+
  theme_bw()+
  facet_wrap(~cluster)+
  ggtitle("LFC across time for H3K27ac top 100 by Set")+
  theme(axis.text.x = element_text(angle=90) )

# saveRDS(Sets_H3K27ac_top,"data/motif_lists/H3K27ac_toplistbymotif.RDS")
```



###  H3K27me3 motifs samples

#### Set 1

```{r H3K27me3 set1}
# slice_sample(H3K27me3_set2,n = 3, replace=FALSE)
examp_H3K27me3_1 <-#slice_sample(H3K27me3_set1,n = 3, replace=FALSE)
  c("chr17:8878023-8878923", "chr2:23436468-23436797","chr10:132830096-132830712")
examp_H3K27me3_2 <-#slice_sample(H3K27me3_set2,n = 3, replace=FALSE)
  c("chr5:141004122-141004627" ,  "chr6:101195726-101195968",  "chr3:139117155-139117784")



H3K27me3_filt_lcpm %>% 
  as.data.frame() %>% 
  dplyr::filter(row.names(H3K27me3_filt_lcpm) %in% examp_H3K27me3_1) %>% 
  ###pivot and add additional information from dataframe
  rownames_to_column("Peakid") %>% 
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K27me3_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 1 H3K27me3 samples")
```

#### Set 2

```{r H3K27me3 set2}
H3K27me3_filt_lcpm %>% 
  as.data.frame() %>% 
  dplyr::filter(row.names(H3K27me3_filt_lcpm) %in% examp_H3K27me3_2) %>% 
  ###pivot and add additional information from dataframe
  rownames_to_column("Peakid") %>% 
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K27me3_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 2 H3K27me3 samples")
```
```{r H3K27me3 genomic regions}

all_H3K27me3_regions_gr <- all_H3K27me3_regions %>% 
    tidyr::separate(., col = "Peakid", into=c("seqnames","range"), sep = ":", remove=FALSE) %>%
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>%
  GRanges()

H3K27me3_set1_gr <- H3K27me3_set1 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>% 
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>% 
  dplyr::rename("Peakid"=".") %>% 
  GRanges()

H3K27me3_set2_gr <- H3K27me3_set2 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>% 
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>% 
  dplyr::rename("Peakid"=".") %>% 
  GRanges()


H3K27me3_list <- list("Set_1"=H3K27me3_set1_gr,"Set_2"=H3K27me3_set2_gr,"all_H3K27me3_regions"=all_H3K27me3_regions_gr)


# peakAnnoList_H3K27me3 <- lapply(H3K27me3_list, annotatePeak, tssRegion =c(-2000,2000), TxDb=txdb)
# 
# saveRDS(peakAnnoList_H3K27me3, "data/motif_lists/H3K27me3_annotated_peaks.RDS")
peakAnnoList_H3K27me3 <- readRDS("data/motif_lists/H3K27me3_annotated_peaks.RDS")

plotAnnoBar(peakAnnoList_H3K27me3[c(3,1,2)])+
  ggtitle("H3K27me3 Feature Distribution")


# pie_plots_H3K27me3 <-   map(peakAnnoList_H3K27me3, ~ plotAnnoPie(.x))
for(nm in names(peakAnnoList_H3K27me3)) {
  plotAnnoPie(peakAnnoList_H3K27me3[[nm]])
  title(main = nm)  # base R title
}
```

#### LFC H3K27me3
  Set1
```{r H3K27me3 LFC set 1}
# H3K27me3_toptable_list %>%
#   dplyr::filter(genes %in% H3K27me3_set1$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K27me3_24T","H3K27me3_24R","H3K27me3_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K27me3_24T", "H3K27me3_144R"),
#                                  c("H3K27me3_24T","H3K27me3_24R"),
#                                  c("H3K27me3_24R", "H3K27me3_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K27me3 set1")


H3K27me3_toptable_list %>%
  dplyr::filter(genes %in% H3K27me3_set1$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K27me3_24T","H3K27me3_24R","H3K27me3_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K27me3_24T", "H3K27me3_144R"),
                                 c("H3K27me3_24T","H3K27me3_24R"),
                                 c("H3K27me3_24R", "H3K27me3_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K27me3 set1")
```


```{r H3K27me line plot LFC set 1}
   
  set1_27me3 <- H3K27me3_toptable_list %>%
  dplyr::filter(genes %in% H3K27me3_set1$.) %>% 
   dplyr::select(group,genes,logFC) %>%
   mutate(group=factor(group, levels = c("H3K27me3_24T","H3K27me3_24R","H3K27me3_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set1=median(logFC),.groups="drop")

```

set2
```{r H3K27me3 LFC set 2}

# H3K27me3_toptable_list %>%
#   dplyr::filter(genes %in% H3K27me3_set2$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K27me3_24T","H3K27me3_24R","H3K27me3_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K27me3_24T", "H3K27me3_144R"),
#                                  c("H3K27me3_24T","H3K27me3_24R"),
#                                  c("H3K27me3_24R", "H3K27me3_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K27me3 set2")


H3K27me3_toptable_list %>%
  dplyr::filter(genes %in% H3K27me3_set2$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K27me3_24T","H3K27me3_24R","H3K27me3_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K27me3_24T", "H3K27me3_144R"),
                                 c("H3K27me3_24T","H3K27me3_24R"),
                                 c("H3K27me3_24R", "H3K27me3_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K27me3 set2")


```

```{r H3K27me3 line plot LFC set 2}
   
  H3K27me3_toptable_list %>%
  dplyr::filter(genes %in% H3K27me3_set2$.) %>% 
  dplyr::select(group,genes,logFC) %>%
   mutate(group=factor(group, levels = c("H3K27me3_24T","H3K27me3_24R","H3K27me3_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set2=median(logFC),.groups="drop") %>% 
  left_join(set1_27me3,by = "group") %>% 
  pivot_longer(cols=!group, values_to = "median_LFC", names_to = "set") %>% 
  ggplot(., aes(x=group, y=median_LFC, group=set, color=set))+
  geom_point()+
  geom_line(aes(alpha = 0.8))+
  theme_bw()+
  ggtitle("LFC across time for H3K27me3 sets")


```

##### Top 100

This is the calculation of Top 100 in each category, then united
```{r H3K27me3 top100}
cormotif_initial_H3K27me3 <- readRDS("data/Cormotif_data/Cormotif_initial_H3K27me3.RDS")
row.names(cormotif_initial_H3K27me3$bestmotif$clustlike) <- row.names(H3K27me3_filt_lcpm)
H3K27me3_clustlike <- cormotif_initial_H3K27me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  rownames_to_column("Peakid")

H3K27me3_clusters <- list(
  Set1="V1",
  Set2="V2")


Sets_H3K27me3 <- H3K27me3_toptable_list %>% 
  mutate(cluster=case_when(genes %in% H3K27me3_set1$. ~ "Set1",
                           genes %in% H3K27me3_set2$. ~ "Set2",
                           TRUE ~ "not_assigned")) %>% 
  left_join(., H3K27me3_clustlike,by=c("genes"="Peakid")) %>% 
  dplyr::filter(cluster!="not_assigned")

###This code uses the H3K27me3_clusters list and filters through Sets_H3K27me3 data frame by selecting
Sets_H3K27me3_top <- map_dfr(names(H3K27me3_clusters), function(cl) {
  col <- H3K27me3_clusters[[cl]]
  Sets_H3K27me3 %>%
    filter(cluster == cl) %>%
    # group_by(group) %>% 
    slice_max(order_by = .data[[col]], n = 100)
}) %>%
  ungroup()

## now plot

Sets_H3K27me3_top %>% 
  mutate(group=factor(group, levels = c("H3K27me3_24T","H3K27me3_24R","H3K27me3_144R")),
         cluster=factor(cluster, levels=c("Set1","Set2"))) %>% 
  ggplot(., aes(x= group,y = logFC, group = interaction(cluster, genes), color=cluster))+
  geom_point()+
  geom_line(aes(alpha = 0.8))+
  theme_bw()+
  facet_wrap(~cluster)+
  ggtitle("LFC across time for H3K27me3 top 100 by Set")+
  theme(axis.text.x = element_text(angle=90) )

# saveRDS(Sets_H3K27me3_top,"data/motif_lists/H3K27me3_toplistbymotif.RDS")
```





###  H3K36me3 motifs samples

#### Set 1

```{r H3K36me3 set1}
# slice_sample(H3K36me3_set1,n = 3, replace=FALSE)
examp_H3K36me3_1 <-#slice_sample(H3K36me3_set1,n = 3, replace=FALSE)
  c("chr6:106975625-106975872","chr17:17259340-17260414","chr14:91506581-91506995")
examp_H3K36me3_2 <-#slice_sample(H3K36me3_set2,n = 3, replace=FALSE)
  # c("chr2:191013033-191013351","chr5:66597066-66597500","chr16:22232346-22233214")
c("chr17:3656403-3665271","chr6:118573949-118574870","chr1:164799555-164800172")



H3K36me3_filt_lcpm %>% 
  as.data.frame() %>% 
  dplyr::filter(row.names(H3K36me3_filt_lcpm) %in% examp_H3K36me3_1) %>% 
  ###pivot and add additional information from dataframe
  rownames_to_column("Peakid") %>% 
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K36me3_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 1 H3K36me3 samples")
```

#### Set 2

```{r H3K36me3 set2}
H3K36me3_filt_lcpm %>% 
  as.data.frame() %>% 
  dplyr::filter(row.names(H3K36me3_filt_lcpm) %in% examp_H3K36me3_2) %>% 
  ###pivot and add additional information from dataframe
  rownames_to_column("Peakid") %>% 
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K36me3_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 2 H3K36me3 samples")
```
```{r H3K36me3 genomic regions}


all_H3K36me3_regions_gr <- all_H3K36me3_regions %>% 
    tidyr::separate(., col = "Peakid", into=c("seqnames","range"), sep = ":", remove=FALSE) %>%
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>%
  GRanges()

H3K36me3_set1_gr <- H3K36me3_set1 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>% 
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>% 
  dplyr::rename("Peakid"=".") %>% 
  GRanges()

H3K36me3_set2_gr <- H3K36me3_set2 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>% 
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>% 
  dplyr::rename("Peakid"=".") %>% 
  GRanges()


H3K36me3_list <- list("Set_1"=H3K36me3_set1_gr,"Set_2"=H3K36me3_set2_gr, "all_H3K36me3_regions"=all_H3K36me3_regions_gr)


# peakAnnoList_H3K36me3 <- lapply(H3K36me3_list, annotatePeak, tssRegion =c(-2000,2000), TxDb=txdb)
# # 
# saveRDS(peakAnnoList_H3K36me3, "data/motif_lists/H3K36me3_annotated_peaks.RDS")
peakAnnoList_H3K36me3 <- readRDS("data/motif_lists/H3K36me3_annotated_peaks.RDS")

plotAnnoBar(peakAnnoList_H3K36me3[c(3,1,2)])+
  ggtitle("H3K36me3 Feature Distribution")
# pie_plots_H3K36me3 <-   map(peakAnnoList_H3K36me3, ~ plotAnnoPie(.x))
for(nm in names(peakAnnoList_H3K36me3)) {
  plotAnnoPie(peakAnnoList_H3K36me3[[nm]])
  title(main = nm)  # base R title
}
```

#### LFC H3K36me3
  Set1
```{r H3K36me3 LFC set 1}
# H3K36me3_toptable_list %>%
#   dplyr::filter(genes %in% H3K36me3_set1$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K36me3_24T","H3K36me3_24R","H3K36me3_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K36me3_24T", "H3K36me3_144R"),
#                                  c("H3K36me3_24T","H3K36me3_24R"),
#                                  c("H3K36me3_24R", "H3K36me3_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K36me3 set1")


H3K36me3_toptable_list %>%
  dplyr::filter(genes %in% H3K36me3_set1$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K36me3_24T","H3K36me3_24R","H3K36me3_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K36me3_24T", "H3K36me3_144R"),
                                 c("H3K36me3_24T","H3K36me3_24R"),
                                 c("H3K36me3_24R", "H3K36me3_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K36me3 set1")
```


```{r H3K36me3 line plot LFC set 1}
   
 set1_36me3 <- H3K36me3_toptable_list %>%
  dplyr::filter(genes %in% H3K36me3_set1$.) %>% 
  dplyr::select(group,genes,logFC) %>%
   mutate(group=factor(group, levels = c("H3K36me3_24T","H3K36me3_24R","H3K36me3_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set1=median(logFC),.groups="drop") 

```

set2
```{r H3K36me3 LFC set 2}
# 
# H3K36me3_toptable_list %>%
#   dplyr::filter(genes %in% H3K36me3_set2$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K36me3_24T","H3K36me3_24R","H3K36me3_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K36me3_24T", "H3K36me3_144R"),
#                                  c("H3K36me3_24T","H3K36me3_24R"),
#                                  c("H3K36me3_24R", "H3K36me3_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K36me3 set2")


H3K36me3_toptable_list %>%
  dplyr::filter(genes %in% H3K36me3_set2$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K36me3_24T","H3K36me3_24R","H3K36me3_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K36me3_24T", "H3K36me3_144R"),
                                 c("H3K36me3_24T","H3K36me3_24R"),
                                 c("H3K36me3_24R", "H3K36me3_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K36me3 set2")


```

```{r H3K36me3 line plot LFC set 2}
   
 H3K36me3_toptable_list %>%
  dplyr::filter(genes %in% H3K36me3_set2$.) %>% 
  dplyr::select(group,genes,logFC) %>%
   mutate(group=factor(group, levels = c("H3K36me3_24T","H3K36me3_24R","H3K36me3_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set2=median(logFC),.groups="drop") %>% 
  left_join(set1_36me3,by = "group") %>% 
  pivot_longer(cols=!group, values_to = "median_LFC", names_to = "set") %>% 
  ggplot(., aes(x=group, y=median_LFC, group=set, color=set))+
  geom_point()+
  geom_line(aes(alpha = 0.8))+
  theme_bw()+
  ggtitle("LFC across time for H3K36me3 sets")

```
##### Top 100

This is the calculation of Top 100 in each category, then united
```{r H3K36me3 top100}
cormotif_initial_H3K36me3 <- readRDS("data/Cormotif_data/Cormotif_initial_H3K36me3.RDS")
row.names(cormotif_initial_H3K36me3$bestmotif$clustlike) <- row.names(H3K36me3_filt_lcpm)
H3K36me3_clustlike <- cormotif_initial_H3K36me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  rownames_to_column("Peakid")

H3K36me3_clusters <- list(
  Set1="V1",
  Set2="V2")


Sets_H3K36me3 <- H3K36me3_toptable_list %>% 
  mutate(cluster=case_when(genes %in% H3K36me3_set1$. ~ "Set1",
                           genes %in% H3K36me3_set2$. ~ "Set2",
                           TRUE ~ "not_assigned")) %>% 
  left_join(., H3K36me3_clustlike,by=c("genes"="Peakid")) %>% 
  dplyr::filter(cluster!="not_assigned")

###This code uses the H3K36me3_clusters list and filters through Sets_H3K36me3 data frame by selecting
Sets_H3K36me3_top <- map_dfr(names(H3K36me3_clusters), function(cl) {
  col <- H3K36me3_clusters[[cl]]
  Sets_H3K36me3 %>%
    filter(cluster == cl) %>%
    # group_by(group) %>% 
    slice_max(order_by = .data[[col]], n = 100)
}) %>%
  ungroup()

## now plot

Sets_H3K36me3_top %>% 
  mutate(group=factor(group, levels = c("H3K36me3_24T","H3K36me3_24R","H3K36me3_144R")),
         cluster=factor(cluster, levels=c("Set1","Set2"))) %>% 
  ggplot(., aes(x= group,y = logFC, group = interaction(cluster, genes), color=cluster))+
  geom_point()+
  geom_line(aes(alpha = 0.8))+
  theme_bw()+
  facet_wrap(~cluster)+
  ggtitle("LFC across time for H3K36me3 top 100 by Set")+
  theme(axis.text.x = element_text(angle=90) )

# saveRDS(Sets_H3K36me3_top,"data/motif_lists/H3K36me3_toplistbymotif.RDS")
```







###  H3K9me3 motifs samples

#### Set 1

```{r H3K9me3 set1}
# slice_sample(H3K9me3_set3,n = 3, replace=FALSE)
examp_H3K9me3_1 <-#slice_sample(H3K9me3_set1,n = 3, replace=FALSE)
  c("chr7:101406339-101407813","chr6:147512995-147514238","chr10:66432792-66433011")
examp_H3K9me3_2 <-#slice_sample(H3K9me3_set2,n = 3, replace=FALSE)
  c("chr2:120985518-120985865" ,   "chr7:926998-927207","chr12:108895126-108895841")

examp_H3K9me3_3 <- #slice_sample(H3K9me3_set3,n = 3, replace=FALSE)
  c("chr6:44276324-44278629","chr7:906806-907939"  ,"chr20:63682374-63682947")



H3K9me3_filt_lcpm %>% 
  as.data.frame() %>% 
  dplyr::filter(row.names(H3K9me3_filt_lcpm) %in% examp_H3K9me3_1) %>% 
  ###pivot and add additional information from dataframe
  rownames_to_column("Peakid") %>% 
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K9me3_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 1 H3K9me3 samples")
```

#### Set 2

```{r H3K9me3 set2}

H3K9me3_filt_lcpm %>% 
  as.data.frame() %>% 
  dplyr::filter(row.names(H3K9me3_filt_lcpm) %in% examp_H3K9me3_2) %>% 
  ###pivot and add additional information from dataframe
  rownames_to_column("Peakid") %>% 
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K9me3_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 2 H3K9me3 samples")
```

#### Set 3

```{r H3K9me3 set 3}
H3K9me3_filt_lcpm %>% 
  as.data.frame() %>% 
  dplyr::filter(row.names(H3K9me3_filt_lcpm) %in% examp_H3K9me3_3) %>% 
  ###pivot and add additional information from dataframe
  rownames_to_column("Peakid") %>% 
  pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
  separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("DOX","VEH")),
         time=factor(time, levels=c("24T","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K9me3_",time),
      group_graph = factor(group_graph, levels = c(
        "DOX_24T",  "VEH_24T",
        "DOX_24R",  "VEH_24R",
        "DOX_144R",  "VEH_144R"))) %>% 
  
  ggplot(., aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle("set 3 H3K9me3 samples")
```


```{r H3K9me3 genomic regions}

all_H3K9me3_regions_gr <- all_H3K9me3_regions %>% 
    tidyr::separate(., col = "Peakid", into=c("seqnames","range"), sep = ":", remove=FALSE) %>%
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>%
  GRanges()


H3K9me3_set1_gr <- H3K9me3_set1 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>% 
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>% 
  dplyr::rename("Peakid"=".") %>% 
  GRanges()

H3K9me3_set2_gr <- H3K9me3_set2 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>% 
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>% 
  dplyr::rename("Peakid"=".") %>% 
  GRanges()

H3K9me3_set3_gr <- H3K9me3_set3 %>% 
  tidyr::separate(., col = ".", into=c("seqnames","range"), sep = ":", remove=FALSE) %>% 
  tidyr::separate(., col = "range", into=c("start","end"), sep = "-") %>% 
  dplyr::rename("Peakid"=".") %>% 
  GRanges()
H3K9me3_list <- list("Set_1"=H3K9me3_set1_gr,"Set_2"=H3K9me3_set2_gr,"Set_3"=H3K9me3_set3_gr,"all_H3K9me3_regions"=all_H3K9me3_regions_gr)


# peakAnnoList_H3K9me3 <- lapply(H3K9me3_list, annotatePeak, tssRegion =c(-2000,2000), TxDb=txdb)
# # 
# saveRDS(peakAnnoList_H3K9me3, "data/motif_lists/H3K9me3_annotated_peaks.RDS")
peakAnnoList_H3K9me3 <- readRDS("data/motif_lists/H3K9me3_annotated_peaks.RDS")

plotAnnoBar(peakAnnoList_H3K9me3[c(4,1,2,3)])+
  ggtitle("H3K9me3 Feature Distribution")
# pie_plots_H3K9me3 <-   map(peakAnnoList_H3K9me3, ~ plotAnnoPie(.x))
for(nm in names(peakAnnoList_H3K9me3)) {
  plotAnnoPie(peakAnnoList_H3K9me3[[nm]])
  title(main = nm)  # base R title
}
```


#### LFC H3K9me3
  Set1
```{r H3K9me3 LFC set 1}
# H3K9me3_toptable_list %>%
#   dplyr::filter(genes %in% H3K9me3_set1$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K9me3_24T", "H3K9me3_144R"),
#                                  c("H3K9me3_24T","H3K9me3_24R"),
#                                  c("H3K9me3_24R", "H3K9me3_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K9me3 set1")


H3K9me3_toptable_list %>%
  dplyr::filter(genes %in% H3K9me3_set1$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K9me3_24T", "H3K9me3_144R"),
                                 c("H3K9me3_24T","H3K9me3_24R"),
                                 c("H3K9me3_24R", "H3K9me3_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K9me3 set1")
```


```{r H3K9me3 line plot LFC set 1}
   
  set1_9me3 <- H3K9me3_toptable_list %>%
  dplyr::filter(genes %in% H3K9me3_set1$.) %>% 
  dplyr::select(group,genes,logFC) %>%
   mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set1=median(logFC),.groups="drop")




```

set2
```{r H3K9me3 LFC set 2}

# H3K9me3_toptable_list %>%
#   dplyr::filter(genes %in% H3K9me3_set2$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K9me3_24T", "H3K9me3_144R"),
#                                  c("H3K9me3_24T","H3K9me3_24R"),
#                                  c("H3K9me3_24R", "H3K9me3_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K9me3 set2")


H3K9me3_toptable_list %>%
  dplyr::filter(genes %in% H3K9me3_set2$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K9me3_24T", "H3K9me3_144R"),
                                 c("H3K9me3_24T","H3K9me3_24R"),
                                 c("H3K9me3_24R", "H3K9me3_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K9me3 set2")


```

```{r H3K9me3 line plot LFC set 2}
   
  set2_9me3 <- H3K9me3_toptable_list %>%
  dplyr::filter(genes %in% H3K9me3_set2$.) %>% 
  dplyr::select(group,genes,logFC) %>%
   mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set2=median(logFC),.groups="drop")



```



set3
```{r H3K9me3 LFC set 3}

# H3K9me3_toptable_list %>%
#   dplyr::filter(genes %in% H3K9me3_set3$.) %>% 
#   dplyr::select(group,genes,logFC) %>%
#   mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R"))) %>% 
#   ggplot(., aes(x=group, y=logFC))+
#   geom_boxplot(aes(fill=group))+
#   geom_signif(comparisons = list(c("H3K9me3_24T", "H3K9me3_144R"),
#                                  c("H3K9me3_24T","H3K9me3_24R"),
#                                  c("H3K9me3_24R", "H3K9me3_144R")),
#                               step_increase = 0.1, 
#               map_signif_level = FALSE, 
#               test = "wilcox.test")+
#   theme_bw()+
#   ggtitle("LFC across time for H3K9me3 set3")


H3K9me3_toptable_list %>%
  dplyr::filter(genes %in% H3K9me3_set3$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R")),
         logFC=abs(logFC)) %>% 
  ggplot(., aes(x=group, y=logFC))+
  geom_boxplot(aes(fill=group))+
  geom_signif(comparisons = list(c("H3K9me3_24T", "H3K9me3_144R"),
                                 c("H3K9me3_24T","H3K9me3_24R"),
                                 c("H3K9me3_24R", "H3K9me3_144R")),
                              step_increase = 0.1, 
              map_signif_level = FALSE, 
              test = "wilcox.test")+
  theme_bw()+
  ggtitle("Absolute LFC across time for H3K9me3 set3")


```

```{r H3K9me3 line plot LFC set 3}
   
  H3K9me3_toptable_list %>%
  dplyr::filter(genes %in% H3K9me3_set3$.) %>% 
  dplyr::select(group,genes,logFC) %>%
  mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R")),
         logFC=abs(logFC)) %>% 
  group_by(group) %>% 
  summarize(med_Set3=median(logFC),.groups="drop") %>% 
  left_join(set1_9me3, by="group") %>% 
  left_join(set2_9me3, by="group") %>% 
  pivot_longer(cols=!group, values_to = "median_LFC", names_to = "set") %>% 
  ggplot(., aes(x=group, y=median_LFC, group=set, color=set))+
  geom_point()+
  geom_line(aes(alpha = 0.8))+
  theme_bw()+
  ggtitle("LFC across time for H3K9me3 sets")



```

##### Top 100

This is the calculation of Top 100 in each category, then united
```{r H3K9me3 top100}
cormotif_initial_H3K9me3 <- readRDS("data/Cormotif_data/Cormotif_initial_H3K9me3.RDS")
row.names(cormotif_initial_H3K9me3$bestmotif$clustlike) <- row.names(H3K9me3_filt_lcpm)
H3K9me3_clustlike <- cormotif_initial_H3K9me3$bestmotif$clustlike %>% 
  as.data.frame() %>% 
  rownames_to_column("Peakid")

H3K9me3_clusters <- list(
  Set1="V1",
  Set2="V2",
  Set3="V3")


Sets_H3K9me3 <- H3K9me3_toptable_list %>% 
  mutate(cluster=case_when(genes %in% H3K9me3_set1$. ~ "Set1",
                           genes %in% H3K9me3_set2$. ~ "Set2",
                           genes %in% H3K9me3_set3$. ~ "Set3",
                           TRUE ~ "not_assigned")) %>% 
  left_join(., H3K9me3_clustlike,by=c("genes"="Peakid")) %>% 
  dplyr::filter(cluster!="not_assigned")

###This code uses the H3K9me3_clusters list and filters through Sets_H3K9me3 data frame by selecting
Sets_H3K9me3_top <- map_dfr(names(H3K9me3_clusters), function(cl) {
  col <- H3K9me3_clusters[[cl]]
  Sets_H3K9me3 %>%
    filter(cluster == cl) %>%
    # group_by(group) %>% 
    slice_max(order_by = .data[[col]], n = 100)
}) %>%
  ungroup()

## now plot

Sets_H3K9me3_top %>% 
  mutate(group=factor(group, levels = c("H3K9me3_24T","H3K9me3_24R","H3K9me3_144R")),
         cluster=factor(cluster, levels=c("Set1","Set2","Set3"))) %>% 
  ggplot(., aes(x= group,y = logFC, group = interaction(cluster, genes), color=cluster))+
  geom_point()+
  geom_line(aes(alpha = 0.8))+
  theme_bw()+
  facet_wrap(~cluster)+
  ggtitle("LFC across time for H3K9me3 top 100 by Set")+
  theme(axis.text.x = element_text(angle=90) )

# saveRDS(Sets_H3K9me3_top,"data/motif_lists/H3K9me3_toplistbymotif.RDS")
```


