---
title: "Outlier removal"
author: "Renee Matthews edits"
date: "2025-08-21"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

## Final Anaylsis

#### Loading Packages 
```{r load packages}
library(tidyverse)
library(readr)
library(edgeR)
library(ComplexHeatmap)
library(data.table)
library(dplyr)
library(stringr)
library(ggplot2)
library(viridis)
library(DT)
library(kableExtra)
library(genomation)
library(GenomicRanges)
library(chromVAR) ## For FRiP analysis and differential analysis
library(DESeq2) ## For differential analysis section
library(ggpubr) ## For customizing figures
library(corrplot) ## For correlation plot
library(ggpmisc)
library(gcplyr)
library(Rsubread)
library(limma)
library(ggrastr)
library(cowplot)
library(smplot2)
library(ggVennDiagram)
```

#### Data Initialization
```{r init}
sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")
```

#### Functions
```{r functions}
drug_pal <- c("#8B006D","#DF707E","#F1B72B", "#3386DD","#707031","#41B333")
pca_plot <- function(pca_obj, df,
                     col_var = NULL,
                     shape_var = NULL,
                     text_var = NULL,
                     title = "") {
  
  # variance explained
  a <- prop_var_percent(pca_obj)
  
  ggplot(df, aes_string(x = "PC1", y = "PC2")) +
    geom_point(aes_string(color = col_var, shape = shape_var), size = 5) +
    ggrepel::geom_text_repel(aes_string(label = text_var),
                             vjust = -.5,
                             max.overlaps = 30) +
    labs(
      title = title,
      x = paste0("PC1 (", round(a[1], 1), "%)"),
      y = paste0("PC2 (", round(a[2], 1), "%)")
    ) +
    scale_color_manual(values = c(
      "#8B006D", "#DF707E", "#F1B72B",
      "#3386DD", "#707031", "#41B333"
    ))
}

pca_var_plot <- function(pca) {
  # x: class == prcomp
  pca.var <- pca$sdev ^ 2
  pca.prop <- pca.var / sum(pca.var)
  var.plot <-
    qplot(PC, prop, data = data.frame(PC = 1:length(pca.prop),
                                      prop = pca.prop)) +
    labs(title = 'Variance contributed by each PC',
         x = 'PC', y = 'Proportion of variance')
  plot(var.plot)
}

calc_pca <- function(x) {
  # Performs principal components analysis with prcomp
  # x: a sample-by-gene numeric matrix
  prcomp(x, scale. = TRUE, retx = TRUE)
}

get_regr_pval <- function(mod) {
  # Returns the p-value for the Fstatistic of a linear model
  # mod: class lm
  stopifnot(class(mod) == "lm")
  fstat <- summary(mod)$fstatistic
  pval <- 1 - pf(fstat[1], fstat[2], fstat[3])
  return(pval)
}

prop_var_percent <- function(pca_result){ 
  # Ensure the input is a PCA result object
  if (!inherits(pca_result, "prcomp")) {
    stop("Input must be a result from prcomp()")
  }
  
  # Get the standard deviations from the PCA result
  sdev <- pca_result$sdev
  
  # Calculate the proportion of variance
  proportion_variance <- (sdev^2) / sum(sdev^2)*100
  
  return(proportion_variance)
}

plot_versus_pc <- function(df, pc_num, fac) {
  # df: data.frame
  # pc_num: numeric, specific PC for plotting
  # fac: column name of df for plotting against PC
  pc_char <- paste0("PC", pc_num)
  # Calculate F-statistic p-value for linear model
  pval <- get_regr_pval(lm(df[, pc_char] ~ df[, fac]))
  if (is.numeric(df[, f])) {
    ggplot(df, aes_string(x = f, y = pc_char)) + geom_point() +
      geom_smooth(method = "lm") + labs(title = sprintf("p-val: %.2f", pval))
  } else {
    ggplot(df, aes_string(x = f, y = pc_char)) + geom_boxplot() +
      labs(title = sprintf("p-val: %.2f", pval))
  }
}
x_axis_labels = function(labels, every_nth = 1, ...) {
  axis(side = 1,
       at = seq_along(labels),
       labels = F)
  text(
    x = (seq_along(labels))[seq_len(every_nth) == 1],
    y = par("usr")[3] - 0.075 * (par("usr")[4] - par("usr")[3]),
    labels = labels[seq_len(every_nth) == 1],
    xpd = TRUE,
    ...
  )
}

volcanosig <- function(df, psig.lvl) {
    df <- df %>% 
    mutate(threshold = ifelse(adj.P.Val > psig.lvl, "A", ifelse(adj.P.Val <= psig.lvl & logFC<=0,"B","C")))
      # ifelse(adj.P.Val <= psig.lvl & logFC >= 0,"B", "C")))
    ##This is where I could add labels, but I have taken out
    # df <- df %>% mutate(genelabels = "")
    # df$genelabels[1:topg] <- df$rownames[1:topg]
    
  ggplot(df, aes(x=logFC, y=-log10(P.Value))) + 
    ggrastr::geom_point_rast(aes(color=threshold))+
    # geom_text_repel(aes(label = genelabels), segment.curvature = -1e-20,force = 1,size=2.5,
    # arrow = arrow(length = unit(0.015, "npc")), max.overlaps = Inf) +
    #geom_hline(yintercept = -log10(psig.lvl))+
    xlab(expression("Log"[2]*" FC"))+
    ylab(expression("-log"[10]*"P Value"))+
    scale_color_manual(values = c("black", "red","blue"))+
    theme_cowplot()+
    ylim(0,25)+
    xlim(-6,6)+
    theme(legend.position = "none",
              plot.title = element_text(size = rel(1.5), hjust = 0.5),
              axis.title = element_text(size = rel(0.8))) 
}




```

### Feature Counts  

```{r feature_cts}
H3K27ac_merged <- read_delim("data/peaks/H3K27ac_FINAL_counts.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE, skip = 1)
H3K27me3_merged <- read_delim("data/peaks/H3K27me3_FINAL_counts.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE, skip = 1)
H3K36me3_merged <- read_delim("data/peaks/H3K36me3_FINAL_counts.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE, skip = 1)
H3K9me3_merged <- read_delim("data/peaks/H3K9me3_FINAL_counts.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE, skip = 1)
rename_list <- sampleinfo %>% 
  mutate(stem= "_nobl.bam") %>% 
  mutate(prefix=paste0("/scratch/10819/styu/MW_multiQC/peaks/",Histone_Mark,"/",Treatment,"/",Timepoint,"/")) %>%
  mutate(oldname=paste0(prefix,`Library ID`,"/",`Library ID`,stem)) %>% 
  mutate(newname=paste0(Individual,"_",Treatment,"_",Timepoint)) %>% 
  dplyr::select(oldname,newname)
rename_vec <- setNames(rename_list$newname, rename_list$oldname)
names(H3K27ac_merged)[names(H3K27ac_merged) %in% names(rename_vec)] <- rename_vec[names(H3K27ac_merged)[names(H3K27ac_merged) %in% names(rename_vec)]]
names(H3K27me3_merged)[names(H3K27me3_merged) %in% names(rename_vec)] <- rename_vec[names(H3K27me3_merged)[names(H3K27me3_merged) %in% names(rename_vec)]]
names(H3K36me3_merged)[names(H3K36me3_merged) %in% names(rename_vec)] <- rename_vec[names(H3K36me3_merged)[names(H3K36me3_merged) %in% names(rename_vec)]]
names(H3K9me3_merged)[names(H3K9me3_merged) %in% names(rename_vec)] <- rename_vec[names(H3K9me3_merged)[names(H3K9me3_merged) %in% names(rename_vec)]]
```

### Removing outliers and reanalyzing

This is removing outliers in H3K36me3: Ind1_VEH_144R
and in H3K9me3: Ind1_VEH_24T, Ind3_DOX_24T, Ind5_DOX_144R


```{r removing outliers}


H3K36me3_merged <- H3K36me3_merged %>% 
  dplyr::select(!Ind1_VEH_144R)

H3K9me3_merged<- H3K9me3_merged %>%
  dplyr::select(!Ind1_VEH_24T) %>% 
  dplyr::select(!Ind3_DOX_24T) %>%
  dplyr::select(!Ind5_DOX_144R)
```


### H3K27ac Count Analysis
```{r H3K27ac HeatMap 1,eval=FALSE, fig.width=14,fig.height=14}
H3K27ac_merged_raw <- H3K27ac_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  as.matrix()

H3K27ac_merged_lcpm <- H3K27ac_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  cpm(., log = TRUE)
H3K27ac_merged_cor <- H3K27ac_merged_lcpm %>% 
  cor()

annomat <- data.frame(sample=colnames(H3K27ac_merged_cor)) %>% 
  separate_wider_delim(sample,delim="_",names=c("Ind","Treatment","Timepoint"),cols_remove = FALSE) %>% 
  mutate(Treatment=factor(Treatment, levels = c("VEH","5FU","DOX")),
         Timepoint=factor(Timepoint, levels =c("24T","24R","144R"))) %>% 
  column_to_rownames("sample")
heatmap_first <- ComplexHeatmap::HeatmapAnnotation(df = annomat)

Heatmap(H3K27ac_merged_cor, 
        top_annotation = heatmap_first,
        column_title="Filtered log2cpm H3K27ac, no removal")
```

### H3K27me3 Count Analysis
```{r H3K27me3 HeatMap 1,eval=FALSE, fig.width=14,fig.height=14}
H3K27me3_merged_raw <- H3K27me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  as.matrix()

H3K27me3_merged_lcpm <- H3K27me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  cpm(., log = TRUE)
H3K27me3_merged_cor <- H3K27me3_merged_lcpm %>% 
  cor()

annomat <- data.frame(sample=colnames(H3K27me3_merged_cor)) %>% 
  separate_wider_delim(sample,delim="_",names=c("Ind","Treatment","Timepoint"),cols_remove = FALSE) %>% 
  mutate(Treatment=factor(Treatment, levels = c("VEH","5FU","DOX")),
         Timepoint=factor(Timepoint, levels =c("24T","24R","144R"))) %>% 
  column_to_rownames("sample")
heatmap_first <- ComplexHeatmap::HeatmapAnnotation(df = annomat)

Heatmap(H3K27me3_merged_cor, 
        top_annotation = heatmap_first,
        column_title="Filtered log2cpm H3K27me3 with no removal")
```

### H3K36me3 Count Analysis
```{r H3K36me3 HeatMap 1, fig.width=14,fig.height=14}
H3K36me3_merged_raw <- H3K36me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  as.matrix()

H3K36me3_merged_lcpm <- H3K36me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  cpm(., log = TRUE)
H3K36me3_merged_cor <- H3K36me3_merged_lcpm %>% 
  cor()

annomat <- data.frame(sample=colnames(H3K36me3_merged_cor)) %>% 
  separate_wider_delim(sample,delim="_",names=c("Ind","Treatment","Timepoint"),cols_remove = FALSE) %>% 
  mutate(Treatment=factor(Treatment, levels = c("VEH","5FU","DOX")),
         Timepoint=factor(Timepoint, levels =c("24T","24R","144R"))) %>% 
  column_to_rownames("sample")
heatmap_first <- ComplexHeatmap::HeatmapAnnotation(df = annomat)

Heatmap(H3K36me3_merged_cor, 
        top_annotation = heatmap_first,
        column_title="Filtered log2cpm H3K36me3 with removal of 1 sample")
```

### H3K9me3 Count Analysis
```{r H3K9me3 HeatMap 1, fig.width=14,fig.height=14}
H3K9me3_merged_raw <- H3K9me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  as.matrix()

H3K9me3_merged_lcpm <- H3K9me3_merged %>% 
  dplyr::select(Geneid,contains("Ind")) %>% 
  column_to_rownames("Geneid") %>% 
  cpm(., log = TRUE)
H3K9me3_merged_cor <- H3K9me3_merged_lcpm %>% 
  cor()

annomat <- data.frame(sample=colnames(H3K9me3_merged_cor)) %>% 
  separate_wider_delim(sample,delim="_",names=c("Ind","Treatment","Timepoint"),cols_remove = FALSE) %>% 
  mutate(Treatment=factor(Treatment, levels = c("VEH","5FU","DOX")),
         Timepoint=factor(Timepoint, levels =c("24T","24R","144R"))) %>% 
  column_to_rownames("sample")
heatmap_first <- ComplexHeatmap::HeatmapAnnotation(df = annomat)

Heatmap(H3K9me3_merged_cor, 
        top_annotation = heatmap_first,
        column_title="Filtered log2cpm H3K9me3 with removal of 3 outliers")
```

### Differential Analysis

#### Filtering Sex Chromosomes

Removing chrX and chrY
```{r filtering x and Y}
# H3K27ac_merged_raw <- H3K27ac_merged_raw[rowMeans(H3K27ac_merged_lcpm)>0,]
# H3K27ac_merged_raw <- H3K27ac_merged_raw[!grepl("chrY",rownames(H3K27ac_merged_raw)),]
# H3K27ac_merged_raw <- H3K27ac_merged_raw[!grepl("chrX",rownames(H3K27ac_merged_raw)),]

# H3K27me3_merged_raw <- H3K27me3_merged_raw[rowMeans(H3K27me3_merged_lcpm)>0,]
# H3K27me3_merged_raw <- H3K27me3_merged_raw[!grepl("chrY",rownames(H3K27me3_merged_raw)),]
# H3K27me3_merged_raw <- H3K27me3_merged_raw[!grepl("chrX",rownames(H3K27me3_merged_raw)),]


H3K36me3_merged_raw <- H3K36me3_merged_raw[rowMeans(H3K36me3_merged_lcpm)>0,]
H3K36me3_merged_raw <- H3K36me3_merged_raw[!grepl("chrY",rownames(H3K36me3_merged_raw)),]
H3K36me3_merged_raw <- H3K36me3_merged_raw[!grepl("chrX",rownames(H3K36me3_merged_raw)),]

H3K9me3_merged_raw <- H3K9me3_merged_raw[rowMeans(H3K9me3_merged_lcpm)>0,]
H3K9me3_merged_raw <- H3K9me3_merged_raw[!grepl("chrY",rownames(H3K9me3_merged_raw)),]
H3K9me3_merged_raw <- H3K9me3_merged_raw[!grepl("chrX",rownames(H3K9me3_merged_raw)),]
```

```{r H3K36me3 HeatMap 2, fig.width=14,fig.height=14}


H3K36me3_merged_raw_lcpm <- H3K36me3_merged_raw %>% 
  cpm(., log = TRUE)

H3K36me3_merged_filt_cor <- H3K36me3_merged_raw_lcpm %>% 
  cor()

annomat <- data.frame(sample=colnames(H3K36me3_merged_filt_cor)) %>% 
  separate_wider_delim(sample,delim="_",names=c("Ind","Treatment","Timepoint"),cols_remove = FALSE) %>% 
  mutate(Treatment=factor(Treatment, levels = c("VEH","5FU","DOX")),
         Timepoint=factor(Timepoint, levels =c("24T","24R","144R"))) %>% 
  column_to_rownames("sample")
heatmap_second <- ComplexHeatmap::HeatmapAnnotation(df = annomat)

Heatmap(H3K36me3_merged_filt_cor, 
        top_annotation = heatmap_second,
        column_title="Filtered log2cpm H3K36me3 with removal of 1 sample")
```

```{r H3K9me3 HeatMap 2, fig.width=14,fig.height=14}


H3K9me3_merged_raw_lcpm <- H3K9me3_merged_raw %>% 
  cpm(., log = TRUE)

H3K9me3_merged_filt_cor <- H3K9me3_merged_raw_lcpm %>% 
  cor()

annomat <- data.frame(sample=colnames(H3K9me3_merged_filt_cor)) %>% 
  separate_wider_delim(sample,delim="_",names=c("Ind","Treatment","Timepoint"),cols_remove = FALSE) %>% 
  mutate(Treatment=factor(Treatment, levels = c("VEH","5FU","DOX")),
         Timepoint=factor(Timepoint, levels =c("24T","24R","144R"))) %>% 
  column_to_rownames("sample")
heatmap_second <- ComplexHeatmap::HeatmapAnnotation(df = annomat)

Heatmap(H3K9me3_merged_filt_cor, 
        top_annotation = heatmap_second,
        column_title="Filtered log2cpm H3K9me3 with removal of 3 samples")
```

#### Setting up Matrix
```{r matrix}
# H3K27ac_annomat <- data.frame(timeset=colnames(H3K27ac_merged_raw)) %>%
#   mutate(sample=timeset) %>%
#   separate(timeset, into = c("ind","tx","time")) %>%
#   mutate(tx=factor(tx, levels = c("VEH", "DOX")),
#          time=factor(time, levels =c("24T","24R","144R"))) %>%
#   mutate(ind = gsub("Ind", "", ind)) %>%
#   mutate(txtime = paste0(tx, "_", time)) %>%
#   mutate(group = txtime)
# H3K27ac_annomat$group <- H3K27ac_annomat$group %>%
#   gsub("DOX_24T", "1", .) %>%
#   gsub("DOX_24R", "2", .) %>%
#   gsub("DOX_144R", "3", .) %>%
#   gsub("VEH_24T", "4", .) %>%
#   gsub("VEH_24R", "5", .) %>%
#   gsub("VEH_144R", "6", .)
# 
# H3K27me3_annomat <- data.frame(timeset=colnames(H3K27me3_merged_raw)) %>%
#   mutate(sample=timeset) %>%
#   separate(timeset, into = c("ind","tx","time")) %>%
#   mutate(tx=factor(tx, levels = c("VEH", "DOX")),
#          time=factor(time, levels =c("24T","24R","144R"))) %>%
#   mutate(ind = gsub("Ind", "", ind)) %>%
#   mutate(txtime = paste0(tx, "_", time)) %>%
#   mutate(group = txtime)
# H3K27me3_annomat$group <- H3K27me3_annomat$group %>%
#   gsub("DOX_24T", "1", .) %>%
#   gsub("DOX_24R", "2", .) %>%
#   gsub("DOX_144R", "3", .) %>%
#   gsub("VEH_24T", "4", .) %>%
#   gsub("VEH_24R", "5", .) %>%
#   gsub("VEH_144R", "6", .)

H3K36me3_annomat <- data.frame(timeset=colnames(H3K36me3_merged_raw)) %>% 
  mutate(sample=timeset) %>% 
  separate(timeset, into = c("ind","tx","time")) %>% 
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(ind = gsub("Ind", "", ind)) %>%
  mutate(txtime = paste0(tx, "_", time)) %>%
  mutate(group = txtime)
H3K36me3_annomat$group <- H3K36me3_annomat$group %>%
  gsub("DOX_24T", "1", .) %>%
  gsub("DOX_24R", "2", .) %>%
  gsub("DOX_144R", "3", .) %>%
  gsub("VEH_24T", "4", .) %>%
  gsub("VEH_24R", "5", .) %>%
  gsub("VEH_144R", "6", .)

H3K9me3_annomat <- data.frame(timeset=colnames(H3K9me3_merged_raw)) %>% 
  mutate(sample=timeset) %>% 
  separate(timeset, into = c("ind","tx","time")) %>% 
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(ind = gsub("Ind", "", ind)) %>%
  mutate(txtime = paste0(tx, "_", time)) %>%
  mutate(group = txtime)
H3K9me3_annomat$group <- H3K9me3_annomat$group %>%
  gsub("DOX_24T", "1", .) %>%
  gsub("DOX_24R", "2", .) %>%
  gsub("DOX_144R", "3", .) %>%
  gsub("VEH_24T", "4", .) %>%
  gsub("VEH_24R", "5", .) %>%
  gsub("VEH_144R", "6", .)

# dge_H3K27ac <- edgeR::DGEList(counts = H3K27ac_merged_raw, group = H3K27ac_annomat$group, genes = row.names(H3K27ac_merged_raw))
# dge_H3K27me3 <- edgeR::DGEList(counts = H3K27me3_merged_raw, group = H3K27me3_annomat$group, genes = row.names(H3K27me3_merged_raw))
dge_H3K36me3 <- edgeR::DGEList(counts = H3K36me3_merged_raw, group = H3K36me3_annomat$group, genes = row.names(H3K36me3_merged_raw))
dge_H3K9me3 <- edgeR::DGEList(counts = H3K9me3_merged_raw, group = H3K9me3_annomat$group, genes = row.names(H3K9me3_merged_raw))

# dge_H3K27ac <- edgeR::calcNormFactors(dge_H3K27ac)
# dge_H3K27me3 <- edgeR::calcNormFactors(dge_H3K27me3)
dge_H3K36me3 <- edgeR::calcNormFactors(dge_H3K36me3)
dge_H3K9me3 <- edgeR::calcNormFactors(dge_H3K9me3)

# mm_H3K27ac <- model.matrix(~0 + H3K27ac_annomat$txtime)
# colnames(mm_H3K27ac) <- H3K27ac_annomat$txtime %>% unique()

# mm_H3K27me3 <- model.matrix(~0 + H3K27me3_annomat$txtime)
# colnames(mm_H3K27me3) <- H3K27me3_annomat$txtime %>% unique()

mm_H3K36me3 <- model.matrix(~0 + H3K36me3_annomat$txtime)
colnames(mm_H3K36me3) <- H3K36me3_annomat$txtime %>% unique()

mm_H3K9me3 <- model.matrix(~0 + H3K9me3_annomat$txtime)
colnames(mm_H3K9me3) <- H3K9me3_annomat$txtime %>% unique()
```
### PCA Plots

#### H3K27ac
```{r H3K27ac PCA_var,eval=FALSE}
pca_H3K27ac <- calc_pca(t(H3K27ac_merged_lcpm))
pca_var_plot(pca_H3K27ac)
```

```{r H3K27ac PCA,eval=FALSE}
# pca_H3K27ac <- pca_H3K27ac$x %>% cbind(., H3K27ac_annomat)
pca_H3K27ac_df <- data.frame(pca_H3K27ac$x , H3K27ac_annomat)
pca_plot(
  pca_H3K27ac,
  pca_H3K27ac_df,
  col_var = "time",
  shape_var = "tx",
  text_var = "ind",   # <-- string, not vector
  title = "H3K27ac filtered lcpm PCA"
)      +theme_bw()
```

```{r H3K27ac HeatMap 2,eval=FALSE ,fig.width=14,fig.height=14}


H3K27ac_merged_raw_lcpm <- H3K27ac_merged_raw %>% 
  cpm(., log = TRUE)

H3K27ac_merged_filt_cor <- H3K27ac_merged_raw_lcpm %>% 
  cor()

annomat <- data.frame(sample=colnames(H3K27ac_merged_filt_cor)) %>% 
  separate_wider_delim(sample,delim="_",names=c("Ind","Treatment","Timepoint"),cols_remove = FALSE) %>% 
  mutate(Treatment=factor(Treatment, levels = c("VEH","5FU","DOX")),
         Timepoint=factor(Timepoint, levels =c("24T","24R","144R"))) %>% 
  column_to_rownames("sample")
heatmap_second <- ComplexHeatmap::HeatmapAnnotation(df = annomat)

Heatmap(H3K27ac_merged_filt_cor, 
        top_annotation = heatmap_second,
        column_title="Filtered log2cpm H3K27ac with Standard Merging")
```
#### H3K27me3
```{r H3K27me3 PCA_var,eval=FALSE}
pca_H3K27me3 <- calc_pca(t(H3K27me3_merged_lcpm))
pca_var_plot(pca_H3K27me3)
```

```{r H3K27me3 PCA,eval=FALSE}
pca_H3K27me3_df <- data.frame(pca_H3K27me3$x , H3K27me3_annomat)
pca_plot(
  pca_H3K27me3,
  pca_H3K27me3_df,
  col_var = "time",
  shape_var = "tx",
  text_var = "ind",   # <-- string, not vector
  title = "H3K27me3 filtered lcpm PCA"
)   +theme_bw()     
```

#### H3K36me3
```{r H3K36me3 PCA_var}
pca_H3K36me3 <- calc_pca(t(H3K36me3_merged_lcpm))
pca_var_plot(pca_H3K36me3)
```

```{r H3K36me3 PCA}
pca_H3K36me3_df <- data.frame(pca_H3K36me3$x , H3K36me3_annomat)
pca_plot(
  pca_H3K36me3,
  pca_H3K36me3_df,
  col_var = "time",
  shape_var = "tx",
  text_var = "ind",   # <-- string, not vector
  title = "H3K36me3 filtered lcpm PCA"
)    +theme_bw()    
```

#### H3K9me3
```{r H3K9me3 PCA_var}
  
pca_H3K9me3 <- calc_pca(t(H3K9me3_merged_lcpm))
pca_var_plot(pca_H3K9me3)
```

```{r H3K9me3 PCA}
pca_H3K9me3_df <- data.frame(pca_H3K9me3$x , H3K9me3_annomat)
pca_plot(
  pca_H3K9me3,
  pca_H3K9me3_df,
  col_var = "time",
  shape_var = "tx",
  text_var = "ind",   # <-- string, not vector
  title = "H3K9me3 filtered lcpm PCA"
)  +theme_bw()    
```

### Differential Enrichment and Volcano Plots

#### H3K27ac
```{r H3K27ac volcano 1, eval=FALSE}
y <- voom(dge_H3K27ac, mm_H3K27ac, plot = FALSE)
corfit <- duplicateCorrelation(y, mm_H3K27ac, block = H3K27ac_annomat$ind)
v <- voom(dge_H3K27ac, mm_H3K27ac, block = H3K27ac_annomat$ind, correlation = corfit$consensus.correlation)
fit <- lmFit(v, mm_H3K27ac, block = H3K27ac_annomat$ind, correlation = corfit$consensus.correlation)
cm <- makeContrasts(
  DOX_24T.VEH_24T = DOX_24T-VEH_24T,
  DOX_24R.VEH_24R = DOX_24R-VEH_24R,
  DOX_144R.VEH_144R = DOX_144R-VEH_144R,
  levels = mm_H3K27ac)

fit2<- contrasts.fit(fit, contrasts=cm)
efit2 <- eBayes(fit2)

results = decideTests(efit2)

summary(results)

```

```{r H3K27ac volcano 2, eval=FALSE}
plotSA(efit2, main="Mean-Variance trend for final model for H3K27ac")
```

```{r H3K27ac volcano 3, fig.width=14, eval=FALSE}
V.24T.top= topTable(efit2, coef=1, adjust.method="BH", number=Inf, sort.by="p")
V.24R.top= topTable(efit2, coef=2, adjust.method="BH", number=Inf, sort.by="p")
V.144R.top= topTable(efit2, coef=3, adjust.method="BH", number=Inf, sort.by="p")

H3K27ac_24T <- volcanosig(V.24T.top, 0.05)+ ggtitle("DOX 24T")
H3K27ac_24R <- volcanosig(V.24R.top, 0.05)+ ggtitle("DOX 24R")+ylab("")
H3K27ac_144R <- volcanosig(V.144R.top, 0.05)+ ggtitle("DOX 144R")+ylab("")
plot_grid(H3K27ac_24T, H3K27ac_24R, H3K27ac_144R, rel_widths =c(1,1,1))
```

#### H3K27me3
```{r H3K27me3 volcano 1, eval=FALSE}
y <- voom(dge_H3K27me3, mm_H3K27me3, plot = FALSE)
corfit <- duplicateCorrelation(y, mm_H3K27me3, block = H3K27me3_annomat$ind)
v <- voom(dge_H3K27me3, mm_H3K27me3, block = H3K27me3_annomat$ind, correlation = corfit$consensus.correlation)
fit <- lmFit(v, mm_H3K27me3, block = H3K27me3_annomat$ind, correlation = corfit$consensus.correlation)
cm <- makeContrasts(
  DOX_24T.VEH_24T = DOX_24T-VEH_24T,
  DOX_24R.VEH_24R = DOX_24R-VEH_24R,
  DOX_144R.VEH_144R = DOX_144R-VEH_144R,
  levels = mm_H3K27me3)

fit2<- contrasts.fit(fit, contrasts=cm)
efit2 <- eBayes(fit2)

results = decideTests(efit2)

summary(results)

```

```{r H3K27me3 volcano 2, eval=FALSE}
plotSA(efit2, main="Mean-Variance trend for final model for H3K27me3")
```

```{r H3K27me3 volcano 3, fig.width=14, eval=FALSE}
V.24T.top= topTable(efit2, coef=1, adjust.method="BH", number=Inf, sort.by="p")
V.24R.top= topTable(efit2, coef=2, adjust.method="BH", number=Inf, sort.by="p")
V.144R.top= topTable(efit2, coef=3, adjust.method="BH", number=Inf, sort.by="p")

H3K27me3_24T <- volcanosig(V.24T.top, 0.05)+ ggtitle("DOX 24T")
H3K27me3_24R <- volcanosig(V.24R.top, 0.05)+ ggtitle("DOX 24R")+ylab("")
H3K27me3_144R <- volcanosig(V.144R.top, 0.05)+ ggtitle("DOX 144R")+ylab("")
plot_grid(H3K27me3_24T, H3K27me3_24R, H3K27me3_144R, rel_widths =c(1,1,1))
```

#### H3K36me3
```{r H3K36me3 volcano 1}
y <- voom(dge_H3K36me3, mm_H3K36me3, plot = FALSE)
corfit <- duplicateCorrelation(y, mm_H3K36me3, block = H3K36me3_annomat$ind)
v <- voom(dge_H3K36me3, mm_H3K36me3, block = H3K36me3_annomat$ind, correlation = corfit$consensus.correlation)
fit <- lmFit(v, mm_H3K36me3, block = H3K36me3_annomat$ind, correlation = corfit$consensus.correlation)
cm <- makeContrasts(
  DOX_24T.VEH_24T = DOX_24T-VEH_24T,
  DOX_24R.VEH_24R = DOX_24R-VEH_24R,
  DOX_144R.VEH_144R = DOX_144R-VEH_144R,
  levels = mm_H3K36me3)

fit2<- contrasts.fit(fit, contrasts=cm)
efit2 <- eBayes(fit2)

results = decideTests(efit2)

summary(results)

```

```{r H3K36me3 volcano 2}
plotSA(efit2, main="Mean-Variance trend for final model for H3K36me3")
```

```{r H3K36me3 volcano 3, fig.width=14}
V.24T.top= topTable(efit2, coef=1, adjust.method="BH", number=Inf, sort.by="p")
V.24R.top= topTable(efit2, coef=2, adjust.method="BH", number=Inf, sort.by="p")
V.144R.top= topTable(efit2, coef=3, adjust.method="BH", number=Inf, sort.by="p")

H3K36me3_24T <- volcanosig(V.24T.top, 0.05)+ ggtitle("DOX 24T")
H3K36me3_24R <- volcanosig(V.24R.top, 0.05)+ ggtitle("DOX 24R")+ylab("")
H3K36me3_144R <- volcanosig(V.144R.top, 0.05)+ ggtitle("DOX 144R")+ylab("")
plot_grid(H3K36me3_24T, H3K36me3_24R, H3K36me3_144R, rel_widths =c(1,1,1))
```

#### H3K9me3
```{r H3K9me3 volcano 1}
y <- voom(dge_H3K9me3, mm_H3K9me3, plot = FALSE)
corfit <- duplicateCorrelation(y, mm_H3K9me3, block = H3K9me3_annomat$ind)
v <- voom(dge_H3K9me3, mm_H3K9me3, block = H3K9me3_annomat$ind, correlation = corfit$consensus.correlation)
fit <- lmFit(v, mm_H3K9me3, block = H3K9me3_annomat$ind, correlation = corfit$consensus.correlation)
cm <- makeContrasts(
  DOX_24T.VEH_24T = DOX_24T-VEH_24T,
  DOX_24R.VEH_24R = DOX_24R-VEH_24R,
  DOX_144R.VEH_144R = DOX_144R-VEH_144R,
  levels = mm_H3K9me3)

fit2<- contrasts.fit(fit, contrasts=cm)
efit2 <- eBayes(fit2)

results = decideTests(efit2)

summary(results)

```

```{r H3K9me3 volcano 2}
plotSA(efit2, main="Mean-Variance trend for final model for H3K9me3")
```

```{r H3K9me3 volcano 3, fig.width=14}
V.24T.top= topTable(efit2, coef=1, adjust.method="BH", number=Inf, sort.by="p")
V.24R.top= topTable(efit2, coef=2, adjust.method="BH", number=Inf, sort.by="p")
V.144R.top= topTable(efit2, coef=3, adjust.method="BH", number=Inf, sort.by="p")

H3K9me3_24T <- volcanosig(V.24T.top, 0.05)+ ggtitle("DOX 24T")
H3K9me3_24R <- volcanosig(V.24R.top, 0.05)+ ggtitle("DOX 24R")+ylab("")
H3K9me3_144R <- volcanosig(V.144R.top, 0.05)+ ggtitle("DOX 144R")+ylab("")
plot_grid(H3K9me3_24T, H3K9me3_24R, H3K9me3_144R, rel_widths =c(1,1,1))
```


### Venn Diagrams

#### Venn Set Up
```{r Venn Set Up}
# genes_H3K27ac_24T <- H3K27ac_24T$data$genes[(H3K27ac_24T$data$adj.P.Val < 0.05)]
# genes_H3K27ac_24R <- H3K27ac_24R$data$genes[(H3K27ac_24R$data$adj.P.Val < 0.05)]
# genes_H3K27ac_144R <- H3K27ac_144R$data$genes[(H3K27ac_144R$data$adj.P.Val < 0.05)]

# genes_H3K27me3_24T <- H3K27me3_24T$data$genes[(H3K27me3_24T$data$adj.P.Val < 0.05)]
# genes_H3K27me3_24R <- H3K27me3_24R$data$genes[(H3K27me3_24R$data$adj.P.Val < 0.05)]
# genes_H3K27me3_144R <- H3K27me3_144R$data$genes[(H3K27me3_144R$data$adj.P.Val < 0.05)]

genes_H3K36me3_24T <- H3K36me3_24T$data$genes[(H3K36me3_24T$data$adj.P.Val < 0.05)]
genes_H3K36me3_24R <- H3K36me3_24R$data$genes[(H3K36me3_24R$data$adj.P.Val < 0.05)]
genes_H3K36me3_144R <- H3K36me3_144R$data$genes[(H3K36me3_144R$data$adj.P.Val < 0.05)]

genes_H3K9me3_24T <- H3K9me3_24T$data$genes[(H3K9me3_24T$data$adj.P.Val < 0.05)]
genes_H3K9me3_24R <- H3K9me3_24R$data$genes[(H3K9me3_24R$data$adj.P.Val < 0.05)]
genes_H3K9me3_144R <- H3K9me3_144R$data$genes[(H3K9me3_144R$data$adj.P.Val < 0.05)]
```

#### H3K27ac Venn Diagrams
```{r H3K27ac Venn,eval=FALSE}
ggVennDiagram(list("24T regions"=genes_H3K27ac_24T,"24R regions"=genes_H3K27ac_24R, "144R regions"=genes_H3K27ac_144R))
```

#### H3K27me3 Venn Diagrams
```{r H3K27me3 Venn,eval=FALSE}
ggVennDiagram(list("24T regions"=genes_H3K27me3_24T,"24R regions"=genes_H3K27me3_24R, "144R regions"=genes_H3K27me3_144R))
``` 

#### H3K36me3 Venn Diagrams
```{r H3K36me3 Venn}
ggVennDiagram(list("24T regions"=genes_H3K36me3_24T,"24R regions"=genes_H3K36me3_24R, "144R regions"=genes_H3K36me3_144R))+
  ggtitle("H3K36me3")
``` 

#### H3K9me3 Venn Diagrams
```{r H3K9me3 Venn}
ggVennDiagram(list("24T regions"=genes_H3K9me3_24T,"24R regions"=genes_H3K9me3_24R, "144R regions"=genes_H3K9me3_144R))+ ggtitle("H3K9me3")
``` 
```{r future dataframes, eval=FALSE}
# H3K27ac_toplist <- list("H3K27ac_24T"=H3K27ac_24T$data,"H3K27ac_24R"= H3K27ac_24R$data, "H3K27ac_144R"= H3K27ac_144R$data)
# saveRDS(H3K27ac_toplist, "data/DER_data/H3K27ac_toplist_nooutlier.RDS")

# H3K27me3_toplist <- list("H3K27me3_24T"=H3K27me3_24T$data,"H3K27me3_24R"= H3K27me3_24R$data, "H3K27me3_144R"= H3K27me3_144R$data)
# saveRDS(H3K27me3_toplist,"data/DER_data/H3K27me3_toplist_nooutlier.RDS")


H3K36me3_toplist <- list("H3K36me3_24T"=H3K36me3_24T$data,"H3K36me3_24R"= H3K36me3_24R$data, "H3K36me3_144R"= H3K36me3_144R$data)
# saveRDS(H3K36me3_toplist,"data/DER_data/H3K36me3_toplist_nooutlier.RDS")

H3K9me3_toplist <- list("H3K9me3_24T"=H3K9me3_24T$data,"H3K9me3_24R"= H3K9me3_24R$data, "H3K9me3_144R"= H3K9me3_144R$data)

# saveRDS(H3K9me3_toplist, "data/DER_data/H3K9me3_toplist_nooutlier.RDS")

```

