---
title: "Protein_checks"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(readxl)
library(biomaRt)
library(ggpubr)
library(httr)
library(jsonlite)
library(org.Hs.eg.db)
library(AnnotationDbi)
```

From: 
Johnson O, Paul S, Gutiérrez J ...
DNA-damage-associated protein co-expression network in cardiomyocytes informs on tolerance to genetic variation and disease
iScience, 2025; 28


```{r import proteomics data, eval=FALSE}
mmc3 <- read_excel("C:/Users/renee/Documents/Ward Lab/AC_proteomics/proteins/mmc3.xlsx", skip = 1) %>% 
  dplyr::select(Protein:Is_CVD_PPI_protein)
write.table(mmc3,"data/Other_paper_data/protein_LFC_Johnson_2025.tsv",sep = "\t",
  quote = FALSE,
  row.names = FALSE)


```

### pulling data  

```{r}
protein_data <- read.delim("data/Other_paper_data/protein_LFC_Johnson_2025.tsv")
KZFP_2026_list <- read_excel("data/Other_paper_data/human_protein_coding_KZFPs_Davis_natgen_2026.xlsx")

# ids <- as.character(protein_data$Protein)
# ids <- trimws(ids)

map <- readRDS( "data/RDS_files/uniprot_map.rds")
map_clean <- map %>%
  group_by(Entry) %>%
  summarise(
    gene = sub(" .*", "", Gene.Names),
    .groups = "drop"
  )

H3K27ac_ext200_anno <- readRDS("data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")
remap <- dplyr::select(org.Hs.eg.db,
  keys = protein_data$Protein,   # your UniProt IDs
  keytype = "UNIPROT",
  columns = c("SYMBOL", "ENTREZID"))

remap_clean <- remap %>%
  dplyr::group_by(UNIPROT) %>%
  dplyr::summarise(
    SYMBOL = dplyr::first(na.omit(SYMBOL)),
    ENTREZID = dplyr::first(na.omit(ENTREZID)),
    .groups = "drop")

ext200_anno_map <- select(org.Hs.eg.db,
                          keys=H3K27ac_ext200_anno$geneId,
                          keytype = "ENTREZID",
                          columns = c("SYMBOL","UNIPROT")) %>% 
  distinct()

```


```{r, eval=FALSE}
# human_kzfp <- 
# KZFP_2017_list %>% 
#   dplyr::filter(Species == "Homo_sapiens.GRCh38") %>% 
#   dplyr::select(Label:strand) %>% 
#   dplyr::filter(Type=="protein_coding") %>% 
#   distinct(Label,.keep_all = TRUE)

mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl",host = "useast.ensembl.org")
listAttributes(mart)

results <- getBM(attributes = c("uniprotswissprot", "hgnc_symbol", "entrezgene_id"),
  filters = "uniprotswissprot",
  values = protein_data$Protein,
  mart = mart)

other_protein_id <- getBM(attributes = c("uniprotswissprot", "hgnc_symbol", "entrezgene_id", "ensembl_gene_id"),
  filters = "ensembl_gene_id",
  values = KZFP_2026_list$`Gene ID`,
  mart = mart)

kzfp_uniprot <- other_protein_id %>% 
  dplyr::filter(uniprotswissprot!="")

protein_lookup_table <- results %>%
  ###only keeping protein if has an hgnc_cymbol and entrezid, but keeping the first entrez_id per uniprot id
  filter(!is.na(entrezgene_id)| hgnc_symbol != "") %>%
  group_by(uniprotswissprot) %>%
  slice(1) %>%
  ungroup()
KZFP_targets <- read_excel("data/Other_paper_data/KZFP_primary_targests_Supplemental_Table_S3.xlsx", sheet = "Primary_targets")

```

```{r getting accession, eval=FALSE}


url <- "https://rest.uniprot.org/uniprotkb/stream?format=tsv&fields=accession,gene_names&query=organism_id:9606"
map <- read.delim(url)


test <- merge(protein_data, map, by.x = "Protein", by.y = "Entry", all.x = TRUE)

saveRDS(map, "data/RDS_files/uniprot_map.rds")

```


I am going to check for enrichment of Set 1 Set 2 and Set 3 in the gene groupings of Protein.  First I need to create a list of entrezIDs for each set.  They will have Entrezid, Peakid, cluster for looking up purposes. I will also make a set of genes for each cluster and a "background" gene-i-verse.

```{r}
Set1_H3K27ac_lookup <- H3K27ac_ext200_anno %>% 
  dplyr::select(Peakid,cluster, geneId) %>% 
  dplyr::filter(cluster=="Set_1") 

set1_genes <- H3K27ac_ext200_anno %>% 
  dplyr::filter(cluster =="Set_1") %>% 
  dplyr::pull(geneId) %>% 
  unique()

Set2_H3K27ac_lookup <- H3K27ac_ext200_anno %>% 
  dplyr::select(Peakid,cluster, geneId) %>% 
  dplyr::filter(cluster=="Set_2")

set2_genes <- H3K27ac_ext200_anno %>% 
  dplyr::filter(cluster =="Set_2") %>% 
  dplyr::pull(geneId) %>% 
  unique()



Set3_H3K27ac_lookup <- H3K27ac_ext200_anno %>% 
  dplyr::select(Peakid,cluster, geneId) %>% 
  dplyr::filter(cluster=="Set_3")

set3_genes <- H3K27ac_ext200_anno %>% 
  dplyr::filter(cluster =="Set_3") %>% 
  dplyr::pull(geneId) %>% 
  unique()


background_genes <- unique(H3K27ac_ext200_anno$geneId)
```

Next, I am going to take the protein list and cluster by each group:
```{r}
module_ids<- merge(protein_data, remap_clean, by.x = "Protein", by.y = "UNIPROT", all.x = TRUE)

protein_module_split <- split(module_ids,module_ids$Module)

```

Now I am going to iterate over the split list and for each module list get a count of module genes/overlaps and background counts and perform a Fisher test for general enrichment over background

### Which modules are enriched in assigned genes of each set?

```{r set2 v background}

enrichment_results_2 <- lapply(names(protein_module_split), function(mod) {
  
  mod_df <- protein_module_split[[mod]]
  
  # module genes (Entrez)
  mod_genes <- unique(mod_df$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  
  # overlaps
  in_set2 <- sum(mod_genes %in% set2_genes)
  not_in_set2 <- length(mod_genes) - in_set2
  
  # background counts
  bg_set2 <- sum(background_genes %in% set2_genes)
  bg_not_set2 <- length(background_genes) - bg_set2
  
  # Fisher test matrix
  mat <- matrix(c(
    in_set2,
    not_in_set2,
    bg_set2 - in_set2,
    bg_not_set2 - not_in_set2
  ), nrow = 2)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = ft$estimate,
    in_set2 = in_set2,
    module_size = length(mod_genes)
  )
})
enrichment_df_2 <- do.call(rbind, enrichment_results_2)

enrichment_df_2$FDR <- p.adjust(enrichment_df_2$pvalue, method = "BH")
```


```{r set3 v background}

enrichment_results_3 <- lapply(names(protein_module_split), function(mod) {
  
  mod_df <- protein_module_split[[mod]]
  
  # module genes (Entrez)
  mod_genes <- unique(mod_df$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  
  # overlaps
  in_set3 <- sum(mod_genes %in% set3_genes)
  not_in_set3 <- length(mod_genes) - in_set3
  
  # background counts
  bg_set3 <- sum(background_genes %in% set3_genes)
  bg_not_set3 <- length(background_genes) - bg_set3
  
  # Fisher test matrix
  mat <- matrix(c(
    in_set3,
    not_in_set3,
    bg_set3 - in_set3,
    bg_not_set3 - not_in_set3
  ), nrow = 2)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = ft$estimate,
    in_set3 = in_set3,
    module_size = length(mod_genes)
  )
})
enrichment_df_3 <- do.call(rbind, enrichment_results_3)

enrichment_df_3$FDR <- p.adjust(enrichment_df_3$pvalue, method = "BH")
```

```{r set1 v background}

enrichment_results_1 <- lapply(names(protein_module_split), function(mod) {
  
  mod_df <- protein_module_split[[mod]]
  
  # module genes (Entrez)
  mod_genes <- unique(mod_df$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  
  # overlaps
  in_set1 <- sum(mod_genes %in% set1_genes)
  not_in_set1 <- length(mod_genes) - in_set1
  
  # background counts
  bg_set1 <- sum(background_genes %in% set1_genes)
  bg_not_set1 <- length(background_genes) - bg_set1
  
  # Fisher test matrix
  mat <- matrix(c(
    in_set1,
    not_in_set1,
    bg_set1 - in_set1,
    bg_not_set1 - not_in_set1
  ), nrow = 2)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = ft$estimate,
    in_set1 = in_set1,
    module_size = length(mod_genes)
  )
})
enrichment_df_1 <- do.call(rbind, enrichment_results_1)

enrichment_df_1$FDR <- p.adjust(enrichment_df_1$pvalue, method = "BH")
```

#### Plots of enrichment

```{r Plot for set 1}

ggplot(enrichment_df_1, aes(x=odds_ratio, -log10(FDR), label=Module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 1"
  ) +
  theme_classic()

```


```{r Plot for set 2}

ggplot(enrichment_df_2, aes(x=odds_ratio, -log10(FDR), label=Module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 2"
  ) +
  theme_classic()

```

```{r Plot for set 3}

ggplot(enrichment_df_3, aes(x=odds_ratio, -log10(FDR), label=Module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 3"
  ) +
  theme_classic()

```

### now module enrichment in set 2 v set 1

```{r enrich 2v1}
bg_genes_2v1 <- union(set1_genes,set2_genes)
enrichment_results_2v1 <- lapply(names(protein_module_split), function(mod) {
  
  mod_genes <- unique(protein_module_split[[mod]]$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  
  a <- sum(set2_genes %in% mod_genes)
  b <- sum(set2_genes %in% background_genes) - a
  
  c <- sum(set1_genes %in% mod_genes)
  d <- sum(set1_genes %in% background_genes) - c
  
  mat <- matrix(c(a, b, c, d), nrow = 2)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = ft$estimate,
    set2_in_module = a,
    set1_in_module = c
  )
})
enrichment_df_2v1 <- do.call(rbind, enrichment_results_2v1)

enrichment_df_2v1$FDR <- p.adjust(enrichment_df_2v1$pvalue, method = "BH")

```

```{r Plot for set 2v1}

ggplot(enrichment_df_2v1, aes(x=odds_ratio, -log10(FDR), label=Module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 2 vs set 1"
  ) +
  theme_classic()

datatable(enrichment_df_2v1,options = list(
    pageLength = 10,
    scrollX = TRUE))

```

### now module enrichment in set 3 v set 1

```{r enrich 3v1}
bg_genes_3v1 <- union(set1_genes,set3_genes)
enrichment_results_3v1 <- lapply(names(protein_module_split), function(mod) {
  
  mod_genes <- unique(protein_module_split[[mod]]$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  
  a <- sum(set3_genes %in% mod_genes)
  b <- sum(set3_genes %in% background_genes) - a
  
  c <- sum(set1_genes %in% mod_genes)
  d <- sum(set1_genes %in% background_genes) - c
  
  mat <- matrix(c(a, b, c, d), nrow = 2)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = ft$estimate,
    set3_in_module = a,
    set1_in_module = c
  )
})
enrichment_df_3v1 <- do.call(rbind, enrichment_results_3v1)

enrichment_df_3v1$FDR <- p.adjust(enrichment_df_3v1$pvalue, method = "BH")

```

```{r Plot for set 3v1}

ggplot(enrichment_df_3v1, aes(x=odds_ratio, -log10(FDR), label=Module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 3 vs Set 1"
  ) +
  theme_classic()
datatable(enrichment_df_3v1,options = list(
    pageLength = 10,
    scrollX = TRUE))
```




```{r KZFP in protein}


conversion_ids %>% 
  dplyr::filter(!is.na(gene)) %>% 
  mutate(KZFP_group=case_when(gene %in%KZFP_2026_list$Label~"KZFP",
                              TRUE ~"not_KZFP")) %>% 
  ggplot(., aes(x=KZFP_group,y=logFC))+
  geom_boxplot()+
   stat_compare_means(method = "wilcox.test") +
  theme_bw()+
  ggtitle("LFC of KZFPs in protein data")

conversion_ids %>% 
  dplyr::filter(!is.na(gene)) %>% 
   mutate(KZFP_group=case_when(gene %in%KZFP_2026_list$Label~"KZFP",
                              TRUE ~"not_KZFP")) %>% 
  group_by(KZFP_group) %>% 
  tally


```



```{r}

  conversion_ids

H3K27ac_ext200_anno


```

```{r pwmpull, eval=FALSE}
TF_Information <- read_delim("C:/Users/renee/Downloads/F135_3.00_2026_04_10_5_41_pm_allC2H2ZFs/TF_Information.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

kzpf_motif_IDs <- TF_Information %>% 
  dplyr::filter(TF_Species=="Homo_sapiens") %>% 
  dplyr::filter(TF_Name %in%KZFP_2026_list$Label) %>% 
  dplyr::filter(Motif_ID != ".") %>% 
  pull(Motif_ID)

TF_Information %>% 
  filter(TF_Species == "Homo_sapiens") %>% 
  filter(TF_Name %in% KZFP_2026_list$Label) %>% 
  summarise(
    total = dplyr::n(),
    with_motif = sum(Motif_ID != "."),
    missing = sum(Motif_ID == ".")
  )

KZFP_motif_info <- TF_Information %>% 
  filter(TF_Species == "Homo_sapiens") %>% 
  filter(TF_Name %in% KZFP_2026_list$Label) %>% 
   dplyr::filter(Motif_ID != ".") 
# saveRDS(KZFP_motif_info,"data/RDS_files/KZFP_motif_info.RDS")

KZFP_no_motif_list <- 
   TF_Information %>% 
  filter(TF_Species == "Homo_sapiens") %>% 
  filter(TF_Name %in% KZFP_2026_list$Label) %>% 
  filter(Motif_ID == ".") %>% 
  pull(TF_Name)

KZFP_no_motif_list

files <- list.files(
  "C:/Users/renee/Downloads/F135_3.00_2026_04_10_5_41_pm_allC2H2ZFs/pwms",
  full.names = TRUE)

file_ids <- tools::file_path_sans_ext(basename(files))

meta <- TF_Information %>%
  dplyr::filter(TF_Species == "Homo_sapiens") %>%
  dplyr::filter(TF_Name %in% KZFP_2026_list$Label) %>%
  dplyr::filter(Motif_ID != ".") %>%
  dplyr::select(Motif_ID, TF_Name)

motif_files <- files[file_ids %in% kzpf_motif_IDs]
motif_names <- tools::file_path_sans_ext(basename(motif_files))
alt_ids <- meta$TF_Name[match(motif_names, meta$Motif_ID)]

length(kzpf_motif_IDs)
length(motif_files)
setdiff(kzpf_motif_IDs,
        tools::file_path_sans_ext(basename(files)))

library(universalmotif)

get_tf_name <- function(x) {
  tools::file_path_sans_ext(basename(x))
}


bad <- lapply(motif_files, function(f) {

  df <- read.table(f, header = TRUE, sep = "\t", check.names = FALSE)

  cols <- intersect(c("A","C","G","T"), colnames(df))

  if (length(cols) != 4) {
    return(f)
  }

  mat <- df[, cols, drop = FALSE]
  mat <- data.matrix(mat)

  if (ncol(mat) == 0 || nrow(mat) == 0) {
    return(f)
  }

  NULL
})

bad <- unlist(bad)
bad


motifs <- lapply(motif_files, function(f) {
  
   df <- read.table(f, header = TRUE, sep = "\t", check.names = FALSE)

  # robust column matching (ignores whitespace/case issues)
  col_map <- colnames(df)
  nuc_cols <- col_map[tolower(col_map) %in% c("a","c","g","t")]

  if (length(nuc_cols) != 4) {
    warning(paste("Skipping file (bad columns):", f))
    return(NULL)
  }

  mat <- df[, nuc_cols, drop = FALSE]
  mat <- data.matrix(mat)

  if (nrow(mat) < 2 || ncol(mat) != 4) {
    warning(paste("Skipping file (bad matrix):", f))
    return(NULL)
  }

  pwm <- t(mat)

  if (ncol(pwm) == 0) {
    warning(paste("Skipping file (empty PWM):", f))
    return(NULL)
  }

  m <- universalmotif::create_motif(pwm, alphabet = "DNA")

  attr(m, "name") <- tools::file_path_sans_ext(basename(f))

  m
})


write_meme(motifs, file = "data/Other_paper_data/KZFP_motifs.meme", overwrite = TRUE)

motif_annotation <- data.frame(
  NAME = motif_names,
  ALT_ID = alt_ids
)
readLines("data/Other_paper_data/KZFP_motifs.meme", n = 40)
```
wanting to extract all SET1 SEt2 and Set3 regions that contain a MER61E element then test for enrichment.
```{r import regions} 

mer_gr <- rtracklayer::import("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed")
H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")


peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)

##assigning Peakid as name of summit region
 mcols(H3K27ac_summit_gr)$name <- mcols(H3K27ac_summit_gr)$Peakid
 
 H3K27ac_summit_gr_ext <- resize(H3K27ac_summit_gr, width(H3K27ac_summit_gr)+400, fix = "center")

H3K27ac_MER61E <- subsetByOverlaps(H3K27ac_summit_gr_ext, mer_gr)

H3K27ac_MER61E_df <- as.data.frame(H3K27ac_MER61E) 

Set1 <- H3K27ac_MER61E_df %>% filter(cluster == "Set_1")
Set2 <- H3K27ac_MER61E_df %>% filter(cluster == "Set_2")
Set3 <- H3K27ac_MER61E_df %>% filter(cluster == "Set_3")

Set1_gr <- makeGRangesFromDataFrame(Set1, keep.extra.columns = TRUE)
Set2_gr <- makeGRangesFromDataFrame(Set2, keep.extra.columns = TRUE)
Set3_gr <- makeGRangesFromDataFrame(Set3, keep.extra.columns = TRUE)

width(Set1_gr)
width(Set3_gr)


library(Biostrings)
library(BSgenome.Hsapiens.UCSC.hg38)
Hsapiens <- BSgenome.Hsapiens.UCSC.hg38

gc1 <- letterFrequency(getSeq(Hsapiens, Set1_gr), "GC", as.prob=TRUE)
gc3 <- letterFrequency(getSeq(Hsapiens, Set3_gr), "GC", as.prob=TRUE)

mean(gc1)
mean(gc3)

export(Set1_gr, "data/Bed_exports/H3K27ac_MER61E_Set_1.bed", format = "BED")
export(Set3_gr, "data/Bed_exports/H3K27ac_MER61E_Set_3.bed", format = "BED")
```

