---
title: "Protein_checks"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(readxl)
library(biomaRt)
library(ggpubr)
library(httr)
library(jsonlite)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(MASS)
library(GenomicFeatures)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(JASPAR2024)
library(TFBSTools)
library(universalmotif)
```

From: 
Johnson O, Paul S, Gutiérrez J ...
DNA-damage-associated protein co-expression network in cardiomyocytes informs on tolerance to genetic variation and disease
iScience, 2025; 28


```{r import proteomics data, eval=FALSE}
mmc3 <- read_excel("C:/Users/renee/Documents/Ward Lab/AC_proteomics/proteins/mmc3.xlsx", skip = 1) %>% 
  dplyr::select(Protein:Is_CVD_PPI_protein)
write.table(mmc3,"data/Other_paper_data/protein_LFC_Johnson_2025.tsv",sep = "\t",
  quote = FALSE,
  row.names = FALSE)


```

### pulling data  

```{r}
protein_data <- read.delim("data/Other_paper_data/protein_LFC_Johnson_2025.tsv")
KZFP_2026_list <- read_excel("data/Other_paper_data/human_protein_coding_KZFPs_Davis_natgen_2026.xlsx")

# ids <- as.character(protein_data$Protein)
# ids <- trimws(ids)

map <- readRDS( "data/RDS_files/uniprot_map.rds")
map_clean <- map %>%
  group_by(Entry) %>%
  summarise(
    gene = sub(" .*", "", Gene.Names),
    .groups = "drop"
  )

H3K27ac_ext200_anno <- readRDS("data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")
remap <- AnnotationDbi::select(org.Hs.eg.db,
  keys = protein_data$Protein,   # your UniProt IDs
  keytype = "UNIPROT",
  columns = c("SYMBOL", "ENTREZID"))

remap_clean <- remap %>%
  dplyr::group_by(UNIPROT) %>%
  dplyr::summarise(
    SYMBOL = dplyr::first(na.omit(SYMBOL)),
    ENTREZID = dplyr::first(na.omit(ENTREZID)),
    .groups = "drop")

ext200_anno_map <- AnnotationDbi::select(org.Hs.eg.db,
                          keys=H3K27ac_ext200_anno$geneId,
                          keytype = "ENTREZID",
                          columns = c("SYMBOL","UNIPROT")) %>% 
  distinct()

Module_lookup <- protein_data %>% 
  mutate(greek_module=case_when(Module=="green"~"alpha",
                                Module=="darkgreen"~"beta",
                                Module=="midnightblue"~"gamma",
                                Module=="salmon"~"delta",
                                Module=="lightyellow"~"epsilon",
                                Module=="lightgreen"~"zeta",
                                 Module=="blue"~"eta",
                                 Module=="magenta"~"theta",
                                 Module=="darkred"~"iota",
                                 Module=="brown"~"kappa",
                                 Module=="yellow"~"lambda",
                                 Module=="royalblue"~"mu",
                                Module=="grey"~"Unassigned")) %>% 
  dplyr::select(Module,greek_module) %>% 
  distinct()

alpha   <- "\u03B1"  # α
beta    <- "\u03B2"  # β
gamma   <- "\u03B3"  # γ
delta   <- "\u03B4"  # δ
epsilon <- "\u03B5"  # ε
zeta    <- "\u03B6"  # ζ
eta     <- "\u03B7"  # η
theta   <- "\u03B8"  # θ
iota    <- "\u03B9"  # ι
kappa   <- "\u03BA"  # κ
lambda  <- "\u03BB"  # λ
mu      <- "\u03BC"  # μ


```


```{r, eval=FALSE}
# human_kzfp <- 
# KZFP_2017_list %>% 
#   dplyr::filter(Species == "Homo_sapiens.GRCh38") %>% 
#   dplyr::select(Label:strand) %>% 
#   dplyr::filter(Type=="protein_coding") %>% 
#   distinct(Label,.keep_all = TRUE)

mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl",host = "useast.ensembl.org")
listAttributes(mart)

results <- getBM(attributes = c("uniprotswissprot", "hgnc_symbol", "entrezgene_id"),
  filters = "uniprotswissprot",
  values = protein_data$Protein,
  mart = mart)

other_protein_id <- getBM(attributes = c("uniprotswissprot", "hgnc_symbol", "entrezgene_id", "ensembl_gene_id"),
  filters = "ensembl_gene_id",
  values = KZFP_2026_list$`Gene ID`,
  mart = mart)

kzfp_uniprot <- other_protein_id %>% 
  dplyr::filter(uniprotswissprot!="")

protein_lookup_table <- results %>%
  ###only keeping protein if has an hgnc_cymbol and entrezid, but keeping the first entrez_id per uniprot id
  filter(!is.na(entrezgene_id)| hgnc_symbol != "") %>%
  group_by(uniprotswissprot) %>%
  slice(1) %>%
  ungroup()
KZFP_targets <- read_excel("data/Other_paper_data/KZFP_primary_targests_Supplemental_Table_S3.xlsx", sheet = "Primary_targets")

```

```{r getting accession, eval=FALSE}


url <- "https://rest.uniprot.org/uniprotkb/stream?format=tsv&fields=accession,gene_names&query=organism_id:9606"
map <- read.delim(url)


test <- merge(protein_data, map, by.x = "Protein", by.y = "Entry", all.x = TRUE)

saveRDS(map, "data/RDS_files/uniprot_map.rds")

```


I am going to check for enrichment of Set 1 Set 2 and Set 3 in the gene groupings of Protein.  First I need to create a list of entrezIDs for each set.  They will have Entrezid, Peakid, cluster for looking up purposes. I will also make a set of genes for each cluster and a "background" gene-i-verse.

```{r}
Set1_H3K27ac_lookup <- H3K27ac_ext200_anno %>% 
  dplyr::select(Peakid,cluster, geneId,distanceToTSS) %>% 
  dplyr::filter(cluster=="Set_1") 

set1_genes <- H3K27ac_ext200_anno %>% 
  dplyr::filter(cluster =="Set_1") %>% 
  dplyr::pull(geneId) %>% 
  unique()

Set2_H3K27ac_lookup <- H3K27ac_ext200_anno %>% 
  dplyr::select(Peakid,cluster, geneId,distanceToTSS) %>% 
  dplyr::filter(cluster=="Set_2")

set2_genes <- H3K27ac_ext200_anno %>% 
  dplyr::filter(cluster =="Set_2") %>% 
  dplyr::pull(geneId) %>% 
  unique()



Set3_H3K27ac_lookup <- H3K27ac_ext200_anno %>% 
  dplyr::select(Peakid,cluster, geneId,distanceToTSS) %>% 
  dplyr::filter(cluster=="Set_3")

set3_genes <- H3K27ac_ext200_anno %>% 
  dplyr::filter(cluster =="Set_3") %>% 
  dplyr::pull(geneId) %>% 
  unique()


background_genes <- unique(H3K27ac_ext200_anno$geneId)
```

Next, I am going to take the protein list and cluster by each group:
```{r}
module_ids<- merge(protein_data, remap_clean, by.x = "Protein", by.y = "UNIPROT", all.x = TRUE)

protein_module_split <- split(module_ids,module_ids$Module)

gene_universe <- intersect(module_ids$ENTREZID,H3K27ac_ext200_anno$geneId)
length(intersect(background_genes,gene_universe))

```

Now I am going to iterate over the split list and for each module list get a count of module genes/overlaps and background counts and perform a Fisher test for general enrichment over background

### Which modules are enriched in assigned genes of each set?

```{r set2 v background}

enrichment_results_2 <- lapply(names(protein_module_split), function(mod) {
  
  mod_df <- protein_module_split[[mod]]
  
  # module genes
  mod_genes <- unique(mod_df$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  mod_genes <- intersect(mod_genes, background_genes)
  
  # overlaps
  in_set2 <- sum(mod_genes %in% set2_genes)
  not_in_set2 <- length(mod_genes) - in_set2
  
  # background universe constrained
  bg_set2 <- sum(background_genes %in% set2_genes)
  bg_not_set2 <- length(background_genes) - bg_set2
  
  mat <- matrix(c(
    in_set2,
    not_in_set2,
    bg_set2 - in_set2,
    bg_not_set2 - not_in_set2
  ), nrow = 2,byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = unname(ft$estimate),
    in_set2 = in_set2,
    module_size = length(mod_genes)
  )
})
enrichment_df_2 <- do.call(rbind, enrichment_results_2)

enrichment_df_2$FDR <- p.adjust(enrichment_df_2$pvalue, method = "BH")
enrichment_df_2 <-enrichment_df_2 %>% 
  left_join(Module_lookup)
```


```{r set3 v background}

enrichment_results_3 <- lapply(names(protein_module_split), function(mod) {
  
  mod_df <- protein_module_split[[mod]]
  
  # module genes
  mod_genes <- unique(mod_df$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  mod_genes <- intersect(mod_genes, background_genes)
  
  # overlaps
  in_set3 <- sum(mod_genes %in% set3_genes)
  not_in_set3 <- length(mod_genes) - in_set3
  
  # background universe constrained
  bg_set3 <- sum(background_genes %in% set3_genes)
  bg_not_set3 <- length(background_genes) - bg_set3
  
  mat <- matrix(c(
    in_set3,
    not_in_set3,
    bg_set3 - in_set3,
    bg_not_set3 - not_in_set3
  ), nrow = 2,byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = unname(ft$estimate),
    in_set3 = in_set3,
    module_size = length(mod_genes)
  )
})
enrichment_df_3 <- do.call(rbind, enrichment_results_3)

enrichment_df_3$FDR <- p.adjust(enrichment_df_3$pvalue, method = "BH")
enrichment_df_3 <-enrichment_df_3 %>% 
  left_join(Module_lookup)
```

```{r set1 v background}

enrichment_results_1 <- lapply(names(protein_module_split), function(mod) {
  
   mod_df <- protein_module_split[[mod]]
  
  # module genes
  mod_genes <- unique(mod_df$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  mod_genes <- intersect(mod_genes, background_genes)
  
  # overlaps
  in_set1 <- sum(mod_genes %in% set1_genes)
  not_in_set1 <- length(mod_genes) - in_set1
  
  # background universe constrained
  bg_set1 <- sum(background_genes %in% set1_genes)
  bg_not_set1 <- length(background_genes) - bg_set1
  
  mat <- matrix(c(
    in_set1,
    not_in_set1,
    bg_set1 - in_set1,
    bg_not_set1 - not_in_set1
  ), nrow = 2,byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = unname(ft$estimate),
    in_set1 = in_set1,
    module_size = length(mod_genes)
  )
})
enrichment_df_1 <- do.call(rbind, enrichment_results_1)

enrichment_df_1$FDR <- p.adjust(enrichment_df_1$pvalue, method = "BH")
enrichment_df_1 <-enrichment_df_1 %>% 
  left_join(Module_lookup)
```

#### Plots of enrichment

```{r Plot for set 1}

ggplot(enrichment_df_1, aes(x=odds_ratio, -log10(FDR), label=greek_module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 1"
  ) +
  theme_classic()

```


```{r Plot for set 2}

ggplot(enrichment_df_2, aes(x=odds_ratio, -log10(FDR), label=greek_module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 2"
  ) +
  theme_classic()

```

```{r Plot for set 3}

ggplot(enrichment_df_3, aes(x=odds_ratio, -log10(FDR), label=greek_module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 3"
  ) +
  theme_classic()

```

### now module enrichment in set 2 v set 1

```{r enrich 2v1}
bg_genes_2v1 <- union(set1_genes,set2_genes)
enrichment_results_2v1 <- lapply(names(protein_module_split), function(mod) {
  
  mod_genes <- unique(protein_module_split[[mod]]$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  
  a <- sum(set2_genes %in% mod_genes)
  b <- sum(set2_genes %in% bg_genes_2v1) - a
  
  c <- sum(set1_genes %in% mod_genes)
  d <- sum(set1_genes %in% bg_genes_2v1) - c
  
  mat <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value, 
    odds_ratio = ft$estimate,
    set2_in_module = a,
    set1_in_module = c
  )
})
enrichment_df_2v1 <- do.call(rbind, enrichment_results_2v1)

enrichment_df_2v1$FDR <- p.adjust(enrichment_df_2v1$pvalue, method = "BH")
enrichment_df_2v1 <-enrichment_df_2v1 %>% 
  left_join(Module_lookup)
```

```{r Plot for set 2v1}

ggplot(enrichment_df_2v1, aes(x=odds_ratio, -log10(FDR), label=greek_module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 2 vs set 1"
  ) +
  theme_classic()

datatable(enrichment_df_2v1,options = list(
    pageLength = 10,
    scrollX = TRUE))

```

### now module enrichment in set 3 v set 1

```{r enrich 3v1}
bg_genes_3v1 <- union(set1_genes,set3_genes)
enrichment_results_3v1 <- lapply(names(protein_module_split), function(mod) {
  
  mod_genes <- unique(protein_module_split[[mod]]$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  
  a <- sum(set3_genes %in% mod_genes)
  b <- sum(set3_genes %in% bg_genes_3v1) - a
  
  c <- sum(set1_genes %in% mod_genes)
  d <- sum(set1_genes %in% bg_genes_3v1) - c
  
  mat <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = ft$estimate,
    set3_in_module = a,
    set1_in_module = c
  )
})
enrichment_df_3v1 <- do.call(rbind, enrichment_results_3v1)

enrichment_df_3v1$FDR <- p.adjust(enrichment_df_3v1$pvalue, method = "BH")
enrichment_df_3v1 <-enrichment_df_3v1 %>% 
  left_join(Module_lookup)
```

```{r Plot for set 3v1}

ggplot(enrichment_df_3v1, aes(x=odds_ratio, -log10(FDR), label=paste0(greek_module)))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 3 vs Set 1"
  ) +
  theme_classic()
datatable(enrichment_df_3v1,options = list(
    pageLength = 10,
    scrollX = TRUE))
```
### does distance to TSS change enrichment?
```{r}
set1_2kb_genes <- Set1_H3K27ac_lookup %>% 
  dplyr::filter(distanceToTSS<2000 & distanceToTSS>-2000) %>% 
  pull(geneId) %>% 
  unique()

set2_2kb_genes <- Set2_H3K27ac_lookup %>% 
  dplyr::filter(distanceToTSS<2000 & distanceToTSS>-2000) %>% 
  pull(geneId) %>% 
  unique()

set3_2kb_genes <- Set3_H3K27ac_lookup %>% 
  dplyr::filter(distanceToTSS<2000 & distanceToTSS>-2000) %>% 
  pull(geneId) %>% 
  unique()
```

```{r}
bg_2v1_2kb <- union(set1_2kb_genes, set2_2kb_genes)
enrichment_results_2v1_2kb <- lapply(names(protein_module_split), function(mod) {
  
  mod_genes <- unique(protein_module_split[[mod]]$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  
  a <- sum(set2_2kb_genes %in% mod_genes)
  b <- sum(set2_2kb_genes %in% bg_2v1_2kb) - a
  
  c <- sum(set1_2kb_genes %in% mod_genes)
  d <- sum(set1_2kb_genes %in% bg_2v1_2kb) - c
  
  mat <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = ft$estimate,
    set3_in_module = a,
    set1_in_module = c
  )
})
enrichment_df_2v1_2kb <- do.call(rbind, enrichment_results_2v1_2kb)

enrichment_df_2v1_2kb$FDR <- p.adjust(enrichment_df_2v1_2kb$pvalue, method = "BH")

ggplot(enrichment_df_2v1_2kb, aes(x=odds_ratio, -log10(FDR), label=Module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 2 vs set 1"
  ) +
  theme_classic()

```


```{r}
bg_3v1_2kb <- union(set1_2kb_genes, set3_2kb_genes)
enrichment_results_3v1_2kb <- lapply(names(protein_module_split), function(mod) {
  
  mod_genes <- unique(protein_module_split[[mod]]$ENTREZID)
  mod_genes <- mod_genes[!is.na(mod_genes)]
  
  a <- sum(set3_2kb_genes %in% mod_genes)
  b <- sum(set3_2kb_genes %in% bg_3v1_2kb) - a
  
  c <- sum(set1_2kb_genes %in% mod_genes)
  d <- sum(set1_2kb_genes %in% bg_3v1_2kb) - c
  
  mat <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  
  ft <- fisher.test(mat)
  
  data.frame(
    Module = mod,
    pvalue = ft$p.value,
    odds_ratio = ft$estimate,
    set3_in_module = a,
    set1_in_module = c
  )
})
enrichment_df_3v1_2kb <- do.call(rbind, enrichment_results_3v1_2kb)

enrichment_df_3v1_2kb$FDR <- p.adjust(enrichment_df_3v1_2kb$pvalue, method = "BH")

ggplot(enrichment_df_3v1_2kb, aes(x=odds_ratio, -log10(FDR), label=Module))+
  geom_point(aes(color = odds_ratio > 1), size = 3) +
   # geom_text_repel(data = subset(SVA_results, FDR < 0.05)) +
  geom_text_repel(size=3, max.overlaps = Inf)+
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  labs(
    x = "(Odds Ratio)",
    y = "-log10(FDR)",
    title = "Module Enrichment of assigned nearest genes in Set 3 vs set 1"
  ) +
  theme_classic()

```

### Frequency checks of Set 2 v  Set 1

This section is to investigate the frequency of near-gene assignments of Proteins from Omar's paper to my interested regions. Here I am taking into account region assignment totals in Set 2 vs Set 1.
test will be number of regions in set2 assigned to protein data gene, number of regions in set 2 not assigned to protein data gene, compared to number of regions assigned to set1 vs not assigned to protein data gene in set 1.
  
  
  more specifically I want to test if Set 2 regions are more frequently assigned to differentially abundant proteins than set 1 H3K27ac regions.

```{r}
Protein_list <-   module_ids %>%
  dplyr::select(Protein, Module,Is_DA_at_adj.p, Is_DOXcorrelated, SYMBOL,ENTREZID) 

Protein_list_clean <- Protein_list %>%
  filter(!is.na(ENTREZID)) %>% 
  group_by(ENTREZID) %>%
  summarise(
    in_protein_data = TRUE,
    SYMBOL = paste(unique(na.omit(SYMBOL)), collapse = ";"),
    Protein = paste(unique(na.omit(Protein)), collapse = ";"),
    Module = paste(unique(na.omit(Module)), collapse = ";"),
    any_DA = any(Is_DA_at_adj.p, na.rm = TRUE),
    any_DOXcorrelated = any(Is_DOXcorrelated, na.rm = TRUE),
    .groups = "drop")

region_data <- H3K27ac_ext200_anno %>% 
  dplyr::select(Peakid, geneId, cluster) %>% 
  left_join(Protein_list_clean, by=c("geneId"="ENTREZID")) %>% 
  mutate(
    any_DA = coalesce(any_DA, 0),
    any_DOXcorrelated= coalesce(any_DOXcorrelated,0))

test_df <- region_data %>%
  filter(cluster %in% c("Set_1", "Set_2")) %>%
  mutate(
    assigned_to_DA_protein = any_DA == 1)
  

a <- sum(test_df$cluster == "Set_2" &  test_df$assigned_to_DA_protein)
b <- sum(test_df$cluster == "Set_2" &  !test_df$assigned_to_DA_protein)
c <- sum(test_df$cluster == "Set_1" &  test_df$assigned_to_DA_protein)
d <- sum(test_df$cluster == "Set_1" &  !test_df$assigned_to_DA_protein)

tab <- matrix(
  c(a, b, c, d),
  nrow = 2,
  byrow = TRUE
)
fisher.test(tab, alternative = "greater")  


```

hmmm, nope not enriched. And for future reference, I did this for Set 3 and also did not find enrichment.  similar to the tests on modules above.

### DAP enrichment in 27ac sets

```{r checking 2 v1 DAP}
DAP_IDs <- protein_data %>% 
dplyr::select(Protein,Module,logFC,P.Value.adj, Is_DA_at_adj.p) %>% 
  dplyr::filter(Is_DA_at_adj.p==1) %>% 
   left_join(remap_clean, by =c("Protein"="UNIPROT")) 
 
nonDAP_IDs <- protein_data %>% 
dplyr::select(Protein,Module,logFC,P.Value.adj, Is_DA_at_adj.p) %>% 
  dplyr::filter(Is_DA_at_adj.p!=1) %>% 
   left_join(remap_clean, by =c("Protein"="UNIPROT"))

a=sum(set2_genes%in%DAP_IDs$ENTREZID)
b=sum(!set2_genes%in% DAP_IDs$ENTREZID)
  c=sum(set1_genes%in%DAP_IDs$ENTREZID)
 d=sum(!set1_genes%in%DAP_IDs$ENTREZID)

tab <- matrix(
  c(a, b, c, d),
  nrow = 2,
  byrow = TRUE
)
fisher.test(tab, alternative = "greater") 
```

```{r Checking 3v 1 DAP}
a=sum(set3_genes%in%DAP_IDs$ENTREZID)
b=sum(!set3_genes%in% DAP_IDs$ENTREZID)
  c=sum(set1_genes%in%DAP_IDs$ENTREZID)
 d=sum(!set1_genes%in%DAP_IDs$ENTREZID)

tab <- matrix(
  c(a, b, c, d),
  nrow = 2,
  byrow = TRUE)

fisher.test(tab, alternative = "greater") 
```



### KZFP
just trying to see if KZFPs are in protein data.
```{r KZFP in protein}

protein_data %>% 
  dplyr::select(Protein,Module,logFC,P.Value.adj) %>% 
  left_join(remap_clean, by =c("Protein"="UNIPROT")) %>% 
  mutate(KZFP_group=case_when(SYMBOL %in%KZFP_2026_list$Label~"KZFP",
                              TRUE ~"not_KZFP")) %>% 
  ggplot(., aes(x=KZFP_group,y=logFC))+
  geom_boxplot()+
   stat_compare_means(method = "wilcox.test") +
  theme_bw()+
  ggtitle("LFC of KZFPs in protein data")

protein_data %>% 
  dplyr::select(Protein,Module,logFC,P.Value.adj) %>% 
  left_join(remap_clean, by =c("Protein"="UNIPROT")) %>% 
  mutate(KZFP_group=case_when(SYMBOL %in%KZFP_2026_list$Label~"KZFP",
                              TRUE ~"not_KZFP")) %>%
  group_by(KZFP_group) %>% 
  tally


```

### DAP enrichment
  getting proteins that are DAP vs non-DAP
```{r}
DAP_IDs <- protein_data %>% 
dplyr::select(Protein,Module,logFC,P.Value.adj, Is_DA_at_adj.p) %>% 
  dplyr::filter(Is_DA_at_adj.p==1) %>% 
   left_join(remap_clean, by =c("Protein"="UNIPROT")) 
 
nonDAP_IDs <- protein_data %>% 
dplyr::select(Protein,Module,logFC,P.Value.adj, Is_DA_at_adj.p) %>% 
  dplyr::filter(Is_DA_at_adj.p!=1) %>% 
   left_join(remap_clean, by =c("Protein"="UNIPROT"))


Entrezid_protein_module <- protein_data %>% 
dplyr::select(Protein,Module,logFC,P.Value.adj, Is_DA_at_adj.p) %>% 
     left_join(remap_clean, by =c("Protein"="UNIPROT")) %>% 
  dplyr::filter(!is.na(SYMBOL)) %>%
  distinct(Module,ENTREZID,SYMBOL)


Entrezid_protein_module_unique <- Entrezid_protein_module%>%
  group_by(SYMBOL) %>%
  summarise(Module = dplyr::first(Module),
            ENTREZID=dplyr::first(ENTREZID), .groups = "drop")
```

asking are genes near mer61E_gr enriched in any module or in DAPs?

```{r}
MER61E_mydata <- 
H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(repName=="MER61E")%>% 
  pull(repID) %>% 
  unique()
MER61E_mydata_set3 <- H3K27ac_ext200_anno %>% 
  separate_longer_delim(cols=c(repClass,repFamily,repName,repID), delim = ";") %>% 
  dplyr::filter(cluster=="Set_3") %>% 
  dplyr::filter(repName=="MER61E")%>% 
  pull(repID) %>% 
  unique()

mer61E_gr <- rtracklayer::import("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed") 
# txdb annotation
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
# gene coordinates
genes_gr <- genes(txdb)
# nearest gene index
nearest_idx <- nearest(mer61E_gr, genes_gr)
# nearest gene IDs
nearest_entrez <- genes_gr$gene_id[nearest_idx]
# add to MER61E object
mcols(mer61E_gr)$nearest_gene <- nearest_entrez
distances <- distance(mer61E_gr, genes_gr[nearest_idx])

mcols(mer61E_gr)$distance_to_gene <- distances
## getting the dataframe of nearest genes in and out of my data for counting:
mer61E_gr %>% 
  as.data.frame() %>% 
  mutate(my_MER=if_else(name %in%MER61E_mydata,"in","out")) %>% 
  ggplot(aes(x=distance_to_gene, color=my_MER))+
  geom_density()

mer61E_gr %>% 
  as.data.frame() %>% 
  mutate(my_MER=if_else(name %in%MER61E_mydata,"in","out")) %>% 
  mutate(is_DAP=if_else(nearest_gene%in%DAP_IDs$ENTREZID,"DAP","not_DAP")) %>% 
  left_join(Entrezid_protein_module_unique, by=c("nearest_gene"="ENTREZID")) %>% 
  mutate(DOX_cor= if_else(Module %in% c("green","lightyellow","salmon","darkgreen","midnightblue"),"DOXcor","non_DOXcor")) %>% 
  group_by(my_MER, DOX_cor) %>% tally
  

my_matrix <- matrix(c(3,64,10,190), nrow = 2, byrow = TRUE)
fisher.test(my_matrix)
```


```{r GREAT nearest gene to MER61E}
repID_to_gene_MER61E <- read_delim("data/Other_paper_data/repID_to_gene_MER61E.txt", 
    delim = "\t", escape_double = FALSE, 
    col_names = FALSE, trim_ws = TRUE, skip = 1)

repID_to_gene_MER61E$closest_gene <- vapply(repID_to_gene_MER61E$X2, function(x){
  # handle missing/empty
    if (is.na(x) || x == "") {
      return(NA_character_)
    }

    # split entries
    parts <- str_split(x, ",\\s*")[[1]]

    # extract distances
    dists <- as.numeric(str_extract(parts, "[+-]\\d+"))

    # extract genes
    genes <- str_extract(parts, "^[^ ]+")

    # remove failed parses
    keep <- !is.na(dists) & !is.na(genes)

    dists <- dists[keep]
    genes <- genes[keep]

    # if nothing valid remains
    if (length(genes) == 0) {
      return(NA_character_)
    }

    # closest gene
    genes[which.min(abs(dists))]

  },
  FUN.VALUE = character(1))

repID_to_gene_MER61E %>% 
  filter(!is.na(closest_gene)) %>% 
   mutate(my_MER=if_else(X1 %in%MER61E_mydata,"in","out")) %>% 
  mutate(is_DAP=if_else(closest_gene%in%DAP_IDs$SYMBOL,"DAP","not_DAP")) %>% 
  left_join(Entrezid_protein_module_unique, by=c("closest_gene"="SYMBOL")) %>% 
  mutate(DOX_cor= if_else(Module %in% c("green","lightyellow","salmon","darkgreen","midnightblue"),"DOXcor","non_DOXcor")) %>% 
  group_by(my_MER,DOX_cor) %>% 
  tally

great_matrix <- matrix(c(5,56,17,161), nrow = 2, byrow = TRUE)

fisher.test(great_matrix)


# table(table(repID_to_gene_MER61E$closest_gene))
# table(repID_to_gene_MER61E$closest_gene, useNA = "ifany")
```


```{r}
repID_to_gene_MER61E %>% 
  filter(!is.na(closest_gene)) %>% 
   mutate(my_MER=if_else(X1 %in%MER61E_mydata_set3,"in","out")) %>% 
  mutate(is_DAP=if_else(closest_gene%in%DAP_IDs$SYMBOL,"DAP","not_DAP")) %>% 
  left_join(Entrezid_protein_module_unique, by=c("closest_gene"="SYMBOL")) %>% 
  mutate(DOX_cor= if_else(Module %in% c("green","lightyellow","salmon","darkgreen","midnightblue"),"DOXcor","non_DOXcor")) %>% 
  group_by(my_MER,DOX_cor) %>% 
  tally
great_matrix_set3 <- matrix(c(3,34,19,183), nrow = 2,byrow = TRUE)

fisher.test(great_matrix_set3)
```
### Looking at TF motif/protein enrichment

```{r}
H3K27ac_Set2_data <- readRDS("data/RDS_files/H3K27ac_Set2_enrmotif_data.RDS")
H3K27ac_Set3_data <- readRDS("data/RDS_files/H3K27ac_Set3_enrmotif_data.RDS")

## TF universe
# jaspar2024 <- JASPAR2024::JASPAR2024()
# opts <- list(
#   collection = "CORE",
#   tax_group = "vertebrates",
#   species = 9606,
#   all_versions = FALSE
# )
# 
# pfm_list <- TFBSTools::getMatrixSet(jaspar2024, opts)
# tf_names <- sapply(pfm_list, name)

motif_names_set2H3K27ac <-
H3K27ac_Set2_data %>%
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>%  
  dplyr::select(motif_name) %>% 
  mutate(TF_split=str_split(motif_name,"::")) %>% 
  unnest(TF_split) %>% 
  mutate(motif_name2=toupper(TF_split)) %>% 
  mutate(is_DAP_comp=motif_name2 %in%Entrezid_protein_module_unique$SYMBOL) %>% 
  group_by(motif_name) %>% 
  summarise(
    DAP = if_else(any(is_DAP_comp), "DAP", "non_DAP"),
    .groups = "drop"
  ) %>% 
  mutate(motif_name=toupper(motif_name)) %>% 
  distinct()


motif_names_set3H3K27ac <- H3K27ac_Set3_data %>%
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>%  
  dplyr::select(motif_name) %>% 
  mutate(TF_split=str_split(motif_name,"::")) %>% 
  unnest(TF_split) %>% 
  mutate(motif_name2=toupper(TF_split)) %>% 
  mutate(is_DAP_comp=motif_name2 %in%Entrezid_protein_module_unique$SYMBOL) %>% 
  group_by(motif_name) %>% 
  summarise(
    DAP = if_else(any(is_DAP_comp), "DAP", "non_DAP"),
    .groups = "drop"
  ) %>% 
  mutate(motif_name=toupper(motif_name)) %>% 
  distinct()



# motif_names_set2H3K27ac

```
### Using exact meme file of TFs for enrichment\

```{r}
motifs <- read_meme("data/Other_paper_data/JASPAR2024_CORE_vertebrates_non-redundant_v2.meme")
### pulling the alt_name from each element (841 elements)
tf_names <- sapply(motifs, function(x) x@altname)

### no making the totupper and removing the redundant names (those that are not human are lowercase but maybe the same, ie.LHX1, Lhx1)  This leaves 779 total
tf_names <- tf_names %>% 
  toupper() %>% 
  unique()

H3K27ac_TF_df <- tibble(tf_names) %>% 
  mutate(TF_split=str_split(tf_names,"::")) %>% 
  unnest(TF_split) %>% 
  mutate(is_DAP_comp=TF_split %in%Entrezid_protein_module_unique$SYMBOL) %>% 
  group_by(tf_names) %>% 
  summarise(
    DAP = if_else(any(is_DAP_comp), "DAP", "non_DAP"),
    .groups = "drop"
  ) %>% 
  mutate(Set_2=if_else(tf_names %in% motif_names_set2H3K27ac$motif_name,"yes","no"),
         Set_3=if_else(tf_names %in% motif_names_set3H3K27ac$motif_name,"yes","no")) %>% 
   mutate(
    DAP = factor(DAP, levels = c("DAP", "non_DAP")),
    Set_2 = factor(Set_2, levels = c("yes", "no")),
    Set_3 = factor(Set_3, levels = c("yes", "no"))) %>% 
  mutate(
    set_group = case_when(
      Set_2 == "yes" & Set_3 == "yes" ~ "shared",
      Set_2 == "yes" & Set_3 == "no"  ~ "set_2_only",
      Set_2 == "no"  & Set_3 == "yes" ~ "set_3_only",
      TRUE ~ "neither"))



tf_matrix_set2 <- table(H3K27ac_TF_df$DAP, H3K27ac_TF_df$Set_2)

fisher.test(tf_matrix_set2)

```


```{r}

tf_matrix_set3 <- table(H3K27ac_TF_df$DAP, H3K27ac_TF_df$Set_3)

fisher.test(tf_matrix_set3)

```
almost set3, good odds ratio but not significant with a pvalue=0.1075.

Now I will test if the unique to Set 3, unique to Set 2, or shared TFs are enriched in DAPs vs non-DAPs
```{r H3K27 ac sus only}


table(H3K27ac_TF_df$DAP,H3K27ac_TF_df$set_group=="set_2_only")

tf_uni_matrix_set2 <- matrix(c(11,19,262,487), nrow = 2,byrow = TRUE)

fisher.test(tf_uni_matrix_set2)

```
```{r H3K27ac acute}

table(H3K27ac_TF_df$DAP,H3K27ac_TF_df$set_group=="set_3_only")

tf_uni_matrix_set3 <- matrix(c(6,24,30,719), nrow = 2, byrow = TRUE)

fisher.test(tf_uni_matrix_set3)



```
So we do know that the unique set 3 enriched TFs are also enriched in DAPs significantly.

```{r H3K27ac shared}
table(H3K27ac_TF_df$DAP,H3K27ac_TF_df$set_group=="shared")

tf_uni_matrix_setshare <- matrix(c(0,30,41,708), nrow = 2)

fisher.test(tf_uni_matrix_setshare)


```




### H3K9me3  

```{r}
H3K9me3_Set2_data <- readRDS("data/RDS_files/H3K9me3_Set2_enrmotif_data.RDS")
H3K9me3_Set3_data <- readRDS("data/RDS_files/H3K9me3_Set3_enrmotif_data.RDS")

motif_names_set2H3K9me3 <-
H3K9me3_Set2_data %>%
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>%  
  dplyr::select(motif_name) %>% 
  mutate(TF_split=str_split(motif_name,"::")) %>% 
  unnest(TF_split) %>% 
  mutate(motif_name2=toupper(TF_split)) %>% 
  mutate(is_DAP_comp=motif_name2 %in%Entrezid_protein_module_unique$SYMBOL) %>% 
  group_by(motif_name) %>% 
  summarise(
    DAP = if_else(any(is_DAP_comp), "DAP", "non_DAP"),
    .groups = "drop"
  ) %>% 
  mutate(motif_name=toupper(motif_name)) %>% 
  distinct()


motif_names_set3H3K9me3 <- H3K9me3_Set3_data %>%
  dplyr::filter(str_starts(pattern = "MA",string=ID)) %>%  
  dplyr::select(motif_name) %>% 
  mutate(TF_split=str_split(motif_name,"::")) %>% 
  unnest(TF_split) %>% 
  mutate(motif_name2=toupper(TF_split)) %>% 
  mutate(is_DAP_comp=motif_name2 %in%Entrezid_protein_module_unique$SYMBOL) %>% 
  group_by(motif_name) %>% 
  summarise(
    DAP = if_else(any(is_DAP_comp), "DAP", "non_DAP"),
    .groups = "drop"
  ) %>% 
  mutate(motif_name=toupper(motif_name)) %>% 
  distinct()

H3K9me3_marked_TF <- tibble(tf_names) %>% 
  mutate(TF_split=str_split(tf_names,"::")) %>% 
  unnest(TF_split) %>% 
  mutate(is_DAP_comp=TF_split %in%Entrezid_protein_module_unique$SYMBOL) %>% 
   group_by(tf_names) %>% 
  summarise(
    DAP = if_else(any(is_DAP_comp), "DAP", "non_DAP"),
    .groups = "drop"
  ) %>% 
  mutate(Set_2=if_else(tf_names %in% motif_names_set2H3K9me3$motif_name,"yes","no"),
         Set_3=if_else(tf_names %in% motif_names_set3H3K9me3$motif_name,"yes","no")) %>% 
   mutate(
    DAP = factor(DAP, levels = c("DAP", "non_DAP")),
    Set_2 = factor(Set_2, levels = c("yes", "no")),
    Set_3 = factor(Set_3, levels = c("yes", "no"))) %>% 
  mutate(
    set_group = case_when(
      Set_2 == "yes" & Set_3 == "yes" ~ "shared",
      Set_2 == "yes" & Set_3 == "no"  ~ "set_2_only",
      Set_2 == "no"  & Set_3 == "yes" ~ "set_3_only",
      TRUE ~ "neither"))

table(H3K9me3_marked_TF$DAP,H3K9me3_marked_TF$Set_2)
 
tf_matrix_set2_k9 <- table(H3K9me3_marked_TF$DAP,H3K9me3_marked_TF$Set_2)

fisher.test(tf_matrix_set2_k9)


```


```{r}

table(H3K9me3_marked_TF$DAP,H3K9me3_marked_TF$Set_3)
 
tf_matrix_set3_k9 <- table(H3K9me3_marked_TF$DAP,H3K9me3_marked_TF$Set_3)

fisher.test(tf_matrix_set3_k9)



```
now I will also check if acute/sus/shared enrichment

Acute:(set 2 in H3K9me3)

```{r}
table(H3K9me3_marked_TF$DAP,H3K9me3_marked_TF$set_group=="set_2_only")
 
tf_uni_matrix_set2_k9 <- matrix(c(1,29,44,705), nrow = 2,byrow = TRUE)


fisher.test(tf_uni_matrix_set2_k9)
```
sustained in H3K9me3 (set 3)

```{r}
table(H3K9me3_marked_TF$DAP,H3K9me3_marked_TF$set_group=="set_3_only")
 
tf_uni_matrix_set3_k9 <- matrix(c(0,30,12,737), nrow = 2,byrow = TRUE)

fisher.test(tf_uni_matrix_set3_k9)
```
```{r h3K9me3 shared enr TFs}

table(H3K9me3_marked_TF$DAP,H3K9me3_marked_TF$set_group=="shared")
 
tf_uni_matrix_setshare_k9 <- matrix(c(4,26,169,580), nrow = 2,byrow = TRUE)

fisher.test(tf_uni_matrix_setshare_k9)
```
### fisher test results

```{r}

mat_list <- list("K27ac2"=tf_uni_matrix_set2,"K27ac3"=tf_uni_matrix_set3,"K27acshared"=tf_uni_matrix_setshare,"K9me32"=tf_uni_matrix_set2_k9,"K9me33"=tf_uni_matrix_set3_k9,"K9me3shared"=tf_uni_matrix_setshare_k9)

fisher_from_matrix <- function(mat, name = NA_character_, alternative = "greater") {
  
  ft <- fisher.test(mat, alternative = alternative)
  
  tibble(
    name = name,
    a = mat[1, 1],
    b = mat[1, 2],
    c = mat[2, 1],
    d = mat[2, 2],
    odds_ratio = unname(ft$estimate),
    conf_low = ft$conf.int[1],
    conf_high = ft$conf.int[2],
    p_value = ft$p.value,
    alternative = ft$alternative
  )
}

fisher_results <- imap_dfr(
  mat_list,
  ~ fisher_from_matrix(.x, name = .y, alternative = "two.sided")
) %>%
  mutate(
    FDR = p.adjust(p_value, method = "BH")
  )

```

```{r}

  kableExtra::kable(fisher_results) %>% 
  kableExtra::kable_paper("striped", full_width = FALSE) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "left",
    bootstrap_options = c("striped", "hover")
  ) %>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r pwmpull, eval=FALSE}
TF_Information <- read_delim("C:/Users/renee/Downloads/F135_3.00_2026_04_10_5_41_pm_allC2H2ZFs/TF_Information.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

kzpf_motif_IDs <- TF_Information %>% 
  dplyr::filter(TF_Species=="Homo_sapiens") %>% 
  dplyr::filter(TF_Name %in%KZFP_2026_list$Label) %>% 
  dplyr::filter(Motif_ID != ".") %>% 
  pull(Motif_ID)

TF_Information %>% 
  filter(TF_Species == "Homo_sapiens") %>% 
  filter(TF_Name %in% KZFP_2026_list$Label) %>% 
  summarise(
    total = dplyr::n(),
    with_motif = sum(Motif_ID != "."),
    missing = sum(Motif_ID == ".")
  )

KZFP_motif_info <- TF_Information %>% 
  filter(TF_Species == "Homo_sapiens") %>% 
  filter(TF_Name %in% KZFP_2026_list$Label) %>% 
   dplyr::filter(Motif_ID != ".") 
# saveRDS(KZFP_motif_info,"data/RDS_files/KZFP_motif_info.RDS")

KZFP_no_motif_list <- 
   TF_Information %>% 
  filter(TF_Species == "Homo_sapiens") %>% 
  filter(TF_Name %in% KZFP_2026_list$Label) %>% 
  filter(Motif_ID == ".") %>% 
  pull(TF_Name)

KZFP_no_motif_list

files <- list.files(
  "C:/Users/renee/Downloads/F135_3.00_2026_04_10_5_41_pm_allC2H2ZFs/pwms",
  full.names = TRUE)

file_ids <- tools::file_path_sans_ext(basename(files))

meta <- TF_Information %>%
  dplyr::filter(TF_Species == "Homo_sapiens") %>%
  dplyr::filter(TF_Name %in% KZFP_2026_list$Label) %>%
  dplyr::filter(Motif_ID != ".") %>%
  dplyr::select(Motif_ID, TF_Name)

motif_files <- files[file_ids %in% kzpf_motif_IDs]
motif_names <- tools::file_path_sans_ext(basename(motif_files))
alt_ids <- meta$TF_Name[match(motif_names, meta$Motif_ID)]

length(kzpf_motif_IDs)
length(motif_files)
setdiff(kzpf_motif_IDs,
        tools::file_path_sans_ext(basename(files)))

library(universalmotif)

get_tf_name <- function(x) {
  tools::file_path_sans_ext(basename(x))
}


bad <- lapply(motif_files, function(f) {

  df <- read.table(f, header = TRUE, sep = "\t", check.names = FALSE)

  cols <- intersect(c("A","C","G","T"), colnames(df))

  if (length(cols) != 4) {
    return(f)
  }

  mat <- df[, cols, drop = FALSE]
  mat <- data.matrix(mat)

  if (ncol(mat) == 0 || nrow(mat) == 0) {
    return(f)
  }

  NULL
})

bad <- unlist(bad)
bad


motifs <- lapply(motif_files, function(f) {
  
   df <- read.table(f, header = TRUE, sep = "\t", check.names = FALSE)

  # robust column matching (ignores whitespace/case issues)
  col_map <- colnames(df)
  nuc_cols <- col_map[tolower(col_map) %in% c("a","c","g","t")]

  if (length(nuc_cols) != 4) {
    warning(paste("Skipping file (bad columns):", f))
    return(NULL)
  }

  mat <- df[, nuc_cols, drop = FALSE]
  mat <- data.matrix(mat)

  if (nrow(mat) < 2 || ncol(mat) != 4) {
    warning(paste("Skipping file (bad matrix):", f))
    return(NULL)
  }

  pwm <- t(mat)

  if (ncol(pwm) == 0) {
    warning(paste("Skipping file (empty PWM):", f))
    return(NULL)
  }

  m <- universalmotif::create_motif(pwm, alphabet = "DNA")

  attr(m, "name") <- tools::file_path_sans_ext(basename(f))

  m
})


write_meme(motifs, file = "data/Other_paper_data/KZFP_motifs.meme", overwrite = TRUE)

motif_annotation <- data.frame(
  NAME = motif_names,
  ALT_ID = alt_ids
)
readLines("data/Other_paper_data/KZFP_motifs.meme", n = 40)
```
wanting to extract all SET1 SEt2 and Set3 regions that contain a MER61E element then test for enrichment.
```{r import regions, eval=FALSE} 

mer_gr <- rtracklayer::import("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed")
H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")


peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)

##assigning Peakid as name of summit region
 mcols(H3K27ac_summit_gr)$name <- mcols(H3K27ac_summit_gr)$Peakid
 
 H3K27ac_summit_gr_ext <- resize(H3K27ac_summit_gr, width(H3K27ac_summit_gr)+400, fix = "center")

H3K27ac_MER61E <- subsetByOverlaps(H3K27ac_summit_gr_ext, mer_gr)

H3K27ac_MER61E_df <- as.data.frame(H3K27ac_MER61E) 

Set1 <- H3K27ac_MER61E_df %>% filter(cluster == "Set_1")
Set2 <- H3K27ac_MER61E_df %>% filter(cluster == "Set_2")
Set3 <- H3K27ac_MER61E_df %>% filter(cluster == "Set_3")

Set1_gr <- makeGRangesFromDataFrame(Set1, keep.extra.columns = TRUE)
Set2_gr <- makeGRangesFromDataFrame(Set2, keep.extra.columns = TRUE)
Set3_gr <- makeGRangesFromDataFrame(Set3, keep.extra.columns = TRUE)

width(Set1_gr)
width(Set3_gr)


library(Biostrings)
library(BSgenome.Hsapiens.UCSC.hg38)
Hsapiens <- BSgenome.Hsapiens.UCSC.hg38

gc1 <- letterFrequency(getSeq(Hsapiens, Set1_gr), "GC", as.prob=TRUE)
gc3 <- letterFrequency(getSeq(Hsapiens, Set3_gr), "GC", as.prob=TRUE)

mean(gc1)
mean(gc3)

export(Set1_gr, "data/Bed_exports/H3K27ac_MER61E_Set_1.bed", format = "BED")
export(Set3_gr, "data/Bed_exports/H3K27ac_MER61E_Set_3.bed", format = "BED")
```

