---
title: "repeatmasker_data"
author: "Renee Matthews"
date: "2025-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

libraries
```{r library}
library(tidyverse)
library(DT)
library(genomation)
library(GenomicRanges)
library(BiocParallel)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(ChIPseeker)
library(plyranges)
```




Loading data files

```{r repeatmasker loading}

repeatmasker <- read_delim("data/Other_paper_data/repeatmasker_20250911.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

```

```{r loading motif data}
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")
peakAnnoList_H3K27me3 <- readRDS("data/motif_lists/H3K27me3_annotated_peaks.RDS")
peakAnnoList_H3K36me3 <- readRDS("data/motif_lists/H3K36me3_annotated_peaks.RDS")
peakAnnoList_H3K9me3 <- readRDS("data/motif_lists/H3K9me3_annotated_peaks.RDS")
```

```{r making granges dfs}


rmskr_std_gr <- repeatmasker %>% 
   makeGRangesFromDataFrame(., seqnames="genoName", keep.extra.columns=TRUE,
                                      start.field="genoStart",
                                      end.field="genoEnd",
                                      starts.in.df.are.0based=TRUE) %>% 
  keepStandardChromosomes(., pruning.mode = "coarse")

##checking chromosomes
seqlevels(rmskr_std_gr)

###making granges lists out of motifs

H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})

H3K27me3_sets_gr <- lapply(peakAnnoList_H3K27me3, function(df) {
  as_granges(df)
})

H3K36me3_sets_gr <- lapply(peakAnnoList_H3K36me3, function(df) {
  as_granges(df)
})

H3K9me3_sets_gr <- lapply(peakAnnoList_H3K9me3, function(df) {
  as_granges(df)
})
  
```

### H3K27ac and TEs
```{r H3K27ac set1}

### this is now taking the overlaps and widths of the TE and roi, then calculating percentage overlaps for each (pct_te_overlap and pct_roi_overlap)
H3K27ac_ols <- lapply(H3K27ac_sets_gr, function(df) {
    overlaps <- join_overlap_inner(df, rmskr_std_gr)
   # overlap base pairs
  overlaps$overlap_bp <- width(pintersect(overlaps))

  # ROI lengths
  overlaps$roi_len <- width(df)[match(overlaps$Peakid, df$Peakid)]
  overlaps$pct_roi_overlap <- (overlaps$overlap_bp / overlaps$roi_len) * 100

  # Repeat (TE) lengths
  overlaps$te_len <- width(rmskr_std_gr)[match(overlaps$genoName, rmskr_std_gr$genoName)]
  overlaps$pct_te_overlap <- (overlaps$overlap_bp / overlaps$te_len) * 100

  as.data.frame(overlaps)
})


H3K27ac_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27ac_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K27ac set2}


H3K27ac_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27ac_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K27ac set3}

H3K27ac_ols$Set_3 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27ac_ols$Set_3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```
#### All regions profile  

```{r H3K27ac all regions profile}
H3K27ac_ols$all_H3K27ac %>% 
   as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27ac_ols$all_H3K27ac %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))
length(H3K27ac_sets_gr$all_H3K27ac)
```
 

### H3K27me3 and TEs
```{r H3K27me3 set1}
H3K27me3_ols <- lapply(H3K27me3_sets_gr, function(df) {
  join_overlap_inner(df, rmskr_std_gr)
})

H3K27me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K27me3 set2}


H3K27me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```
#### All regions profile  

```{r H3K27me3 all regions profile}
H3K27me3_ols$all_H3K27me3 %>% 
   as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27me3_ols$all_H3K27me3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))
length(H3K27me3_sets_gr$all_H3K27me3)
```
 

### H3K36me3 and TEs
```{r H3K36me3 set1}
H3K36me3_ols <- lapply(H3K36me3_sets_gr, function(df) {
  join_overlap_inner(df, rmskr_std_gr)
})

H3K36me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K36me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K36me3 set2}


H3K36me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K36me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```
#### All regions profile  

```{r H3K36me3 all regions profile}
H3K36me3_ols$all_H3K36me3 %>% 
   as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K36me3_ols$all_H3K36me3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))
length(H3K36me3_sets_gr$all_H3K36me3)
```
 
### H3K9me3 and TEs
```{r H3K9me3 set1}
H3K9me3_ols <- lapply(H3K9me3_sets_gr, function(df) {
  join_overlap_inner(df, rmskr_std_gr)
})

H3K9me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K9me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K9me3 set2}


H3K9me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K9me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K9me3 set3}

H3K9me3_ols$Set_3 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() #%>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K9me3_ols$Set_3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

#### All regions profile  

```{r H3K9me3 all regions profile}
H3K9me3_ols$all_H3K9me3 %>% 
   as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K9me3_ols$all_H3K9me3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))
length(H3K9me3_sets_gr$all_H3K9me3)
```


### Counting TE types:

Below I am setting up the data frames, first I take all regions and annotate with the overlapping dataframe.  Then I make new columns, TE_status, SINE_status, LINE_status, LTR_status, DNA_status, SVA_status.  These columns mark Peakids if they have this attribute.  Last, I am annotating a cluster column to add which set the Peakid exists in by histone. (field will be Set1, Set2, Set3, no_cluster)
```{r histone types}

All_H3K27ac <- H3K27ac_sets_gr$all_H3K27ac %>% as.data.frame()
All_H3K27me3 <- H3K27me3_sets_gr$all_H3K27me3 %>% as.data.frame()
All_H3K36me3 <- H3K36me3_sets_gr$all_H3K36me3 %>% as.data.frame()
All_H3K9me3 <- H3K9me3_sets_gr$all_H3K9me3 %>% as.data.frame()


H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
H3K27me3_lookup <- imap_dfr(peakAnnoList_H3K27me3[1:2], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
H3K36me3_lookup <- imap_dfr(peakAnnoList_H3K36me3[1:2], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
H3K9me3_lookup <- imap_dfr(peakAnnoList_H3K9me3[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
  
annotated_H3K27ac <- All_H3K27ac %>% 
  left_join(., (H3K27ac_ols$all_H3K27ac %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass), collapse=";"),
                            repFamily=paste0(unique(repFamily),collapse=";"),
                            repName=paste0(unique(repName),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "not_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "not_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"not_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "not_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"not_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "not_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"not_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "not_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"not_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "not_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"not_SVA")) %>% 
  left_join(H3K27ac_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


annotated_H3K27me3 <- All_H3K27me3 %>% 
  left_join(., (H3K27me3_ols$all_H3K27me3 %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass), collapse=";"),
                            repFamily=paste0(unique(repFamily),collapse=";"),
                            repName=paste0(unique(repName),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "not_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "not_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"not_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "not_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"not_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "not_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"not_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "not_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"not_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "not_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"not_SVA")) %>% 
  left_join(H3K27me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


annotated_H3K36me3 <- All_H3K36me3 %>% 
  left_join(., (H3K36me3_ols$all_H3K36me3 %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass), collapse=";"),
                            repFamily=paste0(unique(repFamily),collapse=";"),
                            repName=paste0(unique(repName),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "not_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "not_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"not_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "not_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"not_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "not_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"not_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "not_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"not_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "not_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"not_SVA")) %>% 
  left_join(H3K36me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


annotated_H3K9me3 <- All_H3K9me3 %>% 
  left_join(., (H3K9me3_ols$all_H3K9me3 %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass), collapse=";"),
                            repFamily=paste0(unique(repFamily),collapse=";"),
                            repName=paste0(unique(repName),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "not_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "not_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"not_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "not_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"not_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "not_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"not_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "not_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"not_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "not_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"not_SVA")) %>% 
  left_join(H3K9me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))
           
```

```{r counting annotations}

te_cols <- c("TE_status","SINE_status","LINE_status","DNA_status","LTR_status","SVA_status")
H3K27ac_counts <- map(te_cols, ~ {
  annotated_H3K27ac %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)


H3K27ac_counts_long <- map2_dfr(
  H3K27ac_counts,      # the named list of wide tables
  names(H3K27ac_counts), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / not_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)
H3K27ac_counts_long <- H3K27ac_counts_long %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()


H3K27me3_counts <-  map(te_cols, ~ {
  annotated_H3K27me3 %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)

H3K27me3_counts_long <- map2_dfr(
  H3K27me3_counts,      # the named list of wide tables
  names(H3K27me3_counts), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / not_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)


H3K36me3_counts <-  map(te_cols, ~ {
  annotated_H3K36me3 %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)


H3K36me3_counts_long <- map2_dfr(
  H3K36me3_counts,      # the named list of wide tables
  names(H3K36me3_counts), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / not_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)



H3K9me3_counts <- map(te_cols, ~ {
  annotated_H3K9me3 %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)

H3K9me3_counts_long <- map2_dfr(
  H3K9me3_counts,      # the named list of wide tables
  names(H3K9me3_counts), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / not_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)

H3K9me3_counts_long <- H3K9me3_counts_long %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()
```
 
 
 making the function
```{r Fisher test comparisons function}

# function to test enrichment for one TE type, one cluster pair
test_pair_TE <- function(df_long, te_name, cluster1, cluster2) {
  
  # filter for TE type
  sub_df <- df_long %>% filter(TE_type == te_name)
  
  # get counts for cluster1 and cluster2
  c1 <- sub_df %>% dplyr::filter(cluster == cluster1)
  c2 <- sub_df %>% dplyr::filter(cluster == cluster2)
  
  # 2x2 table: rows = clusters, cols = TE / not_TE
  mat <- matrix(
    c(c2$count, sum(c2$count) - c2$count,
      c1$count, sum(c1$count) - c1$count),
    nrow = 2,
    byrow = TRUE,
    dimnames = list(
      cluster = c(cluster2, cluster1),
      category = c("TE", "not_TE")
    )
  )
  
  ft <- fisher.test(mat)
  
  tibble(
    TE_type     = te_name,
    comparison  = paste(cluster2, "vs", cluster1),
    odds_ratio  = ft$estimate,
    lower_CI    = ft$conf.int[1],
    upper_CI    = ft$conf.int[2],
    p_value     = ft$p.value
  )
}


# Generic pairwise Fisher test
test_pair_TE_generic <- function(df_long, te_name, cluster1, cluster2) {
  
  sub_df <- df_long %>% filter(TE_type == te_name)
  
  # assume "status" column has TE vs not_TE automatically
  statuses <- unique(sub_df$status)
  
  if(length(statuses) != 2) {
    # ensure we have exactly two categories, fill missing with 0
    sub_df <- sub_df %>%
      complete(cluster, status, fill = list(count = 0))
    statuses <- unique(sub_df$status)
  }
  
  # extract counts for cluster1
  c1_counts <- sub_df %>%
    filter(cluster == cluster1) %>%
    arrange(match(status, statuses)) %>%   # ensure same order
    pull(count)
  
  # extract counts for cluster2
  c2_counts <- sub_df %>%
    filter(cluster == cluster2) %>%
    arrange(match(status, statuses)) %>%
    pull(count)
  
  # build 2x2 matrix
  mat <- matrix(
    c(c2_counts, c1_counts),
    nrow = 2,
    byrow = TRUE,
    dimnames = list(
      cluster = c(cluster2, cluster1),
      category = statuses
    )
  )
  
  ft <- fisher.test(mat)
  
  tibble(
    TE_type     = te_name,
    comparison  = paste(cluster2, "vs", cluster1),
    odds_ratio  = ft$estimate,
    lower_CI    = ft$conf.int[1],
    upper_CI    = ft$conf.int[2],
    p_value     = ft$p.value
  )
}
```

applying to dataframes


```{r}
TE_types <- te_cols

cluster_pairs_3 <- list(
  c("Set_2","Set_1"),
  c("Set_3","Set_1")
  # add more pairs if needed or
)
cluster_pairs_2 <- list(
  c("Set_2","Set_1"))


H3K27ac_results <- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_3, function(pair){
    test_pair_TE_generic(H3K27ac_counts_long, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})

H3K27ac_results
###

```

```{r}
# test results
H3K27ac_counts_long %>% 
  dplyr::filter(TE_type=="LINE_status") %>% 
  dplyr::filter(cluster == "Set_1"| cluster == "Set_2") %>% 
  pivot_wider(id_cols = cluster, names_from = status, values_from = count) %>% 
  column_to_rownames("cluster") %>% 
  as.matrix() %>% 
  fisher.test()

H3K27ac_counts_long %>% 
  dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=count, fill=status)) +
  geom_bar(stat="identity", position = "fill") +
  facet_wrap(~TE_type, scales="free_y")
  

```

```{r}
H3K27ac_counts_long %>% 
  dplyr::filter(cluster != "not_assigned") %>%
  # create a clean TE_type by removing "not_" and "_status"
  mutate(
    TE_type_clean = str_remove(TE_type, "_status"),          # remove "_status"
    TE_type_clean = str_remove(TE_type_clean, "^not_"),      # remove leading "not_"
    status = ifelse(str_starts(status, "not_"), "not_TE", "TE") # unify status for plotting
  ) %>%
  ggplot(aes(x = cluster, y = count, fill = status)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~TE_type_clean, scales = "free_y") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Cluster",
    y = "Proportion",
    fill = "TE Status",
    title = "Proportion of TE types across clusters"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```
```{r}
H3K27ac_counts_long_clean <- H3K27ac_counts_long %>%
  filter(cluster != "not_assigned") %>%
  # unify TE_type and remove "not_" and "_status"
  mutate(
    TE_type_clean = str_remove(TE_type, "_status"),        # remove "_status"
    TE_type_clean = str_remove(TE_type_clean, "^not_"),   # remove leading "not_"
    status_clean = ifelse(str_starts(status, "not_"), "not_TE", "TE"),  # unify status
    # ensure TE comes first in factor
    status_clean = factor(status_clean, levels = c("TE", "not_TE")),
    # optional: order TE_type facets
    TE_type_clean = factor(TE_type_clean, levels = c("TE","LINE","SINE","DNA","LTR","SVA"))
  ) %>%
  # ensure every cluster × TE_type × status combination exists
  complete(cluster, TE_type_clean, status_clean, fill = list(count = 0))

# 2️⃣ Plot proportions
ggplot(H3K27ac_counts_long_clean, aes(x = cluster, y = count, fill = status_clean)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~TE_type_clean, scales = "free_y") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Cluster",
    y = "Proportion",
    fill = "TE Status",
    title = "Proportion of TE types across clusters"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

