---
title: "repeatmasker_data"
author: "Renee Matthews"
date: "2025-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 400px;
}
```

libraries
```{r library}
library(tidyverse)
library(DT)
library(genomation)
library(GenomicRanges)
library(BiocParallel)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(ChIPseeker)
library(plyranges)
library(ggsignif)
library(ggpubr)
```




Loading data files

```{r repeatmasker loading}

repeatmasker <- read_delim("data/Other_paper_data/repeatmasker_20250911.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

```

```{r loading motif data}
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")
peakAnnoList_H3K27me3 <- readRDS("data/motif_lists/H3K27me3_annotated_peaks.RDS")
peakAnnoList_H3K36me3 <- readRDS("data/motif_lists/H3K36me3_annotated_peaks.RDS")
peakAnnoList_H3K9me3 <- readRDS("data/motif_lists/H3K9me3_annotated_peaks.RDS")
```

```{r making granges dfs}


rmskr_std_gr <- repeatmasker %>% 
   makeGRangesFromDataFrame(., seqnames="genoName", keep.extra.columns=TRUE,
                                      start.field="genoStart",
                                      end.field="genoEnd",
                                      starts.in.df.are.0based=TRUE) %>% 
  keepStandardChromosomes(., pruning.mode = "coarse")

##checking chromosomes
#seqlevels(rmskr_std_gr)

###making granges lists out of motifs

H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})

H3K27me3_sets_gr <- lapply(peakAnnoList_H3K27me3, function(df) {
  as_granges(df)
})

H3K36me3_sets_gr <- lapply(peakAnnoList_H3K36me3, function(df) {
  as_granges(df)
})

H3K9me3_sets_gr <- lapply(peakAnnoList_H3K9me3, function(df) {
  as_granges(df)
})
  
```

### H3K27ac and TEs
```{r H3K27ac set1}

## this is now taking the overlaps and widths of the TE and roi, then calculating percentage overlaps for each (pct_te_overlap and pct_roi_overlap)
# H3K27ac_ols <- lapply(H3K27ac_sets_gr, function(df) {
#     overlaps <- join_overlap_inner(df, rmskr_std_gr)
#    # overlap base pairs
#   })
H3K27ac_ols <- lapply(H3K27ac_sets_gr, function(df) {
  
   # 1️⃣ find overlaps
  hits <- findOverlaps(df, rmskr_std_gr)
  
  # 2️⃣ intersect the overlapping ranges
  overlaps_gr <- pintersect(df[queryHits(hits)], rmskr_std_gr[subjectHits(hits)])
  
  # 3️⃣ convert to data frame
  overlaps_df <- as.data.frame(overlaps_gr)
  
  # 4️⃣ add overlap length
  overlaps_df$overlap_bp <- width(overlaps_gr)
  
  # 5️⃣ ROI lengths & % coverage
  roi_widths <- width(df)
  names(roi_widths) <- df$Peakid
  overlaps_df$roi_len <- roi_widths[as.character(overlaps_df$Peakid)]
  overlaps_df$pct_roi_overlap <- (overlaps_df$overlap_bp / overlaps_df$roi_len) * 100
  
  # 6️⃣ TE lengths & % coverage
  te_widths <- width(rmskr_std_gr)
  overlaps_df$te_len <- te_widths[subjectHits(hits)]
  overlaps_df$pct_te_overlap <- (overlaps_df$overlap_bp / overlaps_df$te_len) * 100
  
  # 7️⃣ Add TE metadata from rmsk GRanges
  overlaps_df$repName   <- rmskr_std_gr$repName[subjectHits(hits)]
  overlaps_df$repClass  <- rmskr_std_gr$repClass[subjectHits(hits)]
  overlaps_df$repFamily <- rmskr_std_gr$repFamily[subjectHits(hits)]
  overlaps_df$milliDiv <- rmskr_std_gr$milliDiv[subjectHits(hits)]
  overlaps_df$milliDel <- rmskr_std_gr$milliDel[subjectHits(hits)]
  overlaps_df$milliIns <- rmskr_std_gr$milliIns[subjectHits(hits)]
  
  overlaps_df
})


H3K27ac_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27ac_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K27ac set2}


H3K27ac_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27ac_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K27ac set3}

H3K27ac_ols$Set_3 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27ac_ols$Set_3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```
#### All regions profile  

```{r H3K27ac all regions profile}
H3K27ac_ols$all_H3K27ac %>% 
   as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27ac_ols$all_H3K27ac %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))
length(H3K27ac_sets_gr$all_H3K27ac)
```
 

### H3K27me3 and TEs
```{r H3K27me3 set1}
H3K27me3_ols <- lapply(H3K27me3_sets_gr, function(df) {
   # 1️⃣ find overlaps
  hits <- findOverlaps(df, rmskr_std_gr)
  
  # 2️⃣ intersect the overlapping ranges
  overlaps_gr <- pintersect(df[queryHits(hits)], rmskr_std_gr[subjectHits(hits)])
  
  # 3️⃣ convert to data frame
  overlaps_df <- as.data.frame(overlaps_gr)
  
  # 4️⃣ add overlap length
  overlaps_df$overlap_bp <- width(overlaps_gr)
  
  # 5️⃣ ROI lengths & % coverage
  roi_widths <- width(df)
  names(roi_widths) <- df$Peakid
  overlaps_df$roi_len <- roi_widths[as.character(overlaps_df$Peakid)]
  overlaps_df$pct_roi_overlap <- (overlaps_df$overlap_bp / overlaps_df$roi_len) * 100
  
  # 6️⃣ TE lengths & % coverage
  te_widths <- width(rmskr_std_gr)
  overlaps_df$te_len <- te_widths[subjectHits(hits)]
  overlaps_df$pct_te_overlap <- (overlaps_df$overlap_bp / overlaps_df$te_len) * 100
  
  # 7️⃣ Add TE metadata from rmsk GRanges
  overlaps_df$repName   <- rmskr_std_gr$repName[subjectHits(hits)]
  overlaps_df$repClass  <- rmskr_std_gr$repClass[subjectHits(hits)]
  overlaps_df$repFamily <- rmskr_std_gr$repFamily[subjectHits(hits)]
  overlaps_df$milliDiv <- rmskr_std_gr$milliDiv[subjectHits(hits)]
  overlaps_df$milliDel <- rmskr_std_gr$milliDel[subjectHits(hits)]
  overlaps_df$milliIns <- rmskr_std_gr$milliIns[subjectHits(hits)]
  
  overlaps_df
})


H3K27me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K27me3 set2}


H3K27me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```
#### All regions profile  

```{r H3K27me3 all regions profile}
H3K27me3_ols$all_H3K27me3 %>% 
   as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K27me3_ols$all_H3K27me3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))
length(H3K27me3_sets_gr$all_H3K27me3)
```
 

### H3K36me3 and TEs
```{r H3K36me3 set1}
H3K36me3_ols <- lapply(H3K36me3_sets_gr, function(df) {
  # 1️⃣ find overlaps
  hits <- findOverlaps(df, rmskr_std_gr)
  
  # 2️⃣ intersect the overlapping ranges
  overlaps_gr <- pintersect(df[queryHits(hits)], rmskr_std_gr[subjectHits(hits)])
  
  # 3️⃣ convert to data frame
  overlaps_df <- as.data.frame(overlaps_gr)
  
  # 4️⃣ add overlap length
  overlaps_df$overlap_bp <- width(overlaps_gr)
  
  # 5️⃣ ROI lengths & % coverage
  roi_widths <- width(df)
  names(roi_widths) <- df$Peakid
  overlaps_df$roi_len <- roi_widths[as.character(overlaps_df$Peakid)]
  overlaps_df$pct_roi_overlap <- (overlaps_df$overlap_bp / overlaps_df$roi_len) * 100
  
  # 6️⃣ TE lengths & % coverage
  te_widths <- width(rmskr_std_gr)
  overlaps_df$te_len <- te_widths[subjectHits(hits)]
  overlaps_df$pct_te_overlap <- (overlaps_df$overlap_bp / overlaps_df$te_len) * 100
  
  # 7️⃣ Add TE metadata from rmsk GRanges
  overlaps_df$repName   <- rmskr_std_gr$repName[subjectHits(hits)]
  overlaps_df$repClass  <- rmskr_std_gr$repClass[subjectHits(hits)]
  overlaps_df$repFamily <- rmskr_std_gr$repFamily[subjectHits(hits)]
  overlaps_df$milliDiv <- rmskr_std_gr$milliDiv[subjectHits(hits)]
  overlaps_df$milliDel <- rmskr_std_gr$milliDel[subjectHits(hits)]
  overlaps_df$milliIns <- rmskr_std_gr$milliIns[subjectHits(hits)]
  
  overlaps_df
})

H3K36me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K36me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K36me3 set2}


H3K36me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K36me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```
#### All regions profile  

```{r H3K36me3 all regions profile}
H3K36me3_ols$all_H3K36me3 %>% 
   as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K36me3_ols$all_H3K36me3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))
length(H3K36me3_sets_gr$all_H3K36me3)
```
 
### H3K9me3 and TEs
```{r H3K9me3 set1}
H3K9me3_ols <- lapply(H3K9me3_sets_gr, function(df) {
  # 1️⃣ find overlaps
  hits <- findOverlaps(df, rmskr_std_gr)
  
  # 2️⃣ intersect the overlapping ranges
  overlaps_gr <- pintersect(df[queryHits(hits)], rmskr_std_gr[subjectHits(hits)])
  
  # 3️⃣ convert to data frame
  overlaps_df <- as.data.frame(overlaps_gr)
  
  # 4️⃣ add overlap length
  overlaps_df$overlap_bp <- width(overlaps_gr)
  
  # 5️⃣ ROI lengths & % coverage
  roi_widths <- width(df)
  names(roi_widths) <- df$Peakid
  overlaps_df$roi_len <- roi_widths[as.character(overlaps_df$Peakid)]
  overlaps_df$pct_roi_overlap <- (overlaps_df$overlap_bp / overlaps_df$roi_len) * 100
  
  # 6️⃣ TE lengths & % coverage
  te_widths <- width(rmskr_std_gr)
  overlaps_df$te_len <- te_widths[subjectHits(hits)]
  overlaps_df$pct_te_overlap <- (overlaps_df$overlap_bp / overlaps_df$te_len) * 100
  
  # 7️⃣ Add TE metadata from rmsk GRanges
  overlaps_df$repName   <- rmskr_std_gr$repName[subjectHits(hits)]
  overlaps_df$repClass  <- rmskr_std_gr$repClass[subjectHits(hits)]
  overlaps_df$repFamily <- rmskr_std_gr$repFamily[subjectHits(hits)]
  overlaps_df$milliDiv <- rmskr_std_gr$milliDiv[subjectHits(hits)]
  overlaps_df$milliDel <- rmskr_std_gr$milliDel[subjectHits(hits)]
  overlaps_df$milliIns <- rmskr_std_gr$milliIns[subjectHits(hits)]
  
  overlaps_df
})

H3K9me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K9me3_ols$Set_1 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K9me3 set2}


H3K9me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K9me3_ols$Set_2 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

```{r H3K9me3 set3}

H3K9me3_ols$Set_3 %>% 
  as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K9me3_ols$Set_3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))

```

#### All regions profile  

```{r H3K9me3 all regions profile}
H3K9me3_ols$all_H3K9me3 %>% 
   as.data.frame() %>% 
  group_by(Peakid,repClass) %>% 
  summarise() %>% 
  group_by(Peakid) %>% tally %>% 
  group_by(n) %>% tally

H3K9me3_ols$all_H3K9me3 %>% 
  as.data.frame() %>% 
  group_by(Peakid) %>%
  summarize(., 
            repClass=paste0(unique(repClass), collapse=";"))
length(H3K9me3_sets_gr$all_H3K9me3)
```


### Counting TE types:

Below I am setting up the data frames, first I take all regions and annotate with the overlapping dataframe.  Then I make new columns, TE_status, SINE_status, LINE_status, LTR_status, DNA_status, SVA_status.  These columns mark Peakids if they have this attribute.  Last, I am annotating a cluster column to add which set the Peakid exists in by histone. (field will be Set1, Set2, Set3, not_assigned)
```{r histone types 1 bp}

All_H3K27ac <- H3K27ac_sets_gr$all_H3K27ac %>% as.data.frame()
All_H3K27me3 <- H3K27me3_sets_gr$all_H3K27me3 %>% as.data.frame()
All_H3K36me3 <- H3K36me3_sets_gr$all_H3K36me3 %>% as.data.frame()
All_H3K9me3 <- H3K9me3_sets_gr$all_H3K9me3 %>% as.data.frame()


H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
H3K27me3_lookup <- imap_dfr(peakAnnoList_H3K27me3[1:2], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
H3K36me3_lookup <- imap_dfr(peakAnnoList_H3K36me3[1:2], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
H3K9me3_lookup <- imap_dfr(peakAnnoList_H3K9me3[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
  
annotated_H3K27ac <- All_H3K27ac %>% 
  left_join(., (H3K27ac_ols$all_H3K27ac %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass), collapse=";"),
                            repFamily=paste0(unique(repFamily),collapse=";"),
                            repName=paste0(unique(repName),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
  left_join(H3K27ac_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


annotated_H3K27me3 <- All_H3K27me3 %>% 
  left_join(., (H3K27me3_ols$all_H3K27me3 %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass), collapse=";"),
                            repFamily=paste0(unique(repFamily),collapse=";"),
                            repName=paste0(unique(repName),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
  left_join(H3K27me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


annotated_H3K36me3 <- All_H3K36me3 %>% 
  left_join(., (H3K36me3_ols$all_H3K36me3 %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass), collapse=";"),
                            repFamily=paste0(unique(repFamily),collapse=";"),
                            repName=paste0(unique(repName),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
  left_join(H3K36me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


annotated_H3K9me3 <- All_H3K9me3 %>% 
  left_join(., (H3K9me3_ols$all_H3K9me3 %>% 
                  as.data.frame() %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass), collapse=";"),
                            repFamily=paste0(unique(repFamily),collapse=";"),
                            repName=paste0(unique(repName),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
  left_join(H3K9me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))

# annotated_tables <- list("annotated_H3K27ac"=annotated_H3K27ac,
#                          "annotated_H3K27me3"=annotated_H3K27me3,
#                          "annotated_H3K36me3"=annotated_H3K36me3,
#                          "annotated_H3K9me3"=annotated_H3K9me3)
# 
# saveRDS(annotated_tables,"data/TE_annotation/annotated_tables_1bp_ol.RDS")
``` 


## Counts of regions in each set
```{r}
H3K27ac_data <- data.frame("total"=length(All_H3K27ac$Peakid),"Set1"=length(H3K27ac_lookup$cluster[H3K27ac_lookup$cluster=="Set_1"]),"Set2"=length(H3K27ac_lookup$cluster[H3K27ac_lookup$cluster=="Set_2"]),"Set3"=length(H3K27ac_lookup$cluster[H3K27ac_lookup$cluster=="Set_3"]))

H3K9me3_data <- data.frame("total"=length(All_H3K9me3$Peakid),"Set1"=length(H3K9me3_lookup$cluster[H3K9me3_lookup$cluster=="Set_1"]),"Set2"=length(H3K9me3_lookup$cluster[H3K9me3_lookup$cluster=="Set_2"]),"Set3"=length(H3K9me3_lookup$cluster[H3K9me3_lookup$cluster=="Set_3"]))

H3K27me3_data <- data.frame("total"=length(All_H3K27me3$Peakid),"Set1"=length(H3K27me3_lookup$cluster[H3K27me3_lookup$cluster=="Set_1"]),"Set2"=length(H3K27me3_lookup$cluster[H3K27me3_lookup$cluster=="Set_2"]),"Set3"=length(H3K27me3_lookup$cluster[H3K27me3_lookup$cluster=="Set_3"]))

H3K36me3_data <- data.frame("total"=length(All_H3K36me3$Peakid),"Set1"=length(H3K36me3_lookup$cluster[H3K36me3_lookup$cluster=="Set_1"]),"Set2"=length(H3K36me3_lookup$cluster[H3K36me3_lookup$cluster=="Set_2"]),"Set3"=length(H3K36me3_lookup$cluster[H3K36me3_lookup$cluster=="Set_3"]))

Region_dataframe <- bind_rows(list(H3K27ac=H3K27ac_data, H3K27me3=H3K27me3_data, H3K36me3=H3K36me3_data, H3K9me3=H3K9me3_data), .id="histone")

datatable(Region_dataframe)
```
           


50% counting histones
```{r histone types 50per}


annotated_H3K27ac_50_per <- All_H3K27ac %>%
  left_join(., (H3K27ac_ols$all_H3K27ac %>% 
                  as.data.frame() %>%
                  dplyr::filter(pct_te_overlap>50) %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass[pct_te_overlap>50]), collapse=";"),
                            repFamily=paste0(unique(repFamily[pct_te_overlap>50]),collapse=";"),
                            repName=paste0(unique(repName[pct_te_overlap>50]),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
  left_join(H3K27ac_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


annotated_H3K27me3_50_per <- All_H3K27me3 %>% 
  left_join(., (H3K27me3_ols$all_H3K27me3 %>% 
                  as.data.frame() %>%
                  dplyr::filter(pct_te_overlap>50) %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass[pct_te_overlap>50]), collapse=";"),
                            repFamily=paste0(unique(repFamily[pct_te_overlap>50]),collapse=";"),
                            repName=paste0(unique(repName[pct_te_overlap>50]),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
  left_join(H3K27me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


annotated_H3K36me3_50_per <- All_H3K36me3 %>% 
  left_join(., (H3K36me3_ols$all_H3K36me3 %>% 
                  as.data.frame() %>% 
                  dplyr::filter(pct_te_overlap>50) %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass[pct_te_overlap>50]), collapse=";"),
                            repFamily=paste0(unique(repFamily[pct_te_overlap>50]),collapse=";"),
                            repName=paste0(unique(repName[pct_te_overlap>50]),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
  left_join(H3K36me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


annotated_H3K9me3_50_per <- All_H3K9me3 %>% 
  left_join(., (H3K9me3_ols$all_H3K9me3 %>% 
                  as.data.frame() %>% 
                  dplyr::filter(pct_te_overlap>50) %>% 
                  group_by(Peakid) %>% 
                  summarize(., 
                            repClass=paste0(unique(repClass[pct_te_overlap>50]), collapse=";"),
                            repFamily=paste0(unique(repFamily[pct_te_overlap>50]),collapse=";"),
                            repName=paste0(unique(repName[pct_te_overlap>50]),collapse=";"))))%>%
              mutate(TE_status = if_else(is.na(repClass),   "wnot_TE","TE")) %>%
              mutate(SINE_status = case_when(is.na(repClass)~ "wnot_SINE",
                                             str_detect(repClass, "SINE") ~ "SINE",
                                             TRUE ~"wnot_SINE")) %>% 
              mutate(LINE_status = case_when(is.na(repClass)~ "wnot_LINE",
                                             str_detect(repClass, "LINE") ~ "LINE",
                                             TRUE ~"wnot_LINE")) %>%
              mutate(LTR_status = case_when(is.na(repClass)~ "wnot_LTR",
                                            str_detect(repClass, "LTR") ~ "LTR",
                                            TRUE ~"wnot_LTR")) %>%
               mutate(DNA_status = case_when(is.na(repClass)~ "wnot_DNA",
                                 str_detect(repClass, "DNA") ~ "DNA",
                                 TRUE ~"wnot_DNA")) %>% 
          mutate(SVA_status = case_when(is.na(repClass)~ "wnot_SVA",
                                        str_detect(repClass, "Retroposon") ~ "SVA",
                                        TRUE ~"wnot_SVA")) %>% 
  left_join(H3K9me3_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))


# annotated_tables_50_per <- list("annotated_H3K27ac_50_per"=annotated_H3K27ac_50_per,
#                          "annotated_H3K27me3_50_per"=annotated_H3K27me3_50_per,
#                          "annotated_H3K36me3_50_per"=annotated_H3K36me3_50_per,
#                          "annotated_H3K9me3_50_per"=annotated_H3K9me3_50_per)
# 
# saveRDS(annotated_tables_50_per,"data/TE_annotation/annotated_tables_50per_ol.RDS")
           
```



```{r counting annotations 1bp}

te_cols <- c("TE_status","SINE_status","LINE_status","DNA_status","LTR_status","SVA_status")
H3K27ac_counts <- map(te_cols, ~ {
  annotated_H3K27ac %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)


H3K27ac_counts_long <- map2_dfr(
  H3K27ac_counts,      # the named list of wide tables
  names(H3K27ac_counts), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / wnot_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)
H3K27ac_counts_long <- H3K27ac_counts_long %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()


H3K27me3_counts <-  map(te_cols, ~ {
  annotated_H3K27me3 %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)

H3K27me3_counts_long <- map2_dfr(
  H3K27me3_counts,      # the named list of wide tables
  names(H3K27me3_counts), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / not_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)

H3K27me3_counts_long <- H3K27me3_counts_long %>%
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()

H3K36me3_counts <-  map(te_cols, ~ {
  annotated_H3K36me3 %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)


H3K36me3_counts_long <- map2_dfr(
  H3K36me3_counts,      # the named list of wide tables
  names(H3K36me3_counts), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / not_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)

H3K36me3_counts_long <- H3K36me3_counts_long %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()

H3K9me3_counts <- map(te_cols, ~ {
  annotated_H3K9me3 %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)

H3K9me3_counts_long <- map2_dfr(
  H3K9me3_counts,      # the named list of wide tables
  names(H3K9me3_counts), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / wnot_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)

H3K9me3_counts_long <- H3K9me3_counts_long %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()
```

```{r counting annotations 50per}

te_cols <- c("TE_status","SINE_status","LINE_status","DNA_status","LTR_status","SVA_status")
H3K27ac_counts_50per <- map(te_cols, ~ {
  annotated_H3K27ac_50_per %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)


H3K27ac_counts_long_50per <- map2_dfr(
  H3K27ac_counts_50per,      # the named list of wide tables
  names(H3K27ac_counts_50per), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / wnot_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)
H3K27ac_counts_long_50per <- H3K27ac_counts_long_50per %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()


H3K27me3_counts_50per <-  map(te_cols, ~ {
  annotated_H3K27me3_50_per %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)

H3K27me3_counts_long_50per <- map2_dfr(
  H3K27me3_counts_50per,      # the named list of wide tables
  names(H3K27me3_counts_50per), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / not_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)

H3K27me3_counts_long_50per <- H3K27me3_counts_long_50per %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()

H3K36me3_counts_50per <-  map(te_cols, ~ {
  annotated_H3K36me3_50_per %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)


H3K36me3_counts_long_50per <- map2_dfr(
  H3K36me3_counts_50per,      # the named list of wide tables
  names(H3K36me3_counts_50per), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / not_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)

H3K36me3_counts_long_50per <- H3K36me3_counts_long_50per %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()

H3K9me3_counts_50per <- map(te_cols, ~ {
  annotated_H3K9me3_50_per %>%
    group_by(cluster, !!sym(.x)) %>%  # dynamically use column
    tally() %>%
    pivot_wider(
      id_cols = cluster,
      names_from = !!sym(.x),
      values_from = n
    )
}) %>% set_names(te_cols)

H3K9me3_counts_long_50per <- map2_dfr(
  H3K9me3_counts_50per,      # the named list of wide tables
  names(H3K9me3_counts_50per), # names of TE columns
  ~ .x %>%
    pivot_longer(
      cols = -cluster,       # everything except 'cluster'
      names_to = "status",   # new column for TE status (e.g., TE / wnot_TE)
      values_to = "count"
    ) %>%
    mutate(TE_type = .y)      # add TE_type column based on list element name
)

H3K9me3_counts_long_50per <- H3K9me3_counts_long_50per %>% 
group_by(TE_type) %>%
  tidyr::complete(cluster, status, fill = list(count = 0)) %>%
  ungroup()
```
 
 
 
 making the function
```{r Fisher test comparisons function}

# # function to test enrichment for one TE type, one cluster pair
# test_pair_TE <- function(df_long, te_name, cluster1, cluster2) {
#   
#   # filter for TE type
#   sub_df <- df_long %>% filter(TE_type == te_name)
#   
#   # get counts for cluster1 and cluster2
#   c1 <- sub_df %>% dplyr::filter(cluster == cluster1)
#   c2 <- sub_df %>% dplyr::filter(cluster == cluster2)
#   
#   # 2x2 table: rows = clusters, cols = TE / wnot_TE
#   mat <- matrix(
#     c(c2$count, sum(c2$count) - c2$count,
#       c1$count, sum(c1$count) - c1$count),
#     nrow = 2,
#     byrow = TRUE,
#     dimnames = list(
#       cluster = c(cluster2, cluster1),
#       category = c("TE", "wnot_TE")
#     )
#   )
#   
#   ft <- fisher.test(mat)
#   
#   tibble(
#     TE_type     = te_name,
#     comparison  = paste(cluster2, "vs", cluster1),
#     odds_ratio  = ft$estimate,
#     lower_CI    = ft$conf.int[1],
#     upper_CI    = ft$conf.int[2],
#     p_value     = ft$p.value
#   )
# }


# Generic pairwise Fisher test
test_pair_TE_generic <- function(df_long, te_name, cluster1, cluster2) {
  
  sub_df <- df_long %>% filter(TE_type == te_name)
  
  # assume "status" column has TE vs wnot_TE automatically
  statuses <- unique(sub_df$status)
  
  if(length(statuses) != 2) {
    # ensure we have exactly two categories, fill missing with 0
    sub_df <- sub_df %>%
      complete(cluster, status, fill = list(count = 0))
    statuses <- unique(sub_df$status)
  }
  
  # extract counts for cluster1
  c1_counts <- sub_df %>%
    filter(cluster == cluster1) %>%
    arrange(match(status, statuses)) %>%   # ensure same order
    pull(count)
  
  # extract counts for cluster2
  c2_counts <- sub_df %>%
    filter(cluster == cluster2) %>%
    arrange(match(status, statuses)) %>%
    pull(count)
  
  # build 2x2 matrix
  mat <- matrix(
    c(c2_counts, c1_counts),
    nrow = 2,
    byrow = TRUE,
    dimnames = list(
      cluster = c(cluster2, cluster1),
      category = statuses
    )
  )
  
  ft <- tryCatch(
    fisher.test(mat, workspace = 2e8),
    error = function(e) fisher.test(mat, simulate.p.value = TRUE, B = 1e5)
  )
  
  tibble(
    TE_type     = te_name,
    comparison  = paste(cluster2, "vs", cluster1),
    odds_ratio  = ft$estimate,
    lower_CI    = ft$conf.int[1],
    upper_CI    = ft$conf.int[2],
    p_value     = ft$p.value
  )
}
```

applying to dataframes

### H3K27ac proportions  
##### results for 1bp overlap H3K27ac  
```{r H3K27ac df}
TE_types <- te_cols

cluster_pairs_3 <- list(
  c("Set_2","Set_1"),
  c("Set_3","Set_1")
  # add more pairs if needed or
)
cluster_pairs_2 <- list(
  c("Set_2","Set_1"))


H3K27ac_results_1bp <- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_3, function(pair){
    test_pair_TE_generic(H3K27ac_counts_long, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})

htmltools::tagList(
  htmltools::h3("H3K27ac odds ratio 1bp"), 
datatable(H3K27ac_results_1bp,options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE))



H3K27ac_counts_long %>% 
  dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=count, fill=status)) +
  geom_bar(stat="identity", position = "fill") +
  facet_wrap(~TE_type, scales="free_y")+
  ggtitle("H3K27ac 1 bp overlap enrichment")+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

##### results for 50% TE overlap H3K27ac 
```{r H3K27ac 50 per df}
###
H3K27ac_results_50per <- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_3, function(pair){
    test_pair_TE_generic(H3K27ac_counts_long_50per, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})

htmltools::tagList(
  htmltools::h3("H3K27ac odds ratio >50% overlap results"),
datatable(H3K27ac_results_50per,
          options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE))


H3K27ac_counts_long_50per %>% 
  dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=count, fill=status)) +
  geom_bar(stat="identity", position = "fill") +
  facet_wrap(~TE_type, scales="free_y")+
  ggtitle("H3K27ac 50% TE overlap enrichment")+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
  

```



### H3K27me3 proportions  
##### results for 1bp overlap H3K27me3  
```{r H3K27me3 df}


H3K27me3_results_1bp <- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_2, function(pair){
    test_pair_TE_generic(H3K27me3_counts_long, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})

htmltools::tagList(
  htmltools::h3("H3K27me3 odds ratio 1bp"), 
datatable(H3K27me3_results_1bp,options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE))



H3K27me3_counts_long %>% 
  dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=count, fill=status)) +
  geom_bar(stat="identity", position = "fill") +
  facet_wrap(~TE_type, scales="free_y")+
  ggtitle("H3K27me3 1 bp overlap enrichment")+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

##### results for 50% TE overlap H3K27me3 
```{r H3K27me3 50 per df}
###
H3K27me3_results_50per <- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_2, function(pair){
    test_pair_TE_generic(H3K27me3_counts_long_50per, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})

htmltools::tagList(
  htmltools::h3("H3K27me3 odds ratio >50% overlap results"),
datatable(H3K27me3_results_50per,options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE))


H3K27me3_counts_long_50per %>% 
  dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=count, fill=status)) +
  geom_bar(stat="identity", position = "fill") +
  facet_wrap(~TE_type, scales="free_y")+
  ggtitle("H3K27me3 50% TE overlap enrichment")+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
  

```


### H3K36me3 proportions  
##### results for 1bp overlap H3K36me3  
```{r H3K36me3 df}

H3K36me3_results_1bp <- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_2, function(pair){
    test_pair_TE_generic(H3K36me3_counts_long, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})

htmltools::tagList(
  htmltools::h3("H3K36me3 odds ratio 1bp result"), 
datatable(H3K36me3_results_1bp,options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE))



H3K36me3_counts_long %>% 
  dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=count, fill=status)) +
  geom_bar(stat="identity", position = "fill") +
  facet_wrap(~TE_type, scales="free_y")+
  ggtitle("H3K36me3 1 bp overlap enrichment")+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

##### results for 50% TE overlap H3K36me3 
```{r H3K36me3 50 per df}
###
H3K36me3_results_50per <- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_2, function(pair){
    test_pair_TE_generic(H3K36me3_counts_long_50per, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})

htmltools::tagList(
  htmltools::h3("H3K36me3 odds ratio >50% overlap results"),
datatable(H3K36me3_results_50per,options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE))


H3K36me3_counts_long_50per %>% 
  dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=count, fill=status)) +
  geom_bar(stat="identity", position = "fill") +
  facet_wrap(~TE_type, scales="free_y")+
  ggtitle("H3K36me3 50% TE overlap enrichment")+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
  

```


### H3K9me3 proportions  
##### results for 1bp overlap H3K9me3  
```{r H3K9me3 df}

H3K9me3_results_1bp <- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_3, function(pair){
    test_pair_TE_generic(H3K9me3_counts_long, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})

htmltools::tagList(
  htmltools::h3("H3K9me3 odds ratio 1bp result"), 
datatable(H3K9me3_results_1bp,options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE))



H3K9me3_counts_long %>% 
  dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=count, fill=status)) +
  geom_bar(stat="identity", position = "fill") +
  facet_wrap(~TE_type, scales="free_y")+
  ggtitle("H3K9me3 1 bp overlap enrichment")+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

##### results for 50% TE overlap H3K9me3 
```{r H3K9me3 50 per df}
###
H3K9me3_results_50per <- map_dfr(TE_types, function(te){
  map_dfr(cluster_pairs_3, function(pair){
    test_pair_TE_generic(H3K9me3_counts_long_50per, te, cluster1 = pair[2], cluster2 = pair[1])
  })
})

htmltools::tagList(
  htmltools::h3("H3K9me3 odds ratio >50% overlap results"),
datatable(H3K9me3_results_50per,options = list(
    scrollY = "600px",
    scrollCollapse = TRUE,
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE))


H3K9me3_counts_long_50per %>% 
  dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=count, fill=status)) +
  geom_bar(stat="identity", position = "fill") +
  facet_wrap(~TE_type, scales="free_y")+
  ggtitle("H3K9me3 50% TE overlap enrichment")+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
  

```

### H3K27ac divergence from  consensus
```{r H3K27ac divergence, fig.width=8}


  # H3K27ac_ols$all_H3K27ac %>% 
  #   left_join(., H3K27ac_lookup) %>% 
  #    mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
  #   dplyr::filter(cluster != "not_assigned") %>% 
  #   mutate(special_type=case_when(repClass=="SINE"~"SINE",
  #                                 repClass=="LINE"~"LINE",
  #                                 repClass=="LTR"~"LTR",
  #                                 repClass=="DNA"~"DNA",
  #                                 repClass=="Retroposon"~"SVA",
  #                                 is.na(repClass)~"not_TE",
  #                                 TRUE~"Other")) %>% 
  #   mutate(ppr_div=milliDiv) %>% 
  # ggplot(aes(x=ppr_div))+
  # geom_histogram(aes(fill=cluster, alpha = 0.4) ) +
  # facet_wrap(~special_type, scales="free_y") +
  #   theme_classic()
  # 
  # 
  
  H3K27ac_ols$all_H3K27ac %>% 
  as.data.frame() %>% 
    left_join(., H3K27ac_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  ggplot(aes(y=ppr_div,x=special_type))+
  geom_boxplot(aes(fill=cluster) ) +
  # facet_wrap(~special_type, scales="free_y") +
    theme_classic()+
    ggtitle("Divergence from overlapping TE H3K27ac 1bp")
  
  
  H3K27ac_ols$all_H3K27ac %>% 
  as.data.frame() %>% 
    dplyr::filter(pct_te_overlap> 50) %>% 
    left_join(., H3K27ac_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>%
    ggplot(aes(y=ppr_div,x=special_type, fill=cluster))+
  # geom_boxplot(aes(fill=cluster) ) +
    geom_boxplot()+
stat_compare_means(
    aes(group = cluster),               # tell ggpubr we want to compare clusters
    method = "wilcox.test",             # same test as before
    comparisons = list(c("Set1", "Set2"), c("Set1", "Set3")),
    label = "p.signif",                 # show stars instead of p-values
    hide.ns = TRUE                      # hide non-significant comparisons
  ) +
    theme_classic()+
    # facet_wrap(~ special_type, scales = "free_y") +
    ggtitle("Divergence from overlapping TE H3K27ac 50% overlap")
  
  
### try to calculate wilcox.test
  comparisons <- list(
  c("Set_1", "Set_2"),
  c("Set_1", "Set_3")
)

  # 
  H3K27ac_wilcox_results <- H3K27ac_ols$all_H3K27ac %>%
    as.data.frame() %>%
    left_join(., H3K27ac_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  filter(cluster %in% c("Set_1", "Set_2", "Set_3")) %>%
  group_by(special_type) %>%
  group_map(~ {
    stype <- unique(.x$special_type)  # capture the TE type
    map_dfr(comparisons, function(comp) {
      dat <- .x %>% filter(cluster %in% comp)
      test <- wilcox.test(ppr_div ~ cluster, data = dat)
      tibble(
        special_type = stype,
        comparison = paste(comp, collapse = "_vs_"),
        p.value = test$p.value,
        W = test$statistic,
        median_Set1 = median(dat$ppr_div[dat$cluster == "Set_1"]),
        median_other = median(dat$ppr_div[dat$cluster == comp[2]])
      )
    })
  }, .keep = TRUE) %>%
  bind_rows()
  
  H3K27ac_wilcox_results_per50 <- H3K27ac_ols$all_H3K27ac %>%
    as.data.frame() %>%
    dplyr::filter(pct_te_overlap> 50) %>%
    left_join(., H3K27ac_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  filter(cluster %in% c("Set_1", "Set_2", "Set_3")) %>%
  group_by(special_type) %>%
  group_map(~ {
    stype <- unique(.x$special_type)  # capture the TE type
    map_dfr(comparisons, function(comp) {
      dat <- .x %>% filter(cluster %in% comp)
      test <- wilcox.test(ppr_div ~ cluster, data = dat)
      tibble(
        special_type = stype,
        comparison = paste(comp, collapse = "_vs_"),
        p.value = test$p.value,
        W = test$statistic,
        median_Set1 = median(dat$ppr_div[dat$cluster == "Set_1"]),
        median_other = median(dat$ppr_div[dat$cluster == comp[2]])
      )
    })
  }, .keep = TRUE) %>%
  bind_rows()

  
  datatable( H3K27ac_wilcox_results,caption = htmltools::tags$caption(
  style = 'caption-side: top; text-align: center; font-weight: bold;',
  "Wilcox test on percent divergence overlapped TEs by 1bp H3K27ac"
))

 datatable( H3K27ac_wilcox_results_per50,caption = htmltools::tags$caption(
  style = 'caption-side: top; text-align: center; font-weight: bold;',
  "Wilcox test on percent divergence of >50% overlapped TEs H3K27ac"
))
```

### H3K27me3 divergence from consensus
```{r H3K27me3 divergence, fig.width=8}


  # H3K27me3_ols$all_H3K27me3 %>% 
  #   left_join(., H3K27me3_lookup) %>% 
  #    mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
  #   dplyr::filter(cluster != "not_assigned") %>% 
  #   mutate(special_type=case_when(repClass=="SINE"~"SINE",
  #                                 repClass=="LINE"~"LINE",
  #                                 repClass=="LTR"~"LTR",
  #                                 repClass=="DNA"~"DNA",
  #                                 repClass=="Retroposon"~"SVA",
  #                                 is.na(repClass)~"not_TE",
  #                                 TRUE~"Other")) %>% 
  #   mutate(ppr_div=milliDiv) %>% 
  # ggplot(aes(x=ppr_div))+
  # geom_histogram(aes(fill=cluster, alpha = 0.4) ) +
  # facet_wrap(~special_type, scales="free_y") +
  #   theme_classic()
  
  
  
  H3K27me3_ols$all_H3K27me3 %>% 
  as.data.frame() %>% 
    left_join(., H3K27me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  ggplot(aes(y=ppr_div,x=special_type))+
  geom_boxplot(aes(fill=cluster) ) +
  # facet_wrap(~special_type, scales="free_y") +
    theme_classic()+
    ggtitle("Divergence from overlapping TE H3K27me3 1bp")
  
  
  H3K27me3_ols$all_H3K27me3 %>% 
  as.data.frame() %>% 
    dplyr::filter(pct_te_overlap> 50) %>% 
    left_join(., H3K27me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  ggplot(aes(y=ppr_div,x=special_type))+
  geom_boxplot(aes(fill=cluster) ) +
  # facet_wrap(~special_type, scales="free_y") +
    theme_classic()+
    ggtitle("Divergence from overlapping TE H3K27me3 50% overlap")
  
   H3K27me3_wilcox_results <- H3K27me3_ols$all_H3K27me3 %>%
    as.data.frame() %>%
     left_join(., H3K27me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
    dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  filter(cluster %in% c("Set_1", "Set_2")) %>%
  group_by(special_type) %>%
  group_map(~ {
    stype <- unique(.x$special_type)  # capture the TE type
    map_dfr(comparisons, function(comp) {
      dat <- .x %>% filter(cluster %in% comp)
       # Skip if both clusters are not present
      if(length(unique(dat$cluster)) < 2) return(NULL)
      test <- wilcox.test(ppr_div ~ cluster, data = dat)
      tibble(
        special_type = stype,
        comparison = paste(comp, collapse = "_vs_"),
        p.value = test$p.value,
        W = test$statistic,
        median_Set1 = median(dat$ppr_div[dat$cluster == "Set_1"]),
        median_other = median(dat$ppr_div[dat$cluster == comp[2]])
      )
    })
  }, .keep = TRUE) %>%
  bind_rows()
   
   
 H3K27me3_wilcox_results_per50 <- H3K27me3_ols$all_H3K27me3 %>%
    as.data.frame() %>%
    dplyr::filter(pct_te_overlap> 50) %>%
    left_join(., H3K27me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
    dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  filter(cluster %in% c("Set_1", "Set_2")) %>%
  group_by(special_type) %>%
  group_map(~ {
    stype <- unique(.x$special_type)  # capture the TE type
    map_dfr(comparisons, function(comp) {
      dat <- .x %>% filter(cluster %in% comp)
       # Skip if both clusters are not present
      if(length(unique(dat$cluster)) < 2) return(NULL)
      test <- wilcox.test(ppr_div ~ cluster, data = dat)
      tibble(
        special_type = stype,
        comparison = paste(comp, collapse = "_vs_"),
        p.value = test$p.value,
        W = test$statistic,
        median_Set1 = median(dat$ppr_div[dat$cluster == "Set_1"]),
        median_other = median(dat$ppr_div[dat$cluster == comp[2]])
      )
    })
  }, .keep = TRUE) %>%
  bind_rows()


   datatable( H3K27me3_wilcox_results,caption = htmltools::tags$caption(
  style = 'caption-side: top; text-align: center; font-weight: bold;',
  "Wilcox test on percent divergence overlapped TEs by 1bp H3K27me3"
))

 datatable( H3K27me3_wilcox_results_per50,caption = htmltools::tags$caption(
  style = 'caption-side: top; text-align: center; font-weight: bold;',
  "Wilcox test on percent divergence of >50% overlapped TEs H3K27me3"
))
  
    
```


### H3K36me3 divergence from  consensus
```{r H3K36me3 divergence, fig.width=8}

# 
#   H3K36me3_ols$all_H3K36me3 %>% 
#     left_join(., H3K36me3_lookup) %>% 
#      mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
#     dplyr::filter(cluster != "not_assigned") %>% 
#     mutate(special_type=case_when(repClass=="SINE"~"SINE",
#                                   repClass=="LINE"~"LINE",
#                                   repClass=="LTR"~"LTR",
#                                   repClass=="DNA"~"DNA",
#                                   repClass=="Retroposon"~"SVA",
#                                   is.na(repClass)~"not_TE",
#                                   TRUE~"Other")) %>% 
#     mutate(ppr_div=milliDiv) %>% 
#   ggplot(aes(x=ppr_div))+
#   geom_histogram(aes(fill=cluster, alpha = 0.4) ) +
#   facet_wrap(~special_type, scales="free_y") +
#     theme_classic()
#   
  
  
  H3K36me3_ols$all_H3K36me3 %>% 
  as.data.frame() %>% 
    left_join(., H3K36me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  ggplot(aes(y=ppr_div,x=special_type))+
  geom_boxplot(aes(fill=cluster) ) +
  # facet_wrap(~special_type, scales="free_y") +
    theme_classic()+
    ggtitle("Divergence from overlapping TE H3K36me3 1bp")
  
  
  H3K36me3_ols$all_H3K36me3 %>% 
  as.data.frame() %>% 
    dplyr::filter(pct_te_overlap> 50) %>% 
    left_join(., H3K36me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  ggplot(aes(y=ppr_div,x=special_type))+
  geom_boxplot(aes(fill=cluster) ) +
  # facet_wrap(~special_type, scales="free_y") +
    theme_classic()+
    ggtitle("Divergence from overlapping TE H3K36me3 50% overlap")
  
  
  H3K36me3_ols$all_H3K36me3 %>% 
  as.data.frame() %>% 
    dplyr::filter(pct_te_overlap> 50) %>% 
    left_join(., H3K36me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  ggplot(aes(y=ppr_div,x=special_type))+
  geom_boxplot(aes(fill=cluster) ) +
  # facet_wrap(~special_type, scales="free_y") +
    theme_classic()+
    ggtitle("Divergence from overlapping TE H3K36me3 50% overlap")
  
   H3K36me3_wilcox_results <- H3K36me3_ols$all_H3K36me3 %>%
    as.data.frame() %>%
     left_join(., H3K36me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
    dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  filter(cluster %in% c("Set_1", "Set_2")) %>%
  group_by(special_type) %>%
  group_map(~ {
    stype <- unique(.x$special_type)  # capture the TE type
    map_dfr(comparisons, function(comp) {
      dat <- .x %>% filter(cluster %in% comp)
       # Skip if both clusters are not present
      if(length(unique(dat$cluster)) < 2) return(NULL)
      test <- wilcox.test(ppr_div ~ cluster, data = dat)
      tibble(
        special_type = stype,
        comparison = paste(comp, collapse = "_vs_"),
        p.value = test$p.value,
        W = test$statistic,
        median_Set1 = median(dat$ppr_div[dat$cluster == "Set_1"]),
        median_other = median(dat$ppr_div[dat$cluster == comp[2]])
      )
    })
  }, .keep = TRUE) %>%
  bind_rows()
   
   
   H3K36me3_wilcox_results_per50 <- H3K36me3_ols$all_H3K36me3 %>%
    as.data.frame() %>%
    dplyr::filter(pct_te_overlap> 50) %>%
    left_join(., H3K36me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
    dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  filter(cluster %in% c("Set_1", "Set_2")) %>%
  group_by(special_type) %>%
  group_map(~ {
    stype <- unique(.x$special_type)  # capture the TE type
    map_dfr(comparisons, function(comp) {
      dat <- .x %>% filter(cluster %in% comp)
       # Skip if both clusters are not present
      if(length(unique(dat$cluster)) < 2) return(NULL)
      test <- wilcox.test(ppr_div ~ cluster, data = dat)
      tibble(
        special_type = stype,
        comparison = paste(comp, collapse = "_vs_"),
        p.value = test$p.value,
        W = test$statistic,
        median_Set1 = median(dat$ppr_div[dat$cluster == "Set_1"]),
        median_other = median(dat$ppr_div[dat$cluster == comp[2]])
      )
    })
  }, .keep = TRUE) %>%
  bind_rows()
  
   
  datatable( H3K36me3_wilcox_results,caption = htmltools::tags$caption(
  style = 'caption-side: top; text-align: center; font-weight: bold;',
  "Wilcox test on percent divergence overlapped TEs by 1bp H3K36me3"
))

 datatable( H3K36me3_wilcox_results_per50,caption = htmltools::tags$caption(
  style = 'caption-side: top; text-align: center; font-weight: bold;',
  "Wilcox test on percent divergence of >50% overlapped TEs H3K36me3"
))
  
    
```


### H3K9me3 divergence from  consensus
```{r H3K9me3 divergence, fig.width=8}

# 
#   H3K9me3_ols$all_H3K9me3 %>% 
#     left_join(., H3K9me3_lookup) %>% 
#      mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
#     mutate(special_type=case_when(repClass=="SINE"~"SINE",
#                                   repClass=="LINE"~"LINE",
#                                   repClass=="LTR"~"LTR",
#                                   repClass=="DNA"~"DNA",
#                                   repClass=="Retroposon"~"SVA",
#                                   is.na(repClass)~"not_TE",
#                                   TRUE~"Other")) %>% 
#     mutate(ppr_div=milliDiv) %>% 
#   ggplot(aes(x=ppr_div))+
#   geom_histogram(aes(fill=cluster, alpha = 0.4) ) +
#   facet_wrap(~special_type, scales="free_y") +
#     theme_classic()
  
  
  
  H3K9me3_ols$all_H3K9me3 %>% 
  as.data.frame() %>% 
    left_join(., H3K9me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  ggplot(aes(y=ppr_div,x=special_type))+
  geom_boxplot(aes(fill=cluster) ) +
  # facet_wrap(~special_type, scales="free_y") +
    theme_classic()+
    ggtitle("Divergence from overlapping TE H3K9me3 1bp")
  
  
  H3K9me3_ols$all_H3K9me3 %>% 
  as.data.frame() %>% 
    dplyr::filter(pct_te_overlap> 50) %>% 
    left_join(., H3K9me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  ggplot(aes(y=ppr_div,x=special_type))+
  geom_boxplot(aes(fill=cluster) ) +
  # facet_wrap(~special_type, scales="free_y") +
    theme_classic()+
    ggtitle("Divergence from overlapping TE H3K9me3 50% overlap")
  
  
  H3K9me3_wilcox_results <- H3K9me3_ols$all_H3K9me3 %>%
    as.data.frame() %>%
    left_join(., H3K9me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
      dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  filter(cluster %in% c("Set_1", "Set_2", "Set_3")) %>%
  group_by(special_type) %>%
  group_map(~ {
    stype <- unique(.x$special_type)  # capture the TE type
    map_dfr(comparisons, function(comp) {
      dat <- .x %>% filter(cluster %in% comp)
       # Skip if both clusters are not present
      if(length(unique(dat$cluster)) < 2) return(NULL)
      test <- wilcox.test(ppr_div ~ cluster, data = dat)
      tibble(
        special_type = stype,
        comparison = paste(comp, collapse = "_vs_"),
        p.value = test$p.value,
        W = test$statistic,
        median_Set1 = median(dat$ppr_div[dat$cluster == "Set_1"]),
        median_other = median(dat$ppr_div[dat$cluster == comp[2]])
      )
    })
  }, .keep = TRUE) %>%
  bind_rows()
  
  H3K9me3_wilcox_results_per50 <- H3K9me3_ols$all_H3K9me3 %>%
    as.data.frame() %>%
    dplyr::filter(pct_te_overlap> 50) %>%
    left_join(., H3K9me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>%
    dplyr::filter(cluster != "not_assigned") %>% 
    mutate(special_type=case_when(str_detect(repClass,"SINE")~"SINE",
                                  str_detect(repClass, "LINE") ~ "LINE",
                                  str_detect(repClass, "LTR") ~ "LTR",
                                  str_detect(repClass, "DNA") ~ "DNA",
                                  str_detect(repClass, "Retroposon") ~ "SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other")) %>% 
  dplyr::filter(special_type != "not_TE") %>% 
    mutate(ppr_div=milliDiv) %>% 
  filter(cluster %in% c("Set_1", "Set_2","Set_3")) %>%
  group_by(special_type) %>%
  group_map(~ {
    stype <- unique(.x$special_type)  # capture the TE type
    map_dfr(comparisons, function(comp) {
      dat <- .x %>% filter(cluster %in% comp)
       # Skip if both clusters are not present
      if(length(unique(dat$cluster)) < 2) return(NULL)
      test <- wilcox.test(ppr_div ~ cluster, data = dat)
      tibble(
        special_type = stype,
        comparison = paste(comp, collapse = "_vs_"),
        p.value = test$p.value,
        W = test$statistic,
        median_Set1 = median(dat$ppr_div[dat$cluster == "Set_1"]),
        median_other = median(dat$ppr_div[dat$cluster == comp[2]])
      )
    })
  }, .keep = TRUE) %>%
  bind_rows()

 
  datatable( H3K9me3_wilcox_results,caption = htmltools::tags$caption(
  style = 'caption-side: top; text-align: center; font-weight: bold;',
  "Wilcox test on percent divergence overlapped TEs by 1bp H3K9me3"
))

 datatable( H3K9me3_wilcox_results_per50,caption = htmltools::tags$caption(
  style = 'caption-side: top; text-align: center; font-weight: bold;',
  "Wilcox test on percent divergence of >50% overlapped TEs H3K9me3"
))
    
```


### H3K27ac lengths of TE, TSS distance


```{r, H3K27ac length of overlaps}
ungroup_H3K27ac_repclass <- H3K27ac_ols$all_H3K27ac %>%
   left_join(., H3K27ac_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
    mutate(special_type=case_when(str_detect("SINE", repClass)~"SINE",
                                  str_detect("LINE", repClass)~"LINE",
                                  str_detect("LTR", repClass)~"LTR",
                                  str_detect("DNA", repClass)~"DNA",
                                  str_detect("Retroposon", repClass)~"SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other"))%>% 
  dplyr::filter(cluster != "not_assigned") 

# H3K27ac_repclass_sum <- ungroup_H3K27ac_repclass %>% 
#   group_by(Peakid, special_type) %>% 
#   summarize(width= min(width),
#             min_te_len=min(te_len),
#             min_roi_len=min(roi_len),
#             min_overlap=min(overlap_bp),
#             min_dist_TSS=min(distanceToTSS),
#             max_dist_TSS=max(distanceToTSS),
#             cluster=paste0(unique(cluster), collapse = ":")) 
#   

# H3K27ac_repclass_sum %>% 
#   ggplot(aes(x=special_type,y=min_te_len))+
#   geom_boxplot() +
#   facet_wrap(~cluster)+
#   theme_classic()+
#     ggtitle("Overlapping TE minimum length per class in H3K27ac")+
#    coord_cartesian(ylim=c(0,2500))

ungroup_H3K27ac_repclass %>% 
  ggplot(aes(x=special_type,y=te_len))+
  geom_boxplot() +
  facet_wrap(~cluster)+
  theme_classic()+
    ggtitle("All overlapping TE lengths in H3K27ac")+
   coord_cartesian(ylim=c(0,2500))
```


```{r, H3K27ac length and distance}

All_H3K27ac_ROI_df <- H3K27ac_sets_gr$all_H3K27ac %>% 
  as.data.frame() %>% 
   left_join(., H3K27ac_lookup) %>% 
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) 


All_H3K27ac_ROI_df %>% 
  mutate(cluster="all_regions") %>% 
  rbind(.,All_H3K27ac_ROI_df) %>% 
  # dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=width))+
  geom_boxplot()+
  theme_classic()+
    ggtitle(" All assigned ROI lengths in H3K27ac") +
  coord_cartesian(ylim=c(0,3000))


ungroup_H3K27ac_repclass %>% 
  ggplot(aes(x=special_type,y=width))+
  geom_boxplot() +
  facet_wrap(~cluster)+
  theme_classic()+
    ggtitle("Width of TE_ROI overlaps in H3K27ac")+
  coord_cartesian(ylim=c(0,2000))

ungroup_H3K27ac_repclass %>% 
   group_by(Peakid, special_type) %>% 
  summarize(distanceToTSS=min(distanceToTSS),
            cluster=paste0(unique(cluster),collapse = ",")) %>% 
  ggplot(aes(x=cluster,y=distanceToTSS))+
  geom_boxplot() +
  facet_wrap(~special_type)+
  theme_classic()+
    ggtitle("Distance to TSS by class in H3K27ac of overlapping ROIs")+
  coord_cartesian(ylim=c(-35000,35000))


ungroup_H3K27ac_repclass %>% 
  ggplot(aes(x=special_type,y=pct_te_overlap))+
  geom_boxplot(aes(fill=cluster))+
  theme_classic()+
    ggtitle("Percent of TE overlapped in H3K27ac data") 






```

### H3K27me3 lengths of TE, TSS distance



```{r, H3K27me3 length of overlaps}
ungroup_H3K27me3_repclass <- H3K27me3_ols$all_H3K27me3 %>%
   left_join(., H3K27me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
   mutate(special_type=case_when(str_detect("SINE", repClass)~"SINE",
                                  str_detect("LINE", repClass)~"LINE",
                                  str_detect("LTR", repClass)~"LTR",
                                  str_detect("DNA", repClass)~"DNA",
                                  str_detect("Retroposon", repClass)~"SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other"))%>% 
  dplyr::filter(cluster != "not_assigned") 

# H3K27me3_repclass_sum <- ungroup_H3K27me3_repclass %>% 
#   group_by(Peakid, special_type) %>% 
#   summarize(width= min(width),
#             min_te_len=min(te_len),
#             min_roi_len=min(roi_len),
#             min_overlap=min(overlap_bp),
#             min_dist_TSS=min(distanceToTSS),
#             max_dist_TSS=max(distanceToTSS),
#             cluster=paste0(unique(cluster), collapse = ":")) 
#   

# H3K27me3_repclass_sum %>% 
#   ggplot(aes(x=special_type,y=min_te_len))+
#   geom_boxplot() +
#   facet_wrap(~cluster)+
#   theme_classic()+
#     ggtitle("Overlapping TE minimum length per class in H3K27me3")+
#    coord_cartesian(ylim=c(0,2500))

ungroup_H3K27me3_repclass %>% 
  ggplot(aes(x=special_type,y=te_len))+
  geom_boxplot() +
  facet_wrap(~cluster)+
  theme_classic()+
    ggtitle("All overlapping TE lengths in H3K27me3")+
   coord_cartesian(ylim=c(0,2500))
```


```{r, H3K27me3 length and distance}

All_H3K27me3_ROI_df <- H3K27me3_sets_gr$all_H3K27me3 %>% 
  as.data.frame() %>% 
   left_join(., H3K27me3_lookup) %>% 
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) 


All_H3K27me3_ROI_df %>% 
  mutate(cluster="all_regions") %>% 
  rbind(.,All_H3K27me3_ROI_df) %>% 
  # dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=width))+
  geom_boxplot()+
  theme_classic()+
    ggtitle(" All assigned ROI lengths in H3K27me3") +
  coord_cartesian(ylim=c(0,3000))


ungroup_H3K27me3_repclass %>% 
  ggplot(aes(x=special_type,y=width))+
  geom_boxplot() +
  facet_wrap(~cluster)+
  theme_classic()+
    ggtitle("Width of TE_ROI overlaps in H3K27me3")+
  coord_cartesian(ylim=c(0,2000))

ungroup_H3K27me3_repclass %>% 
   group_by(Peakid, special_type) %>% 
  summarize(distanceToTSS=min(distanceToTSS),
            cluster=paste0(unique(cluster),collapse = ",")) %>% 
  ggplot(aes(x=cluster,y=distanceToTSS))+
  geom_boxplot() +
  facet_wrap(~special_type)+
  theme_classic()+
    ggtitle("Distance to TSS by class in H3K27me3 of overlapping ROIs")+
  coord_cartesian(ylim=c(-35000,35000))

ungroup_H3K27me3_repclass %>% 
  ggplot(aes(x=special_type,y=pct_te_overlap))+
  geom_boxplot(aes(fill=cluster))+
  theme_classic()+
    ggtitle("Percent of TE overlapped in H3K27me3 data") 


ungroup_H3K27me3_repclass %>% 
  dplyr::filter(special_type=="Other") %>% 
  ggplot(aes(repFamily, y = pct_te_overlap))+
  geom_boxplot(aes(fill=cluster))+
  theme_classic()+
    ggtitle("Percent of TE (just the Others) overlapped in H3K27me3 data") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1) )





```



### H3K36me3 lengths of TE, TSS distance



```{r, H3K36me3 length of overlaps}
ungroup_H3K36me3_repclass <- H3K36me3_ols$all_H3K36me3 %>%
   left_join(., H3K36me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
    mutate(special_type=case_when(str_detect("SINE", repClass)~"SINE",
                                  str_detect("LINE", repClass)~"LINE",
                                  str_detect("LTR", repClass)~"LTR",
                                  str_detect("DNA", repClass)~"DNA",
                                  str_detect("Retroposon", repClass)~"SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other"))%>% 
  dplyr::filter(cluster != "not_assigned") 

# H3K36me3_repclass_sum <- ungroup_H3K36me3_repclass %>% 
#   group_by(Peakid, special_type) %>% 
#   summarize(width= min(width),
#             min_te_len=min(te_len),
#             min_roi_len=min(roi_len),
#             min_overlap=min(overlap_bp),
#             min_dist_TSS=min(distanceToTSS),
#             max_dist_TSS=max(distanceToTSS),
#             cluster=paste0(unique(cluster), collapse = ":")) 
#   

# H3K36me3_repclass_sum %>% 
#   ggplot(aes(x=special_type,y=min_te_len))+
#   geom_boxplot() +
#   facet_wrap(~cluster)+
#   theme_classic()+
#     ggtitle("Overlapping TE minimum length per class in H3K36me3")+
#    coord_cartesian(ylim=c(0,2500))

ungroup_H3K36me3_repclass %>% 
  ggplot(aes(x=special_type,y=te_len))+
  geom_boxplot() +
  facet_wrap(~cluster)+
  theme_classic()+
    ggtitle("All overlapping TE lengths in H3K36me3")+
   coord_cartesian(ylim=c(0,2500))
```


```{r, H3K36me3 length and distance}

All_H3K36me3_ROI_df <- H3K36me3_sets_gr$all_H3K36me3 %>% 
  as.data.frame() %>% 
   left_join(., H3K36me3_lookup) %>% 
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) 


All_H3K36me3_ROI_df %>% 
  mutate(cluster="all_regions") %>% 
  rbind(.,All_H3K36me3_ROI_df) %>% 
  # dplyr::filter(cluster != "not_assigned") %>% 
  ggplot(aes(x=cluster, y=width))+
  geom_boxplot()+
  theme_classic()+
    ggtitle(" All assigned ROI lengths in H3K36me3") +
  coord_cartesian(ylim=c(0,3000))


ungroup_H3K36me3_repclass %>% 
  ggplot(aes(x=special_type,y=width))+
  geom_boxplot() +
  facet_wrap(~cluster)+
  theme_classic()+
    ggtitle("Width of TE_ROI overlaps in H3K36me3")+
  coord_cartesian(ylim=c(0,2000))

ungroup_H3K36me3_repclass %>% 
   group_by(Peakid, special_type) %>% 
  summarize(distanceToTSS=min(distanceToTSS),
            cluster=paste0(unique(cluster),collapse = ",")) %>% 
  ggplot(aes(x=cluster,y=distanceToTSS))+
  geom_boxplot() +
  facet_wrap(~special_type)+
  theme_classic()+
    ggtitle("Distance to TSS by class in H3K36me3 of overlapping ROIs")+
  coord_cartesian(ylim=c(-25000,25000))


ungroup_H3K36me3_repclass %>% 
  ggplot(aes(x=special_type,y=pct_te_overlap))+
  geom_boxplot(aes(fill=cluster))+
  theme_classic()+
    ggtitle("Percent of TE overlapped in H3K36me3 data") 



ungroup_H3K36me3_repclass %>% 
  dplyr::filter(special_type=="Other") %>% 
  ggplot(aes(repFamily, y = pct_te_overlap))+
  geom_boxplot(aes(fill=cluster))+
  theme_classic()+
    ggtitle("Percent of TE (just the Others) overlapped in H3K36me3 data") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1) )




```



### H3K9me3 lengths of TE, TSS distance



```{r, H3K9me3 length of overlaps}
ungroup_H3K9me3_repclass <- H3K9me3_ols$all_H3K9me3 %>%
   left_join(., H3K9me3_lookup) %>% 
     mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) %>% 
    mutate(special_type=case_when(str_detect("SINE", repClass)~"SINE",
                                  str_detect("LINE", repClass)~"LINE",
                                  str_detect("LTR", repClass)~"LTR",
                                  str_detect("DNA", repClass)~"DNA",
                                  str_detect("Retroposon", repClass)~"SVA",
                                  is.na(repClass)~"not_TE",
                                  TRUE~"Other"))%>% 
  dplyr::filter(cluster != "not_assigned") 

# H3K9me3_repclass_sum <- ungroup_H3K9me3_repclass %>% 
#   group_by(Peakid, special_type) %>% 
#   summarize(width= min(width),
#             min_te_len=min(te_len),
#             min_roi_len=min(roi_len),
#             min_overlap=min(overlap_bp),
#             min_dist_TSS=min(distanceToTSS),
#             max_dist_TSS=max(distanceToTSS),
#             cluster=paste0(unique(cluster), collapse = ":")) 
#   

# H3K9me3_repclass_sum %>% 
#   ggplot(aes(x=special_type,y=min_te_len))+
#   geom_boxplot() +
#   facet_wrap(~cluster)+
#   theme_classic()+
#     ggtitle("Overlapping TE minimum length per class in H3K9me3")+
#    coord_cartesian(ylim=c(0,2500))

ungroup_H3K9me3_repclass %>% 
  ggplot(aes(x=special_type,y=te_len))+
  geom_boxplot() +
  facet_wrap(~cluster)+
  theme_classic()+
    ggtitle("All overlapping TE lengths in H3K9me3")+
   coord_cartesian(ylim=c(0,2500))
```


```{r, H3K9me3 length and distance}

All_H3K9me3_ROI_df <- H3K9me3_sets_gr$all_H3K9me3 %>% 
  as.data.frame() %>% 
   left_join(., H3K9me3_lookup) %>% 
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster)) 


# All_H3K9me3_ROI_df %>% 
#   mutate(cluster="all_regions") %>% 
#   rbind(.,All_H3K9me3_ROI_df) %>% 
#   dplyr::filter(cluster != "not_assigned") %>%
#   ggplot(aes(x=cluster, y=width))+
#   geom_boxplot()+
#   geom_signif(comparisons = list(c("Set_1","Set_2"),
#                                   c("Set_1","Set_3"),
#                                   c("all_regions","Set_1")),
#               map_signif_level = FALSE,
#               test = "wilcox.test",
#               y_position = c(100,300,500))+# note values are scaled internally by geomsignif
#               # step_increase = 0.1)+
#   theme_classic()+
#     ggtitle(" All assigned ROI lengths in H3K9me3")+
#   coord_cartesian(ylim=c(0,4000))

ungroup_H3K9me3_repclass %>% 
   group_by(Peakid, special_type) %>% 
  summarize(distanceToTSS=min(distanceToTSS),
            cluster=paste0(unique(cluster),collapse = ",")) %>% 
  ggplot(aes(x=cluster,y=distanceToTSS))+
  geom_boxplot() +
  facet_wrap(~special_type)+
  theme_classic()+
    ggtitle("Distance to TSS by class in H3K9me3 of overlapping ROIs")+
  coord_cartesian(ylim=c(-35000,35000))


ungroup_H3K9me3_repclass %>% 
  ggplot(aes(x=special_type,y=width))+
  geom_boxplot() +
  facet_wrap(~cluster)+
  theme_classic()+
    ggtitle("Width of TE_ROI overlaps in H3K9me3")+
  coord_cartesian(ylim=c(0,2000))

ungroup_H3K9me3_repclass %>% 
  ggplot(aes(x=cluster,y=distanceToTSS))+
  geom_boxplot(aes(fill=cluster)) +
  geom_signif(comparisons = list(c("Set_1","Set_2"),
                                  c("Set_1","Set_3")),
                                  
              map_signif_level = FALSE,
              test = "wilcox.test",
              y_position = c(10,50,100))+# note values are scaled internally by geomsignif
              # step_increase = 0.1)+
  facet_wrap(~special_type)+
  theme_classic()+
    ggtitle("Distance to TSS by class in H3K9me3 of overlapping ROIs")+
  coord_cartesian(ylim=c(-45000,45000))

ungroup_H3K9me3_repclass %>% 
  ggplot(aes(x=cluster,y=distanceToTSS))+
  geom_boxplot(aes(fill=cluster)) +
  geom_signif(comparisons = list(c("Set_1","Set_2"),
                                  c("Set_1","Set_3")),
                                  
              map_signif_level = FALSE,
              test = "wilcox.test",
              # y_position = c(10,50,100))+# note values are scaled internally by geomsignif
              step_increase = 0.1)+
  facet_wrap(~special_type)+
  theme_classic()+
    ggtitle("Distance to TSS by class in H3K9me3 of overlapping ROIs")


ungroup_H3K9me3_repclass %>% 
  ggplot(aes(x=special_type,y=pct_te_overlap))+
  geom_boxplot(aes(fill=cluster))+
  theme_classic()+
    ggtitle("Percent of TE overlapped in H3K9me3 data") 


ungroup_H3K9me3_repclass %>% 
  dplyr::filter(special_type=="Other") %>% 
  ggplot(aes(repFamily, y = pct_te_overlap))+
  geom_boxplot(aes(fill=cluster))+
  theme_classic()+
    ggtitle("Percent of TE (just the Others) overlapped in H3K9me3 data") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1) )


```




