---
title: "Summit Files"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```


```{r packages}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(BiocParallel)

```

```{r calling summit files}
sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")
##Path to histone summit files
H3K27ac_dir <- "C:/Users/renee/Other_projects_data/DXR_data/final_data/summit_files/H3K27ac" 

##pull all histone files together
H3K27ac_summit_files <- list.files(
  path = H3K27ac_dir,
  pattern = "\\.bed$",
  recursive = TRUE,
  full.names = TRUE
)
length(H3K27ac_summit_files)

# head(H3K27ac_summit_files)
```
First steps:  pulling in the information of Sets and locations
Pulling in the summit files,
concatenation of all summits into one large file

```{r pulling in other data}
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")
H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})

read_summit <- function(file){
  peaks <- read.table(file,header = FALSE)
  colnames(peaks) <- c("chr","start","end","name","score")
  GRanges(
    seqnames = peaks$chr,
    ranges = IRanges(start=peaks$start, end = peaks$start),
    score=peaks$score,
    file=basename(file),
    Library_ID = stringr::str_remove(basename(file), "_FINAL_summits\\.bed$")
  )
}

all_H3K27ac_summits_list<- lapply(H3K27ac_summit_files, read_summit)
all_H3K27ac_summits_gr <- do.call(c, all_H3K27ac_summits_list)  # combine into one GRanges object
H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
```

```{r pick summit per roi}
pick_best_summit_per_roi <- function(roi_gr, summit_gr, score_col = "signalValue") {

  hits <- findOverlaps(roi_gr, summit_gr)

  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    score = mcols(summit_gr)[[score_col]][subjectHits(hits)]
  )

  # Pick summit with highest score for each ROI
  best <- df %>%
    group_by(roi_idx) %>%
    slice_max(order_by = score, n = 1) %>%
    ungroup()

  # Build the final GRanges object
  best_summits <- summit_gr[best$summit_idx]
  best_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]

  best_summits
}
# 
# best_summits <- pick_best_summit_per_roi(
#   ROIs, highest_summits_long_gr, score_col = "peakHeight"
# )
```

```{r pick summit closest to roi center}
pick_center_summit <- function(roi_gr, summit_gr) {

  hits <- findOverlaps(roi_gr, summit_gr)

  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    dist_to_center = abs(
      start(summit_gr)[subjectHits(hits)] -
      round((start(roi_gr)[queryHits(hits)] + end(roi_gr)[queryHits(hits)]) / 2)
    )
  )

  best <- df %>%
    group_by(roi_idx) %>%
    slice_min(order_by = dist_to_center, n = 1) %>%
    ungroup()

  best_summits <- summit_gr[best$summit_idx]
  best_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]

  best_summits
}
```

```{r consensus summit}
pick_consensus_summit <- function(roi_gr, summit_gr, score_col = "score") {
  
  # Find which summits overlap each ROI
  hits <- findOverlaps(roi_gr, summit_gr)
  
  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    summit_pos = start(summit_gr)[subjectHits(hits)],
    score = mcols(summit_gr)[[score_col]]
  )
  
  # Count frequency per summit position within each ROI
  freq_df <- df %>%
    group_by(roi_idx, summit_pos) %>%
    summarise(
      freq = n(),
      max_score = max(score),
      .groups = "drop"
    )
  
  # Pick the summit with highest frequency; break ties with max score
  best <- freq_df %>%
    group_by(roi_idx) %>%
    slice_max(order_by = freq, n = 1, with_ties = TRUE) %>%
    slice_max(order_by = max_score, n = 1) %>%
    ungroup()
  
  # Map back to original GRanges
  best_idx <- df$summit_idx[match(
    paste0(best$roi_idx, "_", best$summit_pos),
    paste0(df$roi_idx, "_", df$summit_pos)
  )]
  
  final_summits <- summit_gr[best_idx]
  final_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]
  
  final_summits
}
```

### Analysis of all summit files, no grouping  


first creating summit clusters that are around 100bps across all summit files
```{r pulling in other data again}
 
#  Merge summits within 100 bp clusters but keep metadata
# ------------------------
# 1 Reduce with revmap to keep track of original indices

clusters <- GenomicRanges::reduce(all_H3K27ac_summits_gr, min.gapwidth = 100,
                   ignore.strand = TRUE, with.revmap = TRUE)

# 2 For each cluster, pick the highest-score summit
scores <- mcols(all_H3K27ac_summits_gr)$score

revmap <- clusters$revmap

# Compute the highest-score index per cluster
highest_idx <- sapply(revmap, function(idx) idx[which.max(scores[idx])])

# Subset GRanges once
highest_per_cluster_gr <- all_H3K27ac_summits_gr[highest_idx]


# ------------------------
# 3 Count merged summits per ROI
# ------------------------
ROIs <- H3K27ac_sets_gr$all_H3K27ac  # your ROI GRanges

# Optimized counting
roi_counts <- countOverlaps(ROIs, highest_per_cluster_gr)
mcols(ROIs)$merged_summit_count <- roi_counts

# ------------------------
# 5 Map ROIs to clusters with sample info
# ------------------------
overlaps <- findOverlaps(ROIs, highest_per_cluster_gr)

roi_hits_df <- as.data.frame(overlaps) %>%
  mutate(
    Peakid = ROIs$Peakid[queryHits],
    Library_ID = mcols(highest_per_cluster_gr)$Library_ID[subjectHits],
    score = mcols(highest_per_cluster_gr)$score[subjectHits],
    file = mcols(highest_per_cluster_gr)$file[subjectHits]
  ) %>%
  left_join(sampleinfo, by = c("Library_ID" = "Library ID"))

## initualize count column on ROI df
mcols(ROIs)$summit_count <- 0

# Tabulate counts
counts <- table(queryHits(overlaps))
mcols(ROIs)$summit_count[as.numeric(names(counts))] <- as.numeric(counts)

# Convert to dataframe for plotting
roi_counts_df <- as.data.frame(ROIs) %>%
  select(Peakid, seqnames, start, end, summit_count) %>% 
  mutate(roi_size=end-start)

# Optional: add Set / cluster info
roi_counts_df <- roi_counts_df %>%
  left_join(H3K27ac_lookup, by = "Peakid")
# ------------------------
# 6 Optional: add Set/cluster info per ROI
# ------------------------
roi_hits_df <- roi_hits_df %>%
  left_join(H3K27ac_lookup, by = "Peakid")

# Now roi_hits_df is ready for plotting or analysis:
# Columns include Peakid, ROI coordinates, Library_ID, score, file, Individual, Treatment, Timepoint, cluster


# Optional: add sample info
roi_hits_df <- as.data.frame(overlaps) %>%
  mutate(
    Peakid = ROIs$Peakid[queryHits],
    Library_ID = mcols(all_H3K27ac_summits_gr)$Library_ID[subjectHits],
    score = mcols(all_H3K27ac_summits_gr)$score[subjectHits],
    file = mcols(all_H3K27ac_summits_gr)$file[subjectHits]
  ) %>%
  left_join(sampleinfo, by = c("Library_ID" = "Library ID"))

```

Plotting some summits from the first process

```{r get summits function plots}
ROIs %>% 
  as.data.frame() %>% 
  # dplyr::filter(merged_summit_count==1) %>% 
  left_join(H3K27ac_lookup, by = "Peakid") %>%  # make sure join key matches
  filter(!is.na(cluster)) %>%
  ggplot(aes(x = cluster, y = merged_summit_count)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6, size = 2, color = "steelblue") +
  theme_bw() +
  xlab("Set") +
  ylab("Number of merged summits per ROI") +
  ggtitle("Distribution of merged summit counts by Set")


roi_counts_df %>%
  ggplot(aes(x = summit_count)) +
  geom_histogram(binwidth = 1, fill="steelblue", color="black") +
  theme_bw() +
  xlab("Number of summits per ROI") +
  ylab("Number of ROIs")

roi_counts_df %>%
  ggplot(aes(x = summit_count)) +
  geom_histogram(binwidth = 1, fill="steelblue", color="black") +
  theme_bw() +
  xlab("Number of summits per ROI") +
  ylab("Number of ROIs")+
  ggtitle("Zoomed in summit per ROI")+
  coord_cartesian(xlim=c(0,50))

roi_counts_df %>% 
  # group_by(Peakid) %>% tally #%>% 
  # left_join(H3K27ac_lookup) %>% 
  ggplot(aes(x = roi_size, y = summit_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "ROI size (bp)",
    y = "# of summits",
    title = "Relationship between ROI size and # of Summit clusters"
  ) +
  theme_classic()#+
  # coord_cartesian(xlim=c(-1,400), ylim=c(0,10))

roi_counts_df %>% 
  # group_by(Peakid) %>% tally #%>% 
  # left_join(H3K27ac_lookup) %>% 
  ggplot(aes(x = roi_size, y = summit_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "ROI size (bp)",
    y = "# of summits",
    title = "Zoomed in Relationship between ROI size and # of Summit clusters"
  ) +
  theme_classic()+
  coord_cartesian(xlim=c(-1,800), ylim=c(0,10))

```


### Merging strategies!

This is now the place I will take my files and now try to apply the same merging strategy as I did the peaks merging strategy.  Step 1: reduce all trt-time summits within 100 bp
```{r reduction by trt-time}
###Adding in sampleinfo dataframe
meta <- as.data.frame(mcols(all_H3K27ac_summits_gr))
meta2 <- meta %>% 
  left_join(., sampleinfo, by=c("Library_ID"="Library ID"))

mcols(all_H3K27ac_summits_gr) <- meta2

mcols(all_H3K27ac_summits_gr)$group <- 
  paste(all_H3K27ac_summits_gr$Treatment,
        all_H3K27ac_summits_gr$Timepoint,
        sep = "_")

###now splitting into grouped granges 

gr_by_group <- split(all_H3K27ac_summits_gr,
                                    all_H3K27ac_summits_gr$group)
# gr_by_group <- as(gr_by_group, "CompressedGRangesList")
### now reducing within some width by 100 bp with revmap
groups <- unique(all_H3K27ac_summits_gr$group)

reduced_groups <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 100, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_200 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 200, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_300 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 300, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_400 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 400, ignore.strand = TRUE, with.revmap = TRUE)
})

names(reduced_groups) <- groups
names(reduced_groups_200) <- groups
names(reduced_groups_300) <- groups
names(reduced_groups_400) <- groups

# redux_main_100 <- sum(sapply(reduced_groups, length))
# redux_main_200 <- sum(sapply(reduced_groups_200, length))
# redux_main_300 <- sum(sapply(reduced_groups_300, length))
# redux_main_400 <- sum(sapply(reduced_groups_400, length))

```

Plotting effect of reduction bp number on total number of clusters
```{r}
gap_sizes <- seq(0, 500, by = 10)


# Function to apply reduce for each gap and return counts
results <- lapply(gap_sizes, function(g) {
  reduced <- GenomicRanges::reduce(all_H3K27ac_summits_gr, min.gapwidth = g)
  data.frame(gap = g, n_regions = length(reduced))
})

# Combine results
min_gap_summary <- bind_rows(results)

ggplot(min_gap_summary, aes(x = gap, y= n_regions))+
  geom_line(linewidth=2)+
  geom_point()+
  theme_bw(base_size=14)+
  labs(
    title = "Effect of Gap Size on Reduced Summits",
    x = "Gap size (bp)",
    y = "Number of merged regions"
  )

```
```{r get highes per group funciton}
# Function to pick highest summit per cluster from a reduced GRanges list
get_highest_per_group <- function(reduced_groups, orig_summits_gr, group_col = "group") {
  
  groups <- names(reduced_groups)
  highest_per_group <- vector("list", length(reduced_groups))
  names(highest_per_group) <- groups
  
  for(i in seq_along(reduced_groups)) {
    gr <- reduced_groups[[i]]
    
    # original summits for this group
    orig <- orig_summits_gr[mcols(orig_summits_gr)[[group_col]] == groups[i]]
    scores <- orig$score
    
    # revmap is a CompressedIntegerList
    revmap <- mcols(gr)$revmap
    
    # skip if revmap is NULL
    if(is.null(revmap)) next
    
    # unlist all indices once
    all_idx <- unlist(revmap, use.names = FALSE)
    
    # repeat cluster index for each element in revmap
    cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
    
    # scores for all indices
    all_scores <- scores[all_idx]
    
    # For each cluster, pick the index of the max score
    max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
      ii[which.max(all_scores[ii])]
    })
    
    # Convert back to original indices
    orig_idx <- all_idx[unlist(max_idx_per_cluster)]
    
    # subset original GRanges
    highest_per_group[[i]] <- orig[orig_idx]
  }
  
  # Flatten any nested GRangesList
  flatten_gr <- function(x) {
    if (inherits(x, "GRanges")) return(x)
    if (inherits(x, "GRangesList")) return(unlist(x, use.names = FALSE))
    if (is.list(x)) return(do.call(c, lapply(x, flatten_gr)))
    stop("Unexpected object type")
  }
  
  highest_summits_gr <- flatten_gr(highest_per_group)
  
  # Return as long GRanges with group column
  highest_summits_df <- bind_rows(
    lapply(names(highest_per_group), function(gr_name) {
      as.data.frame(highest_per_group[[gr_name]]) %>%
        mutate(group = gr_name)
    })
  )
  
  highest_summits_long_gr <- highest_summits_df %>% GRanges()
  
  return(highest_summits_long_gr)
}
```


```{r}
reduced_sets <- list(
  "100bp" = reduced_groups,
  "200bp" = reduced_groups_200,
  "300bp" = reduced_groups_300,
  "400bp" = reduced_groups_400
)

highest_summits_all <- lapply(reduced_sets, get_highest_per_group, orig_summits_gr = all_H3K27ac_summits_gr)
```
  
  Looking at number of summits across groups as a function of min.gap number

```{r across groups mingap}
 ##Compute counts per group for each reduced set
summit_counts_group <- lapply(names(highest_summits_all), function(gap_name) {
  gr <- highest_summits_all[[gap_name]]
  
  # make sure group column exists
  if(!"group" %in% colnames(mcols(gr))) stop("GRanges must have 'group' column")
  
  df <- as.data.frame(gr) %>%
    count(group, name = "n_summits") %>%
    mutate(gap = gap_name)
  
  return(df)
}) %>% bind_rows()


ggplot(summit_counts_group, aes(x = gap, y = n_summits, group = group, color = group)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  theme_classic(base_size = 14) +
  labs(
    title = "Effect of min-gap width on number of highest summits per group",
    x = "Min-gap width (bp)",
    y = "Number of highest summits",
    color = "Group"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
### Summit average per ROI
What is the average number of summits per ROI as a function of min.gapwidth

```{r double reduce function}
double_reduce_summits <- function(summits_gr, ROIs, groups,
                                  min_gap_within_seq = c(100,200,300),
                                  min_gap_across_seq = c(100,200,300),
                                  BPPARAM = MulticoreParam(4)) {
  
  # Pre-split summits by group to avoid repeated subsetting
  summits_by_group <- split(summits_gr, summits_gr$group)
  
  # Create all gap combinations
  gap_combos <- expand.grid(min_gap_within = min_gap_within_seq,
                            min_gap_across = min_gap_across_seq,
                            stringsAsFactors = FALSE)
  
  # Apply in parallel for each combination
  results <- bplapply(seq_len(nrow(gap_combos)), function(i) {
    g1 <- gap_combos$min_gap_within[i]
    g2 <- gap_combos$min_gap_across[i]
    
    # -----------------
    # Step 1: Reduce within group
    # -----------------
    highest_per_group <- lapply(groups, function(grp) {
      gr_sub <- summits_by_group[[grp]]
      if(length(gr_sub) == 0) return(NULL)
      
      # reduce within group with revmap
      red <- GenomicRanges::reduce(gr_sub, min.gapwidth = g1, ignore.strand = TRUE, with.revmap = TRUE)
      revmap <- mcols(red)$revmap
      if(length(revmap) == 0) return(NULL)
      
      # pick highest score per cluster
      cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
      all_idx <- unlist(revmap, use.names = FALSE)
      all_scores <- gr_sub$score[all_idx]
      max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
        ii[which.max(all_scores[ii])]
      })
      gr_sub[all_idx[unlist(max_idx_per_cluster)]]
    })
    
    highest_per_group <- highest_per_group[!sapply(highest_per_group, is.null)]
    
    # -----------------
    # Step 2: Merge across groups
    # -----------------
    if(length(highest_per_group) == 0) return(NULL)
    all_highest <- do.call(c, highest_per_group)
    consensus <- GenomicRanges::reduce(all_highest, min.gapwidth = g2, ignore.strand = TRUE)
    
    # -----------------
    # Step 3: Count summits per ROI (vectorized)
    # -----------------
    hits <- findOverlaps(ROIs, consensus)
    counts <- as.data.frame(table(queryHits(hits)))
    colnames(counts) <- c("ROI_idx", "n_summits")
    counts$ROI_idx <- as.integer(as.character(counts$ROI_idx))
    counts$min_gap_within <- g1
    counts$min_gap_across <- g2
    counts
  }, BPPARAM = BPPARAM)
  
  # Combine results
  results_df <- bind_rows(results)
  
  return(results_df)
}
```

```{r calculating of double reduce}

BPPARAM <- SnowParam(workers = 4, type = "SOCK")  
register(BPPARAM)

heatmap_data <- double_reduce_summits(
  summits_gr = all_H3K27ac_summits_gr,
  ROIs = H3K27ac_sets_gr$all_H3K27ac,
  groups = unique(all_H3K27ac_summits_gr$group),
  min_gap_within_seq = c(100,200,300,400),
  min_gap_across_seq = c(100,200,300,400),
  BPPARAM = BPPARAM
)

heatmap_avg <- heatmap_data %>%
  group_by(min_gap_within, min_gap_across) %>%
  summarise(mean_summits = mean(n_summits, na.rm = TRUE), .groups = "drop")

```

```{r plot mingap averages}
ggplot(heatmap_avg, aes(x = factor(min_gap_within),
                        y = factor(min_gap_across),
                        fill = mean_summits)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(mean_summits, 1)), color = "black", size = 4) +
  scale_fill_viridis_c(option = "plasma") +
  labs(
    x = "Within-group min-gap (bp)",
    y = "Across-group min-gap (bp)",
    fill = "Mean # summits per ROI",
    title = "Effect of min-gap width on summits per ROI"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

after i figure my final min.gaps, here is where the places are
```{r}

# --- Step 1: Reduce within groups and pick highest summit per cluster ---
get_highest_per_group_final <- function(summits_gr, groups = NULL, min_gap_within = 100) {
  
  if(is.null(groups)) {
    groups <- unique(summits_gr$group)
  }
  
  highest_per_group <- lapply(groups, function(grp) {
    gr_sub <- summits_gr[summits_gr$group == grp]
    if(length(gr_sub) == 0) return(NULL)
    
    # Reduce within group
    reduced <- GenomicRanges::reduce(gr_sub, min.gapwidth = min_gap_within, ignore.strand = TRUE, with.revmap = TRUE)
    
    # Pick highest scoring summit per cluster
    revmap <- reduced$revmap
    scores <- gr_sub$score
    
    all_idx <- unlist(revmap, use.names = FALSE)
    cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
    all_scores <- scores[all_idx]
    
    max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
      ii[which.max(all_scores[ii])]
    })
    
    orig_idx <- all_idx[unlist(max_idx_per_cluster)]
    
    gr_sub[orig_idx]
  })
  
  # Flatten list of GRanges
  highest_per_group <- do.call(c, highest_per_group)
  return(highest_per_group)
}

# --- Step 2: Reduce across groups to get final consensus ---
get_consensus_summits <- function(highest_per_group_gr, min_gap_across = 400) {
  
  # Reduce across groups
  reduced <- GenomicRanges::reduce(highest_per_group_gr, min.gapwidth = min_gap_across, ignore.strand = TRUE, with.revmap = TRUE)
  
  # Pick highest scoring summit per cluster across all groups
  revmap <- reduced$revmap
  scores <- highest_per_group_gr$score
  
  all_idx <- unlist(revmap, use.names = FALSE)
  cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
  all_scores <- scores[all_idx]
  
  max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
    ii[which.max(all_scores[ii])]
  })
  
  orig_idx <- all_idx[unlist(max_idx_per_cluster)]
  
  highest_per_group_gr[orig_idx]
}

add_ROI_to_consensus <- function(consensus_gr, ROIs_gr) {
  
  hits <- findOverlaps(consensus_gr, ROIs_gr)
   # Extract metadata from consensus_gr
  consensus_meta <- mcols(consensus_gr)[queryHits(hits), , drop = FALSE]
  
  
  df <- data.frame(
    summit_chr  = seqnames(consensus_gr)[queryHits(hits)],
    summit_pos  = start(consensus_gr)[queryHits(hits)],
    summit_score = consensus_gr$score[queryHits(hits)],
    Peakid      = ROIs_gr$Peakid[subjectHits(hits)],
    roi_start   = start(ROIs_gr)[subjectHits(hits)],
    roi_end     = end(ROIs_gr)[subjectHits(hits)]
  )
  
  # Bind the metadata from consensus_gr
  df <- cbind(df, as.data.frame(consensus_meta))
  
  df %>%
    mutate(rel_pos = (summit_pos - roi_start)/(roi_end - roi_start),
           roi_center = roi_start + (roi_end - roi_start)/2,
           dist_center = summit_pos - roi_center)
}


```

```{r looking at 100-400}
highest_100 <- get_highest_per_group(all_H3K27ac_summits_gr, min_gap_within = 100)

final_consensus <- get_consensus_summits(highest_100, min_gap_across = 400)

final_df_100_400 <- add_ROI_to_consensus(final_consensus, H3K27ac_sets_gr$all_H3K27ac)


ggplot(final_df_100_400, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```


```{r distance from center plot}

final_df_100_400 %>% 
  group_by(Peakid) %>% 
  slice_max(summit_score) %>% 
ggplot(., aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  geom_vline(xintercept = 250, linetype = "dashed",color="yellow") +  # left 500 
  geom_vline(xintercept = -250, linetype = "dashed",color="yellow") +  # right 500
  
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
 coord_cartesian(xlim=c(-2000,2000))

final_df_100_400 %>% 
  group_by(Peakid) %>% 
  slice_max(summit_score) %>% 
ggplot(., aes(x = dist_center)) +
  geom_density()+
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  geom_vline(xintercept = 250, linetype = "dashed",color="blue") +  # left 500 
  geom_vline(xintercept = -250, linetype = "dashed",color="blue") +  # right 500
 coord_cartesian(xlim=c(-2000,2000))

```
```{r looking at 200-400}
highest_200 <- get_highest_per_group(all_H3K27ac_summits_gr, min_gap_within = 200)

final_consensus2 <- get_consensus_summits(highest_200, min_gap_across = 400)

final_df_200_400 <- add_ROI_to_consensus(final_consensus2, H3K27ac_sets_gr$all_H3K27ac)


ggplot(final_df_100_400, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color="green") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+coord_cartesian(xlim=c(-2000,2000))
```

#### function get highest scoring summit clustered by group:

```{r, eval=FALSE}
#' Get highest summit per reduced cluster per group using plyranges
#'
#' @param gr GRanges with metadata columns: group, score
#' @param min_gap numeric, minimum gap width to merge summits
#' @param ignore.strand logical, whether to ignore strand when reducing
#'
#' @return GRanges of highest scoring summit per reduced cluster per group
get_highest_summits_with_cluster_id <- function(gr, min_gap = 100, ignore.strand = TRUE) {
  
  gr %>%
    # Reduce summits per group
    group_by(group) %>%
    reduce_ranges(min_gap = min_gap, ignore.strand = ignore.strand, with_revmap = TRUE) %>%
    # Add a cluster_id based on row_number()
    mutate(cluster_id = paste0(group, "_cluster_", row_number())) %>%
    # join back original summits
    join_overlap_inner(gr, suffix = c(".cluster", ".orig")) %>%
    # slice max per cluster_id
    group_by(cluster_id) %>%
    slice_max(score.orig, with_ties = FALSE) %>%
    ungroup() %>%
    # select only the original summit coordinates
    select(seqnames, start = start.orig, end = end.orig, score = score.orig,
           group, cluster_id)
}
```


#### function reduce_and_pick_highest

```{r}

#### this quickly reduces any granges summit  collection by the gap that is passed in the function. It returns the highest scoring summit from the cluster of summits.
reduce_and_pick_highest <- function(gr, gap = 100) {

  # Preextract score once
  sc <- gr$score

  # Fast reduce with revmap
  red <- GenomicRanges::reduce(
    gr,
    min.gapwidth = gap,
    ignore.strand = TRUE,
    with.revmap = TRUE
  )

  # This is vectorized and very fast
  idx_list <- red$revmap

  # Pick highest-scoring summit within each cluster
  sel <- IntegerList(
    lapply(idx_list, function(idx) idx[which.max(sc[idx])])
  )

  # Convert IntegerList → integer vector
  best_idx <- unlist(sel, use.names = FALSE)

  # Subset original GRanges
  gr[best_idx]
}

```

```{r}
redux_100 <- reduce_and_pick_highest(all_H3K27ac_summits_gr, gap = 100)
redux_200 <- reduce_and_pick_highest(all_H3K27ac_summits_gr, gap = 200)
redux_300 <- reduce_and_pick_highest(all_H3K27ac_summits_gr, gap = 300)

redux_Peak_100 <- add_ROI_to_consensus(redux_100, H3K27ac_sets_gr$all_H3K27ac)
redux_Peak_200 <- add_ROI_to_consensus(redux_200, H3K27ac_sets_gr$all_H3K27ac)
redux_Peak_300 <- add_ROI_to_consensus(redux_300, H3K27ac_sets_gr$all_H3K27ac)
```


```{r plotting alternates}
ggplot(redux_Peak_100, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("Only single reduction 100bp")

ggplot(redux_Peak_200, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("Only single reduction 200bp")

ggplot(redux_Peak_300, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("Only single reduction 300bp")

```
### Creating Final 100-400 reduction summit set  

Here is how I ran the get_highest_per_group_final

```{r 100 400 highest per group}

### step 1:  reduce within group using 100 bp gap
highest_100_final <- get_highest_per_group_final(all_H3K27ac_summits_gr, min_gap_within = 100)


### step 2: reduce across groups (converge)
final_consensus_final <- get_consensus_summits(highest_100_final, min_gap_across = 400)

### step 3: add in ROI information

final_df_summits <- add_ROI_to_consensus(final_consensus_final, H3K27ac_sets_gr$all_H3K27ac)
```

```{r}
final_df_summits %>% 
  ggplot(., aes(x=rel_pos))+
  geom_histogram(bins = 100)+
  geom_vline(xintercept = 0.5)+
  ggtitle("Relative position of consensus summits by ROI")


```


#### Debugging why some ROIs do not have summits when they clearly do in IGV
```{r}


### step 1: Finding which ROIs are without summits
hits <- findOverlaps(ROIs, final_consensus_final)

roi_has_hit <- unique(queryHits(hits))

roi_index <- seq_along(ROIs)

# ROIs with no summits
no_summit_idx <- setdiff(roi_index, roi_has_hit)

missing_ROIs <- ROIs[no_summit_idx]
missing_ROIs


```

```{r}
### Step 2: Check if ROIs ever had a summit to begin with
raw_hits <- findOverlaps(ROIs, all_H3K27ac_summits_gr)
missing_raw_idx <- setdiff(roi_index, unique(queryHits(raw_hits)))

ROIs[missing_raw_idx]

### This returns nothing, which means I have a reduction problem.


```

No ROIs were summit-less, which means there is a problem in the reduction process.  Let's see if I can suss that out.

```{r}

### This code checked for all ~14,000 ROIs without summits to see if there were any nearby (It runs FOREVER)
# storage <- for(i in no_summit_idx) {
#   roi <- ROIs[i]
#   nearby = subsetByOverlaps(all_H3K27ac_summits_gr, roi, maxgap = 1000)
# 
#   cat("\nROI", ROIs$Peakid[i], "had", length(nearby), "nearby summits\n")
# }

## now we are seeing where those summits may be
first_level <- get_highest_per_group_final(all_H3K27ac_summits_gr, min_gap_within = 100)

# ROIs with summits after first reduction
hits1 <- findOverlaps(ROIs, first_level)
roi1 <- unique(queryHits(hits1))

# After second reduction
hits2 <- findOverlaps(ROIs, final_consensus_final)
roi2 <- unique(queryHits(hits2))

lost_after_consensus <- setdiff(roi1, roi2)
ROIs[lost_after_consensus]







i <- no_summit_idx[1]

roi <- ROIs[i]

inside_raw  <- subsetByOverlaps(all_H3K27ac_summits_gr, roi)
inside_lvl1 <- subsetByOverlaps(first_level, roi)
inside_lvl2 <- subsetByOverlaps(final_consensus_final, roi)

nearby_lvl1 <- subsetByOverlaps(first_level, roi, maxgap = 500)
nearby_lvl2 <- subsetByOverlaps(final_consensus_final, roi, maxgap = 500)

list(
  raw = length(inside_raw),
  reduced_within = length(inside_lvl1),
  reduced_across = length(inside_lvl2),
  nearby_after_lvl1 = length(nearby_lvl1),
  nearby_after_lvl2 = length(nearby_lvl2)
)

```

```{r}
debug_roi <- function(roi_index, ROIs, summits) {
  library(GenomicRanges)

  roi <- ROIs[roi_index]

  # Summits overlapping the ROI ± 500 bp
  nearby <- summits[
    summits@seqnames == roi@seqnames &
    summits@ranges@start >= roi@ranges@start - 500 &
    summits@ranges@end   <= roi@ranges@end + 500
  ]

  cat("\n### ROI:", roi_index, as.character(roi), "\n")
  cat("Nearby summits:", length(nearby), "\n")

  # Cluster summits within sets (level 1)
  lvl1 <- reduce(split(nearby, nearby$set), min.gapwidth = 1)
  lvl1_flat <- unlist(lvl1)
  cat("Level 1 reduced:", length(lvl1_flat), "\n")

  # Collapse across sets (level 2)
  lvl2 <- reduce(lvl1_flat, min.gapwidth = 1)
  cat("Level 2 reduced (ACROSS):", length(lvl2), "\n")

  # Show coordinates for inspection
  return(list(
    roi        = roi,
    nearby     = nearby,
    lvl1       = lvl1_flat,
    lvl2       = lvl2
  ))
}

```

```{r revised function}

# -------------------------
# Step A: reduce within groups and pick top per cluster
# -------------------------
get_highest_per_group_revised <- function(summits_gr, group_col = "group", score_col = "score", min_gap_within = 100) {
  groups <- unique(mcols(summits_gr)[[group_col]])
  
  highest_per_group <- lapply(groups, function(g) {
    gr_sub <- summits_gr[mcols(summits_gr)[[group_col]] == g]
    if (length(gr_sub) == 0) return(NULL)
    
    # reduce and get revmap (mapping reduced -> original indices)
    reduced <- GenomicRanges::reduce(gr_sub, min.gapwidth = min_gap_within, ignore.strand = TRUE, with.revmap = TRUE)
    revmap <- mcols(reduced)$revmap
    if (is.null(revmap) || length(revmap) == 0) return(NULL)
    
    # flatten revmap and pick highest-scoring original index per reduced cluster
    cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
    orig_idx <- unlist(revmap, use.names = FALSE)
    scores <- mcols(gr_sub)[[score_col]]
    all_scores <- scores[orig_idx]
    
    max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
      ii[which.max(all_scores[ii])]
    })
    chosen_orig_idx <- orig_idx[unlist(max_idx_per_cluster, use.names = FALSE)]
    
    # return the original GRanges entries (preserve metadata)
    gr_sub[chosen_orig_idx]
  })
  
  # remove NULLs and concatenate into a GRanges
  highest_per_group <- highest_per_group[!sapply(highest_per_group, is.null)]
  if (length(highest_per_group) == 0) return(GRanges())
  do.call(c, highest_per_group)
}


# -------------------------
# Step B: reduce across groups and pick top per cluster
# -------------------------
get_consensus_summits_revised <- function(highest_per_group_gr, score_col = "score", min_gap_across = 400) {
  if (length(highest_per_group_gr) == 0) return(GRanges())
  
  # Reduce across groups (ensure we pass a flat GRanges)
  consensus_red <- GenomicRanges::reduce(highest_per_group_gr, min.gapwidth = min_gap_across, ignore.strand = TRUE, with.revmap = TRUE)
  revmap <- mcols(consensus_red)$revmap
  if (is.null(revmap) || length(revmap) == 0) return(GRanges())
  
  cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
  orig_idx <- unlist(revmap, use.names = FALSE)
  scores <- mcols(highest_per_group_gr)[[score_col]]
  all_scores <- scores[orig_idx]
  
  max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
    ii[which.max(all_scores[ii])]
  })
  chosen_orig_idx <- orig_idx[unlist(max_idx_per_cluster, use.names = FALSE)]
  
  # Return original summit entries (preserve metadata)
  highest_per_group_gr[chosen_orig_idx]
}


# -------------------------
# Step C: assign exactly one summit to each ROI (overlap preferred, fallback to nearest)
# returns a data.frame and a GRanges of assigned summits
# -------------------------
assign_best_summit_to_ROI_revised <- function(consensus_gr, ROIs_gr, max_dist = 500) {
  # Prepare output containers
  nROIs <- length(ROIs_gr)
  if (nROIs == 0) return(list(df = tibble::tibble(), gr = GRanges()))
  
  # 1) Exact overlaps: find all overlaps (ROI -> consensus)
  ov <- findOverlaps(ROIs_gr, consensus_gr)  # query = ROI, subject = consensus
  ov_df <- NULL
  if (length(ov) > 0) {
    ov_df <- data.frame(
      roi_idx = queryHits(ov),
      cons_idx = subjectHits(ov),
      summit_pos = start(consensus_gr)[subjectHits(ov)],
      summit_score = mcols(consensus_gr)$score[subjectHits(ov)],
      stringsAsFactors = FALSE
    )
    # For each ROI choose the overlapping consensus with highest score
    ov_df <- ov_df %>%
      group_by(roi_idx) %>%
      slice_max(summit_score, with_ties = FALSE) %>%
      ungroup()
  } else {
    ov_df <- data.frame(roi_idx = integer(0), cons_idx = integer(0),
                        summit_pos = integer(0), summit_score = numeric(0))
  }
  
  # 2) Find nearest consensus for ROIs that have no overlap
  assigned_roi_idx <- unique(ov_df$roi_idx)
  roi_unassigned <- setdiff(seq_len(nROIs), assigned_roi_idx)
  
  nearest_df <- data.frame(roi_idx = integer(0), cons_idx = integer(0),
                           summit_pos = integer(0), summit_score = numeric(0), distance = integer(0))
  if (length(roi_unassigned) > 0 && length(consensus_gr) > 0) {
    # distanceToNearest gives one nearest consensus per ROI (but we will subset to roi_unassigned)
    dn <- distanceToNearest(ROIs_gr, consensus_gr)
    dn_df <- data.frame(
      roi_idx = queryHits(dn),
      cons_idx = subjectHits(dn),
      distance = mcols(dn)$distance,
      stringsAsFactors = FALSE
    )
    # keep only the ROIs we need and apply max_dist cutoff
    dn_df <- dn_df %>% filter(roi_idx %in% roi_unassigned & distance <= max_dist)
    if (nrow(dn_df) > 0) {
      dn_df <- dn_df %>%
        mutate(summit_pos = start(consensus_gr)[cons_idx],
               summit_score = mcols(consensus_gr)$score[cons_idx]) %>%
        group_by(roi_idx) %>%
        # if multiple at same distance, choose higher score; distanceToNearest returns single nearest usually
        slice_max(summit_score, with_ties = FALSE) %>%
        ungroup()
      nearest_df <- as.data.frame(dn_df)
    }
  }
  
  # 3) Combine overlap assignments and nearest fallback
  assigned_df <- bind_rows(
    ov_df %>% select(roi_idx, cons_idx, summit_pos, summit_score),
    nearest_df %>% select(roi_idx, cons_idx, summit_pos, summit_score)
  )
  
  # 4) For ROIs that still have no assignment (no overlap and no nearest within max_dist), we will set NA
  unassigned_final <- setdiff(seq_len(nROIs), assigned_df$roi_idx)
  
  # 5) Build output dataframe with ROI metadata and consensus metadata
  if (nrow(assigned_df) > 0) {
    out_df <- assigned_df %>%
      left_join(
        tibble::tibble(roi_idx = seq_len(nROIs),
                       Peakid = ROIs_gr$Peakid,
                       roi_seqname = as.character(seqnames(ROIs_gr)),
                       roi_start = start(ROIs_gr),
                       roi_end = end(ROIs_gr)),
        by = "roi_idx"
      ) %>%
      mutate(dist_center = summit_pos - (roi_start + (roi_end - roi_start)/2),
             rel_pos = (summit_pos - roi_start) / (roi_end - roi_start)) %>%
      # attach consensus metadata columns (preserve all mcols)
      rowwise() %>%
      mutate(cons_meta = list(if (!is.na(cons_idx)) as.list(mcols(consensus_gr)[cons_idx, , drop = FALSE]) else list(NULL))) %>%
      ungroup()
    
    # explode the cons_meta list into columns (safe)
    cons_meta_df <- do.call(rbind, lapply(out_df$cons_meta, function(x) {
      if (is.null(x)) return(as.data.frame(matrix(NA, nrow = 1, ncol = 0)))
      as.data.frame(x, stringsAsFactors = FALSE)
    }))
    if (ncol(cons_meta_df) > 0) {
      out_df <- bind_cols(out_df %>% select(-cons_meta), cons_meta_df)
    } else {
      out_df <- out_df %>% select(-cons_meta)
    }
  } else {
    out_df <- tibble::tibble(
      roi_idx = integer(0), cons_idx = integer(0), summit_pos = integer(0),
      summit_score = numeric(0), Peakid = character(0), roi_seqname = character(0),
      roi_start = integer(0), roi_end = integer(0), dist_center = numeric(0), rel_pos = numeric(0)
    )
  }
  
  # 6) Append rows for ROIs that remain unassigned (all NA for summit info)
  if (length(unassigned_final) > 0) {
    missing_rows <- tibble::tibble(
      roi_idx = unassigned_final,
      cons_idx = NA_integer_,
      summit_pos = NA_integer_,
      summit_score = NA_real_,
      Peakid = ROIs_gr$Peakid[unassigned_final],
      roi_seqname = as.character(seqnames(ROIs_gr))[unassigned_final],
      roi_start = start(ROIs_gr)[unassigned_final],
      roi_end = end(ROIs_gr)[unassigned_final],
      dist_center = NA_real_,
      rel_pos = NA_real_
    )
    out_df <- bind_rows(out_df, missing_rows)
  }
  
  # Ensure ordering by ROI index to make downstream joins easy
  out_df <- out_df %>% arrange(roi_idx)
  
  # 7) Also return a GRanges of assigned summits (only those assigned)
  assigned_rows <- out_df %>% filter(!is.na(cons_idx))
  if (nrow(assigned_rows) > 0) {
    assigned_gr <- GRanges(
      seqnames = assigned_rows$roi_seqname,
      ranges = IRanges(start = assigned_rows$summit_pos, end = assigned_rows$summit_pos),
      Peakid = assigned_rows$Peakid
    )
    # Attach consensus metadata columns (if present in out_df)
    meta_cols <- setdiff(colnames(assigned_rows), c("roi_idx","cons_idx","summit_pos","summit_score","Peakid","roi_seqname","roi_start","roi_end","dist_center","rel_pos"))
    if (length(meta_cols) > 0) {
      mcols(assigned_gr) <- assigned_rows[, meta_cols, drop = FALSE]
    }
  } else {
    assigned_gr <- GRanges()
  }
  
  return(list(df = out_df, gr = assigned_gr))
}


# -------------------------
# Wrapper: run whole pipeline and produce final df + GRanges
# -------------------------
make_consensus_and_assign <- function(summits_gr, ROIs_gr,
                                      group_col = "group", score_col = "score",
                                      min_gap_within = 100, min_gap_across = 400, max_dist = 500) {
  highest_within <- get_highest_per_group_revised(summits_gr, group_col = group_col, score_col = score_col, min_gap_within = min_gap_within)
  consensus <- get_consensus_summits_revised(highest_within, score_col = score_col, min_gap_across = min_gap_across)
  assigned <- assign_best_summit_to_ROI_revised(consensus, ROIs_gr, max_dist = max_dist)
  
  # tidy return
  list(
    highest_within = highest_within,
    consensus = consensus,
    assigned_df = assigned$df,
    assigned_gr = assigned$gr
  )
}

```

```{r}
res <- make_consensus_and_assign(
  summits_gr = all_H3K27ac_summits_gr,
  ROIs_gr = H3K27ac_sets_gr$all_H3K27ac,
  group_col = "group",
  score_col = "score",
  min_gap_within = 100,
  min_gap_across = 400,
  max_dist = 500
)

# data.frame with one row per ROI (assigned or NA)
final_df <- res$assigned_df

# GRanges of ROIs that received an assigned summit (with metadata)
assigned_summits_gr <- res$assigned_gr

# Quick checks
n_missing <- sum(is.na(final_df$summit_pos))
cat("ROIs lacking any summit within max_dist:", n_missing, "\n")


```
Almost done, now to look at the 6,324 ROIs without summits
```{r}
ROIs_no_summit <- final_df %>% 
  dplyr::filter(is.na(file)) %>% 
  dplyr::select(Peakid)

length(ROIs_no_summit$Peakid)

no_summits_revised <- ROIs[(ROIs$Peakid %in% ROIs_no_summit$Peakid)]


missing_gr <- H3K27ac_sets_gr$all_H3K27ac[H3K27ac_sets_gr$all_H3K27ac$Peakid %in% ROIs_no_summit$Peakid]

nearest_hits <- nearest(missing_gr, all_H3K27ac_summits_gr)
nearest_dist <- start(all_H3K27ac_summits_gr)[nearest_hits] - start(missing_gr)

ROIs_no_summit$nearest_summit_dist <- nearest_dist
head(ROIs_no_summit)


ROIs_no_summit %>% 
  ggplot(., aes(x=nearest_summit_dist))+
  geom_histogram()+
  ggtitle("nearest summit distance to ROIs without summits")
```
#Fix Peakid loss

```{r}
# Example: fallback assignment
fallback_summits <- lapply(seq_along(ROIs_no_summit$Peakid), function(i) {
  roi <- ROIs[ROIs$Peakid == ROIs_no_summit$Peakid[i]]
  summits_in_roi <- all_H3K27ac_summits_gr[queryHits(findOverlaps(all_H3K27ac_summits_gr, roi))]
  if(length(summits_in_roi) == 0) return(NULL)
  
  # Pick highest scoring summit
  summits_in_roi[which.max(summits_in_roi$score)]
})

fallback_summits <- do.call(c, fallback_summits)

common_cols <- intersect(colnames(mcols(assigned_summits_gr)),
                         colnames(mcols(fallback_summits)))

assigned_summits_gr <- assigned_summits_gr[, common_cols, drop = FALSE]
fallback_summits <- fallback_summits[, common_cols, drop = FALSE]


mcols(fallback_summits)
mcols(assigned_summits_gr)

final_consensus_gr <- c(assigned_summits_gr, fallback_summits)

# check metadata
mcols(final_consensus_gr)       # per-range metadata


hits_peakid <- findOverlaps(final_consensus_gr, ROIs)

# assign Peakid from ROIs to the consensus summits
mcols(final_consensus_gr)$Peakid <- NA_character_  # initialize

mcols(final_consensus_gr)$Peakid[queryHits(hits_peakid)] <- ROIs$Peakid[subjectHits(hits_peakid)]

# check
head(mcols(final_consensus_gr))

final_consensus_annotated_df <- as.data.frame(final_consensus_gr) %>%
  left_join(H3K27ac_lookup, by = "Peakid") %>%
  mutate(cluster = if_else(is.na(cluster), "not_assigned", cluster))
```

# Fix resave with peakid and set
```{r redoing cluster membership and export files, eval=FALSE}


### adding in cluster membership for export

Final_df_annotated <- final_df  %>% 
  left_join(H3K27ac_lookup)

SET_1_df <- Final_df_annotated %>% 
  dplyr::filter(cluster=="Set_1") %>% 
  dplyr::select(roi_seqname, summit_pos,Peakid) %>% 
  na.omit()
Set_1_gr <- GRanges(
  seqnames = SET_1_df$roi_seqname,
  ranges = IRanges(start = SET_1_df$summit_pos, end = SET_1_df$summit_pos),
  name = SET_1_df$Peakid
)

rtracklayer::export(Set_1_gr, "data/Bed_exports/Set_1_summits.bed")


SET_2_df <- Final_df_annotated %>% 
  dplyr::filter(cluster=="Set_2") %>% 
  dplyr::select(roi_seqname, summit_pos,Peakid) %>% 
  na.omit()
Set_2_gr <- GRanges(
  seqnames = SET_2_df$roi_seqname,
  ranges = IRanges(start = SET_2_df$summit_pos, end = SET_2_df$summit_pos),
  name = SET_2_df$Peakid
)

rtracklayer::export(Set_2_gr, "data/Bed_exports/Set_2_summits.bed")

SET_3_df <- Final_df_annotated %>% 
  dplyr::filter(cluster=="Set_3") %>% 
  dplyr::select(roi_seqname, summit_pos,Peakid) %>% 
  na.omit()

Set_3_gr <- GRanges(
  seqnames = SET_3_df$roi_seqname,
  ranges = IRanges(start = SET_3_df$summit_pos, end = SET_3_df$summit_pos),
  name = SET_3_df$Peakid
)

rtracklayer::export(Set_3_gr, "data/Bed_exports/Set_3_summits.bed")

# rtracklayer::export(all_H3K27ac_summits_gr,"data/Bed_exports/All_summits_bedfile.bed")

# rtracklayer::export(ROIs,"data/Bed_exports/All_ROIs.bed")

```

