---
title: "Summit Files"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```


```{r packages}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(BiocParallel)
library(parallel)
library(future.apply)
```

```{r calling summit files}
sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")
##Path to histone summit files
H3K27ac_dir <- "C:/Users/renee/Other_projects_data/DXR_data/final_data/summit_files/H3K27ac" 

##pull all histone files together
H3K27ac_summit_files <- list.files(
  path = H3K27ac_dir,
  pattern = "\\.bed$",
  recursive = TRUE,
  full.names = TRUE
)
length(H3K27ac_summit_files)

# head(H3K27ac_summit_files)
```
First steps:  pulling in the information of Sets and locations
Pulling in the summit files,
concatenation of all summits into one large file

```{r pulling in other data}
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")
H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})

read_summit <- function(file){
  peaks <- read.table(file,header = FALSE)
  colnames(peaks) <- c("chr","start","end","name","score")
  GRanges(
    seqnames = peaks$chr,
    ranges = IRanges(start=peaks$start, end = peaks$start),
    score=peaks$score,
    file=basename(file),
    Library_ID = stringr::str_remove(basename(file), "_FINAL_summits\\.bed$")
  )
}

all_H3K27ac_summits_list<- lapply(H3K27ac_summit_files, read_summit)
all_H3K27ac_summits_gr <- do.call(c, all_H3K27ac_summits_list)  # combine into one GRanges object
H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
```

```{r pick summit per roi}
pick_best_summit_per_roi <- function(roi_gr, summit_gr, score_col = "signalValue") {

  hits <- findOverlaps(roi_gr, summit_gr)

  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    score = mcols(summit_gr)[[score_col]][subjectHits(hits)]
  )

  # Pick summit with highest score for each ROI
  best <- df %>%
    group_by(roi_idx) %>%
    slice_max(order_by = score, n = 1) %>%
    ungroup()

  # Build the final GRanges object
  best_summits <- summit_gr[best$summit_idx]
  best_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]

  best_summits
}
# 
# best_summits <- pick_best_summit_per_roi(
#   ROIs, highest_summits_long_gr, score_col = "peakHeight"
# )
```

```{r pick summit closest to roi center}
pick_center_summit <- function(roi_gr, summit_gr) {

  hits <- findOverlaps(roi_gr, summit_gr)

  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    dist_to_center = abs(
      start(summit_gr)[subjectHits(hits)] -
      round((start(roi_gr)[queryHits(hits)] + end(roi_gr)[queryHits(hits)]) / 2)
    )
  )

  best <- df %>%
    group_by(roi_idx) %>%
    slice_min(order_by = dist_to_center, n = 1) %>%
    ungroup()

  best_summits <- summit_gr[best$summit_idx]
  best_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]

  best_summits
}
```

```{r consensus summit}
pick_consensus_summit <- function(roi_gr, summit_gr, score_col = "score") {
  
  # Find which summits overlap each ROI
  hits <- findOverlaps(roi_gr, summit_gr)
  
  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    summit_pos = start(summit_gr)[subjectHits(hits)],
    score = mcols(summit_gr)[[score_col]]
  )
  
  # Count frequency per summit position within each ROI
  freq_df <- df %>%
    group_by(roi_idx, summit_pos) %>%
    summarise(
      freq = n(),
      max_score = max(score),
      .groups = "drop"
    )
  
  # Pick the summit with highest frequency; break ties with max score
  best <- freq_df %>%
    group_by(roi_idx) %>%
    slice_max(order_by = freq, n = 1, with_ties = TRUE) %>%
    slice_max(order_by = max_score, n = 1) %>%
    ungroup()
  
  # Map back to original GRanges
  best_idx <- df$summit_idx[match(
    paste0(best$roi_idx, "_", best$summit_pos),
    paste0(df$roi_idx, "_", df$summit_pos)
  )]
  
  final_summits <- summit_gr[best_idx]
  final_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]
  
  final_summits
}
```

### Analysis of all summit files, no grouping  


first creating summit clusters that are around 100bps across all summit files
```{r pulling in other data again}
 
#  Merge summits within 100 bp clusters but keep metadata
# ------------------------
# 1 Reduce with revmap to keep track of original indices

clusters <- GenomicRanges::reduce(all_H3K27ac_summits_gr, min.gapwidth = 100,
                   ignore.strand = TRUE, with.revmap = TRUE)

# 2 For each cluster, pick the highest-score summit
scores <- mcols(all_H3K27ac_summits_gr)$score

revmap <- clusters$revmap

# Compute the highest-score index per cluster
highest_idx <- sapply(revmap, function(idx) idx[which.max(scores[idx])])

# Subset GRanges once
highest_per_cluster_gr <- all_H3K27ac_summits_gr[highest_idx]


# ------------------------
# 3 Count merged summits per ROI
# ------------------------
ROIs <- H3K27ac_sets_gr$all_H3K27ac  # your ROI GRanges

# Optimized counting
roi_counts <- countOverlaps(ROIs, highest_per_cluster_gr)
mcols(ROIs)$merged_summit_count <- roi_counts

# ------------------------
# 5 Map ROIs to clusters with sample info
# ------------------------
overlaps <- findOverlaps(ROIs, highest_per_cluster_gr)

roi_hits_df <- as.data.frame(overlaps) %>%
  mutate(
    Peakid = ROIs$Peakid[queryHits],
    Library_ID = mcols(highest_per_cluster_gr)$Library_ID[subjectHits],
    score = mcols(highest_per_cluster_gr)$score[subjectHits],
    file = mcols(highest_per_cluster_gr)$file[subjectHits]
  ) %>%
  left_join(sampleinfo, by = c("Library_ID" = "Library ID"))

## initualize count column on ROI df
mcols(ROIs)$summit_count <- 0

# Tabulate counts
counts <- table(queryHits(overlaps))
mcols(ROIs)$summit_count[as.numeric(names(counts))] <- as.numeric(counts)

# Convert to dataframe for plotting
roi_counts_df <- as.data.frame(ROIs) %>%
  select(Peakid, seqnames, start, end, summit_count) %>% 
  mutate(roi_size=end-start)

# Optional: add Set / cluster info
roi_counts_df <- roi_counts_df %>%
  left_join(H3K27ac_lookup, by = "Peakid")
# ------------------------
# 6 Optional: add Set/cluster info per ROI
# ------------------------
roi_hits_df <- roi_hits_df %>%
  left_join(H3K27ac_lookup, by = "Peakid")

# Now roi_hits_df is ready for plotting or analysis:
# Columns include Peakid, ROI coordinates, Library_ID, score, file, Individual, Treatment, Timepoint, cluster


# Optional: add sample info
roi_hits_df <- as.data.frame(overlaps) %>%
  mutate(
    Peakid = ROIs$Peakid[queryHits],
    Library_ID = mcols(all_H3K27ac_summits_gr)$Library_ID[subjectHits],
    score = mcols(all_H3K27ac_summits_gr)$score[subjectHits],
    file = mcols(all_H3K27ac_summits_gr)$file[subjectHits]
  ) %>%
  left_join(sampleinfo, by = c("Library_ID" = "Library ID"))

```

Plotting some summits from the first process

```{r get summits function plots}
ROIs %>% 
  as.data.frame() %>% 
  # dplyr::filter(merged_summit_count==1) %>% 
  left_join(H3K27ac_lookup, by = "Peakid") %>%  # make sure join key matches
  filter(!is.na(cluster)) %>%
  ggplot(aes(x = cluster, y = merged_summit_count)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6, size = 2, color = "steelblue") +
  theme_bw() +
  xlab("Set") +
  ylab("Number of merged summits per ROI") +
  ggtitle("Distribution of merged summit counts by Set")


roi_counts_df %>%
  ggplot(aes(x = summit_count)) +
  geom_histogram(binwidth = 1, fill="steelblue", color="black") +
  theme_bw() +
  xlab("Number of summits per ROI") +
  ylab("Number of ROIs")

roi_counts_df %>%
  ggplot(aes(x = summit_count)) +
  geom_histogram(binwidth = 1, fill="steelblue", color="black") +
  theme_bw() +
  xlab("Number of summits per ROI") +
  ylab("Number of ROIs")+
  ggtitle("Zoomed in summit per ROI")+
  coord_cartesian(xlim=c(0,50))

roi_counts_df %>% 
  # group_by(Peakid) %>% tally #%>% 
  # left_join(H3K27ac_lookup) %>% 
  ggplot(aes(x = roi_size, y = summit_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "ROI size (bp)",
    y = "# of summits",
    title = "Relationship between ROI size and # of Summit clusters"
  ) +
  theme_classic()#+
  # coord_cartesian(xlim=c(-1,400), ylim=c(0,10))

roi_counts_df %>% 
  # group_by(Peakid) %>% tally #%>% 
  # left_join(H3K27ac_lookup) %>% 
  ggplot(aes(x = roi_size, y = summit_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "ROI size (bp)",
    y = "# of summits",
    title = "Zoomed in Relationship between ROI size and # of Summit clusters"
  ) +
  theme_classic()+
  coord_cartesian(xlim=c(-1,800), ylim=c(0,10))

```


### Merging strategies!

This is now the place I will take my files and now try to apply the same merging strategy as I did the peaks merging strategy.  Step 1: reduce all trt-time summits within 100 bp
```{r reduction by trt-time}
###Adding in sampleinfo dataframe
meta <- as.data.frame(mcols(all_H3K27ac_summits_gr))
meta2 <- meta %>% 
  left_join(., sampleinfo, by=c("Library_ID"="Library ID"))

mcols(all_H3K27ac_summits_gr) <- meta2

mcols(all_H3K27ac_summits_gr)$group <- 
  paste(all_H3K27ac_summits_gr$Treatment,
        all_H3K27ac_summits_gr$Timepoint,
        sep = "_")

###now splitting into grouped granges 

gr_by_group <- split(all_H3K27ac_summits_gr,
                                    all_H3K27ac_summits_gr$group)
# gr_by_group <- as(gr_by_group, "CompressedGRangesList")
### now reducing within some width by 100 bp with revmap
groups <- unique(all_H3K27ac_summits_gr$group)

reduced_groups <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 100, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_200 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 200, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_300 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 300, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_400 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 400, ignore.strand = TRUE, with.revmap = TRUE)
})

names(reduced_groups) <- groups
names(reduced_groups_200) <- groups
names(reduced_groups_300) <- groups
names(reduced_groups_400) <- groups

# redux_main_100 <- sum(sapply(reduced_groups, length))
# redux_main_200 <- sum(sapply(reduced_groups_200, length))
# redux_main_300 <- sum(sapply(reduced_groups_300, length))
# redux_main_400 <- sum(sapply(reduced_groups_400, length))

```

### Reduction plots  

Plotting effect of reduction bp number on total number of clusters
```{r}
gap_sizes <- seq(0, 500, by = 10)


# Function to apply reduce for each gap and return counts
results <- lapply(gap_sizes, function(g) {
  reduced <- GenomicRanges::reduce(all_H3K27ac_summits_gr, min.gapwidth = g)
  data.frame(gap = g, n_regions = length(reduced))
})

# Combine results
min_gap_summary <- bind_rows(results)

ggplot(min_gap_summary, aes(x = gap, y= n_regions))+
  geom_line(linewidth=2)+
  geom_point()+
  theme_bw(base_size=14)+
  labs(
    title = "Effect of Gap Size on Reduced Summits",
    x = "Gap size (bp)",
    y = "Number of merged regions"
  )

```
```{r get highes per group funciton}
# Function to pick highest summit per cluster from a reduced GRanges list
get_highest_per_group <- function(reduced_groups, orig_summits_gr, group_col = "group") {
  
  groups <- names(reduced_groups)
  highest_per_group <- vector("list", length(reduced_groups))
  names(highest_per_group) <- groups
  
  for(i in seq_along(reduced_groups)) {
    gr <- reduced_groups[[i]]
    
    # original summits for this group
    orig <- orig_summits_gr[mcols(orig_summits_gr)[[group_col]] == groups[i]]
    scores <- orig$score
    
    # revmap is a CompressedIntegerList
    revmap <- mcols(gr)$revmap
    
    # skip if revmap is NULL
    if(is.null(revmap)) next
    
    # unlist all indices once
    all_idx <- unlist(revmap, use.names = FALSE)
    
    # repeat cluster index for each element in revmap
    cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
    
    # scores for all indices
    all_scores <- scores[all_idx]
    
    # For each cluster, pick the index of the max score
    max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
      ii[which.max(all_scores[ii])]
    })
    
    # Convert back to original indices
    orig_idx <- all_idx[unlist(max_idx_per_cluster)]
    
    # subset original GRanges
    highest_per_group[[i]] <- orig[orig_idx]
  }
  
  # Flatten any nested GRangesList
  flatten_gr <- function(x) {
    if (inherits(x, "GRanges")) return(x)
    if (inherits(x, "GRangesList")) return(unlist(x, use.names = FALSE))
    if (is.list(x)) return(do.call(c, lapply(x, flatten_gr)))
    stop("Unexpected object type")
  }
  
  highest_summits_gr <- flatten_gr(highest_per_group)
  
  # Return as long GRanges with group column
  highest_summits_df <- bind_rows(
    lapply(names(highest_per_group), function(gr_name) {
      as.data.frame(highest_per_group[[gr_name]]) %>%
        mutate(group = gr_name)
    })
  )
  
  highest_summits_long_gr <- highest_summits_df %>% GRanges()
  
  return(highest_summits_long_gr)
}
```


```{r}
reduced_sets <- list(
  "100bp" = reduced_groups,
  "200bp" = reduced_groups_200,
  "300bp" = reduced_groups_300,
  "400bp" = reduced_groups_400
)

highest_summits_all <- lapply(reduced_sets, get_highest_per_group, orig_summits_gr = all_H3K27ac_summits_gr)
```
  
  Looking at number of summits across groups as a function of min.gap number

```{r across groups mingap}
 ##Compute counts per group for each reduced set
summit_counts_group <- lapply(names(highest_summits_all), function(gap_name) {
  gr <- highest_summits_all[[gap_name]]
  
  # make sure group column exists
  if(!"group" %in% colnames(mcols(gr))) stop("GRanges must have 'group' column")
  
  df <- as.data.frame(gr) %>%
    count(group, name = "n_summits") %>%
    mutate(gap = gap_name)
  
  return(df)
}) %>% bind_rows()


ggplot(summit_counts_group, aes(x = gap, y = n_summits, group = group, color = group)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  theme_classic(base_size = 14) +
  labs(
    title = "Effect of min-gap width on number of highest summits per group",
    x = "Min-gap width (bp)",
    y = "Number of highest summits",
    color = "Group"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
### Summit average per ROI
What is the average number of summits per ROI as a function of min.gapwidth

```{r double reduce function}
double_reduce_summits <- function(summits_gr, ROIs, groups,
                                  min_gap_within_seq = c(100,200,300),
                                  min_gap_across_seq = c(100,200,300),
                                  BPPARAM = MulticoreParam(4)) {
  
  # Pre-split summits by group to avoid repeated subsetting
  summits_by_group <- split(summits_gr, summits_gr$group)
  
  # Create all gap combinations
  gap_combos <- expand.grid(min_gap_within = min_gap_within_seq,
                            min_gap_across = min_gap_across_seq,
                            stringsAsFactors = FALSE)
  
  # Apply in parallel for each combination
  results <- bplapply(seq_len(nrow(gap_combos)), function(i) {
    g1 <- gap_combos$min_gap_within[i]
    g2 <- gap_combos$min_gap_across[i]
    
    # -----------------
    # Step 1: Reduce within group
    # -----------------
    highest_per_group <- lapply(groups, function(grp) {
      gr_sub <- summits_by_group[[grp]]
      if(length(gr_sub) == 0) return(NULL)
      
      # reduce within group with revmap
      red <- GenomicRanges::reduce(gr_sub, min.gapwidth = g1, ignore.strand = TRUE, with.revmap = TRUE)
      revmap <- mcols(red)$revmap
      if(length(revmap) == 0) return(NULL)
      
      # pick highest score per cluster
      cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
      all_idx <- unlist(revmap, use.names = FALSE)
      all_scores <- gr_sub$score[all_idx]
      max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
        ii[which.max(all_scores[ii])]
      })
      gr_sub[all_idx[unlist(max_idx_per_cluster)]]
    })
    
    highest_per_group <- highest_per_group[!sapply(highest_per_group, is.null)]
    
    # -----------------
    # Step 2: Merge across groups
    # -----------------
    if(length(highest_per_group) == 0) return(NULL)
    all_highest <- do.call(c, highest_per_group)
    consensus <- GenomicRanges::reduce(all_highest, min.gapwidth = g2, ignore.strand = TRUE)
    
    # -----------------
    # Step 3: Count summits per ROI (vectorized)
    # -----------------
    hits <- findOverlaps(ROIs, consensus)
    counts <- as.data.frame(table(queryHits(hits)))
    colnames(counts) <- c("ROI_idx", "n_summits")
    counts$ROI_idx <- as.integer(as.character(counts$ROI_idx))
    counts$min_gap_within <- g1
    counts$min_gap_across <- g2
    counts
  }, BPPARAM = BPPARAM)
  
  # Combine results
  results_df <- bind_rows(results)
  
  return(results_df)
}
```

```{r calculating of double reduce}

BPPARAM <- SnowParam(workers = 4, type = "SOCK")  
register(BPPARAM)

heatmap_data <- double_reduce_summits(
  summits_gr = all_H3K27ac_summits_gr,
  ROIs = H3K27ac_sets_gr$all_H3K27ac,
  groups = unique(all_H3K27ac_summits_gr$group),
  min_gap_within_seq = c(100,200,300,400),
  min_gap_across_seq = c(100,200,300,400),
  BPPARAM = BPPARAM
)

heatmap_avg <- heatmap_data %>%
  group_by(min_gap_within, min_gap_across) %>%
  summarise(mean_summits = mean(n_summits, na.rm = TRUE), .groups = "drop")

```

## Exploring differences in min gaps in step 1 and step 2 reductions

```{r plot mingap averages}
ggplot(heatmap_avg, aes(x = factor(min_gap_within),
                        y = factor(min_gap_across),
                        fill = mean_summits)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(mean_summits, 1)), color = "black", size = 4) +
  scale_fill_viridis_c(option = "plasma") +
  labs(
    x = "Within-group min-gap (bp)",
    y = "Across-group min-gap (bp)",
    fill = "Mean # summits per ROI",
    title = "Effect of min-gap width on summits per ROI"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

after i figure my final min.gaps, here is where the places are located
```{r}

# --- Step 1: Reduce within groups and pick highest summit per cluster ---
get_highest_per_group_final <- function(summits_gr, groups = NULL, min_gap_within = 100) {
  
  if(is.null(groups)) {
    groups <- unique(summits_gr$group)
  }
  
  highest_per_group <- lapply(groups, function(grp) {
    gr_sub <- summits_gr[summits_gr$group == grp]
    if(length(gr_sub) == 0) return(NULL)
    
    # Reduce within group
    reduced <- GenomicRanges::reduce(gr_sub, min.gapwidth = min_gap_within, ignore.strand = TRUE, with.revmap = TRUE)
    
    # Pick highest scoring summit per cluster
    revmap <- reduced$revmap
    scores <- gr_sub$score
    
    all_idx <- unlist(revmap, use.names = FALSE)
    cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
    all_scores <- scores[all_idx]
    
    max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
      ii[which.max(all_scores[ii])]
    })
    
    orig_idx <- all_idx[unlist(max_idx_per_cluster)]
    
    gr_sub[orig_idx]
  })
  
  # Flatten list of GRanges
  highest_per_group <- do.call(c, highest_per_group)
  return(highest_per_group)
}

# --- Step 2: Reduce across groups to get final consensus ---
get_consensus_summits <- function(highest_per_group_gr, min_gap_across = 400) {
  
  # Reduce across groups
  reduced <- GenomicRanges::reduce(highest_per_group_gr, min.gapwidth = min_gap_across, ignore.strand = TRUE, with.revmap = TRUE)
  
  # Pick highest scoring summit per cluster across all groups
  revmap <- reduced$revmap
  scores <- highest_per_group_gr$score
  
  all_idx <- unlist(revmap, use.names = FALSE)
  cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
  all_scores <- scores[all_idx]
  
  max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
    ii[which.max(all_scores[ii])]
  })
  
  orig_idx <- all_idx[unlist(max_idx_per_cluster)]
  
  highest_per_group_gr[orig_idx]
}

add_ROI_to_consensus <- function(consensus_gr, ROIs_gr) {
  
  hits <- findOverlaps(consensus_gr, ROIs_gr)
   # Extract metadata from consensus_gr
  consensus_meta <- mcols(consensus_gr)[queryHits(hits), , drop = FALSE]
  
  
  df <- data.frame(
    summit_chr  = seqnames(consensus_gr)[queryHits(hits)],
    summit_pos  = start(consensus_gr)[queryHits(hits)],
    summit_score = consensus_gr$score[queryHits(hits)],
    Peakid      = ROIs_gr$Peakid[subjectHits(hits)],
    roi_start   = start(ROIs_gr)[subjectHits(hits)],
    roi_end     = end(ROIs_gr)[subjectHits(hits)]
  )
  
  # Bind the metadata from consensus_gr
  df <- cbind(df, as.data.frame(consensus_meta))
  
  df %>%
    mutate(rel_pos = (summit_pos - roi_start)/(roi_end - roi_start),
           roi_center = roi_start + (roi_end - roi_start)/2,
           dist_center = summit_pos - roi_center)
}


```

```{r looking at 100-400}
highest_100 <- get_highest_per_group_final(all_H3K27ac_summits_gr, min_gap_within = 100)

final_consensus <- get_consensus_summits(highest_100, min_gap_across = 400)

final_df_100_400 <- add_ROI_to_consensus(final_consensus, H3K27ac_sets_gr$all_H3K27ac)


ggplot(final_df_100_400, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("100-400 run comparison")
```


```{r distance from center plot}

final_df_100_400 %>% 
  group_by(Peakid) %>% 
  slice_max(summit_score) %>% 
ggplot(., aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  geom_vline(xintercept = 250, linetype = "dashed",color="yellow") +  # left 500 
  geom_vline(xintercept = -250, linetype = "dashed",color="yellow") +  # right 500
  
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  ggtitle("100-400 run comparison")+
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
 coord_cartesian(xlim=c(-2000,2000))

final_df_100_400 %>% 
  group_by(Peakid) %>% 
  slice_max(summit_score) %>% 
ggplot(., aes(x = dist_center)) +
  geom_density()+
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  geom_vline(xintercept = 250, linetype = "dashed",color="blue") +  # left 500 
  geom_vline(xintercept = -250, linetype = "dashed",color="blue") +  # right 500
 coord_cartesian(xlim=c(-2000,2000))+
  ggtitle("100-400 run comparison")

```

```{r looking at 200-400}
highest_200 <- get_highest_per_group_final(all_H3K27ac_summits_gr, min_gap_within = 200)

final_consensus2 <- get_consensus_summits(highest_200, min_gap_across = 400)

final_df_200_400 <- add_ROI_to_consensus(final_consensus2, H3K27ac_sets_gr$all_H3K27ac)


ggplot(final_df_200_400, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color="green") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  ggtitle("200-400 run comparison")+
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+coord_cartesian(xlim=c(-2000,2000))
```

#### function get highest scoring summit clustered by group:

```{r, eval=FALSE}
#' Get highest summit per reduced cluster per group using plyranges
#'
#' @param gr GRanges with metadata columns: group, score
#' @param min_gap numeric, minimum gap width to merge summits
#' @param ignore.strand logical, whether to ignore strand when reducing
#'
#' @return GRanges of highest scoring summit per reduced cluster per group
get_highest_summits_with_cluster_id <- function(gr, min_gap = 100, ignore.strand = TRUE) {
  
  gr %>%
    # Reduce summits per group
    group_by(group) %>%
    reduce_ranges(min_gap = min_gap, ignore.strand = ignore.strand, with_revmap = TRUE) %>%
    # Add a cluster_id based on row_number()
    mutate(cluster_id = paste0(group, "_cluster_", row_number())) %>%
    # join back original summits
    join_overlap_inner(gr, suffix = c(".cluster", ".orig")) %>%
    # slice max per cluster_id
    group_by(cluster_id) %>%
    slice_max(score.orig, with_ties = FALSE) %>%
    ungroup() %>%
    # select only the original summit coordinates
    select(seqnames, start = start.orig, end = end.orig, score = score.orig,
           group, cluster_id)
}
```


#### function reduce_and_pick_highest

```{r}

#### this quickly reduces any granges summit  collection by the gap that is passed in the function. It returns the highest scoring summit from the cluster of summits.
reduce_and_pick_highest <- function(gr, gap = 100) {

  # Preextract score once
  sc <- gr$score

  # Fast reduce with revmap
  red <- GenomicRanges::reduce(
    gr,
    min.gapwidth = gap,
    ignore.strand = TRUE,
    with.revmap = TRUE
  )

  # This is vectorized and very fast
  idx_list <- red$revmap

  # Pick highest-scoring summit within each cluster
  sel <- IntegerList(
    lapply(idx_list, function(idx) idx[which.max(sc[idx])])
  )

  # Convert IntegerList → integer vector
  best_idx <- unlist(sel, use.names = FALSE)

  # Subset original GRanges
  gr[best_idx]
}

```

```{r, eval=FALSE}
redux_100 <- reduce_and_pick_highest(all_H3K27ac_summits_gr, gap = 100)
redux_200 <- reduce_and_pick_highest(all_H3K27ac_summits_gr, gap = 200)
redux_300 <- reduce_and_pick_highest(all_H3K27ac_summits_gr, gap = 300)

redux_Peak_100 <- add_ROI_to_consensus(redux_100, H3K27ac_sets_gr$all_H3K27ac)
redux_Peak_200 <- add_ROI_to_consensus(redux_200, H3K27ac_sets_gr$all_H3K27ac)
redux_Peak_300 <- add_ROI_to_consensus(redux_300, H3K27ac_sets_gr$all_H3K27ac)
```


```{r plotting alternates, eval=FALSE}
ggplot(redux_Peak_100, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("Only single reduction 100bp")

ggplot(redux_Peak_200, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("Only single reduction 200bp")

ggplot(redux_Peak_300, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("Only single reduction 300bp")

```
### Creating Final 100-400 reduction summit set  

Here is how I ran the get_highest_per_group_final

```{r 100 400 highest per group}

### step 1:  reduce within group using 100 bp gap
highest_100_final <- get_highest_per_group_final(all_H3K27ac_summits_gr, min_gap_within = 100)


### step 2: reduce across groups (converge)
final_consensus_final <- get_consensus_summits(highest_100_final, min_gap_across = 400)

### step 3: add in ROI information

final_df_summits <- add_ROI_to_consensus(final_consensus_final, H3K27ac_sets_gr$all_H3K27ac)
```

```{r, eval=FALSE}
final_df_summits %>% 
  ggplot(., aes(x=rel_pos))+
  geom_histogram(bins = 100)+
  geom_vline(xintercept = 0.5)+
  ggtitle("Relative position of consensus summits by ROI")


```


#### Debugging why some ROIs do not have summits when they clearly do in IGV  

```{r deb first step}


### step 1: Finding which ROIs are without summits
hits <- findOverlaps(ROIs, final_consensus_final)

roi_has_hit <- unique(queryHits(hits))

roi_index <- seq_along(ROIs)

# ROIs with no summits
no_summit_idx <- setdiff(roi_index, roi_has_hit)

missing_ROIs <- ROIs[no_summit_idx]
missing_ROIs


```

```{r debug step2}
### Step 2: Check if ROIs ever had a summit to begin with
raw_hits <- findOverlaps(ROIs, all_H3K27ac_summits_gr)
missing_raw_idx <- setdiff(roi_index, unique(queryHits(raw_hits)))

ROIs[missing_raw_idx]

### This returns nothing, which means I have a reduction problem.


```

No ROIs were summit-less, which means there is a problem in the reduction process.  Let's see if I can suss that out.

```{r debug step3}

### This code checked for all ~14,000 ROIs without summits to see if there were any nearby (It runs FOREVER)
# storage <- for(i in no_summit_idx) {
#   roi <- ROIs[i]
#   nearby = subsetByOverlaps(all_H3K27ac_summits_gr, roi, maxgap = 1000)
# 
#   cat("\nROI", ROIs$Peakid[i], "had", length(nearby), "nearby summits\n")
# }

## now we are seeing where those summits may be
first_level <- get_highest_per_group_final(all_H3K27ac_summits_gr, min_gap_within = 100)

# ROIs with summits after first reduction
hits1 <- findOverlaps(ROIs, first_level)
roi1 <- unique(queryHits(hits1))

# After second reduction
hits2 <- findOverlaps(ROIs, final_consensus_final)
roi2 <- unique(queryHits(hits2))

lost_after_consensus <- setdiff(roi1, roi2)
ROIs[lost_after_consensus]







i <- no_summit_idx[4]

roi <- ROIs[i]

inside_raw  <- subsetByOverlaps(all_H3K27ac_summits_gr, roi)
inside_lvl1 <- subsetByOverlaps(first_level, roi)
inside_lvl2 <- subsetByOverlaps(final_consensus_final, roi)

nearby_lvl1 <- subsetByOverlaps(first_level, roi, maxgap = 500)
nearby_lvl2 <- subsetByOverlaps(final_consensus_final, roi, maxgap = 500)

list(
  raw = length(inside_raw),
  reduced_within = length(inside_lvl1),
  reduced_across = length(inside_lvl2),
  nearby_after_lvl1 = length(nearby_lvl1),
  nearby_after_lvl2 = length(nearby_lvl2)
)

```

```{r debug step4}
debug_roi <- function(roi_index, ROIs, summits) {
  library(GenomicRanges)

  roi <- ROIs[roi_index]

  # Summits overlapping the ROI ± 500 bp
  nearby <- summits[
    summits@seqnames == roi@seqnames &
    summits@ranges@start >= roi@ranges@start - 500 &
    summits@ranges@end   <= roi@ranges@end + 500
  ]

  cat("\n### ROI:", roi_index, as.character(roi), "\n")
  cat("Nearby summits:", length(nearby), "\n")

  # Cluster summits within sets (level 1)
  lvl1 <- reduce(split(nearby, nearby$set), min.gapwidth = 1)
  lvl1_flat <- unlist(lvl1)
  cat("Level 1 reduced:", length(lvl1_flat), "\n")

  # Collapse across sets (level 2)
  lvl2 <- reduce(lvl1_flat, min.gapwidth = 1)
  cat("Level 2 reduced (ACROSS):", length(lvl2), "\n")

  # Show coordinates for inspection
  return(list(
    roi        = roi,
    nearby     = nearby,
    lvl1       = lvl1_flat,
    lvl2       = lvl2
  ))
}

```

### newly revised function  


```{r revised function}

# -------------------------
# Step 1: Reduce within groups and pick top per cluster
# -------------------------
get_highest_per_group_fast <- function(summits_gr, group_col = "group", score_col = "score", min_gap_within = 100) {
  groups <- unique(mcols(summits_gr)[[group_col]])
  
  highest_list <- lapply(groups, function(g) {
    gr_sub <- summits_gr[mcols(summits_gr)[[group_col]] == g]
    if(length(gr_sub) == 0) return(GRanges())
    
    red <- GenomicRanges::reduce(gr_sub, min.gapwidth = min_gap_within, ignore.strand = TRUE, with.revmap = TRUE)
    revmap <- mcols(red)$revmap
    
    # For each reduced cluster, pick the summit with highest score
    idx <- unlist(lapply(revmap, function(x) {
      scores <- mcols(gr_sub)$score[x]
      x[which.max(scores)]
    }))
    
    gr_sub[idx]
  })
  
  do.call(c, highest_list)
}

# -------------------------
# Step 2: Reduce across groups and pick top per cluster
# -------------------------
get_consensus_summits_fast <- function(highest_per_group_gr, score_col = "score", min_gap_across = 400) {
  if(length(highest_per_group_gr) == 0) return(GRanges())
  
  red <- GenomicRanges::reduce(highest_per_group_gr, min.gapwidth = min_gap_across, ignore.strand = TRUE, with.revmap = TRUE)
  revmap <- mcols(red)$revmap
  
  idx <- unlist(lapply(revmap, function(x) {
    scores <- mcols(highest_per_group_gr)$score[x]
    x[which.max(scores)]
  }))
  
  highest_per_group_gr[idx]
}

# -------------------------
# Step 3: Assign summits to ROIs (with fallback)
# -------------------------
assign_best_summit_to_ROI_fast <- function(consensus_gr, ROIs_gr, max_dist = 500) {
  library(GenomicRanges)
  library(dplyr)
  
  assigned_list <- vector("list", length = length(unique(seqnames(ROIs_gr))))
  chr_names <- unique(seqnames(ROIs_gr))
  
  cat("Processing", length(chr_names), "chromosomes...\n")
  
  for (i in seq_along(chr_names)) {
    chr <- chr_names[i]
    cat(sprintf("Chromosome %s (%d/%d)\n", chr, i, length(chr_names)))
    
    rois_chr <- ROIs_gr[seqnames(ROIs_gr) == chr]
    cons_chr <- consensus_gr[seqnames(consensus_gr) == chr]
    nROIs <- length(rois_chr)
    
    if (nROIs == 0 || length(cons_chr) == 0) next
    
    # 1) Exact overlaps
    ov <- findOverlaps(rois_chr, cons_chr)
    assigned_df <- NULL
    if (length(ov) > 0) {
      assigned_df <- tibble(
        roi_idx = queryHits(ov),
        cons_idx = subjectHits(ov),
        summit_pos = start(cons_chr)[subjectHits(ov)],
        summit_score = mcols(cons_chr)$score[subjectHits(ov)]
      ) %>%
        group_by(roi_idx) %>%
        slice_max(summit_score, with_ties = FALSE) %>%
        ungroup()
    }
    
    # 2) Nearest fallback
    assigned_idx <- assigned_df$roi_idx %||% integer(0)
    roi_unassigned <- setdiff(seq_len(nROIs), assigned_idx)
    
    if (length(roi_unassigned) > 0) {
      dn <- distanceToNearest(rois_chr[roi_unassigned], cons_chr)
      dn_df <- tibble(
        roi_idx = queryHits(dn),
        cons_idx = subjectHits(dn),
        distance = mcols(dn)$distance,
        summit_pos = start(cons_chr)[subjectHits(dn)],
        summit_score = mcols(cons_chr)$score[subjectHits(dn)]
      ) %>%
        filter(distance <= max_dist) %>%
        group_by(roi_idx) %>%
        slice_max(summit_score, with_ties = FALSE) %>%
        ungroup()
      
      assigned_df <- bind_rows(assigned_df, dn_df)
    }
    
    # Attach ROI metadata
    if (!is.null(assigned_df)) {
      roi_meta <- tibble(
        roi_idx = seq_len(nROIs),
        Peakid = rois_chr$Peakid,
        roi_seqname = as.character(seqnames(rois_chr)),
        roi_start = start(rois_chr),
        roi_end = end(rois_chr)
      )
      assigned_df <- left_join(roi_meta, assigned_df, by = "roi_idx")
    }
    
    assigned_list[[as.character(chr)]] <- assigned_df
  }
  
  # Combine chromosomes
  final_df <- bind_rows(assigned_list)
  
  # Create GRanges of assigned summits
  assigned_rows <- final_df %>% filter(!is.na(cons_idx))
  assigned_gr <- GRanges(
    seqnames = assigned_rows$roi_seqname,
    ranges = IRanges(start = assigned_rows$summit_pos, end = assigned_rows$summit_pos),
    Peakid = assigned_rows$Peakid
  )
  
  list(df = final_df, gr = assigned_gr)
}

# -------------------------
# Wrapper function
# -------------------------
# make_consensus_and_assign_fast <- function(summits_gr, ROIs_gr,
#                                            group_col = "group", score_col = "score",
#                                            min_gap_within = 100, min_gap_across = 400, max_dist = 500) {
#   highest_within <- get_highest_per_group_fast(summits_gr, group_col, score_col, min_gap_within)
#   consensus <- get_consensus_summits_fast(highest_within, score_col, min_gap_across)
#   assigned <- assign_summits_to_ROIs_fast(consensus, ROIs_gr, max_dist)
#   
#   list(
#     highest_within = highest_within,
#     consensus = consensus,
#     assigned_df = assigned$df,
#     assigned_gr = assigned$gr
#   )
# }

```
this function took hours to run.  I next worked with chatgpt to make a parallel version:

```{r parallel function}


# -------------------------
# Step 1: Reduce within groups (parallel, Windows-safe)
# -------------------------
get_highest_per_group_parallel <- function(summits_gr, group_col = "group",
                                           score_col = "score", min_gap_within = 100,
                                           workers = 2) {
  # Split GRanges by group to reduce memory per worker
  group_list <- split(summits_gr, mcols(summits_gr)[[group_col]])
  
  plan(multisession, workers = workers)  # Windows-compatible
  results <- future_lapply(group_list, function(gr_sub) {
    if (length(gr_sub) == 0) return(GRanges())
    
    red <- GenomicRanges::reduce(gr_sub, min.gapwidth = min_gap_within, ignore.strand = TRUE, with.revmap = TRUE)
    revmap <- mcols(red)$revmap
    
    idx <- unlist(lapply(revmap, function(x) {
      scores <- mcols(gr_sub)[[score_col]][x]
      x[which.max(scores)]
    }))
    
    gr_sub[idx]
  }, future.seed = TRUE)
  
  do.call(c, results)
}

# -------------------------
# Step 2: Reduce across groups (parallel, Windows-safe)
# -------------------------
get_consensus_summits_parallel <- function(highest_per_group_gr, score_col = "score",
                                           min_gap_across = 400, workers = 2) {
  if (length(highest_per_group_gr) == 0) return(GRanges())
  
  # Split by chromosome to reduce memory per worker
  chr_list <- split(highest_per_group_gr, seqnames(highest_per_group_gr))
  
  plan(multisession, workers = workers)
  results <- future_lapply(chr_list, function(gr_sub) {
    red <- GenomicRanges::reduce(gr_sub, min.gapwidth = min_gap_across, ignore.strand = TRUE, with.revmap = TRUE)
    revmap <- mcols(red)$revmap
    
    idx <- unlist(lapply(revmap, function(x) {
      scores <- mcols(gr_sub)[[score_col]][x]
      x[which.max(scores)]
    }))
    
    gr_sub[idx]
  }, future.seed = TRUE)
  
  do.call(c, results)
}

####### step 3 #####################3
assign_best_summit_to_ROI_parallel <- function(consensus_gr, ROIs_gr, max_dist = 500, workers = 2) {
  # Split ROIs by chromosome
  roi_list <- split(ROIs_gr, seqnames(ROIs_gr))
  cons_list <- split(consensus_gr, seqnames(consensus_gr))
  
  plan(multisession, workers = workers)
  
  results <- future_lapply(names(roi_list), function(chr) {
    rois_chr <- roi_list[[chr]]
    cons_chr <- cons_list[[chr]]
    nROIs <- length(rois_chr)
    
    if (nROIs == 0) return(tibble())
    
    # --- 1) Exact overlaps ---
    ov <- findOverlaps(rois_chr, cons_chr)
    assigned_df <- if(length(ov) > 0) {
      tibble(
        roi_idx = queryHits(ov),
        cons_idx = subjectHits(ov),
        summit_pos = start(cons_chr)[subjectHits(ov)],
        summit_score = mcols(cons_chr)$score[subjectHits(ov)]
      ) %>%
        group_by(roi_idx) %>%
        slice_max(summit_score, with_ties = FALSE) %>%
        ungroup()
    } else {
      tibble()
    }
    
    # --- 2) Nearest fallback for unassigned ---
    assigned_idx <- assigned_df$roi_idx %||% integer(0)
    roi_unassigned <- setdiff(seq_len(nROIs), assigned_idx)
    
    if(length(roi_unassigned) > 0 && length(cons_chr) > 0) {
      dn <- distanceToNearest(rois_chr[roi_unassigned], cons_chr)
      dn_df <- tibble(
        roi_idx = queryHits(dn),
        cons_idx = subjectHits(dn),
        distance = mcols(dn)$distance,
        summit_pos = start(cons_chr)[subjectHits(dn)],
        summit_score = mcols(cons_chr)$score[subjectHits(dn)]
      ) %>%
        filter(distance <= max_dist) %>%
        group_by(roi_idx) %>%
        slice_max(summit_score, with_ties = FALSE) %>%
        ungroup()
      
      assigned_df <- bind_rows(assigned_df, dn_df)
    }
    
    # --- 3) Attach ROI metadata ---
    roi_meta <- tibble(
      roi_idx = seq_len(nROIs),
      Peakid = rois_chr$Peakid,
      roi_seqname = as.character(seqnames(rois_chr)),
      roi_start = start(rois_chr),
      roi_end = end(rois_chr)
    )
    
    out_df <- left_join(roi_meta, assigned_df, by = "roi_idx") %>%
      mutate(
        dist_center = summit_pos - (roi_start + (roi_end - roi_start)/2),
        rel_pos = (summit_pos - roi_start)/(roi_end - roi_start)
      )
    
    out_df
  }, future.seed = TRUE)
  
  final_df <- bind_rows(results) %>% arrange(roi_idx)
  
  # --- 4) Create GRanges of assigned summits ---
  assigned_rows <- final_df %>% filter(!is.na(cons_idx))
  assigned_gr <- if(nrow(assigned_rows) > 0) {
    gr <- GRanges(
      seqnames = assigned_rows$roi_seqname,
      ranges = IRanges(start = assigned_rows$summit_pos, end = assigned_rows$summit_pos),
      Peakid = assigned_rows$Peakid
    )
    
    meta_cols <- setdiff(colnames(assigned_rows),
                         c("roi_idx","cons_idx","summit_pos","summit_score",
                           "Peakid","roi_seqname","roi_start","roi_end","dist_center","rel_pos","distance"))
    if(length(meta_cols) > 0) mcols(gr)[, meta_cols] <- assigned_rows[, meta_cols, drop = FALSE]
    gr
  } else {
    GRanges()
  }
  
  list(df = final_df, gr = assigned_gr)
}
```

 
```{r apply revised function}
# Allow up to 10 GB for tranferring large objects to parallel workers(adjust as needed)
options(future.globals.maxSize = 10 * 1024^3)
workers <- parallel::detectCores() - 1
### Step 1
temp_highest_per_group <- get_highest_per_group_parallel(all_H3K27ac_summits_gr,
                                                         group_col = "group",
                                                         score_col = "score",
                                                         min_gap_within = 100,
                                                         workers = workers)


# Concatenate into one GRanges (Step 1.2)
flat_highest_gr <- (c(temp_highest_per_group$DOX_144R, temp_highest_per_group$DOX_24R,temp_highest_per_group$DOX_24T,temp_highest_per_group$VEH_144R,temp_highest_per_group$VEH_24R,temp_highest_per_group$VEH_24T))

### STep 2
consensus_summits <- get_consensus_summits_parallel(
  flat_highest_gr,
  score_col = "score",
  min_gap_across = 400,
  workers = workers
)

####Concatenate into one GRanges
consensus_summits_gr <- (c(consensus_summits$chr1,consensus_summits$chr2,
                           consensus_summits$chr3,
                           consensus_summits$chr4,
                           consensus_summits$chr5,
                           consensus_summits$chr6,
                           consensus_summits$chr7,
                           consensus_summits$chr8,
                           consensus_summits$chr9,
                           consensus_summits$chr10,
                           consensus_summits$chr11,
                           consensus_summits$chr12,
                           consensus_summits$chr13,
                           consensus_summits$chr14,
                           consensus_summits$chr15,consensus_summits$chr16,
                           consensus_summits$chr17,consensus_summits$chr18,
                           consensus_summits$chr19,consensus_summits$chr20,
                           consensus_summits$chr21,consensus_summits$chr22))

final_data <- assign_best_summit_to_ROI_parallel(consensus_summits_gr, ROIs,max_dist = 500,workers = workers)
# data.frame with one row per ROI (assigned or NA)
final_df <- final_data$df

# GRanges of ROIs that received an assigned summit (with metadata)
assigned_summits_gr <- final_data$gr

# Quick checks
n_missing <- sum(is.na(final_df$summit_pos))
cat("ROIs lacking any summit within max_dist:", n_missing, "\n")



```

Almost done, now to look at the 13,015 ROIs without summits

```{r looking to assign missing}

na_rois <- final_df %>% filter(is.na(summit_pos)) 

na_rois_gr <- GRanges(
  seqnames = na_rois$roi_seqname,
  ranges = IRanges(start = na_rois$roi_start, end = na_rois$roi_end),
  Peakid = na_rois$Peakid
)
ov <- findOverlaps(na_rois_gr, all_H3K27ac_summits_gr)

overlap_df <- tibble(
  roi_idx = queryHits(ov),
  summit_idx = subjectHits(ov),
  summit_pos = start(all_H3K27ac_summits_gr)[subjectHits(ov)],
  summit_score = mcols(all_H3K27ac_summits_gr)$score[subjectHits(ov)]
)

best_summits <- overlap_df %>%
  group_by(roi_idx) %>%
  slice_max(summit_score, with_ties = FALSE) %>%
  ungroup()

assigned_df <- tibble(
  roi_idx = seq_along(na_rois_gr),
  Peakid = na_rois_gr$Peakid,
  roi_seqname = as.character(seqnames(na_rois_gr)),
  roi_start = start(na_rois_gr),
  roi_end = end(na_rois_gr)
) %>%
  left_join(best_summits, by = "roi_idx")

assigned_df_calc <- assigned_df %>% mutate(
        dist_center = summit_pos - (roi_start + (roi_end - roi_start)/2),
        rel_pos = (summit_pos - roi_start)/(roi_end - roi_start)
        
      )
complete_summit_df <- final_df %>% 
  dplyr::select(!distance) %>% 
   dplyr::filter(!is.na(summit_pos)) %>% 
  bind_rows(.,assigned_df_calc) %>% 
  left_join(., H3K27ac_lookup, by = "Peakid") %>% 
  dplyr::select(!summit_idx) %>% 
  dplyr::select(!cons_idx)

complete_summit_gr <- GRanges(
  seqnames=complete_summit_df$roi_seqname,
  ranges=IRanges(start= complete_summit_df$summit_pos, 
                 end= complete_summit_df$summit_pos)
)

# Columns to exclude from metadata (used to define GRanges)
exclude_cols <- c("roi_seqname", "summit_pos")

# Only keep columns that exist in df
meta_cols <- intersect(setdiff(colnames(complete_summit_df), exclude_cols), colnames(complete_summit_df))

# Assign metadata
mcols(complete_summit_gr) <- complete_summit_df[, meta_cols, drop = FALSE]

mcols(complete_summit_gr) 

```


Now to join the two data frames together as GRANGES and export to files

```{r fixing lost summits defunct, eval=FALSE}
assign_fallback_summits <- function(ROIs_gr, summits_gr, max_dist = 500) {
  # Find nearest summit for each ROI
  dn <- GenomicRanges::distanceToNearest(ROIs_gr, summits_gr)
  
  # Keep only those within max_dist
  dn <- dn[mcols(dn)$distance <= max_dist]
  
  if(length(dn) == 0) {
    return(list(df = tibble::tibble(), gr = GRanges()))
  }
  
  nearest_df <- tibble(
    roi_idx = queryHits(dn),
    cons_idx = subjectHits(dn),
    summit_pos = start(summits_gr)[subjectHits(dn)],
    summit_score = mcols(summits_gr)$score[subjectHits(dn)]
  )
  
  # If multiple nearest, pick highest score per ROI
  nearest_df <- nearest_df %>%
    group_by(roi_idx) %>%
    slice_max(summit_score, with_ties = FALSE) %>%
    ungroup()
  
  # Build full dataframe with ROI metadata
  out_df <- tibble(
    roi_idx = seq_along(ROIs_gr),
    Peakid = ROIs_gr$Peakid,
    roi_seqname = as.character(seqnames(ROIs_gr)),
    roi_start = start(ROIs_gr),
    roi_end = end(ROIs_gr)
  ) %>%
    left_join(nearest_df, by = "roi_idx") %>%
    mutate(
      dist_center = summit_pos - (roi_start + (roi_end - roi_start)/2),
      rel_pos = (summit_pos - roi_start)/(roi_end - roi_start)
    )
  
  # Build GRanges of assigned fallbacks (with Peakid)
  assigned_rows <- out_df %>% filter(!is.na(cons_idx))
  if(nrow(assigned_rows) > 0) {
    assigned_gr <- GRanges(
      seqnames = assigned_rows$roi_seqname,
      ranges = IRanges(start = assigned_rows$summit_pos, end = assigned_rows$summit_pos)
    )
    mcols(assigned_gr)$Peakid <- assigned_rows$Peakid
    
    # attach any additional summit metadata
    meta_cols <- setdiff(colnames(assigned_rows),
                         c("roi_idx","cons_idx","summit_pos","summit_score",
                           "Peakid","roi_seqname","roi_start","roi_end","dist_center","rel_pos"))
    if(length(meta_cols) > 0) mcols(assigned_gr)[, meta_cols] <- assigned_rows[, meta_cols, drop = FALSE]
  } else {
    assigned_gr <- GRanges()
  }
  
  list(df = out_df, gr = assigned_gr)
}


```


```{r redoing cluster membership and export files, eval=FALSE}


### adding in cluster membership for export

SET_1_gr <-   complete_summit_gr %>%
  as.data.frame() %>% 
  dplyr::filter(cluster=="Set_1") %>% 
  dplyr::select(seqnames, start, end,Peakid) %>% 
  na.omit() %>% 
  GRanges()


rtracklayer::export(SET_1_gr, "data/Bed_exports/H3K27ac_Set_1_summits.bed")


SET_2_gr <-   complete_summit_gr %>%
  as.data.frame() %>% 
  dplyr::filter(cluster=="Set_2") %>% 
  dplyr::select(seqnames, start, end,Peakid) %>% 
  na.omit() %>% 
  GRanges()



rtracklayer::export(SET_2_gr, "data/Bed_exports/H3K27ac_Set_2_summits.bed")

SET_3_gr <-   complete_summit_gr %>%
  as.data.frame() %>% 
  dplyr::filter(cluster=="Set_3") %>% 
  dplyr::select(seqnames, start, end,Peakid) %>% 
  na.omit() %>% 
  GRanges()



rtracklayer::export(SET_3_gr, "data/Bed_exports/H3K27ac_Set_3_summits.bed")
  

rtracklayer::export(complete_summit_gr, "data/Bed_exports/H3K27ac_complete_final_summits.bed")
# rtracklayer::export(all_H3K27ac_summits_gr,"data/Bed_exports/All_summits_bedfile.bed")

# rtracklayer::export(ROIs,"data/Bed_exports/All_ROIs.bed")

```

