---
title: "Summit Files"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```


```{r packages}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(BiocParallel)

```

```{r calling summit files}
sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")
##Path to histone summit files
H3K27ac_dir <- "C:/Users/renee/Other_projects_data/DXR_data/final_data/summit_files/H3K27ac" 

##pull all histone files together
H3K27ac_summit_files <- list.files(
  path = H3K27ac_dir,
  pattern = "\\.bed$",
  recursive = TRUE,
  full.names = TRUE
)
length(H3K27ac_summit_files)

# head(H3K27ac_summit_files)
```
First steps:  pulling in the information of Sets and locations
Pulling in the summit files,
concatenation of all summits into one large file

```{r pulling in other data}
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")
H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})

read_summit <- function(file){
  peaks <- read.table(file,header = FALSE)
  colnames(peaks) <- c("chr","start","end","name","score")
  GRanges(
    seqnames = peaks$chr,
    ranges = IRanges(start=peaks$start, end = peaks$start),
    score=peaks$score,
    file=basename(file),
    Library_ID = stringr::str_remove(basename(file), "_FINAL_summits\\.bed$")
  )
}

all_H3K27ac_summits_list<- lapply(H3K27ac_summit_files, read_summit)
all_H3K27ac_summits_gr <- do.call(c, all_H3K27ac_summits_list)  # combine into one GRanges object
H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
```

```{r pick summit per roi}
pick_best_summit_per_roi <- function(roi_gr, summit_gr, score_col = "signalValue") {

  hits <- findOverlaps(roi_gr, summit_gr)

  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    score = mcols(summit_gr)[[score_col]][subjectHits(hits)]
  )

  # Pick summit with highest score for each ROI
  best <- df %>%
    group_by(roi_idx) %>%
    slice_max(order_by = score, n = 1) %>%
    ungroup()

  # Build the final GRanges object
  best_summits <- summit_gr[best$summit_idx]
  best_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]

  best_summits
}
# 
# best_summits <- pick_best_summit_per_roi(
#   ROIs, highest_summits_long_gr, score_col = "peakHeight"
# )
```

```{r pick summit closest to roi center}
pick_center_summit <- function(roi_gr, summit_gr) {

  hits <- findOverlaps(roi_gr, summit_gr)

  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    dist_to_center = abs(
      start(summit_gr)[subjectHits(hits)] -
      round((start(roi_gr)[queryHits(hits)] + end(roi_gr)[queryHits(hits)]) / 2)
    )
  )

  best <- df %>%
    group_by(roi_idx) %>%
    slice_min(order_by = dist_to_center, n = 1) %>%
    ungroup()

  best_summits <- summit_gr[best$summit_idx]
  best_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]

  best_summits
}
```

```{r consensus summit}
pick_consensus_summit <- function(roi_gr, summit_gr, score_col = "score") {
  
  # Find which summits overlap each ROI
  hits <- findOverlaps(roi_gr, summit_gr)
  
  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    summit_pos = start(summit_gr)[subjectHits(hits)],
    score = mcols(summit_gr)[[score_col]]
  )
  
  # Count frequency per summit position within each ROI
  freq_df <- df %>%
    group_by(roi_idx, summit_pos) %>%
    summarise(
      freq = n(),
      max_score = max(score),
      .groups = "drop"
    )
  
  # Pick the summit with highest frequency; break ties with max score
  best <- freq_df %>%
    group_by(roi_idx) %>%
    slice_max(order_by = freq, n = 1, with_ties = TRUE) %>%
    slice_max(order_by = max_score, n = 1) %>%
    ungroup()
  
  # Map back to original GRanges
  best_idx <- df$summit_idx[match(
    paste0(best$roi_idx, "_", best$summit_pos),
    paste0(df$roi_idx, "_", df$summit_pos)
  )]
  
  final_summits <- summit_gr[best_idx]
  final_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]
  
  final_summits
}
```

### Analysis of all summit files, no grouping  


first creating summit clusters that are around 100bps across all summit files
```{r pulling in other data again}
 
#  Merge summits within 100 bp clusters but keep metadata
# ------------------------
# 1 Reduce with revmap to keep track of original indices

clusters <- GenomicRanges::reduce(all_H3K27ac_summits_gr, min.gapwidth = 100,
                   ignore.strand = TRUE, with.revmap = TRUE)

# 2 For each cluster, pick the highest-score summit
scores <- mcols(all_H3K27ac_summits_gr)$score

revmap <- clusters$revmap

# Compute the highest-score index per cluster
highest_idx <- sapply(revmap, function(idx) idx[which.max(scores[idx])])

# Subset GRanges once
highest_per_cluster_gr <- all_H3K27ac_summits_gr[highest_idx]


# ------------------------
# 3 Count merged summits per ROI
# ------------------------
ROIs <- H3K27ac_sets_gr$all_H3K27ac  # your ROI GRanges

# Optimized counting
roi_counts <- countOverlaps(ROIs, highest_per_cluster_gr)
mcols(ROIs)$merged_summit_count <- roi_counts

# ------------------------
# 5 Map ROIs to clusters with sample info
# ------------------------
overlaps <- findOverlaps(ROIs, highest_per_cluster_gr)

roi_hits_df <- as.data.frame(overlaps) %>%
  mutate(
    Peakid = ROIs$Peakid[queryHits],
    Library_ID = mcols(highest_per_cluster_gr)$Library_ID[subjectHits],
    score = mcols(highest_per_cluster_gr)$score[subjectHits],
    file = mcols(highest_per_cluster_gr)$file[subjectHits]
  ) %>%
  left_join(sampleinfo, by = c("Library_ID" = "Library ID"))

## initualize count column on ROI df
mcols(ROIs)$summit_count <- 0

# Tabulate counts
counts <- table(queryHits(overlaps))
mcols(ROIs)$summit_count[as.numeric(names(counts))] <- as.numeric(counts)

# Convert to dataframe for plotting
roi_counts_df <- as.data.frame(ROIs) %>%
  select(Peakid, seqnames, start, end, summit_count) %>% 
  mutate(roi_size=end-start)

# Optional: add Set / cluster info
roi_counts_df <- roi_counts_df %>%
  left_join(H3K27ac_lookup, by = "Peakid")
# ------------------------
# 6 Optional: add Set/cluster info per ROI
# ------------------------
roi_hits_df <- roi_hits_df %>%
  left_join(H3K27ac_lookup, by = "Peakid")

# Now roi_hits_df is ready for plotting or analysis:
# Columns include Peakid, ROI coordinates, Library_ID, score, file, Individual, Treatment, Timepoint, cluster


# Optional: add sample info
roi_hits_df <- as.data.frame(overlaps) %>%
  mutate(
    Peakid = ROIs$Peakid[queryHits],
    Library_ID = mcols(all_H3K27ac_summits_gr)$Library_ID[subjectHits],
    score = mcols(all_H3K27ac_summits_gr)$score[subjectHits],
    file = mcols(all_H3K27ac_summits_gr)$file[subjectHits]
  ) %>%
  left_join(sampleinfo, by = c("Library_ID" = "Library ID"))

```

Plotting some summits from the first process

```{r get summits function plots}
ROIs %>% 
  as.data.frame() %>% 
  # dplyr::filter(merged_summit_count==1) %>% 
  left_join(H3K27ac_lookup, by = "Peakid") %>%  # make sure join key matches
  filter(!is.na(cluster)) %>%
  ggplot(aes(x = cluster, y = merged_summit_count)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6, size = 2, color = "steelblue") +
  theme_bw() +
  xlab("Set") +
  ylab("Number of merged summits per ROI") +
  ggtitle("Distribution of merged summit counts by Set")


roi_counts_df %>%
  ggplot(aes(x = summit_count)) +
  geom_histogram(binwidth = 1, fill="steelblue", color="black") +
  theme_bw() +
  xlab("Number of summits per ROI") +
  ylab("Number of ROIs")

roi_counts_df %>%
  ggplot(aes(x = summit_count)) +
  geom_histogram(binwidth = 1, fill="steelblue", color="black") +
  theme_bw() +
  xlab("Number of summits per ROI") +
  ylab("Number of ROIs")+
  ggtitle("Zoomed in summit per ROI")+
  coord_cartesian(xlim=c(0,50))

roi_counts_df %>% 
  # group_by(Peakid) %>% tally #%>% 
  # left_join(H3K27ac_lookup) %>% 
  ggplot(aes(x = roi_size, y = summit_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "ROI size (bp)",
    y = "# of summits",
    title = "Relationship between ROI size and # of Summit clusters"
  ) +
  theme_classic()#+
  # coord_cartesian(xlim=c(-1,400), ylim=c(0,10))

roi_counts_df %>% 
  # group_by(Peakid) %>% tally #%>% 
  # left_join(H3K27ac_lookup) %>% 
  ggplot(aes(x = roi_size, y = summit_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "ROI size (bp)",
    y = "# of summits",
    title = "Zoomed in Relationship between ROI size and # of Summit clusters"
  ) +
  theme_classic()+
  coord_cartesian(xlim=c(-1,800), ylim=c(0,10))

```


### Merging strategies!

This is now the place I will take my files and now try to apply the same merging strategy as I did the peaks merging strategy.  Step 1: reduce all trt-time summits within 100 bp
```{r reduction by trt-time}
###Adding in sampleinfo dataframe
meta <- as.data.frame(mcols(all_H3K27ac_summits_gr))
meta2 <- meta %>% 
  left_join(., sampleinfo, by=c("Library_ID"="Library ID"))

mcols(all_H3K27ac_summits_gr) <- meta2

mcols(all_H3K27ac_summits_gr)$group <- 
  paste(all_H3K27ac_summits_gr$Treatment,
        all_H3K27ac_summits_gr$Timepoint,
        sep = "_")

###now splitting into grouped granges 

gr_by_group <- split(all_H3K27ac_summits_gr,
                                    all_H3K27ac_summits_gr$group)
# gr_by_group <- as(gr_by_group, "CompressedGRangesList")
### now reducing within some width by 100 bp with revmap
groups <- unique(all_H3K27ac_summits_gr$group)

reduced_groups <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 100, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_200 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 200, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_300 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 300, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_400 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 400, ignore.strand = TRUE, with.revmap = TRUE)
})

names(reduced_groups) <- groups
names(reduced_groups_200) <- groups
names(reduced_groups_300) <- groups
names(reduced_groups_400) <- groups

# redux_main_100 <- sum(sapply(reduced_groups, length))
# redux_main_200 <- sum(sapply(reduced_groups_200, length))
# redux_main_300 <- sum(sapply(reduced_groups_300, length))
# redux_main_400 <- sum(sapply(reduced_groups_400, length))

```

Plotting effect of reduction bp number on total number of clusters
```{r}
gap_sizes <- seq(0, 500, by = 10)


# Function to apply reduce for each gap and return counts
results <- lapply(gap_sizes, function(g) {
  reduced <- GenomicRanges::reduce(all_H3K27ac_summits_gr, min.gapwidth = g)
  data.frame(gap = g, n_regions = length(reduced))
})

# Combine results
min_gap_summary <- bind_rows(results)

ggplot(min_gap_summary, aes(x = gap, y= n_regions))+
  geom_line(linewidth=2)+
  geom_point()+
  theme_bw(base_size=14)+
  labs(
    title = "Effect of Gap Size on Reduced Summits",
    x = "Gap size (bp)",
    y = "Number of merged regions"
  )

```
```{r get highes per group funciton}
# Function to pick highest summit per cluster from a reduced GRanges list
get_highest_per_group <- function(reduced_groups, orig_summits_gr, group_col = "group") {
  
  groups <- names(reduced_groups)
  highest_per_group <- vector("list", length(reduced_groups))
  names(highest_per_group) <- groups
  
  for(i in seq_along(reduced_groups)) {
    gr <- reduced_groups[[i]]
    
    # original summits for this group
    orig <- orig_summits_gr[mcols(orig_summits_gr)[[group_col]] == groups[i]]
    scores <- orig$score
    
    # revmap is a CompressedIntegerList
    revmap <- mcols(gr)$revmap
    
    # skip if revmap is NULL
    if(is.null(revmap)) next
    
    # unlist all indices once
    all_idx <- unlist(revmap, use.names = FALSE)
    
    # repeat cluster index for each element in revmap
    cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
    
    # scores for all indices
    all_scores <- scores[all_idx]
    
    # For each cluster, pick the index of the max score
    max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
      ii[which.max(all_scores[ii])]
    })
    
    # Convert back to original indices
    orig_idx <- all_idx[unlist(max_idx_per_cluster)]
    
    # subset original GRanges
    highest_per_group[[i]] <- orig[orig_idx]
  }
  
  # Flatten any nested GRangesList
  flatten_gr <- function(x) {
    if (inherits(x, "GRanges")) return(x)
    if (inherits(x, "GRangesList")) return(unlist(x, use.names = FALSE))
    if (is.list(x)) return(do.call(c, lapply(x, flatten_gr)))
    stop("Unexpected object type")
  }
  
  highest_summits_gr <- flatten_gr(highest_per_group)
  
  # Return as long GRanges with group column
  highest_summits_df <- bind_rows(
    lapply(names(highest_per_group), function(gr_name) {
      as.data.frame(highest_per_group[[gr_name]]) %>%
        mutate(group = gr_name)
    })
  )
  
  highest_summits_long_gr <- highest_summits_df %>% GRanges()
  
  return(highest_summits_long_gr)
}
```
```{r}
reduced_sets <- list(
  "100bp" = reduced_groups,
  "200bp" = reduced_groups_200,
  "300bp" = reduced_groups_300,
  "400bp" = reduced_groups_400
)

highest_summits_all <- lapply(reduced_sets, get_highest_per_group, orig_summits_gr = all_H3K27ac_summits_gr)
```
  
  Looking at number of summits across groups as a function of min.gap number

```{r across groups mingap}
 ##Compute counts per group for each reduced set
summit_counts_group <- lapply(names(highest_summits_all), function(gap_name) {
  gr <- highest_summits_all[[gap_name]]
  
  # make sure group column exists
  if(!"group" %in% colnames(mcols(gr))) stop("GRanges must have 'group' column")
  
  df <- as.data.frame(gr) %>%
    count(group, name = "n_summits") %>%
    mutate(gap = gap_name)
  
  return(df)
}) %>% bind_rows()


ggplot(summit_counts_group, aes(x = gap, y = n_summits, group = group, color = group)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  theme_classic(base_size = 14) +
  labs(
    title = "Effect of min-gap width on number of highest summits per group",
    x = "Min-gap width (bp)",
    y = "Number of highest summits",
    color = "Group"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
### Summit average per ROI
What is the average number of summits per ROI as a function of min.gapwidth

```{r double reduce function}
double_reduce_summits <- function(summits_gr, ROIs, groups,
                                  min_gap_within_seq = c(100,200,300),
                                  min_gap_across_seq = c(100,200,300),
                                  BPPARAM = MulticoreParam(4)) {
  
  # Pre-split summits by group to avoid repeated subsetting
  summits_by_group <- split(summits_gr, summits_gr$group)
  
  # Create all gap combinations
  gap_combos <- expand.grid(min_gap_within = min_gap_within_seq,
                            min_gap_across = min_gap_across_seq,
                            stringsAsFactors = FALSE)
  
  # Apply in parallel for each combination
  results <- bplapply(seq_len(nrow(gap_combos)), function(i) {
    g1 <- gap_combos$min_gap_within[i]
    g2 <- gap_combos$min_gap_across[i]
    
    # -----------------
    # Step 1: Reduce within group
    # -----------------
    highest_per_group <- lapply(groups, function(grp) {
      gr_sub <- summits_by_group[[grp]]
      if(length(gr_sub) == 0) return(NULL)
      
      # reduce within group with revmap
      red <- GenomicRanges::reduce(gr_sub, min.gapwidth = g1, ignore.strand = TRUE, with.revmap = TRUE)
      revmap <- mcols(red)$revmap
      if(length(revmap) == 0) return(NULL)
      
      # pick highest score per cluster
      cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
      all_idx <- unlist(revmap, use.names = FALSE)
      all_scores <- gr_sub$score[all_idx]
      max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
        ii[which.max(all_scores[ii])]
      })
      gr_sub[all_idx[unlist(max_idx_per_cluster)]]
    })
    
    highest_per_group <- highest_per_group[!sapply(highest_per_group, is.null)]
    
    # -----------------
    # Step 2: Merge across groups
    # -----------------
    if(length(highest_per_group) == 0) return(NULL)
    all_highest <- do.call(c, highest_per_group)
    consensus <- GenomicRanges::reduce(all_highest, min.gapwidth = g2, ignore.strand = TRUE)
    
    # -----------------
    # Step 3: Count summits per ROI (vectorized)
    # -----------------
    hits <- findOverlaps(ROIs, consensus)
    counts <- as.data.frame(table(queryHits(hits)))
    colnames(counts) <- c("ROI_idx", "n_summits")
    counts$ROI_idx <- as.integer(as.character(counts$ROI_idx))
    counts$min_gap_within <- g1
    counts$min_gap_across <- g2
    counts
  }, BPPARAM = BPPARAM)
  
  # Combine results
  results_df <- bind_rows(results)
  
  return(results_df)
}
```

```{r calculating of double reduce}

BPPARAM <- SnowParam(workers = 4, type = "SOCK")  
register(BPPARAM)

heatmap_data <- double_reduce_summits(
  summits_gr = all_H3K27ac_summits_gr,
  ROIs = H3K27ac_sets_gr$all_H3K27ac,
  groups = unique(all_H3K27ac_summits_gr$group),
  min_gap_within_seq = c(100,200,300,400),
  min_gap_across_seq = c(100,200,300,400),
  BPPARAM = BPPARAM
)

heatmap_avg <- heatmap_data %>%
  group_by(min_gap_within, min_gap_across) %>%
  summarise(mean_summits = mean(n_summits, na.rm = TRUE), .groups = "drop")

```

```{r plot mingap averages}
ggplot(heatmap_avg, aes(x = factor(min_gap_within),
                        y = factor(min_gap_across),
                        fill = mean_summits)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(mean_summits, 1)), color = "black", size = 4) +
  scale_fill_viridis_c(option = "plasma") +
  labs(
    x = "Within-group min-gap (bp)",
    y = "Across-group min-gap (bp)",
    fill = "Mean # summits per ROI",
    title = "Effect of min-gap width on summits per ROI"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

after i figure my final min.gaps, here is where the places are
```{r}

# --- Step 1: Reduce within groups and pick highest summit per cluster ---
get_highest_per_group <- function(summits_gr, groups = NULL, min_gap_within = 100) {
  
  if(is.null(groups)) {
    groups <- unique(summits_gr$group)
  }
  
  highest_per_group <- lapply(groups, function(grp) {
    gr_sub <- summits_gr[summits_gr$group == grp]
    if(length(gr_sub) == 0) return(NULL)
    
    # Reduce within group
    reduced <- GenomicRanges::reduce(gr_sub, min.gapwidth = min_gap_within, ignore.strand = TRUE, with.revmap = TRUE)
    
    # Pick highest scoring summit per cluster
    revmap <- reduced$revmap
    scores <- gr_sub$score
    
    all_idx <- unlist(revmap, use.names = FALSE)
    cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
    all_scores <- scores[all_idx]
    
    max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
      ii[which.max(all_scores[ii])]
    })
    
    orig_idx <- all_idx[unlist(max_idx_per_cluster)]
    
    gr_sub[orig_idx]
  })
  
  # Flatten list of GRanges
  highest_per_group <- do.call(c, highest_per_group)
  return(highest_per_group)
}

# --- Step 2: Reduce across groups to get final consensus ---
get_consensus_summits <- function(highest_per_group_gr, min_gap_across = 400) {
  
  # Reduce across groups
  reduced <- GenomicRanges::reduce(highest_per_group_gr, min.gapwidth = min_gap_across, ignore.strand = TRUE, with.revmap = TRUE)
  
  # Pick highest scoring summit per cluster across all groups
  revmap <- reduced$revmap
  scores <- highest_per_group_gr$score
  
  all_idx <- unlist(revmap, use.names = FALSE)
  cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
  all_scores <- scores[all_idx]
  
  max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
    ii[which.max(all_scores[ii])]
  })
  
  orig_idx <- all_idx[unlist(max_idx_per_cluster)]
  
  highest_per_group_gr[orig_idx]
}

add_ROI_to_consensus <- function(consensus_gr, ROIs_gr) {
  
  hits <- findOverlaps(consensus_gr, ROIs_gr)
  
  df <- data.frame(
    summit_chr  = seqnames(consensus_gr)[queryHits(hits)],
    summit_pos  = start(consensus_gr)[queryHits(hits)],
    summit_score = consensus_gr$score[queryHits(hits)],
    Peakid      = ROIs_gr$Peakid[subjectHits(hits)],
    roi_start   = start(ROIs_gr)[subjectHits(hits)],
    roi_end     = end(ROIs_gr)[subjectHits(hits)]
  )
  
  df %>%
    mutate(rel_pos = (summit_pos - roi_start)/(roi_end - roi_start),
           roi_center = roi_start + (roi_end - roi_start)/2,
           dist_center = summit_pos - roi_center)
}


```
```{r looking at 100-400}
highest_100 <- get_highest_per_group(all_H3K27ac_summits_gr, min_gap_within = 100)

final_consensus <- get_consensus_summits(highest_100, min_gap_across = 400)

final_df_100_400 <- add_ROI_to_consensus(final_consensus, H3K27ac_sets_gr$all_H3K27ac)


ggplot(final_df_100_400, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```


```{r distance from center plot}

final_df_100_400 %>% 
  group_by(Peakid) %>% 
  slice_max(summit_score) %>% 
ggplot(., aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  geom_vline(xintercept = 250, linetype = "dashed",color="yellow") +  # left 500 
  geom_vline(xintercept = -250, linetype = "dashed",color="yellow") +  # right 500
  
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
 coord_cartesian(xlim=c(-2000,2000))

final_df_100_400 %>% 
  group_by(Peakid) %>% 
  slice_max(summit_score) %>% 
ggplot(., aes(x = dist_center)) +
  geom_density()+
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  geom_vline(xintercept = 250, linetype = "dashed",color="blue") +  # left 500 
  geom_vline(xintercept = -250, linetype = "dashed",color="blue") +  # right 500
 coord_cartesian(xlim=c(-2000,2000))

```
```{r looking at 200-400}
highest_200 <- get_highest_per_group(all_H3K27ac_summits_gr, min_gap_within = 200)

final_consensus2 <- get_consensus_summits(highest_200, min_gap_across = 400)

final_df_200_400 <- add_ROI_to_consensus(final_consensus2, H3K27ac_sets_gr$all_H3K27ac)


ggplot(final_df_100_400, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color="green") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+coord_cartesian(xlim=c(-2000,2000))
```

#### function get highest scoring summit clustered by group:

```{r}
#' Get highest summit per reduced cluster per group using plyranges
#'
#' @param gr GRanges with metadata columns: group, score
#' @param min_gap numeric, minimum gap width to merge summits
#' @param ignore.strand logical, whether to ignore strand when reducing
#'
#' @return GRanges of highest scoring summit per reduced cluster per group
get_highest_summits_with_cluster_id <- function(gr, min_gap = 100, ignore.strand = TRUE) {
  
  gr %>%
    # Reduce summits per group
    group_by(group) %>%
    reduce_ranges(min_gap = min_gap, ignore.strand = ignore.strand, with_revmap = TRUE) %>%
    # Add a cluster_id based on row_number()
    mutate(cluster_id = paste0(group, "_cluster_", row_number())) %>%
    # join back original summits
    join_overlap_inner(gr, suffix = c(".cluster", ".orig")) %>%
    # slice max per cluster_id
    group_by(cluster_id) %>%
    slice_max(score.orig, with_ties = FALSE) %>%
    ungroup() %>%
    # select only the original summit coordinates
    select(seqnames, start = start.orig, end = end.orig, score = score.orig,
           group, cluster_id)
}
```


#### function reduce_and_pick_highest

```{r}

#### this quickly reduces any granges summit  collection by the gap that is passed in the function. It returns the highest scoring summit from the cluster of summits.
reduce_and_pick_highest <- function(gr, gap = 100) {

  # Preextract score once
  sc <- gr$score

  # Fast reduce with revmap
  red <- GenomicRanges::reduce(
    gr,
    min.gapwidth = gap,
    ignore.strand = TRUE,
    with.revmap = TRUE
  )

  # This is vectorized and very fast
  idx_list <- red$revmap

  # Pick highest-scoring summit within each cluster
  sel <- IntegerList(
    lapply(idx_list, function(idx) idx[which.max(sc[idx])])
  )

  # Convert IntegerList → integer vector
  best_idx <- unlist(sel, use.names = FALSE)

  # Subset original GRanges
  gr[best_idx]
}

```

```{r}
redux_100 <- reduce_and_pick_highest(all_H3K27ac_summits_gr, gap = 100)
redux_200 <- reduce_and_pick_highest(all_H3K27ac_summits_gr, gap = 200)
redux_300 <- reduce_and_pick_highest(all_H3K27ac_summits_gr, gap = 300)

redux_Peak_100 <- add_ROI_to_consensus(redux_100, H3K27ac_sets_gr$all_H3K27ac)
redux_Peak_200 <- add_ROI_to_consensus(redux_200, H3K27ac_sets_gr$all_H3K27ac)
redux_Peak_300 <- add_ROI_to_consensus(redux_300, H3K27ac_sets_gr$all_H3K27ac)
```


```{r plotting alternates}
ggplot(redux_Peak_100, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("Only single reduction 100bp")

ggplot(redux_Peak_200, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("Only single reduction 200bp")

ggplot(redux_Peak_300, aes(x = dist_center, y = Peakid)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  geom_vline(xintercept = 500, linetype = "dashed",color="red") +  # left 500 
  geom_vline(xintercept = -500, linetype = "dashed",color="red") +  # right 500
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  ggtitle("Only single reduction 300bp")

```

