---
title: "Summit Files"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```


```{r packages}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)

```

```{r calling summit files}
sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")
##Path to histone summit files
H3K27ac_dir <- "C:/Users/renee/Other_projects_data/DXR_data/final_data/summit_files/H3K27ac" 

##pull all histone files together
H3K27ac_summit_files <- list.files(
  path = H3K27ac_dir,
  pattern = "\\.bed$",
  recursive = TRUE,
  full.names = TRUE
)
length(H3K27ac_summit_files)

# head(H3K27ac_summit_files)
```
First steps:  pulling in the information of Sets and locations
Pulling in the summit files,
concatenation of all summits into one large file

```{r pulling in other data}
peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")
H3K27ac_sets_gr <- lapply(peakAnnoList_H3K27ac, function(df) {
  as_granges(df)
})

read_summit <- function(file){
  peaks <- read.table(file,header = FALSE)
  colnames(peaks) <- c("chr","start","end","name","score")
  GRanges(
    seqnames = peaks$chr,
    ranges = IRanges(start=peaks$start, end = peaks$start),
    score=peaks$score,
    file=basename(file),
    Library_ID = stringr::str_remove(basename(file), "_FINAL_summits\\.bed$")
  )
}

all_H3K27ac_summits_list<- lapply(H3K27ac_summit_files, read_summit)
all_H3K27ac_summits_gr <- do.call(c, all_H3K27ac_summits_list)  # combine into one GRanges object
H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y)
)
```

```{r pick summit per roi}
pick_best_summit_per_roi <- function(roi_gr, summit_gr, score_col = "signalValue") {

  hits <- findOverlaps(roi_gr, summit_gr)

  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    score = mcols(summit_gr)[[score_col]][subjectHits(hits)]
  )

  # Pick summit with highest score for each ROI
  best <- df %>%
    group_by(roi_idx) %>%
    slice_max(order_by = score, n = 1) %>%
    ungroup()

  # Build the final GRanges object
  best_summits <- summit_gr[best$summit_idx]
  best_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]

  best_summits
}
# 
# best_summits <- pick_best_summit_per_roi(
#   ROIs, highest_summits_long_gr, score_col = "peakHeight"
# )
```

```{r pick summit closest to roi center}
pick_center_summit <- function(roi_gr, summit_gr) {

  hits <- findOverlaps(roi_gr, summit_gr)

  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    dist_to_center = abs(
      start(summit_gr)[subjectHits(hits)] -
      round((start(roi_gr)[queryHits(hits)] + end(roi_gr)[queryHits(hits)]) / 2)
    )
  )

  best <- df %>%
    group_by(roi_idx) %>%
    slice_min(order_by = dist_to_center, n = 1) %>%
    ungroup()

  best_summits <- summit_gr[best$summit_idx]
  best_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]

  best_summits
}
```

```{r consensus summit}
pick_consensus_summit <- function(roi_gr, summit_gr, score_col = "score") {
  
  # Find which summits overlap each ROI
  hits <- findOverlaps(roi_gr, summit_gr)
  
  df <- data.frame(
    roi_idx = queryHits(hits),
    summit_idx = subjectHits(hits),
    summit_pos = start(summit_gr)[subjectHits(hits)],
    score = mcols(summit_gr)[[score_col]]
  )
  
  # Count frequency per summit position within each ROI
  freq_df <- df %>%
    group_by(roi_idx, summit_pos) %>%
    summarise(
      freq = n(),
      max_score = max(score),
      .groups = "drop"
    )
  
  # Pick the summit with highest frequency; break ties with max score
  best <- freq_df %>%
    group_by(roi_idx) %>%
    slice_max(order_by = freq, n = 1, with_ties = TRUE) %>%
    slice_max(order_by = max_score, n = 1) %>%
    ungroup()
  
  # Map back to original GRanges
  best_idx <- df$summit_idx[match(
    paste0(best$roi_idx, "_", best$summit_pos),
    paste0(df$roi_idx, "_", df$summit_pos)
  )]
  
  final_summits <- summit_gr[best_idx]
  final_summits$ROI_name <- roi_gr$Peakid[best$roi_idx]
  
  final_summits
}
```

first creating summit clusters that are around 100bps across all summits
```{r pulling in other data}
 
#  Merge summits within 100 bp clusters but keep metadata
# ------------------------
# Reduce with revmap to keep track of original indices

clusters <- GenomicRanges::reduce(all_H3K27ac_summits_gr, min.gapwidth = 100,
                   ignore.strand = TRUE, with.revmap = TRUE)

# For each cluster, pick the highest-score summit
scores <- mcols(all_H3K27ac_summits_gr)$score

revmap <- clusters$revmap

# Compute the highest-score index per cluster
highest_idx <- sapply(revmap, function(idx) idx[which.max(scores[idx])])

# Subset GRanges once
highest_per_cluster_gr <- all_H3K27ac_summits_gr[highest_idx]


# ------------------------
# 3️⃣ Count merged summits per ROI
# ------------------------
ROIs <- H3K27ac_sets_gr$all_H3K27ac  # your ROI GRanges

# Optimized counting
roi_counts <- countOverlaps(ROIs, highest_per_cluster_gr)
mcols(ROIs)$merged_summit_count <- roi_counts

# ------------------------
# 5️⃣ Map ROIs to clusters with sample info
# ------------------------
overlaps <- findOverlaps(ROIs, highest_per_cluster_gr)

roi_hits_df <- as.data.frame(overlaps) %>%
  mutate(
    Peakid = ROIs$Peakid[queryHits],
    Library_ID = mcols(highest_per_cluster_gr)$Library_ID[subjectHits],
    score = mcols(highest_per_cluster_gr)$score[subjectHits],
    file = mcols(highest_per_cluster_gr)$file[subjectHits]
  ) %>%
  left_join(sampleinfo, by = c("Library_ID" = "Library ID"))

## initualize count column on ROI df
mcols(ROIs)$summit_count <- 0

# Tabulate counts
counts <- table(queryHits(overlaps))
mcols(ROIs)$summit_count[as.numeric(names(counts))] <- as.numeric(counts)

# Convert to dataframe for plotting
roi_counts_df <- as.data.frame(ROIs) %>%
  select(Peakid, seqnames, start, end, summit_count) %>% 
  mutate(roi_size=end-start)

# Optional: add Set / cluster info
roi_counts_df <- roi_counts_df %>%
  left_join(H3K27ac_lookup, by = "Peakid")
# ------------------------
# 6️⃣ Optional: add Set/cluster info per ROI
# ------------------------
roi_hits_df <- roi_hits_df %>%
  left_join(H3K27ac_lookup, by = "Peakid")

# Now roi_hits_df is ready for plotting or analysis:
# Columns include Peakid, ROI coordinates, Library_ID, score, file, Individual, Treatment, Timepoint, cluster


# Optional: add sample info
roi_hits_df <- as.data.frame(overlaps) %>%
  mutate(
    Peakid = ROIs$Peakid[queryHits],
    Library_ID = mcols(all_H3K27ac_summits_gr)$Library_ID[subjectHits],
    score = mcols(all_H3K27ac_summits_gr)$score[subjectHits],
    file = mcols(all_H3K27ac_summits_gr)$file[subjectHits]
  ) %>%
  left_join(sampleinfo, by = c("Library_ID" = "Library ID"))

```

Plotting some summits from the first process

```{r get summits function plots}
ROIs %>% 
  as.data.frame() %>% 
  # dplyr::filter(merged_summit_count==1) %>% 
  left_join(H3K27ac_lookup, by = "Peakid") %>%  # make sure join key matches
  filter(!is.na(cluster)) %>%
  ggplot(aes(x = cluster, y = merged_summit_count)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.6, size = 2, color = "steelblue") +
  theme_bw() +
  xlab("Set") +
  ylab("Number of merged summits per ROI") +
  ggtitle("Distribution of merged summit counts by Set")


roi_counts_df %>%
  ggplot(aes(x = summit_count)) +
  geom_histogram(binwidth = 1, fill="steelblue", color="black") +
  theme_bw() +
  xlab("Number of summits per ROI") +
  ylab("Number of ROIs")

roi_counts_df %>%
  ggplot(aes(x = summit_count)) +
  geom_histogram(binwidth = 1, fill="steelblue", color="black") +
  theme_bw() +
  xlab("Number of summits per ROI") +
  ylab("Number of ROIs")+
  ggtitle("Zoomed in summit per ROI")+
  coord_cartesian(xlim=c(0,50))

roi_counts_df %>% 
  # group_by(Peakid) %>% tally #%>% 
  # left_join(H3K27ac_lookup) %>% 
  ggplot(aes(x = roi_size, y = summit_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "ROI size (bp)",
    y = "# of summits",
    title = "Relationship between ROI size and # of Summit clusters"
  ) +
  theme_classic()#+
  # coord_cartesian(xlim=c(-1,400), ylim=c(0,10))

roi_counts_df %>% 
  # group_by(Peakid) %>% tally #%>% 
  # left_join(H3K27ac_lookup) %>% 
  ggplot(aes(x = roi_size, y = summit_count)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "ROI size (bp)",
    y = "# of summits",
    title = "Zoomed in Relationship between ROI size and # of Summit clusters"
  ) +
  theme_classic()+
  coord_cartesian(xlim=c(-1,800), ylim=c(0,10))

```


```{r, eval=FALSE}
ROIs_tor <- H3K27ac_sets_gr$all_H3K27ac

hits <- as.data.frame(findOverlaps(ROIs_tor, all_H3K27ac_summits_gr))

df <- hits %>%
  mutate(
    Peakid      = ROIs_tor$Peakid[queryHits],
    summit_start = start(all_H3K27ac_summits_gr)[subjectHits],
    summit_end   = end(all_H3K27ac_summits_gr)[subjectHits],
    roi_start    = start(ROIs_tor)[queryHits],
    roi_end      = end(ROIs_tor)[queryHits],
    roi_width    = roi_end - roi_start
  )

df <- df %>%
  mutate(
    rel_pos = (summit_start - roi_start) / roi_width
  )
```

```{r takes too long,eval=FALSE}
df <- df %>%
  left_join(H3K27ac_lookup, by = "Peakid")   # adds cluster column

roi_order <- df %>%
  count(Peakid, name = "summit_count") %>%
  arrange(desc(summit_count))

df <- df %>%
  mutate(roi_factor = factor(Peakid, levels = roi_order$Peakid))

roi_df <- as.data.frame(ROIs_tor) %>%
  dplyr::select(Peakid, seqnames, start, end, width)

df <- df %>% mutate(
  rel_centered = rel_pos - 0.5
)


ggplot(df, aes(
  x = rel_pos,
  y = roi_factor,
  color = cluster
)) +
  geom_point(alpha = 0.6, size = 1.5) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_x_continuous(
    name = "Relative summit position (0 = ROI start, 1 = ROI end)",
    limits = c(0, 1)
  ) +
  ylab("ROIs (ordered by number of summits)") +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )



```
### Merging strategies!

This is now the place I will take my files and now try to apply the same merging strategy as I did the peaks merging strategy.  Step 1: reduce all trt-time summits within 100 bp
```{r reduction by trt-time}

meta <- as.data.frame(mcols(all_H3K27ac_summits_gr))
meta2 <- meta %>% 
  left_join(., sampleinfo, by=c("Library_ID"="Library ID"))

mcols(all_H3K27ac_summits_gr) <- meta2

mcols(all_H3K27ac_summits_gr)$group <- 
  paste(all_H3K27ac_summits_gr$Treatment,
        all_H3K27ac_summits_gr$Timepoint,
        sep = "_")

###now splitting into grouped granges 

gr_by_group <- split(all_H3K27ac_summits_gr,
                                    all_H3K27ac_summits_gr$group)
# gr_by_group <- as(gr_by_group, "CompressedGRangesList")
### now reducing within some width by 100 bp with revmap
groups <- unique(all_H3K27ac_summits_gr$group)

reduced_groups <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 100, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_200 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 200, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_300 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 300, ignore.strand = TRUE, with.revmap = TRUE)
})

reduced_groups_400 <- lapply(groups, function(g) {
  gr_sub <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == g]
  GenomicRanges::reduce(gr_sub, min.gapwidth = 400, ignore.strand = TRUE, with.revmap = TRUE)
})

names(reduced_groups) <- groups
names(reduced_groups_200) <- groups
names(reduced_groups_300) <- groups
names(reduced_groups_400) <- groups

# redux_main_100 <- sum(sapply(reduced_groups, length))
# redux_main_200 <- sum(sapply(reduced_groups_200, length))
# redux_main_300 <- sum(sapply(reduced_groups_300, length))
# redux_main_400 <- sum(sapply(reduced_groups_400, length))

```

```{r highest scoring summit per cluster 100}
highest_per_group <- vector("list", length(reduced_groups))
names(highest_per_group) <- groups

for(i in seq_along(reduced_groups)) {
  gr <- reduced_groups[[i]]
  
  # original summits for this group
  orig <- all_H3K27ac_summits_gr[all_H3K27ac_summits_gr$group == groups[i]]
  scores <- orig$score
  
  # revmap is a CompressedIntegerList
  revmap <- gr$revmap
  
  # unlist all indices once
  all_idx <- unlist(revmap, use.names = FALSE)
  
  # repeat cluster index for each element in revmap
  cluster_idx <- rep(seq_along(revmap), times = elementNROWS(revmap))
  
  # scores for all indices
  all_scores <- scores[all_idx]
  
  # For each cluster, pick the index of the max score
  max_idx_per_cluster <- tapply(seq_along(all_scores), cluster_idx, function(ii) {
    ii[which.max(all_scores[ii])]
  })
  
  # Convert back to original indices
  orig_idx <- all_idx[unlist(max_idx_per_cluster)]
  
  # subset original GRanges
  highest_per_group[[i]] <- orig[orig_idx]
}

# Recursive function to flatten any nested lists/GRangesList
flatten_gr <- function(x) {
  if (inherits(x, "GRanges")) {
    return(x)
  } else if (inherits(x, "GRangesList")) {
    return(unlist(x, use.names = FALSE))
  } else if (is.list(x)) {
    # recursively flatten list elements
    return(do.call(c, lapply(x, flatten_gr)))
  } else {
    stop("Unexpected object type")
  }
}

# Apply to your list of highest per group
highest_summits_gr <- flatten_gr(highest_per_group)


highest_summits_df <- bind_rows(
  lapply(names(highest_summits_gr), function(gr_name) {
    as.data.frame(highest_summits_gr[[gr_name]]) %>%
      mutate(group = gr_name)
  })
)
highest_summits_long_gr <- highest_summits_df %>% GRanges()


```

```{r up to tornado plots?}

# ROIs
ROIs <- H3K27ac_sets_gr$all_H3K27ac

# Overlaps
hits_group <- findOverlaps(ROIs, highest_summits_long_gr)

df_group <- data.frame(
  Peakid    = ROIs$Peakid[queryHits(hits_group)],
  summit_pos = start(highest_summits_long_gr)[subjectHits(hits_group)],
  roi_start  = start(ROIs)[queryHits(hits_group)],
  roi_end    = end(ROIs)[queryHits(hits_group)]
)

# Relative position within ROI (0=start, 1=end)
df_group <- df_group %>%
  mutate(rel_pos = (summit_pos - roi_start) / (roi_end - roi_start))


roi_order_group <- df_group %>%
  count(Peakid, name = "summit_count") %>%
  arrange(desc(summit_count))


df_group <- df_group %>%
  mutate(roi_factor = factor(Peakid, levels = roi_order_group$Peakid)) %>% 
  left_join(., H3K27ac_lookup, by = "Peakid") %>% 
  mutate(roi_center = roi_start + (roi_end - roi_start) / 2  ) %>% 
  mutate(dist_center = summit_pos - roi_center)


# df_group %>% 
# ggplot(., aes(x = rel_pos, y = roi_factor, color = cluster)) +
#   geom_point(size = 0.5, alpha = 0.7) +   # each summit
#   scale_x_continuous(limits = c(0,1)) +
#   labs(x = "Relative position within ROI", y = "ROIs", color = "Set/Cluster") +
#   theme_minimal() +
#   theme(
#     axis.text.y = element_blank(),   # hide ROI labels if too many
#     axis.ticks.y = element_blank()
#   )
```

```{r rel pos}
# df_group %>%
#   dplyr::filter(cluster=="Set_3") %>% 
# ggplot(., aes(x = rel_pos, y = roi_factor, color = cluster)) +
#   geom_point(size = 0.5, alpha = 0.7) +   # each summit
#   scale_x_continuous(limits = c(0,1)) +
#   labs(x = "Relative position within ROI", y = "ROIs", color = "Set/Cluster") +
#   theme_minimal() +
#   theme(
#     axis.text.y = element_blank(),   # hide ROI labels if too many
#     axis.ticks.y = element_blank()
#   )
```

```{r distance from center plot}
ggplot(df_group, aes(x = dist_center, y = roi_factor, color=cluster)) +
  ggrastr::geom_point_rast(size = 0.6, alpha = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +  # ROI center
  labs(
    x = "Distance from ROI center (bp)",
    y = "ROI",
    color = "Group"
  ) +
  # theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )+
  coord_cartesian(xlim=c(-2000,2000))
```
#### function get highest scoring summit clustered by group:

```{r}
#' Get highest summit per reduced cluster per group using plyranges
#'
#' @param gr GRanges with metadata columns: group, score
#' @param min_gap numeric, minimum gap width to merge summits
#' @param ignore.strand logical, whether to ignore strand when reducing
#'
#' @return GRanges of highest scoring summit per reduced cluster per group
get_highest_summits_with_cluster_id <- function(gr, min_gap = 100, ignore.strand = TRUE) {
  
  gr %>%
    # Reduce summits per group
    group_by(group) %>%
    reduce_ranges(min_gap = min_gap, ignore.strand = ignore.strand, with_revmap = TRUE) %>%
    # Add a cluster_id based on row_number()
    mutate(cluster_id = paste0(group, "_cluster_", row_number())) %>%
    # join back original summits
    join_overlap_inner(gr, suffix = c(".cluster", ".orig")) %>%
    # slice max per cluster_id
    group_by(cluster_id) %>%
    slice_max(score.orig, with_ties = FALSE) %>%
    ungroup() %>%
    # select only the original summit coordinates
    select(seqnames, start = start.orig, end = end.orig, score = score.orig,
           group, cluster_id)
}
```


```{r}
redux_main_100 <- get_highest_summits_plyranges(all_H3K27ac_summits_gr,min_gap = 100)
redux_main_200 <- get_highest_summits_plyranges(all_H3K27ac_summits_gr,min_gap = 200)
redux_main_300 <- get_highest_summits_plyranges(all_H3K27ac_summits_gr,min_gap = 300)
redux_main_400 <- get_highest_summits_plyranges(all_H3K27ac_summits_gr,min_gap = 400)

```


#### function reduce_and_pick_highest

```{r}

#### this quickly reduces any granges summit  collection by the gap that is passed in the function. It returns the highest scoring summit from the cluster of summits.
reduce_and_pick_highest <- function(gr, gap = 100) {

  # Preextract score once
  sc <- gr$score

  # Fast reduce with revmap
  red <- GenomicRanges::reduce(
    gr,
    min.gapwidth = gap,
    ignore.strand = TRUE,
    with.revmap = TRUE
  )

  # This is vectorized and very fast
  idx_list <- red$revmap

  # Pick highest-scoring summit within each cluster
  sel <- IntegerList(
    lapply(idx_list, function(idx) idx[which.max(sc[idx])])
  )

  # Convert IntegerList → integer vector
  best_idx <- unlist(sel, use.names = FALSE)

  # Subset original GRanges
  gr[best_idx]
}

```

```{r}

redux_100 <- reduce_and_pick_highest(highest_summits_long_gr, gap = 100)
redux_200 <- reduce_and_pick_highest(highest_summits_long_gr, gap = 200)
redux_300 <- reduce_and_pick_highest(highest_summits_long_gr, gap = 300)

redux_100 %>% as.data.frame()

```

```{r}
pick_best_summit_per_roi(ROIs,redux_100,score_col = "score")
```

