---
title: "Supplemental Figure draft"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```


```{r library entries}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(ComplexHeatmap)
library(circlize)
library(ggalluvial)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(edgeR)
library(RColorBrewer)
library(ggVennDiagram)
library(ChIPseeker)
```


Data import:
```{r data import}
sampleinfo <- read_delim("data/sample_info.tsv", delim = "\t")
peak_ct <- read_delim("data/peaks/peaks_cts_FINAL.txt", delim = "\t")
H3K27ac_peaks <- read_delim("data/peaks/H3K27ac_FINAl_results.tsv",delim = "\t")

H3K27me3_peaks <- read_delim("data/peaks/H3K27me3_final_results.tsv",delim = "\t")

H3K9me3_peaks <- read_delim("data/peaks/H3K9me3_final_results.tsv",delim = "\t")

H3K36me3_peaks <- read_delim("data/peaks/H3K36me3_final_results.tsv",delim = "\t")

trimmed_multiqc_fastqc <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/Steven_bed_Files/data/multiqc_data_trim/multiqc_fastqc.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
fastq_fragments <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/fastq_backup_stats.tsv",delim = "\t", escape_double = FALSE,
    trim_ws = TRUE)
frag_count_total_no_multi <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/Steven_bed_Files/data/bam_no_multi/frag_count_total_no_multi.tsv",
    delim = "\t", escape_double = FALSE,
    trim_ws = TRUE)

fragment_count_totals_final <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/Steven_bed_Files/data/bam_final/fragment_count_totals_final.tsv",
    delim = "\t", escape_double = FALSE,
    trim_ws = TRUE)

# creating the fragmenfile

H3K27acfrag_summary_file <- fastq_fragments %>% 
  mutate(file=str_remove(file, "_R[12]\\.fastq\\.gz$")) %>% 
  distinct() %>% 
  left_join(sampleinfo,by=c("file"="Library ID")) %>% 
  dplyr::filter(Treatment != "5FU" & Histone_Mark =="H3K27ac") %>% 
  dplyr::select(file, Histone_Mark:Timepoint,num_seqs) %>% 
  left_join(frag_count_total_no_multi, by=c("file"="sample")) %>% 
  left_join(fragment_count_totals_final, by=c("file"="sample")) %>% 
  dplyr::rename("all_fragments"=num_seqs,
                "mapped_fragments"= fragment_count_total.x,
                "unique_noM_dedup"=fragment_count_total.y)


H3K27ac_toplist <- readRDS( "data/DER_data/H3K27ac_toplist.RDS")

peakAnnoList_H3K27ac <- readRDS("data/motif_lists/H3K27ac_annotated_peaks.RDS")

H3K27ac_lookup <- imap_dfr(peakAnnoList_H3K27ac[1:3], ~
  tibble(Peakid = .x@anno$Peakid, cluster = .y))
```

```{r echo=TRUE, file='code/corMotifcustom.R'}

```

```{r needed functions}

volcanosig <- function(df, psig.lvl) {
    df <- df %>% 
    mutate(threshold = case_when(
        adj.P.Val > psig.lvl ~ "A",
        adj.P.Val <= psig.lvl & logFC <= 0 ~ "B",
        adj.P.Val <= psig.lvl & logFC > 0 ~ "C"))
   
      ## count significant red and blue points
  red_n <- sum(df$threshold == "B", na.rm = TRUE)
  blue_n <- sum(df$threshold == "C", na.rm = TRUE)
    
  ggplot(df, aes(x=logFC, y=-log10(P.Value))) + 
    ggrastr::geom_point_rast(aes(color=threshold))+
    annotate("text",x = -5.5, y = 24, label = paste0("n = ", red_n), color = "red", hjust = 0,size = 4) +
    annotate("text", x = 5.5, y = 24, label = paste0("n = ", blue_n), color = "blue", hjust = 1,size = 4) +
    xlab(expression("Log"[2]*" FC"))+
    ylab(expression("-log"[10]*"P Value"))+
    scale_color_manual(values = c("black", "red","blue"))+
    theme_classic()+
    ylim(0,25)+
    xlim(-6,6)+
    theme(legend.position = "none",
              plot.title = element_text(size = rel(1.5), hjust = 0.5),
              axis.title = element_text(size = rel(0.8))) 
}

standardizeAnnoList <- function(annoList, categories, color_pal = NULL) {
  if (is.null(color_pal)) {
    if (!requireNamespace("RColorBrewer", quietly = TRUE)) {
      stop("Please install RColorBrewer to use the default palette")
    }
    color_pal <- RColorBrewer::brewer.pal(length(categories), "Set3")
    names(color_pal) <- categories
  }

  for (nm in names(annoList)) {
    annoStat <- annoList[[nm]]@annoStat
    
    # pad missing categories
    missing_cats <- setdiff(categories, as.character(annoStat$Feature))
    if (length(missing_cats) > 0) {
      pad <- data.frame(Feature = missing_cats, Frequency = 0)
      annoStat <- rbind(annoStat, pad)
    }
    
    # factor + reorder rows
    annoStat$Feature <- factor(annoStat$Feature, levels = categories)
    annoStat <- annoStat[order(match(annoStat$Feature, categories)), ]
    
    annoList[[nm]]@annoStat <- annoStat
  }
  
  return(list(
    annoList = annoList,
    color_pal = color_pal
  ))
}



```
  

Basics of supplemental data in the context of H3K27ac data first.

### Current Fig 1  

```{r SF1, fig.width=8}
H3K27acfrag_summary_file %>% 
  pivot_longer(cols=c(all_fragments:unique_noM_dedup), names_to = "type", values_to = "frag_counts") %>% 
  mutate(Timepoint=factor(Timepoint, levels = c("24T","24R","144R"),labels = c("0R","24R","144R")),
         Treatment=factor(Treatment, levels = c("VEH","DOX"))) %>% 
  ggplot(., aes(x=type, y=frag_counts))+
  geom_boxplot(aes(fill=Treatment))+
  geom_point(aes(col=Individual, size=2))+
   theme_bw()+
  facet_wrap(~Treatment+Timepoint,nrow = 3, ncol = 6 )+
 scale_fill_brewer(palette="Set2")+
  scale_color_brewer(palette = "Dark2")+
  theme(strip.text = element_text(face = "bold",  hjust = 0, size = 8),
        strip.background = element_rect(fill = "white", linetype = "solid",
                                        color = "black", linewidth = 1),
        panel.spacing = unit(1, 'points'),
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1))+
  coord_cartesian(ylim=c(0,19000000))

```


### Current Fig 2
Fragment lengths
```{r SF2, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

file_list_filter <- list.files(path="data/bam_no_multi",
           pattern ="frag_len_count\\.txt$",full.names = TRUE)
read_and_label <- function(file) {
  df <- read_delim(file, delim = "\t", col_names = c("Col1", "Col2"))  # Adjust delimiter if needed
  df <- df %>%
    mutate(File = basename(file),  # Add filename column
  weight = Col2/sum(Col2))
  return(df)
}
combined_df <- map_df(file_list_filter, read_and_label)

annotated_combo_df <- combined_df %>% 
  mutate(sample = gsub("_frag_len_count.txt","",File)) %>% 
  left_join(., sampleinfo, by = c("sample"="Library ID"))
annotated_combo_df <- annotated_combo_df[(!annotated_combo_df$Treatment %in% "5FU"),]

annotated_combo_df %>% 
  mutate(Timepoint=factor(Timepoint, levels = c("24T","24R","144R"),labels = c("0R","24R","144R")),
         Treatment=factor(Treatment, levels = c("VEH","DOX"))) %>% 
  mutate(sample_name=paste0(Individual,"_",Treatment,"_",Timepoint),
         trt_time=paste0(Treatment,"_",Timepoint),
         trt_time=factor(trt_time, levels = c("VEH_0R","VEH_24R","VEH_144R","DOX_0R","DOX_24R","DOX_144R"))) %>% 
   arrange( Treatment,Individual, Timepoint) %>%
  mutate(sample_name = factor(sample_name, levels = unique(sample_name))) %>%
  dplyr::filter(Histone_Mark=="H3K27ac") %>%
  ggplot(., aes(x=sample_name, y=Col1, weight = weight,fill = Treatment))+
geom_violin(bw = 5) +
    scale_y_continuous(breaks = seq(0, 800, 50)) +
   theme_bw(base_size = 15) +
    ggpubr::rotate_x_text(angle = 90) +
  ggtitle("Fragment lengths for H3K27ac")+
    ylab("Fragment Length") +
    xlab("")+
  scale_fill_brewer(palette="Set2")

```

### Current Fig 3

```{r SF3}

```

### Current Fig 4
peak counts
```{r SF4 peak counts}

Peak_n_FRIP <- H3K27ac_peaks %>% 
  left_join(sampleinfo, by =c("Sample"='Library ID')) %>% 
  mutate(Timepoint=factor(Timepoint, levels = c("24T","24R","144R"),labels = c("0R","24R","144R")),
         Treatment=factor(Treatment, levels = c("VEH","DOX"))) %>% 
  left_join(peak_ct, by=c("Sample"="Sample")) %>% 
  mutate(sample_name=paste0(Individual,"_",Treatment,"_",Timepoint),
         trt_time=paste0(Treatment,"_",Timepoint),
         trt_time=factor(trt_time, levels = c("VEH_0R","VEH_24R","VEH_144R","DOX_0R","DOX_24R","DOX_144R")))

ggplot(Peak_n_FRIP, aes(x=trt_time, y = Count))+
  geom_boxplot(aes(fill=Treatment, alpha=.8))+
  geom_point(aes(color=Individual, size=4))+
  labs(title="Count of H3K27ac Peaks")+
  ylab("Number of Peaks")+
  xlab("Treatment and Time")+
  theme_bw()+
  scale_color_brewer(palette="Dark2")+
  scale_fill_brewer(palette="Set2")+
  coord_cartesian(ylim = c(0,3e5))
  

```

### Current Fig 5  

FRiP
```{r SF5 Frip}
ggplot(Peak_n_FRIP, aes(x=trt_time, y = FRiP))+
  geom_boxplot(aes(fill=Treatment, alpha=.8))+
  geom_point(aes(color=Individual, size=4))+
  labs(title="Fraction of Reads in Peaks for H3K27ac")+
  ylab("Reads in Peaks/ Total Reads")+
  xlab("Treatment and Time")+
  theme_bw()+
  scale_color_brewer(palette="Dark2")+
  scale_fill_brewer(palette="Set2")+
  coord_cartesian(ylim=c(0,1))

```

### Current Fig 6

```{r}
H3K27ac_merged <- read_delim("data/peaks/H3K27ac_FINAL_counts.txt",
    delim = "\t", escape_double = FALSE,
    trim_ws = TRUE, skip = 1)
rename_list <- sampleinfo %>% 
  mutate(stem= "_nobl.bam") %>% 
  mutate(prefix=paste0("/scratch/10819/styu/MW_multiQC/peaks/",Histone_Mark,"/",Treatment,"/",Timepoint,"/")) %>%
  mutate(oldname=paste0(prefix,`Library ID`,"/",`Library ID`,stem)) %>% 
  mutate(newname=paste0(Individual,"_",Treatment,"_",Timepoint)) %>% 
  dplyr::select(oldname,newname)
rename_vec <- setNames(rename_list$newname, rename_list$oldname)
names(H3K27ac_merged)[names(H3K27ac_merged) %in% names(rename_vec)] <- rename_vec[names(H3K27ac_merged)[names(H3K27ac_merged) %in% names(rename_vec)]]


H3K27ac_merged_raw <- H3K27ac_merged %>%
  dplyr::select(Geneid,contains("Ind")) %>%
  column_to_rownames("Geneid") %>%
  as.matrix()
H3K27ac_merged_lcpm <- H3K27ac_merged %>%
  dplyr::select(Geneid,contains("Ind")) %>%
  column_to_rownames("Geneid") %>%
  cpm(., log = TRUE)

H3K27ac_merged_raw <- H3K27ac_merged_raw[rowMeans(H3K27ac_merged_lcpm)>0,]
H3K27ac_merged_raw <- H3K27ac_merged_raw[!grepl("chrY",rownames(H3K27ac_merged_raw)),]
H3K27ac_merged_raw <- H3K27ac_merged_raw[!grepl("chrX",rownames(H3K27ac_merged_raw)),]

H3K27ac_annomat <- data.frame(timeset=colnames(H3K27ac_merged_raw)) %>%
  mutate(sample=timeset) %>%
  separate(timeset, into = c("ind","tx","time")) %>%
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(ind = gsub("Ind", "", ind)) %>%
  mutate(txtime = paste0(tx, "_", time)) %>%
  mutate(group = txtime)
H3K27ac_annomat$group <- H3K27ac_annomat$group %>%
  gsub("DOX_24T", "1", .) %>%
  gsub("DOX_24R", "2", .) %>%
  gsub("DOX_144R", "3", .) %>%
  gsub("VEH_24T", "4", .) %>%
  gsub("VEH_24R", "5", .) %>%
  gsub("VEH_144R", "6", .)

dge_H3K27ac <- edgeR::DGEList(counts = H3K27ac_merged_raw, group = H3K27ac_annomat$group, genes = row.names(H3K27ac_merged_raw))

lcpm_dge_H3K27ac <- cpm(dge_H3K27ac, log = TRUE)
```


```{r}

get_regr_pval <- function(mod) {
  # Returns the p-value for the Fstatistic of a linear model
  # mod: class lm
  stopifnot(class(mod) == "lm")
  fstat <- summary(mod)$fstatistic
  pval <- 1 - pf(fstat[1], fstat[2], fstat[3])
  return(pval)
}



plot_versus_pc <- function(df, pc_num, fac) {
  # df: data.frame
  # pc_num: numeric, specific PC for plotting
  # fac: column name of df for plotting against PC
  pc_char <- paste0("PC", pc_num)
  # Calculate F-statistic p-value for linear model
  pval <- get_regr_pval(lm(df[[ pc_char]] ~ df[[ fac]]))
  # Pick palette based on factor
  pal <- dplyr::case_when(
    fac == "ind" ~ "Dark2",
    fac == "tx"  ~ "Set2",
    TRUE ~ NA_character_)
  
  
  if (is.numeric(df[[fac]])) {
    p <- ggplot(df, aes(x=.data[[fac]], y= .data[[pc_char]]))+
     geom_point() +
      geom_smooth(method = "lm") + 
      theme_bw()+
      labs(title = sprintf("p-val: %.2f", pval))
  } else {
     p <- ggplot(df, aes(x = .data[[fac]], y = .data[[pc_char]], fill = .data[[fac]])) +
      geom_boxplot() +
       theme_bw()+
      labs(title = sprintf("p-val: %.3f", pval))
     if (!is.na(pal)) {
      p <- p + scale_fill_brewer(palette = pal)
    }
  }
  
  p
}
```

```{r contributes}

H3K27ac_pr <- prcomp(t(lcpm_dge_H3K27ac)) 

H3K27ac_pr_anno <- H3K27ac_pr$x %>% 
  as.data.frame() %>% 
  rownames_to_column("sample") %>% 
  separate(sample, into = c("ind","tx","time")) %>%
  mutate(tx=factor(tx, levels = c("VEH", "DOX")),
         time=factor(time, levels =c("24T","24R","144R"))) %>%
  mutate(txtime = paste0(tx, "_", time)) %>% 
  mutate(sex=if_else(ind %in% c("Ind1","Ind2"),"F","M"))%>% 
  mutate(sex=factor(sex, levels=c("F","M")))


facs <- c("ind", "tx", "time", "sex")
names(facs) <- c("Individual", "Treatment", "Time","Sex")

plots_H3K27ac <- lapply(names(facs), function(f_label) {
  f <- facs[f_label]
  f_v_pc1 <- plot_versus_pc(H3K27ac_pr_anno, 1, f)
  f_v_pc2 <- plot_versus_pc(H3K27ac_pr_anno, 2, f)
  f_v_pc3 <- plot_versus_pc(H3K27ac_pr_anno, 3, f)
  f_v_pc4 <- plot_versus_pc(H3K27ac_pr_anno, 4, f)

  gridExtra::grid.arrange(f_v_pc1, f_v_pc2, f_v_pc3,f_v_pc4, ncol = 2, top = paste0(f_label," H3K27ac"))
})
```


### Current Fig 7
heatmap of log2cpm

```{r heatmap, fig.width=10, fig.height=10}
H3K27ac_merged_cor <- H3K27ac_merged_lcpm %>% 
  cor()

annomat <- data.frame(sample=colnames(H3K27ac_merged_cor)) %>% 
  separate_wider_delim(sample,delim="_",names=c("Ind","Treatment","Timepoint"),cols_remove = FALSE) %>% 
  mutate(Treatment=factor(Treatment, levels = c("VEH","DOX")),
         Timepoint=factor(Timepoint, levels =c("24T","24R","144R")),
         Ind=factor(Ind, levels=c("Ind1","Ind2","Ind3","Ind4","Ind5"))) %>% 
  column_to_rownames("sample")

dark2_cols <- brewer.pal(8, "Dark2")
set2_cols  <- brewer.pal(8, "Set2")

ind_cols <- setNames(
  dark2_cols[seq_along(levels(annomat$Ind))],
  levels(annomat$Ind))

treatment_cols <- setNames(
  set2_cols[c(1, 2)],
  levels(annomat$Treatment))

timepoint_cols <- setNames(
  set2_cols[c(4, 5, 7)],
  levels(annomat$Timepoint))

heatmap_first <- ComplexHeatmap::HeatmapAnnotation(
   df = annomat,
  col = list(
    Ind = ind_cols,
    Treatment = treatment_cols,
    Timepoint = timepoint_cols))

        
Heatmap(
  H3K27ac_merged_cor, 
  top_annotation = heatmap_first,
  column_title = "Filtered log2cpm H3K27ac"
)
```
 
### Current Fig 8

volcano plots, venn, and proportions 

A:
```{r fig8a}
H3K27ac_toptable_list <- bind_rows(H3K27ac_toplist, .id = "group")




H3K27ac_0R <- volcanosig(H3K27ac_toplist$H3K27ac_24T, 0.05)+ ggtitle("DOX 0R")
H3K27ac_24R <- volcanosig(H3K27ac_toplist$H3K27ac_24R, 0.05)+ ggtitle("DOX 24R")+ylab("")
H3K27ac_144R <- volcanosig(H3K27ac_toplist$H3K27ac_144R, 0.05)+ ggtitle("DOX 144R")+ylab("")
cowplot::plot_grid(H3K27ac_0R, H3K27ac_24R, H3K27ac_144R, rel_widths =c(1,1,1))

```

venn diagram
```{r fig8b}
DOX_0R <- H3K27ac_toplist$H3K27ac_24T
DOX_0R <- DOX_0R %>% dplyr::filter(adj.P.Val<0.05)

DOX_24R <- H3K27ac_toplist$H3K27ac_24R
DOX_24R <- DOX_24R %>% dplyr::filter(adj.P.Val<0.05)

DOX_144R <- H3K27ac_toplist$H3K27ac_144R
DOX_144R <- DOX_144R %>% dplyr::filter(adj.P.Val<0.05)

ggVennDiagram(list("DOX_0R"=DOX_0R$genes,
                   "DOX_24R"=DOX_24R$genes,
                   "DOX_144R"=DOX_144R$genes),
              label="count",
              set_color = c("DOX_0R"=timepoint_cols["24T"],
                            "DOX_24R"  =timepoint_cols["24R"],
                            "DOX_144R" = timepoint_cols["144R"]))



```



```{r fig8c}
H3K27ac_toptable_list %>% 
  dplyr::select(group,genes,logFC,adj.P.Val) %>% 
  mutate(sig_val=if_else(adj.P.Val<0.05,"sig","not_sig")) %>% 
  mutate(
    sig_val = factor(sig_val, levels = c("not_sig", "sig")),
    group = factor(group, levels = c("H3K27ac_24T","H3K27ac_24R","H3K27ac_144R"))) %>%
  ggplot(aes(x = group, fill = sig_val)) +
  geom_bar(position = "fill") +
  theme_bw() +
  ggtitle("H3K27ac") +
  ylab("Proportion") +
  scale_y_continuous(labels = scales::percent)


```

### Current Fig 9

```{r Fig9AB}
cormotif_H3K27ac <- readRDS("data/Cormotif_data/Cormotif_initial_H3K27ac.RDS")

plotIC(cormotif_H3K27ac)
plotMotif(cormotif_H3K27ac)

grey_vals <- seq(from = 1, to = 0, by = -0.1)

legend_df <- data.frame(
  motif.q = factor(seq(0, 1, by = 0.1)),
  x = 1,
  y = seq(0, 1, by = 0.1)
)

p <- ggplot(legend_df, aes(x = x, y = y, fill = motif.q)) +
  geom_tile() +
  scale_fill_manual(values = gray(grey_vals),
                    name = "Probability") +
  theme_void() +
  theme(legend.position = "right")

legend_only <- cowplot::get_legend(p)

cowplot::plot_grid(legend_only)

```

proportion plot
```{r}
all_H3K27ac_regions <- peakAnnoList_H3K27ac$all_H3K27ac %>% as.data.frame() %>% dplyr::select(Peakid)

all_H3K27ac_regions %>%  
  mutate(group="H3K27ac") %>% 
  left_join(., H3K27ac_lookup) %>% 
   tidyr::replace_na(list(cluster = "NA")) %>% 
  ggplot(., aes(y=group,fill=cluster))+
  geom_bar()+
  geom_text(aes( label = scales::percent(..prop..),
                   x= ..prop.. ), stat= "count", vjust = -.5)+
  ggtitle("H3K27ac proportions of ROIs")+
  theme_classic()


```

```{r}
examp_H3K27ac_1 <-
    c("chr1:77923740-77924468","chr20:25484622-25485357","chr1:229148017-229149226")
examp_H3K27ac_2 <-
  c("chr5:85451856-85453893","chr10:114271262-114272616", "chr3:63343167-63344341" )
examp_H3K27ac_3 <- 
  c("chr8:8664034-8666371","chr3:141152796-141154087","chr21:43701994-43704465")


```


```{r}
plot_example_lcpm <- function(examples, set){
  
  df <- lcpm_dge_H3K27ac %>% 
    as.data.frame() %>% 
    dplyr::filter(row.names(lcpm_dge_H3K27ac) %in% examples) %>%
    rownames_to_column("Peakid") %>% 
    pivot_longer(., cols = !Peakid, names_to = "sample", values_to = "lcpm" ) %>% 
    separate_wider_delim(., cols=sample, delim="_", names=c("ind","tx","time")) %>% 
  mutate(ind=factor(ind, levels=c("Ind1", "Ind2", "Ind3", "Ind4","Ind5")),
         tx=factor(tx,levels = c("VEH","DOX")),
         time=factor(time, levels=c("24T","24R","144R"), labels=c("0R","24R","144R")),
         group_graph= paste0(tx, "_", time),
         group=paste0("H3K27ac_",time),
      group_graph = factor(group_graph, levels = c(        "VEH_0R", "VEH_24R","VEH_144R", "DOX_0R","DOX_24R","DOX_144R")))
  
  geneplot <- ggplot(df, aes(x=group_graph, y=lcpm)) +
  geom_boxplot(aes(fill=tx))+
  facet_wrap(~Peakid, scales="free")+
  theme_bw()+
  theme(axis.text.x=element_text(angle = 90))+
  ggtitle(paste0(set," H3K27ac samples"))+
  scale_fill_brewer(palette="Set2")

  
  plot(geneplot)
}
```

```{r fig9exmp plots}
plot_example_lcpm(examp_H3K27ac_1,"Set 1")

plot_example_lcpm(examp_H3K27ac_2,"Set 2")

plot_example_lcpm(examp_H3K27ac_3,"Set 3")
```


### Current Fig 10  



```{r Fig10A}

categories <- c(
  "Distal Intergenic", "Downstream (<=300)", 
  "Other Intron", "1st Intron", 
  "Other Exon", "1st Exon", 
  "3' UTR", "5' UTR", 
  "Promoter (1-2kb)", "Promoter (<=1kb)"
)

# Optional: custom color palette
color_pal <- c(
  "#6a3d9a", "#cab2d6",
  "#ff7f00", "#fdbf6f",
  "#e31a1c", "#fb9a99",
  "#33a02c", "#b2df8a",
  "#1f78b4", "#a6cee3"
)
names(color_pal) <- categories

res <- standardizeAnnoList(peakAnnoList_H3K27ac, categories, color_pal)
peakAnnoList_H3K27ac_std <- res$annoList
color_pal_std <- res$color_pal

# bar plot
plotAnnoBar(peakAnnoList_H3K27ac_std[c(4,1,2,3)]) +
  scale_fill_manual(values=color_pal_std) +
  ggtitle("H3K27ac Feature Distribution")
```

```{r Fig10B}

H3K27ac_cre_df <- readRDS("data/RDS_files/H3K27ac_cre_df.RDS")
scale_fill_cCREs <-  function(...){
  ggplot2:::manual_scale(
    'fill', 
    values = setNames(c( "#E64B35FF", "#4DBBD5FF", "#00A087FF" ,"#3C5488FF", "#F39B7FFF", "#8491B4FF", "#91D1C2FF", "#DC0000FF", "#7E6148FF", "#B09C85FF","#008B45FF", "#631879FF",  "#5F559BFF", "#A20056FF", "#808180FF", "#1B1919FF"), c(unique(H3K27ac_cre_df$cCRE_class),"NA","CA-only","Low-DNase")), 
    ...
  )
}


H3K27ac_cre_counts <- readRDS("data/RDS_files/H3K27ac_cre_LVheart_counts.RDS")
ggplot(H3K27ac_cre_counts, aes(x = cluster, y = percent, fill = cCRE_class)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    y = "Proportion of ROIs",
    x = "Set",
    fill = "cCRE class"
  ) +
  theme_bw()+
     ggtitle("H3K27ac ROI summits crossing heart LV all cCREs ")+scale_fill_cCREs()
 
 
```

### Current Fig 11

```{r SF11, fig.width=7,fig.height=8}
### make the summary of fragments for all histones
other_frag_summary_file <- fastq_fragments %>% 
  mutate(file=str_remove(file, "_R[12]\\.fastq\\.gz$")) %>% 
  distinct() %>% 
  left_join(sampleinfo,by=c("file"="Library ID")) %>% 
  dplyr::filter(Treatment != "5FU" & Histone_Mark %in% c("H3K9me3","H3K27me3","H3K36me3")) %>% 
  dplyr::select(file, Histone_Mark:Timepoint,num_seqs) %>% 
  left_join(frag_count_total_no_multi, by=c("file"="sample")) %>% 
  left_join(fragment_count_totals_final, by=c("file"="sample")) %>% 
  dplyr::rename("all_fragments"=num_seqs,
                "mapped_fragments"= fragment_count_total.x,
                "unique_noM_dedup"=fragment_count_total.y)

other_frag_summary_file %>% 
   pivot_longer(cols=c(all_fragments:unique_noM_dedup), names_to = "type", values_to = "frag_counts") %>% 
  mutate(Timepoint=factor(Timepoint, levels = c("24T","24R","144R"),labels = c("0R","24R","144R")),
         Treatment=factor(Treatment, levels = c("VEH","DOX"))) %>% 
  mutate(group=paste0(Timepoint,"_",Treatment)) %>% 
  mutate(group=factor(group,levels = c("0R_VEH","24R_VEH","144R_VEH","0R_DOX", "24R_DOX", "144R_DOX"))) %>% 
  ggplot(., aes(x=type, y=frag_counts))+
  geom_boxplot(aes(fill=Treatment))+
  geom_point(aes(col=Individual, size=2))+
   theme_bw()+
  facet_wrap(Histone_Mark~group,nrow = 3, ncol = 6 )+
 scale_fill_brewer(palette="Set2")+
  scale_color_brewer(palette = "Dark2")+
  theme(strip.text = element_text(face = "bold",  hjust = 0, size = 8),
        strip.background = element_rect(fill = "white", linetype = "solid",
                                        color = "black", linewidth = 1),
        panel.spacing = unit(1, 'points'),
        axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1))+
  coord_cartesian(ylim=c(0,19000000))

```
 
### Current Fig 12  

```{r SF12,fig.width=10,fig.height=10}
annotated_combo_df %>% 
  mutate(Timepoint=factor(Timepoint, levels = c("24T","24R","144R"),labels = c("0R","24R","144R")),
         Treatment=factor(Treatment, levels = c("VEH","DOX"))) %>% 
  mutate(sample_name=paste0(Individual,"_",Treatment,"_",Timepoint),
         trt_time=paste0(Treatment,"_",Timepoint),
         trt_time=factor(trt_time, levels = c("VEH_0R","VEH_24R","VEH_144R","DOX_0R","DOX_24R","DOX_144R"))) %>% 
   arrange( Treatment,Individual, Timepoint) %>%
  mutate(sample_name = factor(sample_name, levels = unique(sample_name))) %>%
  dplyr::filter(Histone_Mark!="H3K27ac") %>%
  ggplot(., aes(x=sample_name, y=Col1, weight = weight,fill = Treatment))+
geom_violin(bw = 5) +
    scale_y_continuous(breaks = seq(0, 800, 50)) +
   theme_bw(base_size = 15) +
  facet_wrap(~Histone_Mark, nrow=3, ncol=1)+
    ggpubr::rotate_x_text(angle = 90) +
  ggtitle("Fragment lengths for other histones")+
    ylab("Fragment Length") +
    xlab("")+
  scale_fill_brewer(palette="Set2")


```
### Current Fig 13 
 
```{r SF13}

```
 
 
### Current Fig 14  

```{r SF14,fig.width=6,fig.height=10}
Peak_n_FRIP_other <- H3K27me3_peaks %>% 
  bind_rows(H3K9me3_peaks) %>% 
  bind_rows(H3K36me3_peaks) %>% 
  left_join(sampleinfo, by =c("Sample"='Library ID')) %>% 
  mutate(Timepoint=factor(Timepoint, levels = c("24T","24R","144R"),labels = c("0R","24R","144R")),
         Treatment=factor(Treatment, levels = c("VEH","DOX"))) %>% 
  left_join(peak_ct, by=c("Sample"="Sample")) %>% 
  mutate(sample_name=paste0(Individual,"_",Treatment,"_",Timepoint),
         trt_time=paste0(Treatment,"_",Timepoint),
         trt_time=factor(trt_time, levels = c("VEH_0R","VEH_24R","VEH_144R","DOX_0R","DOX_24R","DOX_144R")))


ggplot(Peak_n_FRIP_other, aes(x=trt_time, y = Count))+
  geom_boxplot(aes(fill=Treatment, alpha=.8))+
  geom_point(aes(color=Individual, size=4))+
  labs(title="Count of Histone peaks")+
  ylab("Number of Peaks")+
  xlab("Treatment and Time")+
  facet_wrap(~Histone,nrow=3, ncol=1)+
  theme_bw()+
  scale_color_brewer(palette="Dark2")+
  scale_fill_brewer(palette="Set2")+
  coord_cartesian(ylim = c(0,3e5))
  
```

### Current Fig 15

```{r SF15,fig.width=6,fig.height=10}
ggplot(Peak_n_FRIP_other, aes(x=trt_time, y = FRiP))+
  geom_boxplot(aes(fill=Treatment, alpha=.8))+
  geom_point(aes(color=Individual, size=4))+
  labs(title="Fraction of Reads in Peaks other histones")+
  ylab("Reads in Peaks/ Total Reads")+
  xlab("Treatment and Time")+
  theme_bw()+
  facet_wrap(~Histone,nrow=3, ncol=1)+
  scale_color_brewer(palette="Dark2")+
  scale_fill_brewer(palette="Set2")+
  coord_cartesian(ylim=c(0,1))

```

