---
title: "TP53_motif_interrogation"
author: "Renee Matthews"
date: "`r Sys.Date()`"
output: 
  html_document:
  toc: true
  toc_float: true
  toc_depth: 4
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	dev = c("png","pdf")
)
```

```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

```{r package loading}
library(tidyverse)
library(GenomicRanges)
library(plyranges)
library(genomation)
library(readr)
library(rtracklayer)
library(stringr)
library(ggrepel)
library(DT)
library(data.table)
library(readxl)
library(ChIPseeker)
library(smplot2)
library(biomaRt)
library(TFBSTools)
library(JASPAR2022)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(universalmotif)
library(kableExtra)

```









## Exploring TP53 motifs across MER61 subfamily
Testing files
```{r data set loading}

##loading TE dataframe
te_df <- read.table("data/Bed_exports/repeatmasker_sets/ERV1/MER61E_ERV1_LTR.bed", header=FALSE) # or your BED
colnames(te_df) <- c("chr","start","end","name","score","strand")
##checking for NAs
sum(is.na(te_df$start))
sum(is.na(te_df$end))
##cleaning TE dataframe (usually not need, this is just in case)
te_df_clean <- te_df[!is.na(te_df$start) & !is.na(te_df$end), ]
##Convert to Granges
te_gr <- GRanges(
  seqnames = te_df_clean$chr,
  ranges = IRanges(start=te_df_clean$start, end=te_df_clean$end),
  strand = te_df_clean$strand)


### data about bed files for future reference

ERV1_summary_char <- read.csv(file.path("data/Other_paper_data/ERV1_file_summary.csv"))

# ERV1_summary_char %>% 
#   dplyr::filter(filename=="MER61E_ERV1_LTR.bed")

RNA_full_toptable <-  readRDS("data/Other_paper_data/RNA_full_toptable.RDS")


Canonical_p53_sites <- read.delim("C:/Users/renee/Other_projects_data/DXR_data/TE_sea_jaspar_results/MER61E_ERV1_LTR_sea_jaspar/sites.tsv", header=TRUE)

Symbol_lookup <- RNA_full_toptable %>% 
  dplyr::select (Entrez_ID,SYMBOL) %>% 
  distinct(Entrez_ID,SYMBOL)


```
 overlapping and computing relative positions


```{r}
p53_canonical_motifs <- function(fimo_root_dir, te_bed_base_dir) {
 
  
  # Initialize results list
  results_list <- list()
  
  # Find all directories ending with "_fimo"
  fimo_dirs <- list.dirs(fimo_root_dir, recursive = FALSE)
  fimo_dirs <- fimo_dirs[grepl("_sea_jaspar$", basename(fimo_dirs))]
  
  for (fimo_dir in fimo_dirs) {
    base_name <- sub("_sea_jaspar$", "", basename(fimo_dir))
    
    # Find FIMO file
    sites_file <- file.path(fimo_dir, "sites.tsv")
    if (!file.exists(sites_file)) {
      message("Skipping ", fimo_dir, ": sites.tsv not found")
      next
    }
  
  
  te_file <- file.path(te_bed_base_dir, paste0(base_name, ".bed"))
    if (!file.exists(te_file)) {
      message("Skipping ", base_name, ": TE BED file not found")
      next
    }
  
    
    # ------------------------------
    # Load TE BED file
    # ------------------------------
    te_df <- read.table(te_file, header = FALSE, stringsAsFactors = FALSE)
    colnames(te_df) <- c("chr","start","end","name","score","strand")
    te_df <- te_df[!is.na(te_df$start) & !is.na(te_df$end) & (te_df$start <= te_df$end), ]
    te_gr <- GRanges(
      seqnames = te_df$chr,
      ranges = IRanges(start = te_df$start, end = te_df$end),
      strand = te_df$strand
    )
    # ------------------------------
    # Load TP53 sites
    # ------------------------------
    cano_sites <- read.delim(sites_file, stringsAsFactors = FALSE)
    
    # Filter TP53 motifs
    if (!"motif_ALT_ID" %in% colnames(cano_sites)) {
      message("No motif_ALT_ID column in ", sites_file)
      next
    }
    cano_sites <- cano_sites[
      grepl("TP53", cano_sites$motif_ALT_ID, ignore.case = TRUE),
    ]
    # Skip if no motifs found
    if (nrow(cano_sites) == 0) {
      message("No TP53 motifs in ", base_name)
      results_list[[paste0(base_name, "_no_hits")]] <- data.frame(
        TE_name = base_name,
        motif_name = "TP53",
        total_TEs = length(te_gr),
        TEs_with_TP53 = 0,
        total_TP53_hits = 0,
        occupancy = 0
      )
      next
    }
    
    # Clean FIMO entries
    cano_sites <- cano_sites[!is.na(cano_sites$site_Start) & !is.na(cano_sites$site_End) & (cano_sites$site_Start <= cano_sites$site_End), ]
    
    motif_name <- unique(cano_sites$motif_ALT_ID)
    
    # Loop over each motif separately (in case multiple motifs exist)
    for (mn in motif_name) {
      fimo_motif <- cano_sites[cano_sites$motif_ALT_ID == mn, ]
      
      fimo_gr <- GRanges(
        seqnames = fimo_motif$seq_ID,
        ranges = IRanges(start = fimo_motif$site_Start, end = fimo_motif$site_End),
        strand = fimo_motif$site_Strand,
        score = fimo_motif$site_Score
      )
      
      # ------------------------------
      # Overlaps
      # ------------------------------
      hits <- findOverlaps(fimo_gr, te_gr)
      n_te_total <- length(te_gr)
      
      if (length(hits) == 0) {
        message("No overlaps for motif ", mn, " in ", base_name)
        results_list[[paste0(base_name, "_", mn)]] <- data.frame(
          TE_name = base_name,
          motif_name = mn,
          total_TEs = n_te_total,
          TEs_with_TP53 = 0,
          total_TP53_hits = 0,
          occupancy = 0
        )
        next
      }
      
      n_te_with_hits <- length(unique(subjectHits(hits)))
      n_total_hits <- length(hits)
      occupancy <- ifelse(n_te_total > 0, n_te_with_hits / n_te_total, NA)
      
    
    # ------------------------------
      # Compute relative positions
      # ------------------------------
      rel_pos <- (start(fimo_gr)[queryHits(hits)] - start(te_gr)[subjectHits(hits)]) /
                 width(te_gr)[subjectHits(hits)]
      rel_pos <- pmin(pmax(rel_pos, 0), 1)
      df <- data.frame(rel_pos = rel_pos)
      
      # ------------------------------
      # Store results
      # ------------------------------
      results_list[[paste0(base_name, "_", mn)]] <- data.frame(
        TE_name = base_name,
        motif_name = mn,
        total_TEs = n_te_total,
        TEs_with_TP53 = n_te_with_hits,
        total_TP53_hits = n_total_hits,
        occupancy = occupancy
      )
      
      # ------------------------------
      # Plot density
      # ------------------------------
      cat(paste0("### ", base_name, " ", mn, "\n"))
      cat(base_name, "| total TEs:", n_te_total, "| with cTP53:", n_te_with_hits, 
          "| hits:", n_total_hits, "\n")
      cat("Sequences with full TP53 motif: ", round(occupancy * 100, 2), "%\n\n")
      
      p <- ggplot(df, aes(x = rel_pos)) +
        geom_density(fill = "steelblue", alpha = 0.5) +
        labs(x = "Relative position in TE", y = "Motif density") +
        theme_classic() +
        ggtitle(paste0("canonical TP53 motif distribution: ", base_name,
                       "\nSequences with motif: ", round(occupancy * 100, 1), "%"))
      print(p)
      cat("\n\n")
    }
  } # end fimo_dirs loop
  
  # Combine all results into one data frame
  results_df <- do.call(rbind, results_list)
  
  print(results_df)
  return(results_df)
}
    




```

```{r}
full_p53 <- p53_canonical_motifs("C:/Users/renee/Other_projects_data/DXR_data/TE_sea_jaspar_results","data/Bed_exports/repeatmasker_sets/ERV1")

kable(full_p53, caption = "TP53 full motif occupancy per TE family") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE) 
```


```{r pwm based on TingWang, eval=FALSE}


A <- c(0.48,-0.06,0.72,-3.59, 0.99,-0.07,-2.17,-2.17,-2.41,-0.65) 
C <- c(-1.21, -1.89,-2.29,1.34,-1.68,-1.74,-2.95, 0.55, 1.03, 0.44)
G <- c(0.44,1.03,0.55,-2.95,-1.74, -1.68,1.34,-2.29,-1.89, -1.21)
T <- c(-0.65,-2.41,-2.17,-2.17,-0.07, 0.99,-3.59,0.72,-0.06,0.48)

TP53_half_mat <- rbind(A,C,G,T)

# background frequencies
bg <- c(A=0.25, C=0.25, G=0.25, T=0.25)

# convert log-odds → probabilities
TP53_half_prob_mat <- apply(TP53_half_mat, 2, function(col) {
  p <- bg * (2^col)
  p / sum(p)  # normalize
})

# TP53_half_prob_mat <- t(TP53_half_prob_mat)

rownames(TP53_half_prob_mat) <- c("A","C","G","T")

colSums(TP53_half_prob_mat) #all ones?  sanity checking as they should be
TP53_half_prob_mat <- as.matrix(TP53_half_prob_mat)

TP53_half_motif <- create_motif(TP53_half_prob_mat, name = "TP53_half_motif_paper_pwm", alphabet = "DNA")


write_meme(TP53_half_motif, file = "data/Other_paper_data/TP53_half_motif_paper_pwm.meme")

```

### looking a TP53 half sites from Ting Wang paper

```{r plot p53 locations}

plot_p53_motifs_all <- function(fimo_root_dir, te_bed_base_dir) {
 
  
  # Initialize results list
  results_list <- list()
  detail_list <- list()
  # Find all directories ending with "_fimo"
  fimo_dirs <- list.dirs(fimo_root_dir, recursive = FALSE)
  fimo_dirs <- fimo_dirs[grepl("_fimo$", basename(fimo_dirs))]
  
  for (fimo_dir in fimo_dirs) {
    base_name <- sub("_fimo$", "", basename(fimo_dir))
    
    # Find FIMO file
    fimo_file <- file.path(fimo_dir, "fimo.tsv")
    if (!file.exists(fimo_file)) {
      message("Skipping ", fimo_dir, ": fimo.tsv not found")
      next
    }
    
    # Find corresponding TE BED file
    te_file <- file.path(te_bed_base_dir, paste0(base_name, ".bed"))
    if (!file.exists(te_file)) {
      message("Skipping ", base_name, ": TE BED file not found")
      next
    }
    
    # ------------------------------
    # Load TE BED file
    # ------------------------------
    te_df <- read.table(te_file, header = FALSE, stringsAsFactors = FALSE)
    colnames(te_df) <- c("chr","start","end","name","score","strand")
    te_df <- te_df[!is.na(te_df$start) & !is.na(te_df$end) & (te_df$start <= te_df$end), ]
    te_gr <- GRanges(
      seqnames = te_df$chr,
      ranges = IRanges(start = te_df$start, end = te_df$end),
      strand = te_df$strand,
      name = te_df$name)
    
    
    # ------------------------------
    # Load FIMO
    # ------------------------------
    fimo <- read.delim(fimo_file, stringsAsFactors = FALSE)
    
    # Filter TP53 motifs
    if ("motif_id" %in% colnames(fimo)) {
      fimo <- fimo[grepl("TP53", fimo$motif_id, ignore.case = TRUE), ]
    } else {
      message("No motif_id column in ", fimo_file)
      next
    }
    
    # Skip if no motifs found
    if (nrow(fimo) == 0) {
      message("No TP53 motifs in ", base_name)
      results_list[[paste0(base_name, "_no_hits")]] <- data.frame(
        TE_name = base_name,
        motif_name = NA,
        total_TEs = length(te_gr),
        TEs_with_TP53 = 0,
        total_TP53_hits = 0,
        occupancy = 0
      )
      next
    }
    
    # Clean FIMO entries
    fimo <- fimo[!is.na(fimo$start) & !is.na(fimo$stop) & (fimo$start <= fimo$stop), ]
    
    motif_name <- unique(fimo$motif_id)
    
    # Loop over each motif separately (in case multiple motifs exist)
    for (mn in motif_name) {
      fimo_motif <- fimo[fimo$motif_id == mn, ]
      
      fimo_gr <- GRanges(
        seqnames = fimo_motif$sequence_name,
        ranges = IRanges(start = fimo_motif$start, end = fimo_motif$stop),
        strand = fimo_motif$strand,
        score = fimo_motif$score
      )
      
      # ------------------------------
      # Overlaps
      # ------------------------------
      hits <- findOverlaps(fimo_gr, te_gr)
      n_te_total <- length(te_gr)
      
      if (length(hits) == 0) {
        message("No overlaps for motif ", mn, " in ", base_name)
        results_list[[paste0(base_name, "_", mn)]] <- data.frame(
          TE_name = base_name,
          motif_name = mn,
          total_TEs = n_te_total,
          TEs_with_TP53 = 0,
          total_TP53_hits = 0,
          occupancy = 0
        )
        next
      }
      
      n_te_with_hits <- length(unique(subjectHits(hits)))
      n_total_hits <- length(hits)
      occupancy <- n_te_with_hits / n_te_total
      te_names <- mcols(te_gr)$name[subjectHits(hits)]
     
      
      # ------------------------------
      # Compute relative positions
      # ------------------------------
      rel_pos <- (start(fimo_gr)[queryHits(hits)] - start(te_gr)[subjectHits(hits)]) /
                 width(te_gr)[subjectHits(hits)]
      rel_pos <- pmin(pmax(rel_pos, 0), 1)
      df <- data.frame(rel_pos = rel_pos)
      
      # ------------------------------
      # Store results
      # ------------------------------
      results_list[[paste0(base_name, "_", mn)]] <- data.frame(
        TE_name = base_name,
        motif_name = mn,
        total_TEs = n_te_total,
        TEs_with_TP53 = n_te_with_hits,
        total_TP53_hits = n_total_hits,
        occupancy = occupancy
      )
      # ------------------------------
# Store details ONLY if hits exist
# ------------------------------
if (length(hits) > 0 && length(subjectHits(hits)) > 0) {

  detail_list[[paste0(base_name, "_", mn)]] <- data.frame(
    TE_name = base_name,
    motif_name = mn,
    Peakid = mcols(te_gr)$name[subjectHits(hits)]
  )

}
      
      # ------------------------------
      # Plot density
      # ------------------------------
      cat(paste0("### ", base_name, " ", mn, "\n"))
      cat(base_name, "| total TEs:", n_te_total, "| with TP53:", n_te_with_hits, 
          "| hits:", n_total_hits, "\n")
      cat("Sequences with TP53 motif: ", round(occupancy * 100, 2), "%\n\n")
      
      p <- ggplot(df, aes(x = rel_pos)) +
        geom_density(fill = "steelblue", alpha = 0.5) +
        labs(x = "Relative position in TE", y = "Motif density") +
        theme_classic() +
        ggtitle(paste0("TP53 motif distribution: ", base_name,
                       "\nSequences with motif: ", round(occupancy * 100, 1), "%"))
      print(p)
      cat("\n\n")
    }
  } # end fimo_dirs loop
  
  # Combine all results into one data frame
  results_df <- do.call(rbind, results_list)
  detail_df <- dplyr::bind_rows(detail_list)
  print(results_df)
  return(list(
  summary = results_df,
  detail  = detail_df
))
}
```

```{r for_tp53motifs}
Astorage <- plot_p53_motifs_all("C:/Users/renee/Other_projects_data/DXR_data/fimo_results", "data/Bed_exports/repeatmasker_sets/ERV1")
table_summary <- Astorage$summary
kable(table_summary, caption = "TP53 motif occupancy per TE family") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%          # make TE_name bold
  column_spec(6, color = "white", background = spec_color(table_summary$occupancy, end = 0.9))  # color occupancy
```

```{r Getting numbers togther}
H3K27ac_summit_gr <- readRDS("data/RDS_files/H3K27ac_complete_summit_gr.RDS")
mcols(H3K27ac_summit_gr)$name <- mcols(H3K27ac_summit_gr)$Peakid
 
 H3K27ac_summit_gr_ext <- resize(H3K27ac_summit_gr, width(H3K27ac_summit_gr)+400, fix = "center")

# need to know # of half sites by TE subfamily integrant, number of whole sites, which sequences have which sites, how many overlap p53 ENcode (using HepG2 cells Myers,Experiment: ENCSR980, accession ENCFF701DTE  bed narrowpeak file)
BB_p53_gr <-  import("data/Other_paper_data/ENCFF701DTE.bed", format = "narrowPeak")
mcols(BB_p53_gr)$peak_id <- paste0("p53_peak_", seq_along(BB_p53_gr))

p53_Crossover_hits <- findOverlaps(H3K27ac_summit_gr_ext,BB_p53_gr)

overlap_df <- data.frame(
  summit_id = queryHits(p53_Crossover_hits),
  ext_range= mcols(H3K27ac_summit_gr_ext)$name[queryHits(p53_Crossover_hits)],
  cluster= mcols(H3K27ac_summit_gr_ext)$cluster[queryHits(p53_Crossover_hits)],
  p53_peak  = mcols(BB_p53_gr)$peak_id[subjectHits(p53_Crossover_hits)]
)
p53bound_region <- overlap_df %>% 
  dplyr::distinct(ext_range,cluster)
    
p53_counts <- p53bound_region %>%
  count(cluster, name = "p53_bound")
total_counts <- as.data.frame(table(H3K27ac_summit_gr_ext$cluster))
colnames(total_counts) <- c("cluster", "total")
p53_counts

counts_df <- left_join(total_counts, p53_counts, by = "cluster") %>%
  mutate(p53_bound = ifelse(is.na(p53_bound), 0, p53_bound),
         not_bound = total - p53_bound)

matrix <- matrix(
  c(
    counts_df$p53_bound[counts_df$cluster == "Set_3"],
    counts_df$not_bound[counts_df$cluster == "Set_3"],
    counts_df$p53_bound[counts_df$cluster == "Set_1"],
    counts_df$not_bound[counts_df$cluster == "Set_1"]
  ),
  nrow = 2,
  byrow = TRUE
)

set2_matrix <- matrix(
  c(
    counts_df$p53_bound[counts_df$cluster == "Set_2"],
    counts_df$not_bound[counts_df$cluster == "Set_2"],
    counts_df$p53_bound[counts_df$cluster == "Set_1"],
    counts_df$not_bound[counts_df$cluster == "Set_1"]
  ),
  nrow = 2,
  byrow = TRUE
)

fisher.test(matrix)
fisher.test(set2_matrix)
```




In order to determine if MER61E integrants are enriched in p53 binding sites compared to MER61 integrants in Set_3 v set1, I need to first find all members of MER61 that are in Set_3.
```{r finding}
p53_TE_ids <- Astorage$detail #contains the TEs that have p53sites
fimo_MER61int <-   read_delim("C:/Users/renee/Other_projects_data/DXR_data/fimo_results/MER61-int_ERV1_LTR_fimo/fimo.tsv")
    
    
Mer_list <- c("MER61-int","MER61A","MER61B","MER61C","MER61D","MER61E","MER61F")
annotated_H3K27ac_ext200 <- readRDS("data/TE_annotation/annotated_H3K27ac_ext200_TE_overlaps.RDS")
all_mer_members <- annotated_H3K27ac_ext200 %>% 
  dplyr::select(Peakid,repClass:repID,cluster ) %>%
  separate_longer_delim(cols = c(repClass, repFamily,repName, repID), delim = ";") %>% 
  
  dplyr::filter(repName %in% Mer_list) %>% 
  mutate(p53=case_when(repID %in% p53_TE_ids$Peakid~"has_p53",
                       TRUE~"no_p53")) %>% 
  mutate(p53_bin=if_else(p53 == "has_p53", 1, 0)) 

all_mer_members_clean <- all_mer_members %>%
  group_by(Peakid) %>%
  summarise(
    repName = paste(unique(repName), collapse = ";"),
    cluster = first(cluster),
    p53_bin = max(p53_bin)   # 1 if any hit
  )

df <- all_mer_members %>%
  dplyr::filter(repName == "MER61E",
         cluster %in% c("Set_1", "Set_3"))

tab <- table(df$cluster, df$p53_bin)
tab
fisher.test(tab)


df_S1n3 <- all_mer_members_clean %>%
  filter(cluster %in% c("Set_1", "Set_3")) %>%
  mutate(
    cluster = factor(cluster, levels = c("Set_1", "Set_3"))
  )


Mer_list <- c("MER61-int","MER61A","MER61B",
              "MER61C","MER61D","MER61E","MER61F")

enrich_results <- lapply(Mer_list, function(m) {

  df_sub <- df_S1n3 %>%
    mutate(is_family = repName == m)

  tab <- table(df_sub$is_family, df_sub$cluster)

  ft <- fisher.test(tab)

  data.frame(
    repName = m,
    odds_ratio = unname(ft$estimate),
    conf_low = ft$conf.int[1],
    conf_high = ft$conf.int[2],
    p_value = ft$p.value,
    n_family = sum(df_sub$is_family),
    n_total = nrow(df_sub)
  )
})

enrichment_df <- bind_rows(enrich_results)

enrichment_df
```


```{r plot OR of MER61 enrichment}
# enrichment_df

ggplot(enrichment_df, aes(x = repName, y = odds_ratio)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  # scale_y_log10() +
  theme_classic() +
  labs(
    y = "Odds ratio (Set3 vs Set1,  not log scale)",
    x = "MER61 subfamily",
    title = "Odds Ratio of MER61 element being in Set 3 v Set 1 compared to rest of Subfamily"
  )+
  coord_flip()



```

```{r finding p53 enrichment}

results <- lapply(Mer_list, function(m) {

  df_sub <- df_S1n3 %>%
    filter(repName == m) %>%
    mutate(has_p53 = p53_bin == 1)

  # skip if not testable
  if (n_distinct(df_sub$cluster) < 2 || n_distinct(df_sub$has_p53) < 2) {
    return(data.frame(
      repName = m,
      odds_ratio = NA,
      p_value = NA,
      note = "insufficient variation"
    ))
  }

  tab <- table(df_sub$cluster, df_sub$has_p53)

  ft <- fisher.test(tab)

  data.frame(
    repName = m,
    odds_ratio = unname(ft$estimate),
    conf_low = ft$conf.int[1],
    conf_high = ft$conf.int[2],
    p_value = ft$p.value,
    n_total = nrow(df_sub)
  )
})

df_S1n3 %>%
  group_by(repName, cluster) %>%
  summarize(
    n = dplyr::n(),
    has_p53 = sum(p53_bin == 1)
  )

p53_enrichment_df <- bind_rows(results)

p53_enrichment_df
```


```{r plot OR of MER61 enrichment}
# enrichment_df

ggplot(p53_enrichment_df, aes(x = repName, y = odds_ratio)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10() +
  theme_classic() +
  labs(
    y = "Odds ratio (Set3 vs Set1, log scale)",
    x = "MER61 subfamily",
    title = "Odds Ratio of MER61 element having a p53 site in Set 3 to Set 1"
  )+
  coord_flip()


# ggplot(enrichment_df, aes(x = repName, y = odds_ratio)) +
#   geom_point(size = 3) +
#   geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
#   geom_hline(yintercept = 1, linetype = "dashed") +
#   scale_y_log10() +
#   theme_classic() +
#   labs(
#     y = "Odds ratio (Set3 vs Set1,  log scale)",
#     x = "MER61 subfamily",
#     title = "Odds Ratio of MER61 element having a p53 site in Set 3 to Set 1"
#   )+
#   coord_flip()
```
now I am building the data frame to test of all PeakIDs that overlap MER61Es and overlap p53 binding regions, is set 3 enriched in both MER61Es and p53 binding regions.
```{r}
Peakid_df <- data.frame(Peakid=annotated_H3K27ac_ext200$Peakid,
                        cluster = annotated_H3K27ac_ext200$cluster)

MER61EPeakIDs <- all_mer_members %>% 
  dplyr::filter(repName=="MER61E") %>% 
  distinct(Peakid)

p53bound_region

Peakid_df$has_MER61E <- Peakid_df$Peakid %in% MER61EPeakIDs$Peakid

Peakid_df$has_p53 <- Peakid_df$Peakid %in% p53bound_region$ext_range

Peakid_df$double_hit <- Peakid_df$has_MER61E & Peakid_df$has_p53

table(Peakid_df$cluster, Peakid_df$double_hit)
```

What about finding if the number of no-p53 sites is enriched in set3 vs set1

```{r}

no_p53_results <- lapply(Mer_list, function(m) {

  df_sub <- df_S1n3 %>%
    filter(repName == m) %>%
    mutate(
      cluster = factor(cluster, levels = c("Set_1","Set_3")),
      no_p53 = factor(p53_bin == 0, levels = c(FALSE, TRUE))
    )
  
  tab <- table(df_sub$cluster, df_sub$no_p53)
  
   # skip if not full 2x2
  if (any(dim(tab) < 2) || any(rowSums(tab) == 0) || any(colSums(tab) == 0)) {
    return(data.frame(
      repName = m,
      odds_ratio = NA,
      conf_low = NA,
      conf_high = NA,
      p_value = NA,
      note = "insufficient data"
    ))
  }

  ft <- fisher.test(tab)

  data.frame(
    repName = m,
    odds_ratio = unname(ft$estimate),
    conf_low = ft$conf.int[1],
    conf_high = ft$conf.int[2],
   p_value = ft$p.value,
    note = "ok"
  )
})

no_p53_enrichment_df <- bind_rows(no_p53_results)

ggplot(no_p53_enrichment_df, aes(x = repName, y = odds_ratio)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_log10() +
  theme_classic() +
  labs(
    y = "Odds ratio (Set3 vs Set1,  not log scale)",
    x = "MER61 subfamily",
    title = "Odds Ratio of MER61 element NOT having a p53 site in Set 3 to Set 1"
  )+
  coord_flip()
df_S1n3 %>%
  group_by(repName, cluster) %>%
  summarize(
    n = dplyr::n(),
    no_p53 = sum(p53_bin == 0)
  )

no_p53_enrichment_df
```


### looking at p53 groupings in MER61E elements.
```{r}
Astorage$detail %>% 
  group_by(Peakid, TE_name) %>% tally() %>% 
  dplyr::filter(n>1) %>% 
  group_by(TE_name) %>% 
  count(TE_name)

```




### KZFP motif similarity

```{r}
KZFP_2026_list <- read_excel("data/Other_paper_data/human_protein_coding_KZFPs_Davis_natgen_2026.xlsx")
KZFP_motifs <- read_meme("data/Other_paper_data/KZFP_motifs.meme")
KZFP_motif_info <- readRDS("data/RDS_files/KZFP_motif_info.RDS")

KZFP_target_TEs <-read_excel("data/Other_paper_data/KZFP_primary_targests_Supplemental_Table_S3.xlsx", sheet = "Primary_targets")
KZFP_motif_names <-KZFP_motif_info %>% 
  dplyr::select(Motif_ID,TF_Name, MSource_Identifier)
KZFP_sim <- compare_motifs(
  KZFP_motifs,
  method = "PCC")

KZFP_dist_mat <- 1 - KZFP_sim
na_per_row <- rowSums(is.na(KZFP_dist_mat))
# KZFP_hc <- hclust(as.dist(KZFP_dist_mat), method = "average")

keep <- na_per_row < (0.2 * ncol(KZFP_dist_mat))  # allow up to 20% NA
KZFP_dist_mat_filt <- KZFP_dist_mat[keep, keep]
sum(is.na(KZFP_dist_mat_filt))
sum(is.nan(KZFP_dist_mat_filt))
sum(is.infinite(KZFP_dist_mat_filt))

max_val <- max(KZFP_dist_mat_filt, na.rm = TRUE)## adding max distance to NAs
KZFP_dist_mat_filt[is.na(KZFP_dist_mat_filt)] <- max_val
diag(KZFP_dist_mat_filt) <- 0  # just to be safe

KZFP_hc <- hclust(as.dist(KZFP_dist_mat_filt), method = "average")

ComplexHeatmap::Heatmap(KZFP_dist_mat_filt,
        cluster_rows = KZFP_hc,
        cluster_columns = KZFP_hc,
        show_row_names = FALSE,
        show_column_names = FALSE)

plot(KZFP_hc, labels = FALSE)
for (h in c(0.3, 0.4, 0.5,0.55,0.6)) {
  cat("h =", h, "clusters =", length(unique(cutree(KZFP_hc, h = h))), "\n")
}
```
```{r}
clusters <- cutree(KZFP_hc, h = 0.55)
cluster_df <- data.frame(
  Motif_ID = names(clusters),
  cluster = clusters
)

cluster_annot <- cluster_df %>%
  dplyr::left_join(KZFP_motif_info, by = "Motif_ID")

# split(cluster_df$motif, cluster_df$cluster)


cluster_list <- split(cluster_annot$TF_Name, cluster_annot$cluster)

table(cluster_annot$cluster)
library(dendextend)

dend <- as.dendrogram(KZFP_hc)
all(names(clusters) == labels(dend))

dend <- color_branches(dend, h = 0.55)

plot(dend)




```

What are the enriched motifs in KZFPs?

Set 2 v Set 1:

```{r}

KZFP_set2vset1 <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/AME_H3K27ac_set2vset1_output/ame.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>%  
  slice_head(n = length(.$ID)-3)

KZFP_set2vset1 <- KZFP_set2vset1 %>% 
  left_join(., KZFP_motif_names, by=c("motif_ID"="Motif_ID"))

KZFP_set3vset1 <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/AME_H3K27ac_set3vset1_output/ame.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>%  
  slice_head(n = length(.$ID)-3)

KZFP_set3vset1 <- KZFP_set3vset1 %>% 
  left_join(., KZFP_motif_names, by=c("motif_ID"="Motif_ID"))

K3v1_set <- KZFP_set3vset1$motif_ID

K2v1_set <- KZFP_set2vset1$motif_ID

intersect_IDs <- intersect(K2v1_set, rownames(KZFP_dist_mat_filt))### some IDs were filtered out in the na removal stage above
filt_mat_2v1<- KZFP_dist_mat_filt[intersect_IDs,intersect_IDs]

filt_mat_3v1 <- KZFP_dist_mat_filt[K3v1_set,K3v1_set]

KZFP_hc_2v1 <- hclust(as.dist(filt_mat_2v1), method = "average")

ComplexHeatmap::Heatmap(filt_mat_2v1,
        cluster_rows = KZFP_hc_2v1,
        cluster_columns = KZFP_hc_2v1,
        show_row_names = FALSE,
        show_column_names = FALSE)


```


```{r}

KZFP_hc_3v1 <- hclust(as.dist(filt_mat_3v1), method = "average")

ComplexHeatmap::Heatmap(filt_mat_3v1,
        cluster_rows = KZFP_hc_3v1,
        cluster_columns = KZFP_hc_3v1,
        show_row_names = FALSE,
        show_column_names = FALSE)


```
```{r}
set2_ids <- rownames(filt_mat_2v1)
set3_ids <- rownames(filt_mat_3v1)

shared_ids <- Reduce(intersect, list(set2_ids, set3_ids))


KZFP_set2vset1 %>% 
  dplyr::filter(`E-value`<0.05) %>% 
   mutate(log10Evalue= log(`E-value`, base = 10)*(-1), 
  TF_name_unique = make.unique(`TF_Name`)) %>% 
  ggplot(., aes (y= reorder(TF_name_unique,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`%TP`/4), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~.*4,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched KZFP motif")+
  ggtitle(paste("SET2 enrichment over SET1 KZFP motifs"))

Set2_KZFP_names <- KZFP_set2vset1 %>% 
  dplyr::filter(`E-value`<0.05) %>% 
  dplyr::select(TF_Name)
```
```{r}
KZFP_target_TEs %>% 
  dplyr::filter(KZFP %in%Set2_KZFP_names$TF_Name) %>% 
  dplyr::select(KZFP:`merged padj.final`,`TE age`,`TE family`) %>% 
  kable(., caption = "SET 2 KZFP TE targets") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE) 


```
```{r SET2 KZFP RNA expression}
KZFP_set2vset1 %>% 
  dplyr::filter(`E-value`<0.05) %>% 
  distinct(TF_Name) %>% 
  dplyr::filter(TF_Name %in% Symbol_lookup$SYMBOL) %>% 
  left_join(RNA_full_toptable, by=c("TF_Name"="SYMBOL")) %>%
  mutate(timepoint=factor(timepoint, levels = c("24T","24R","144R"))) %>% 
  ggplot(aes(y=logFC,x=timepoint))+
  geom_boxplot()+
  geom_point(aes(color=TF_Name))+
  theme_bw()+
  ggtitle("Set 2 KZFP LFC of expressed KZFPs")
```

```{r}
KZFP_set3vset1 %>% 
  dplyr::filter(`E-value`<0.05) %>% 
   mutate(log10Evalue= log(`E-value`, base = 10)*(-1), 
  TF_name_unique = make.unique(`TF_Name`)) %>% 
  ggplot(., aes (y= reorder(TF_name_unique,log10Evalue))) +
  geom_col(aes(x=log10Evalue))+
  geom_point(aes(x=`%TP`/4), size =4)+
  scale_x_continuous(expand=c (0,.125),sec.axis = sec_axis(transform= ~.*4,name="Percent of peaks with motif"))+
  theme_classic()+
  ylab("Enriched KZFP motif")+
  ggtitle(paste("SET3 enrichment over SET1 KZFP motifs"))

Set3_KZFP_names <- KZFP_set3vset1 %>% 
  dplyr::filter(`E-value`<0.05) %>% 
  dplyr::select(TF_Name)


```

```{r SET2 KZFP RNA expression}
KZFP_set3vset1 %>% 
  dplyr::filter(`E-value`<0.05) %>% 
  distinct(TF_Name) %>% 
  dplyr::filter(TF_Name %in% Symbol_lookup$SYMBOL) %>% 
  left_join(RNA_full_toptable, by=c("TF_Name"="SYMBOL")) %>%
  mutate(timepoint=factor(timepoint, levels = c("24T","24R","144R"))) %>% 
  ggplot(aes(y=logFC,x=timepoint))+
  geom_boxplot()+
  geom_point(aes(color=TF_Name))+
  theme_bw()+
  ggtitle("Set 3 KZFP LFC of expressed KZFPs")
```



```{r}
KZFP_target_TEs %>% 
  dplyr::filter(KZFP %in%Set3_KZFP_names$TF_Name) %>% 
  dplyr::select(KZFP:`merged padj.final`,`TE age`,`TE family`) %>% 
  kable(., caption = "SET 3 KZFP TE targets") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE) 
```
All KZFPs from human found in expressed data

```{r}
expressed_KZFPs <- KZFP_2026_list %>% 
  dplyr::filter(Label %in% Symbol_lookup$SYMBOL) %>% 
  pull(Label)
# saveRDS(expressed_KZFPs,"data/RDS_files/expressed_KZFPs.RDS")

```



### KZFP motif binding sites shared with JASPAR TF binding sites

```{r}
tomtom_share_results <- read_delim("C:/Users/renee/Other_projects_data/DXR_data/JAS_KZFP_compare_results/tomtom.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
opts <- list(species = 9606)  # human
pfm_list <- getMatrixSet(JASPAR2022, opts)

jaspar_map <- data.frame(
  motif_ID = sapply(pfm_list, ID),
  TF_name  = sapply(pfm_list, function(x) x@name)
)

tomtom_share_results %>% 
  left_join(., KZFP_motif_names,by=c("Query_ID"="Motif_ID")) %>% 
  left_join(.,jaspar_map, by=c("Target_ID"="motif_ID")) %>% 
  dplyr::filter(`q-value`<0.05) %>% 
  filter(TF_Name == TF_name) %>% 
  distinct(TF_Name)
  
```

```{r}
# sim_mat <-
# tomtom_share_results %>%
#   left_join(., KZFP_motif_names,by=c("Query_ID"="Motif_ID")) %>% 
#   left_join(.,jaspar_map, by=c("Target_ID"="motif_ID")) %>% 
  # mutate(sim=-log10(`q-value`)) %>% 
  # dplyr::select(Query_ID, Target_ID, sim) %>%
  # pivot_wider(names_from = Target_ID,
  #             values_from = sim,values_fill = 0)


tomtom_share_results %>%
  left_join(., KZFP_motif_names,by=c("Query_ID"="Motif_ID")) %>%
  left_join(.,jaspar_map, by=c("Target_ID"="motif_ID")) %>%
mutate(sim=-log10(`q-value`)) %>% 
  filter(Query_consensus == Target_consensus) %>% 
  distinct(TF_Name)

tomtom_share_results %>%
  left_join(., KZFP_motif_names,by=c("Query_ID"="Motif_ID")) %>%
  left_join(.,jaspar_map, by=c("Target_ID"="motif_ID")) %>%
mutate(sim=-log10(`q-value`)) %>% 
  filter(Query_consensus == Target_consensus) 
```


